### 2015-03-10 resistanceModelNotesAndy.Rmd
initial commit, just few edits from Beths code, restructuring comments to enable bookmark navigation in RStudio

---
title: "resistanceModelNotesAndy"
output: html_document
---

Notes on insecticide resistance model for liverpool.
andy south 27/1/15

My first go at keeping notes like this as an rmarkdown file.

**27/1/15**  7.5hrs

Editing `malaria_code_andy2.r`, indenting functions etc.

Much of the script from line 464 is setting up parameters for example runs.

input[a,b] : a=parameter number and b=scenario number

464         set input params
1058-2716   run model
2718        Actions needing full results.list (e.g. curtis plots use multiple scenarios)

edited bookmarks to navigate around the file in RStudio.

**9/3/15** 4-5 1hr

**10/3/15** 4.15 -

```{r}
listFunctions <- function(filename) {
  temp.env <- new.env()
  sys.source(filename, envir = temp.env)
  functions <- lsf.str(envir=temp.env)
  rm(temp.env)
  return(functions)
}

listFunctions("malaria_code_andy.R")
```

So there are only 11 functions.

I can create an initial github repo for the files from Beth (&this notes file)
That means I will have version control for my initial restructuring. Later I may want to create a new repo for the reformatted code as a package.

What to call the first repo ? resistance

1. created repo on github
1. created RStudio project from github repo

### TO DO

1. create a Github repo and just put my edited file from Beth in to start
1. create a readme.md file
1. ? create an RStudio project for the initial version
1. later, structure as R package, maybe move to a new github repo

### 2015-03-10 malaria_code_andy.R

@@ -1,0 +1,2747 @@
+# malaria_code_andy.r 
+# andy south 10/3/2015
+# a few edits from :
+
+# Malaria insecticide resistance model ###
+# Bethany Levick, LSTM 2013 ##
+
+
+
+#### SET WORKING DRIVE  ####
+# If using .csv file to enter parameters, ensure file is stored in same directory as this document
+# Graphical outputs and the fitness parameter matrices will also be stored in this directory
+# Copy and paste this line into the R console, save the code document, then enter the source line 
+# setwd( "C:/Users/Bethany/Dropbox/Levick et al manuscript/")  
+
+#andy this allows me to disable bits that stop my file listing from working
+tempDisabled <- TRUE
+if( !tempDisabled ) setwd( "C:\\Dropbox\\Ian and Andy\\andy\\malaria\\")
+
+#### INPUT PARAMETERS ####
+#Enter file name containing parameters before saving code document 
+#If not using input file, set params.csv to FALSE, and enter a calibration number
+#Save file once these are set, and then run code
+#andy commented out next 4 lines & replaced with below to run inbuilt scenario
+# params.csv <- TRUE
+#input <- read.csv( "full-dom_SD.csv", header=T, stringsAsFactors=F )	
+# produce.plots <- FALSE	## If true, will automatically produce the plots listed below and save to same directory as code is stored
+# calibration <- 100	## If inputting through a .csv, this will be overwritten by calibration row of input file
+
+#andy to run an inbuilt scenario
+params.csv <- TRUE
+produce.plots <- TRUE	## If true, will automatically produce the plots listed below and save to same directory as code is stored
+calibration <- 102	## If inputting through a .csv, this will be overwritten by calibration row of input file
+#input <- read.csv( "input.parameters.csv", header=T, stringsAsFactors=F )  
+if( !tempDisabled ) input <- read.csv( "input.parameters.col1.csv", header=T, stringsAsFactors=F ) 
+
+#BUT object 'input' not found
+#so I tried reading first the input csv and then just the saved first column
+#then with params.csv <- FALSE
+#Error in matrix(nrow = max_gen, ncol = 11) : non-numeric matrix extent
+#with params.csv <- TRUE
+#Error in `rownames<-`(`*tmp*`, value = c("Calibration (100 default)",  : 
+#                                           attempt to set 'rownames' on an object with no dimensions
[…]

### 2015-03-11 resistanceModelNotesAndy.Rmd
recording progress

@@
 
-Editing `malaria_code_andy2.r`, indenting functions etc.
+Editing `malaria_code_andy.r`, indenting functions etc.
 
@@
 
-**10/3/15** 4.15 -
+**10/3/15** 4.15 - 7.15 3hrs
 
@@
 
-listFunctions("malaria_code_andy.R")
+listFunctions("malaria_code_fromBeth.R")
 ```
@@
 So there are only 11 functions.
+
+allele.freq : function (mat)  
+curtis_f1 : function (nrelaxmat, relaxmat, gencol, r1col)  
+curtis_f2 : function (combmat, bmat, amat, gencol, r1col, r2col)  
+curtis_ld : function (resultsmat, relaxedmat, gencol, ldcol)  
+haplotype : function (mat)  
+HW : function (P, mat)  
+linkage : function (mat)  
+make.genotypemat : function (P_1, P_2)  
+make.matrix : function (mat, rnames)  
+singlealleleFrequency : function (locus, max_gen, results.list, input)  
+timetoFifty : function (locus, max_gen, results.list, input
 
@@
 1. created RStudio project from github repo
++ To get ssh push working.
++ RStudio Tools, Shell
++ git remote set-url origin git@github.com:AndySouth/resistance.git
+
+**11/3/15** 9.30
+
+moved my dropbox folder back to C.
 
@@
 
-1. create a Github repo and just put my edited file from Beth in to start
-1. create a readme.md file
-1. ? create an RStudio project for the initial version
+1. copy R functions out of `malaria_code_andy` into their own files in an R folder
+1. add roxygen documentation for functions
 1. later, structure as R package, maybe move to a new github repo

### 2015-03-11 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-**11/3/15** 9.30
+**11/3/15** 9.30-12.30 3hrs 2.15-6.45 7.5hrs
 
 moved my dropbox folder back to C.
+
+Plan
+1. move Beths functions out into a single file for each in an R folder
+1. edit Beths commments into roxygen format
+1. only later think about changing functions and documentation to follow best practice
+
+Changed plan slightly by appending plot onto start of plot functions.
+
+To source files in the R folder
+`lapply(dir("R"),function(x) source(paste0("R//",x)))`
 
@@
 1. later, structure as R package, maybe move to a new github repo
+
+### suggestion from Beth
+looking at it now I realise this is not ideal
+a solution may be to provide a .csv file with the Curtis scenarios set up
+and then the need to set the calibration number and params.csv in the R script can be removed altogether
+so it is all set up in the .csv file - I think similar can be done with produce.plots
+so the only bit of R code the user will need to change is the setwd() and the read.csv()

### 2015-03-12 resistanceModelNotesAndy.Rmd
progress notes

@@
 **27/1/15**  7.5hrs
+
+```
+#working version of Beths code
+#run it outside of this project because it changes wd
+source("C:\\Dropbox\\Ian and Andy\\andy\\malaria\\Beth code\\malaria_code_beth.r")
+```
 
@@
 
+**12/3/15** 9.30-14 4.5hrs 3-5 6-7 7.5hrs
+
+The model generates 3 matrices : results, genotype, fitness
+1. Results : freq of R allele at each loci in each sex and linkage disequilibrium of R allele in each sex, per generation
+2. Genotype : frequencies of each of the ten genotypes, per generation
+3. Fitness : fitness scores of each genotype/niche combination (table 4. of Main Document)
+
+When a file is input with multiple scenarios, each of the three matrices is stored in a list, where scenario number gives position of the matrix in the list.
+
+results.list : 
+fitness.list : 
+genotype.list :
+
+Can I put these into a single list so that they can be returned from a single function ?
+
+done ~ put input object creation into a function
+But there is still a fair bit of code that reads parameter values out of the input matrix and into named variables.
+
+I might be able to restructure these collections of single variables into arrays.
+e.g. for the exposure levels of m&f
+
+```{r}
+#to create an array a[sex][locus1][locus2]
+sex <- c("F","M")
+locus1 <- c("0","A","a")
+locus2 <- c("0","B","b")
+dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
+dim1 <- sapply(dimnames1, function(x) length(x))
+a <- array(0,dim=dim1, dimnames=dimnames1)
+a
+length(a)
+#to access one element
+a['M','a','b']
+#to check that male exposures total 1
+sum(a['M',,])
+```
+
+```
+  ## Exposure levels of males and females to each insecticide niche ##
+  # males
+  a.m_00 <- input[8,i]
+
+  a.m_a0 <- input[9,i]
+  a.m_A0 <- input[10,i]
+
+  a.m_0b <- input[11,i]
+  a.m_0B <- input[12,i]
+
+  a.m_ab <- input[13,i]
+  a.m_AB <- input[14,i]
+
+  a.m_Ab <- input[15,i]
+  a.m_aB <- input[16,i]
+
+  #a.m <- sum(a.m_00, a.m_a0, a.m_A0, a.m_0b, a.m_0B, a.m_ab, a.m_AB, a.m_Ab, a.m_aB)
+  #if ( a.m != 1 ){		 
+  #	print( paste("Error in male exposures: must total one: ", a.m) )
+
+  #	}else{
+  #		print( paste( "Male exposures total 1: ", a.m ))
+  #		}
+
+  # females
+  a.f_00 <- input[17,i]
+
+  a.f_a0 <- input[18,i]
+  a.f_A0 <- input[19,i]
+
+  a.f_0b <- input[20,i]
+  a.f_0B <- input[21,i]
+
+  a.f_ab <- input[22,i]
+  a.f_AB <- input[23,i]
+
+  a.f_Ab <- input[24,i]
+  a.f_aB <- input[25,i]
+```
+
+I should check that what we are doing fits in with what others are doing. We may be able to capitalise on existing tools.
+
+[popgen on CRAN](http://cran.r-project.org/web/packages/popgen/popgen.pdf)
+seems a bit old and doesn't do a huge amount, no vignette.
+
+[gstudio](http://cran.r-project.org/web/packages/gstudio/) looks more promising. e.g. it defines locus objects.
+[gstudio documentation](http://dyerlab.github.io/gstudio/)
+A locus object is set like this, but I'm not sure if this would help us.
+It does have some stuff on drift at end of help.
+```
+require(gstudio)
+loc <- locus(c("C", "A"))
+loc
+```
+
+pegas provides functions for the analysis of allelic data and of haplotype data from DNA sequences.
+It requires and complements two other R-packages: ape and adegenet.
+[pegas](http://ape-package.ird.fr/pegas.html)
+[pegas data structures](http://ape-package.ird.fr/pegas/DefinitionDataClassesPegas.pdf)
+includes a class called loci. : An object of class "loci" is a data frame where rows represent individuals
+and columns are loci and optional additional variables.
+Again I can't quite see how it would be useful for us.
+
+There is a [HardyWeinberg](http://cran.r-project.org/web/packages/HardyWeinberg/vignettes/HardyWeinberg.pdf) package on CRAN.
+"The HardyWeinberg package consists of a set of tools for analyzing diallelic
+genetic markers, and is particularly focused on the graphical representation of their (dis)equilibrium condition in various ways."
+
+There is a CRAN [genetics task view](http://cran.r-project.org/web/views/Genetics.html).
+
+The [genetics package](http://cran.r-project.org/web/packages/genetics/genetics.pdf) which also has a locus class, seems to be more detailed, about chromosomes etc. and probably not of use to us.
+
+#### Structure of the results matrices
+
+```{r}
+max_gen <- 2 #just as exampel
+
+# Set up results matrix - prints overall freq of R and S allele per locus per sex, LD and overall allele freq (i.e. 1)
+results <- matrix ( nrow = max_gen, ncol = 11 )
+colnames( results ) <- c( "Gen", "m.R1", "m.R2", "m.LD","f.R1", "f.R2", "f.LD", "M", "F", "dprime", "r2" )
+
+# set up fitness by niche matrix - records fitness scores for each niche for each genotype
+fitness <- matrix ( nrow = 10, ncol = 9, c(rep(0,90)))
+colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
+rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
+						"RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
+						"RR1SS2", "RR1RS2", "RR1RR2")
+
+# set up genotype matrix - records frequencies of each of the 9 two locus genotypes each generation
+genotype <- matrix( nrow=max_gen, ncol=11 )
+colnames(genotype) <- c("gen", "SS1SS2", "SS2RS2", "SS1RR2", 
+						"RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
+						"RR1SS2", "RR1RS2", "RR1RR2")
+
+```
+
+Somehow it seems I've broken it, I'm getting NAs in results matrix.
+
+The input files are different (but same structure) between Beths version that works and my new version ...
+calibration is 100 in my failing version and 1012 in Beths, i think this is because I'm reading in the csv when I shouldn't.
+
+I seem not to get to line 889 in my version
+  	# male
+  	f.m.SS1SS2 <- genotype.freq[1,]
+
+Then suddenly it seemed to start working when I put a browser in.
+It's something tricky around line 711, if it doesn't get to 889 it doesn't fill in the results matrix ??
+
+Aha! It seems to have been this without the comments before the tilde's that stoppped it from working
+```
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 		
+  ### Loop to run the model from the initial conditions generated above ####
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ 
+```
+
 ### TO DO
 
-1. copy R functions out of `malaria_code_andy` into their own files in an R folder
-1. add roxygen documentation for functions
+1. check what the tilde does in R
+1. copy the 'run' bit out of `malaria_code_andy.r` into a function 
 1. later, structure as R package, maybe move to a new github repo
+1. re-look at R popgen packages when I've got more of a handle on what we are doing.
 

### 2015-03-16 resistanceModelNotesAndy.Rmd
progress notes

@@
 ```{r}
-max_gen <- 2 #just as exampel
+max_gen <- 2 #just as example
 
@@
 
+**13/3/15** 9.15-12.15 3hrs
+
+A reproducible version of yesterdays bug.
+
+```
+x<-1
+~
+for(i in 1:10) x<-x+1
+cat(x)
+```
+[R bug reporting](https://bugs.r-project.org/bugzilla3/enter_bug.cgi)
+
+1. can I create a list of the 3 lists for results ?
+Yes
+Replace this  
+results.list <- list()			
+fitness.list <- list()
+genotype.list <- list()	
+with  
+listOut <- list( results=list(), fitness=list(), genotype=list() )
+
+Then find & replace  
+results.list with listOut$results  
+fitness.list with listOut$fitness
+genotype.list with listOut$genotype
+
+**16/3/15** monday 18.30
+
+cut out lines 92 to 1740 into a runModel() function, later can subdivide it up
+
+In max(combmat[, gencol]) : no non-missing arguments to max; returning -Inf
+fixed by returning listOut from runModel()
+
 ### TO DO
 
-1. check what the tilde does in R
-1. copy the 'run' bit out of `malaria_code_andy.r` into a function 
+1. add a savePlots argument, to enable plotting without saving
+1. submit tilde bug report
 1. later, structure as R package, maybe move to a new github repo

### 2015-03-18 resistanceModelNotesAndy.Rmd
prog notes

@@
 
-**16/3/15** monday 18.30
+**16/3/15** mon 18.30-19.30 1hr
 
@@
 
+**17/3/15** tues to liverpool 8.30-5.30 9hrs
+
+done: add a savePlots argument, to enable plotting without saving
+
+Assuming that resistance to each insecticide was encoded at one locus respectively, and that these can either have a homozygous susceptible (SS), heterozygous (RS) or homozygous resistant (RR) genotypes at each locus, there are then ten possible genotypes (including cis & trans forms of the double heterozygous).
+
+We consider two insecticides (A and B) at three possible levels; absent, low(ab) and high(AB), resulting in 9 niches;
+
+Locus1 relates to resistance for insecticide A, and locus2 for B. Locus2 does not effect fitness under exposure to A.
+
+Males and females can be exposed differently.
+
+Curtis states that if resistance to one insecticide is present at very low levels in the population, then resistance will rise slower if this insecticide is used on its own, rather than in a combination with a more established insecticide.
+
+This is the important Curtis statement “the use of a mixture where the initial gene frequencies are unequal leads to more rapid increase in the frequency of the rarer of the genes”
+
+A key aspect of Curtis’ prediction was that this relationship would be determined by an increasing association between the presence of the two resistance alleles, measured as high levels of linkage disequilibrium.
+
+The results given above seem to be in contrast with Curtis’ assumption, and suggest that in fact using the insecticides in combination can slow the spread of resistance.
+
+My summary :
+Curtis suggests resistance spreads more quickly with 2 insecticides, due to  increasing association between the alleles. Beths work suggests that combination can slow spread of resistance.
+
+Good figure 1 in manuscript :
+The rate of spread of IR depends on whether the resistance mutation is dominant, semi-dominant, or recessive.
+
+**aha! I've only just understood this bit**
+**At high concentrations only the RR individuals survive so resistance is recessive; as concentrations decline some RS mosquitoes survive making resistance semi-dominant; at low concentration both RR and RS mosquitoes survive, making resistance dominant.**
+**so main problems of resistance arise when insecticide concentrations decline**
+
+From manuscript about scenarios to run:
+**row in params csv**
+
+Insecticide coverage for females: 0.1, 0.2, 0.6, 0.8, 0.9 (n=5) **9 & 15**  
+Insecticide coverage for males: as for females, or half that of females (n=2) **18 & 24**  
+‘New’ insecticide resistance starting freq: 0.0001, 0.001, 0.01 (n=3) **6**  
+‘Old’ insecticide resistance starting freq: 0.05, 0.1, 0.2, 0.5, 0.8 (n=5) **7**  
+Resistance heterozygous fitness: 0.5, 0.75. 1    (n=3 for each insecticide) **?28 & 30**  
+Sensitive heterozygous fitness:0, 0.2, 0.4 (n=3 for each insecticide) **?27 & 29**  
+Resistance allele dominance: 0, 0.2, 0.5, 0.9, 1 for each locus (n=5 for each locus) **35,38 do to start**  
+Fitness costs: 0, 0.05, 0.1 (n=3) **43&44**  
+Fitness dominance: 0, 0.5, 1 (n=3) **33&36**
+
+BUT now we are going to do it by sampling from a uniform distribution before doing decision trees. Then you can progressively add results to a file.
+
+I could start by creating a document with the results at the extremes of the parameter space.
+
+Send Ian & Beth devtools installation instructions and a document.
+
+~60,000 combinations…. Too many???
+
+would latin cube hypersampling be better?
+
+The simulation output was the ratio of  the time (in generation) for resistance  to the ‘new’ insecticide  to reach a given allele threshold frequency when the ‘new’ insecticide was deployed as combination, compared to the time taken to reach the same threshold if deployed on its own; a ratio >1 indicates that combination is better.
+Three threshold frequencies were investigated: 0.1, 0.25, 0.5.
+
+These ratios was then analysed by regression to find out how the input parameters alter the ratio, and hence under what circumstance combination would be best/worst.
+Classification tress can also be done….
+
+The intention was not to consider particular plausible deployment situations, but was to systematically investigate a wide range of parameter space to determine whether CD was better than single-insecticide deployment at reducing the rate at which the minor resistance form spreads.
+
+Show my progress :
+
+1. github repo
+1. move functions to their own files and structure documentation
+1. move parts of main file into functions
+1. combine 3 outputs to a single list so that can be returned from a function
+1. nearly able to create a package with structured documentation etc.
+
+Questions for Beth & Ian
+
+1. Curtis Table1 vi that shows resistance spreading faster with mixture.
+1. is coverage same as exposure ? yes
+1. what subset of parameters would they like me to vary to start
+1. locations of sensitivity params in params.csv
+1. what outputs are wanted for sensitivity scenarios ? num generations until resistance reaches threshold (0.1, 0.25 & 0.5) Curtis just used 0.5.
+
+Talking with Ian & Beth
+
+Sequential use.
+
+For some scenarios, we'll need to run with one insecticide and then stop and start other.
+
+When does mixture fail, until one resistance allele reaches critical point, then continue using 2nd insecticide until resistance for that one reaches critical point too.
+
+Could create a version of timetoFifty that accepts threshold param.
+
+Ian has created list of defaults in most recent manuscript.
+
+The Curtis version is going to be simpler. Because not all param variations are included.
+
+**17/3/15** weds in liverpool 8.45-
+
+Ian created `suggested_sensitivity_analysis.doc`
+
+I created a variables table in there with Ians parameter ranges.
+
+My poor paraphrasing of what I think Ian said about the sequential scenarios.
+The starting conditions from after one insecticide has 'failed' are the same as starting from scratch for the other insecticide because the insecticides to do effect selection on the other locus.
+
+moving `malaria_code_andy.r` into a function resistanceMaster()
+exposed a couple of variable scoping problems.
+
+plus now the graphs look different and I get :
+Warning message: In data.matrix(mat) : NAs introduced by coercion
+
+this is from createInputMatrix()
+input <- make.matrix(input, input$Input)
+
+this might be something to do with the location of "input.parameters.csv" which I need to sort ...
+
+devtools commands for setting up a package :
+create requires that the directory doesn't exist yet; it will be created. 
+setup assumes an existing directory from which it will infer the package name.
+
+probably need to set rstudio rg to false because mine is already a rs project.
+
+setup(rstudio=FALSE)
+
+Then in RStudio I had to set project options, build tools to package
+(so maybe I should have done just done detools::setup())
+
+but on build :
+ERROR: The build directory does not contain a DESCRIPTION
+file so cannot be built as a package.
+Build directory: C:/rsprojects/resistance
+
+but there is a DESCRIPTION there ???
+
+devtools::load_all(".")
+Error: Line starting 'person("Bethany", "L ...' is malformed!
+sorted, just needed tab before 2ry authors
+
 ### TO DO  
-
-1. add a savePlots argument, to enable plotting without saving
+1. add myself as maintainer in DESCRIPTION
+1. sort location in package of "input.parameters.csv" functions should be able to run without it
+1. check out the part about running sequential insecticide scenarios
+1. start to write script to run sensitivity analysis, does it want to work by creating a large input.csv file, or by setting input object directly.
+1. spend some time developing my understanding of how the model works
 1. submit tilde bug report

### 2015-03-18 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-listFunctions("malaria_code_fromBeth.R")
+#listFunctions("C:\\Dropbox\\Ian and Andy\\andy\\malaria\\Beth code\\malaria_code_beth.r")
 ```
@@
 
-**17/3/15** weds in liverpool 8.45-
+**17/3/15** weds in liverpool 8.30-6 9.5hrs
 
@@
 
+package does build now
+
+potential refactoring of code, if i think carefully, using arrays I should be able to replace this :
+```
+    if(calibration==103){		## no selection calibration
+        ## male
+        # SS1
+        fs.m.SS1SS2 <- f.m.SS1SS2
+        fs.m.SS1RS2 <- f.m.SS1RS2
+        fs.m.SS1RR2 <- f.m.SS1RR2
+        # RS1
+        fs.m.RS1SS2 <- f.m.RS1SS2 
+        fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
+        fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
+        fs.m.RS1RR2 <- f.m.RS1RR2
+        # RR2 
+        fs.m.RR1SS2 <- f.m.RR1SS2
+        fs.m.RR1RS2 <- f.m.RR1RS2
+        fs.m.RR1RR2 <- f.m.RR1RR2
+
+        ## female
+        # SS1
+        fs.f.SS1SS2 <- f.f.SS1SS2
+        fs.f.SS1RS2 <- f.f.SS1RS2
+        fs.f.SS1RR2 <- f.f.SS1RR2
+        # RS1
+        fs.f.RS1SS2 <- f.f.RS1SS2 
+        fs.f.RS1RS2_cis <- f.f.RS1RS2_cis
+        fs.f.RS1RS2_trans <- f.f.RS1RS2_trans
+        fs.f.RS1RR2 <- f.f.RS1RR2
+        # RR2 
+        fs.f.RR1SS2 <- f.f.RR1SS2
+        fs.f.RR1RS2 <- f.f.RR1RS2
+        fs.f.RR1RR2 <- f.f.RR1RR2
+```
+
+with something like :
+`fs <- f`
+
+where arrays are set up as :
+```{r}
+#to create an array a[sex][locus1][locus2]
+namesLoci <- c('SS','RS','RR')
+locus1 <- paste0(namesLoci,'1')
+locus2 <- paste0(namesLoci,'2')
+sex <- c("F","M")
+
+dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
+dim1 <- sapply(dimnames1, function(x) length(x))
+a <- array(0,dim=dim1, dimnames=dimnames1)
+a
+length(a)
+#to access one element
+a['M','SS1','RR2']
+#to check that males total 1
+sum(a['M',,])
+```
+
+Just need to work out how best to deal with cis & trans.
+This would be a less pleasing alternative
+```{r}
+
+namesLoci <- c('SS1SS2','SS1RS2','SS1RR2',
+               'RS1SS2','RS1RS2','RS1RR2',
+               'RR1SS2','RR1RS2','RR1RR2')
+
+#adding cis & trans
+namesLoci <- c('SS1SS2','SS1RS2','SS1RR2',
+               'RS1SS2','RS1RS2cis','RS1RS2trans','RS1RR2',
+               'RR1SS2','RR1RS2','RR1RR2')
+
+sex <- c("F","M")
+
+dimnames1 <- list( sex=sex, namesLoci=namesLoci )
+
+dim1 <- sapply(dimnames1, function(x) length(x))
+a <- array(0,dim=dim1, dimnames=dimnames1)
+a
+length(a)
+#to access one element
+a['M','SS1SS2']
+#to check that males total 1
+sum(a['M',])
+str(a)
+```
+
 ### TO DO  
 1. add myself as maintainer in DESCRIPTION
-1. sort location in package of "input.parameters.csv" functions should be able to run without it
+1. sort licence in description
+1. put back in checks for things that should sum to 1
+1. edit runModel carefully
+1. check whether functions can run without "input.parameters.csv" 
 1. check out the part about running sequential insecticide scenarios

### 2015-03-20 resistanceModelNotesAndy.Rmd
progress notes

@@
 **16/3/15** mon 18.30-19.30 1hr
-
 cut out lines 92 to 1740 into a runModel() function, later can subdivide it up
 
-In max(combmat[, gencol]) : no non-missing arguments to max; returning -Inf
-fixed by returning listOut from runModel()
-
 **17/3/15** tues to liverpool 8.30-5.30 9hrs
-
-done: add a savePlots argument, to enable plotting without saving
 
@@
 
-~60,000 combinations…. Too many???
+Approx. 60,000 combinations. Too many???
 
@@
 These ratios was then analysed by regression to find out how the input parameters alter the ratio, and hence under what circumstance combination would be best/worst.
-Classification tress can also be done….
+Classification tress can also be done.
 
@@
 
+**19/3/15** fri 9.45-14 4.25hrs 3.30-7 3.5hrs 7.75hrs
+
+checking whether I can run without the input file. Fixed now.
+
+```{r}
+resistanceMaster(params.csv = FALSE)
+```
+
+This is what the fitness scores that can be saved for each scenario as *two-locus_fitness-scores.csv look like :
+```
+	-,-	a,-	A,-	-,b	-,B	a,b	A,B	A,b	a,B
+SS1SS2	1	0	0	0	0	0	0	0	0
+SS1RS2	1	0	0	0	0	0	1.86E-05	0	0
+SS1RR2	1	0	0	0	0	0	0.1161	0	0
+RS1SS2	1	0	0	0	0	0	0	0	0
+RS1RS2	1	0	0	0	0	0	2.13E-05	0	0
+RS1RR2	1	0	0	0	0	0	0.132913	0	0
+RR1SS2	1	0	0	0	0	0	0	0	0
+RR1RS2	1	0	0	0	0	0	3.44E-05	0	0
+RR1RR2	1	0	0	0	0	0	0.215	0	0
+```
+
+Starting resistSimple() to be able to run single simpleScenarios.
+
 ### TO DO
-1. add myself as maintainer in DESCRIPTION
-1. sort licence in description
+
+1. check whether setting input[13,1] <- to the next line at 58 in createInputMatrix() was a bug
+
+1. finish tidying documentation of reading in input vars in createInputMatrix() add avr names from runModel()
++ can use these later for roxygen doc of function inputs to resistSimple
++ and I could use find&replace to rename later if needed
+
+1. sort that resistSimple will only work with calibration of 1012
+
+1. make it easier to run a single scenario and default scenarios
++ might do this by first adding a function rather than changing existing stuff
+1. i think the main visual output is the frequency of resistance over time
+1. start creating a vignette to help me understand how the model works
+
+1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
 1. put back in checks for things that should sum to 1
 1. edit runModel carefully
-1. check whether functions can run without "input.parameters.csv" 
 1. check out the part about running sequential insecticide scenarios
@@
 ### suggestion from Beth
-looking at it now I realise this is not ideal
 a solution may be to provide a .csv file with the Curtis scenarios set up
 and then the need to set the calibration number and params.csv in the R script can be removed altogether
-so it is all set up in the .csv file - I think similar can be done with produce.plots
-so the only bit of R code the user will need to change is the setwd() and the read.csv()
+the only bit of R code the user will need to change is the setwd() and the read.csv()
+
+### variable naming conventions
+**SS1, RS1, RR1** homozygous susceptible, heterozygous or homozygous resistant genotypes at each locus.   
+With 2 loci there are ten possible genotypes (including cis & trans forms of the double heterozygous).
+
+** A, B, a, b, -** 2 insecticides at 3 levels; low(ab), high(AB), absent(--) resulting in 9 niches.
+
+Locus1 relates to resistance for insecticide A, and locus2 for B. Locus2 does not effect exposure to A.
+
+```
+colnames(fitness) <- c( "-,-", "a,-", "A,-", "b,-", "B,-", "a,b", "A,B", "A,b", "a,B" )
+rownames(fitness) <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
+                        "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
+                        "RR1SS2", "RR1RS2", "RR1RR2")
+```

### 2015-03-21 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-Starting resistSimple() to be able to run single simpleScenarios.
+Starting resistSimple() to be able to run single simple Scenarios.
+
+**19/3/15** sat 10.15-2.15 4hrs
+
+[recent popgen hack](https://github.com/NESCent/r-popgen-hackathon/wiki)
+Doesn't seem that useful for this project.
+
+Good progress :
+check out e.g.
+`resistSimple(P_1=0.01,P_2=0.3)`
 
@@
 
-1. check whether setting input[13,1] <- to the next line at 58 in createInputMatrix() was a bug
+1. create setInputOneScenario to enable setting inputs for a single scenario
++ sensible defaults
++ later will allow customised scenarios e.g. setInputCurtisFig1
++ will reduce code volume and repetition in createInputMatrix
 
@@
 + and I could use find&replace to rename later if needed
+
+1. check whether setting input[13,1] <- to the next line at 58 in createInputMatrix() was a bug
 

### 2015-03-23 resistanceModelNotesAndy.Rmd
progress notes
noting progress

@@
 
-**19/3/15** sat 10.15-2.15 4hrs
+**20/3/15** sat 10.15-2.15 4hrs
 
@@
 
+**22/3/15** mon 9.45
+
 ### TO DO
+
+sort this
+To git@github.com:AndySouth/resistance.git
+ ! [rejected]        master -> master (non-fast-forward)
+error: failed to push some refs to 'git@github.com:AndySouth/resistance.git'
+hint: Updates were rejected because the tip of your current branch is behind
+hint: its remote counterpart. Integrate the remote changes (e.g.
+hint: 'git pull ...') before pushing again.
+hint: See the 'Note about fast-forwards' in 'git push --help' for details.
 

### 2015-03-23 resistanceModelNotesAndy.Rmd
progress notes

--- 2015-03-21 resistanceModelNotesAndy.Rmd
@@
 Sensitive heterozygous fitness:0, 0.2, 0.4 (n=3 for each insecticide) **?27 & 29**  
-Resistance allele dominance: 0, 0.2, 0.5, 0.9, 1 for each locus (n=5 for each locus) **35,38 do to start**  
+Resistance allele dominance: 0, 0.2, 0.5, 0.9, 1 for each locus (n=5 for each locus) **32, 35 do to start**  
 Fitness costs: 0, 0.05, 0.1 (n=3) **43&44**  
@@
 ```{r}
+require(devtools)    
+install_github('AndySouth/resistance') 
+library(resistance)
 resistanceMaster(params.csv = FALSE)
@@
 
-**19/3/15** sat 10.15-2.15 4hrs
+**19/3/15** sat 10.15
 
@@
 Doesn't seem that useful for this project.
+
+Created setInputOneScenario to enable setting inputs for a single scenario
++ sensible defaults
++ later will allow customised scenarios e.g. setInputCurtisFig1
++ will reduce code volume and repetition in createInputMatrix
 
@@
 
+**22/3/15** mon 9.45-11.45 2hrs 5-6
+
+created a modified plot allele freq function to be able to see overlapping lines
+
+Default settings :
+```{r}
+resistSimple()
+```
+
+Reducing dominance coefficient of locus1 from 1 shifts curves to right & separates locus 1 & 2
+```{r}
+resistSimple(h.RS1_A0=0.17)
+```
+
+But then setting dominance coefficient of locus2 to anything less than 0.7 grounds the curves (actually by pushing them further to right).
+```{r}
+resistSimple(h.RS1_A0=0.17,h.RS2_0B=0.7)
+```
+
+Reducing the selection coefficient for locus1 in A (from 1) pushes curves to right, but strangely doesn't separate locus 1 & 2 ??
+```{r}
+resistSimple(s.RR1_A0=0.17)
+```
+
+Put variable names into the suggested sensitivity analysis table.
+
+Thinking about creating a shiny app(s) within the resistance package.
+[Advice from RStudio](http://rstudio.github.io/shiny/tutorial/#deployment-local)
+
+Put your Shiny application directory under the package’s inst directory, then create and export a function that contains something like this:
+
+`shiny::runApp(system.file('appdir', package='packagename'))`
+
+I might be able to use something like [this](http://shiny.rstudio.com/gallery/plot-plus-three-columns.html) to put inputs in columns.
+
+Or perhaps better to keep it to a single column to start ?
+
+Starting frequency of resistance
+Locus1
+Locus2
+
+Exposure to insecticide (same for M&F & in Curtis for both insecticides)
+
+Fitness of susceptibles in presence of insecticide.
+Insecticide1
+Insecticide2
+
+Dominance of resistance
+Locus1
+Locus2
+
+Selective advantage of resistance
+Locus1
+Locus2
+
+It could perhaps be 2 columns, Locus1 & Locus2
+
+Or to fit better with widget titles, 2 rows (l1&2) with 5 columns.
+
 ### TO DO
 
-1. create setInputOneScenario to enable setting inputs for a single scenario
-+ sensible defaults
-+ later will allow customised scenarios e.g. setInputCurtisFig1
-+ will reduce code volume and repetition in createInputMatrix
+1. look at modifying other parameters
+
+1. start creating a shinyapp and/or vignette to help me understand how the model works
 
@@
 
-1. sort that resistSimple will only work with calibration of 1012
-
 1. make it easier to run a single scenario and default scenarios
@@
 1. i think the main visual output is the frequency of resistance over time
-1. start creating a vignette to help me understand how the model works
 

### 2015-03-24 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-**19/3/15** sat 10.15
+**19/3/15** sat 10.15-2.15 4hrs
 
@@
 
-**22/3/15** mon 9.45-11.45 2hrs 5-6
+**22/3/15** mon 9.45-11.45 2hrs 5-8 5hrs
 
@@
 
+**23/3/15** tue 9.45-
+
+created function to run shiny UI from package, and add it to readme
+
+there was a bug that dominance had no effect, because I was setting `h.RS1_00` rather than `h.RS1_A0 & h.RS1_0B`
+
+Cool! Seems to be working well now.
+
 ### TO DO
 
-1. look at modifying other parameters
+1. create a vignette from first shiny app documenting how the model works
 
-1. start creating a shinyapp and/or vignette to help me understand how the model works
-
-1. finish tidying documentation of reading in input vars in createInputMatrix() add avr names from runModel()
-+ can use these later for roxygen doc of function inputs to resistSimple
-+ and I could use find&replace to rename later if needed
-
-1. check whether setting input[13,1] <- to the next line at 58 in createInputMatrix() was a bug
-
-1. make it easier to run a single scenario and default scenarios
-+ might do this by first adding a function rather than changing existing stuff
-1. i think the main visual output is the frequency of resistance over time
+1. finish tidying documentation of reading in input vars in createInputMatrix()
 
@@
 1. submit tilde bug report
-1. later, structure as R package, maybe move to a new github repo
 1. re-look at R popgen packages when I've got more of a handle on what we are doing.

### 2015-03-24 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-**23/3/15** tue 9.45-
+**23/3/15** tue 9.45-2.15 4.5hrs
 
@@
 
+Sent message to Ian & Beth to test installing from Github and the shinyUI.
+
 ### TO DO
+
+1. start to write script to run sensitivity analysis, does it want to work by creating a large input.csv file, or by setting input object directly.
+
+1. submit travel claim & maybe first invoice after creating hours spreadsheet
 
@@
 1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
-1. put back in checks for things that should sum to 1
 1. edit runModel carefully
 1. check out the part about running sequential insecticide scenarios
-1. start to write script to run sensitivity analysis, does it want to work by creating a large input.csv file, or by setting input object directly.
+
 1. spend some time developing my understanding of how the model works

### 2015-04-13 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-My first go at keeping notes like this as an rmarkdown file.
+Timelines and deliverables
+**Days 1 to 12:** 
+Read and assimilate the Curtis paper cited above.
+Run the LSTM code written by Levick, replicate the results from the Curtis paper, and prepare the code for sensitivity analysis by firstly refactoring into an R package and creating a Github repository.
+Identify appropriate parameter ranges and any rules that should be applied while sampling (e.g.
+male exposure to insecticide within homes should not be higher than for the more endophilic females).
+Run sensitivity analysis and present the results as classification trees showing what parameter combinations favour which deployment strategy.
+
+**Deliverables:**  the provision of sensitivity analyses and classification trees that will allow us to publish the model in a peer-reviewed journal.
+
+**Days 13 to 25.**  This component is to use the full model rather than the sub-model used to replicate the Curtis results.
+Liaise with IVCC to identify the policy question they consider the most pressing and the likely end users of the simulation package.
+Perform sensitivity analysis on the full model as above.
+Attempt to put a user-friendly interface onto the simulation package using the “Shiny R” package.
+
+**Deliverables:** sensitivity analysis of the complete model and a reasoned consideration (or delivery) of the feasibility of distributing the package with a graphical interface for use by users with no computing experience.
 
@@
 
-**17/3/15** weds in liverpool 8.30-6 9.5hrs
+**18/3/15** weds in liverpool 8.30-6 9.5hrs
 
@@
 
-**22/3/15** mon 9.45-11.45 2hrs 5-8 5hrs
+**23/3/15** mon 9.45-11.45 2hrs 5-8 5hrs
 
@@
 
-**23/3/15** tue 9.45-2.15 4.5hrs
+**24/3/15** tue 9.45-2.15 4.5hrs
 
@@
 
+**25/3/15** weds 12.30-13.45 1.25 2.30-4 1.5 2.75
+
+created a time record
+
+Barbosa paper methods: 
+The analysis was performed using R & package lhs. It does not
+allow for the specification of each variable distribution
+beforehand, so sampling was performed assuming a uniform
+distribution. Once the sample was generated, the
+uniform sample from a column (variable) could be transformed
+to the required distribution (Table 3) by using
+quantile functions (using the qtriangle comand in R).
+
+A data set of 3,000 replications was generated, with random
+parameter sets. Ten replicates of this procedure were performed as suggested
+in [24] to investigate the predictive precision of model
+using LHS as the sampling method. This was achieved
+by analysing each replicate separately and verifying that
+results were consistent across ten replicates.
+
+...
+
+Results included a counter-intuitive outcome that the
+inclusion of a synergist could lead to an increase in the
+rate of the spread of resistance (i.e.
+y > 1).
+Further investigation of this result was pursed by performing a logistic regression with a binary dependent variable (1 if y > 1 and 0 if y < 1), therefore quantifying how changes in the
+parameters values affect the odds of getting the unexpected
+outcome y > 1.
+
+...
+
+Classification trees are used to predict membership of
+cases in the classes of a categorical dependent variable
+(1 if y > 1 or 0 if y < 1) from their input parameters
+and were implemented using an algorithm that grows a
+binary tree [27]. At each internal node in the tree, a test
+is applied to the input parameters to identify the binary
+distinction which gives the most information about the
+class membership.
+
+These are the references that refer to the classification trees :
+27. Breiman L: Classification and Regression Trees. Belmont (Calif.): Wadsworth International Group; 1984: 358.
+28. Therneau T, Atkinson EJ: An introduction to recursive partitioning using the RPART routines 2011:1–67. [r.789695.n4.nabble.com/attachment/3209029/0/zed.pdf].
+
+seems that [recursive partitioning using rpart](http://r.789695.n4.nabble.com/attachment/3209029/0/zed.pdf) is the R software used.
+
+[rpart package](http://cran.r-project.org/web/packages/rpart/index.html)
+Recursive partitioning for classification, regression and survival trees. An implementation of most of the functionality of the 1984 book by Breiman et al.
+
+[rpart intro vignette](http://cran.r-project.org/web/packages/rpart/vignettes/longintro.pdf) updated version of the reference.
+
+First example from the vignette (p10). This works on a binary response yes/no. method = 'class' is for categorical.
+```{r}
+library(rpart)
+#uses stagec, a dataframe
+str(stagec)
+#creates a factor for response var to improve labelling
+progstat <- factor(stagec$pgstat, levels = 0:1, labels = c("No", "Prog"))
+#fit the model by adding the columns
+cfit <- rpart(progstat ~ age + eet + g2 + grade + gleason + ploidy, data = stagec, method = 'class')
+plot(cfit) #plots tree
+text(cfit) #labels tree
+```
+
+So that seems to suggest that I'll need to output a dataframe with one row per simulation, columns for outputs (generations until resistance reached) and inputs.
+Will need to run it by Ian exactly what the outputs are.
+
+p25 in vignette has an example with multiple categories.
+Seems there are methods that work on continuous data too, which might be our case ...
+
+**30/3/15** mon 10.30 - 11.30 1hr
+
+RSTMH digital methods conference.
+KDR insecticide resistance gene.
+Turning 'big' data about how resistance spreads into epidemiological models and then into DSS about control and eradication is very complex. 
+Malaria genomics: tracking a diverse and evolving parasite population
+Professor Dominic Kwiatkowski, University of Oxford and Wellcome Trust Sanger Institute
+Paul Ben Allistere, Panoptes web app
+https://www.malariagen.net/apps/ag1000g/phase1-AR2/index.html#genomebrowser
+
+**2/4/15** 1hr
+
+emails from Ian & Marlize Coleman about insecticide resistance gaming.
+
+Ian : Marlize Coleman has funding from B&MGF to build computer games of insecticide deployment policies and wants to tie them to the outputs of various IR models.
+
+Marlize : At the moment **we need someone to look at current mathematical models that have any relevance to insecticide resistance management and determine if there is any potential to use them, or just components of the model/s, to create scenarios for a IR game**.
+We planning to start development of the game in July so we need to get a game scope sorted pretty soon.
+It will also be important for this person to communicate and discuss these scenarios/simulations with the game development team to ensure that they have a good grasp of the backend mathematical detail.
+
+We want to expand on this version by adding a bit more complexity in the backend and we’d like to create a bit more flexibility e.g.
+choice of vectors, disease prevalence/incidence, interventions, specific learning outcomes of interest etc.
+
+If you are interested, we can discuss more detail.
+Probably 3 months full time or **6 months 2-3 days a week**.
+You might be better placed to tell me how much effort will be required.
+Perhaps more time in the first few months and less during game development for assistance to developers.
+
+Me to Ian : Will be good to talk to you about potential IR models after I've spoken to Marlize. Do you imagine that a simplified version of Curtis might be useful for her ?
+
+Ian : I definitely think the 2locus Curtis model should be included.
+I don't think it necessarily needs to be simple.
+We may restrict the user to certain parameter values then preload the results from all possible permutations.
+We can discuss later.
+BW Ian
+
+**13/4/15** mon 9.30 - 13.30 4hrs 2.30-6 3.5hrs 7.5hrs
+
+looking at new serious gaming project ... & can offset these hours on that
+
+publish shinyCurtis1 to shinyapps.io to enable sharing with Marlize
+using publish button from RStudio worked well
+It gives this warning on installation, but then does work from web.
+Warning message:
+In FUN(c("shiny", "resistance")[[2L]], ...) :
+  Package 'resistance' not available in repository or locally
+
+https://andysouth.shinyapps.io/shinyCurtis1/
+
+Options :
+1. run R code from within the game
+1. translate model modules into the game language
+1. save model outputs as input files that can modify game behaviour
+
+Points to discuss with Marlize :
+
+* I am mostly committed until end of June, but might be able to shuffle.
+If we know what they want from the design, we could work on the model behind to produce this after the game design has been started.
+* might they want to run actual model code or translate it ?
+* I don't know how much info there is out there on the effect on resistance of different insecticide applications, but am happy to research in discussion with Ian, and produce working versions.
+* I would be a good intermediary between the biologists and the coders, as I'm somewhere in between myself !
+
+Imperial malaria models :
+http://www1.imperial.ac.uk/resources/27F31FE6-4CA6-4C97-85BD-58B342BC2CDA/usingthemalariatoolssoftwareforesp.pdf
+http://www1.imperial.ac.uk/malariamodelling/toolsdata/tools/
+On download I got an error msg saying : "The program can't start because VCOMP120.DLL is missing from your computer. Try reinstalling the program to fix"
+
+http://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1000324
+Reducing Plasmodium falciparum Malaria Transmission in Africa: A Model-Based Evaluation of Intervention Strategies
+
 ### TO DO
+
+1. can I create a plotting method that can accept multiple scenarios
+
+1. look at the method used in the Barbosa paper
+1. ask Ian if he has the code for the Barbosa classification trees analysis.
 
@@
 
-1. submit travel claim & maybe first invoice after creating hours spreadsheet
+1.
+from the first stage of the sensitivity analysis I'll want to output a dataframe with 1 row per replicate containing both the inputs and the outputs.
+Getting it to put the inputs into a dataframe like this would be a good start.
+
+1. submit travel claim
 

### 2015-04-15 resistanceModelNotesAndy.Rmd
progress notes

@@
 Then in RStudio I had to set project options, build tools to package
-(so maybe I should have done just done detools::setup())
+(so maybe I should have done just done devtools::setup())
 
@@
 
+emailed them to see about the error
+
+**14/4/15** tue 11.45 - 14.15 2.5hrs 3-5.30 2.5 5hrs
+
+can I create a plotting method that can accept multiple scenarios ?
+initially it could just plot all scenarios as additional lines in the existing colours for locus & sex
+use `plotallele.freq.andy`
+might need to modify to accept more than one matrix of results outputs  
+I thought tricky bit might be to set plot bounds based on multiple lines (but actually the y is already hardcoded from 0-1 for allele frequency).
+Might need to be a bit careful with x because different scenarios can have different numbers of generations.
+
+done : fixed why `resistanceMaster()` currently fails. I suspect due to input scenario calibration combination, it used to work ??
+
+Error in 1:ddt_cutoff : result would be too long a vector
+In addition: Warning messages:
+1: In data.matrix(mat) : NAs introduced by coercion
+2: In min(which((amat[, r1col]) > 0.5)) :
+  no non-missing arguments to min; returning Inf
+
+the error comes from `plotcurtis_f2()`
+I suspect it's because the input.parameters.csv isn't the one compatible with producing the curtis plot.
+`plotcurtis_f2()` needs results matrices from 3 scenarios: 1 hch, 2 ddt, 3 combined. Yet the current csv I have in the package seems to have 7 scenarios
+Input,Curtis Fig 2 Comb - calibration,DDT high,DDT high,DDT high,HCH high,HCH high,HCH high  
+Why did it work before then ??
+Aha! it's because in `createInputMatrix()` this sets up the 3 Curtis scenarios if an input file is not specified :
+  `if( !params.csv && calibration == 1012 ){
+    input <- matrix( ncol=3, nrow=52 )
+    colnames(input) <- c("B (HCH)", "A (DDT)","Combination")`
+So maybe I should have no input csv as the default.
+
+setting params.csv=FALSE as the default for resistanceMaster() sorted it.
+
+Back to `plotallele.freq.andy`, I might be able easily to modify it to accept a list of matrices for multiple scenarios, as well as it's current single scenario mode.
+
+this gives me the number of generations per scenario
+`sapply(listOut$results,nrow)`
+70  70 160
+`max_gen <- max( sapply(listOut$results,nrow) )`
+
+done : create a plotting method that can accept multiple scenarios
+
+now try to use it to plot a parameter range.
+To run multiple scenarios one way would be to call `setInputOneScenario()` repeatedly to create single scenarios & then rbind these together into an input object that I can pass to `runModel()`
+
+```{r}
+
+input <- NULL
+vals <- seq(0,1,0.2) #create a range of input values
+for(i in vals)
+{
+ inputOneScenario <- setInputOneScenario( h.RS1_A0=i )  
+ input <- cbind(input, inputOneScenario)
+}
+
+listOut <- runModel(input)
+plotallele.freq.andy(listOut)
+
+```
+This is getting close to allowing the user to specify an input parameter and what range of that parameter you want to plot. I could then potentially create a UI allowing the user to do that.
+Would it be useful to put the above code into a function ? 
+Actually the above code is very close to what I would need for the sensitivity analysis.
+
+I could create a `setInputMultiScenario()` function that accepts a vector of values for a param, and creates an input object with all other params set to their default.
+It could maybe even accept multiple args, but then would need to return the n*n scenarios.
+Also this is going away from the way that Ian suggested doing it by smapling from a distribution, maybe I should try that route first ?
+
+In that case do I want to randomly change all params at once ?
+Also might be good to get it to output a file at the end of each 100 runs, so that if it does crash during a long run, everything is not lost.
+These could just be RDA files saved from the listOut object.
+Do I want it to output listOut() each time or will I process so that the outputs are smaller ?
+(Might be better to keep the raw outputs to save any confusion later.)
+
+The sensitivity analysis function may be called a bit like this :
+
+```
+#specify ranges for parameters, if a range is not specified the default will be used
+#could offer option to setother params at diff fixed values
+#this would only accomodate uniform distributions, OK for now
+#think about random seeds later
+#be careful about some params which must total one
+input <- setInputSensiScenarios( n=100, h.RS1_A0=c(0,1) )
+```
+
+Am I reinventing the wheel here, can anything else help me with this ?
+
+**15/4/15** wed 1.30-2.30 1hrs 3.45-
+
 ### TO DO
 
-1. can I create a plotting method that can accept multiple scenarios
+1. see setInputSensiScenarios
+1. start to write script to run sensitivity analysis, probably by creating a large input object rather than a large input.params.csv.
 
-1. look at the method used in the Barbosa paper
+1. write a function to calculate timeToThreshold for all scenarios based on listOut, allow user to specify which locus.
+
 1. ask Ian if he has the code for the Barbosa classification trees analysis.
-
-1. start to write script to run sensitivity analysis, does it want to work by creating a large input.csv file, or by setting input object directly.
 
@@
 Getting it to put the inputs into a dataframe like this would be a good start.
+
+1. add some defensive checks, e.g. for sensible input variables
 
@@
 and then the need to set the calibration number and params.csv in the R script can be removed altogether
-the only bit of R code the user will need to change is the setwd() and the read.csv()
 

### 2015-04-16 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-**15/4/15** wed 1.30-2.30 1hrs 3.45-
+**15/4/15** wed 1.30-2.30 1hrs 3.45-6.15 2.5 3.5hrs
+
+setInputSensiScenarios
+function to create sensitivity analysis scenarios, probably by creating a large input object.
+
+**16/4/15** thurs 10-14 4hrs 2.30-5 2.5 6.5hrs
+
+To get input params into 1 row per scenario I can transpose, but the result is a matrix and has no names.
+But I need to find a way of getting column names set.
+
+```{r}
+input <- setInputSensiScenarios( nScenarios=3, P_1=c(0,0.5), P_2=c(0,0.5) )
+tinput <- t(input)
+str(tinput)
+#num [1:3, 1:52] 100 100 100 100 100 100 1 1 1 0 ...
+```
+
+putting names into `setInputOneScenario()` fixed this.
+
+Having a problem with sensiAnalysis1()
+Error in eval(..1) : ..1 used in an incorrect context, no ... to look in
+
+This may be because setInputSensiScenarios() expects a range for each argument, whereas setInputOneScenario expects a single value for the same named args. (but I don't really think it is this)
+
+Also stackoverflow pointed that using c for a varname (and there is one in resistance) is dangerous because R can expect the function.
+
+replaced c with recomb_rate for "Recombination Rate" to avoid R conflicts with c() 
+previous ... error still remained
+
+Aha! i think problem is that I specify defaults in setInputSensiScenarios when I don't need to (and that stops them making it into the ...)
+
+cool yes, getting there.
 
@@
 
-1. see setInputSensiScenarios
-1. start to write script to run sensitivity analysis, probably by creating a large input object rather than a large input.params.csv.
+1. continue with getting classificatrion trees working in sensiAn1
+
+1. update R, new lengths() function may be useful for me
+but doesn't find latest version yet
+library(installr)
+updateR()
+
+1. in susana's script the 'outcome' is 0/1 if used in our case that would seem to lose a lot of information about the outcomes ? Even timeToFifty would be better for us.
 
 1. write a function to calculate timeToThreshold for all scenarios based on listOut, allow user to specify which locus.
-
-1. ask Ian if he has the code for the Barbosa classification trees analysis.
 
@@
 Getting it to put the inputs into a dataframe like this would be a good start.
+I think this may just be a case of transposing the input object ?
+
+1. try out classification trees code from Susana
+
+1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
 

### 2015-04-27 resistanceModelNotesAndy.Rmd
progress notes

@@
 
+**22/4/15** jury service 11.30 - 12.30 1 hr
+
+```{r}
+out <- sensiAn1( nScenarios = 10, h.RS1_A0=c(0.1,1))
+```
+
+**24/4/15** 4-5 1hr
+
+**27/4/15** 9.15-13.15 4hrs
+
+trying to test rpart in sensiAn1
+getting :
+cfit <- rpart(res ~ h.RS1_A0 + h.RS2_0B, data = forCT, method = 'class')
+Error in cbind(yval2, yprob, nodeprob) : 
+  number of rows of matrices must match (see arg 2)
+
+This was because there were some NAs in the results that can from where a resistance frequency of 0.5 was not reached.
+so I temp converted them to 999
+
+Now get :
+Error in plot.rpart(cfit) : fit is not a tree, just a root
+
+Is this because I haven't used enough samples or because it doesn't work with just 2 predictors ?
+
+Trying with more samples, still get same error :
+
+`inputAndResults <- sensiAn1( 100, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1) )`
+
+more samples and more inputs I do get a tree
+`inputAndResults <- sensiAn1(500, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))`
+
+Some of susanas code for finessing the tree :
+tree->rpart(outcome ~ h_n+ h_o+h_I+s_o+s_I+φ_o+φ_I+α_o+α_I+α_fn+α_mo+α_mI+α_mn+z+beta_f+β_m ,method='class',data=working_example,control=rpart.control(minsplit=50))
+
+# control=rpart.control(minsplit=50)
+# minsplit	the minimum number of observations that must exist in a node in order for a split to be attempted.
+# then prune the tree to avoid overfitting the data
+pruned_tree<- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+
 ### TO DO
 
-1. continue with getting classificatrion trees working in sensiAn1
+1. continue with getting classification trees working in sensiAn1
 

### 2015-04-29 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-**27/4/15** 9.15-13.15 4hrs
+**27/4/15** 9.15-13.15 4hrs 1.45-2.45 5hrs
 
@@
 
+**28/4/15** 10-14.30 4.5hrs 3-6 7.5hrs
+
+reading Ians recent manuscript update.
+I think this from p13 is key on what I need to do with the analysis :
+
+** The simulation output was the ratio of  the time (in generation) for resistance  to the ‘new’ insecticide  to reach a given allele threshold frequency when the ‘new’ insecticide was deployed as combination, compared to the time taken to reach the same threshold if deployed on its own; a ratio >1 indicates that combination is better.
+Five threshold frequencies were investigated: 0.05, 0.1, 0.2, 0.5, 0.8.
+**
+
+In order to do that I'd need to repeat each scenario for a single & combined insecticide treatment.
+Single is just the new rarer insecticide, Combined also has an older insecticide to which initial resistance is more common.
+
+Searching for the Curtis paper. Cited by 181 papers according to Google.
+https://www.google.co.uk/webhp?sourceid=chrome-instant&ion=1&espv=2&ie=UTF-8#q=%22theoretical%20models%20of%20the%20use%20of%20insecticide%20mixtures%22
+
+https://books.google.co.uk/books?id=ooHjBwAAQBAJ&pg=PA177&lpg=PA177&dq=%22theoretical+models+of+the+use+of+insecticide+mixtures%22&source=bl&ots=b3CPTQc5LL&sig=FCB5MCvX1xUjtfgog1z_dq1pH1Y&hl=en&sa=X&ei=bXw_VbriFoWpsAGnvYG4Ag&ved=0CEEQ6AEwBA#v=onepage&q=Curtis&f=false
+
+Pesticide Resistance in Arthropods (1990) edited by Richard Roush, Bruce E. Tabashnik
+
+these 2 chapters look useful :
+Roush, R. T., and J. C. Daly. 1990. The role of population genetics in resistance research
+and management, pp. 97-152.In: Pesticide resistance in arthropods. R. T. Roush and
+B. E. tabashnik (Eds.) Chapman and Hall, New York.
+
+Tabashnik, B. E. 1990. Modeling and evaluation of resistance management tactics, pp.
+153-182. In: Pesticide resistance in arthropods. R. T. Roush and B. E. Tabashnik
+(Eds.) Chapman and Hall, New York.
+
+page 131 talks about insecticide mixtures. Refers to some modelling by Roush (1989a)
+
+Roush, R. T. 1989. Designing resistance management programs: How can you choose?
+Pestic. Sci. 26:423-441.
+
+p467 in Insect Resistance Management: Biology, Economics, and Prediction edited by David W. Onstad (2013)
+seems to have some simulations addressing similar issues to us.
+
+https://books.google.co.uk/books?id=6hp384ZH0_kC&pg=PA453&dq=%22Modeling+and+evaluation+of+resistance+management+tactics%22&source=gbs_toc_r&cad=4#v=onepage&q=%22Modeling%20and%20evaluation%20of%20resistance%20management%20tactics%22&f=false
+
+Now have sorted passing user args to rpart in sensiAn1(). It does produce reasonable looking trees just by e.g.
+
+`inputAndResults <- sensiAn1(100, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))`
+
+**29/4/15** 10.30-12.30 2hrs 3.15-
+
+working on refactoring variables to arrays
+
 ### TO DO
 
-1. continue with getting classification trees working in sensiAn1
-
-1. update R, new lengths() function may be useful for me
-but doesn't find latest version yet
-library(installr)
-updateR()
+1. is there a way I can easily run single & combined deployments for the same scenario ?
+* is it `niche_00` and `niche_AB`
+* or in `a.m_AB` insecticide exposure male hi1 hi2
+* do I also need to change "Starting frequency of resistance" ?
+* perhaps I can set up a simple example input to send to Beth ?
+* or ask Beth to send me the input.csv used to produce tables 8 & 9 in 'Brief_full-genotype-tables.pdf'
+* maybe comparing the files in full-dom_CD.csv & SD in C:\Dropbox\Levick et al IR manuscript will help me.
+
+1. before May 15th move Spark and Glimmer shiny apps
+
+1.
+go through the paper and check which bits relate to which in the code.
+Maybe create a table ?
+Could I create the table from R ?
+but actually much of the functionality is still in runModel() and needs to be put into functions.
+
+1. Maybe create a model diagram perhaps from diagrammer ?
+
+1. maybe separate out classification trees from sensiAn1
 
@@
 
-1.
-from the first stage of the sensitivity analysis I'll want to output a dataframe with 1 row per replicate containing both the inputs and the outputs.
-Getting it to put the inputs into a dataframe like this would be a good start.
-I think this may just be a case of transposing the input object ?
-
-1. try out classification trees code from Susana
-
 1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
@@
 1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
-1. edit runModel carefully
+1. edit/refactor runModel carefully
 1. check out the part about running sequential insecticide scenarios
@@
 1. submit tilde bug report
-1. re-look at R popgen packages when I've got more of a handle on what we are doing.
 
 ### suggestion from Beth
-a solution may be to provide a .csv file with the Curtis scenarios set up
+could provide a .csv file with the Curtis scenarios set up
 and then the need to set the calibration number and params.csv in the R script can be removed altogether
+
+### How best to refactor variables to arrays ?
+Refactoring to arrays will reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions.
+cis/trans makes it slightly tricky, but not all variables have cis/trans
+only used for genotype frequencies (f) and in output matrices fitness & genotype 
+in the wild in Beths code :
+```
+fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
+fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
+```
+
+e.g. W doesn't have cis/trans
+I could refactor
+```{r}
+        # relaxed selection fitnesses
+        ## Males
+        W.m.SS1SS2 <- 0.1 
+        W.m.SS1RS2 <- 0.1
+        W.m.SS1RR2 <- 0.1
+
+        W.m.RS1SS2 <- 0.1
+        W.m.RS1RS2 <- 0.1  
+        W.m.RS1RR2 <- 0.1
+
+        W.m.RR1SS2 <- 0.1 
+        W.m.RR1RS2 <- 0.1 
+        W.m.RR1RR2 <- 0.1
+
+        ## Female
+        W.f.SS1SS2 <- 0.1
+        W.f.SS1RS2 <- 0.1
+        W.f.SS1RR2 <- 0.1
+
+        W.f.RS1SS2 <- 0.1
+        W.f.RS1RS2 <- 0.1 
+        W.f.RS1RR2 <- 0.1
+
+        W.f.RR1SS2 <- 0.1
+        W.f.RR1RS2 <- 0.1 
+        W.f.RR1RR2 <- 0.1
+# to :
+#to create an array a[sex][locus1][locus2]
+namesLoci <- c('SS','RS','RR')
+locus1 <- paste0(namesLoci,'1')
+locus2 <- paste0(namesLoci,'2')
+sex <- c("f","m")
+
+dimnames1 <- list( sex=sex, locus1=locus1, locus2=locus2 )
+dim1 <- sapply(dimnames1, function(x) length(x))
+W <- array(0,dim=dim1, dimnames=dimnames1)
+
+#to fill al elements as above
+W <- 0.1
+#check length
+length(W)
+#to access one element
+W['m','SS1','RR2']
+#to check that males total 1
+sum(a['m',,])
+```
 
@@
 ```
+#### inputs
+** a ** exposure to insecticide
+** phi ** fitness of one locus (baseline)
+** W.SS1_00 & W.SS2_00** fitness of SS in no insecticide
+** h ** dominance coefficient
+** s ** selection coefficient
+** z ** fitness cost of resistance allele in no insecticide
+** niche_ ** insecticide niche toggle
+
+#### calculated
+** W ** fitnesses by niche and sex
+** f ** genotype frequency
+** fs ** genotype frequency following selection
+** G ** gametes
+
+Different W fitness variables
+`W.RR1_00 : [locus][niche]` single locus
+`W.RR1SS2_0b :   [locus1][locus2][niche]` niche
+`W.m.SS1SS2 : [sex][locus1][locus2]` sex
+
+Question.
+What if `a.m_A0 > 0` but `niche_A0 == 0` ?
+The ifelse clauses allow niches to be toggled on/off, i.e. a fitness can be given for A0
+but if only the niche A,B is toggled on, the fitness scores for A0 and Ab will be set to 0

### 2015-04-30 resistanceModelNotesAndy.Rmd
progress notes

@@
 ```{r}
-require(devtools)    
-install_github('AndySouth/resistance') 
+#require(devtools)    
+#install_github('AndySouth/resistance') 
 library(resistance)
@@
 
-```{r}
+```
 out <- sensiAn1( nScenarios = 10, h.RS1_A0=c(0.1,1))
@@
 
-**29/4/15** 10.30-12.30 2hrs 3.15-
-
-working on refactoring variables to arrays
-
-### TO DO
-
-1. is there a way I can easily run single & combined deployments for the same scenario ?
-* is it `niche_00` and `niche_AB`
-* or in `a.m_AB` insecticide exposure male hi1 hi2
-* do I also need to change "Starting frequency of resistance" ?
-* perhaps I can set up a simple example input to send to Beth ?
-* or ask Beth to send me the input.csv used to produce tables 8 & 9 in 'Brief_full-genotype-tables.pdf'
-* maybe comparing the files in full-dom_CD.csv & SD in C:\Dropbox\Levick et al IR manuscript will help me.
-
-1. before May 15th move Spark and Glimmer shiny apps
-
-1.
-go through the paper and check which bits relate to which in the code.
-Maybe create a table ?
-Could I create the table from R ?
-but actually much of the functionality is still in runModel() and needs to be put into functions.
-
-1. Maybe create a model diagram perhaps from diagrammer ?
-
-1. maybe separate out classification trees from sensiAn1
-
-1. in susana's script the 'outcome' is 0/1 if used in our case that would seem to lose a lot of information about the outcomes ? Even timeToFifty would be better for us.
-
-1. write a function to calculate timeToThreshold for all scenarios based on listOut, allow user to specify which locus.
-
-1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
-
-1. add some defensive checks, e.g. for sensible input variables
-
-1. submit travel claim
-
-1. create a vignette from first shiny app documenting how the model works
-
-1. finish tidying documentation of reading in input vars in createInputMatrix()
-
-1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
-1. edit/refactor runModel carefully
-1. check out the part about running sequential insecticide scenarios
-
-1. spend some time developing my understanding of how the model works
-1. submit tilde bug report
-
-### suggestion from Beth
-could provide a .csv file with the Curtis scenarios set up
-and then the need to set the calibration number and params.csv in the R script can be removed altogether
-
-### How best to refactor variables to arrays ?
-Refactoring to arrays will reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions.
-cis/trans makes it slightly tricky, but not all variables have cis/trans
-only used for genotype frequencies (f) and in output matrices fitness & genotype 
-in the wild in Beths code :
-```
-fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
-fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
-```
+**29/4/15** 10.30-12.30 2hrs 3.15-5.15 4hrs
 
@@
 #to fill al elements as above
-W <- 0.1
+W[] <- 0.1
 #check length
@@
 #to check that males total 1
-sum(a['m',,])
+sum(W['m',,])
+```
+
+working on refactoring variables to arrays, made flexible createArray() function.
+good progress
+started runModel2() to contain the refactoring
+
+**30/4/15** 9.45-
+
+Need to be careful about different W fitness variables
+`W.RR1_00 : [locus][niche]` single locus (Wlocus)
+`W.RR1SS2_0b :   [locus1][locus2][niche]` niche (Wniche)
+`W.m.SS1SS2 : [sex][locus1][locus2]` sex (Windiv)
+
+Once I have array refactoring working by .
+```{r}
+
+    a <- createArray( sex=c('m','f'), niche1=c('0','a','A'), niche2=c('0','b','B') )
+    Wniche <- createArray( locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2'), niche1=c('0','a','A'), niche2=c('0','b','B') )    
+    Windiv <- createArray( sex=c('m','f'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2') )
+```
+I should be able to replace 
+```
+    W.m.SS1SS2 <- (a.m_00 * W.SS1SS2_00) + 
+      (a.m_a0 * W.SS1SS2_a0) + (a.m_A0 * W.SS1SS2_A0) + 
+      (a.m_0b * W.SS1SS2_0b) + (a.m_0B * W.SS1SS2_0B) + 
+      (a.m_ab * W.SS1SS2_ab) + (a.m_AB * W.SS1SS2_AB) + 
+      (a.m_Ab * W.SS1SS2_Ab) + (a.m_aB * W.SS1SS2_aB) 
+```
+with
+```{r}
+    Windiv['m','SS1','SS2'] <- sum( a['m',,] * Wniche['SS1','SS2',,])
+```
+To prove that the above works
+```{r}
+    #set constant
+    a['m',,] <- 2 
+    #set variable
+    Wniche['SS1','SS2',,] <- 1:9
+    Wniche['SS1','SS2',,]
+
+    #showing that each element gets multiplied by 2
+    a['m',,] * Wniche['SS1','SS2',,]
+
+    #show result of sum
+    sum( a['m',,] * Wniche['SS1','SS2',,])
+
+    Windiv['m','SS1','SS2'] <- sum( a['m',,] * Wniche['SS1','SS2',,])
+
+```
+Gives a result of 90 for just mSS1SS2 which is correct
+``` Windiv
+, , locus2 = SS2
+
+   locus1
+sex SS1 RS1 RR1
+  m  90   0   0
+  f   0   0   0
+
+, , locus2 = RS2
+
+   locus1
+sex SS1 RS1 RR1
+  m   0   0   0
+  f   0   0   0
+
+, , locus2 = RR2
+
+   locus1
+sex SS1 RS1 RR1
+  m   0   0   0
+  f   0   0   0
+```
+
+Or indeed might be able to replace the whole 120 lines of code from 510-630 with :
+No these give: 
+```
+#Error in a[, , ] * Wniche[, , , ] : non-conformable arrays
+    Windiv['m',,] <- sum( a['m',,] * Wniche[,,,])
+    Windiv['f',,] <- sum( a['f',,] * Wniche[,,,])
+    #or even just
+    Windiv[,,] <- sum( a[,,] * Wniche[,,,]) 
+```
+Just needs a little more thought ...
+It's because a needs to be applied repeatedly to each genotype.
+Maybe worry about that later ..
+
+Want to develop a sequential testing process, so that I can compare runModel() and runModel2().
+Maybe I could use testthat for it ?
+
+Just need to do this to start :
+`devtools::use_testthat()`
+
+see testRefactoring() which does things like this :
+
+```
+input <- setInputOneScenario()
+tst <- runModel(input)
+tst2 <- runModel2(input)
+input <- setInputSensiScenarios(20, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
+tst <- runModel2(input, produce.plots=FALSE)
+tst2 <- runModel(input, produce.plots=FALSE)
+identical(tst,tst2)
+```
+
+### TO DO
+
+1. is there a way I can easily run single & combined deployments for the same scenario ?
+* is it `niche_00` and `niche_AB`
+* or in `a.m_AB` insecticide exposure male hi1 hi2
+* do I also need to change "Starting frequency of resistance" ?
+* perhaps I can set up a simple example input to send to Beth ?
+* or ask Beth to send me the input.csv used to produce tables 8 & 9 in 'Brief_full-genotype-tables.pdf'
+* maybe comparing the files in full-dom_CD.csv & SD in C:\Dropbox\Levick et al IR manuscript will help me.
+
+1. before May 15th move Spark and Glimmer shiny apps
+1. I should have included unit testing on my CV
+
+1.
+go through the paper and check which bits relate to which in the code.
+Maybe create a table ?
+Could I create the table from R ?
+but actually much of the functionality is still in runModel() and needs to be put into functions.
+
+1. Maybe create a model diagram perhaps from diagrammer ?
+
+1. maybe separate out classification trees from sensiAn1
+
+1. in susana's script the 'outcome' is 0/1 if used in our case that would seem to lose a lot of information about the outcomes ? Even timeToFifty would be better for us.
+
+1. write a function to calculate timeToThreshold for all scenarios based on listOut, allow user to specify which locus.
+
+1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
+
+1. add some defensive checks, e.g. for sensible input variables
+
+1. submit travel claim
+
+1. create a vignette from first shiny app documenting how the model works
+
+1. finish tidying documentation of reading in input vars in createInputMatrix()
+
+1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
+1. edit/refactor runModel carefully
+1. check out the part about running sequential insecticide scenarios
+
+1. spend some time developing my understanding of how the model works
+1. submit tilde bug report
+
+### suggestion from Beth
+could provide a .csv file with the Curtis scenarios set up
+and then the need to set the calibration number and params.csv in the R script can be removed altogether
+
+### How best to refactor variables to arrays ?
+Refactoring to arrays will reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions.
+cis/trans makes it slightly tricky, but not all variables have cis/trans
+only used for genotype frequencies (f) and in output matrices fitness & genotype 
+in the wild in Beths code :
+```
+fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
+fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
 ```

### 2015-05-01 resistanceModelNotesAndy.Rmd
progress notes

@@
 
-**30/4/15** 9.45-
+**30/4/15** 9.45-13.30 3.75hrs 2.15-5.30 3.25hrs 6-7 8hrs
 
 Need to be careful about different W fitness variables
-`W.RR1_00 : [locus][niche]` single locus (Wlocus)
-`W.RR1SS2_0b :   [locus1][locus2][niche]` niche (Wniche)
-`W.m.SS1SS2 : [sex][locus1][locus2]` sex (Windiv)
+`W.RR1_00 : [locus,niche]` single locus (Wlocus)
+`W.RR1SS2_0b :   [locus1, locus2, niche]` niche (Wniche)
+`W.m.SS1SS2 : [sex, locus1, locus2]` sex (Windiv)
 
@@
 
+So now I can keep developing runModel2() and use Ctrl T to test I'm not breaking it.
+
+Think, does my Wniche contain too many elements at 81 ?
+i.e. does it contain impossible combinations ? I think I'm OK Beths code fills 9*9 groups of individual variables.
+
+`W.RR1SS2_0b` ... `*81 3*3*3*3` to: Wniche[locus1, locus2, niche1, niche2]
+
+I may need to add a genotype option to createArray() to cope with f that needs cis & trans
+
+If I convert niche_0B to niche[niche1,niche2] I can simplify further the calculation of 
+Two genotype fitnesses in two insecticide Niche
+by having a loop that goes through each niche.
+
+Coolio! reduced 250 lines of code to ~ 20 with this
+```
+    ## Two genotype fitnesses in two insecticide Niche ##
+
+    #!r to replace 250+ lines below
+    for( niche1 in dimnames(Wniche)$niche1)
+    {
+      for( niche2 in dimnames(Wniche)$niche2)
+      {
+        #if this niche toggled off set fitness to 0
+        if (niche[niche1,niche2] == 0)
+        {
+          Wniche[,,niche1,niche2] <- 0
+        } else{
+          #otherwise set fitness to product of the 2 loci
+          for( locus1 in dimnames(Wniche)$locus1)
+          {
+            for( locus2 in dimnames(Wniche)$locus2)
+            {    
+              Wniche[locus1,locus2,niche1,niche2] <- Wloci[locus1,niche1,niche2] * Wloci[locus2,niche1,niche2]
+            }
+          }          
+        }
+      }
+    }
+```
+
+Trying to get fitnesses from my array into the existing output object fbn at the end of runModel2
+
+       -,- a,- A,- -,b -,B a,b A,B A,b a,B
+SS1SS2   0   0   0   0   0   0   0   0   0
+SS1RS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+SS1RR2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+RS1SS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+RS1RS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+RS1RR2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+RR1SS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+RR1RS2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+RR1RR2  NA  NA  NA  NA  NA  NA  NA  NA  NA
+
+STill not quite working ...
+
+**1/5/15** 9-14.30 5.5hrs 3.15-6 2.75 8.25hrs
+
+Fitness matrix different in my refactored version. But do I know which is correct ?
+
+Browse[2]> fbn
+       -,- a,- A,- -,b -,B a,b A,B A,b a,B
+SS1SS2   1   0   0   0   0   0   0   0   0
+SS1RS2   1   0   0   0   0   0   0   0   0
+SS1RR2   1   0   0   0   0   0   0   0   0
+RS1SS2   1   0   0   0   0   0   0   0   0
+RS1RS2   1   0   0   0   0   0   1   0   0
+RS1RR2   1   0   0   0   0   0   1   0   0
+RR1SS2   1   0   0   0   0   0   0   0   0
+RR1RS2   1   0   0   0   0   0   1   0   0
+RR1RR2   1   0   0   0   0   0   1   0   0
+Browse[2]> fbn2
+       -,- a,- A,- -,b -,B a,b A,B A,b a,B
+SS1SS2   0   0   0   0   0   0   0   0   0
+SS1RS2   0   0   0   0   0   0   0   0   0
+SS1RR2   0   0   0   0   0   0   0   0   0
+RS1SS2   0   0   0   0   0   0   0   0   0
+RS1RS2   0   0   0   0   0   0   0   0   0
+RS1RR2   0   0   0   0   0   0   0   0   0
+RR1SS2   0   0   0   0   0   0   0   0   0
+RR1RS2   0   0   0   0   0   0   0   0   0
+RR1RR2   0   0   0   0   0   0   0   0   0
+
+Aha difference is because Wniche just has 0s in, may be a problem earlier in the program.
+Aha issue is that Wloci is not read in yet.
+
+Reading in of Wloci can also be refactored, but I first need to read in h,z,phi & s as arrays.
+
+Then I can try to refactor this :
+```
+    ## Calculated fitnesses ####
+
+    # absence of insecticide
+    ## fitness of SS in absence of insecticide is entered above as a parameter
+    W.RS1_00 <- 1 - (h.RS1_00 * z.RR1_00)
+    W.RR1_00 <- 1 - z.RR1_00
+
+    W.RS2_00 <- 1 - (h.RS2_00 * z.RR2_00)
+    W.RR2_00 <- 1 - z.RR2_00
+
+    # low levels of insecticide a
+    W.SS1_a0 <- 1 - phi.SS1_a0
+    W.RS1_a0 <- W.SS1_a0 + (h.RS1_a0 * s.RR1_a0)
+    W.RR1_a0 <- W.SS1_a0 + s.RR1_a0
+
+    # high levels of insecticide A
+    W.SS1_A0 <- 1 - phi.SS1_A0
+    W.RS1_A0 <- W.SS1_A0 + (h.RS1_A0 * s.RR1_A0)
+    W.RR1_A0 <- W.SS1_A0 + s.RR1_A0
+
+    # low levels of insecticide b
+    W.SS2_0b <- 1 - phi.SS2_0b
+    W.RS2_0b <- W.SS2_0b + (h.RS2_0b * s.RR2_0b)
+    W.RR2_0b <- W.SS2_0b + s.RR2_0b
+
+    # high levels of insecticide B
+    W.SS2_0B <- 1 - phi.SS2_0B
+    W.RS2_0B <- W.SS2_0B + (h.RS2_0B * s.RR2_0B)
+    W.RR2_0B <- W.SS2_0B + s.RR2_0B
+```
+
+h only has 6 values 2*3. Because it's only for heterozygous loci & niches.
+
+```
+    # h = dominance coefficient
+    h.RS1_00 <- input[32,i]
+    h.RS1_a0 <- input[33,i]
+    h.RS1_A0 <- input[34,i]
+
+    h.RS2_00 <- input[35,i]
+    h.RS2_0b <- input[36,i]
+    h.RS2_0B <- input[37,i]
+```
+
+Perhaps best to store as
+h[locusNum,niche1,niche2]
+
+This nearly works, but includes un-needed homozygous niches aa etc 
+createArray( loci=c('RS1','RS2'), niche1=c('0','a','A'), niche2=c('0','b','B') )
+
+I could add a niches arg to createArray()
+createArray( loci=c('RS1','RS2'), niches=c('00','a0','A0') )
+
+BUT no even that doesn't work. 
+Because it only needs to contain the insecticide niche associated with the locus.
+
+Actually it just needs :
+h[locusNum, nicheLevel] : where insecticide is no, lo, hi
+... seeing how that works.
+
+and what about z - fitness cost of resistance allele in no insecticide
+z only has 2 values
+    z.RR1_00 <- input[42,i]
+    z.RR2_00 <- input[43,i]
+z[locusNum]
+Because by definition it is the fitness cost of one allele in absence of the corresponding insecticide. (so `RR1_00` is a little misleading `z.R1_0` more like it)
+
+phi - fitness of SS in each insecticide/concentration 
+just 4 values
+    phi.SS1_a0 <- input[26,i]
+    phi.SS1_A0 <- input[27,i]
+
+    phi.SS2_0b <- input[28,i]
+    phi.SS2_0B <- input[29,i]
+
+phi[locusNum, nicheLevel] where nicheLevel is lo, hi
+
+I have Wloci being 54
+> str(Wloci)
+ num [1:6, 1:3, 1:3] 0 0 0 0 0 0 0 0 0 0 ...
+ - attr(*, "dimnames")=List of 3
+  ..$ loci  : chr [1:6] "SS1" "RS1" "RR1" "SS2" ...
+  ..$ niche1: chr [1:3] "0" "a" "A"
+  ..$ niche2: chr [1:3] "0" "b" "B"
+> length(Wloci)
+[1] 54
+
+but actually I think it only needs 1 niche that corresponds to the chosen locus.
+for locus1 it only has as and locus2 it only has bs
+
+look at how Wloci is used to think about how best to structure
+
+This is where it is used 
+`Wniche[locus1,locus2,niche1,niche2] <- Wloci[locus1,niche1,niche2] * Wloci[locus2,niche1,niche2]`
+I think I need to change from 
+`Wloci[locus1,niche1,niche2] * Wloci[locus2,niche1,niche2]`
+to
+`Wloci[locus1,niche1] * Wloci[locus2,niche2]`
+
+I might want to change niche names to no,lo,hi throughout. Having them different for each locus (0aA & 0bB) causes the code difficulty.
+
+But that is a bigger change making it more difficult to check that the code logic remains the same.
+
+A temporray solution could be to have a lookup from nicheLevel to niche.
+e.g. no,lo,hi to 0,a,A & 0,b,B for locus1 & 2 respectively
+
+Later maybe
+Wloci[(locusNum),allele=(SS,SR,RR),exposure=(no,lo,hi)]
+for now
+Wloci[loci=(SS1,SR1,RR1,SS2,SR2,RR2),exposure=(no,lo,hi)]
+
+s selection coeeficient : just 4 values
+    # s = selection coefficient
+    s.RR1_a0 <- input[38,i]
+    s.RR1_A0 <- input[39,i]
+
+    s.RR2_0b <- input[40,i]
+    s.RR2_0B <- input[41,i]
+
+It's the selection coefficient of resistance in presence of the corresponding insecticide (at lo or hi)
+s[locusNum,exposure]
+but exposure is only lo or hi. Would a -ve value here for no be the same as z (the fitness cost of resistance in absence of insecticide ?)
+
+cool I think I have generic createArray2 function working
+
+I'm getting closer to fbn & fbn2 being the same
+
+Differences in fbn2:
+1) The 00 column has 0s, in fbn it is all 1s
+2) Column aB has the values that should be in AB
+
+e.g. SS1RR2
+W.SS1RR2_00
+Wniche['SS1','RR2','0','0']
+
+Wloci[locus2,exposure2] == 0
+but
+W.RR2_00 == 1
+
+aha think due to typo here :
+Wloci[ paste0('RR',locusNum), 'no'] <- 1 - z[locusNum]
+
+hurrah fixed first problem, now on to 2)
+
+Wniche looks correct
+So i think it may be due to ordering of columns in fbn
+Which goes AB, Ab, aB, where I think I assume aB, Ab, AB
+
+Wniche['RR1','RR2',,]
+      niche2
+niche1 0 b B
+     0 1 0 0
+     a 0 0 0
+     A 0 0 1
+
+as.vector( Wniche['RR1','RR2',,])
+[1] 1 0 0 0 0 0 0 0 1
+
+It may be tricky for me to get output in exactly the same order as Beths.
+How can I use the niche names to name the output matrix ?
+
+Difference in orders
+colnames(fbn) <- c("-,-", "a,-", "A,-", "-,b", "-,B", "a,b", "A,B", "A,b", "a,B")
+my order is         00     a0     A0     0b     ab     Ab     0B    aB     AB
+
+which is 1,2,3,4,6,8,5,9,7
+
+I might just need to allow that the fitness list is different in my refactor test ?
+
+Still how can I name the output columns from the data ?
+Fixed !!
+
+Then I started to delete some of the old code, and found that it broke some things I didn't want to break yet.
+
+Particularly this which suggests I need to get into the cis/trans issue soon:
+
+W.bar.m <- Windiv['m']
+W.bar.m <- (f.m.SS1SS2 * W.m.SS1SS2) ...
+
+revert changes for now and come back to
+
 ### TO DO
+
+1. what to call createArray2() ? maybe namedArray, or arrayNamed() ? could even put back in the option to pass the fill value ? submit to stackoverflow ?
 
@@
 Different W fitness variables
-`W.RR1_00 : [locus][niche]` single locus
-`W.RR1SS2_0b :   [locus1][locus2][niche]` niche
-`W.m.SS1SS2 : [sex][locus1][locus2]` sex
+`W.RR1_00 : [locus, niche1, niche2]` single locus (Wloci)
+`W.RR1SS2_0b :   [locus1, locus2, niche1, niche2]` niche (Wniche)
+`W.m.SS1SS2 : [sex, locus1, locus2]` sex (Windiv)
+
+### refactoring, what I have done
+Replaced the following collections of individual variables with arrays :
+(To reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions. Also potentially allows it to cope with more loci and/or niches later).
+
+```
+a.m_a0 ... *18 2*3*3  to: a[sex, niche1, niche2]
+W.m.SS1SS2 ... *18 2*3*3 to:  Windiv[sex, locus1, locus2]
+W.RR1SS2_0b ... *81 3*3*3*3 to: Wniche[locus1, locus2, niche1, niche2]
+
+still todo
+W.RR1_00 etc. with Wloci[loci, niche1, niche2]
+
+because of cis/trans:
+f.m.SS1SS2 ... *19 2*3*3+1 to: f[sex, genotype]
+
+```
+Want to do:
+1. rename a to exposure
+1. rename W to ?fitness : fitnessLocus, fitnessNiche and fitnessIndiv
 

### 2015-05-20 resistanceModelNotesAndy.Rmd
notes

@@
 W.bar.m <- Windiv['m']
-W.bar.m <- (f.m.SS1SS2 * W.m.SS1SS2) ...
+W.bar.m <- (f.m.SS1SS2 * W.m.SS1SS2)
 
@@
 
+**5/5/15** 10.30 - 14 3.5hrs 2.45-6.15 7hrs
+
+removed refactored fbn code
+
+sent update email to Ian & Beth
+
+starting to look at cis/trans
+this is the first place they appear in the generation loop
+```
+# set genotype frequencies as variables
+# from genotype frequency matrix generated above from initial value of P (freq. of R allele)
+# f = frequency before selection
+
+# male
+f.m.SS1SS2 <- genotype.freq[1,]
+f.m.SS1RS2 <- genotype.freq[2,]
+f.m.SS1RR2 <- genotype.freq[3,]
+f.m.RS1SS2 <- genotype.freq[4,]
+f.m.RS1RS2_cis <- genotype.freq[5,]		### cis
+f.m.RS1RS2_trans <- genotype.freq[6,]	### trans
+```
+
+How is genotype.freq created ?
+```
+# set up genotype matrix - records frequencies of each of the 9 two locus genotypes each generation
+genotype <- matrix( nrow=max_gen, ncol=11 )
+colnames(genotype) <- c("gen", "SS1SS2", "SS2RS2", "SS1RR2", 
+                        "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
+                        "RR1SS2", "RR1RS2", "RR1RR2")
+## make.genotypemat function will use this data and make a matrix of the genotype frequencies
+## frequencies of genotypes before selection - in HW equilibrium and same in male and female
+## needs name of matrix and takes corresponding frequency of resistant allele in function call
+genotype.freq <- make.genotypemat ( P_1, P_2 )
+```
+
+Within make.genotypemat, currently seems that cis & trans are always set to the same ??
+```
+## two forms of RS1RS2
+mat[5,] <- ( loc1[2,]*loc2[2,] ) * 0.5 #cis
+mat[6,] <- ( loc1[2,]*loc2[2,] ) * 0.5 #trans
+
+#this is what genotype.freq looks like
+                      Freq.
+SS1SS2       9.960060e-01
+SS1RS2       1.994006e-03
+SS1RR2       9.980010e-07
+RS1SS2       1.994006e-03
+RS1RS2_cis   1.996002e-06
+RS1RS2_trans 1.996002e-06
+RS1RR2       1.998000e-09
+RR1SS2       9.980010e-07
+RR1RS2       1.998000e-09
+RR1RR2       1.000000e-12
+```
+
+aarg createArray2() fails if you pass named variables to it, e.g. sex=sex rather than sex=c('m','f')
+seems like mget might solve, but :
+mget('sex', inherits=TRUE)
+Error: value for ‘sex’ not found
+
+may be able to use this from SO
+get_args <- function () {
+as.list( match.call(
+    def = sys.function( -1 ),
+    call = sys.call(-1)) )[-1]
+
+}
+
+BUT from the commandLine it does seem to work ??
+```
+> sex2 = c('f','m')
+> createArray2( sex = sex2 )
+sex
+f m 
+0 0 
+```
+Ahhh so someohow it does work if the variable is in the Global environment.
+
+eventually by trial & error found this solution :
+  #fails
+  #dimnames1 <- lapply(listArgs,function(x){eval.parent(x,n=1)})
+  #works
+  #eval.parent(listArgs$sex)
+  #[1] "f" "m"
+  #fails
+  #dimnames1 <- lapply(listArgs,eval.parent)  
+  #works!! n=3 but I'm not sure why !!!
+  dimnames1 <- lapply(listArgs,function(x){eval.parent(x,n=3)})
+
+**11/5/15** 1.30-5.45 4.25hrs   
+checking current testthat difference
+changing from `expect_identical` to `expect_equal` sorted it 
+refactoring gamete calculations into createGametes() function
+
+**12/5/15** 9.45-14 4.25hrs 2.30-5.45 3.25 7.5hrs
+
+sorting random mating & maybe getting to nub of cis/trans ?
+
+Beths random mating code :
+```
+      f.m.SS1SS2 <- 0
+      f.m.SS1RS2 <- 0
+      f.m.SS1RR2 <- 0
+
+      f.m.RS1SS2 <- 0
+      f.m.RS1RS2_cis <- 0			#RS1RS2
+      f.m.RS1RS2_trans <- 0		#RS1SR2
+      f.m.RS1RR2 <- 0
+
+      f.m.RR1SS2 <- 0
+      f.m.RR1RS2 <- 0 
+      f.m.RR1RR2 <- 0
+
+      # SS male with SS female
+      f.m.SS1SS2 <- f.m.SS1SS2 + ( G.m.S1.S2 * G.f.S1.S2 )
+      # SS male with SR female
+      f.m.SS1RS2 <- f.m.SS1RS2 + ( G.m.S1.S2 * G.f.S1.R2 )
+      # SS male with RS female
+      f.m.RS1SS2 <- f.m.RS1SS2 + ( G.m.S1.S2 * G.f.R1.S2 )
+      # SS male with RR female
+      f.m.RS1RS2_cis <- f.m.RS1RS2_cis + ( G.m.S1.S2 * G.f.R1.R2 )
+
+      # SR male with SS female
+      f.m.SS1RS2 <- f.m.SS1RS2 + ( G.m.S1.R2 * G.f.S1.S2 )
+      # SR male with SR female
+      f.m.SS1RR2 <- f.m.SS1RR2 + ( G.m.S1.R2 * G.f.S1.R2 )
+      # SR male with RS female
+      f.m.RS1RS2_trans <- f.m.RS1RS2_trans + ( G.m.S1.R2 * G.f.R1.S2 )
+      # SR male with RR female
+      f.m.RS1RR2 <- f.m.RS1RR2 + ( G.m.S1.R2 * G.f.R1.R2 )
+
+      # RS male with SS female
+      f.m.RS1SS2 <- f.m.RS1SS2 + ( G.m.R1.S2 * G.f.S1.S2 )
+      # RS male with SR female
+      f.m.RS1RS2_trans <- f.m.RS1RS2_trans + ( G.m.R1.S2 * G.f.S1.R2 )
+      # RS male with RS female
+      f.m.RR1SS2 <- f.m.RR1SS2 + ( G.m.R1.S2 * G.f.R1.S2 )
+      # RS male with RR female
+      f.m.RR1RS2 <- f.m.RR1RS2 + ( G.m.R1.S2 * G.f.R1.R2 )
+
+      # RR male with SS female
+      f.m.RS1RS2_cis <- f.m.RS1RS2_cis + ( G.m.R1.R2 * G.f.S1.S2 ) 
+      # RR male with SR female
+      f.m.RS1RR2 <- f.m.RS1RR2 + ( G.m.R1.R2 * G.f.S1.R2 )
+      # RR male with RS female
+      f.m.RR1RS2 <- f.m.RR1RS2 + ( G.m.R1.R2 * G.f.R1.S2 )
+      # RR male with RR female
+      f.m.RR1RR2 <- f.m.RR1RR2 + ( G.m.R1.R2 * G.f.R1.R2 )
+```
+
+Beth only has 10 combinations (including cis/trans), this is because RS is usually same as SR, and in all the names R goes first.
+```
+genotypes <- c( "SS1SS2", "SS2RS2", "SS1RR2", 
+                "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
+                "RR1SS2", "RR1RS2", "RR1RR2")
+```
+
+for randomMating this loop creates 16 combinations
+including cis/trans
+
+```
+  counter <- 0
+
+  for( m2 in c('S2','R2'))
+  {
+    for( m1 in c('S1','R1'))
+    {
+      for( f2 in c('S2','R2'))
+      {
+        for( f1 in c('S1','R1'))
+        {
+          counter <- counter+1
+          cat(paste(counter, m1,f1,m2,f2,"\n"))
+        }
+      }
+    }
+  }
+```
+1 S1 S1 S2 S2 
+2 S1 R1 S2 S2 
+3 S1 S1 S2 R2 
+4 S1 R1 S2 R2 
+5 R1 S1 S2 S2 
+6 R1 R1 S2 S2 
+7 R1 S1 S2 R2 
+8 R1 R1 S2 R2 
+9 S1 S1 R2 S2 
+10 S1 R1 R2 S2 
+11 S1 S1 R2 R2 
+12 S1 R1 R2 R2 
+13 R1 S1 R2 S2 
+14 R1 R1 R2 S2 
+15 R1 S1 R2 R2 
+16 R1 R1 R2 R2
+
+reformat slightly :
+1 SS1 SS2
+2 SR1 SS2
+3 SS1 SR2
+4 SR1 SR2
+5 RS1 SS2
+6 RR1 SS2
+7 RS1 SR2
+8 RR1 SR2
+9 SS1 RS2
+10 SR1 RS2
+11 SS1 RR2
+12 SR1 RR2
+13 RS1 RS2
+14 RR1 RS2
+15 RS1 RR2
+16 RR1 RR2
+
+Perhaps I can do the calculations on these 16 & then aggregate to produce the 10 ?
+
+```
+#these are other ways of creating the 16
+
+#as an array
+arr <- createArray2(loc1all1=c('S','R'),loc1all2=c('S','R'),loc2all1=c('S','R'),loc2all2=c('S','R'))
+
+#as a dataframe
+dF <- expand.grid(loc1all1=c('S','R'),loc1all2=c('S','R'),loc2all1=c('S','R'),loc2all2=c('S','R'))
+> dF
+   loc1all1 loc1all2 loc2all1 loc2all2
+1         S        S        S        S
+2         R        S        S        S
+3         S        R        S        S
+4         R        R        S        S
+5         S        S        R        S
+6         R        S        R        S
+7         S        R        R        S
+8         R        R        R        S
+9         S        S        S        R
+10        R        S        S        R
+11        S        R        S        R
+12        R        R        S        R
+13        S        S        R        R
+14        R        S        R        R
+15        S        R        R        R
+16        R        R        R        R
+
+genotypes <- paste0(dF[,1],dF[,2]," ",dF[,3],dF[,4])
+#[1] "SS SS" "RS SS" "SR SS" "RR SS" "SS RS" "RS RS" "SR RS" "RR RS" "SS SR" "RS SR" "SR SR" "RR SR" "SS RR" "RS RR" "SR RR" "RR RR"
+
+> length(genotypes)
+[1] 16
+
+#or could miss out the space
+
+l1a1 <- substr(genotypes,1,1)
+l1a2 <- substr(genotypes,2,2)
+l2a1 <- substr(genotypes,4,4)
+l2a2 <- substr(genotypes,5,5)
+
+#these are homozygous at both loci and have 1:1 between expanded & contracted genotypes
+which(l1a1==l1a2 & l2a1==l2a2)
+[1]  1  4 13 16
+
+#this gets the cis/trans
+which(!(l1a1==l1a2 | l2a1==l2a2))
+[1]  6  7 10 11
+
+But I might not to be this fancy. If I do all the working in the expanded genotype storage, and then just have a single function to do the conversion ??
+
+genotypesContract()
+or genotypesLong2Short()
+
+```
+
+done : convert from the expanded genotype storage to the contracted one.
+
+Beware I do seem to have occasional small difference between runModel & runModel2. I set the random seed so that the difference is consistent.
+
+It now seems only to turn up if I have >20 scenarios.
+
+It's not to do with randomMating() or genotypesLong2Short because they are not fully implemented yet.
+
+Failure(@testRefactoring.r#21): refactoring named variables with arrays doesn't change results 
+runModel2(input, produce.plots = FALSE) not equal to runModel(input, produce.plots = FALSE)
+Component "results": Component 25: Mean relative difference: 6.391493e-08
+
+It does show some very small differences, but I suspect they are just rounding errors, e.g. this is how they start out. (and the M,F columns just contain 1 despite showing this difference here)
+
+```
+listOut2$results[[25]] - listOut$results[[25]]
+
+       Gen          m.R1          m.R2          m.LD          f.R1          f.R2          f.LD             M             F
+[1,]   0  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00
+  [2,]   0  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00
+  [3,]   0  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00 -2.220446e-16 -2.220446e-16
+  [4,]   0  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00  0.000000e+00
+  [5,]   0 -2.168404e-19 -2.168404e-19  0.000000e+00 -2.168404e-19 -2.168404e-19  0.000000e+00 -1.110223e-16 -1.110223e-16
+```
+
+tested that my new W.bar calc works :
+W.bar
+        m         f 
+0.1000036 0.1000036 
+Browse[2]> W.bar.m
+   a.m_00 
+0.1000036 
+Browse[2]> W.bar.f
+   a.f_00 
+0.1000036
+
+**13/5/15** 10-14 2
+
+dealing with cis/trans :
+```
+namesLoci <- c('SS1SS2','SS1RS2','SS1RR2',
+               'RS1SS2','RS1RS2cis','RS1RS2trans','RS1RR2',
+               'RR1SS2','RR1RS2','RR1RR2')
+#or
+namesLoci <- rownames( genotype.freq )
+sex <- c("F","M")
+f <- createArray2( sex=sex, loci=namesLoci )
+#i then may be able to refactor more later
+```
+Plus see randomMating() which uses something like below, and then genotypesLong2Short.r to convert from the expanded genotype format.
+```
+  counter <- 0
+
+  for( m2 in c('S2','R2'))
+  {
+    for( m1 in c('S1','R1'))
+    {
+      for( f2 in c('S2','R2'))
+      {
+        for( f1 in c('S1','R1'))
+        {
+          counter <- counter+1
+          cat(paste(counter, m1,f1,m2,f2,"\n"))
+        }
+      }
+    }
+  }
+```
+
+deleting old code from runModel2()
+
+* done to f[m,loci] : sort storage of f.m.RS1RS2_cis
+* done : refactor gamete calculations : G.m.R1.R2 <- G.m.R1.R2 + f.m.RS1RS2_cis ...
+* done : sort calc of W.bar.m from Windiv
+* done : remove reliance on W.m.SS1SS2 etc. (Windiv)
+* done : remove reliance on a.m_00 etc.
+
+So far down from 1590 lines to 685, with only a bit in new functions
+
+**20/5/15** 11- in liverpool meeting with Beth and train back
+
+New bit from the manuscript by Ian about including sex-linked genes by a bit of trickery.
+
+It is also possible to allow one of the loci to be sex-linked; here we assume that sex linkage is at Locus 1, and ‘conventional’ sex-determination i.e.
+that the females is hemozygous sex (XX), the male is XY, and that the gene exists only on the X chromosome.
+This option is important as Anopheles gambiae has only 2 autosomes and one sex chromosome; approximately ??% of the genes are on the X in this species.
+
+The female genotype frequencies are as given on Table S1 and as described in Equation 4 and Equation 5  because females have two copies of the X chromosome.
+Locus 1 is homozygous in the male so heterozygotes are impossible at this locus; in this case the allele inherited by males at locus 1  is the maternal-derived one (because they get their X chromosome from their mother and the Y from the father).
+Hence the male genotypes can be derived by a process similar to that of females and autosomes as shown on Table S2.
+The males will be simulated  as “RR” or “SS” at the locus even though, in reality” they will be either “R-“ or “S-“.
+An inherent assumption is therefore that the hemizgous males “R-“ and “S-“ have the same fitness as the equivalent “RR” and “SS” females; this seems reasonable given dosage compensation and is an assumption commonly made.
+
+The male “diploid” genotypes are derived by a similar process to that above and the genotypes are as given on Table S2.
+The calculation become slightly more complicated to those given above but the same process occurs i.e.
+Equation 4 becomes
+
+Equation 6
+
+The next genotype can be produced by two combinations so Equation 5 becomes
+
+Equation 7 ...
+
+IH thinks this short-cut will work because selection will be OK, and the males will also correctly pass on their maternal X-linked allele (Table 4). We need to discuss this.
+
+It just means we need edit the code slightly.
+i.e.
+calculate the female genotypes then if locus 1 in autosomal just set the male frequencies equal to the females (as we do already), else re-calculate the male frequencies according to Table S2.
+
+end of bit from manuscript
+I think this is the important part for implementation :
+
+we assume that sex linkage is at Locus 1, females XX, the male XY, 
+and that the gene exists only on the X chromosome.
+
+If sex-linked Locus 1 is homozygous in the male so heterozygotes are impossible at this locus 
+the allele inherited by males at locus 1 is the maternal-derived one 
+(because they get their X chromosome from their mother and the Y from the father). 
+males will be simulated  as RR or SS at the locus even though, in reality they will be either R- or S-
+
+I think this change needs to be made in the randomMating() function.
+Currently i think randomMating() doesn't differentiate between the sex of the offspring.
+
+Does sex linkage have further implications in terms of the generation of gametes ? maybe not
+
 ### TO DO
 
+1. try to stop man pages being tracked by git, thy ar already in git ignore but I think I may need to do git delete or something delete them from the repo to stop git from tracking them
+1. further develop sex-linkage in randomMating(), allow option to be passed as arg, plus may need to look at calculating genotypes of both m&f in runModel2()
+
+Remember.
+It is a 2 locus model. One locus for each insecticide. At each locus there are 2 alleles. This gives a total of 9 genotypes if the position of the alleles is ignored.
+
+Good simple test
+```
+input <- setInputOneScenario(max_gen=5)
+tst2 <- runModel2(input)
+tst <- runModel(input)
+```
+
+1. submit stackoverflow question about createArray2(), including about n=3 in eval.parent
 1. what to call createArray2() ? maybe namedArray, or arrayNamed() ? could even put back in the option to pass the fill value ? submit to stackoverflow ?
@@
 
-1. before May 15th move Spark and Glimmer shiny apps
-1. I should have included unit testing on my CV
+1. I should include unit testing on my CV
 
@@
 `W.RR1SS2_0b :   [locus1, locus2, niche1, niche2]` niche (Wniche)
-`W.m.SS1SS2 : [sex, locus1, locus2]` sex (Windiv)
+`W.m.SS1SS2 : [sex, locus1, locus2]` (Windiv)
 
@@
 
-Question.
-What if `a.m_A0 > 0` but `niche_A0 == 0` ?
+Questions.
+1. check on calibration 104
+1. What if `a.m_A0 > 0` but `niche_A0 == 0` ?
 The ifelse clauses allow niches to be toggled on/off, i.e. a fitness can be given for A0

### 2015-05-22 resistanceModelNotesAndy.Rmd
notes

@@
 
-**20/5/15** 11- in liverpool meeting with Beth and train back
+**20/5/15** 11-4 in liverpool meeting with Beth and train back
 
@@
 
+**22/5/15**
+To stop things being tracked by git that you have just added to gitignore
+```
+git rm --cached `git ls-files -i -X .gitignore` 
+``` 
+Then I just did commit from RStudio.
+
+1. somehow I've broken the check
+Failure(@testRefactoring.r#12): refactoring named variables with arrays doesn't change results 
+runModel2(input, produce.plots = FALSE) not equal to runModel(input, produce.plots = FALSE)
+Component "results": Component 1: 'is.NA' value mismatch: 186 in current 0 in target
+Its because of Warning messages like this:
+1: In runModel2(input) : Male frequencies before selection total != 1 0.997966163681722
+
+Perhaps I need to make the checks less sensitive ??
+Is it because I re-enabled checks that had been disabled by Beth ?
+I reduced the tolerance in checks, runs OK, but tests still fail.
+
+Points to some difference having crept in ...
+
+difference is in 
+genotype outputs in generations after 1
+all those starting with RS1
+
+to do with object genotype ....
+Aha! problems were due to my new sex-linked changes that were operating without me realising.
+So I should probaly remove the tolerances I put in. and work out how to reassign the impossible genotype freqs.
+
 ### TO DO
 
-1. try to stop man pages being tracked by git, thy ar already in git ignore but I think I may need to do git delete or something delete them from the repo to stop git from tracking them
+1. check on implementation of sexLinkage in randomMating()
+
+1.
+think I can remove genotype.freq it is the same as f['m',], I might want to create an fstart[] outside of the generations loop.
+This was prompted because previously genotype.freq was the same for males and females and it adds an unecessary extra layer of potential confusion
+
 1. further develop sex-linkage in randomMating(), allow option to be passed as arg, plus may need to look at calculating genotypes of both m&f in runModel2()

### 2015-06-04 resistanceModelNotesAndy.Rmd
progress notes

@@
 So I should probaly remove the tolerances I put in. and work out how to reassign the impossible genotype freqs.
+done
+
+**4/6/15** thurs 12.15-
+
+checking on implementation of sexLinkage in randomMating()
+
+sexLinked probably needs to be included as an extra paramter in input (but should make it robust to cope if an input file doesn't have the sex-linked option)
+
+be careful because this could break previous model runs
+
+genotype.freq mostly removed it was the same as f['m',].
+Creation of f & fs moved to before the generations loop.
+This was prompted because previously genotype.freq was the same for males and females and it adds an unecessary extra layer of potential confusion.
+
+added calculating genotypes of both m&f in runModel2() to allow sex linkage.
+
+bug in random mating with sex linkage genotypes aren't summing to 1
+
+fGenotypeExpanded for non sex-linked looks like the top one
+In the lower one the R1 homozgotes should sum to 1 too
+
+    l1a2
+l1a1     S1     R1
+  S1 0.0625 0.0625
+  R1 0.0625 0.0625
+
+    l1a2
+l1a1    S1     R1
+  S1 0.125 0.0000
+  R1 0.000 0.0625
+
+Fixed now :
+with this bit of code :
+```
+#heterozygotes at locus1 are impossible so add to the homozygotes instead
+if(f1!=m1)
+{
+  #add to the homozygotes : f1,f1
+  fGenotypeExpanded[f1,f1,f2,m2] <- fGenotypeExpanded[f1,f1,f2,m2] + fThisGenotype
+} else #i.e. if not heterozygous at locus 1
+{
+  #just add to this genotype : f1,m1
+  #have to add rather than set in case this is a homozygote that has already been added to by previous condition
+  fGenotypeExpanded[f1,m1,f2,m2] <- fGenotypeExpanded[f1,m1,f2,m2] + fThisGenotype
+}
+```
+
+Created a new shiny app in shinyCurtis2 and run from runUI2.R to look at graphs with & without sex linkage side by side.
+Initially these gave identical results, which was because resistSimple() called the old runModel() which doesn't allow sex linkage.
+Try it again now.
+
+Cool, now I am getting a difference between locus1 & locus2. However males & females still have the same results, is that what would be expected ?
+
+Maybe it's because of the way that results are output from Beths one sex system ?
+Where does the graph get the data from ?
+
+genplot <- plotallele.freq.andy( listOut$results[[1]] )
+
+  # Males
+  lines( mat[,1], mat[,2], col="darkblue", lwd=2, lty=13 )
+  lines( mat[,1], mat[,3], col="green", lwd=2, lty=14 )
+  # Females
+  lines( mat[,1], mat[,5], col="red",lwd=2, lty=15 )
+  lines( mat[,1], mat[,6], col="orange", lwd=2, lty=16 )
+
+so its columns 2,3,5 & 6 I'm interested in
+
+so where is listOut$results[[1]] filled ?
+in runModel2()     listOut$results[[i]] <- results
+& further back :
+```
+      ## frequency of resistance alleles
+      m.R1 <- sum(f['m',grep("RR1",colnames(f))]) + ( 0.5 * sum(f['m',grep("RS1",colnames(f))]))
+      m.R2 <- sum(f['m',grep("RR2",colnames(f))]) + ( 0.5 * sum(f['m',grep("RS2",colnames(f))]))
+      f.R1 <- sum(f['f',grep("RR1",colnames(f))]) + ( 0.5 * sum(f['f',grep("RS1",colnames(f))]))
+      f.R2 <- sum(f['f',grep("RR2",colnames(f))]) + ( 0.5 * sum(f['f',grep("RS2",colnames(f))]))   
+      results[k,2] <- m.R1
+      results[k,3] <- m.R2
+      results[k,5] <- f.R1
+      results[k,6] <- f.R2
+```
+
+Seems that this should be doing the right thing. Maybe the allele frequencies overall don't change, because for the males they just redistributed from heterozygotes to homozygotes.
 
@@
 
-1. check on implementation of sexLinkage in randomMating()
-
-1.
-think I can remove genotype.freq it is the same as f['m',], I might want to create an fstart[] outside of the generations loop.
-This was prompted because previously genotype.freq was the same for males and females and it adds an unecessary extra layer of potential confusion
-
-1. further develop sex-linkage in randomMating(), allow option to be passed as arg, plus may need to look at calculating genotypes of both m&f in runModel2()
+1. be careful that runModel & runModel2 should no longer generate same results when sex linkage is set (but i don't think this should break current testthat tests)
 
@@
 but actually much of the functionality is still in runModel() and needs to be put into functions.
-
-1. Maybe create a model diagram perhaps from diagrammer ?
 

### 2015-06-08 resistanceModelNotesAndy.Rmd
progress notes

@@
 
+**5/6/15**
+
+Instructions emailed to Ian & Beth to run UI2
+require(devtools)
+install_github('AndySouth/resistance')
+require(resistance)
+runUI2()
+
+Ians feedback on UI2 (sex-linkage)
+
+The R script runs fine. The results are as I would have expected given that locus 1 is sex-linked and locus 2 is autosomal i.e. resistance evolves faster at the sex-linked locus.
+
+The dynamics at locus 2 appear unchanged...I therefore assume the drugs were deployed individually i.e. not as a combination?
+
+But I thought they were applied as a combination.
+The exposure comes from here in the UI :
+
+ h5("Exposure to each insecticide"),
+ numericInput("a.m_AB", "same for both insecticides in Curtis", 0.5, min = 0.1, max = 0.9, step = 0.1)
+
+and is set here in server :
+a.m_AB = input$a.m_AB,
+a.f_AB = input$a.m_AB, #set f to same as m
+
+are these the correct parameters to set ??
+yes i think so.
+
+Thanks Ian,
+In relation to your 2nd comment, from the UI, both a.m_AB and a.f_AB
+are set from the box with "Exposure to each insecticide".
+
+It's my understanding that should result in exposure to high levels of
+both insecticides (i.e. in combination).
+
+Is that correct Beth ?
+
+Didn't get a reply from Beth yet.
+From Ian :
+OK, I was just speculating that if resistnce at locus 1 was spreading faster due to sex-linkage then so might resistnce at locus2 through genic linkage and hitch-hiking.
+I'll give its some more thought over the weekend.
+
+**8/6/15** on train to liverpool 4.30-9.30
+**The plan for running the sensitivity analysis for the first paper**
+
+Edited from Ians text in the paper.
+
+For each parameter combination 2 strategies : 
+A) single ‘new’ insecticide alone
+B) ‘new’ insecticide in combination with ‘older’ insecticide
+
+The simulation output was the ratio of number of generations for resistance  to the ‘new’ insecticide  to reach a given allele frequency when in combination compared to on its own.
+
+A ratio >1 indicates that combination is better.
+
+Five threshold frequencies were investigated: 0.05, 0.1, 0.2, 0.5, 0.8.
+
+But in the document :
+"Suggested sensitivity analysis.docx" it talks about sequential and 2 mixture optoins.
+
+Sequential use:, use one insecticide until resistance frequency reaches critical point, then switch insecticide until second resistance allele frequency reaches the critical point.
+Record time in total generations for the two insecticides.
+
+Mixture option 1. Deploy mixture and record time in generations until EITHER frequency reaches the critical point
+
+Mixture, option 2: deploy mixture until one resistance allele reaches the critical point, then stop using that insecticide in the mixture.
+Record time until both resistance allele frequencies reach critical point.
+
+Easy by post-processing : Mixture 1 : set exposure to both insecticides
+Trickier : Sequential
+Trickiest : Mixture 2
+
+Mixture 1 : set exposure to both insecticides and post process to get time until critical points for EITHER insecticide is reached.
+Sequential : set callibration to 1013, add bit to runModel2() to switch from insecticide1 to insecticide2 when critical point for the first insecticide is reached.
+Post process results file to find the time that the 2nd critical point is reached.
+
+Tried adding a sequential insecticide scenario as an example to runModel2(), doesn't seem to be working yet, or maybe I just haven't passed it the correct args.
+
+aha, i think probably because the niche toggles not switched on. Can I set all of these to on by default so that this problem doesn't happen again ?? done
+
+resistance to 2nd insecticide still seems not to rise.
+
+It may be because I need to reset the fitnesses (which happens before gen loop):
+
+Windiv[sex,locus1,locus2] <- sum( a[sex,,] * Wniche[locus1,locus2,,])
+and could put this into a function called setFitness() or the like.
+
 ### TO DO
+
+1. Questions for Ian
++ do we want to do sequential & mixtures for the first sensitivity analysis ?
++ what are the maximum generations that we are running for ? How do we deal with runs where critical point isn't reached in max generations.
++ is it OK to asses just resistance in males ?
+
+1. write out how I'm going to run the sensitivity analysis so I can do a final check with Ian while I'm in Liverpool.
+
+1. think about whether we are able to run sequential insecticide scenarios with the current setup.
 
@@
 
-1. submit travel claim
-
 1. create a vignette from first shiny app documenting how the model works
@@
 1. check out the part about running sequential insecticide scenarios
-
-1. spend some time developing my understanding of how the model works
 1. submit tilde bug report

### 2015-06-09 resistanceModelNotesAndy.Rmd
progress notes

@@
 
+** 9/6/15 ** 4-5
+
+Chatting with Ian & working out that both sequential and mixture2 scenarios can also be done by post-processing of the output file.
+
+Now looking at setting up sensiAnPaper1() to run the sensitivity analysis for the paper.
+Realising that I might not be able to call sensiAn1 and setInputSensiScenarios() because of some of the dependencies between parameters.
+
+Particularly that `P_2` needs to be set as a proprtion of `P_1`.
+
+Actually sensiAnPaper1() can be much more straightforward if it has hardcoded args and doesn't need to be so flexible.
+
 ### TO DO

### 2015-06-10 resistanceModelNotesAndy.Rmd
notes

@@
 
+be careful that runModel & runModel2 should no longer generate same results when sex linkage is set (but i don't think this should break current testthat tests)
+
 **8/6/15** on train to liverpool 4.30-9.30
@@
 
+** 10/6/15 ** train back from liverpool
+
+modifying sensiAnPaper1() to run the different insecticideUsed scenarios
+started sensiPostProc() to do post-processing
+
 ### TO DO
 
-1. Questions for Ian
-+ do we want to do sequential & mixtures for the first sensitivity analysis ?
-+ what are the maximum generations that we are running for ? How do we deal with runs where critical point isn't reached in max generations.
-+ is it OK to asses just resistance in males ?
-
-1. write out how I'm going to run the sensitivity analysis so I can do a final check with Ian while I'm in Liverpool.
-
-1. think about whether we are able to run sequential insecticide scenarios with the current setup.
-
-1. be careful that runModel & runModel2 should no longer generate same results when sex linkage is set (but i don't think this should break current testthat tests)
+1.
+work on sensiPostProc() It will need to be quite clever.
+It will take the output files from insecticide1, insecticide2 and mixture and use these to calculate timeToResistance critical points for sequential, mixture1 & mixture2 scenarios.
+
+1.
+do the output files from runModel2() have the input params attached or do I need to save inputs & outputs separately ?
+The latter may be a good plan to avoid having to rerun long simulations if I want to investigate outputs further later.
+
+1. maybe rename sensiAnPaper1 to something more informative
 

### 2015-06-30 sensiAnPaper1All.Rmd
change overall sensitivity analysis file from function to Rmd

---
title: "sensiAnPaper1All"
output: html_document
---

### All code for the sensitivity analysis for the first paper

```{r}
  library(resistance)

outFolder <- "C:\\Dropbox\\resistanceResults\\"

  nScenarios <- 1000
  ## do model runs
  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture' )
  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1' )
  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2' )

  ## save results objects as rda for re-use
  save(listOutMix, file=paste(outFolder,'listOutMix.rda'))

```

  ## post processing
  #sensiPostProc( inAndOutMix = inAndOutMix,
  #               inAndOutI1 = inAndOutI1,
  #               inAndOutI2 = inAndOutI2 )

  #return(listOut)

}

### 2015-06-30 R/sensiAnPaperAll.r

@@ -1,31 +1,0 @@
-#' overall sensitivity analysis for the first paper
-#' 
-#' unlikely to be run in one go
-#' but held here as a record of everything done
-#' 
-#' @param nScenarios number of scenarios 
-#' 
-#' @return todo
-#' @export
-#' 
-sensiAnPaperAll <- function( nScenarios = 10 )
-{
-
-  nScenarios <- 1000
-  ## do model runs
-  inAndOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture' )
-  inAndOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1' )
-  inAndOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2' )
-  
-  ## save results objects as rda ?
-  
-  
-  ## post processing
-  #sensiPostProc( inAndOutMix = inAndOutMix,
-  #               inAndOutI1 = inAndOutI1,
-  #               inAndOutI2 = inAndOutI2 )
-  
-  
-  #return(listOut)
-  
-}

### 2015-06-30 resistanceModelNotesAndy.Rmd
progress notes

@@
 
+** 24/6/15 ** in liverpool before talking to Ian
+
+checking where I have got to.
+
+I'm going to renam sensiAnPaper1 to sensiAnPaperPart()
+then create a sensiAnPaperAll() to run everything (although in practice unlikely to run it all at once)
+
+** 29/6/15 **
+little on sensiPostProc()
+
+** 30/6/15 **
+moved sensiAnPaper1All() into an rmarkdown file - because in reality I'm not likely to run it all at once.
+
+Beths work just uses male allele frequencies to assess criticalPoints m.R1 & f.R1 are the same for now, but it might be safer for me to use an average.
+
+~ added the input matrix onto the listOut output object in runModel2() to help with post processing
+Slight issue that this has now broken my testthat tests !!
+
+How can m.R2 and f.R2 be greater than 1 (e.g. they start at 1.7455 in first run) ?
+Is this because the frequency of R2 is set too high at start ?
+
+m.R2 is calculated at line 801 in the older runModel.r and line 376 in runModel2
+
+Frequency of resistance allele to the 2nd insecticide is starting above 1.
+
+m.R2 <- ( f.m.SS1RR2 + f.m.RS1RR2 + f.m.RR1RR2 ) + ( 0.5 * (f.m.SS1RS2 + f.m.RS1RS2_cis + f.m.RS1RS2_trans + f.m.RR1RS2 ) )
+
+#refactored
+m.R2 <- sum(f['m',grep("RR2",colnames(f))]) + ( 0.5 * sum(f['m',grep("RS2",colnames(f))]))
+
+Can I create a reproducible example ? or should I just ask first ?
+
+What is the starting value of P_2 the starting frequency of resistance to insecticide2 ?
+For the first dodgy run
+listOutMix$input['P_2',1]
+    P_2 
+1.74551
+
+Aha yes, so maybe i should put something in to stop it going above 1 ??
+
+I think problem was because I set the upper limit on P_2 to 100 rather than 1 (based upon Ians table).
+
+FRom Ian :
+A trick one that..
+
+I think my first instinct was to not restrict one insecticide resistance frequency to be always  less than the other.
+But we would have to censor the inputs because if frequency at locus 1  is 0.1 then the frequency at locus 2 cannot be 100 times higher.
+
+I think we should go with your suggestion i.e.
+recognise that frequency at one locus will always be less than at the other and the lower frequency occurs at Locus 2.
+I would maybe go from 0.0001 to 1 under these circumstances so we get a wide range of starting frequency ratios.
+
+Cool findResistancePoints() seems to be working
+
+> findResistancePoints(listOut, locus=1)
+           [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10] [,11] [,12] [,13] [,14] [,15] [,16] [,17] [,18] [,19]
+gen_cP0.1     9   10    5   36    4    3    5   19   48    77    12    10     3     2     4     3    31    36     5   
+gen_cP0.25   20   20   27   70   15   16   11   41   72   120    19    16    17     5     8     9    54    71    20   
+gen_cP0.5    27   29   39   89   28   25   18   56   97   146    25    23    36     9    12    16    73    95    39   
+> findResistancePoints(listOut, locus=2)
+           [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10] [,11] [,12] [,13] [,14] [,15] [,16] [,17] [,18] [,19] 
+gen_cP0.1     4   31   21   76   25   14    7   30   23   301    59     8    34     2    25    91    19    26    97   
+gen_cP0.25    8   48   45  128   56   26   16   40   35   334    79    12    42     5    30   118    29    53   134   
+gen_cP0.5    13   61   62  154   77   36   24   51   48   357    90    17    51     9    34   133    39    81   166   
+> findResistancePoints(listOut, locus='either')
+           [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10] [,11] [,12] [,13] [,14] [,15] [,16] [,17] [,18] [,19]
+gen_cP0.1     4   10    5   36    4    3    5   19   23    77    12     8     3     2     4     3    19    26     5 
+gen_cP0.25    8   20   27   70   15   16   11   40   35   120    19    12    17     5     8     9    29    53    20   
+gen_cP0.5    13   29   39   89   28   25   18   51   48   146    25    17    36     9    12    16    39    81    39   
+> findResistancePoints(listOut, locus='both')
+           [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10] [,11] [,12] [,13] [,14] [,15] [,16] [,17] [,18] [,19]
+gen_cP0.1     9   31   21   76   25   14    7   30   48   301    59    10    34     2    25    91    31    36    97   
+gen_cP0.25   20   48   45  128   56   26   16   41   72   334    79    16    42     5    30   118    54    71   134   
+gen_cP0.5    27   61   62  154   77   36   24   56   97   357    90    23    51     9    34   133    73    95   166
+
+but the 2nd mixture scenario is the tricky bit because
+
+need to know which of the 2 insecticides reached the critical point first, and the resistance level of the other. I think I'll need to create a special function.
+
+findResistancePointsMixResponsive()
+
 ### TO DO
+
+1. create findResistancePointsMixResponsive()
+
+1. fix the testthat tests to avoid new input addition
 
@@
 It will take the output files from insecticide1, insecticide2 and mixture and use these to calculate timeToResistance critical points for sequential, mixture1 & mixture2 scenarios.
-
-1.
-do the output files from runModel2() have the input params attached or do I need to save inputs & outputs separately ?
-The latter may be a good plan to avoid having to rerun long simulations if I want to investigate outputs further later.
-
-1. maybe rename sensiAnPaper1 to something more informative
+Try to make it as modular as possible.
 
@@
 
-1. I should include unit testing on my CV
-
-1.
-go through the paper and check which bits relate to which in the code.
-Maybe create a table ?
-Could I create the table from R ?
-but actually much of the functionality is still in runModel() and needs to be put into functions.
-
 1. maybe separate out classification trees from sensiAn1
@@
 1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
-1. edit/refactor runModel carefully
-1. check out the part about running sequential insecticide scenarios
 1. submit tilde bug report
+
+Want to do:
+1. rename a to exposure
+1. rename W to ?fitness : fitnessLocus, fitnessNiche and fitnessIndiv
 
@@
 W.RR1SS2_0b ... *81 3*3*3*3 to: Wniche[locus1, locus2, niche1, niche2] 
-
-still todo
 W.RR1_00 etc. with Wloci[loci, niche1, niche2]
@@
 f.m.SS1SS2 ... *19 2*3*3+1 to: f[sex, genotype]
-
-```
-Want to do:
-1. rename a to exposure
-1. rename W to ?fitness : fitnessLocus, fitnessNiche and fitnessIndiv
-
-Questions.
-1. check on calibration 104
-1. What if `a.m_A0 > 0` but `niche_A0 == 0` ?
-The ifelse clauses allow niches to be toggled on/off, i.e. a fitness can be given for A0
-but if only the niche A,B is toggled on, the fitness scores for A0 and Ab will be set to 0
+```
+
+### outputs ###
+From runModel2()
+A list of 3 lists of one or more scenarios: 
+1. results : by generation freq of R allele at each loci in each sex plus linkage stuff
+    columns include Gen, m.R1, m.R2, f.R1, f.R2
+2. genotype : frequencies of each of the ten genotypes, per generation
+3. fitness : fitness scores of each genotype/niche combination (table 4. of Main Document)
+4. input
+e.g. 
+listOut$results[1] gives a results matrix for the first scenario
+listOutMix$input gives the inputs for all scenarios

### 2015-06-30 sensiAnPaper1All.Rmd
develop analysis

@@
 
-  nScenarios <- 1000
+  nScenarios <- 100
   ## do model runs
@@
   save(listOutMix, file=paste(outFolder,'listOutMix.rda'))
+  save(listOutI1, file=paste(outFolder,'listOutI1.rda'))
+  save(listOutI2, file=paste(outFolder,'listOutI2.rda'))    
+```
 
-```
+### now want to calculate the times to reach critical points and add them on to the input file 
+### for different scenarios
+### or will sensiPostProc() sort that ?
 

### 2015-07-01 sensiAnPaper1All.Rmd
deleted post processing function and code moved into Rmd instead

@@
 ### All code for the sensitivity analysis for the first paper
+
+The main question is under which scenarios are mixtures better than sequential ?
+
+This breaks down into these subsidiary quetsions :
+1. Under which scenarios is the time to resistance for either insecticide quicker for
 
@@
 ### now want to calculate the times to reach critical points and add them on to the input file 
-### for different scenarios
-### or will sensiPostProc() sort that ?
+### for different insecticide strategies, sequential, mix1 and mix2
+```{r}
+  #1) sequential : time to resistance for each insecticide in isolation
+  #inputs : inAndOutI1, inAndOutI2
+  #find time to criticalPoint for insecticide1
+  #find time to criticalPoint for insecticide2
+  #add together
+  resistPoints1 <- findResistancePoints(listOutI1, locus=1)
+  resistPoints2 <- findResistancePoints(listOutI2, locus=2)
 
-  ## post processing
-  #sensiPostProc( inAndOutMix = inAndOutMix,
-  #               inAndOutI1 = inAndOutI1,
-  #               inAndOutI2 = inAndOutI2 )
+  resistPointsSeq <- resistPoints1 + resistPoints2
+
+  #2) mixture1 : time to resistance for either insecticide when used in a mixture
+  #inputs : inAndOutMix
+  #find time to criticalPoint for EITHER insecticide in mixture  
+  resistPointsMix1 <- findResistancePoints(listOutMix, locus='either')  
+  #todo - to be comparable I think this should be for when resistance to BOTH insecticides is reached
+
+  #3) mixture2 : when resistance to one insecticide in the mixture reached, switch to sole use of the 
+  #   other until that too reaches the critical point. Record total time.
+  # what I actually need to do is start with mixture find the first critical point
+  # (need to know which of the insecticides it is)
+  # then I need to go to the single run for the other insecticide & starting at 
+  # it's current resistance point find out how many more generations to go
+  #inputs : inAndOutI1, inAndOutI2, inAndOutMix
+
+  #resistPointsMix2 <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
+```
+
+### calculate the difference between sequential & mixture scenarios
+```{r}
+
+  # easy to calc for which scenarios mixture1 is 'better' i.e. longer to resistance
+  #boolean
+  resistBetterMix1SeqBoolean <- resistPointsMix1 > resistPointsSeq
+  #converted to 0/1
+  resistBetterMix1Seq <- plyr::aaply(resistBetterMix1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix1Seq) <- gsub('gen','betterMix1Seq', rownames(resistBetterMix1Seq))
+
+```
+
+### here I want a file with the inputs (just one value for exposure)
+### with extra columns for seq
+```{r}
+treeInput <- rbind( listOutMix$input, resistBetterMix1Seq )
+
+```
 

### 2015-07-01 R/sensiPostProc.r

@@ -1,59 +1,0 @@
-#' post-processing for sensitivity analysis for the first paper
-#' 
-#' takes output files for sensitivity analysis runs for insecticide1, insecticide2 and mixture
-#' uses these to calculate time to resistance critical points (which can be passed as an argument)
-#' for 3 scenarios of insecticide use :   
-#' 1) sequential : time to resistance for each insecticide in isolation
-#' 2) mixture1 : time to resistance for either insecticide when used in a mixture
-#' 3) mixture2 : when resistance to one insecticide in th mixture reached, switch to sole use of the 
-#' other until that too reaches the critical point. Record total time.
-#' 
-#' @param criticalPoints vector of resistance critical points where insecticide is deemed no longer useful
-#' @param inAndOutMix output object for mixture from sensiAnPaperPart()
-#' @param inAndOutI1 output object for insecticide1 (commoner and first) from sensiAnPaperPart()
-#' @param inAndOutI2 output object for insecticide2 (rarer and secound) from sensiAnPaperPart()
-#' 
-#' @examples
-#' #sensiPostProc()
-#' @return todo
-#' @export
-#' 
-sensiPostProc <- function( criticalPoints = c(0.1,0.25,0.5),
-                           inAndOutMix,
-                           inAndOutI1,
-                           inAndOutI2 )
-{
-  
-  #? should I do a loop for each criticalPoint ?
-  #? for a start just reference criticalPoint below
-  
-  
-  #1) sequential : time to resistance for each insecticide in isolation
-  #inputs : inAndOutI1, inAndOutI2
-  #find time to criticalPoint for insecticide1
-  #find time to criticalPoint for insecticide2
-  #add together
-  resistPoints1 <- findResistancePoints(listOutI1, locus=1)
-  resistPoints2 <- findResistancePoints(listOutI2, locus=2)  
-  
-  resistPointsSeq <- resistPoints1 + resistPoints2
-  
-  
-  #2) mixture1 : time to resistance for either insecticide when used in a mixture
-  #inputs : inAndOutMix
-  #find time to criticalPoint for EITHER insecticide in mixture  
-  resistPointsMix1 <- findResistancePoints(listOutMix, locus='either')  
-  
-  #3) mixture2 : when resistance to one insecticide in the mixture reached, switch to sole use of the 
-  #   other until that too reaches the critical point. Record total time.
-  # what I actually need to do is start with mixture find the first critical point
[…]

### 2015-07-01 sensiAnPaper1All.Rmd
sensitivity analysis improvements

@@
 
+  #default setting of the random seed in sensiAnPaperPart() should make the input files mostly the same with exception of exposure
+
   nScenarios <- 100
+  #nScenarios <- 10
   ## do model runs
@@
 
-```
-
-### here I want a file with the inputs (just one value for exposure)
-### with extra columns for seq
-```{r}
-treeInput <- rbind( listOutMix$input, resistBetterMix1Seq )
+  #in 100 runs for all 3 critical points sequential was better than mix1 all the time
 
@@
 
-  #return(listOut)
+### bind results onto input file
+```{r}
 
-}
+  treeInput <- listOutMix$input
+
+  #input files in listOutMix, listOutIn1 & listOutI2 are the same if the runs are done with default randomSeed 
+  #except that exposure will be in a.f_AB, a.f_A0 and a.f_B0 respectively (& a.m* too)
+  #I could just rename one to exposure
+  rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+  #hardcode which variables to include in the tree analysis here
+  #to keep it simple and transparent
+  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  treeInput <- treeInput[ treePredictors, ]
+
+  treeInput <- rbind( treeInput, resistBetterMix1Seq )
+  #todo add other results rows here
+
+```
+
+### do tree analysis
+```{r}
+  require(rpart)
+
+  #transpose
+  treeInput <- t(treeInput)
+  #convert to a dataframe
+  treeInput <- data.frame(treeInput)
+
+  treePredictorString <- paste(treePredictors, collapse="+")
+  treeResponse <- 'betterMix1Seq_cP0.5'
+
+  #at 0.5 level
+  tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
+  #getting error number of rows of matrices must match (see arg 2)
+  #may be becuase all vals are 0 ??
+  #yes when I added some 1s it worked OK
+  #treeInput$betterMix1Seq_cP0.5[1:5] <- 1
+
+  #using as.formula
+  #tree <- rpart::rpart(as.formula(paste("res ~",argString)), data = inputAndResults, method = 'class')  
+  #hardcoded args
+  #tree <- rpart(res ~ h.RS1_A0 + h.RS2_0B, data = inputAndResults, method = 'class')
+  #tree <- rpart(betterMix1Seq_cP0.5 ~ h.RS1_A0 + h.RS2_0B, data = treeInput, method = 'class')
+
+  plot(tree) #plots tree
+  text(tree) #labels tree
+
+    #pruning tree
+    #   treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+    #   plot(treePruned) #plots tree
+    #   text(treePruned) #labels tree
+
+```

### 2015-07-06 sensiAnPaper1All.Rmd
add to sensitivity analysis

@@
 
-```{r}
+<!-- eval=FALSE to stop slow code being executed --> 
+```{r, eval=FALSE}
   library(resistance)
@@
   ## save results objects as rda for re-use
-  save(listOutMix, file=paste(outFolder,'listOutMix.rda'))
-  save(listOutI1, file=paste(outFolder,'listOutI1.rda'))
-  save(listOutI2, file=paste(outFolder,'listOutI2.rda'))    
+  save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))
+
+  ## to load previously saved runs, e.g. these that I added 100 onto filename
+  load(file=paste0(outFolder,'listOutMix_100.rda'))
+  load(file=paste0(outFolder,'listOutI1_100.rda'))
+  load(file=paste0(outFolder,'listOutI2_100.rda'))
+
 ```
@@
   #add together
-  resistPoints1 <- findResistancePoints(listOutI1, locus=1)
-  resistPoints2 <- findResistancePoints(listOutI2, locus=2)
+  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
+  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)
 
-  resistPointsSeq <- resistPoints1 + resistPoints2
+  resistPointsSeq <- resistPointsI1 + resistPointsI2
 
@@
   #resistPointsMix2 <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
+
+  #4) mixture3 : time to resistance for both insecticides when used in a mixture
+  #inputs : inAndOutMix
+  #find time to criticalPoint for BOTH insecticide in mixture  
+  resistPointsMix3 <- findResistancePoints(listOutMix, locus='both')
+
 ```
@@
 
-  # easy to calc for which scenarios mixture1 is 'better' i.e. longer to resistance
-  #boolean
+  require('plyr')
+
+  ## is mix1 better than sequential ?
+  #T/F
   resistBetterMix1SeqBoolean <- resistPointsMix1 > resistPointsSeq
-  #converted to 0/1
+  #convert to 0/1
   resistBetterMix1Seq <- plyr::aaply(resistBetterMix1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
@@
   #in 100 runs for all 3 critical points sequential was better than mix1 all the time
+
+  ## is mix3 better than sequential ?
+  #T/F
+  resistBetterMix3SeqBoolean <- resistPointsMix3 > resistPointsSeq
+  #convert to 0/1
+  resistBetterMix3Seq <- plyr::aaply(resistBetterMix3SeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix3Seq) <- gsub('gen','betterMix3Seq', rownames(resistBetterMix3Seq))
+
+  #in 100 runs only one critical point run combination was mix3 better
+
+  #check1
+  #plot the 2 individual results & the mixture, just to check what is going on
+  #the mixture one looks superficially similar to the individual ones superimposed
+  x11()
+  plotallele.freq.andy(listOutMix)
+  x11()
+  plotallele.freq.andy(listOutI1)
+  x11()
+  plotallele.freq.andy(listOutI2)
+
+  #check2
+  #what is the absolute difference between mix3 & sequential ?
+  #looks plausible
+  resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
+
+  #check3 is their a difference between the mix & the individual strategies
+  #looks mostly plausible, occasionally there are 0s
+  resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
+  resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2
+
+  ## is mix2 better than sequential ?  
+  # I might expect it to be similar to mix3
 

### 2015-07-06 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 31/6/15 **
+Quick review of what needs to go into the final decision tree.
+
+slight rethink : I want the input files to be mostly the same for I1,I2 & mixture, so that they can be added together in combined scenarios.
+
+I added a set.seed into sensiAnPaperPart()
+This shows the input files for diffeerent insecticide strategies are now the same.
+
+tst <- listOutMix$input[c('P_1','P_2','a.f_AB','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B'),]
+#note a.f_A0 rather than a.f_AB
+tst2 <- listOutI1$input[c('P_1','P_2','a.f_A0','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B'),]
+identical(as.numeric(tst),as.numeric(tst2))
+
 ### TO DO
 
+1. CHECK MODEL RUNS ARE DOING WHAT I THINK. Currently seems that resistance is quicker for mixtures under all scenarios (or very nearly all)
+
 1. create findResistancePointsMixResponsive()
 
+1. review how to answer these questions.
+Main question is : Tree1 is mixture better (0/1), Tree2 is mixture better by >20%.
+
 1. fix the testthat tests to avoid new input addition
 
-1.
-work on sensiPostProc() It will need to be quite clever.
-It will take the output files from insecticide1, insecticide2 and mixture and use these to calculate timeToResistance critical points for sequential, mixture1 & mixture2 scenarios.
-Try to make it as modular as possible.
+1. work further on sensiAnPaper1All
+
+1. decide whether to do main runs with P_2 dependent on P_1 or not
 
@@
 It is a 2 locus model. One locus for each insecticide. At each locus there are 2 alleles. This gives a total of 9 genotypes if the position of the alleles is ignored.
+
+Main question is : Tree1 is mixture better (0/1), Tree2 is mixture better by >20%.
 
@@
 
-1. maybe separate out classification trees from sensiAn1
-
 1. in susana's script the 'outcome' is 0/1 if used in our case that would seem to lose a lot of information about the outcomes ? Even timeToFifty would be better for us.
-
-1. write a function to calculate timeToThreshold for all scenarios based on listOut, allow user to specify which locus.
 
@@
 listOutMix$input gives the inputs for all scenarios
+listOutMix$input[c('P_1','P_2'),] gives selected inputs

### 2015-07-06 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 6/7/15 **
+`plotcurtis_f2_generic` : a function allowing mixture & individual scenarios to be compared, similar to Curtis Fig2.
+
+In creating a shiny app to display this, I could allow user to select parameters, but then the run might take a little while because it has to do 3 runs (I1,I2,mix) & post processing.
+Quicker may be to display my sensitivity analysis results & allow user to step through scenarios.
++ include printing out the input values used (& maybe even their relative position within ranges).
+
 ### TO DO
+
+1. should I create a little shiny app allowing the mixture & sequential scenarios to be compared ?
 

### 2015-07-07 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 7/7/15 **
+resistanceMaster() does reproduce curtis fig2
+so the parameters in system.file("extdata","input.parameters.csv", package="resistance")
+must work
+
+what is different about the input params in the sensitivity analysis ?
+
+resistanceMaster() still calls runModel() if I change it to call runModel2() does it still work ??
+
+No I get this :
+Error in runModel2(input, calibration) : 
+  Error in male exposures: must total one:  1.9
+
+but in the input files, looks like it is -- 0.1 & AB 0.9 for all.
+
+No not in the hardcoded, curtis fig2 input which comes from the program. It has a 0.9 for ab which slips in too.
+
+                                             B (HCH)   A (DDT) Combination
+Calibration (100 default)                  1012.0000 1012.0000   1.012e+03
+Number of generations                        70.0000   70.0000   1.600e+02
+Collect fitness scores in matrix (1/0)        1.0000    1.0000   1.000e+00
+Export fitness scores matrix to .csv (1/0)    0.0000    0.0000   0.000e+00
+Frequency of R at locus 1                     0.0100    0.0100   1.000e-02
+Frequency of R at locus 2                     0.0100    0.0100   1.000e-02
+Recombination Rate                            0.5000    0.5000   5.000e-01
+Exposure Males -,-                            0.1000    0.1000   1.000e-01
+Exposure Males a,-                            0.0000    0.0000   0.000e+00
+Exposure Males A,-                            0.0000    0.9000   0.000e+00
+Exposure Males -,b                            0.0000    0.0000   0.000e+00
+Exposure Males -,B                            0.9000    0.0000   0.000e+00
+Exposure males a,b                            0.0000    0.0000   9.000e-01
+Exposure Males A,B                            0.0000    0.0000   9.000e-01
+...
+
+exactly where is that input object created ?
+createInputMatrix()
+fixed bug, but it didn't effect the output
+
+can I compare the inputs for the 3 scenarios by Beth to those in my first scenario ? in a csv.
+
+This gets the file to create Curtis Fig2.
+```
+inputCurtis <- createInputMatrix(FALSE)
+
+```
+
+Can I paste the input for the first scenario onto it ?
+Or might be better just to output all 100 scenarios.
+
+see :
+
+seems that the dominance coefficient of locus2 in -B is much lower in Beths than in the sensi scenarios.
+
+Dominance coefficient L2 in -,B	0.0016	0.0016	0.00016	0.944675269	0.497699242
+
+First see if the curtis fig2 persists when dominance for the mixture is reduced.
+
+Good news that this correction seemed to make the graph closer to Curtis. 
+    #input[37,3] <- 0.00016	#Dominance coefficient in B
+    #? should this be the same as the value in ,2 & ,1
+    input[37,3] <- 0.0016	#Dominance coefficient in B
+
 ### TO DO

### 2015-07-07 sensiAnPaper1All.Rmd
more sensitivity testing

@@
 
-<!-- eval=FALSE to stop slow code being executed --> 
-```{r, eval=FALSE}
+```{r}
   library(resistance)
@@
   outFolder <- "C:\\Dropbox\\resistanceResults\\"
+```
+
+<!-- eval=FALSE to stop slow code being executed --> 
+```{r, eval=FALSE}
 
@@
 
+```
+
+```{r}
   ## to load previously saved runs, e.g. these that I added 100 onto filename
@@
   #the mixture one looks superficially similar to the individual ones superimposed
-  x11()
+  #x11()
   plotallele.freq.andy(listOutMix)
-  x11()
+  #x11()
   plotallele.freq.andy(listOutI1)
-  x11()
+  #x11()
   plotallele.freq.andy(listOutI2)
+
+  #I should be able to compare single scenarios
+  #yes this works, but not too useful
+  plotallele.freq.andy( list(results=list(listOutMix$results[[1]], listOutI1$results[[1]] )) )
+  #this is a modifdeied version of Beths function
+  #allowing curtis fig2 to be applied to any scenario
+  plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )
 
@@
   # I might expect it to be similar to mix3
+
+  # comparing sensitivity inputs to those used by Beth to reproduce curtis fig2
+  inputCurtis <- createInputMatrix(FALSE)
+  inputSensiMix <- listOutMix$input
+
+  input <- cbind(inputCurtis, inputSensiMix)
+  write.csv(input, file=paste0(outFolder,'inputCurtisSensiCompare.csv'))
 
@@
 ### do tree analysis
-```{r}
+```{r, eval=FALSE}
   require(rpart)

### 2015-07-08 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 7/7/15 **
+new version of UI comparing scenarios with greater differences in dominance.
+Still doesn't produce something similar to curtis fig2 where mixture does much better.
+
 ### TO DO
 
-1. should I create a little shiny app allowing the mixture & sequential scenarios to be compared ?
+1. should I create a little shiny app mimicing curtis fig2, & allowing usr to modify params, to see how to keep the mixtures doing better ?
 

### 2015-07-08 resistanceModelNotesAndy.Rmd
notes

@@
 
+created runcurtis_f2() to do runs to recreate curtis fig2. recreates it fine.
+
+created shinyFig2Curtis, that calls runcurtis_f2()
+
+for some reason it currently doesn't recreate the curtis plot.
+what inputs are wrong in shinyFig2Curtis ?
+aha! the difference was just `P_1 & P_2` which were set to 0.1 rather than 0.01
+aha2, even after that fixed, something still not quite right
+
+this was because phi.SS2_0B = 1 was set to 0 instead. This is the fitness of exposed susceptibles.
+(1 suggests that the susceptible is completely fit in presence of the insecticide which doesn't seem right).
+
+yes phi.SS2_0B seems key :
+compare these 2:
+```{r}
+#curtis default phi.SS2_0B = 1 : mixture better than sequence
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+#reduce phi.SS2_0B to 0.6 : mixture same as sequence
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.6 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+```
+
+How does that relate to this comment from the paper ?
+In particular, Curtis [1] allowed the resistant and sensitive homozygotes to survive DDT exposure with finesses of 50% and 27% respectively (caption to his Figure 2).
+
 ### TO DO
 
-1. should I create a little shiny app mimicing curtis fig2, & allowing usr to modify params, to see how to keep the mixtures doing better ?
+1. why can't shinyapps find resistance for this :
+Deployment completed: https://andysouth.shinyapps.io/shinyFig2Curtis
+Warning message:
+In FUN(X[[i]], ...) :
+  Package 'resistance' not available in repository or locally
+
+1. work out what to do if critical threshold not reached in findResistancePoints()
+causes : Error in plot.window(...) : need finite 'xlim' values
 

### 2015-07-13 resistanceModelNotesAndy.Rmd
notes

@@
 
+packageDescription('resistance')
+
+GithubSHA1: 7858f1edc0022233c0cd26f84be64e1a9bd0cd04 
+            7858f1edc0022233c0cd26f84be64e1a9bd0cd04
+
+This shows that the installed package is the latest version from github.
+
+problem was because I hadn't updated tghe NAMESPACE Ctrl Shift D
+
+reran model with new ranges for phi vals (0.6 to 1)
+(but still need to sort finite 'xlim' values errors)
+
+e.g. scenario 2
+
+Email from Ian, showing section vi of sensitivity doc.
+
+We have to ensure the fitness of the RS and RR genotypes does not exceed 1.0 as this is the reference (maximum value of the SS genotype in the absence of insecticide.
+The fitness of the SS homozygote is determined by the ϕ values in Table 3 i.e.
+as 1- ϕ.
+Hence the values of s must be less than ϕ.
+I suggested we set s=x ϕ where x comes for a uniform distribution between 0.2 and 1.0”
+
+This relates to this bit of code :
+
+      for( exposure in c('lo','hi') )
+      {
+        Wloci[ paste0('SS',locusNum), exposure] <-  1 - phi[locusNum, exposure]
+
+        Wloci[ paste0('RS',locusNum), exposure] <- (1 - phi[locusNum, exposure]) + 
+                                                   (h[locusNum, exposure] * s[locusNum, exposure])
+
+        Wloci[ paste0('RR',locusNum), exposure] <- (1 - phi[locusNum, exposure]) + 
+                                                   (s[locusNum, exposure])
+      }
+
+Wloci[RR] <- (1-phi) + s
+
+so s must be < phi to keep Wloci[RR] below 1
+
+added in checks for fitnesses >1 or <0 in runModel2
+
+initial tree looks promising :
+for mix1 better than sequential:
+phi.SS2_0B < 0.66
+phi.SS1_AO < 0.68
+P_2 > 0.033
+
+** 10/7/15 **
+installed rpart.plot
+for prp function used by susana
+
+created loop to do all plots, with titles and save
+
+repeated runs for 500 scenarios, took about an hour.
+
+** 11/7/15 **
+running 1000 scenarios
+started at 10.15, check timestamp on outputs to see how long it took
+
+** 13//7/15 **
+
+done created findResistancePointsMixResponsive() for mix2
+
+analysing 1000 runs
+
+  #1000runs
+  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #              0.178                0.232                0.238 
+  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #              0.471                0.568                0.551 
+  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #              0.794                0.726                0.673 
+rounded
+  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #              0.18                0.23                0.24 
+  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #              0.47                0.57                0.55 
+  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #              0.79                0.73                0.67
+
+essentially
+mix1 : 0.2
+mix2 : 0.5
+mix3 : 0.7
+
+alas the trees are more complicated, but that might be fixed by pruning them ?
+maybe not ... 
+does work well for :
+betterMix1Seq_cP0.1
+
+BUT wierdly seems that plots look better (ie much fewer branches) when done individually than when saved in the loop.
+
+is it something to do with interaction between prp and savePlot() ?
+
+Now I seem to be getting a different result each time ?????
+
 ### TO DO
 
-1. why can't shinyapps find resistance for this :
-Deployment completed: https://andysouth.shinyapps.io/shinyFig2Curtis
-Warning message:
-In FUN(X[[i]], ...) :
-  Package 'resistance' not available in repository or locally
+1. try to get some simple trees
+
+1. decide how to format table of results
+1. put some results together for Ian, not sure how to do trees
+1. include a couple of contrasting graphs, to make a story. e.g. one where mixture better and one where sequential. Maybe number points on the figs.
+
+1. it might be that the tyree results are not useful ??
+
+1. sort order of insecticides for sequential figures
 
@@
 
-1. CHECK MODEL RUNS ARE DOING WHAT I THINK. Currently seems that resistance is quicker for mixtures under all scenarios (or very nearly all)
-
-1. create findResistancePointsMixResponsive()
-
 1. review how to answer these questions.
@@
 1. fix the testthat tests to avoid new input addition
-
-1. work further on sensiAnPaper1All
-
-1. decide whether to do main runs with P_2 dependent on P_1 or not
 
@@
 listOutMix$input[c('P_1','P_2'),] gives selected inputs
+
+# current best UIs
+https://andysouth.shinyapps.io/shinyMixSeqCompare1
+https://andysouth.shinyapps.io/shinyFig2Curtis

### 2015-07-13 sensiAnPaper1All.Rmd
sensitivity analysis contd

--- 2015-07-06 sensiAnPaper1All.Rmd
@@
   nScenarios <- 100
-  #nScenarios <- 10
+  #nScenarios <- 1000
   ## do model runs
@@
 
+```
+
+```{r}
   ## to load previously saved runs, e.g. these that I added 100 onto filename
-  load(file=paste0(outFolder,'listOutMix_100.rda'))
-  load(file=paste0(outFolder,'listOutI1_100.rda'))
-  load(file=paste0(outFolder,'listOutI2_100.rda'))
+  load(file=paste0(outFolder,'listOutMix_100_4.rda'))
+  load(file=paste0(outFolder,'listOutI1_100_4.rda'))
+  load(file=paste0(outFolder,'listOutI2_100_4.rda'))
 
@@
   #inputs : inAndOutI1, inAndOutI2, inAndOutMix
-
-  #resistPointsMix2 <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
+  resistPointsMix2 <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
 
@@
 
-  #in 100 runs for all 3 critical points sequential was better than mix1 all the time
+  rowSums(resistBetterMix1Seq)/ncol(resistBetterMix1Seq)
+  #100runs
+  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #                .25                   .26                   .30   
+  #1000runs
+  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #               .178                  .232                  .238
+
+  ## is mix2 better than sequential ?
+  #T/F
+  resistBetterMix2SeqBoolean <- resistPointsMix2 > resistPointsSeq
+  #convert to 0/1
+  resistBetterMix2Seq <- plyr::aaply(resistBetterMix2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix2Seq) <- gsub('gen','betterMix2Seq', rownames(resistBetterMix2Seq))
+
+  rowSums(resistBetterMix2Seq)/ncol(resistBetterMix2Seq)  
+  #1000runs
+  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #             0.471                0.568                0.551
 
@@
 
-  #in 100 runs only one critical point run combination was mix3 better
+  #tally runs where mix3 better than seq
+  rowSums(resistBetterMix3Seq)/ncol(resistBetterMix3Seq)
+  #100runs
+  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #                  .83                   .75                   .72 
+  #1000runs
+  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #                .794                  .726                  .673
 
@@
   #the mixture one looks superficially similar to the individual ones superimposed
-  x11()
-  plotallele.freq.andy(listOutMix)
-  x11()
-  plotallele.freq.andy(listOutI1)
-  x11()
-  plotallele.freq.andy(listOutI2)
+  #x11()
+#   plotallele.freq.andy(listOutMix)
+#   #x11()
+#   plotallele.freq.andy(listOutI1)
+#   #x11()
+#   plotallele.freq.andy(listOutI2)
+
+  #I should be able to compare single scenarios
+  #yes this works, but not too useful
+  # plotallele.freq.andy( list(results=list(listOutMix$results[[1]], listOutI1$results[[1]] )) )
+  #this is a modifdeied version of Beths function
+  #allowing curtis fig2 to be applied to any scenario
+  #plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )
 
   #check2
-  #what is the absolute difference between mix3 & sequential ?
-  #looks plausible
-  resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
-
-  #check3 is their a difference between the mix & the individual strategies
-  #looks mostly plausible, occasionally there are 0s
-  resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
-  resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2
+#   #what is the absolute difference between mix3 & sequential ?
+#   resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
+#   
+#   #check3 is their a difference between the mix & the individual strategies
+#   #looks mostly plausible, occasionally there are 0s
+#   resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
+#   resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2
 
@@
   # I might expect it to be similar to mix3
+
+  # comparing sensitivity inputs to those used by Beth to reproduce curtis fig2
+#   inputCurtis <- createInputMatrix(FALSE)
+#   inputSensiMix <- listOutMix$input
+#   
+#   input <- cbind(inputCurtis, inputSensiMix)
+#   write.csv(input, file=paste0(outFolder,'inputCurtisSensiCompare.csv'))
 
@@
 
+  #bind results onto predictors
   treeInput <- rbind( treeInput, resistBetterMix1Seq )
-  #todo add other results rows here
+  treeInput <- rbind( treeInput, resistBetterMix2Seq )
+  treeInput <- rbind( treeInput, resistBetterMix3Seq )
 
@@
 ### do tree analysis
-```{r}
-  require(rpart)
+```{r, eval=FALSE}
+  require(rpart.plot)
 
@@
   treeInput <- data.frame(treeInput)
+  #create string with predictor names in
+  treePredictorString <- paste(treePredictors, collapse="+")
 
-  treePredictorString <- paste(treePredictors, collapse="+")
-  treeResponse <- 'betterMix1Seq_cP0.5'
+  treeResponses <- c( rownames(resistBetterMix1Seq),
+                      rownames(resistBetterMix2Seq),
+                      rownames(resistBetterMix3Seq) )
 
-  #at 0.5 level
+  #to do trees & plots for all response variables
+  for( treeResponse in treeResponses )
+  {
+    x11()
+    #png(paste0(outFolder,"tree",treeResponse,".png"))
+
     tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
-  #getting error number of rows of matrices must match (see arg 2)
-  #may be becuase all vals are 0 ??
-  #yes when I added some 1s it worked OK
-  #treeInput$betterMix1Seq_cP0.5[1:5] <- 1
+    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+    #adds in no. scenarios correctly and incorrectly classified
+    #prp(tree, extra=1, varlen = 10, main=treeResponse)
+    prp(treePruned, extra=1, varlen = 10, main=treeResponse)
 
-  #using as.formula
-  #tree <- rpart::rpart(as.formula(paste("res ~",argString)), data = inputAndResults, method = 'class')  
-  #hardcoded args
-  #tree <- rpart(res ~ h.RS1_A0 + h.RS2_0B, data = inputAndResults, method = 'class')
-  #tree <- rpart(betterMix1Seq_cP0.5 ~ h.RS1_A0 + h.RS2_0B, data = treeInput, method = 'class')
+    #dev.off()
+    savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')     
+  }
 
-  plot(tree) #plots tree
-  text(tree) #labels tree
+#   treeResponse <- 'betterMix1Seq_cP0.5'
+#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
+#   
+#   #if get error number of rows of matrices must match (see arg 2)
+#   #is because all vals are 0
+# 
+#   plot(tree) #plots tree
+#   text(tree) #labels tree
+#   #branch predictors 2 phis & P_2  
+#     
+#   #for mix3 (resistance to both)
+#   x11()
+#   treeResponse <- 'betterMix3Seq_cP0.5'
+#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  
+#   plot(tree) #plots tree
+#   text(tree) #labels tree  
+#   
+#   #branch predictors 2 phis & P_1
+#   
+#   #or from rpart.plot
+#   #prp(tree)
+#   
+#   #adds in no. scenarios correctly and incorrectly classified
+#   prp(tree, extra=1, varlen = 10, main=treeResponse)
+#   savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
 
+  #prp(tree, extra=1,box.col=c("pink","palegreen3")[tree$frame$yval],cex=0.5,yesno.yshift=-2)
+
+    #susanas code
     #pruning tree
@@
     #   text(treePruned) #labels tree 
+    # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)
 

### 2015-07-14 sensiAnPaper1All.Rmd
add 20% better calculations

@@
   ## to load previously saved runs, e.g. these that I added 100 onto filename
-  load(file=paste0(outFolder,'listOutMix_100_4.rda'))
-  load(file=paste0(outFolder,'listOutI1_100_4.rda'))
-  load(file=paste0(outFolder,'listOutI2_100_4.rda'))
+  load(file=paste0(outFolder,'listOutMix_1000.rda'))
+  load(file=paste0(outFolder,'listOutI1_1000.rda'))
+  load(file=paste0(outFolder,'listOutI2_1000.rda'))
 
@@
   #plot the 2 individual results & the mixture, just to check what is going on
-  #the mixture one looks superficially similar to the individual ones superimposed
   #x11()
@@
 #   write.csv(input, file=paste0(outFolder,'inputCurtisSensiCompare.csv'))
+
+```
+
+### finding when mixtures are >20% better
+### repeating calculation of the difference between sequential & mixture scenarios
+```{r}
+
+  require('plyr')
+
+  ## is mix1 better than sequential ?
+  #T/F
+  resistBetterMix1SeqBoolean20 <- resistPointsMix1 > (1.2 * resistPointsSeq)
+  #convert to 0/1
+  resistBetterMix1Seq20 <- plyr::aaply(resistBetterMix1SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix1Seq20) <- gsub('gen','betterMix1Seq20', rownames(resistBetterMix1Seq20))
+
+  rowSums(resistBetterMix1Seq20)/ncol(resistBetterMix1Seq20)
+  #1000runs
+ #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+ #                 0.136                  0.175                  0.173
+
+  ## is mix2 better than sequential ?
+  #T/F
+  resistBetterMix2SeqBoolean20 <- resistPointsMix2 > (1.2 * resistPointsSeq)
+  #convert to 0/1
+  resistBetterMix2Seq20 <- plyr::aaply(resistBetterMix2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix2Seq20) <- gsub('gen','betterMix2Seq20', rownames(resistBetterMix2Seq20))
+
+  rowSums(resistBetterMix2Seq20)/ncol(resistBetterMix2Seq20)  
+  #1000runs
+#  betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+#                  0.241                  0.326                  0.313
+
+  ## is mix3 better than sequential ?
+  #T/F
+  resistBetterMix3SeqBoolean20 <- resistPointsMix3 > (1.2 * resistPointsSeq)
+  #convert to 0/1
+  resistBetterMix3Seq20 <- plyr::aaply(resistBetterMix3SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix3Seq20) <- gsub('gen','betterMix3Seq20', rownames(resistBetterMix3Seq20))
+
+  #tally runs where mix3 better than seq
+  rowSums(resistBetterMix3Seq20)/ncol(resistBetterMix3Seq20)
+  #1000runs
+#  betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+#                  0.672                  0.560                  0.488
 

### 2015-07-14 sensiAnPaper1All.Rmd
limited trees

@@
     #dev.off()
-    savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')     
+    #savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
+    savePlot(paste0(outFolder,"tree",treeResponse,".jpg"), type='jpg')
+
+    #trying to limit num levels a different way using maxdepth
+    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
+    #adds in no. scenarios correctly and incorrectly classified
+    prp(tree, extra=1, varlen = 10, main=treeResponse)
+    savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')
+
   }

### 2015-07-21 resistanceModelNotesAndy.Rmd
notes

@@
 P_2 > 0.033
+P_1
 
@@
 
+** 14/7/15 **
+
+putting results document together.
+
+Just checking that the results file for the 1000 runs looks like the figures & UI for the 100 runs
+
+run3: sequential best (seq 28+29, mix3 50)
+run4: mixture best (seq 93+32, mix3 172)
+run5: resistance to one in mixture develops more slowly than both in sequence (seq 29+103, mix1 152)
+
+resistPointsI1[,1:5]
+           [,1] [,2] [,3] [,4] [,5]
+gen_cP0.1    38  999   11   57    7
+gen_cP0.25   48  999   21   82   22
+gen_cP0.5    53  999   28   93   29
+resistPointsI2[,1:5]
+           [,1] [,2] [,3] [,4] [,5]
+gen_cP0.1    64  210    7    4   91
+gen_cP0.25   90  250   21   21   99
+gen_cP0.5   102  268   29   32  103
+resistPointsMix1[,1:5]
+           [,1] [,2] [,3] [,4] [,5]
+gen_cP0.1    75  999    9   10   34
+gen_cP0.25   93  999   30   57  116
+gen_cP0.5   102  999   41   86  152
+resistPointsMix3[,1:5]
+           [,1] [,2] [,3] [,4] [,5]
+gen_cP0.1   168  999   21  129  191
+gen_cP0.25  199  999   38  159  202
+gen_cP0.5   213  999   50  172  207
+
+Now can I do the 20% better results ? 
+ betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+                 0.136                  0.175                  0.173 
+ betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+                 0.241                  0.326                  0.313 
+ betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+                 0.672                  0.560                  0.488
+
+rounded
+ betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+                 0.14                  0.18                  0.17 
+ betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+                 0.24                  0.33                  0.31 
+ betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+                 0.67                  0.56                  0.49
+
+Low phi values lead to sequential being better.
+e.g. https://andysouth.shinyapps.io/shinyMixSeqCompare1 
+runs 
+3: phi 0.53, 0.79
+19: phi 0.65, 0.44
+25: phi 0.60, 0.56
+
+** 15/7/15 **
+
+installed triangle package
+this to generate random deviates
+rtriangle(n, a=0, b=1, c=(a+b)/2)
+
+changed label 'combination' to 'mixture' in my generic curtis fig2 plot
+
+** 16/7/15 **
+reran 1000 sensitivity scenarios with new agreed input ranges
+& reran 100 scenarios and put them into shinyMixSeqCompare1
+
+check that the example runs still support the args the way they did before
+
+the 3 example plots that I pasted into the word doc seem to be identical to what they were before
+ARSE is it possible I didn't reload the package so that it used the old function version with the old input ranges ?
+
+yes I think that's what I did !
+
+rereran 100 runs and put them into shinyMixSeqCompare1
+
+sorted order of insecticides for sequential figures
+
+run shinyMixSeqCompare1 in a window rather than the browser to enable copy and pasting figures into word doc
+
+copied 3 figs into the word doc. Note that run 9 is interesting. Mixture does much better than sequential.
+
 ### TO DO
 
-1. try to get some simple trees
-
-1. decide how to format table of results
-1. put some results together for Ian, not sure how to do trees
-1. include a couple of contrasting graphs, to make a story. e.g. one where mixture better and one where sequential. Maybe number points on the figs.
-
-1. it might be that the tyree results are not useful ??
-
-1. sort order of insecticides for sequential figures
-
-1. work out what to do if critical threshold not reached in findResistancePoints()
-causes : Error in plot.window(...) : need finite 'xlim' values
-
-1. review how to answer these questions.
-Main question is : Tree1 is mixture better (0/1), Tree2 is mixture better by >20%.
+1. set 1000 runs to do tonight
+
+1. redo the word results doc
 
@@
 
-### suggestion from Beth
-could provide a .csv file with the Curtis scenarios set up
-and then the need to set the calibration number and params.csv in the R script can be removed altogether
-
-### How best to refactor variables to arrays ?
+### refactoring variables to arrays ?
 Refactoring to arrays will reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions.

### 2015-07-21 sensiAnPaper1All.Rmd
add an extra predictor (ratio of P_1 and P_2)

@@
 ```{r}
+
+  #library(devtools)
+  #install_github('AndySouth/resistance')
   library(resistance)
@@
   #               .178                  .232                  .238 
+  #new100
+  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #              0.18                 0.24                 0.25
 
@@
   #             0.471                0.568                0.551 
+  #new100
+  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #              0.58                 0.57                 0.52
 
@@
   #                .794                  .726                  .673  
+  #new100
+  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #              0.83                 0.79                 0.72
 
@@
 
+  #add an extra predictor, the lower starting freq of resistance divided by the larger
+  startResistanceRatio <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
+  treeInput <- rbind(treeInput,startResistanceRatio)
+
   #hardcode which variables to include in the tree analysis here
   #to keep it simple and transparent
-  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
   treeInput <- treeInput[ treePredictors, ]

### 2015-07-22 sensiAnPaper1All.Rmd
added trees for diff > 20

@@
 
-  nScenarios <- 100
-  #nScenarios <- 1000
+  #nScenarios <- 100
+  nScenarios <- 1000
   ## do model runs
@@
   #              0.18                 0.24                 0.25 
+  #new1000
+  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #             0.223                0.245                0.243
 
@@
   #              0.58                 0.57                 0.52 
+  #new1000
+  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #             0.558                0.563                0.534
 
@@
   #              0.83                 0.79                 0.72 
+  #new1000
+  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #             0.820                0.747                0.677
 
@@
  #                 0.136                  0.175                  0.173 
+  #new1000
+ #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+#                 0.163                  0.178                  0.182
 
@@
 #                  0.241                  0.326                  0.313 
+  #new1000
+# betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+#                 0.283                  0.302                  0.283
 
@@
 #                  0.672                  0.560                  0.488   
+  #new1000
+# betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+#                 0.622                  0.536                  0.458
 
@@
   treeInput <- rbind( treeInput, resistBetterMix3Seq )
+  #add >20% results
+  treeInput <- rbind( treeInput, resistBetterMix1Seq20 )
+  treeInput <- rbind( treeInput, resistBetterMix2Seq20 )
+  treeInput <- rbind( treeInput, resistBetterMix3Seq20 )
 
@@
                       rownames(resistBetterMix2Seq),
-                      rownames(resistBetterMix3Seq) )
+                      rownames(resistBetterMix3Seq),
+                      rownames(resistBetterMix1Seq20),
+                      rownames(resistBetterMix2Seq20),
+                      rownames(resistBetterMix3Seq20)
+                      )
 

### 2015-07-22 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 21/7/15 **
 check that the example runs still support the args the way they did before
@@
 
+I'm a bit worried that the post-processed results for the new100 runs seem not to have changed even though
+
+** 22/7/15 **
+analysing 1000 rereruns
+
+betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+               0.223                0.245                0.243 
+betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+               0.558                0.563                0.534 
+betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+               0.820                0.747                0.677
+
+rounded
+betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+               0.22                0.25                0.24
+betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+               0.56                0.56                0.53
+betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+               0.82                0.75                0.68
+
+20% better :
+ betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+                 0.163                  0.178                  0.182 
+ betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+                 0.283                  0.302                  0.283 
+ betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+                 0.622                  0.536                  0.458 
+rounded               
+ betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+                 0.16                  0.18                  0.18 
+ betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+                 0.28                  0.30                  0.28 
+ betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+                 0.62                  0.54                  0.46
+
+reran trees
+
+Best ones :
+
+to try to find an approximate general rule
+
+Favouring mixtures over sequential use.
+insecticide effectiveness (phi) > 0.7
+exposure < 0.7
+
+e.g. see treebetterMix2Seq_cP0.5.jpg
+
+I haven't done 20% better trees yet. Should I ?
+
 ### TO DO
 
-1. set 1000 runs to do tonight
+1. choose 1 tree to present to Ian & maybe this should be the most useful applied one ?
+(which might be mix2 at 20% & 50% resistance)
 
 1. redo the word results doc
+
+1. I could add the combined mixture/individual (strategy 2) as a dotted line to plotcurtis_f2_generic()
+(and then could even add labels of 1,2,3 to where they cross the 50% line)
 

### 2015-07-30 sensiAnPaper1All.Rmd
improve tree pruning by allowing for error

@@
 
+  #renaming other rownames to make nicer plots
+  rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_insecticide1"
+  rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_insecticide2"
+
   #add an extra predictor, the lower starting freq of resistance divided by the larger
@@
   #to keep it simple and transparent
-  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #prettifying names
+  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
   treeInput <- treeInput[ treePredictors, ]
+
+  #replace 0s & 1s with mixture/sequence to make plot clearer
+  repMix <- function(x){
+    x[x==0] <- 'sequence'
+    x[x==1] <- 'mixture'
+    x
+  }
+
+  resistBetterMix1Seq <- repMix(resistBetterMix1Seq)
+  resistBetterMix2Seq <- repMix(resistBetterMix2Seq)
+  resistBetterMix3Seq <- repMix(resistBetterMix3Seq)
+
+  resistBetterMix1Seq20 <- repMix(resistBetterMix1Seq20)
+  resistBetterMix2Seq20 <- repMix(resistBetterMix2Seq20)
+  resistBetterMix3Seq20 <- repMix(resistBetterMix3Seq20)
 
@@
 ### do tree analysis
-```{r, eval=FALSE}
+```{r, eval=TRUE}
   require(rpart.plot)
@@
   treeInput <- t(treeInput)
+
   #convert to a dataframe
-  treeInput <- data.frame(treeInput)
+  treeInput <- data.frame(treeInput, stringsAsFactors = FALSE)
+
+  #convert predictor columns back to numeric
+  #following this above : replace 0s & 1s with mixture/sequence to make plot clearer
+  filter <- !substr(names(treeInput),1,6) %in% 'better'
+  treeInput[,filter] <- lapply(treeInput[,filter],as.numeric) 
+  filter <- !filter
+  treeInput[,filter] <- lapply(treeInput[,filter],as.factor)
+
   #create string with predictor names in
@@
     tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
-    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+    #treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+
+    #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
+    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
+
+    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
+
+    treePruned <- prune(tree, cp=cpOptimal)
+
+    #to set box colours
+    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
+
     #adds in no. scenarios correctly and incorrectly classified
     #prp(tree, extra=1, varlen = 10, main=treeResponse)
-    prp(treePruned, extra=1, varlen = 10, main=treeResponse)
+    #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
+    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, col=cols, border.col=cols)
+    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
 
@@
     #trying to limit num levels a different way using maxdepth
-    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
+    #tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
     #adds in no. scenarios correctly and incorrectly classified
-    prp(tree, extra=1, varlen = 10, main=treeResponse)
-    savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')
+    #prp(tree, extra=1, varlen = 10, main=treeResponse)
+    #savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')
 
   }
+```
 
+### testing tree analysis for paper
+```{r, eval=TRUE}
+
+    x11()
+
+    treeResponse <- "betterMix2Seq_cP0.5"
+
+    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
+    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+
+    #to set box colours
+    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
+
+    #adds in no. scenarios correctly and incorrectly classified
+    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
+
+    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
+
+    treePruned <- prune(tree, cp=cpOptimal)
+
+    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
+
+    #comparing it with 0/1 gives the same big tree
+    treeInput$betterMix2Seq_cP0.5numeric <- as.numeric(treeInput$betterMix2Seq_cP0.5)
+
+    treeResponse <- "betterMix2Seq_cP0.5numeric"
+
+    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
+    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+
+    #to set box colours
+    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
+
+    prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
+
+    #testing alternative pruning
+    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
+    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+
+```
+
+```    
 #   treeResponse <- 'betterMix1Seq_cP0.5'
@@
     # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)
-
 ```

### 2015-07-30 resistanceModelNotesAndy.Rmd
notes

@@
 I haven't done 20% better trees yet. Should I ?
+done now
+
+** 23/7/15 **
+prettifying trees
+changed response from 0/1 to mixture/sequence
+
+seems that it may have made the trees bigger, ARSE.
+
+is it to do with pruning ?
+
+tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
+
+** 30/7/15 **
+
+checking on trees, and why they seem to change when I tweak formatting options
+add a check into sensiAnPaper1All.Rmd
+
+The tree is the same at the top (if reversed) but has more branches when 0/1 was changed to sequence/mixture.
+
+treeResponse <- "betterMix2Seq_cP0.5"
+
+tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
+#[1] 0.01
+
+What was this error before I made the conversion to text ?
+
+I tried converting back to numeric and I got the same big tree ...
+
+but the value of the error is lightly different :
+tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
+[1] 0.01072961
+
+This is what the cptable looks like for the numeric version
+numeric
+> tree$cptable
+          CP nsplit rel error    xerror       xstd
+1 0.34549356      0 1.0000000 1.0000000 0.03385148
+2 0.12875536      1 0.6545064 0.8047210 0.03285260
+3 0.05150215      2 0.5257511 0.6931330 0.03173289
+4 0.04184549      3 0.4742489 0.5107296 0.02889882
+5 0.01824034      5 0.3905579 0.4527897 0.02768814
+6 0.01287554      7 0.3540773 0.4163090 0.02683382
+7 0.01180258      8 0.3412017 0.4098712 0.02667504
+8 0.01072961     12 0.2939914 0.4012876 0.02645941
+9 0.01000000     13 0.2832618 0.4098712 0.02667504
+
+mixture/sequence
+> tree$cptable
+          CP nsplit rel error    xerror       xstd
+1 0.34549356      0 1.0000000 1.0000000 0.03385148
+2 0.12875536      1 0.6545064 0.7532189 0.03238844
+3 0.05150215      2 0.5257511 0.6287554 0.03088571
+4 0.04184549      3 0.4742489 0.4957082 0.02860113
+5 0.01824034      5 0.3905579 0.4699571 0.02806478
+6 0.01287554      7 0.3540773 0.4527897 0.02768814
+7 0.01180258      8 0.3412017 0.4420601 0.02744467
+8 0.01072961     12 0.2939914 0.4377682 0.02734549
+9 0.01000000     13 0.2832618 0.4442060 0.02749387
+
+Aha! when I repeat identical model, cptable is different 
+> tree$cptable
+          CP nsplit rel error    xerror       xstd
+1 0.34549356      0 1.0000000 1.0000000 0.03385148
+2 0.12875536      1 0.6545064 0.7618026 0.03247193
+3 0.05150215      2 0.5257511 0.6351931 0.03097749
+4 0.04184549      3 0.4742489 0.5085837 0.02885696
+5 0.01824034      5 0.3905579 0.4635193 0.02792537
+6 0.01287554      7 0.3540773 0.4248927 0.02704169
+7 0.01180258      8 0.3412017 0.4163090 0.02683382
+8 0.01072961     12 0.2939914 0.4098712 0.02667504
+9 0.01000000     13 0.2832618 0.4098712 0.02667504
+
+Can I get back to the nice smaller, neater tree ??
+
+http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
+
+A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
+
+http://stackoverflow.com/questions/15318409/how-to-prune-a-tree-in-r
+
+This does that :
+
+nsplit <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"nsplit"][1]
 
@@
 
-1. choose 1 tree to present to Ian & maybe this should be the most useful applied one ?
-(which might be mix2 at 20% & 50% resistance)
-
-1. redo the word results doc
+1. what is xerror in tree$cptable it seems to be not constant between analyses ??? It is the cross validation parameter.
+
+1. compare sequence/mixture tree with 0/1 for 1 of the chosen trees
+
+1. is using the optimal error the way that susannah has suggested a problem ?
 

### 2015-07-31 R/plotcurtis_f2_generic.r
defend against error if resistance not reached

@@ -82,6 +82,12 @@ plotcurtis_f2_generic <- function( combmat, bmat, amat, gencol=1, r1col=2, r2col
   hch_cutoff <- min(which((amat[,r1col])>max(criticalPoints)))
   ddt_cutoff <- min(which((bmat[,r2col])>max(criticalPoints)))
   
+  # check if resistance has not been reached
+  # and prevent an error, by seeting to max generations from input
+  # beware this may have other unintended consequences
+  if (is.infinite(hch_cutoff)) hch_cutoff <- nrow(amat)
+  if (is.infinite(ddt_cutoff)) ddt_cutoff <- nrow(bmat)  
+  
   # set results columns as vectors from single insecticide treatment
   hch <- amat[1:hch_cutoff,r1col]	# trim vector to point where freq of R > 0.5
   hch <- log10( 100 * hch )

### 2015-07-31 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 31/7/15 **
+give more sensible error than "need finite 'xlim' values" from shiny curtis fig2 app
+set starting freq of resistance to 0.0001 for each to generate the error.
+
+this is when resistance is not reached within max generations.
+
+runcurtis_f2()
+plotcurtis_f2_generic()
+Error in 1:ddt_cutoff : result would be too long a vector
+
 ### TO DO
 
-1. what is xerror in tree$cptable it seems to be not constant between analyses ??? It is the cross validation parameter.
-
-1. compare sequence/mixture tree with 0/1 for 1 of the chosen trees
-
-1. is using the optimal error the way that susannah has suggested a problem ?
+1. give more sensible error than "need finite 'xlim' values" from shiny curtis fig2 app
 

### 2016-01-13 resistanceModelNotesAndy.Rmd
notes

@@
 
+fixed by adding in this code to plotcurtis_f2_generic() 
+  # check if resistance has not been reached
+  # and prevent an error, by seeting to max generations from input
+  # beware this may have other unintended consequences
+  if (is.infinite(hch_cutoff)) hch_cutoff <- nrow(amat)
+  if (is.infinite(ddt_cutoff)) ddt_cutoff <- nrow(bmat)
+
+uploaded new version to shinyapps
+
+** 26/8/15 **
+submit invoice
+from contract : maximum of 25 days at 217 per day
+
+I spent 30.6 days (can claim some against tsetse work, or from future contract ...)
+
+** 6/10/15 **
+Chatting to Joe Lines at OpenMalaria TAG
+~ liked the shiny front end
+~ asked when he would be able to play with the model
+~ would like to see plots of parameter space, 2D & 3D when mixture better than sequential 
+~ suggested that Curtis does not say mixtures are bad (& that Ian may have slightly skewed impression)
+~~ I said that sentence in GPIRM might suggest that mixtures not as good
+~ he said with classification trees he would like to see an option to influence which branches are done first based upon the operational ability to alter them
+# to do re Joe
+~ send Joe link to the other shiny app
+~ send him the sentence from GPIRM
+~ ask Joe for a paper copy of GPIRM
+
+Pre-existing resistance to one insecticide in the mixture could accelerate the development of double resistant vector populations because of linkage disequilibrium (Figure A9.1).
+
+GPIRM p121 "theoretical models suggest that mixtures might delay resistance longer than rotations
+or broad mosaics (1)."
+
+p122
+"there is general agreement that LLIN mixture products, when developed, could
+still contain pyrethroids, since the pyrethroid may still contribute
+to the effectiveness of the net despite resistance (this needs to
+be confirmed). Some agree that the rationale applies to IRS, and
+that pyrethroid mixtures could be considered in areas of pyrethroid
+resistance; others believe that this is inadvisable and likely to lead
+to rapid evolution of double resistance."
+
+p122 
+Pre-existing resistance to one insecticide in the mixture could accelerate the development of double resistant vector populations because of linkage disequilibrium (Figure A9.1).
+
+At current prices, rotations and mosaics are estimated to have the lowest cost effect among the IRS-based IRM interventions.
+They would increase the annual cost by 20–50%, whereas use of mixtures for IRS once developed, are projected to increase the cost by 30–100%, depending on the duration of the transmission
+season. Because mixtures have not yet been developed or marketed, the final cost could be higher or lower than current projections.
+
+1. World Health Organization. The technical basis for coordinated action against insecticide resistance: preserving the effectiveness of modern malaria vector control. Geneva, 2011.
+
+Curtis is ref. 7. 
+Perhaps I hadn't remembered very well what it said. Or maybe I was remembering more what is in the WHO 2011 publication ? Curtis paper not cited in WHO 2011.
+
+** 7/12/15 **
+
+email from Ian :
+
+I had a major session on the 2-locus insecticide-resistance manuscript over the weekend (in between DIY; what a dull life I live!).
+Anyway I am now fairly happy with it .
+I plan to submit a IR-modelling grant application on 12th January so would like to have the paper submitted by then.
+I am not sure if I will manage but at least it is a target...
+I assume you can see the current version in Dropbox.
+
+Andy,  can you commit a couple to days to it ,or however long it takes, charging to the malaria grant (i.e.
+the one working with Swiss TPH).
+Here is a shopping list, but I think #3 is the only one that may require a significant amount of coding
+
+As I note in the cover page:  The big change I made was that calculating fitness (equations 1 and 2) was previously described as occurring within the generations.
+In fact it does not depend on current genotype frequencies so only needs to be done once prior to each deployment scenario  i.e.
+before the simulation is iterated over the generations.
+Check the code...if it is  still calculated each generation, it can be moved out, calculated once prior to the iterations,  and computational speed may increase substantially.
+Andy, you might like to check that it does not currently depend on current genotype frequencies in the code.
+
+The classification trees are currently based on 1,000 samples.
+Can you increase them to 5,000.
+The tree for "combinations only better if last >20% longer" needs to be re-drawn with more transparent parameter names.
+I have updated Table 5 with some suggested variable names for the trees...
+feel free to change and/or shorten them if you wish.
+I assume it is trivial to rename the variables in the tree graphing package.
+I think we need the pruned trees (i.e.
+within 1 s.e.) for the main manuscript and the full, unpruned trees for the supplementary material.
+
+At the moment we simply re-create the results from Curtis.
+I think a slight extension will greatly improve the ms by demonstrating the flexibility of the methodology.
+I have proposed two small extensions: as described on  Page 14/15.
+One is for biological realism i.e.
+male Anopheles mosquitoes are less likely to enter the house so will typically have lower exposure to the insecticide than females.
+The second is operational: even if we propose to deploy a mixture then various factors (theft, sloth, stock-outs etc) mean that one insecticide will often be absent from the mixture so the other insecticide is sprayed alone.
+As before, a sensitivity analysis with 5,000 runs.
+I think we just use these as examples then wrap up the ms by arguing that many other examples could be explored but we lack time/space so have focussed on describing the basic methodology to be used in future more policy-orientated publications
+
+That should keep you (or more accurately your computer) busy over Christmas.
+We can phone and discuss if you want, just nominate a time.
+I am away in Geneva Wed to Friday this week inclusive so could not talk then.
+Sorry I can't be here on Thursday for a pint but I'm sure you will manage.
+
+Best wishes,
+Ian
+
+Other bits from the manuscript itself :
+
+Jo Lines/Andy conversation at TAG.
+Suggest a few graphs where just change one parameter at a time.
+Need to resurrect our default parameters and easy to do….include them on Table 5 along with Curtis default parameters in other columns.
+
+Andy.
+Can we include the synergism/antagonism/cross-resistance factors Λ in the code; see calibration section of methods, page 6 and Table S1
+
+I think the classification trees each  need 10,000 rather than 1,000 runs. Don’t start clogging up your computer until we decide the final format.
+
+Recreate the “mixtures are only better if last 20% longer” classification tree for Figure 3.
+I know you have already done it but the nodes all had technical symbols that are difficult to interpret.
+They need simple names like in Figure 3A.
+I assume its just a question of changing the parameter names in a “R” plotting script.
+
+Run a new set of simulations using the “Curtis extended scenario” described on Page 16 to produce new classification trees again using the two criteria used previously i.e.
+(a) mixtures better if last >1 generation more than sequential (b) mixtures only better if last >20% longer
+
+page 8 :
+Our simulations will assume that the effects of insecticides in a mixture are multiplicative e.g.
+if the probability of surviving exposure to insecticide A alone is 0.3 and of surviving insecticide B alone is 0.2, then the probability of surviving expose to a  mixture of A and B is 0.3*0.2=0.06.
+In other words there is no synergy or antagonistic interactions between the insecticides.
+We do build flexibility into our methodology by scaling the fitnesses in niches where mosquitoes encounter both insecticides by a factor ΛAB, ΛAb, ΛaB or Λab (Table S1).
+If insecticides are synergistic then Λ<1, but if insecticides are antagonistic, for example because they share the same target site or share a similar mechanism of (cross)resistance, then  Λ>1.
+This is a flexible approach but other strategies are also possible, for example if the insecticides are highly antagonistic then the fitness may be defined as the locus with the higher individual fitness factor, w.
+
+Where would this 'synergism/antagonism/cross-resistance' bit need to go in code ?
+scaling the fitnesses in niches where mosquitoes encounter both insecticides by a factor ΛAB, ΛAb, ΛaB or Λab (Table S1).
+If insecticides are synergistic then Λ<1, but if insecticides are antagonistic, for example because they share the same target site or share a similar mechanism of (cross)resistance, then  Λ>1
+
+in runModel2.r, probably Wniche[locus1, locus2, niche1, niche2]
+Around line 300, just after this :
+ ## Two locus fitnesses in two insecticide Niche
+
+    for( nicheNum1 in 1:3 ) #todo get this 1:3 from somewhere
+    {
+      for( nicheNum2 in 1:3 ) #todo get this 1:3 from somewhere
+      { 
+        #temporary solution
+        #to get both niche (one of 0aAbB)
+        #and exposure (one of no,lo,hi)
+        niche1 <- dimnames(Wniche)$niche1[ nicheNum1 ]
+        niche2 <- dimnames(Wniche)$niche2[ nicheNum2 ]
+        exposure1 <- dimnames(Wloci)$exposure[ nicheNum1 ]
+        exposure2 <- dimnames(Wloci)$exposure[ nicheNum2 ]
+
+        #if this niche toggled off set fitness to 0
+        if (niche[niche1,niche2] == 0)
+        {
+          Wniche[,,niche1,niche2] <- 0
+        } else{
+          #otherwise set fitness to product of the 2 loci
+          for( locus1 in dimnames(Wniche)$locus1)
+          {
+            for( locus2 in dimnames(Wniche)$locus2)
+            {   
+              #6/1/16 i think ians new insecticide interaction parameter can just go here
+              #does in need to be just one param or 4 ?
+              #ΛAB, ΛAb, ΛaB or Λab 
+              #Wniche[locus1,locus2,niche1,niche2] <- interaction * Wloci[locus1,exposure1] * Wloci[locus2,exposure2]
+              Wniche[locus1,locus2,niche1,niche2] <- Wloci[locus1,exposure1] * Wloci[locus2,exposure2]
+            }
+          }          
+        }
+      }
+    }
+
+** 18/12/15 **
+~ what do I need to do to get labels correct in all trees ?
+
+Looking at the requests from Ian.
+I think it would be better if he & I sat down to go through it before I spend time doing 1000s more runs.
+
+** 6/1/16 ** on train to liverpool
+
+Questions for Ian :
+~ I recomend a few days, maybe a week, finishing the restructurng of the model. It still doesn't conform well to good practice.
+done ~ we could go through the model itself, checking Ians code points.
+hat ~ What can we call the insecticide interaction parameter see 290 in runModel2() ?
+4 ~ Is it just 1 param or do we need 4 ?
+no ~ should I remove W.bar arounf line 460 in runModel2()
+~ can we go over the curtis extended scenario p16 ?
+
+Linkage dis is bad in the context of mixtures because mosquitos are resistant to both insecticides more often than would be expected by chance.
+So problem is when the mixture is in use.
+LD decreases by half each generation under free recombination (recomb rate 0.5) .
+As genes get closer on chromosome recomb rate decreases.
+
+9/1/16 on train back from liverpool game rollout workshop
+
+modifying setExposure() function to cope with Ians request for lower male exposure
+I can add it in as a param with a default of 1.
+
+13/1/16
+starting on the new 1 parameter at a time sensi analysis in sensiAnPaperPartOneAtATime.r
+
 ### TO DO
 
-1. give more sensible error than "need finite 'xlim' values" from shiny curtis fig2 app
+See sensiAnPaperPartOneAtATime.r
+  #TODO I need to think more carefully about how to do this
+  #and maybe lookup to see if I can steal any existing sensi analysis type code
+  #and look back at my existing notes
+1. Graphs where just change one param at a time. Jo Lines/Andy conversation at TAG.
+
+Extended Curtis requests from Ian (still not quite sure how I will do)
+1. 'Male exposure' as a proportion of female (0.5 to 1)
+Exposure is held in array a, read from input file by sex, line 90.
+    a['m','0','0'] <- input[8,i]
+I can add a parameter to the setExposure() function which is called from sensiAnPaperPart.r
+
+1.
+'correct deployment' for mixtures (0.5 to 1) the proportion correctly exposed to the mixture, the rest are divided equally by being exposed to high concentrations of each individual each e.g.
+exposure 0.8 and correct deployment 0.7 : 0.8x0.7=0.56 are exposed to niche AB, while [0.8x0.3]/2=0.12 are exposed to niche A and 0.12 are exposed to niche B.
+Again this can be added as a parameter to setExposure()
+
+Main actions required from Ians last email :
+
+1.
+Can we include the synergism/antagonism/cross-resistance factors Λ in the code; see calibration section of methods and Table S1.
+This is about how effect of >1 insecticide are combined.
+see my notes above for where this is in R code.
+1.
+Run a new set of simulations using the “Curtis extended scenario” described on Page 16 to produce new classification trees again using the two criteria used previously i.e.
+(a) mixtures better if last >1 generation more than sequential (b) mixtures only better if last >20% longer
+1. Graphs where just change one param at a time. Jo Lines/Andy conversation at TAG.
+1.
+Recreate the “mixtures are only better if last 20% longer” classification tree for Figure 3.
+Replace technical symbols with simple names like in Figure 3A.
+(see sensiAnPaper1All.Rmd rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure")
+1. once we have decided what to do increase sensitivity analysis to 5000 or 10000 runs & run new trees
+
+We the run a “Curtis extended scenario ” to further illustrate the processes. The simulations have two additional parameters in addition to those described above and on Table 5 i.e.
+
+•	A “male exposure” parameter which defines the proportion of males exposed to an insecticide niche as a percentage of the females; it is assumed to be the same for each insecticide niche and varies uniformly from 50% to 100%.
+So, for example if females are 20% unexposed, 56% exposed to ‘AB’ and 12% exposed to each of niches’ A’ and ‘B’.
+If male exposure in 60% of the females their expose to niches ‘AB’,  ‘A’, and ‘B’ are  0.56*0.6= 33.6%, 0.12*0.6=7.2% and 0.12*0.6=7.2% respectively and the proportion unexposed rises to 0.2+(0.8*0.40)= 52%.
+
+•	A “correct deployment” parameter in the simulations examining the use of mixtures.
+Correct deployment is often not achieved for a variety of operational reasons and exposure to insecticides may not be as anticipated.
+In particular, mixtures may be mandated but not achieved for various reasons (such as stock-outs) such that the ‘mixtures’ are often replaced by the single insecticide that is available.
+We vary this factor uniformly between 50% and 100% and the incorrectly exposed mosquitoes are equally divided into exposure of the single insecticides.
+So, for example, if exposure is 80% and correct deployment” is 70% then within that 80% exposure to insecticides 0.8x0.7= 56% are exposed to niche AB, while [0.8x0.3]/2= 12%are exposed to niche A and [0.8x0.3]/2= 12% are exposed to niche B.
+In reality, poor deployment will result in exposure to lots of other niches i.e.
+a, b, Ab, Ab, ab but we only examine high levels of insecticide here to avoid having to additionally define survival probabilities to low insecticide levels ‘a’ and ‘b’.
+
+1. look at investigating response by changing single parameters at a time (or can I use existing runs)
+results files are in 'C:\\Users\\andy\\Documents\\resistanceResultsBackup\\listOutI1_1000.rda'
+listOut2, & listOutMix
+
+1. create a vignette that shows how best to run current model
+
+1. I could ad something to shinyCurtis with a message at the top saying whether mixture or sequence better and maybe an arrow pointing to it.
 
@@
 
+1. create functions fitnessLocus(), fitnessNiche(), fitnessIndiv()
+
 Remember.
@@
 Main question is : Tree1 is mixture better (0/1), Tree2 is mixture better by >20%.
+
+### current state of play 18/12/15
+1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
+1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
+1. runModel2.r          : runs model for the passed scenarios
 
@@
 
-1. in susana's script the 'outcome' is 0/1 if used in our case that would seem to lose a lot of information about the outcomes ? Even timeToFifty would be better for us.
-
 1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
@@
 1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
-1. submit tilde bug report
 
@@
 ** a ** exposure to insecticide
+** W.SS1_00 & W.SS2_00** fitness of SS in no insecticide
+** s ** selection coefficient
 ** phi ** fitness of one locus (baseline)
-** W.SS1_00 & W.SS2_00** fitness of SS in no insecticide
 ** h ** dominance coefficient
-** s ** selection coefficient
 ** z ** fitness cost of resistance allele in no insecticide

### 2016-01-14 sensiAnPaper1All.Rmd
adding ggplots of individual inputs

@@
 
+  #get the inputs here to use in ggplot investigation of model responses to inputs
+  #used in a later chunk
+  ggInput <- treeInput
+
   #replace 0s & 1s with mixture/sequence to make plot clearer
@@
 ```
+
+### investigating response of time-to-resistance to individual input parameters
+### (while they all are varying so it may not work well)
+```{r, eval=TRUE}
+
+  #uses ggInput calculated in earlier trunk
+
+  #first needs to transpose rows to cols
+  ggInput_T <- t(ggInput)
+
+  #function to transpose and add strategy column with the passed value
+  #also add on the input columns  
+  addStrategyColumn <- function(x, value, inputs){
+    x <- as.data.frame( t(x) )
+    x$strategy <- value
+    #transpose inputs
+    inputs <- as.data.frame( t(inputs) )
+    #cbind onto the outputs
+    x <- cbind(inputs,x)
+    x
+  }
+
+  resistPointsI1_T <- addStrategyColumn(resistPointsI1,"insecticide 1",ggInput) 
+  resistPointsI2_T <- addStrategyColumn(resistPointsI2,"insecticide 2",ggInput)   
+  resistPointsSeq_T <- addStrategyColumn(resistPointsSeq,"Sequential",ggInput) 
+  resistPointsMix1_T <- addStrategyColumn(resistPointsMix1,"Mixture 1",ggInput) 
+  resistPointsMix2_T <- addStrategyColumn(resistPointsMix2,"Mixture 2",ggInput) 
+  resistPointsMix3_T <- addStrategyColumn(resistPointsMix3,"Mixture 3",ggInput)
+
+  ggInsOuts <- rbind( resistPointsI1_T, resistPointsI2_T, resistPointsSeq_T, 
+                      resistPointsMix1_T, resistPointsMix2_T, resistPointsMix3_T)
+
+  #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
+
+  #cool working!
+
+  #this one shows no discernible pattern
+  ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=strategy)) + geom_point() + facet_wrap( ~ strategy) + geom_smooth()
+
+  #doing; for all inputs
+  names_inputs <- names(ggInsOuts)[1:10]
+
+  for(i in names_inputs)
+  {
+    x11()
+
+    y <- 'gen_cP0.5'
+    #color <- 'strategy'
+    #coloring by exposure shows its effect even when other params are varying
+    color <- 'exposure'
+
+    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
+             geom_point() + 
+             facet_wrap( ~ strategy) + 
+             geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+  }
+
+  #can I facet by particular parameter ranges ?
+  #or even subset the data for specific parameter ranges
+
+```

### 2016-01-14 resistanceModelNotesAndy.Rmd
notes

@@
 
-9/1/16 on train back from liverpool game rollout workshop
+** 9/1/16 ** on train back from liverpool game rollout workshop
 
@@
 
-13/1/16
+** 13/1/16 **
 starting on the new 1 parameter at a time sensi analysis in sensiAnPaperPartOneAtATime.r
+
+** 14/1/16 **
+first see if I can make any informative plots with the existing sensitivity analysis data
+
+These outputs are the frequencies by generation of the resistance alleles for the different insecticides (1&2) 
+Gen, m.R1, m.R2, f.R1, f.R2
+
+I want to get to a single file that has inputs & time_to_resistance thresholds for each scenario.
+
+I must have something nearly like this already from sensiAnPaper1All.Rmd
+
+Yes : resistPointsI1,2,Mix etc.
+For each scenario these have gen_cP0.1,0.25,0.5 which is the number of generations taken to reach those resistance thresholds.
+
+The next chunk works out which strategy is better than which.
+
+Maybe I want to bind all the resistPoints together into one file and bind onto the inputs, so that we can look at how the inputs effect the time to resistance for the different strategies.
+
+I want input on the x axis & time to resistance on the y.
+(but there will be the issue that the y values will be affected by the changing values of other parameters too) Cross that bridge when I come to it.
+
+how to get data in format for ggplot2
+check the diamonds dataset in ggplot2
+
+so I want to get as a dataframe with scenarios in rows, & columns for inputs & outputs
+
+so I can do something like :
+ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point()
+
+If I have a column that is strategy (insecticide1, insectide2, sequential, mixture)
+Then I would be able to facet by these later.
+
+ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
+
+This could be good as I can see the different strategies side by side.
+
+So I want columns :
+input1, input2 ... strategy, outputs : gen_cP0.1, 0.25, 0.5
+
+so the dataframe will have length=num strategies*num scenarios
+
+I'll need to rbind the strategies on top of each other.
+
+Some success with ggplot plots.
 
@@
 
+Files :
+listOut* I1,I2, Mix : raw results for insecticide1,2 & mixture
+                    : results are in $results[[scenarioNum]]
+                    : e.g. listOutI1$results[[1]] has columns
+                    : colnames(listOutI1$results[[1]])
+                    : "Gen" "m.R1" "m.R2"  "m.LD" "f.R1"  "f.R2" "f.LD" "M"  "F"  "dprime" "r2"
+treeInput : a matrix with scenarios in columns, inputs & outputs in rows
+
 Good simple test
@@
 
-1. is there a way I can easily run single & combined deployments for the same scenario ?
-* is it `niche_00` and `niche_AB`
-* or in `a.m_AB` insecticide exposure male hi1 hi2
-* do I also need to change "Starting frequency of resistance" ?
-* perhaps I can set up a simple example input to send to Beth ?
-* or ask Beth to send me the input.csv used to produce tables 8 & 9 in 'Brief_full-genotype-tables.pdf'
-* maybe comparing the files in full-dom_CD.csv & SD in C:\Dropbox\Levick et al IR manuscript will help me.
-
 1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
@@
 
-### refactoring variables to arrays ?
-Refactoring to arrays will reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions.
+### refactoring variables to arrays
+Refactoring to arrays reduces code volume, reduces risk of mistyped variable names and allow objects to be passed between functions.
 cis/trans makes it slightly tricky, but not all variables have cis/trans
@@
 ** A, B, a, b, -** 2 insecticides at 3 levels; low(ab), high(AB), absent(--) resulting in 9 niches.  
-
 Locus1 relates to resistance for insecticide A, and locus2 for B. Locus2 does not effect exposure to A.

### 2016-01-15 effectOfIndividualInputs.Rmd
plots of effect of individual inputs

--- 2015-07-22 sensiAnPaper1All.Rmd
@@
 ---
-title: "sensiAnPaper1All"
-output: html_document
+  title: "effectOfIndividualInputs"
+  output: pdf_document
+  author: "Andy South"
+  date: "`r Sys.Date()`"
 ---
 
-### All code for the sensitivity analysis for the first paper
+# Investigation into effects of individual inputs on time to reach resistance thresholds.
+# version 2 - in progress
 
-The main question is under which scenarios are mixtures better than sequential ?
+## Based on 1000 runs with all inputs varying at once which explains the variability at particular input values.
+## Within each figure the plots are divided into 6 sub-plots according to the insecticide use strategy and when the resistance threshold is reached
+1. insecticide 1 : sole use
+1. insecticide 2 : sole use
+1. Mixture 1 : threshold reached for either insecticide in mixture
+1. Mixture 2 : once threshold reached for either insecticide in mixture, switch to sole use of other until it too reaches threshold
+1. Mixture 3 : threshold reached for both insecticides in mixture
+1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold
 
-This breaks down into these subsidiary quetsions :
-1. Under which scenarios is the time to resistance for either insecticide quicker for
+These plots show gen_cP0.2, which is the number of generations to reach 20% resistance.
 
-```{r}
+Red dashed lines are a smoothed mean.
 
-  #library(devtools)
-  #install_github('AndySouth/resistance')
+The plots show a set of points at 1000 generations.
+The model is run for 500 generations, any runs which have not reached the resistance threshold by this time are given a value of 1000.
+This has little effect on questions of whether mixtures or sequential use is better, but does effect the plots presented here.
+
+```{r,echo=FALSE,results='hide'}
+
   library(resistance)
@@
   outFolder <- "C:\\Dropbox\\resistanceResults\\"
-```
 
-<!-- eval=FALSE to stop slow code being executed --> 
-```{r, eval=FALSE}
-
-  #default setting of the random seed in sensiAnPaperPart() should make the input files mostly the same with exception of exposure
-
-  #nScenarios <- 100
-  nScenarios <- 1000
-  ## do model runs
-  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture' )
-  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1' )
-  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2' )
-
-  ## save results objects as rda for re-use
-  save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
-  save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
-  save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))
-
-```
-
-```{r}
-  ## to load previously saved runs, e.g. these that I added 100 onto filename
+  ## to load previously saved runs
   load(file=paste0(outFolder,'listOutMix_1000.rda'))
@@
   load(file=paste0(outFolder,'listOutI2_1000.rda'))
-
 ```
 
+```{r,echo=FALSE,results='hide'}
+### chunk copied from sensiAnPaper1All
 ### now want to calculate the times to reach critical points and add them on to the input file 
 ### for different insecticide strategies, sequential, mix1 and mix2
-```{r}
+
   #1) sequential : time to resistance for each insecticide in isolation
@@
 
-### calculate the difference between sequential & mixture scenarios
-```{r}
-
-  require('plyr')
-
-  ## is mix1 better than sequential ?
-  #T/F
-  resistBetterMix1SeqBoolean <- resistPointsMix1 > resistPointsSeq
-  #convert to 0/1
-  resistBetterMix1Seq <- plyr::aaply(resistBetterMix1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
-  #rename rows ready for binding on to inputs
-  rownames(resistBetterMix1Seq) <- gsub('gen','betterMix1Seq', rownames(resistBetterMix1Seq))
-
-  rowSums(resistBetterMix1Seq)/ncol(resistBetterMix1Seq)
-  #100runs
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
-  #                .25                   .26                   .30   
-  #1000runs
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
-  #               .178                  .232                  .238 
-  #new100
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
-  #              0.18                 0.24                 0.25 
-  #new1000
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
-  #             0.223                0.245                0.243
-
-  ## is mix2 better than sequential ?
-  #T/F
-  resistBetterMix2SeqBoolean <- resistPointsMix2 > resistPointsSeq
-  #convert to 0/1
-  resistBetterMix2Seq <- plyr::aaply(resistBetterMix2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
-  #rename rows ready for binding on to inputs
-  rownames(resistBetterMix2Seq) <- gsub('gen','betterMix2Seq', rownames(resistBetterMix2Seq))
-
-  rowSums(resistBetterMix2Seq)/ncol(resistBetterMix2Seq)  
-  #1000runs
-  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
-  #             0.471                0.568                0.551 
-  #new100
-  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
-  #              0.58                 0.57                 0.52 
-  #new1000
-  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
-  #             0.558                0.563                0.534
-
-  ## is mix3 better than sequential ?
-  #T/F
-  resistBetterMix3SeqBoolean <- resistPointsMix3 > resistPointsSeq
-  #convert to 0/1
-  resistBetterMix3Seq <- plyr::aaply(resistBetterMix3SeqBoolean,.margins=c(1,2),.fun=as.numeric)
-  #rename rows ready for binding on to inputs
-  rownames(resistBetterMix3Seq) <- gsub('gen','betterMix3Seq', rownames(resistBetterMix3Seq))
-
-  #tally runs where mix3 better than seq
-  rowSums(resistBetterMix3Seq)/ncol(resistBetterMix3Seq)
-  #100runs
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
-  #                  .83                   .75                   .72 
-  #1000runs
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
-  #                .794                  .726                  .673  
-  #new100
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
-  #              0.83                 0.79                 0.72 
-  #new1000
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
-  #             0.820                0.747                0.677
-
-  #check1
-  #plot the 2 individual results & the mixture, just to check what is going on
-  #x11()
-#   plotallele.freq.andy(listOutMix)
-#   #x11()
-#   plotallele.freq.andy(listOutI1)
-#   #x11()
-#   plotallele.freq.andy(listOutI2)
-
-  #I should be able to compare single scenarios
-  #yes this works, but not too useful
-  # plotallele.freq.andy( list(results=list(listOutMix$results[[1]], listOutI1$results[[1]] )) )
-  #this is a modifdeied version of Beths function
-  #allowing curtis fig2 to be applied to any scenario
-  #plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )
-
-  #check2
-#   #what is the absolute difference between mix3 & sequential ?
-#   resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
-#   
-#   #check3 is their a difference between the mix & the individual strategies
-#   #looks mostly plausible, occasionally there are 0s
-#   resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
-#   resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2
-
-  ## is mix2 better than sequential ?  
-  # I might expect it to be similar to mix3
-
-  # comparing sensitivity inputs to those used by Beth to reproduce curtis fig2
-#   inputCurtis <- createInputMatrix(FALSE)
-#   inputSensiMix <- listOutMix$input
-#   
-#   input <- cbind(inputCurtis, inputSensiMix)
-#   write.csv(input, file=paste0(outFolder,'inputCurtisSensiCompare.csv'))
-
-```
-
-### finding when mixtures are >20% better
-### repeating calculation of the difference between sequential & mixture scenarios
-```{r}
-
-  require('plyr')
-
-  ## is mix1 better than sequential ?
-  #T/F
-  resistBetterMix1SeqBoolean20 <- resistPointsMix1 > (1.2 * resistPointsSeq)
-  #convert to 0/1
-  resistBetterMix1Seq20 <- plyr::aaply(resistBetterMix1SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
-  #rename rows ready for binding on to inputs
-  rownames(resistBetterMix1Seq20) <- gsub('gen','betterMix1Seq20', rownames(resistBetterMix1Seq20))
-
-  rowSums(resistBetterMix1Seq20)/ncol(resistBetterMix1Seq20)
-  #1000runs
- #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
- #                 0.136                  0.175                  0.173 
-  #new1000
- #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
-#                 0.163                  0.178                  0.182
-
-  ## is mix2 better than sequential ?
-  #T/F
-  resistBetterMix2SeqBoolean20 <- resistPointsMix2 > (1.2 * resistPointsSeq)
-  #convert to 0/1
-  resistBetterMix2Seq20 <- plyr::aaply(resistBetterMix2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
-  #rename rows ready for binding on to inputs
-  rownames(resistBetterMix2Seq20) <- gsub('gen','betterMix2Seq20', rownames(resistBetterMix2Seq20))
-
-  rowSums(resistBetterMix2Seq20)/ncol(resistBetterMix2Seq20)  
-  #1000runs
-#  betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
-#                  0.241                  0.326                  0.313 
-  #new1000
-# betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
-#                 0.283                  0.302                  0.283
-
-  ## is mix3 better than sequential ?
-  #T/F
-  resistBetterMix3SeqBoolean20 <- resistPointsMix3 > (1.2 * resistPointsSeq)
-  #convert to 0/1
-  resistBetterMix3Seq20 <- plyr::aaply(resistBetterMix3SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
-  #rename rows ready for binding on to inputs
-  rownames(resistBetterMix3Seq20) <- gsub('gen','betterMix3Seq20', rownames(resistBetterMix3Seq20))
-
-  #tally runs where mix3 better than seq
-  rowSums(resistBetterMix3Seq20)/ncol(resistBetterMix3Seq20)
-  #1000runs
-#  betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
-#                  0.672                  0.560                  0.488   
-  #new1000
-# betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
-#                 0.622                  0.536                  0.458
-
-```
-
+```{r,echo=FALSE,results='hide'}
+### chunk copied from part of one in sensiAnPaper1All
 ### bind results onto input file
-```{r}
 
@@
   #to keep it simple and transparent
-  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #prettifying names
+  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
   treeInput <- treeInput[ treePredictors, ]
 
-  #bind results onto predictors
-  treeInput <- rbind( treeInput, resistBetterMix1Seq )
-  treeInput <- rbind( treeInput, resistBetterMix2Seq )
-  treeInput <- rbind( treeInput, resistBetterMix3Seq )
-  #add >20% results
-  treeInput <- rbind( treeInput, resistBetterMix1Seq20 )
-  treeInput <- rbind( treeInput, resistBetterMix2Seq20 )
-  treeInput <- rbind( treeInput, resistBetterMix3Seq20 )
+  #get the inputs here to use in ggplot investigation of model responses to inputs
+  #used in a later chunk
+  ggInput <- treeInput
+```
+
+## This plot shows exposure which has the most noticeable pattern.
+```{r, eval=TRUE, echo=FALSE, warning=FALSE}
+
+  library(ggplot2)
+
+  #uses ggInput calculated in earlier chunk
+
+  #first needs to transpose rows to cols
+  #ggInput_T <- t(ggInput)
+
+  #function to transpose and add strategy column with the passed value
+  #also add on the input columns  
+  addStrategyColumn <- function(x, value, inputs){
+    x <- as.data.frame( t(x) )
+    x$strategy <- value
+    #transpose inputs
+    inputs <- as.data.frame( t(inputs) )
+    #cbind onto the outputs
+    x <- cbind(inputs,x)
+    x
+  }
+
+  resistPointsI1_T <- addStrategyColumn(resistPointsI1,"insecticide 1",ggInput) 
+  resistPointsI2_T <- addStrategyColumn(resistPointsI2,"insecticide 2",ggInput)   
+  resistPointsSeq_T <- addStrategyColumn(resistPointsSeq,"Sequential",ggInput) 
+  resistPointsMix1_T <- addStrategyColumn(resistPointsMix1,"Mixture 1",ggInput) 
+  resistPointsMix2_T <- addStrategyColumn(resistPointsMix2,"Mixture 2",ggInput) 
+  resistPointsMix3_T <- addStrategyColumn(resistPointsMix3,"Mixture 3",ggInput)
+
+  ggInsOuts <- rbind( resistPointsI1_T, resistPointsI2_T, resistPointsSeq_T, 
+                      resistPointsMix1_T, resistPointsMix2_T, resistPointsMix3_T)
+
+  #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
+
+  #cool working!
+
+  #exposure has the most noticeable effect
+  ggplot(ggInsOuts, aes(x=exposure, y=gen_cP0.25, color=strategy)) + 
+    geom_point() + 
+    facet_wrap( ~ strategy) + 
+    #geom_smooth(aes(x=exposure, y=gen_cP0.5, color=NULL))
+    geom_smooth(colour='red', linetype='dashed',size=0.5)
+
+  #effectiveness of insecticide1 little noticeable effect
+#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=gen_cP0.25, color=strategy)) + 
+#     geom_point() + 
+#     facet_wrap( ~ strategy) + 
+#     #geom_smooth(aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=NULL)) 
+#     geom_smooth(colour='red', linetype='dashed',size=0.5)
 
@@
 
-### do tree analysis
-```{r, eval=FALSE}
-  require(rpart.plot)
+## In the following plots all inputs are plotted individually, the points are coloured by exposure, which allows us to see that, even as other inputs change, low times to resistance are associated with high exposure.
+```{r, eval=TRUE, echo=FALSE}
 
-  #transpose
-  treeInput <- t(treeInput)
-  #convert to a dataframe
-  treeInput <- data.frame(treeInput)
-  #create string with predictor names in
-  treePredictorString <- paste(treePredictors, collapse="+")
+  #doing; for all inputs
+  names_inputs <- names(ggInsOuts)[1:10]
 
-  treeResponses <- c( rownames(resistBetterMix1Seq),
-                      rownames(resistBetterMix2Seq),
-                      rownames(resistBetterMix3Seq),
-                      rownames(resistBetterMix1Seq20),
-                      rownames(resistBetterMix2Seq20),
-                      rownames(resistBetterMix3Seq20)
-                      )
+  for(i in names_inputs)
+  {
+    #x11()
 
-  #to do trees & plots for all response variables
-  for( treeResponse in treeResponses )
-  {
-    x11()
-    #png(paste0(outFolder,"tree",treeResponse,".png"))
+    y <- 'gen_cP0.25'    
+    #y <- 'gen_cP0.5'
+    #color <- 'strategy'
+    #coloring by exposure shows its effect even when other params are varying
+    color <- 'exposure'
 
-    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
-    treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
-    #adds in no. scenarios correctly and incorrectly classified
-    #prp(tree, extra=1, varlen = 10, main=treeResponse)
-    prp(treePruned, extra=1, varlen = 10, main=treeResponse)
-
-    #dev.off()
-    #savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
-    savePlot(paste0(outFolder,"tree",treeResponse,".jpg"), type='jpg')
-
-    #trying to limit num levels a different way using maxdepth
-    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class', maxdepth=3) 
-    #adds in no. scenarios correctly and incorrectly classified
-    prp(tree, extra=1, varlen = 10, main=treeResponse)
-    savePlot(paste0(outFolder,"tree",treeResponse,"depth3.jpg"), type='jpg')
-
+    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
+             geom_point() + 
+             facet_wrap( ~ strategy) + 
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             geom_smooth(colour='red', linetype='dashed',size=0.5) )
   }
 
-#   treeResponse <- 'betterMix1Seq_cP0.5'
-#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
-#   
-#   #if get error number of rows of matrices must match (see arg 2)
-#   #is because all vals are 0
-# 
-#   plot(tree) #plots tree
-#   text(tree) #labels tree
-#   #branch predictors 2 phis & P_2  
-#     
-#   #for mix3 (resistance to both)
-#   x11()
-#   treeResponse <- 'betterMix3Seq_cP0.5'
-#   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  
-#   plot(tree) #plots tree
-#   text(tree) #labels tree  
-#   
-#   #branch predictors 2 phis & P_1
-#   
-#   #or from rpart.plot
-#   #prp(tree)
-#   
-#   #adds in no. scenarios correctly and incorrectly classified
-#   prp(tree, extra=1, varlen = 10, main=treeResponse)
-#   savePlot(paste0(outFolder,"tree",treeResponse,".png"), type='png')
-
-  #prp(tree, extra=1,box.col=c("pink","palegreen3")[tree$frame$yval],cex=0.5,yesno.yshift=-2)
-
-    #susanas code
-    #pruning tree
-    #   treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
-    #   plot(treePruned) #plots tree
-    #   text(treePruned) #labels tree 
-    # prp(prunned_tree, extra=1,box.col=c("pink", "palegreen3")[prunned_tree$frame$yval],cex=0.5,yesno.yshift=-2)
+  #can I facet by particular parameter ranges ?
+  #or even subset the data for specific parameter ranges
 

### 2016-01-15 effectOfIndividualInputs.Rmd
progressing

@@
   #I could just rename one to exposure
+  #BEWARE risk if future changes
   rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+  #hardcode which variables to include in analysis to keep it simple and transparent
+  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+
+  treeInput <- treeInput[ treePredictors, ]
+
+  #add an extra predictor, the lower starting freq of resistance divided by the larger
+  startResistanceRatio <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
+  treeInput <- rbind(treeInput,startResistanceRatio)
 
@@
 
-  #add an extra predictor, the lower starting freq of resistance divided by the larger
-  startResistanceRatio <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
-  treeInput <- rbind(treeInput,startResistanceRatio)
+  rownames(treeInput)[rownames(treeInput)=="P_1"] <- "start_freq_allele1"
+  rownames(treeInput)[rownames(treeInput)=="P_2"] <- "start_freq_allele2"
 
-  #hardcode which variables to include in the tree analysis here
-  #to keep it simple and transparent
-  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
-  #prettifying names
-  treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
-  treeInput <- treeInput[ treePredictors, ]
+  rownames(treeInput)[rownames(treeInput)=="h.RS1_A0"] <- "dominance_allele1"
+  rownames(treeInput)[rownames(treeInput)=="h.RS2_0B"] <- "dominance_allele2"
+
+  rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
+  rownames(treeInput)[rownames(treeInput)=="s.RR2_0B"] <- "selection_coef_allele2"
+
+  #subsetting now done above before names prettified
+  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #treeInput <- treeInput[ treePredictors, ]
 
@@
   ggInput <- treeInput
+
 ```
@@
 
+  #prettify output names
+  names(ggInsOuts)[names(ggInsOuts)=="gen_cP0.5"] <- "time_to_resistance0.5"
+  names(ggInsOuts)[names(ggInsOuts)=="gen_cP0.25"] <- "time_to_resistance0.25"
+  names(ggInsOuts)[names(ggInsOuts)=="gen_cP0.1"] <- "time_to_resistance0.1"
+
   #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
@@
   #exposure has the most noticeable effect
-  ggplot(ggInsOuts, aes(x=exposure, y=gen_cP0.25, color=strategy)) + 
+  ggplot(ggInsOuts, aes(x=exposure, y=time_to_resistance0.25, color=strategy)) + 
     geom_point() + 
@@
   #effectiveness of insecticide1 little noticeable effect
-#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=gen_cP0.25, color=strategy)) + 
+#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=time_to_resistance0.25, color=strategy)) + 
 #     geom_point() + 
@@
 
-    y <- 'gen_cP0.25'    
-    #y <- 'gen_cP0.5'
+    y <- 'time_to_resistance0.25'    
+    #y <- 'time_to_resistance0.5'
     #color <- 'strategy'
@@
   #can I facet by particular parameter ranges ?
-  #or even subset the data for specific parameter ranges
+  #might need to create a categorical var from the range and use that to facet 
+  #or subset the data for specific parameter ranges
 

### 2016-01-15 resistanceModelNotesAndy.Rmd
progressing

@@
 
+** 15/1/16 **
+sent first version of ggplot doc (effectOfIndividualInputs.pdf) to Ian & Beth
+
+prettifying parameter names
+
 ### TO DO
+
+# try changing plot symbol to open circles
+# maybe add in a loop to plot for all 3 resistance thresholds
+# maybe try removing reps with num generations 1000 ?
+# maybe try faceting by parameter ranges ?
+# lookup to see if I can steal any existing sensi analysis type code
 
@@
   #TODO I need to think more carefully about how to do this
-  #and maybe lookup to see if I can steal any existing sensi analysis type code
   #and look back at my existing notes

### 2016-01-15 sensiAnPaper1All.Rmd
progressing

@@
 ```{r}
-  ## to load previously saved runs, e.g. these that I added 100 onto filename
+  ## to load previously saved runs
   load(file=paste0(outFolder,'listOutMix_1000.rda'))
@@
 
-  #uses ggInput calculated in earlier trunk
+  #uses ggInput calculated in earlier chunk
 
   #first needs to transpose rows to cols
-  ggInput_T <- t(ggInput)
+  #ggInput_T <- t(ggInput)
 

### 2016-01-20 effectOfIndividualInputs.Rmd
remove from plots runs where thresholds not reached

@@
 
+  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
+  #>1000 excludes sequential strategy that had a 999 in
+  #BEWARE would need to increase 1000 if I increase max generations in the runs
+  didntReachThresh <- which(ggInsOuts$gen_cP0.5 == 999 | ggInsOuts$gen_cP0.5 > 1000 |
+                            ggInsOuts$gen_cP0.25 == 999 | ggInsOuts$gen_cP0.25 > 1000 |
+                            ggInsOuts$gen_cP0.1 == 999 | ggInsOuts$gen_cP0.1 > 1000 )
+
+  #subset by runs not making threshold
+  ggInsOuts <- ggInsOuts[-didntReachThresh,]
+
   #prettify output names

### 2016-01-22 sensiAnPaper1All.Rmd
add predictor resist_start_1_div_2

@@
 
+  nScenarios <- 10  
   #nScenarios <- 100
-  nScenarios <- 1000
+  #nScenarios <- 1000
   ## do model runs
@@
   #add an extra predictor, the lower starting freq of resistance divided by the larger
-  startResistanceRatio <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
-  treeInput <- rbind(treeInput,startResistanceRatio)
+  resist_start_hi_div_lo <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
+  treeInput <- rbind(treeInput,resist_start_hi_div_lo)
+
+  #20160122 add test for Ian of resistance1/resistance2
+  resist_start_1_div_2 <- treeInput['P_1',]/treeInput['P_2',]
+  treeInput <- rbind(treeInput,resist_start_1_div_2)
 

### 2016-01-26 effectOfIndividualInputs.Rmd
adding option for extended experiment & start of PRCCs

--- 2016-01-15 effectOfIndividualInputs.Rmd
@@
 
-The plots show a set of points at 1000 generations.
+Runs where resistance thresholds have not been reached have been removed.
+
+Previous plots showed a set of points at 1000 generations.
 The model is run for 500 generations, any runs which have not reached the resistance threshold by this time are given a value of 1000.
-This has little effect on questions of whether mixtures or sequential use is better, but does effect the plots presented here.
+This has little effect on questions of whether mixtures or sequential use is better.
 
@@
   ## to load previously saved runs
-  load(file=paste0(outFolder,'listOutMix_1000.rda'))
-  load(file=paste0(outFolder,'listOutI1_1000.rda'))
-  load(file=paste0(outFolder,'listOutI2_1000.rda'))
+#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
+#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
+#   load(file=paste0(outFolder,'listOutI2_1000.rda'))
+
+  ## trying with the extended experiment _ex100
+  experiment <- 'extended'
+  load(file=paste0(outFolder,'listOutMix_ex1000.rda'))
+  load(file=paste0(outFolder,'listOutI1_ex1000.rda'))
+  load(file=paste0(outFolder,'listOutI2_ex1000.rda'))
+
 ```
@@
 
+  #add these for extended analysis
+  if (experiment=='extended')
+    treePredictors <- c(treePredictors,'maleExposureProp','correctMixDeployProp')
+
   treeInput <- treeInput[ treePredictors, ]
@@
 
+  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
+  #>1000 excludes sequential strategy that had a 999 in
+  #BEWARE would need to increase 1000 if I increase max generations in the runs
+  didntReachThresh <- which(ggInsOuts$gen_cP0.5 == 999 | ggInsOuts$gen_cP0.5 > 500 |
+                            ggInsOuts$gen_cP0.25 == 999 | ggInsOuts$gen_cP0.25 > 500 |
+                            ggInsOuts$gen_cP0.1 == 999 | ggInsOuts$gen_cP0.1 > 500 )
+
+  #subset by runs not making threshold
+  ggInsOuts <- ggInsOuts[-didntReachThresh,]
+
   #prettify output names
@@
   ggplot(ggInsOuts, aes(x=exposure, y=time_to_resistance0.25, color=strategy)) + 
-    geom_point() + 
+    geom_point(shape=3) + 
     facet_wrap( ~ strategy) + 
@@
   #doing; for all inputs
-  names_inputs <- names(ggInsOuts)[1:10]
+  if (experiment=='extended') {
+    num_inputs <- 13
+    } else {
+    num_inputs <- 11  
+    }
+
+  names_inputs <- names(ggInsOuts)[1:num_inputs]
 
@@
     print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-             geom_point() + 
+             geom_point(shape=3) + 
              facet_wrap( ~ strategy) + 
              #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             geom_smooth(colour='red', linetype='dashed',size=0.5) )
+             geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             labs(title = i))
   }
@@
 ```
+
+## Looking more closely at effect of the ratio between the level of resistance to the first insecticide divided by that of the second one. Restricting the x axis to lower values.
+```{r, eval=TRUE, echo=FALSE}
+
+  i <- 'resist_start_1_div_2'
+  color <- "start_freq_allele2"
+
+  print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
+           geom_point(shape=3) + 
+           facet_wrap( ~ strategy) + 
+           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+           geom_smooth(colour='red', linetype='dashed',size=0.5) +
+           labs(title = i) +
+           xlim(0,100)
+         )
+
+```
+
+#trying a PRCC analysis using the package sensitivity
+```{r, eval=TRUE, echo=FALSE}
+
+  library(sensitivity)
+
+# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
+# 
+# ## S3 method for class 'pcc'
+# plot(x, ylim = c(-1,1), ...)
+# 
+# Arguments
+# X a data frame containing model input variables
+# y a vector containing the responses 
+# rank logical. If TRUE, the analysis is done on the ranks.
+# nboot the number of bootstrap replicates.
+# conf the confidence level of the bootstrap confidence intervals.
+
+  #first attempt
+  by_strategy <- ggInsOuts[ggInsOuts$strategy=='insecticide 1',]
+
+  x <- by_strategy[,1:num_inputs]
+  y <- by_strategy['time_to_resistance0.25']
+
+  pcc_res <- pcc(x, y, rank=TRUE)
+
+  #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+  #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+  #plot(pcc_res)
+
+  #results are here I can probably rbind them together into a df that I can ggplot
+  #pcc_res$PRCC
+
+  to_plot <- pcc_res$PRCC
+  to_plot$test <- c(1:nrow(to_plot))
+  to_plot$names <- rownames(to_plot)
+
+  ggplot( to_plot, aes_string(x='names',y='original') ) + 
+           geom_point(shape=3) +
+           theme(axis.text.x = element_text(angle = 45))
+
+```

### 2016-01-26 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 20/1/16 ** on train to liverpool
+
+removed from plots runs where resistance threshold not reached
+
+** 22/1/16 ** with Ian in Liverpool
+I said realistically it would be a month before I could get started on the
+
+conventional wisdom of poor spraying increasing resistance is because poor spray thought not to kill heterozygotes.
+we see some o this effect in the dominannce param.
+Later we can loook at by the low exposure niches e.g.
+a vs A.
+We can look at in the next paper.
+
+Ian plans to submit to PLOS computationally biology with the aim that it's a flexible model that can address a bunch of situations.
+Do we want to be slightly carful to protect our IP ?
+e.g.
+in the first version w might not want to promote use by others (we'd need to do some more wrk to make it properly useable).
+I can discuss this with Ian when we are closer to publication.
+
+~ I did think that correctDeploymentProp only effects mixtureswhich would constrain tree analyses for whether mix better than seq, but actually analysis can still be done, you would just expect mix to be closer to seq.
+
+~ added maleExposureProp & correctMixDeployProp into tree & plot analyses, by adding on to the end of the input object.
+inputs are got from   treeInput <- listOutMix$input which is created by runModel2() just from the input object that is passed to it.
+
+~ added as an extra argument to sensiAnPaperPart() that would make less code repetition
+
+** 25/1/16 **
+added to effectOfIndividualInputs.Rmd
+~~ new predictors for males & deploy
+~~ calc of new ratio, copy from sensiAnPaperAll
+
+13.40 started 1000 reps of the 'extended' experiment 
+< 50 mins to run
+
+This is what Susanan did for her 2012 model :
+
+Simulations to understand the influence of each parameter
+on the outcome variables (mean fitness and change
+in resistance allele frequency) were performed using latin
+hypercube sampling (LHS) to generate a data set and
+partial rank correlation coefficients (PRCC) calculated to
+provide a quantitative measure of the impact of each
+parameter [24]. LHS techniques were first developed to
+explore the behavior of complex models in economics,
+engineering, chemistry and physics and have been used in
+models predicting the impact of insecticide-treated nets
+on malaria transmission [25].
+
+The analysis was performed using R software [26] and
+implementation of LHS using package lhs. It does not
+allow for the specification of each variable distribution
+beforehand, so sampling was performed assuming a uniform
+distribution. Once the sample was generated, the
+uniform sample from a column (variable) could be transformed
+to the required distribution (Table 3) by using
+quantile functions (using the qtriangle comand in R).
+
+A data set of 3,000 replications was generated, with random
+parameter sets and the corresponding values of the
+outcome variables using equations 8, 9, 12 and 13. Ten
+replicates of this procedure were performed as suggested
+in [24] to investigate the predictive precision of model
+using LHS as the sampling method. This was achieved
+by analysing each replicate separately and verifying that
+results were consistent across ten replicates.
+
+** 26/1/16 **
+looking into prcc analysis of model outputs &  whether I might want tio use lhs.
+
+~ what is difference between lhs and what I have done
+~ check lhs package 
+~ lhs would ensure a more uniform distributions of samples across parameter space, i would think results are unlikely to be different to our monte-carlo approach.
+~ it might be tricky for the params where we haven't used a uniform distribution (although we probably could get around)
+~ some advice on using lhs package
+http://r.789695.n4.nabble.com/latin-hypercube-sampling-td4659028.html
+
+~ add PRCC code
+This provides some (dated) advice on running a sensitivity anlaysis including LHS & prcc
+https://cran.r-project.org/web/packages/pse/vignettes/pse_tutorial.pdf
+
+pcc function in package sensitivity may do what we want :
+https://cran.r-project.org/web/packages/sensitivity/sensitivity.pdf
+
+pcc computes the Partial Correlation Coefficients (PCC), or Partial Rank Correlation Coefficients
+(PRCC), which are sensitivity indices based on linear (resp. monotonic) assumptions, in the case of
+(linearly) correlated factors.
+
+~ or could use epiR package
+epi.prcc(dat) # where final column in dat is the output, all other columns are inputs
+
 ### TO DO
 
-# try changing plot symbol to open circles
+~ If I would like to get prcc plots in a similar format to my existing plots.
+~ get data in form : 
+input1, input2, inputN, strategy 
+prcc vals,,             indecticide1
+
+# would it be useful for Ian for me to create an interactive shiny app to view the 1000 sim data. E.g. could just have the ggplot code with options to set the x value & colour by different inputs.
+
 # maybe add in a loop to plot for all 3 resistance thresholds
-# maybe try removing reps with num generations 1000 ?
 # maybe try faceting by parameter ranges ?

### 2016-01-26 sensiAnPaper1All.Rmd
add options for extended experiments

@@
 
-  nScenarios <- 10  
+  #nScenarios <- 10  
   #nScenarios <- 100
-  #nScenarios <- 1000
+  nScenarios <- 1000
+
+  #to run 'extended' or not
+  experiment <- 'extended' #'curtis'
+
   ## do model runs
-  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture' )
-  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1' )
-  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2' )
+  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
+  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment )
+  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment )
 
@@
 
+  save(listOutMix, file=paste0(outFolder,'listOutMix_ex1000.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1_ex1000.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2_ex1000.rda'))
+
 ```
@@
   ## to load previously saved runs
-  load(file=paste0(outFolder,'listOutMix_1000.rda'))
-  load(file=paste0(outFolder,'listOutI1_1000.rda'))
-  load(file=paste0(outFolder,'listOutI2_1000.rda'))
+#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
+#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
+#   load(file=paste0(outFolder,'listOutI2_1000.rda'))
+
+  load(file=paste0(outFolder,'listOutMix_ex1000.rda'))
+  load(file=paste0(outFolder,'listOutI1_ex1000.rda'))
+  load(file=paste0(outFolder,'listOutI2_ex1000.rda'))
 

### 2016-01-26 R/sensiAnPaperPart.r

@@ -108,7 +108,9 @@ sensiAnPaperPart <- function( nScenarios = 10,
                         h.RS2_0B = h.RS2_0B,
                         
                         s.RR1_A0 = s.RR1_A0,
-                        s.RR2_0B = s.RR2_0B
+                        s.RR2_0B = s.RR2_0B,
+                        maleExposureProp = maleExposureProp,
+                        correctMixDeployProp = correctMixDeployProp  
                         )   
     
     

### 2016-01-26 R/setInputOneScenario.r

@@ -59,6 +59,9 @@
 #' @param niche_Ab insecticide niche toggle hi1 lo2 0=off 1=on
 #' @param niche_aB insecticide niche toggle lo1 hi2 0=off 1=on
 #' @param sexLinked whether resistance is sex linked, default=0(FALSE)
+#' @param maleExposureProp male exposure as a propoertion of female, default 1 for same, likely <1  
+#' @param correctMixDeployProp proportion of correct deployment of mixtures, 
+#'    if <1 other portion divided between single insecticides
 #' 
 #' @return named vector
 #' @export
@@ -116,11 +119,13 @@ setInputOneScenario <- function( calibration = 100,
                                  niche_AB	=	1	,
                                  niche_Ab	=	1	,
                                  niche_aB	=	1 ,
-                                 sexLinked = 0
+                                 sexLinked = 0,
+                                 maleExposureProp = 1,
+                                 correctMixDeployProp = 1  
                                  )
 {
   
-  input <- matrix( ncol=1, nrow=53 )
+  input <- matrix( ncol=1, nrow=55 )
   
   input[1] <- calibration
   input[2] <- max_gen
@@ -198,14 +203,21 @@ setInputOneScenario <- function( calibration = 100,
   input[	52	] <-	niche_aB
   
   input[	53	] <-	sexLinked
+  
+  #22/1/16 new variables for extended experiment
+  #they aren't used in runModel2() but are needed for post run analyses
+  input[	54	] <-	maleExposureProp
+  input[	55	] <-	correctMixDeployProp
     
   a.m <- sum(a.m_00, a.m_a0, a.m_A0, a.m_0b, a.m_0B, a.m_ab, a.m_AB, a.m_Ab, a.m_aB)
-  if ( a.m != 1 ){		 
+  #if ( a.m != 1 ){
+  #these warnings allow for rounding differences
+  if ( !isTRUE( all.equal(1, a.m  ))){
   	stop("male exposures must total one, currently: ", a.m )
   	}
   
   a.f <- sum(a.f_00, a.f_a0, a.f_A0, a.f_0b, a.f_0B, a.f_ab, a.f_AB, a.f_Ab, a.f_aB)
-  if ( a.f != 1 ){		 
+  if ( !isTRUE( all.equal(1, a.f  ))){
     stop( "female exposures must total one, currently: ", a.f )
   	}
  

### 2016-02-01 resistanceModelNotesAndy.Rmd
notes

@@
 
+I got sensitivity::pcc() working for a single strategy.
+
+** 28/1/16 **
+~ fixed diagonal axis labels.
+~ get sensitivity::pcc() working for all strategies & putting data into format for faceted plots.
+
+** 29/1/16 **
+
+~ why does correctMixDeployProp seem to have an effect in the PRCC of insecticide1 when it shouldn't have any effect outside of mixtures ?
+
+In the plots againts time_to_resistance there is only a bit of a pattern for Mixture1 (time_to_R increaeses between 0.9 & 1 correct deployment).
+Points to it being a problem in the PRCC itself.
+
+~~ in setExposure() looks fine, its not applied to insecticide1 or 2
+
+other things it could be :
+~~ switched with another var in the call to setExposure
+~~ something about how runs are done in sensiAnPaperPart.r 
+~~ rows getting reordered within effectOfIndividualInputs.Rmd ?
+
+it is +vely correlated with time to resistance, so better application leads to longer times for resistance to be reached (not sure whether that is waht I would expect ?)
+
+aha PRCC results are influenced by the presence of the other vars, if just do for exposure & deployProp deployProp goes down to 0.2. 
+x2 <- x[c('correctMixDeployProp','exposure')]
+pcc_res2 <- pcc(x2, y, rank=TRUE)
+pcc_res2$PRCC
+                       original
+correctMixDeployProp  0.2167706
+exposure             -0.5248086
+
+I still don't quite understand why an essentially random var is correlated with the output ??
+
+This shows that there's very little correlation between the in & out, for insecticide1
+> cor(xy$time_to_resistance0.25,xy$correctMixDeployProp)
+[1] 0.003849606
+
+** 1/2/16 **
+
+Checking again on whats happening with correctMixDeployProp
+
+is the correlation similar for mixture1 as it is for insecticide1 ?
+
+could problem be that the inputs are got from a mix run ? shouldn't matter
+  treeInput <- listOutMix$input
+
+could it be something to do with the random number seed being disrupted by the extra params ? shouldn't happen
+
+i think the inputs should be identical.
+
+I think it may be a correlation in the analysis.
+
+What I have done is generate inputs for 'male proportion' and 'correct mix deploy'.
+Correct mix deployment is not used in the individual insecticide runs, but it is still included as a potential predictor in the PRCC analysis.
+
+In the mixture scenarios 'correct mix deploy' effects exposure ...
+
+exposure is not a single variable in the simulation, but is it used as a single value as a predictor in the sensitivity & prcc analyses ?
+
+aha the exposure predictor is taken from the female AB (actually I think that is OK ??)
+although maybe not because it will be reduced by correct deploy and maybe by male (or maybe not ???)
+
+  rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+correctDeploy has a strong interaction with exposure in mixtures.
+
+When correct deploy is low in mixture runs exposure to AB is much lower because of corresponding exposure to A0 and 0B.
+
+But i think this shouldn't effect single insecticide runs.
+
+Am I sure I'm not mixing up rows somewhere in effectOfIndividualInputs.Rmd ??
+
+I could try setting the values of results for I1 and see if it propogates through.
+
 ### TO DO
 
-~ If I would like to get prcc plots in a similar format to my existing plots.
+~ rerun the trees for Ian
+
+~ set 10000 runs going before yoga, but whether extended or normal curtis ?
+
+~ I might be able to do a PRCC of whether mixtures better than sequential (but actually thats what the trees do)
+
+~ I'd like to show prcc on an axis from 0-1 with diff sysmbols for neg & pos. We care about which are most imp. params, not whether they are pos or neg.
+
+~ If I would like to get prcc plots in a similar format to my existing plots. (but may not be space for axis labels)
 ~ get data in form : 

### 2016-02-01 effectOfIndividualInputs.Rmd
hunting anomaly

@@
 ## In the following plots all inputs are plotted individually, the points are coloured by exposure, which allows us to see that, even as other inputs change, low times to resistance are associated with high exposure.
-```{r, eval=TRUE, echo=FALSE}
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
 
@@
 ## Looking more closely at effect of the ratio between the level of resistance to the first insecticide divided by that of the second one. Restricting the x axis to lower values.
-```{r, eval=TRUE, echo=FALSE}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE}
 
@@
 #trying a PRCC analysis using the package sensitivity
-```{r, eval=TRUE, echo=FALSE}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
 
@@
 
-  #first attempt
-  by_strategy <- ggInsOuts[ggInsOuts$strategy=='insecticide 1',]
+  strategies <- unique(ggInsOuts$strategy)
+
+  for(strategy in strategies)
+  {
+
+    by_strategy <- ggInsOuts[ggInsOuts$strategy==strategy,]
 
@@
     to_plot <- pcc_res$PRCC
-  to_plot$test <- c(1:nrow(to_plot))
-  to_plot$names <- rownames(to_plot)
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
 
-  ggplot( to_plot, aes_string(x='names',y='original') ) + 
-           geom_point(shape=3) +
-           theme(axis.text.x = element_text(angle = 45))
+    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             geom_hline(yintercept = 0, linetype=3) +
+             ylim(-1,1) +
+             ggtitle(paste(strategy,"PRCC")) +
+             xlab(NULL)
+          )
+  }
 
 ```
+
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, fig.height=3.5}
+
+#trying to find apparent anomaly in PRCC analysis
+#what causes correctMixDeployProp to get a high PRCC score for I1 & 2 when its not used in those runs ??
+
+tstI1 <- round(listOutI1$input,3)
+tstMix <- round(listOutMix$input,3)
+tstDif <- round(tstMix - tstI1,3)
+
+```

### 2016-02-01 effectOfIndividualInputs.Rmd
exposure now taken from single random start val.

@@
   #BEWARE risk if future changes
-  rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+  #1/2/16 don't need this now because I've saved exposure from the single original random value
+  #rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
 

### 2016-02-09 paper1_results_figs.Rmd
new figures

--- 2016-02-01 effectOfIndividualInputs.Rmd
@@
 ---
-  title: "effectOfIndividualInputs"
+  title: "paper1_results_figs"
   output: pdf_document
@@
 
-# Investigation into effects of individual inputs on time to reach resistance thresholds.
+# Figures for paper 1 
 # version 2 - in progress
 
-## Based on 1000 runs with all inputs varying at once which explains the variability at particular input values.
-## Within each figure the plots are divided into 6 sub-plots according to the insecticide use strategy and when the resistance threshold is reached
-1. insecticide 1 : sole use
-1. insecticide 2 : sole use
-1. Mixture 1 : threshold reached for either insecticide in mixture
-1. Mixture 2 : once threshold reached for either insecticide in mixture, switch to sole use of other until it too reaches threshold
+## Based on 10000 runs with all inputs varying at once which explains the variability at particular input values.
+
+Comparing these insecticide use strategies:
+1. Single use
+1. Mixture 1 : threshold reached for either insecticide
+1. Mixture 2 : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
 1. Mixture 3 : threshold reached for both insecticides in mixture
@@
 
-These plots show gen_cP0.2, which is the number of generations to reach 20% resistance.
+## Notable results
 
-Red dashed lines are a smoothed mean.
-
-Runs where resistance thresholds have not been reached have been removed.
-
-Previous plots showed a set of points at 1000 generations.
-The model is run for 500 generations, any runs which have not reached the resistance threshold by this time are given a value of 1000.
-This has little effect on questions of whether mixtures or sequential use is better.
+1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
+1.
+In between these extremes, there is less difference between the alternative strategies of a) sequential and b) mixture switching to sole use.
+The latter takes slightly longer to reach the threshold.
+1. Exposure has the greatest effect on time-to-resistance, but as it has a similar effect across all strategies it has little influence on which strategy is preferable.
+1.
+Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) mixture switching to sole use.
+Below an effectiveness of ~0.55 to 0.6 sequential takes longer, and above this 'mixture switching to sole use' takes longer.
 
@@
   experiment <- 'extended'
-  load(file=paste0(outFolder,'listOutMix_ex1000.rda'))
-  load(file=paste0(outFolder,'listOutI1_ex1000.rda'))
-  load(file=paste0(outFolder,'listOutI2_ex1000.rda'))
+
+  #remember these don't have new exposure column
+  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
+
+# ex10000 runs showed effect of selection    
+#   load(file=paste0(outFolder,'listOutMix_ex10000.rda'))
+#   load(file=paste0(outFolder,'listOutI1_ex10000.rda'))
+#   load(file=paste0(outFolder,'listOutI2_ex10000.rda'))
 
@@
 ```{r,echo=FALSE,results='hide'}
-### chunk copied from sensiAnPaper1All
-### now want to calculate the times to reach critical points and add them on to the input file 
+### calculate times to reach critical points and add them on to the input file 
 ### for different insecticide strategies, sequential, mix1 and mix2
@@
   resistPointsI2 <- findResistancePoints(listOutI2, locus=2)  
-
   resistPointsSeq <- resistPointsI1 + resistPointsI2
@@
 
-  #subsetting now done above before names prettified
-  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
-  #treeInput <- treeInput[ treePredictors, ]
-
   #get the inputs here to use in ggplot investigation of model responses to inputs
@@
 
-## This plot shows exposure which has the most noticeable pattern.
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE} 
+# getting data into format for ggplot
 
@@
     x$strategy <- value
+
+    #inputs <- as.numeric(inputs)
     #transpose inputs
-    inputs <- as.data.frame( t(inputs) )
+    inputs <- as.data.frame( t(inputs), stringsAsFactors=FALSE )
     #cbind onto the outputs
@@
 
-  #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
-
-  #cool working!
-
-  #exposure has the most noticeable effect
-  ggplot(ggInsOuts, aes(x=exposure, y=time_to_resistance0.25, color=strategy)) + 
-    geom_point(shape=3) + 
-    facet_wrap( ~ strategy) + 
-    #geom_smooth(aes(x=exposure, y=gen_cP0.5, color=NULL))
-    geom_smooth(colour='red', linetype='dashed',size=0.5)
-
-  #effectiveness of insecticide1 little noticeable effect
-#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=time_to_resistance0.25, color=strategy)) + 
-#     geom_point() + 
-#     facet_wrap( ~ strategy) + 
-#     #geom_smooth(aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=NULL)) 
-#     geom_smooth(colour='red', linetype='dashed',size=0.5)
-
-```
-
-## In the following plots all inputs are plotted individually, the points are coloured by exposure, which allows us to see that, even as other inputs change, low times to resistance are associated with high exposure.
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
   #doing; for all inputs
@@
 
+  #format data in a different way to enable PRCC on difference between sequential & mix2
+  #resistPointsMix2_T
+  #resistPointsSeq_T
+
+  #columns to remove from first
+  indices1 <- which(names(resistPointsMix2_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
+  #columns to add to 2nd  
+  indices2 <- which(names(resistPointsMix2_T) %in% c("gen_cP0.25"))  
+  #rename column in 2nd
+  tmp <- resistPointsSeq_T[indices2]
+  names(tmp) <- "gen_cP0.25seq"
+
+  df_diff <- cbind(resistPointsMix2_T[-indices1], tmp)
+
+  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
+  #>1000 excludes sequential strategy that had a 999 in
+  #BEWARE would need to increase 1000 if I increase max generations in the runs
+  didntReachThresh <- which( df_diff$gen_cP0.25 > 500 | df_diff$gen_cP0.25seq > 500 )
+
+  #subset by runs not making threshold
+  df_diff <- df_diff[-didntReachThresh,]
+
+  #df_diff$mix2_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
+  df_diff["mix2_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
+
+  #TODO : later do PRCC on df_diff["mix2_minus_seq0.25"]
+
+```
+
+# Time-to-resistance across all input values
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE}
+
+#names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
+names_results <- c("time_to_resistance0.25")
+
+ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 2"), ]
+
+for(i in names_results)
+{
+  #x11()
+
+  print( ggplot(ggSubset, aes_string(x='strategy',y=i, color='strategy')) + 
+      #ylim(0,450) +
+      coord_cartesian( ylim=c(0, 350)) +
+      geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
+      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="2nd in mix\nswitch to sole", "Mixture 3"="2nd in mix" ) )
+
+  ) #end print
+}
+
+```
+
+# Difference between time-to-resistance for sequential and mixture2
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE}
+
+#names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
+# names_results <- c("mix2_minus_seq0.25")
+#   
+# ggSubset <- df_diff #[ !ggInsOuts$strategy %in% c("insecticide 2"), ]  
+#   
+# for(i in names_results)
+# {
+#   #x11()
+#   
+#   print( ggplot(ggSubset, aes_string(y=i)) + 
+#       #ylim(0,450) +
+#       #coord_cartesian( ylim=c(0, 350)) +
+#       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 )
+# 
+#   ) #end print
+# }
+
+```
+
+## How time-to-resistance is effected by input values for each insecticide use strategy
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
   names_inputs <- names(ggInsOuts)[1:num_inputs]
+
+  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
 
@@
     y <- 'time_to_resistance0.25'
-    #y <- 'time_to_resistance0.5'
-    #color <- 'strategy'
-    #coloring by exposure shows its effect even when other params are varying
-    color <- 'exposure'
 
-    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-             geom_point(shape=3) + 
-             facet_wrap( ~ strategy) + 
+    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
+    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
+            #points not wanted if 10000
+             #geom_point(shape=3, show.legend=FALSE) + 
+             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             geom_smooth(linetype='dashed',size=1) +
+             #facet_wrap( ~ strategy) +
              #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             geom_smooth(colour='red', linetype='dashed',size=0.5) +
              labs(title = i))
@@
 
-  #can I facet by particular parameter ranges ?
-  #might need to create a categorical var from the range and use that to facet 
-  #or subset the data for specific parameter ranges
+```
+
+## How difference in time-to-resistance for sequential and mixture3 is effected by input values
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
+  ggSubset <- df_diff
+
+  names_inputs <- names(ggSubset)[1:num_inputs]
+
+  for(i in names_inputs)
+  {
+    #x11()
+
+    y <- "mix2_minus_seq0.25"
+
+    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="gen_cP0.25seq")) + #, colour="strategy")) + 
+            #points not wanted if 10000
+             geom_point(shape=3, show.legend=FALSE) + 
+             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             geom_smooth(linetype='dashed',size=1) +
+             #facet_wrap( ~ strategy) +
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             labs(title = i))
+  }
 
@@
 
-## Looking more closely at effect of the ratio between the level of resistance to the first insecticide divided by that of the second one. Restricting the x axis to lower values.
-```{r, eval=TRUE, echo=FALSE, warning=FALSE}
-
-  i <- 'resist_start_1_div_2'
-  color <- "start_freq_allele2"
-
-  print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-           geom_point(shape=3) + 
-           facet_wrap( ~ strategy) + 
-           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-           geom_smooth(colour='red', linetype='dashed',size=0.5) +
-           labs(title = i) +
-           xlim(0,100)
-         )
-
-```
-
-#trying a PRCC analysis using the package sensitivity
+# PRCC analysis for time-to-resistance subsetted by each insecticide use strategy
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
@@
 # pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
-# 
-# ## S3 method for class 'pcc'
-# plot(x, ylim = c(-1,1), ...)
 # 
@@
 
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, fig.height=3.5}
+## response of resistance thresholds to individual inputs faceted by strategy
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
 
-#trying to find apparent anomaly in PRCC analysis
-#what causes correctMixDeployProp to get a high PRCC score for I1 & 2 when its not used in those runs ??
+  names_inputs <- names(ggInsOuts)[1:num_inputs]
 
-tstI1 <- round(listOutI1$input,3)
-tstMix <- round(listOutMix$input,3)
-tstDif <- round(tstMix - tstI1,3)
+  #todo find out to take sample this doesn't work
+  #ggSample <- sample(ggInsOuts,500)
+
+  for(i in names_inputs)
+  {
+    #x11()
+
+    y <- 'time_to_resistance0.25'    
+    #y <- 'time_to_resistance0.5'
+    color <- 'strategy'
+    #coloring by exposure shows its effect even when other params are varying
+    #color <- 'exposure'
+
+    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
+             #points not wanted if 10000
+             #geom_point(shape=3, show.legend=FALSE) + 
+             geom_smooth(colour='red', linetype='dashed',size=0.5, show.legend=FALSE) +
+             facet_wrap( ~ strategy) + #, show_guide = FALSE) + 
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             labs(title = i))
+  }
 
 ```
+
+## Andys questions 
+1. In mixtures what causes resistance to the 2nd insecticide to arise more slowly when the first insecticide is removed from the mixture ?

### 2016-02-09 resistanceModelNotesAndy.Rmd
notes

@@
 
+Instead I saved the original random value of exposure and used that in PRCC.
+
+I think what was happening was that correctDeployProp influenced the AB exposure value for mixtures and that was the predictor I took for exposure. Instead I want the intended exposure value.
+
+I still don't quite understand how this effects the results in the single insecticide runs.
+
+but with my 100 test runs what I'm seeing in the PRCC is no effect of anything for insect1 & 2. and a similar pattern as before for mixture1.
+
+This may point me towards another problem that is causing this.
+I may need to exclude corectDeploy as a predictor from those runs in which it wasn't used ??
+
+** 2/2/16 **
+
+~ ran 10000 runs over night
+~ run effectOfIndividualInputs
+~ now it's fine !!! PRCC looking good.
+
+New trees from 10,000 : 
+Predictors :
+effectiveness_insecticide1 :9
+effectiveness_insecticide2 :9
+exposure :                  7
+
+dominance_allele2 :         3
+dominance_allele1 :         4
+resist_start_1_div_2 :      1
+start_freq_allele2 :        1
+start_freq_allele1 :        2
+
+Best tree to show (if indeed trees are the best) : 
+Can't find a best one ! I don't like these trees !!
+Try to find a better method for Ian. Could it be PRCC.
+
+Violin plots. look good. They are very similar for time to resistance 0.1, 0.25 or 0.5. 
+~ need a plot of proportion of runs in which mixture or sequential better. Maybe I could do a violin plot of the time to threshold, with diff strategies on x axis.
+
+~ faceting them by exposure is cool
+~ faceting by effectivity shows very little
+
+** 5/2/16 **
+
+trying to get plots working
+
+## GOOD RESULTS ! :
+
+###1 effectiveness < 0.6 sequential best. > 0.6 mix best. 
+Plot of time_to_resistance against effectiveness  with 4 strategies on the same plot is key.
+
+It shows the line for mix2 crossing that of sequential. Effectiveness 0.4-0.6 resistance arises faster for mix2. > 0.6 resistance arises faster for sequential.
+
+I should adapt this plot perhaps just to show mix2 and sequential.
+
+###2 selection coeff < 0.2 sequential best. > 0.2 mix best.
+selection coefficient curves also cross for mix2 and sequential. < ~0.2 sequential is better. > 0.2 mixture is better. 
+But check that this is multiplied as shown below. So beware could this be caused by the effectiveness result ?
+    ##effectiveness
+    phi.SS1_A0 <- runif(1, min=0.4, max=1)
+    phi.SS2_0B <- runif(1, min=0.4, max=1)
+
+    # Ian suggested this should be dependent on effectiveness to ensure fitness of RR stays below 1 
+    s.RR1_A0 <- runif(1, min=0.2, max=1) * phi.SS1_A0
+    s.RR2_0B <- runif(1, min=0.2, max=1) * phi.SS2_0B
+
+I think perhaps I should remove multiplication of selection coefficient by effectiveness ?
+
+###3
+Effect of exposure is greater overall on times to resistance, but it has a more similar effect on mixtures & sequential thus it is of less importance when considering the differences between the two strategies.
+
+In majority of cases the point where resistance to just one of the insecticides in a mixture is reached comes before that for both insecticides in sequential use. [this seems obvious]
+Also it always takes longer to reach resistance for both insecticides when starting with both in a mixture. Is this obvious, or a meaningfull result or an error in our methodology ?
+
+I think for the paper we should focus on mix2 versus sequential. We could have that as what we call mixture, and then refer in smallere analyses to the other mix scenarios.
+Maybe I should have a quick look at curtis paper to remind myself ?
+
+Tried removing multiplication of selection coefficient by effectiveness.
+But got warnings about fitnesses > 1.
+
+6/2/16
+
+problem is with locus fitness values going over 1
+
+This is the problem that causes it :
+
+Wloci[ paste0('RR',locusNum), exposure] <- (1 - phi[locusNum, exposure]) + 
+                                                   (s[locusNum, exposure])
+
+so it just occurs when s > phi, selection coeff > effectiveness
+
+current range :
+effectiveness : 0.4 - 1
+selection coeff : 0.2 - 0.5
+
+Options : 
+1) discard sims in which selection > effectiveness
+2) set selection range to 0.1-0.45 and effectiveness from 0.45-1
+
+Doing 2) might be good to check whether the line crossing still goes on ...
+
+Yes saving 2) as listOutMix_ex2_1000.rda etc.
+
+ Done in effectOfIndividualInputs
+
+** 7/2/16 **
+with new runs
+
+Selection coefficient curves for mix2 & sequential no longer cross. Seems likely that this previous pattern was generated by correlation with effectiveness.
+
 ### TO DO
 
-~ rerun the trees for Ian
-
-~ set 10000 runs going before yoga, but whether extended or normal curtis ?
-
-~ I might be able to do a PRCC of whether mixtures better than sequential (but actually thats what the trees do)
+~ run PRCC for difference between mix2 & sequential
+~ might need to structure data differently
+
+Tighten up the story before sending to Ian, I'm getting there. Maybe put into a summary Rmd pdf.
+
+## IDEA
+# rather than trees of e.g. mix1 better than sequence I could show time-to-resistance mix1-sequence.
+
+OR I could even put time_to_resistance on the same plot for each stratgey (maybe with exposure on the x-axis)
+This is very similar to some plots I have already, I just need to remove the facetting by strategy and remove individual strategies.
+
+## BEWARE what am I doing with putting mixture better than seq on each repeated input ....  
+## e.g. it doesn't seem to quite make sense that I have mixture or sequence better for insecticide1
+## but maybe it is because all the inpiuts are repeated
+
+## might be good to facet by mixture sequence better and just have points of x=exposure, y=effectiveness_insecticide1
+
+~ do PRCCs on whether mixture or sequential better
+
+~ tomorrow when on good form in morning work out how best to summarise for Ian
+
+~ add other resist thresholds into PRCC.
+
+~ can I get plots to show just a subset of points ?
+
+~ produce a graphich of the 3 diff mixture scenarios & sequential
+~~ may be able to use the seriousGame plot method
+~~ I could even show resistance curves next to it ?
+
+### My summary of results & thoughts
+
+Resistance to two insecticides arises more slowly if they are used in combination rather than sequentially. See violin plots.
 

### 2016-02-11 paper1_results_figs.Rmd
add PRCC for difference

@@
 # Figures for paper 1 
-# version 2 - in progress
+# version 3 - in progress
 
@@
 The latter takes slightly longer to reach the threshold.
-1. Exposure has the greatest effect on time-to-resistance, but as it has a similar effect across all strategies it has little influence on which strategy is preferable.
+1. Exposure and effectiveness seem to be the most important inputs.
+1. Exposure has the greatest effect on time-to-resistance.
 1.
@@
 
-# Difference between time-to-resistance for sequential and mixture2
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE}
-
-#names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
-# names_results <- c("mix2_minus_seq0.25")
-#   
-# ggSubset <- df_diff #[ !ggInsOuts$strategy %in% c("insecticide 2"), ]  
-#   
-# for(i in names_results)
-# {
-#   #x11()
-#   
-#   print( ggplot(ggSubset, aes_string(y=i)) + 
-#       #ylim(0,450) +
-#       #coord_cartesian( ylim=c(0, 350)) +
-#       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 )
-# 
-#   ) #end print
-# }
-
-```
-
 ## How time-to-resistance is effected by input values for each insecticide use strategy
@@
 
-## How difference in time-to-resistance for sequential and mixture3 is effected by input values
+## How difference in time-to-resistance between sequential and mixture2 is effected by input values
 
@@
 
-    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="gen_cP0.25seq")) + #, colour="strategy")) + 
+    print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
             #points not wanted if 10000
-             geom_point(shape=3, show.legend=FALSE) + 
+             geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
              #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             geom_smooth(linetype='dashed',size=1) +
+             geom_smooth(linetype='dashed',size=1, colour="red") +
              #facet_wrap( ~ strategy) +
              #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i))
+             labs(title = i) +
+             geom_hline(yintercept = 0, colour="white")
+           )
   }
+
+```
+
+# PRCC difference in time-to-resistance between sequential and mixture2
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
+
+  library(sensitivity)
+
+    x <- df_diff[,1:num_inputs]
+    y <- df_diff["mix2_minus_seq0.25"]
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    #results are here I can probably rbind them together into a df that I can ggplot
+    #pcc_res$PRCC
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
+
+    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             geom_hline(yintercept = 0, linetype=3) +
+             ylim(-1,1) +
+             ggtitle(paste("PRCC difference between sequential and mixture2")) +
+             xlab(NULL)
+          )
 

### 2016-02-11 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 9/2/16 **
+sent new figs of PRCC of diff between seq & mix2 to Ian
+
+** 10/2/16 **
+
+~ reran trees they show that effectiveness > ~ 0.7 & exposure < 0.7 favours mixture2 over sequential
+~ this is consistent with my graphs of difference vs effect & exposure (therefore are the trees really necessary ?)
+
+mix2 can already be displayed in plotcurtis_f2_generic() using addCombinedStrategy
+I had just set default to F because the plot gets even more confusing.
+
+Maybe add option to just plot mix2 ?
+
+Also offer option to plot from : runcurtis_f2()
+
+This which simply reduces effectiveness of insecticide2 from 1 to 0.85, makes seq better than mix (where mix was better for curtis).
+
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
 ### TO DO
 
-~ run PRCC for difference between mix2 & sequential
-~ might need to structure data differently
+~ add default curtis values to UI labels in server.r
+~ check on plotting of combined strategy
+~
+
+~ play with shiny app to help my understanding of how modifying the params alters results
+
+~ check code in findResistancePointsMixResponsive()
 
@@
 
-1. I could add the combined mixture/individual (strategy 2) as a dotted line to plotcurtis_f2_generic()
-(and then could even add labels of 1,2,3 to where they cross the 50% line)
-
 1. fix the testthat tests to avoid new input addition

### 2016-02-11 sensiAnPaper1All.Rmd
changing scenarios

@@
 
-  #nScenarios <- 10  
+  nScenarios <- 3  
   #nScenarios <- 100
-  nScenarios <- 1000
+  #nScenarios <- 1000  
+  #nScenarios <- 10000
 
@@
   ## save results objects as rda for re-use
-  save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
-  save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
-  save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))
+  #save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
+  #save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
+  #save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))
 
-  save(listOutMix, file=paste0(outFolder,'listOutMix_ex1000.rda'))
-  save(listOutI1, file=paste0(outFolder,'listOutI1_ex1000.rda'))
-  save(listOutI2, file=paste0(outFolder,'listOutI2_ex1000.rda'))
+  save(listOutMix, file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
+
+#   save(listOutMix, file=paste0(outFolder,'listOutMix_ex2_3.rda'))
+#   save(listOutI1, file=paste0(outFolder,'listOutI1_ex2_3.rda'))
+#   save(listOutI2, file=paste0(outFolder,'listOutI2_ex2_3.rda'))
 
@@
 
-```{r}
+```{r, eval=FALSE}
   ## to load previously saved runs
@@
 
-  load(file=paste0(outFolder,'listOutMix_ex1000.rda'))
-  load(file=paste0(outFolder,'listOutI1_ex1000.rda'))
-  load(file=paste0(outFolder,'listOutI2_ex1000.rda'))
+  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
 
@@
 
-```
-
-### do tree analysis
-```{r, eval=TRUE}
-  require(rpart.plot)
-
   #transpose
@@
 
+```
+
+### do tree analysis
+```{r, eval=TRUE}
+
+  require(rpart.plot)
+
   #convert predictor columns back to numeric
   #following this above : replace 0s & 1s with mixture/sequence to make plot clearer
-  filter <- !substr(names(treeInput),1,6) %in% 'better'
-  treeInput[,filter] <- lapply(treeInput[,filter],as.numeric) 
-  filter <- !filter
-  treeInput[,filter] <- lapply(treeInput[,filter],as.factor)
+  results_columns <- substr(names(treeInput),1,6) %in% 'better'
+  input_columns <- !results_columns
+  treeInput[,input_columns] <- lapply(treeInput[,input_columns],as.numeric) 
+  treeInput[,results_columns] <- lapply(treeInput[,results_columns],as.factor)
 
   #create string with predictor names in
-  treePredictorString <- paste(treePredictors, collapse="+")
+  #treePredictorString <- paste(treePredictors, collapse="+")
+  treePredictorString <- paste(colnames(treeInput)[input_columns], collapse="+")
 
@@
 ```
-
-### investigating response of time-to-resistance to individual input parameters
-### (while they all are varying so it may not work well)
-```{r, eval=TRUE}
-
-  #uses ggInput calculated in earlier chunk
-
-  #first needs to transpose rows to cols
-  #ggInput_T <- t(ggInput)
-
-  #function to transpose and add strategy column with the passed value
-  #also add on the input columns  
-  addStrategyColumn <- function(x, value, inputs){
-    x <- as.data.frame( t(x) )
-    x$strategy <- value
-    #transpose inputs
-    inputs <- as.data.frame( t(inputs) )
-    #cbind onto the outputs
-    x <- cbind(inputs,x)
-    x
-  }
-
-  resistPointsI1_T <- addStrategyColumn(resistPointsI1,"insecticide 1",ggInput) 
-  resistPointsI2_T <- addStrategyColumn(resistPointsI2,"insecticide 2",ggInput)   
-  resistPointsSeq_T <- addStrategyColumn(resistPointsSeq,"Sequential",ggInput) 
-  resistPointsMix1_T <- addStrategyColumn(resistPointsMix1,"Mixture 1",ggInput) 
-  resistPointsMix2_T <- addStrategyColumn(resistPointsMix2,"Mixture 2",ggInput) 
-  resistPointsMix3_T <- addStrategyColumn(resistPointsMix3,"Mixture 3",ggInput)
-
-  ggInsOuts <- rbind( resistPointsI1_T, resistPointsI2_T, resistPointsSeq_T, 
-                      resistPointsMix1_T, resistPointsMix2_T, resistPointsMix3_T)
-
-  #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
-
-  #cool working!
-
-  #this one shows no discernible pattern
-  ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=strategy)) + geom_point() + facet_wrap( ~ strategy) + geom_smooth()
-
-  #doing; for all inputs
-  names_inputs <- names(ggInsOuts)[1:10]
-
-  for(i in names_inputs)
-  {
-    x11()
-
-    y <- 'gen_cP0.5'
-    #color <- 'strategy'
-    #coloring by exposure shows its effect even when other params are varying
-    color <- 'exposure'
-
-    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-             geom_point() + 
-             facet_wrap( ~ strategy) + 
-             geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-  }
-
-  #can I facet by particular parameter ranges ?
-  #or even subset the data for specific parameter ranges
-
-```

### 2016-02-11 resistanceModelNotesAndy.Rmd
notes

@@
 
-** 9/1/16 ** on train back from liverpool game rollout workshop
-
-modifying setExposure() function to cope with Ians request for lower male exposure
-I can add it in as a param with a default of 1.
-
-** 13/1/16 **
-starting on the new 1 parameter at a time sensi analysis in sensiAnPaperPartOneAtATime.r
-
-** 14/1/16 **
-first see if I can make any informative plots with the existing sensitivity analysis data
-
-These outputs are the frequencies by generation of the resistance alleles for the different insecticides (1&2) 
-Gen, m.R1, m.R2, f.R1, f.R2
-
-I want to get to a single file that has inputs & time_to_resistance thresholds for each scenario.
-
-I must have something nearly like this already from sensiAnPaper1All.Rmd
-
-Yes : resistPointsI1,2,Mix etc.
-For each scenario these have gen_cP0.1,0.25,0.5 which is the number of generations taken to reach those resistance thresholds.
-
-The next chunk works out which strategy is better than which.
-
-Maybe I want to bind all the resistPoints together into one file and bind onto the inputs, so that we can look at how the inputs effect the time to resistance for the different strategies.
-
-I want input on the x axis & time to resistance on the y.
-(but there will be the issue that the y values will be affected by the changing values of other parameters too) Cross that bridge when I come to it.
-
-how to get data in format for ggplot2
-check the diamonds dataset in ggplot2
-
-so I want to get as a dataframe with scenarios in rows, & columns for inputs & outputs
-
-so I can do something like :
-ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point()
-
-If I have a column that is strategy (insecticide1, insectide2, sequential, mixture)
-Then I would be able to facet by these later.
-
-ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
-
-This could be good as I can see the different strategies side by side.
-
-So I want columns :
-input1, input2 ... strategy, outputs : gen_cP0.1, 0.25, 0.5
-
-so the dataframe will have length=num strategies*num scenarios
-
-I'll need to rbind the strategies on top of each other.
-
-Some success with ggplot plots.
-
-** 15/1/16 **
-sent first version of ggplot doc (effectOfIndividualInputs.pdf) to Ian & Beth
-
-prettifying parameter names
-
-** 20/1/16 ** on train to liverpool
-
-removed from plots runs where resistance threshold not reached
-
-** 22/1/16 ** with Ian in Liverpool
-I said realistically it would be a month before I could get started on the
-
-conventional wisdom of poor spraying increasing resistance is because poor spray thought not to kill heterozygotes.
-we see some o this effect in the dominannce param.
-Later we can loook at by the low exposure niches e.g.
-a vs A.
-We can look at in the next paper.
-
-Ian plans to submit to PLOS computationally biology with the aim that it's a flexible model that can address a bunch of situations.
-Do we want to be slightly carful to protect our IP ?
-e.g.
-in the first version w might not want to promote use by others (we'd need to do some more wrk to make it properly useable).
-I can discuss this with Ian when we are closer to publication.
-
-~ I did think that correctDeploymentProp only effects mixtureswhich would constrain tree analyses for whether mix better than seq, but actually analysis can still be done, you would just expect mix to be closer to seq.
-
-~ added maleExposureProp & correctMixDeployProp into tree & plot analyses, by adding on to the end of the input object.
-inputs are got from   treeInput <- listOutMix$input which is created by runModel2() just from the input object that is passed to it.
-
-~ added as an extra argument to sensiAnPaperPart() that would make less code repetition
-
-** 25/1/16 **
-added to effectOfIndividualInputs.Rmd
-~~ new predictors for males & deploy
-~~ calc of new ratio, copy from sensiAnPaperAll
-
-13.40 started 1000 reps of the 'extended' experiment 
-< 50 mins to run
-
-This is what Susanan did for her 2012 model :
-
-Simulations to understand the influence of each parameter
-on the outcome variables (mean fitness and change
-in resistance allele frequency) were performed using latin
-hypercube sampling (LHS) to generate a data set and
-partial rank correlation coefficients (PRCC) calculated to
-provide a quantitative measure of the impact of each
-parameter [24]. LHS techniques were first developed to
-explore the behavior of complex models in economics,
-engineering, chemistry and physics and have been used in
-models predicting the impact of insecticide-treated nets
-on malaria transmission [25].
-
-The analysis was performed using R software [26] and
-implementation of LHS using package lhs. It does not
-allow for the specification of each variable distribution
-beforehand, so sampling was performed assuming a uniform
-distribution. Once the sample was generated, the
-uniform sample from a column (variable) could be transformed
-to the required distribution (Table 3) by using
-quantile functions (using the qtriangle comand in R).
-
-A data set of 3,000 replications was generated, with random
-parameter sets and the corresponding values of the
-outcome variables using equations 8, 9, 12 and 13. Ten
-replicates of this procedure were performed as suggested
-in [24] to investigate the predictive precision of model
-using LHS as the sampling method. This was achieved
-by analysing each replicate separately and verifying that
-results were consistent across ten replicates.
-
-** 26/1/16 **
-looking into prcc analysis of model outputs &  whether I might want tio use lhs.
-
-~ what is difference between lhs and what I have done
-~ check lhs package 
-~ lhs would ensure a more uniform distributions of samples across parameter space, i would think results are unlikely to be different to our monte-carlo approach.
-~ it might be tricky for the params where we haven't used a uniform distribution (although we probably could get around)
-~ some advice on using lhs package
-http://r.789695.n4.nabble.com/latin-hypercube-sampling-td4659028.html
-
-~ add PRCC code
-This provides some (dated) advice on running a sensitivity anlaysis including LHS & prcc
-https://cran.r-project.org/web/packages/pse/vignettes/pse_tutorial.pdf
-
-pcc function in package sensitivity may do what we want :
-https://cran.r-project.org/web/packages/sensitivity/sensitivity.pdf
-
-pcc computes the Partial Correlation Coefficients (PCC), or Partial Rank Correlation Coefficients
-(PRCC), which are sensitivity indices based on linear (resp. monotonic) assumptions, in the case of
-(linearly) correlated factors.
-
-~ or could use epiR package
-epi.prcc(dat) # where final column in dat is the output, all other columns are inputs
-
-I got sensitivity::pcc() working for a single strategy.
-
-** 28/1/16 **
-~ fixed diagonal axis labels.
-~ get sensitivity::pcc() working for all strategies & putting data into format for faceted plots.
-
-** 29/1/16 **
-
-~ why does correctMixDeployProp seem to have an effect in the PRCC of insecticide1 when it shouldn't have any effect outside of mixtures ?
-
-In the plots againts time_to_resistance there is only a bit of a pattern for Mixture1 (time_to_R increaeses between 0.9 & 1 correct deployment).
-Points to it being a problem in the PRCC itself.
-
-~~ in setExposure() looks fine, its not applied to insecticide1 or 2
-
-other things it could be :
-~~ switched with another var in the call to setExposure
-~~ something about how runs are done in sensiAnPaperPart.r 
-~~ rows getting reordered within effectOfIndividualInputs.Rmd ?
-
-it is +vely correlated with time to resistance, so better application leads to longer times for resistance to be reached (not sure whether that is waht I would expect ?)
-
-aha PRCC results are influenced by the presence of the other vars, if just do for exposure & deployProp deployProp goes down to 0.2. 
-x2 <- x[c('correctMixDeployProp','exposure')]
-pcc_res2 <- pcc(x2, y, rank=TRUE)
-pcc_res2$PRCC
-                       original
-correctMixDeployProp  0.2167706
-exposure             -0.5248086
-
-I still don't quite understand why an essentially random var is correlated with the output ??
-
-This shows that there's very little correlation between the in & out, for insecticide1
-> cor(xy$time_to_resistance0.25,xy$correctMixDeployProp)
-[1] 0.003849606
-
-** 1/2/16 **
-
-Checking again on whats happening with correctMixDeployProp
-
-is the correlation similar for mixture1 as it is for insecticide1 ?
-
-could problem be that the inputs are got from a mix run ? shouldn't matter
-  treeInput <- listOutMix$input
-
-could it be something to do with the random number seed being disrupted by the extra params ? shouldn't happen
-
-i think the inputs should be identical.
-
-I think it may be a correlation in the analysis.
-
-What I have done is generate inputs for 'male proportion' and 'correct mix deploy'.
-Correct mix deployment is not used in the individual insecticide runs, but it is still included as a potential predictor in the PRCC analysis.
-
-In the mixture scenarios 'correct mix deploy' effects exposure ...
-
-exposure is not a single variable in the simulation, but is it used as a single value as a predictor in the sensitivity & prcc analyses ?
-
-aha the exposure predictor is taken from the female AB (actually I think that is OK ??)
-although maybe not because it will be reduced by correct deploy and maybe by male (or maybe not ???)
-
-  rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
-
-correctDeploy has a strong interaction with exposure in mixtures.
-
-When correct deploy is low in mixture runs exposure to AB is much lower because of corresponding exposure to A0 and 0B.
-
-But i think this shouldn't effect single insecticide runs.
-
-Am I sure I'm not mixing up rows somewhere in effectOfIndividualInputs.Rmd ??
-
-I could try setting the values of results for I1 and see if it propogates through.
-
-Instead I saved the original random value of exposure and used that in PRCC.
-
-I think what was happening was that correctDeployProp influenced the AB exposure value for mixtures and that was the predictor I took for exposure. Instead I want the intended exposure value.
-
-I still don't quite understand how this effects the results in the single insecticide runs.
-
-but with my 100 test runs what I'm seeing in the PRCC is no effect of anything for insect1 & 2. and a similar pattern as before for mixture1.
-
-This may point me towards another problem that is causing this.
-I may need to exclude corectDeploy as a predictor from those runs in which it wasn't used ??
-
-** 2/2/16 **
-
-~ ran 10000 runs over night
-~ run effectOfIndividualInputs
-~ now it's fine !!! PRCC looking good.
-
-New trees from 10,000 : 
-Predictors :
-effectiveness_insecticide1 :9
-effectiveness_insecticide2 :9
-exposure :                  7
-
-dominance_allele2 :         3
-dominance_allele1 :         4
-resist_start_1_div_2 :      1
-start_freq_allele2 :        1
-start_freq_allele1 :        2
-
-Best tree to show (if indeed trees are the best) : 
-Can't find a best one ! I don't like these trees !!
-Try to find a better method for Ian. Could it be PRCC.
-
-Violin plots. look good. They are very similar for time to resistance 0.1, 0.25 or 0.5. 
-~ need a plot of proportion of runs in which mixture or sequential better. Maybe I could do a violin plot of the time to threshold, with diff strategies on x axis.
-
-~ faceting them by exposure is cool
-~ faceting by effectivity shows very little
-
-** 5/2/16 **
-
-trying to get plots working
-
-## GOOD RESULTS ! :
-
-###1 effectiveness < 0.6 sequential best. > 0.6 mix best. 
-Plot of time_to_resistance against effectiveness  with 4 strategies on the same plot is key.
-
-It shows the line for mix2 crossing that of sequential. Effectiveness 0.4-0.6 resistance arises faster for mix2. > 0.6 resistance arises faster for sequential.
-
-I should adapt this plot perhaps just to show mix2 and sequential.
-
-###2 selection coeff < 0.2 sequential best. > 0.2 mix best.
-selection coefficient curves also cross for mix2 and sequential. < ~0.2 sequential is better. > 0.2 mixture is better. 
-But check that this is multiplied as shown below. So beware could this be caused by the effectiveness result ?
-    ##effectiveness
-    phi.SS1_A0 <- runif(1, min=0.4, max=1)
-    phi.SS2_0B <- runif(1, min=0.4, max=1)
-
-    # Ian suggested this should be dependent on effectiveness to ensure fitness of RR stays below 1 
-    s.RR1_A0 <- runif(1, min=0.2, max=1) * phi.SS1_A0
-    s.RR2_0B <- runif(1, min=0.2, max=1) * phi.SS2_0B
-
-I think perhaps I should remove multiplication of selection coefficient by effectiveness ?
-
-###3
-Effect of exposure is greater overall on times to resistance, but it has a more similar effect on mixtures & sequential thus it is of less importance when considering the differences between the two strategies.
-
-In majority of cases the point where resistance to just one of the insecticides in a mixture is reached comes before that for both insecticides in sequential use. [this seems obvious]
-Also it always takes longer to reach resistance for both insecticides when starting with both in a mixture. Is this obvious, or a meaningfull result or an error in our methodology ?
-
-I think for the paper we should focus on mix2 versus sequential. We could have that as what we call mixture, and then refer in smallere analyses to the other mix scenarios.
-Maybe I should have a quick look at curtis paper to remind myself ?
-
-Tried removing multiplication of selection coefficient by effectiveness.
-But got warnings about fitnesses > 1.
-
-6/2/16
-
-problem is with locus fitness values going over 1
-
-This is the problem that causes it :
-
-Wloci[ paste0('RR',locusNum), exposure] <- (1 - phi[locusNum, exposure]) + 
-                                                   (s[locusNum, exposure])
-
-so it just occurs when s > phi, selection coeff > effectiveness
-
-current range :
-effectiveness : 0.4 - 1
-selection coeff : 0.2 - 0.5
-
-Options : 
-1) discard sims in which selection > effectiveness
-2) set selection range to 0.1-0.45 and effectiveness from 0.45-1
-
-Doing 2) might be good to check whether the line crossing still goes on ...
-
-Yes saving 2) as listOutMix_ex2_1000.rda etc.
-
- Done in effectOfIndividualInputs
-
-** 7/2/16 **
-with new runs
-
-Selection coefficient curves for mix2 & sequential no longer cross. Seems likely that this previous pattern was generated by correlation with effectiveness.
-
-** 9/2/16 **
-sent new figs of PRCC of diff between seq & mix2 to Ian
-
-** 10/2/16 **
-
-~ reran trees they show that effectiveness > ~ 0.7 & exposure < 0.7 favours mixture2 over sequential
-~ this is consistent with my graphs of difference vs effect & exposure (therefore are the trees really necessary ?)
-
-mix2 can already be displayed in plotcurtis_f2_generic() using addCombinedStrategy
-I had just set default to F because the plot gets even more confusing.
-
-Maybe add option to just plot mix2 ?
-
-Also offer option to plot from : runcurtis_f2()
-
-This which simply reduces effectiveness of insecticide2 from 1 to 0.85, makes seq better than mix (where mix was better for curtis).
-
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-
-### TO DO
-
-~ add default curtis values to UI labels in server.r
-~ check on plotting of combined strategy
-~
-
-~ play with shiny app to help my understanding of how modifying the params alters results
-
-~ check code in findResistancePointsMixResponsive()
-
-Tighten up the story before sending to Ian, I'm getting there. Maybe put into a summary Rmd pdf.
-
-## IDEA
-# rather than trees of e.g. mix1 better than sequence I could show time-to-resistance mix1-sequence.
-
-OR I could even put time_to_resistance on the same plot for each stratgey (maybe with exposure on the x-axis)
-This is very similar to some plots I have already, I just need to remove the facetting by strategy and remove individual strategies.
-
-## BEWARE what am I doing with putting mixture better than seq on each repeated input ....  
-## e.g. it doesn't seem to quite make sense that I have mixture or sequence better for insecticide1
-## but maybe it is because all the inpiuts are repeated
-
-## might be good to facet by mixture sequence better and just have points of x=exposure, y=effectiveness_insecticide1
-
-~ do PRCCs on whether mixture or sequential better
-
-~ tomorrow when on good form in morning work out how best to summarise for Ian
-
-~ add other resist thresholds into PRCC.
-
-~ can I get plots to show just a subset of points ?
-
-~ produce a graphich of the 3 diff mixture scenarios & sequential
-~~ may be able to use the seriousGame plot method
-~~ I could even show resistance curves next to it ?
-
-### My summary of results & thoughts
-
-Resistance to two insecticides arises more slowly if they are used in combination rather than sequentially. See violin plots.
-
-~ I'd like to show prcc on an axis from 0-1 with diff sysmbols for neg & pos. We care about which are most imp. params, not whether they are pos or neg.
-
-~ If I would like to get prcc plots in a similar format to my existing plots. (but may not be space for axis labels)
-~ get data in form : 
-input1, input2, inputN, strategy 
-prcc vals,,             indecticide1
-
-# would it be useful for Ian for me to create an interactive shiny app to view the 1000 sim data. E.g. could just have the ggplot code with options to set the x value & colour by different inputs.
-
-# maybe add in a loop to plot for all 3 resistance thresholds
-# maybe try faceting by parameter ranges ?
-# lookup to see if I can steal any existing sensi analysis type code
-
-See sensiAnPaperPartOneAtATime.r
-  #TODO I need to think more carefully about how to do this
-  #and look back at my existing notes
-1. Graphs where just change one param at a time. Jo Lines/Andy conversation at TAG.
-
-Extended Curtis requests from Ian (still not quite sure how I will do)
+Extended Curtis requests from Ian
 1. 'Male exposure' as a proportion of female (0.5 to 1)
@@
 
-1. look at investigating response by changing single parameters at a time (or can I use existing runs)
-results files are in 'C:\\Users\\andy\\Documents\\resistanceResultsBackup\\listOutI1_1000.rda'
-listOut2, & listOutMix
+** 9/1/16 ** on train back from liverpool game rollout workshop
+
+modifying setExposure() function to cope with Ians request for lower male exposure
+I can add it in as a param with a default of 1.
+
+** 13/1/16 **
+starting on the new 1 parameter at a time sensi analysis in sensiAnPaperPartOneAtATime.r
+
+** 14/1/16 **
+first see if I can make any informative plots with the existing sensitivity analysis data
+
+These outputs are the frequencies by generation of the resistance alleles for the different insecticides (1&2) 
+Gen, m.R1, m.R2, f.R1, f.R2
+
+I want to get to a single file that has inputs & time_to_resistance thresholds for each scenario.
+
+I must have something nearly like this already from sensiAnPaper1All.Rmd
+
+Yes : resistPointsI1,2,Mix etc.
+For each scenario these have gen_cP0.1,0.25,0.5 which is the number of generations taken to reach those resistance thresholds.
+
+The next chunk works out which strategy is better than which.
+
+Maybe I want to bind all the resistPoints together into one file and bind onto the inputs, so that we can look at how the inputs effect the time to resistance for the different strategies.
+
+I want input on the x axis & time to resistance on the y.
+(but there will be the issue that the y values will be affected by the changing values of other parameters too) Cross that bridge when I come to it.
+
+how to get data in format for ggplot2
+check the diamonds dataset in ggplot2
+
+so I want to get as a dataframe with scenarios in rows, & columns for inputs & outputs
+
+so I can do something like :
+ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point()
+
+If I have a column that is strategy (insecticide1, insectide2, sequential, mixture)
+Then I would be able to facet by these later.
+
+ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
+
+This could be good as I can see the different strategies side by side.
+
+So I want columns :
+input1, input2 ... strategy, outputs : gen_cP0.1, 0.25, 0.5
+
+so the dataframe will have length=num strategies*num scenarios
+
+I'll need to rbind the strategies on top of each other.
+
+Some success with ggplot plots.
+
+** 15/1/16 **
+sent first version of ggplot doc (effectOfIndividualInputs.pdf) to Ian & Beth
+
+prettifying parameter names
+
+** 20/1/16 ** on train to liverpool
+
+removed from plots runs where resistance threshold not reached
+
+** 22/1/16 ** with Ian in Liverpool
+I said realistically it would be a month before I could get started on the
+
+conventional wisdom of poor spraying increasing resistance is because poor spray thought not to kill heterozygotes.
+we see some o this effect in the dominannce param.
+Later we can loook at by the low exposure niches e.g.
+a vs A.
+We can look at in the next paper.
+
+Ian plans to submit to PLOS computationally biology with the aim that it's a flexible model that can address a bunch of situations.
+Do we want to be slightly carful to protect our IP ?
+e.g.
+in the first version w might not want to promote use by others (we'd need to do some more wrk to make it properly useable).
+I can discuss this with Ian when we are closer to publication.
+
+~ I did think that correctDeploymentProp only effects mixtureswhich would constrain tree analyses for whether mix better than seq, but actually analysis can still be done, you would just expect mix to be closer to seq.
+
+~ added maleExposureProp & correctMixDeployProp into tree & plot analyses, by adding on to the end of the input object.
+inputs are got from   treeInput <- listOutMix$input which is created by runModel2() just from the input object that is passed to it.
+
+~ added as an extra argument to sensiAnPaperPart() that would make less code repetition
+
+** 25/1/16 **
+added to effectOfIndividualInputs.Rmd
+~~ new predictors for males & deploy
+~~ calc of new ratio, copy from sensiAnPaperAll
+
+13.40 started 1000 reps of the 'extended' experiment 
+< 50 mins to run
+
+This is what Susanan did for her 2012 model :
+
+Simulations to understand the influence of each parameter
+on the outcome variables (mean fitness and change
+in resistance allele frequency) were performed using latin
+hypercube sampling (LHS) to generate a data set and
+partial rank correlation coefficients (PRCC) calculated to
+provide a quantitative measure of the impact of each
+parameter [24]. LHS techniques were first developed to
+explore the behavior of complex models in economics,
+engineering, chemistry and physics and have been used in
+models predicting the impact of insecticide-treated nets
+on malaria transmission [25].
+
+The analysis was performed using R software [26] and
+implementation of LHS using package lhs. It does not
+allow for the specification of each variable distribution
+beforehand, so sampling was performed assuming a uniform
+distribution. Once the sample was generated, the
+uniform sample from a column (variable) could be transformed
+to the required distribution (Table 3) by using
+quantile functions (using the qtriangle comand in R).
+
+A data set of 3,000 replications was generated, with random
+parameter sets and the corresponding values of the
+outcome variables using equations 8, 9, 12 and 13. Ten
+replicates of this procedure were performed as suggested
+in [24] to investigate the predictive precision of model
+using LHS as the sampling method. This was achieved
+by analysing each replicate separately and verifying that
+results were consistent across ten replicates.
+
+** 26/1/16 **
+looking into prcc analysis of model outputs &  whether I might want tio use lhs.
+
+~ what is difference between lhs and what I have done
+~ check lhs package 
+~ lhs would ensure a more uniform distributions of samples across parameter space, i would think results are unlikely to be different to our monte-carlo approach.
+~ it might be tricky for the params where we haven't used a uniform distribution (although we probably could get around)
+~ some advice on using lhs package
+http://r.789695.n4.nabble.com/latin-hypercube-sampling-td4659028.html
+
+~ add PRCC code
+This provides some (dated) advice on running a sensitivity anlaysis including LHS & prcc
+https://cran.r-project.org/web/packages/pse/vignettes/pse_tutorial.pdf
+
+pcc function in package sensitivity may do what we want :
+https://cran.r-project.org/web/packages/sensitivity/sensitivity.pdf
+
+pcc computes the Partial Correlation Coefficients (PCC), or Partial Rank Correlation Coefficients
+(PRCC), which are sensitivity indices based on linear (resp. monotonic) assumptions, in the case of
+(linearly) correlated factors.
+
+~ or could use epiR package
+epi.prcc(dat) # where final column in dat is the output, all other columns are inputs
+
+I got sensitivity::pcc() working for a single strategy.
+
+** 28/1/16 **
+~ fixed diagonal axis labels.
+~ get sensitivity::pcc() working for all strategies & putting data into format for faceted plots.
+
+** 29/1/16 **
+
+~ why does correctMixDeployProp seem to have an effect in the PRCC of insecticide1 when it shouldn't have any effect outside of mixtures ?
+
+In the plots againts time_to_resistance there is only a bit of a pattern for Mixture1 (time_to_R increaeses between 0.9 & 1 correct deployment).
+Points to it being a problem in the PRCC itself.
+
+~~ in setExposure() looks fine, its not applied to insecticide1 or 2
+
+other things it could be :
+~~ switched with another var in the call to setExposure
+~~ something about how runs are done in sensiAnPaperPart.r 
+~~ rows getting reordered within effectOfIndividualInputs.Rmd ?
+
+it is +vely correlated with time to resistance, so better application leads to longer times for resistance to be reached (not sure whether that is waht I would expect ?)
+
+aha PRCC results are influenced by the presence of the other vars, if just do for exposure & deployProp deployProp goes down to 0.2. 
+x2 <- x[c('correctMixDeployProp','exposure')]
+pcc_res2 <- pcc(x2, y, rank=TRUE)
+pcc_res2$PRCC
+                       original
+correctMixDeployProp  0.2167706
+exposure             -0.5248086
+
+I still don't quite understand why an essentially random var is correlated with the output ??
+
+This shows that there's very little correlation between the in & out, for insecticide1
+> cor(xy$time_to_resistance0.25,xy$correctMixDeployProp)
+[1] 0.003849606
+
+** 1/2/16 **
+
+Checking again on whats happening with correctMixDeployProp
+
+is the correlation similar for mixture1 as it is for insecticide1 ?
+
+could problem be that the inputs are got from a mix run ? shouldn't matter
+  treeInput <- listOutMix$input
+
+could it be something to do with the random number seed being disrupted by the extra params ? shouldn't happen
+
+i think the inputs should be identical.
+
+I think it may be a correlation in the analysis.
+
+What I have done is generate inputs for 'male proportion' and 'correct mix deploy'.
+Correct mix deployment is not used in the individual insecticide runs, but it is still included as a potential predictor in the PRCC analysis.
+
+In the mixture scenarios 'correct mix deploy' effects exposure ...
+
+exposure is not a single variable in the simulation, but is it used as a single value as a predictor in the sensitivity & prcc analyses ?
+
+aha the exposure predictor is taken from the female AB (actually I think that is OK ??)
+although maybe not because it will be reduced by correct deploy and maybe by male (or maybe not ???)
+
+  rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+correctDeploy has a strong interaction with exposure in mixtures.
+
+When correct deploy is low in mixture runs exposure to AB is much lower because of corresponding exposure to A0 and 0B.
+
+But i think this shouldn't effect single insecticide runs.
+
+Am I sure I'm not mixing up rows somewhere in effectOfIndividualInputs.Rmd ??
+
+I could try setting the values of results for I1 and see if it propogates through.
+
+Instead I saved the original random value of exposure and used that in PRCC.
+
+I think what was happening was that correctDeployProp influenced the AB exposure value for mixtures and that was the predictor I took for exposure. Instead I want the intended exposure value.
+
+I still don't quite understand how this effects the results in the single insecticide runs.
+
+but with my 100 test runs what I'm seeing in the PRCC is no effect of anything for insect1 & 2. and a similar pattern as before for mixture1.
+
+This may point me towards another problem that is causing this.
+I may need to exclude corectDeploy as a predictor from those runs in which it wasn't used ??
+
+** 2/2/16 **
+
+~ ran 10000 runs over night
+~ run effectOfIndividualInputs
+~ now it's fine !!! PRCC looking good.
+
+New trees from 10,000 : 
+Predictors :
+effectiveness_insecticide1 :9
+effectiveness_insecticide2 :9
+exposure :                  7
+
+dominance_allele2 :         3
+dominance_allele1 :         4
+resist_start_1_div_2 :      1
+start_freq_allele2 :        1
+start_freq_allele1 :        2
+
+Best tree to show (if indeed trees are the best) : 
+Can't find a best one ! I don't like these trees !!
+Try to find a better method for Ian. Could it be PRCC.
+
+Violin plots. look good. They are very similar for time to resistance 0.1, 0.25 or 0.5. 
+~ need a plot of proportion of runs in which mixture or sequential better. Maybe I could do a violin plot of the time to threshold, with diff strategies on x axis.
+
+~ faceting them by exposure is cool
+~ faceting by effectivity shows very little
+
+** 5/2/16 **
+
+trying to get plots working
+
+## GOOD RESULTS ! :
+
+###1 effectiveness < 0.6 sequential best. > 0.6 mix best. 
+Plot of time_to_resistance against effectiveness  with 4 strategies on the same plot is key.
+
+It shows the line for mix2 crossing that of sequential. Effectiveness 0.4-0.6 resistance arises faster for mix2. > 0.6 resistance arises faster for sequential.
+
+I should adapt this plot perhaps just to show mix2 and sequential.
+
+###2 selection coeff < 0.2 sequential best. > 0.2 mix best.
+selection coefficient curves also cross for mix2 and sequential. < ~0.2 sequential is better. > 0.2 mixture is better. 
+But check that this is multiplied as shown below. So beware could this be caused by the effectiveness result ?
+    ##effectiveness
+    phi.SS1_A0 <- runif(1, min=0.4, max=1)
+    phi.SS2_0B <- runif(1, min=0.4, max=1)
+
+    # Ian suggested this should be dependent on effectiveness to ensure fitness of RR stays below 1 
+    s.RR1_A0 <- runif(1, min=0.2, max=1) * phi.SS1_A0
+    s.RR2_0B <- runif(1, min=0.2, max=1) * phi.SS2_0B
+
+I think perhaps I should remove multiplication of selection coefficient by effectiveness ?
+
+###3
+Effect of exposure is greater overall on times to resistance, but it has a more similar effect on mixtures & sequential thus it is of less importance when considering the differences between the two strategies.
+
+In majority of cases the point where resistance to just one of the insecticides in a mixture is reached comes before that for both insecticides in sequential use. [this seems obvious]
+Also it always takes longer to reach resistance for both insecticides when starting with both in a mixture. Is this obvious, or a meaningfull result or an error in our methodology ?
+
+I think for the paper we should focus on mix2 versus sequential. We could have that as what we call mixture, and then refer in smallere analyses to the other mix scenarios.
+Maybe I should have a quick look at curtis paper to remind myself ?
+
+Tried removing multiplication of selection coefficient by effectiveness.
+But got warnings about fitnesses > 1.
+
+6/2/16
+
+problem is with locus fitness values going over 1
+
+This is the problem that causes it :
+
+Wloci[ paste0('RR',locusNum), exposure] <- (1 - phi[locusNum, exposure]) + 
+                                                   (s[locusNum, exposure])
+
+so it just occurs when s > phi, selection coeff > effectiveness
+
+current range :
+effectiveness : 0.4 - 1
+selection coeff : 0.2 - 0.5
+
+Options : 
+1) discard sims in which selection > effectiveness
+2) set selection range to 0.1-0.45 and effectiveness from 0.45-1
+
+Doing 2) might be good to check whether the line crossing still goes on ...
+
+Yes saving 2) as listOutMix_ex2_1000.rda etc.
+
+ Done in effectOfIndividualInputs
+
+** 7/2/16 **
+with new runs
+
+Selection coefficient curves for mix2 & sequential no longer cross. Seems likely that this previous pattern was generated by correlation with effectiveness.
+
+** 9/2/16 **
+sent new figs of PRCC of diff between seq & mix2 to Ian
+
+** 10/2/16 **
+
+~ reran trees they show that effectiveness > ~ 0.7 & exposure < 0.7 favours mixture2 over sequential
+~ this is consistent with my graphs of difference vs effect & exposure (therefore are the trees really necessary ?)
+
+mix2 can already be displayed in plotcurtis_f2_generic() using addCombinedStrategy
+I had just set default to F because the plot gets even more confusing.
+
+Maybe add option to just plot mix2 ?
+
+Also offer option to plot from : runcurtis_f2()
+
+This which simply reduces effectiveness of insecticide2 from 1 to 0.85, makes seq better than mix (where mix was better for curtis).
+
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+** 11/2/16 **
+~ added default curtis values to UI labels in server.r
+
+### TO DO
+
+~ check on plotting of combined strategy, i think I may not be doing it right
+
+~ for the paper we could have a couple of figures that show how a relatively small change to effectiveness can alter the curtis plot.
+
+effectI2 from 1 to 0.85, & resistance arises faster for mix1 than seq
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+effectI2 from 1 to 0.5, & resistance arises faster for mix2 than seq
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+~ play with shiny app to help my understanding of how modifying the params alters results
+
+~ check code in findResistancePointsMixResponsive()
+
+## might be good to facet by mixture sequence better and just have points of x=exposure, y=effectiveness_insecticide1
+
+~ can I get plots to show just a subset of points ?
+
+~ produce a graphich of the 3 diff mixture scenarios & sequential
+~~ may be able to use the seriousGame plot method
+~~ I could even show resistance curves next to it ?
+
+### My summary of results & thoughts
+
+Resistance to two insecticides arises more slowly if they are used in combination rather than sequentially. See violin plots.
+
+~ I'd like to show prcc on an axis from 0-1 with diff sysmbols for neg & pos. We care about which are most imp. params, not whether they are pos or neg.
+
+~ If I would like to get prcc plots in a similar format to my existing plots. (but may not be space for axis labels)
+~ get data in form : 
+input1, input2, inputN, strategy 
+prcc vals,,             indecticide1
+
+# would it be useful for Ian for me to create an interactive shiny app to view the 1000 sim data. E.g. could just have the ggplot code with options to set the x value & colour by different inputs.
+# maybe try faceting by parameter ranges ?
 
 1. create a vignette that shows how best to run current model
-
-1. I could ad something to shinyCurtis with a message at the top saying whether mixture or sequence better and maybe an arrow pointing to it.
-
 1. fix the testthat tests to avoid new input addition
-
 1. create functions fitnessLocus(), fitnessNiche(), fitnessIndiv()
@@
 1. submit stackoverflow question about createArray2(), including about n=3 in eval.parent
-1. what to call createArray2() ? maybe namedArray, or arrayNamed() ? could even put back in the option to pass the fill value ? submit to stackoverflow ?
-
+1. what to call createArray2() ? maybe namedArray, or arrayNamed() ? could even put back in the option to pass the fill value ? submit to stackoverflow 
 1. maybe add option in setInputSensiScenarios to save input object as a csv or rda.
-
 1. add some defensive checks, e.g. for sensible input variables
-
-1. create a vignette from first shiny app documenting how the model works
-
-1. finish tidying documentation of reading in input vars in createInputMatrix()
-
-1. the curtis fig2 plot is produced from >1 scenario, do I want to get it into 1 'reactive' scenario
 
@@
 1. rename W to ?fitness : fitnessLocus, fitnessNiche and fitnessIndiv
-
-### refactoring variables to arrays
-Refactoring to arrays reduces code volume, reduces risk of mistyped variable names and allow objects to be passed between functions.
-cis/trans makes it slightly tricky, but not all variables have cis/trans
-only used for genotype frequencies (f) and in output matrices fitness & genotype 
-in the wild in Beths code :
-```
-fs.m.RS1RS2_cis <- f.m.RS1RS2_cis
-fs.m.RS1RS2_trans <- f.m.RS1RS2_trans
-```
 
@@
 `W.m.SS1SS2 : [sex, locus1, locus2]` (Windiv)
-
-### refactoring, what I have done
-Replaced the following collections of individual variables with arrays :
-(To reduce code volume, reduce risk of mistyped variable names and allow objects to be passed between functions. Also potentially allows it to cope with more loci and/or niches later).
 

### 2016-02-12 paper1_results_figs.Rmd
added curtis inputs line to plots

@@
 # Figures for paper 1 
-# version 3 - in progress
+# version 4 - in progress
 
@@
 
-  #remember these don't have new exposure column
   load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
@@
 
-# ex10000 runs showed effect of selection    
-#   load(file=paste0(outFolder,'listOutMix_ex10000.rda'))
-#   load(file=paste0(outFolder,'listOutI1_ex10000.rda'))
-#   load(file=paste0(outFolder,'listOutI2_ex10000.rda'))
+  # load(file=paste0(outFolder,'listOutMix_ex2_1000.rda'))
+  # load(file=paste0(outFolder,'listOutI1_ex2_1000.rda'))
+  # load(file=paste0(outFolder,'listOutI2_ex2_1000.rda'))
+
+# very quick test data
+  # load(file=paste0(outFolder,'listOutMix_ex2_3.rda'))
+  # load(file=paste0(outFolder,'listOutI1_ex2_3.rda'))
+  # load(file=paste0(outFolder,'listOutI2_ex2_3.rda'))
 
@@
 
+  #create a curtisInputs dataframe for use later
+  curtisInputs <- data.frame("exposure"=0.9,
+                             "effectiveness_insecticide1"=0.73,
+                             "effectiveness_insecticide2"=1,
+                             "start_freq_allele1"=0.01,
+                             "start_freq_allele2"=0.01,
+                             "dominance_allele1"=0.17,
+                             "dominance_allele2"=0.0016,
+                             "selection_coef_allele1"=0.23,
+                             "selection_coef_allele2"=0.43,
+                             "maleExposureProp"=1,
+                             "correctMixDeployProp"=1,
+                             "resist_start_hi_div_lo"=1,
+                             "resist_start_1_div_2"=1)
+
 ```
@@
 # getting data into format for ggplot
-
-  library(ggplot2)
 
@@
 
+  library(ggplot2)
+
 #names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
@@
              #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i))
+             labs(title = i) +
+             #add line for Curtis inputs
+             geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
+             #tryting to get the vline into the legend fails
+             #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
+
+           )
   }
@@
 
+```
+
+# Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
+### Curtis' Fig 2. inputs : mix2 > combined > mix1 > seq
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+```
+
+### decrease effectiveness of I2 from 1 to 0.85 : mix2 > combined > seq > mix1
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+```
+
+### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > combined > mix1
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
 ```

### 2016-02-12 resistanceModelNotesAndy.Rmd
notes

@@
 ~ added default curtis values to UI labels in server.r
+~ fixed plotting of combined strategy, i think I may not be doing it right
+
+** 12/2/16 **
+~ added position of Curtis inputs to the figures
 
 ### TO DO
-
-~ check on plotting of combined strategy, i think I may not be doing it right
 

### 2016-02-29 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 18/2/206 **
+Response email from Ian :
+
+Anyway I think the plots all look very nice, and give intuitive answers (but see point #1 below). My comments are
+
+(1)    It is very strange that “correct deployment” parameter has no impact on time until resistance in mixtures.
+
+(2)    Re your question at the end of the document i.e.
+“In mixtures what causes resistance to the 2nd insecticide to arise more slowly when the first insecticide is removed from the mixture ?”.
+Isn’t it the opposite i.e.
+if we compare Mixture 2 (where the first insecticide is removed) then doesn’t resistance spread more rapidly compared to when the first insecticide is retained (as in mixture 3)??
+
+(3)    This question stimulated me to look again at your definitions of mixtures etc on the front page of the document.
+I know its only a draft and we both know what it means but I took the liberty of expanding the definition as in the attachment.
+Can you check they are correct and hence that I have correctly understood your question addressed in the previous point (and, if so, you may want to cut-and-paste them into the document).
+I also suggested a potential change (i.e.
+that mixture 1 and 3 be combined because they only differ in their output measure).
+This would clarify the captions on the plots but it just depends how easily you could cut-and-paste such changes.
+
+One other thing that has just occurred to me.
+One option not considered by Curtis, but commonly proposed now, is to deploy  “mosaics” where insecticides are deployed singly but at the same time…so some fields/huts may have insecticide “A” while some will have insecticide “B”.
+Could we simulate these mosaics simply  by setting the “correct deployment” parameter to zero??
+It would be  another good selling point for the manuscript.
+If it is straightforward to set up, I would do it, If not we could just note how we could do it in the ms.
+It really follows on from point #1 above… if correct deployment has no effect even when it reaches zero then it implies mosaics are as good as mixtures.
+
+Ians recomendations for captions :
+Single use of either insecticide. The outcome measure is time for resistance allele frequency to reach the user-defined threshold value (0.25 in these case)
+
+Sequential : sole use of one insecticide, switch to other when threshold reached. Outcome measure is the time until frequency of resistance to the second insecticide reaches the threshold value.
+
+Mixture 1: Insecticides always used as a mixture. Outcome measure is the time until threshold reached for either insecticide i.e. the frequency of resistance alleles exceeds 0.25 at either locus
+
+Mixture 2 : insecticides initially deployed as mixtures until the threshold is reached for either insecticide, then switch to sole use of the other insecticide.
+Outcome measure is time until resistance allele frequencies exceed the threshold value for the second insecticide
+
+Mixture 3 : Insecticides always used as a mixture. The outcome measure is the time until the threshold allele frequency exceeds the threshold value at both loci.
+
+Actually, it would be much better if we combined mixture 1 and 3 and just note the difference in outcome measure. Mixture 3 could be renamed “Adaptive mixture”. In the figures the key could then be
+
+Single
+Sequential
+Mixture (either AI)*
+Mixture (both AIs)
+Adaptive mixture
+
+* I assume you know this but AI is short for “active ingredient” which the commercial developers much prefer to “insecticide” for reasons that elude me
+
+Which would be much easier for readers (and reviewers) to remember
+
+**26/2/16 **
+
+look at Ians point :
+Re your question at the end of the document i.e.
+“In mixtures what causes resistance to the 2nd insecticide to arise more slowly when the first insecticide is removed from the mixture ?”.
+Isn’t it the opposite i.e.
+if we compare Mixture 2 (where the first insecticide is removed) then doesn’t resistance spread more rapidly compared to when the first insecticide is retained (as in mixture 3)??
+
+Resistance spreads more rapidly when the 2nd insecticide is removed from a mixture. See Fig 1.
+
+Perhaps I should rename all references to mix 1,2 & 3 to avoid confusion of getting tangled up with this again.
+
+Sequential                                     or seq
+Mixture (either AI)   [mixture1] or 1st in mix or mix1
+Mixture (both AIs)    [mixture3] or 2nd in mix or mix2
+Adaptive mixture      [mixture2] or Adaptive mix or comb
+
+Hi Ian,
+
+Firstly, yes you are right. Resistance spreads more rapidly in the responsive mixture strategy than when the first insecticide is retained.
+
+I agree that we need to rationalise the names to try to minimise this confusion. I would prefer to not use AI in labels if possible, but we can refer to in text. (but we can discuss).
+
+If we could agree consistent long, medium and short names as started below that would be very helpful.
+I can then go through my scripts and rename.
+I'll have to be very careful because I will need to rename mix2 to mixA & mix3 to mix2.
+So if I miss any of the first mix2s we'll get in a right pickle.
+
+Also can we agree on what to call the strategy that changes from mix to sole ?
+My fault that I've been variously calling it adaptive, responsive, combination.
+Is there a precedent ?
+Adaptive seems potentially confusing as it could be confused with being selected for.
+I have a slight preference for responsive.
+
+Feel free to edit these below. Once we are agreed I'll print out and bluetac to my wall so that I stick to it !!
+
+Long Names (in text where space not an issue)
+Mixture (1st insecticide)   
+Mixture (2nd insecticide)
+Adaptive mixture (2nd insecticide)
+Sequential (2nd insecticide)
+
+Medium Names (plot labels where medium space)
+mix 1st              
+mix 2nd             
+mix adaptive
+sequential 2nd
+
+Short Names (plot labels where v little space)
+mix1
+mix2
+mixA
+seq
+
+cheers,
+Andy
+
+** 29/2/2016 **
+
+checking on correctMixDeployProp after Ian commented :
+
+(1)    It is very strange that “correct deployment” parameter has no impact on time until resistance in mixtures.
+
+In the plots the only effect was on mix1 that seemed to rise slightly.
+
+in setExposure() this is how it is done :
+
+    if ( correctMixDeployProp < 1 )
+    {
+      a['f','A','0'] <- (1 - correctMixDeployProp)/2 * exposure
+
+      a['m','A','0'] <- a['f','A','0'] * maleExposureProp
+
+      #m&f together
+      a[,'0','B'] <- a[,'A','0']    
+      a[,'0','0'] <- 1 - (a[,'A','B'] + a[,'A','0'] + a[,'0','B'])
+
+    }
+
+Code seems as expected.
+
+Could anything about how the sequential & adaptive mixture scenarios are patched together influence this ?
+I don't think so.
+
+#' @param correctMixDeployProp proportion of times that mixture is deployed correctly, 
+#'    assumes that when not deployed correctly the single insecticides are used instead
+
+~ Thinking about correctMixDeployProp, how do we choose which of the 2 insecticides is used singly ?
+It is divided between them.
+Maybe this is the issue ?
+Perhps splitting between the 2 single applications in the same simulation has a similar effect as the 'mixture' ?
+
+~ maybe write a function to produce a graphical output from setExposure() could put it into a UI to send to Ian.
+~~ effectively it just needs to visualise the array a.
+~~ I could firstly vis as a table rather than a graph
+~~ outputs : 2 tables side by side, a["f",,] & a["m",,]
+https://andysouth.shinyapps.io/shinyNiche/
+
 ### TO DO
 
-~ for the paper we could have a couple of figures that show how a relatively small change to effectiveness can alter the curtis plot.
+~ checking on correctMixDeployProp
+~ add correctMixDeployProp into curtis shiny UI to enable checking mechanism
+~~ first add correctMixDeployProp into runcurtis_f2
+
+~ carefully implement strategy name changes
 
@@
 runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-
-~ play with shiny app to help my understanding of how modifying the params alters results
 

### 2016-02-29 paper1_results_figs.Rmd
tweak

@@
 Comparing these insecticide use strategies:
+
 1. Single use
-1. Mixture 1 : threshold reached for either insecticide
-1. Mixture 2 : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
-1. Mixture 3 : threshold reached for both insecticides in mixture
+1. Mixture (either AI) : threshold reached for either insecticide
+1. Adaptive mixture : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
+1. Mixture (both AIs) : threshold reached for both insecticides in mixture
 1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold
+
+Renamed from previous naming system :
+Mixture (either AI)   [mixture1]
+Mixture (both AIs)    [mixture3]
+Adaptive mixture      [mixture2]
 
@@
 1.
-In between these extremes, there is less difference between the alternative strategies of a) sequential and b) mixture switching to sole use.
+In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture switching to sole use.
 The latter takes slightly longer to reach the threshold.
@@
 1.
-Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) mixture switching to sole use.
+Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) adaptive mixture switching to sole use.
 Below an effectiveness of ~0.55 to 0.6 sequential takes longer, and above this 'mixture switching to sole use' takes longer.
@@
 
-# Time-to-resistance across all input values
+# Fig. x1 Time-to-resistance across all input values
 
@@
       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
-      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="2nd in mix\nswitch to sole", "Mixture 3"="2nd in mix" ) )
+      #scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="2nd in mix\nswitch to sole", "Mixture 3"="2nd in mix" ) )
+      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="adaptive mix", "Mixture 3"="2nd in mix" ) )
 
@@
 
-## How time-to-resistance is effected by input values for each insecticide use strategy
+## Fig. x2 How time-to-resistance is effected by input values for each insecticide use strategy
 
@@
 
-## How difference in time-to-resistance between sequential and mixture2 is effected by input values
+## How difference in time-to-resistance between sequential and adaptive mixture is effected by input values
 

### 2016-03-01 paper1_results_figs.Rmd
findreplace mix2_minus with mixA_minus

@@
 
-  #df_diff$mix2_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
-  df_diff["mix2_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
+  #df_diff$mixA_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
+  df_diff["mixA_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
 
-  #TODO : later do PRCC on df_diff["mix2_minus_seq0.25"]
+  #TODO : later do PRCC on df_diff["mixA_minus_seq0.25"]
 
@@
 
-    y <- "mix2_minus_seq0.25"
+    y <- "mixA_minus_seq0.25"
 
@@
     x <- df_diff[,1:num_inputs]
-    y <- df_diff["mix2_minus_seq0.25"]
+    y <- df_diff["mixA_minus_seq0.25"]
 
@@
              ylim(-1,1) +
-             ggtitle(paste("PRCC difference between sequential and mixture2")) +
+             ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
              xlab(NULL)
@@
 ```
-
-## Andys questions 
-1. In mixtures what causes resistance to the 2nd insecticide to arise more slowly when the first insecticide is removed from the mixture ?

### 2016-03-01 paper1_results_figs.Rmd
name changes from mix1,2,3 to mix 1,adaptive,2

@@
 # Figures for paper 1 
-# version 4 - in progress
+# version 5 - in progress
 
@@
   #find time to criticalPoint for EITHER insecticide in mixture  
-  resistPointsMix1 <- findResistancePoints(listOutMix, locus='either')  
+  resistPointsMix_1 <- findResistancePoints(listOutMix, locus='either')  
   #todo - to be comparable I think this should be for when resistance to BOTH insecticides is reached
@@
   #inputs : inAndOutI1, inAndOutI2, inAndOutMix
-  resistPointsMix2 <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
+  resistPointsMix_A <- findResistancePointsMixResponsive(listOutMix, listOutI1, listOutI2)
 
@@
   #find time to criticalPoint for BOTH insecticide in mixture  
-  resistPointsMix3 <- findResistancePoints(listOutMix, locus='both')
+  resistPointsMix_2 <- findResistancePoints(listOutMix, locus='both')
 
@@
   resistPointsSeq_T <- addStrategyColumn(resistPointsSeq,"Sequential",ggInput) 
-  resistPointsMix1_T <- addStrategyColumn(resistPointsMix1,"Mixture 1",ggInput) 
-  resistPointsMix2_T <- addStrategyColumn(resistPointsMix2,"Mixture 2",ggInput) 
-  resistPointsMix3_T <- addStrategyColumn(resistPointsMix3,"Mixture 3",ggInput)
+  resistPointsMix_1_T <- addStrategyColumn(resistPointsMix_1,"Mix either",ggInput) 
+  resistPointsMix_A_T <- addStrategyColumn(resistPointsMix_A,"Mix adaptive",ggInput) 
+  resistPointsMix_2_T <- addStrategyColumn(resistPointsMix_2,"Mix both",ggInput)
 
   ggInsOuts <- rbind( resistPointsI1_T, resistPointsI2_T, resistPointsSeq_T, 
-                      resistPointsMix1_T, resistPointsMix2_T, resistPointsMix3_T)
+                      resistPointsMix_1_T, resistPointsMix_A_T, resistPointsMix_2_T)
 
@@
   #format data in a different way to enable PRCC on difference between sequential & mix2
-  #resistPointsMix2_T
+  #resistPointsMix_A_T
   #resistPointsSeq_T
@@
   #columns to remove from first
-  indices1 <- which(names(resistPointsMix2_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
+  indices1 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
   #columns to add to 2nd  
-  indices2 <- which(names(resistPointsMix2_T) %in% c("gen_cP0.25"))  
+  indices2 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.25"))  
   #rename column in 2nd
@@
 
-  df_diff <- cbind(resistPointsMix2_T[-indices1], tmp)
+  df_diff <- cbind(resistPointsMix_A_T[-indices1], tmp)
 
@@
       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
-      #scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="2nd in mix\nswitch to sole", "Mixture 3"="2nd in mix" ) )
-      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mixture 1","Sequential","Mixture 2","Mixture 3"), labels = c("insecticide 1"="Sole use", "Mixture 1"="1st in mix","Sequential"="2nd in sequence", "Mixture 2"="adaptive mix", "Mixture 3"="2nd in mix" ) )
+      #scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Sole use", "Mix either"="1st in mix","Sequential"="2nd in sequence", "Mix adaptive"="2nd in mix\nswitch to sole", "Mix both"="2nd in mix" ) )
+      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Sole use", "Mix either"="1st in mix","Sequential"="2nd in sequence", "Mix adaptive"="adaptive mix", "Mix both"="2nd in mix" ) )
 
@@
 # Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
-### Curtis' Fig 2. inputs : mix2 > combined > mix1 > seq
+### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
@@
 
-### decrease effectiveness of I2 from 1 to 0.85 : mix2 > combined > seq > mix1
+### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
@@
 
-### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > combined > mix1
+### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
@@
 
-# PRCC difference in time-to-resistance between sequential and mixture2
+# PRCC difference in time-to-resistance between sequential and mixture adaptive
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}

### 2016-03-01 sensiAnPaper1All.Rmd
name changes from mix1,2,3 to mix 1,adaptive,2

@@
 
-```{r, eval=FALSE}
+```{r, eval=TRUE}
   ## to load previously saved runs
@@
   #T/F
-  resistBetterMix1SeqBoolean <- resistPointsMix1 > resistPointsSeq
+  resistBetterMix_1SeqBoolean <- resistPointsMix_1 > resistPointsSeq
   #convert to 0/1
-  resistBetterMix1Seq <- plyr::aaply(resistBetterMix1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_1Seq <- plyr::aaply(resistBetterMix_1SeqBoolean,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix1Seq) <- gsub('gen','betterMix1Seq', rownames(resistBetterMix1Seq))
+  rownames(resistBetterMix_1Seq) <- gsub('gen','betterMix_1Seq', rownames(resistBetterMix_1Seq))
 
-  rowSums(resistBetterMix1Seq)/ncol(resistBetterMix1Seq)
+  rowSums(resistBetterMix_1Seq)/ncol(resistBetterMix_1Seq)
   #100runs
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
   #                .25                   .26                   .30   
   #1000runs
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
   #               .178                  .232                  .238 
   #new100
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
   #              0.18                 0.24                 0.25 
   #new1000
-  #betterMix1Seq_cP0.1 betterMix1Seq_cP0.25  betterMix1Seq_cP0.5 
+  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
   #             0.223                0.245                0.243
@@
   #T/F
-  resistBetterMix2SeqBoolean <- resistPointsMix2 > resistPointsSeq
+  resistBetterMix_ASeqBoolean <- resistPointsMix_A > resistPointsSeq
   #convert to 0/1
-  resistBetterMix2Seq <- plyr::aaply(resistBetterMix2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_ASeq <- plyr::aaply(resistBetterMix_ASeqBoolean,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix2Seq) <- gsub('gen','betterMix2Seq', rownames(resistBetterMix2Seq))
+  rownames(resistBetterMix_ASeq) <- gsub('gen','betterMix_ASeq', rownames(resistBetterMix_ASeq))
 
-  rowSums(resistBetterMix2Seq)/ncol(resistBetterMix2Seq)  
+  rowSums(resistBetterMix_ASeq)/ncol(resistBetterMix_ASeq)  
   #1000runs
-  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #betterMix_ASeq_cP0.1 betterMix_ASeq_cP0.25  betterMix_ASeq_cP0.5 
   #             0.471                0.568                0.551 
   #new100
-  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #betterMix_ASeq_cP0.1 betterMix_ASeq_cP0.25  betterMix_ASeq_cP0.5 
   #              0.58                 0.57                 0.52 
   #new1000
-  #betterMix2Seq_cP0.1 betterMix2Seq_cP0.25  betterMix2Seq_cP0.5 
+  #betterMix_ASeq_cP0.1 betterMix_ASeq_cP0.25  betterMix_ASeq_cP0.5 
   #             0.558                0.563                0.534
@@
   #T/F
-  resistBetterMix3SeqBoolean <- resistPointsMix3 > resistPointsSeq
+  resistBetterMix_2SeqBoolean <- resistPointsMix_2 > resistPointsSeq
   #convert to 0/1
-  resistBetterMix3Seq <- plyr::aaply(resistBetterMix3SeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_2Seq <- plyr::aaply(resistBetterMix_2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix3Seq) <- gsub('gen','betterMix3Seq', rownames(resistBetterMix3Seq))
+  rownames(resistBetterMix_2Seq) <- gsub('gen','betterMix_2Seq', rownames(resistBetterMix_2Seq))
 
   #tally runs where mix3 better than seq
-  rowSums(resistBetterMix3Seq)/ncol(resistBetterMix3Seq)
+  rowSums(resistBetterMix_2Seq)/ncol(resistBetterMix_2Seq)
   #100runs
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
   #                  .83                   .75                   .72 
   #1000runs
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
   #                .794                  .726                  .673  
   #new100
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
   #              0.83                 0.79                 0.72 
   #new1000
-  #betterMix3Seq_cP0.1 betterMix3Seq_cP0.25  betterMix3Seq_cP0.5 
+  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
   #             0.820                0.747                0.677
@@
 #   #what is the absolute difference between mix3 & sequential ?
-#   resistDifferenceMix3Seq <- resistPointsMix3 - resistPointsSeq
+#   resistDifferenceMix_2Seq <- resistPointsMix_2 - resistPointsSeq
 #   
@@
 #   #looks mostly plausible, occasionally there are 0s
-#   resistDifferenceMix1I1 <- resistPointsMix1 - resistPointsI1
-#   resistDifferenceMix1I2 <- resistPointsMix1 - resistPointsI2
+#   resistDifferenceMix_1I1 <- resistPointsMix_1 - resistPointsI1
+#   resistDifferenceMix_1I2 <- resistPointsMix_1 - resistPointsI2
 
@@
   #T/F
-  resistBetterMix1SeqBoolean20 <- resistPointsMix1 > (1.2 * resistPointsSeq)
+  resistBetterMix_1SeqBoolean20 <- resistPointsMix_1 > (1.2 * resistPointsSeq)
   #convert to 0/1
-  resistBetterMix1Seq20 <- plyr::aaply(resistBetterMix1SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_1Seq20 <- plyr::aaply(resistBetterMix_1SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix1Seq20) <- gsub('gen','betterMix1Seq20', rownames(resistBetterMix1Seq20))
+  rownames(resistBetterMix_1Seq20) <- gsub('gen','betterMix_1Seq20', rownames(resistBetterMix_1Seq20))
 
-  rowSums(resistBetterMix1Seq20)/ncol(resistBetterMix1Seq20)
+  rowSums(resistBetterMix_1Seq20)/ncol(resistBetterMix_1Seq20)
   #1000runs
- #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+ #betterMix_1Seq20_cP0.1 betterMix_1Seq20_cP0.25  betterMix_1Seq20_cP0.5 
  #                 0.136                  0.175                  0.173 
   #new1000
- #betterMix1Seq20_cP0.1 betterMix1Seq20_cP0.25  betterMix1Seq20_cP0.5 
+ #betterMix_1Seq20_cP0.1 betterMix_1Seq20_cP0.25  betterMix_1Seq20_cP0.5 
 #                 0.163                  0.178                  0.182
@@
   #T/F
-  resistBetterMix2SeqBoolean20 <- resistPointsMix2 > (1.2 * resistPointsSeq)
+  resistBetterMix_ASeqBoolean20 <- resistPointsMix_A > (1.2 * resistPointsSeq)
   #convert to 0/1
-  resistBetterMix2Seq20 <- plyr::aaply(resistBetterMix2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_ASeq20 <- plyr::aaply(resistBetterMix_ASeqBoolean20,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix2Seq20) <- gsub('gen','betterMix2Seq20', rownames(resistBetterMix2Seq20))
+  rownames(resistBetterMix_ASeq20) <- gsub('gen','betterMix_ASeq20', rownames(resistBetterMix_ASeq20))
 
-  rowSums(resistBetterMix2Seq20)/ncol(resistBetterMix2Seq20)  
+  rowSums(resistBetterMix_ASeq20)/ncol(resistBetterMix_ASeq20)  
   #1000runs
-#  betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+#  betterMix_ASeq20_cP0.1 betterMix_ASeq20_cP0.25  betterMix_ASeq20_cP0.5 
 #                  0.241                  0.326                  0.313 
   #new1000
-# betterMix2Seq20_cP0.1 betterMix2Seq20_cP0.25  betterMix2Seq20_cP0.5 
+# betterMix_ASeq20_cP0.1 betterMix_ASeq20_cP0.25  betterMix_ASeq20_cP0.5 
 #                 0.283                  0.302                  0.283
@@
   #T/F
-  resistBetterMix3SeqBoolean20 <- resistPointsMix3 > (1.2 * resistPointsSeq)
+  resistBetterMix_2SeqBoolean20 <- resistPointsMix_2 > (1.2 * resistPointsSeq)
   #convert to 0/1
-  resistBetterMix3Seq20 <- plyr::aaply(resistBetterMix3SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_2Seq20 <- plyr::aaply(resistBetterMix_2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix3Seq20) <- gsub('gen','betterMix3Seq20', rownames(resistBetterMix3Seq20))
+  rownames(resistBetterMix_2Seq20) <- gsub('gen','betterMix_2Seq20', rownames(resistBetterMix_2Seq20))
 
   #tally runs where mix3 better than seq
-  rowSums(resistBetterMix3Seq20)/ncol(resistBetterMix3Seq20)
+  rowSums(resistBetterMix_2Seq20)/ncol(resistBetterMix_2Seq20)
   #1000runs
-#  betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+#  betterMix_2Seq20_cP0.1 betterMix_2Seq20_cP0.25  betterMix_2Seq20_cP0.5 
 #                  0.672                  0.560                  0.488   
   #new1000
-# betterMix3Seq20_cP0.1 betterMix3Seq20_cP0.25  betterMix3Seq20_cP0.5 
+# betterMix_2Seq20_cP0.1 betterMix_2Seq20_cP0.25  betterMix_2Seq20_cP0.5 
 #                 0.622                  0.536                  0.458
@@
 
-  resistBetterMix1Seq <- repMix(resistBetterMix1Seq)
-  resistBetterMix2Seq <- repMix(resistBetterMix2Seq)
-  resistBetterMix3Seq <- repMix(resistBetterMix3Seq)
+  resistBetterMix_1Seq <- repMix(resistBetterMix_1Seq)
+  resistBetterMix_ASeq <- repMix(resistBetterMix_ASeq)
+  resistBetterMix_2Seq <- repMix(resistBetterMix_2Seq)
 
-  resistBetterMix1Seq20 <- repMix(resistBetterMix1Seq20)
-  resistBetterMix2Seq20 <- repMix(resistBetterMix2Seq20)
-  resistBetterMix3Seq20 <- repMix(resistBetterMix3Seq20)
+  resistBetterMix_1Seq20 <- repMix(resistBetterMix_1Seq20)
+  resistBetterMix_ASeq20 <- repMix(resistBetterMix_ASeq20)
+  resistBetterMix_2Seq20 <- repMix(resistBetterMix_2Seq20)
 
   #bind results onto predictors
-  treeInput <- rbind( treeInput, resistBetterMix1Seq )
-  treeInput <- rbind( treeInput, resistBetterMix2Seq )
-  treeInput <- rbind( treeInput, resistBetterMix3Seq )
+  treeInput <- rbind( treeInput, resistBetterMix_1Seq )
+  treeInput <- rbind( treeInput, resistBetterMix_ASeq )
+  treeInput <- rbind( treeInput, resistBetterMix_2Seq )
   #add >20% results
-  treeInput <- rbind( treeInput, resistBetterMix1Seq20 )
-  treeInput <- rbind( treeInput, resistBetterMix2Seq20 )
-  treeInput <- rbind( treeInput, resistBetterMix3Seq20 )
+  treeInput <- rbind( treeInput, resistBetterMix_1Seq20 )
+  treeInput <- rbind( treeInput, resistBetterMix_ASeq20 )
+  treeInput <- rbind( treeInput, resistBetterMix_2Seq20 )
 
@@
 
-  treeResponses <- c( rownames(resistBetterMix1Seq),
-                      rownames(resistBetterMix2Seq),
-                      rownames(resistBetterMix3Seq),
-                      rownames(resistBetterMix1Seq20),
-                      rownames(resistBetterMix2Seq20),
-                      rownames(resistBetterMix3Seq20)
+  treeResponses <- c( rownames(resistBetterMix_1Seq),
+                      rownames(resistBetterMix_ASeq),
+                      rownames(resistBetterMix_2Seq),
+                      rownames(resistBetterMix_1Seq20),
+                      rownames(resistBetterMix_ASeq20),
+                      rownames(resistBetterMix_2Seq20)
                       )
@@
 
-    treeResponse <- "betterMix2Seq_cP0.5"
+    treeResponse <- "betterMix_ASeq_cP0.5"
 
@@
     #comparing it with 0/1 gives the same big tree
-    treeInput$betterMix2Seq_cP0.5numeric <- as.numeric(treeInput$betterMix2Seq_cP0.5)
+    treeInput$betterMix_ASeq_cP0.5numeric <- as.numeric(treeInput$betterMix_ASeq_cP0.5)
 
-    treeResponse <- "betterMix2Seq_cP0.5numeric"
+    treeResponse <- "betterMix_ASeq_cP0.5numeric"
 
@@
 ```    
-#   treeResponse <- 'betterMix1Seq_cP0.5'
+#   treeResponse <- 'betterMix_1Seq_cP0.5'
 #   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
@@
 #   x11()
-#   treeResponse <- 'betterMix3Seq_cP0.5'
+#   treeResponse <- 'betterMix_2Seq_cP0.5'
 #   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')  

### 2016-03-01 resistanceModelNotesAndy.Rmd
notes

@@
 
+~ checking on correctMixDeployProp
+~ added correctMixDeployProp into curtis shiny UI to enable checking mechanism
+~~ first added correctMixDeployProp into runcurtis_f2
+
 ### TO DO
 
-~ checking on correctMixDeployProp
-~ add correctMixDeployProp into curtis shiny UI to enable checking mechanism
-~~ first add correctMixDeployProp into runcurtis_f2
-
 ~ carefully implement strategy name changes
+Find & replace todo : in sensiAnPaper1All.Rmd & paper1_results_figs.Rmd
+
+resistPointsMix2    resistPointsMix_A
+resistPointsMix3    resistPointsMix_2
+resistBetterMix2SeqBoolean  resistBetterMix_ASeqBoolean
+resistBetterMix2Seq         resistBetterMix_ASeq
+
+#### can I replace :
+Mix1 - Mix_1
+Mix2 - Mix_A
+Mix3 - Mix_2
+"Mixture 1" - "Mix either"
+"Mixture 2" - "Mix adaptive"
+"Mixture 3" - "Mix both"
+done mix2_minus_seq0.25          mixA_minus_seq0.25
+
+Long Names (in text where space not an issue)
+Mixture (either insecticide)   
+Mixture (both insecticides)
+Adaptive mixture (both insecticides)
+Sequential (both insecticides)
+
+Medium Names (plot labels where medium space)
+mix either              
+mix both             
+mix adaptive
+sequential
+
+Short Names (plot labels where v little space)
+mix1
+mix2
+adapt
+seq
 

### 2016-03-02 paper1_results_figs_slimmed.Rmd
add an edited main figs doc

--- 2016-03-01 paper1_results_figs.Rmd
@@
 ---
-  title: "paper1_results_figs"
+title: "paper1_results_figs_slimmed"
 output: pdf_document
+#output: word_document
 author: "Andy South"
@@
 
-# Figures for paper 1 
-# version 5 - in progress
-
-## Based on 10000 runs with all inputs varying at once which explains the variability at particular input values.
-
-Comparing these insecticide use strategies:
-
-1. Single use
-1. Mixture (either AI) : threshold reached for either insecticide
-1. Adaptive mixture : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
-1. Mixture (both AIs) : threshold reached for both insecticides in mixture
-1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold
-
-Renamed from previous naming system :
-Mixture (either AI)   [mixture1]
-Mixture (both AIs)    [mixture3]
-Adaptive mixture      [mixture2]
+## Main Figures for paper 1 : version 1 - in progress
+### This file is automagically generated by R from paper1_results_figs_slimmed.Rmd
 
@@
 1.
-In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture switching to sole use.
+In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture, switching to sole use.
 The latter takes slightly longer to reach the threshold.
-1. Exposure and effectiveness seem to be the most important inputs.
+1. Exposure and dominance have the greatest on time-to-resistance as shown in PRCC analyses.
 1. Exposure has the greatest effect on time-to-resistance.
@@
 Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) adaptive mixture switching to sole use.
-Below an effectiveness of ~0.55 to 0.6 sequential takes longer, and above this 'mixture switching to sole use' takes longer.
+Below an effectiveness of ~0.55 to 0.6 sequential takes longer tp reach the resistance threshold, and above this 'mixture switching to sole use' takes longer.
 
@@
 
+\pagebreak
+
 # Fig. x1 Time-to-resistance across all input values
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE}
+1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
+1.
+In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture, switching to sole use.
+The latter takes slightly longer to reach the threshold.
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=4.5}
 
@@
 
-## Fig. x2 How time-to-resistance is effected by input values for each insecticide use strategy
+\pagebreak
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+# Fig x2 PRCC analysis for time-to-resistance for sequential and mixture strategies
 
-  names_inputs <- names(ggInsOuts)[1:num_inputs]
+1. Exposure and dominance have the greatest on time-to-resistance as shown in PRCC analyses.
 
-  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- 'time_to_resistance0.25'
-
-    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
-    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
-            #points not wanted if 10000
-             #geom_point(shape=3, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             geom_smooth(linetype='dashed',size=1) +
-             #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i) +
-             #add line for Curtis inputs
-             geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
-             #tryting to get the vline into the legend fails
-             #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
-
-           )
-  }
-
-```
-
-## How difference in time-to-resistance between sequential and adaptive mixture is effected by input values
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
-  ggSubset <- df_diff
-
-  names_inputs <- names(ggSubset)[1:num_inputs]
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- "mixA_minus_seq0.25"
-
-    print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
-            #points not wanted if 10000
-             geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             geom_smooth(linetype='dashed',size=1, colour="red") +
-             #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i) +
-             geom_hline(yintercept = 0, colour="white")
-           )
-  }
-
-```
-
-# Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
-### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-# PRCC difference in time-to-resistance between sequential and mixture adaptive
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
-
-  library(sensitivity)
-
-    x <- df_diff[,1:num_inputs]
-    y <- df_diff["mixA_minus_seq0.25"]
-
-    pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
-
-    to_plot <- pcc_res$PRCC
-    #rename column 1 from 'original to PRCC
-    names(to_plot)[1] <- 'PRCC'
-    to_plot$inputs <- rownames(to_plot)
-
-    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
-             geom_point(shape=1, colour='red') +
-             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
-             geom_hline(yintercept = 0, linetype=3) +
-             ylim(-1,1) +
-             ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
-             xlab(NULL)
-          )
-
-```
-
-# PRCC analysis for time-to-resistance subsetted by each insecticide use strategy
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8}
 
@@
 
-  for(strategy in strategies)
+  #in main paper I want to just display sequential and mixture strategies
+  strategies <- strategies[-1:-2]
+
+  plotlist <- list(length(strategies))
+
+  for(strategy_num in 1:length(strategies))
   {  
+    strategy <- strategies[strategy_num]
 
@@
 
-    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+    #just labelling axis for final plot
+    if (strategy_num == length(strategies))
+    {
+      axis_text_x <- element_text(angle = 45,hjust = 1, vjust = 1)
+      height <- 0.9
+    } else
+    {
+      axis_text_x <- element_blank() 
+      height = 0.5
+    }
+
+    library(cowplot) # to enable setting relative heights of plots
+
+    #draw_plot(p,x=0,y=0,width=.5,height = .5)
+
+    # print( ggdraw() + #panel_border(remove = TRUE) +
+    #          draw_plot( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
+    # #print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+    #          geom_point(shape=1, colour='red') +
+    #          theme(axis.text.x = axis_text_x) +
+    #          #theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+    #          geom_hline(yintercept = 0, linetype=3) +
+    #          ylim(-1,1) +
+    #          #ggtitle(paste(strategy,"PRCC")) +
+    #          ggtitle(paste(strategy,"time-to-resistance 0.25")) +
+    #          xlab(NULL)
+    #       , height= height))
+
+    plotlist[[strategy_num]] <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
              geom_point(shape=1, colour='red') +
-             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             theme(axis.text.x = axis_text_x) +
              geom_hline(yintercept = 0, linetype=3) +
              ylim(-1,1) +
-             ggtitle(paste(strategy,"PRCC")) +
+             ggtitle(paste(strategy,"time-to-resistance 0.25")) +
              xlab(NULL) 
-          )
   }
 
-```
-
-## response of resistance thresholds to individual inputs faceted by strategy
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
-  names_inputs <- names(ggInsOuts)[1:num_inputs]
-
-  #todo find out to take sample this doesn't work
-  #ggSample <- sample(ggInsOuts,500)
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- 'time_to_resistance0.25'    
-    #y <- 'time_to_resistance0.5'
-    color <- 'strategy'
-    #coloring by exposure shows its effect even when other params are varying
-    #color <- 'exposure'
-
-    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-             #points not wanted if 10000
-             #geom_point(shape=3, show.legend=FALSE) + 
-             geom_smooth(colour='red', linetype='dashed',size=0.5, show.legend=FALSE) +
-             facet_wrap( ~ strategy) + #, show_guide = FALSE) + 
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i))
-  }
+    plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))
 

### 2016-03-02 paper1_results_figs.Rmd
tweak

@@
 
+---
+
 # Fig. x1 Time-to-resistance across all input values
@@
 
+---
 ## Fig. x2 How time-to-resistance is effected by input values for each insecticide use strategy
@@
              ylim(-1,1) +
-             ggtitle(paste(strategy,"PRCC")) +
+             #ggtitle(paste(strategy,"PRCC")) +
+             ggtitle(paste(strategy,"time-to-resistance 0.25")) +
              xlab(NULL)

### 2016-03-02 resistanceModelNotesAndy.Rmd
notes

@@
 ** 26/8/15 **
-submit invoice
-from contract : maximum of 25 days at 217 per day
-
+submit invoice : from contract : maximum of 25 days at 217 per day
 I spent 30.6 days (can claim some against tsetse work, or from future contract ...)
@@
 
-### TO DO
-
 ~ carefully implement strategy name changes
@@
 
-#### can I replace :
+#### replacements done :
 Mix1 - Mix_1
@@
 done mix2_minus_seq0.25          mixA_minus_seq0.25
+
+done ~ record time spent & when was the last time I claimed on the previous contract ?
+Previous contract claimed 15 Sept 2015.
+submit invoice : from contract : maximum of 25 days at 217 per day
+I spent 30.6 days (can claim some against tsetse work, or from future contract ...)
+So time I spent in January on toggl should be on new contract/
+To the end of Feb :
+Hours		105:45:00
+Hours rounded		105.00
+Days (/7.5)		14.00
+Days remaining (from 85)		71.00
+
+### TO DO
+
+~ can I facet the PRCC mutliplots so I just get one axis ?
+
+~ start creating a proposed slimmed down figs doc (maybe as a word doc using pandoc) proposed for the paper
+~~ beware that if I send it to Ian as a doc, he will insert text into it that I may not be able to find.
+~~ maybe send it as a pdf first & change to a doc later to allow Ian to copy & paste the figs.
+~~ see paper1_results_figs_slimmed.Rmd
 

### 2016-03-04 paper1_results_figs.Rmd
working on figures

@@
 
----
 ## Fig. x2 How time-to-resistance is effected by input values for each insecticide use strategy
@@
 
-## How difference in time-to-resistance between sequential and adaptive mixture is effected by input values
+## Fig. x3 How difference in time-to-resistance between sequential and adaptive mixture is effected by input values
 
@@
   }
+
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
+  #can I do a 2D hexbin plot with x=exposure, y=effectiveness
+  #but do I just want to use 1 effectiveness, or maybe add both effectivenesses together ?
+  #bin colour : proportion of runs in which sequential better & could have diverging scale about 0.5
+  #I probably still need to tweak data first
+  #currently I have differences in times to resistance : mixA_minus_seq0.25
+  #what I think I want is the proportion of runs in which mix is better
+  #I can get that by setting to 0 & 1 and calc average
+  # tst_runs <- c(1,0,0,0)
+  # tst_runs <- c(1,1,1,0)
+  # mean(tst_runs)
+
+  ggSubset <- df_diff
+  #if mix-seq >1 resistance took longer for mix so mix is better
+  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
+  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
+  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)
+
+  #takes a long time to plot, initial version gave me count
+  #I want it to plot the mean
+  #check stackoverflow ..
+
+    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+             geom_hex()
+             #points not wanted if 10000
+             #geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
+             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             #geom_smooth(linetype='dashed',size=1, colour="red") +
+             #facet_wrap( ~ strategy) +
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             #labs(title = i) +
+             #geom_hline(yintercept = 0, colour="white")
+           )
+
+    #interesting plot showing where sequence is better is restricted to higher values of exposure and effectiveness 
+    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+             #geom_hex()
+             geom_point(alpha=0.2)  
+           )
+
+    #can I multiply the two effectivenesses together ?
+    #aha! yes that's even better
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_better_seq0.25))+
+             #geom_hex()
+             geom_point(alpha=0.2)  
+           )
+
+    #i think adding the effectivenesses may give an even better plot
+    #and makes it easier for the reader to interpret
+    #aha! yes that's even better
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mixA_better_seq0.25))+
+
+             geom_point(alpha=0.2)  
+           )
+
+    #facetted 0/1, labelled mix better, seq better
+    #if I get this sorted I could try to have +ve & -ve on the same scale
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mixA0.25))+
+             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
+
+             theme(legend.position="none") +
+             geom_point(alpha=1)  
+           )
+
+    #what if I do the difference rather than just 0/1
+    #aha! this is close to being the killer plot, just need to get the colour scale right 
+    #but the previous one may be clearer
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_minus_seq0.25))+
+             #geom_hex()
+             geom_point(alpha=1) +
+             #no adding grey outlines obscures the pattern
+             #geom_point(shape=21, colour='grey') +
+             #this one does look good but slight potential to see negatives above even though they aren't there
+             #could I modify the shapes to be different
+             #would a repeated, faceted plot solve that ?
+             #scale looks good but colours could do with being denser
+             #and I'm not sure that making the points that are close to zero white is a good idea ?
+             #although it does mean points that are only a few generations different do not obscure the pattern
+             #maybe I could explicitly do Ians 20% thing & colour those white
+             scale_colour_gradient2()
+
+             #didn't manage to get the colorBrewer scale to diverge from 0
+             #scale_colour_distiller("RdYlGn",type='div')
+            )
+
+    #COOLIO !! I really like this figure.
+    library(viridis)
+    #try facetted showing generation differences, labelled mix better, seq better
+    #if I get this sorted I could try to have +ve & -ve on the same scale
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+
+             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
+
+             #YlGnBu looks really good except that the yellow points are very faint
+             #?viridis
+             #scale_colour_distiller(palette='YlGnBu', direction=1) +
+             #option="plasma" looks good too
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
+             theme_bw() +
+             #theme_minimal() +
+             geom_point(alpha=1)  
+           )
 

### 2016-03-04 paper1_results_figs_slimmed.Rmd
working on figures

@@
 
-## Main Figures for paper 1 : version 1 - in progress
+## Main Figures for paper 1 : version 2 - in progress
 ### This file is automagically generated by R from paper1_results_figs_slimmed.Rmd
@@
 1. Exposure and dominance have the greatest on time-to-resistance as shown in PRCC analyses.
-1. Exposure has the greatest effect on time-to-resistance.
 1.
-Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) adaptive mixture switching to sole use.
-Below an effectiveness of ~0.55 to 0.6 sequential takes longer tp reach the resistance threshold, and above this 'mixture switching to sole use' takes longer.
+Effectiveness of the insecticides is the one input that can change the average relative positioning of the strategies a) sequential and b) adaptive mixture.
+Below an effectiveness of ~0.55 to 0.6 sequential takes longer to reach the resistance threshold, and above this the adaptive mixture takes longer.
+1. Exposure and effectiveness come out as the most important effect on the difference between the sequential and adpative mixture strategies.
 
@@
 \pagebreak
-
-# Fig. x1 Time-to-resistance across all input values
+## Fig. x1 Time-to-resistance across all input values
 
@@
 \pagebreak
-
-# Fig x2 PRCC analysis for time-to-resistance for sequential and mixture strategies
+## Fig x2 PRCC analysis for time-to-resistance 0.25 for sequential and mixture strategies
 
@@
              geom_hline(yintercept = 0, linetype=3) +
+             #theme_bw() +
+             #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
              ylim(-1,1) +
-             ggtitle(paste(strategy,"time-to-resistance 0.25")) +
+             ggtitle(paste(strategy)) +
              xlab(NULL) 
@@
 ```
+
+\pagebreak
+## Fig. x3 How time-to-resistance is effected by input values for each insecticide use strategy
+
+1.
+Effectiveness of the insecticides is the one input that can change the average relative positioning of the strategies a) sequential and b) adaptive mixture.
+Below an effectiveness of ~0.55 to 0.6 sequential takes longer to reach the resistance threshold, and above this the adaptive mixture takes longer.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=2.5}
+
+  #names_inputs <- names(ggInsOuts)[1:num_inputs]
+
+  #remove insecticide1 & 2 strategies
+  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
+  #subset which output plots to show in the main paper
+  in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
+              "dominance_allele1","selection_coef_allele1","maleExposureProp")
+  out_cols <- c("time_to_resistance0.25","strategy")
+  ggSubset <- ggSubset[, c(in_cols,out_cols) ]
+
+  #num_inputs <- length(in_cols)
+  #names_inputs <- in_cols
+
+  #to rearrange order of legend entries
+  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either")) #, labels=c("MBB", "MAA", "MCC"))
+
+  #TODO could I facet the plots below rather than using the loop
+  #probably need to put all inputs into a single column and then facet by that
+
+  #for(i in names_inputs)
+  for(i in in_cols)
+    {
+    #x11()
+
+    y <- 'time_to_resistance0.25'
+
+    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
+    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
+            #points not wanted if 10000
+             #geom_point(shape=3, show.legend=FALSE) + 
+             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             geom_smooth(linetype='dashed',size=1) +
+             #facet_wrap( ~ strategy) +
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             #labs(title = i) +
+             #add line for Curtis inputs
+             geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
+             #tryting to get the vline into the legend fails
+             #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
+
+           )
+  }
+
+```
+
+\pagebreak
+## Fig. x4 PRCC difference in time-to-resistance 0.25 between sequential and adaptive mixture strategies
+
+1. Exposure and effectiveness come out as the most important effect on the difference between the sequential and adpative mixture strategies.
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
+
+  library(sensitivity)
+
+    x <- df_diff[,1:num_inputs]
+    y <- df_diff["mixA_minus_seq0.25"]
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    #results are here I can probably rbind them together into a df that I can ggplot
+    #pcc_res$PRCC
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
+
+    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             geom_hline(yintercept = 0, linetype=3) +
+             #theme_bw() +
+             ylim(-1,1) +
+             #ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
+             xlab(NULL)
+          )
+
+```
+
+\pagebreak
+## Fig. x5 Effectiveness and exposure together determine most of whether resistance arises faster under adaptive mixture or sequential strategies
+
+```{r, eval=TRUE, echo=FALSE, fig.height=5}
+
+  ggSubset <- df_diff
+  #if mix-seq >1 resistance took longer for mix so mix is better
+  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
+  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
+  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)
+
+    #COOLIO !! I really like this figure.
+    library(viridis)
+    #try facetted showing generation differences, labelled mix better, seq better
+    #if I get this sorted I could try to have +ve & -ve on the same scale
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+
+             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix\nslower to reach resistance',
+                                                                        '1'='sequential\nslower to reach resistance'))) +
+
+             #YlGnBu looks really good except that the yellow points are very faint
+             #?viridis
+             #scale_colour_distiller(palette='YlGnBu', direction=1) +
+             #option="plasma" looks good too
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in\ngenerations\nto reach\nresistance\nthreshold")) + 
+             theme_bw() +
+             #theme_minimal() +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') +             
+             geom_point(alpha=1, size=0.5)  
+           )
+
+```

### 2016-03-04 resistanceModelNotesAndy.Rmd
working on figures

@@
 
+** 2/3/2016 **
+
 done ~ record time spent & when was the last time I claimed on the previous contract ?
@@
 
+~ modifying PRCC mutliplots to get just one axis
+
+** 3/3/2016 **
+
+msg from Ian :
+
+Aha! I've understood what he means by setting correct_deploy to 0, to study mosaics.
+
 ### TO DO  
-
-~ can I facet the PRCC mutliplots so I just get one axis ?
+~ check stackoverflow for getting mean into hexbin plot
+
+~ communicate my concern to Ian that figures I've spent quite a long time on just go into supplementary material where few will see them.
+Can we think of a way of getting these into a paper themselves ?
+e.g.
+if this first paper is targeted at policy people can we get a 2nd paper that is targetted at biologists and explores mechanisms ?
+
+~ can I change order of strategies in legend ?
+~ can I facet the smoothed scatter plots to get them to take up less space and avoid repeating legend ?
+~~ or I may be able to use a similar plot_grid() method
+
+~~ if I wanted to facet by inputs might need to restructure the data, put inputs into a single column to replace below :
+  for(i in names_inputs)
+  {
+    print( ggplot(ggSubset, aes_string(x=i
+
+~ Ian wants trees. Maybe just do one tree of resistance0.25 on whether seq or maxAdapt is better.
+
+adaptive & sequential
+20% longer
 
@@
 ~~ see paper1_results_figs_slimmed.Rmd
+
+~ I could try a 2D plot with x=exposure, y=effectiveness & hexbin & colour cells according to proportion of runs where mix or seq is better.
+
+~~ seq is sometimes better than mix both. Where does that occur. The above plot might even help to answer that.
 

### 2016-03-08 paper1_results_figs.Rmd
figures

@@
 
+  #tricky
+  #be a bit careful what to do with zeros where results are same for both strategies
+  #I use these better vars to facet the plots
+  #may want the zeroes in both facets
+  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
+  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
+  #set these to 1 (they were 0)
+  df_same$mixA_better_seq0.25 <- 1
+  df_same$seq_better_mixA0.25 <- 1
+  #rbind them back onto other dataset
+  ggSubset <- rbind( ggSubset, df_same)
+
   #takes a long time to plot, initial version gave me count
@@
     library(viridis)
-    #try facetted showing generation differences, labelled mix better, seq better
-    #if I get this sorted I could try to have +ve & -ve on the same scale
+    #facet showing generation differences, labelled mix better, seq better
+    #+ve & -ve on the same colour scale
+    #failed attempt at getting Curtis point in legend
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff, size="Curtis"))+
     print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
@@
              #YlGnBu looks really good except that the yellow points are very faint
-             #?viridis
              #scale_colour_distiller(palette='YlGnBu', direction=1) +
@@
              #theme_minimal() +
-             geom_point(alpha=1)  
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             #failing to add legend for curtis point
+             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
+             #scale_size_manual(name = "Curtis Fig.2", values="test") +
+             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #add text for Curtis point
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
            )
+
+    #NOW what if I subset the 0 values & fit a linear model to it
+    #could that be used to decide between the 2 strategies ?
+    #I might even be able to do the lm plot within ggplot
+
+    #trickiness I have is that I have previously put 0 values in the adptive mix better plot 
+    #which is maybe a little misleading anyway
+    #and also means that I only get the lm line in the one plot
+    #maybe what I can do is add the zeros into both plots then  will get lines in both
+    #might even be good to overplot the zero values after
+    #or even remove the zero values ?
+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+
+             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
+
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
+             theme_bw() +
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #!!!NOW can I add a lm fitted to the 0 points abs_diff
+             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE) +
+             #can I subtley highlight the zero points which are repeated in each plot ?
+             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
+             #add text for Curtis point
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
+           )
+
+# doing the lm outside of ggplot    
+df_same = ggSubset[which(ggSubset$abs_diff==0),]
+x1=df_same$exposure 
+y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
+
+l_mod <- lm(y1 ~ x1)
+l_mod$coefficients
+#(Intercept)           x 
+#  0.9790948   0.6095132
+
+x2 <- seq(0,1,0.1)
+#approximation
+y2 <- 0.6*x2+0.98
+plot(x2,y2)
 

### 2016-03-08 paper1_results_figs_slimmed.Rmd
figures

@@
 
-## Main Figures for paper 1 : version 2 - in progress
+## Main Figures for paper 1 : version 3 - in progress
 ### This file is automagically generated by R from paper1_results_figs_slimmed.Rmd
@@
 1. Exposure and effectiveness come out as the most important effect on the difference between the sequential and adpative mixture strategies.
+1.
+Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
 
@@
 \pagebreak
+
 ## Fig. x1 Time-to-resistance across all input values
@@
 \pagebreak
+
 ## Fig x2 PRCC analysis for time-to-resistance 0.25 for sequential and mixture strategies
@@
 \pagebreak
+
 ## Fig. x3 How time-to-resistance is effected by input values for each insecticide use strategy
@@
 \pagebreak
+
 ## Fig. x4 PRCC difference in time-to-resistance 0.25 between sequential and adaptive mixture strategies
@@
 \pagebreak
-## Fig. x5 Effectiveness and exposure together determine most of whether resistance arises faster under adaptive mixture or sequential strategies
 
-```{r, eval=TRUE, echo=FALSE, fig.height=5}
+## Fig. x5 Effectiveness and exposure together determine most of whether resistance arises slower under adaptive mixture or sequential strategies
+
+1.
+Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=6}
 
@@
 
+  #tricky
+  #be a bit careful what to do with zeros where results are same for both strategies
+  #I use these better vars to facet the plots
+  #may want the zeroes in both facets
+  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
+  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
+  #set these to 1 (they were 0)
+  df_same$mixA_better_seq0.25 <- 1
+  df_same$seq_better_mixA0.25 <- 1
+  #rbind them back onto other dataset
+  ggSubset <- rbind( ggSubset, df_same)
+
     #COOLIO !! I really like this figure.
     library(viridis)
-    #try facetted showing generation differences, labelled mix better, seq better
-    #if I get this sorted I could try to have +ve & -ve on the same scale
+    #facet showing generation differences, labelled mix better, seq better
+    #+ve & -ve on the same colour scale
+    #failed attempt at getting Curtis point in legend
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff, size="Curtis"))+
     print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
 
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix\nslower to reach resistance',
-                                                                        '1'='sequential\nslower to reach resistance'))) +
+             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance slower','1'='sequential\nresistance slower'))) +
 
              #YlGnBu looks really good except that the yellow points are very faint
-             #?viridis
              #scale_colour_distiller(palette='YlGnBu', direction=1) +
@@
              scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in\ngenerations\nto reach\nresistance\nthreshold")) + 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
              theme_bw() +
+             theme(legend.position = "bottom") +
              #theme_minimal() +
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             #failing to add legend for curtis point
+             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
+             #scale_size_manual(name = "Curtis Fig.2", values="test") +
+             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
              ylab('effectiveness of insecticides added') +
              xlab('exposure to both insecticides') + 
-             geom_point(alpha=1, size=0.5)  
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #!!!NOW can I add a lm fitted to the 0 points abs_diff
+             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='lightgrey', lwd=0.5) +
+             #can I subtley highlight the zero points which are repeated in each plot ?
+             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +
+
+             #add text for Curtis point
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='red')
            )       
+```
+Grey triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels).
+The grey line is a linear model through these same points.
+
+## Fig. x6 Classification Tree
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5}
+
+  #code copied from sensiAnPaper1All
+
+  require('plyr')
+
+  ## is mix adaptive better than sequential ?
+  #T/F
+  resistBetterMix_ASeqBoolean <- resistPointsMix_A > resistPointsSeq
+  #convert to 0/1
+  resistBetterMix_ASeq <- plyr::aaply(resistBetterMix_ASeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix_ASeq) <- gsub('gen','betterMix_ASeq', rownames(resistBetterMix_ASeq))
+
+  #replace 0s & 1s with mixture/sequence to make tree plot clearer
+  repMix <- function(x){
+    x[x==0] <- 'sequence'
+    x[x==1] <- 'mixture'
+    x
+  }
+
+  resistBetterMix_ASeq <- repMix(resistBetterMix_ASeq)
+
+  treeInput <- rbind( treeInput, resistBetterMix_ASeq )
+
+  #transpose
+  treeInput <- t(treeInput)
+
+  #convert to a dataframe
+  treeInput <- data.frame(treeInput, stringsAsFactors = FALSE)
+
+  require(rpart.plot)
+
+  #convert predictor columns back to numeric
+  #following this above : replace 0s & 1s with mixture/sequence to make plot clearer
+  results_columns <- substr(names(treeInput),1,6) %in% 'better'
+  input_columns <- !results_columns
+  treeInput[,input_columns] <- lapply(treeInput[,input_columns],as.numeric) 
+  treeInput[,results_columns] <- lapply(treeInput[,results_columns],as.factor)
+
+  #create string with predictor names in
+  treePredictorString <- paste(colnames(treeInput)[input_columns], collapse="+")
+
+  #just doing for one strategy, sensiAnPpaer1All.Rmd does for a bunch
+  treeResponse <- rownames(resistBetterMix_ASeq)[2] #2 is 25%
+
+  tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
+  #treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
+
+  #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
+  #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
+
+  cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
+
+  treePruned <- prune(tree, cp=cpOptimal)
+
+  #to set box colours
+  cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
+
+  #adds in no. scenarios correctly and incorrectly classified
+  #prp(tree, extra=1, varlen = 10, main=treeResponse)
+  #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
+  #prp(treePruned, extra=1, varlen = 0, main=treeResponse, col=cols, border.col=cols)
+  prp(treePruned, extra=1, varlen = 0, main='', box.col=cols, under=TRUE)
 

### 2016-03-08 resistanceModelNotesAndy.Rmd
figures

@@
 
+~ changed order of strategies in legend
+
+~ created new killer figure with added effectivenesses on the y & exposure on x.
+
+** 8/3/2016 **
+
+~ add Curtis fig2 params to killer figure
+~ tricky adding to legend or labelling
+~ can I get legend for killer figure beneath ?
+~ can I derive a line for the border between when seq & mix better by fitting a line to the 0 values ?
+~ get the classification tree that matches my killer figure
+
 ### TO DO
-~ check stackoverflow for getting mean into hexbin plot
-
-~ communicate my concern to Ian that figures I've spent quite a long time on just go into supplementary material where few will see them.
+
+~ improve classification tree heading
+~ add classification tree for >20% better
+
+~ check with Ian what the best resistance threshold to focus on is. In Curtis Fig2 it is 50%. I have been using the 25% in my figs (but I could change). WHO thresholds are :
+
+~ I just had a thought. Is the effectiveness effect additive ? i.e. do 2 insecticides of effectiveness 0.7 have the same effect e.g. as one of 1 and the other of 1.4 ?
+
+~ maybe communicate my concern to Ian that figures I've spent quite a long time on just go into supplementary material where few will see them.
 Can we think of a way of getting these into a paper themselves ?
@@
 
-~ can I change order of strategies in legend ?
 ~ can I facet the smoothed scatter plots to get them to take up less space and avoid repeating legend ?

### 2016-03-08 sensiAnPaper1All.Rmd
figures

@@
   rowSums(resistBetterMix_1Seq)/ncol(resistBetterMix_1Seq)
-  #100runs
-  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
-  #                .25                   .26                   .30   
-  #1000runs
-  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
-  #               .178                  .232                  .238 
-  #new100
-  #betterMix_1Seq_cP0.1 betterMix_1Seq_cP0.25  betterMix_1Seq_cP0.5 
-  #              0.18                 0.24                 0.25 
+
   #new1000
@@
 
-  ## is mix2 better than sequential ?
+  ## is mix adaptive better than sequential ?
   #T/F
@@
   rowSums(resistBetterMix_ASeq)/ncol(resistBetterMix_ASeq)  
-  #1000runs
-  #betterMix_ASeq_cP0.1 betterMix_ASeq_cP0.25  betterMix_ASeq_cP0.5 
-  #             0.471                0.568                0.551 
-  #new100
-  #betterMix_ASeq_cP0.1 betterMix_ASeq_cP0.25  betterMix_ASeq_cP0.5 
-  #              0.58                 0.57                 0.52 
   #new1000
@@
 
-  ## is mix3 better than sequential ?
+  ## is mix2 better than sequential ?
   #T/F
@@
   rowSums(resistBetterMix_2Seq)/ncol(resistBetterMix_2Seq)
-  #100runs
-  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
-  #                  .83                   .75                   .72 
-  #1000runs
-  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
-  #                .794                  .726                  .673  
-  #new100
-  #betterMix_2Seq_cP0.1 betterMix_2Seq_cP0.25  betterMix_2Seq_cP0.5 
-  #              0.83                 0.79                 0.72 
   #new1000
@@
   #check2
-#   #what is the absolute difference between mix3 & sequential ?
+#   #what is the absolute difference between mix2 & sequential ?
 #   resistDifferenceMix_2Seq <- resistPointsMix_2 - resistPointsSeq
@@
 #   resistDifferenceMix_1I2 <- resistPointsMix_1 - resistPointsI2
-
-  ## is mix2 better than sequential ?  
-  # I might expect it to be similar to mix3
 
@@
 
-  #replace 0s & 1s with mixture/sequence to make plot clearer
+  #replace 0s & 1s with mixture/sequence to make tree plot clearer
   repMix <- function(x){

### 2016-03-09 paper1_results_figs.Rmd
figures

@@
 
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
+
   #takes a long time to plot, initial version gave me count
@@
 
+  #modify the killer plot to show where mixtures are 20% better ?
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_diff))+
+
+             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+
+             theme_bw() +
+             #add point for Curtis inputs
+             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) +
+             #add a lm fitted to the 0 points abs_diff
+             geom_smooth(data=df_same, method = "lm", se = FALSE) 
+           )
+
 ```

### 2016-03-09 paper1_results_figs_slimmed.Rmd
figures

@@
 Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
+1.
+The implication of additional operational costs of adaptive mixtures can be investigated.
+If a requirement of a 20% improvement in resistance is imposed then more runs favour sequential than mixture.
 
@@
 
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
+
     #COOLIO !! I really like this figure.
@@
 ```
+
 Grey triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels).
@@
 
-## Fig. x6 Classification Tree
+\pagebreak
+
+## Fig. x6 Modifying the previous plot to show where resistance arises over 20% slower in an adaptive mixture strategy versus sequential.
+
+1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the adaptive mixture under these constraints.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=6}
+
+  #modify the killer plot to show where mixtures are 20% better ?
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_diff))+
+
+             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+
+             theme_bw() +
+             #add point for Curtis inputs
+             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) +
+             #add a lm fitted to the 0 points abs_diff
+             geom_smooth(data=df_same, method = "lm", se = FALSE) 
+           )
+```
+
+## Fig.
+x7 Classification Trees for whether resistance arises slower with an adaptive mixture or sequential strategy.
+The upper plot shows when mixture is slower than sequential, the lower plot is for when the mixture is more than 20% slower.
+
 ```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5}
@@
 
+  #20% better
+  resistBetterMix_ASeqBoolean20 <- resistPointsMix_A > (1.2 * resistPointsSeq)
+  #convert to 0/1
+  resistBetterMix_ASeq20 <- plyr::aaply(resistBetterMix_ASeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  #rename rows ready for binding on to inputs
+  rownames(resistBetterMix_ASeq20) <- gsub('gen','betterMix_ASeq20', rownames(resistBetterMix_ASeq20))
+
   #replace 0s & 1s with mixture/sequence to make tree plot clearer
@@
   resistBetterMix_ASeq <- repMix(resistBetterMix_ASeq)
+  resistBetterMix_ASeq20 <- repMix(resistBetterMix_ASeq20)
 
-  treeInput <- rbind( treeInput, resistBetterMix_ASeq )
+  treeInput <- rbind( treeInput, resistBetterMix_ASeq, resistBetterMix_ASeq20 )
 
@@
 
-  #just doing for one strategy, sensiAnPpaer1All.Rmd does for a bunch
-  treeResponse <- rownames(resistBetterMix_ASeq)[2] #2 is 25%
+  #just doing for a few strategies, sensiAnPpaer1All.Rmd does for a bunch
+  treeResponses <- c( rownames(resistBetterMix_ASeq)[2], #2 is 25%
+                      rownames(resistBetterMix_ASeq20)[2]
+                    )
 
+  #to do trees & plots for all response variables
+  for( treeResponse in treeResponses )
+  {
     tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
-  #treePruned <- prune(tree, cp=tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"])
-
     #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
@@
     #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
-  #prp(treePruned, extra=1, varlen = 0, main=treeResponse, col=cols, border.col=cols)
+    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
     prp(treePruned, extra=1, varlen = 0, main='', box.col=cols, under=TRUE)
+    }
 

### 2016-03-09 resistanceModelNotesAndy.Rmd
figures

@@
 
+** 9/3/2016 **
+
+~ creating 20% tree & plot
+~ got the 0 line onto the first facet of the 20% plot
+
 ### TO DO
 
-~ improve classification tree heading
-~ add classification tree for >20% better
+reread Curtis to check the below ...
+
+From Ian
+~ I think the Curtis result (the one that said sequential was better) assumed effectiveness of both =1 and that exposure =0.9. I think that fits right at the top RH corner of the figure.
 

### 2016-03-11 paper1_results_figs.Rmd
tweak figs

--- 2016-03-08 paper1_results_figs.Rmd
@@
 
-  df_diff <- cbind(resistPointsMix_A_T[-indices1], tmp)
+  dif_mixA_seq <- cbind(resistPointsMix_A_T[-indices1], tmp)
+  dif_mix2_seq <- cbind(resistPointsMix_2_T[-indices1], tmp)
 
@@
   #BEWARE would need to increase 1000 if I increase max generations in the runs
-  didntReachThresh <- which( df_diff$gen_cP0.25 > 500 | df_diff$gen_cP0.25seq > 500 )
+  didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )
+  #subset by runs not making threshold
+  dif_mixA_seq <- dif_mixA_seq[-didntReachThresh,]
 
+  dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
+
+  #for dif_mix2_seq
+  didntReachThresh <- which( dif_mix2_seq$gen_cP0.25 > 500 | dif_mix2_seq$gen_cP0.25seq > 500 )
   #subset by runs not making threshold
-  df_diff <- df_diff[-didntReachThresh,]
+  dif_mix2_seq <- dif_mix2_seq[-didntReachThresh,]
 
-  #df_diff$mixA_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
-  df_diff["mixA_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
-
-  #TODO : later do PRCC on df_diff["mixA_minus_seq0.25"]
+  dif_mix2_seq["mix2_minus_seq0.25"] <- dif_mix2_seq["gen_cP0.25"] - dif_mix2_seq["gen_cP0.25seq"]
 
@@
 
-  ggSubset <- df_diff
+  ggSubset <- dif_mixA_seq
 
@@
 
-  ggSubset <- df_diff
+  ##BEWARE toogle this to look at plots for mixboth vs seq
+  ggSubset <- dif_mixA_seq
+  #ggSubset <- dif_mix2_seq
+
   #if mix-seq >1 resistance took longer for mix so mix is better
@@
   ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
-  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)
+  ggSubset$abs_mixA_minus_seq0.25 <- abs(ggSubset$mixA_minus_seq0.25)
 
@@
   #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
+  df_same <- ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
   #set these to 1 (they were 0)
@@
   df_same$seq_better_mixA0.25 <- 1
-  #rbind them back onto other dataset
+  #rbind them back onto other dataset, probably not needed ??
   ggSubset <- rbind( ggSubset, df_same)
 
-  #takes a long time to plot, initial version gave me count
-  #I want it to plot the mean
-  #check stackoverflow ..
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
 
-    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
-             geom_hex()
-             #points not wanted if 10000
-             #geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             #geom_smooth(linetype='dashed',size=1, colour="red") +
-             #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             #labs(title = i) +
-             #geom_hline(yintercept = 0, colour="white")
-           )
+    #failed hex plot just gave me counts
+    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+    #          geom_hex()
+    #        )
 
@@
     #failed attempt at getting Curtis point in legend
-    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25, size="Curtis"))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
 
@@
 
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
 
@@
              geom_point(alpha=1, size=0.5) + 
-             #!!!NOW can I add a lm fitted to the 0 points abs_diff
-             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE) +
+             #!!!NOW can I add a lm fitted to the 0 points mixA_minus_seq0.25
+             geom_smooth(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], method = "lm", se = FALSE) +
              #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
+             geom_point(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
              #add text for Curtis point
@@
   # doing the lm outside of ggplot    
-df_same = ggSubset[which(ggSubset$abs_diff==0),]
+  df_same = ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
   x1=df_same$exposure 
@@
   y2 <- 0.6*x2+0.98
-plot(x2,y2)
+  #plot(x2,y2)
+
+  #modify the killer plot to show where mixtures are 20% better ?
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
+
+             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+
+             theme_bw() +
+             #add point for Curtis inputs
+             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) +
+             #add a lm fitted to the 0 points
+             geom_smooth(data=df_same, method = "lm", se = FALSE) 
+           )
+
+```
+
+# repeating killer plot for mixtures both versus sequential
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
+  #try repeating the killer plot for mixtures both versus sequential
+
+  ##BEWARE toogle this to look at plots for mixboth vs seq
+  #ggSubset <- dif_mixA_seq
+  ggSubset <- dif_mix2_seq
+
+  #if mix-seq >1 resistance took longer for mix so mix is better
+  ggSubset$mix2_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 > 0, 1, 0 )
+  ggSubset$seq_better_mix20.25 <- ifelse( ggSubset$mix2_minus_seq0.25 < 0, 1, 0 )
+  ggSubset$abs_mix2_minus_seq0.25 <- abs(ggSubset$mix2_minus_seq0.25)
+
+  #tricky
+  #be a bit careful what to do with zeros where results are same for both strategies
+  #I use these better vars to facet the plots
+  #may want the zeroes in both facets
+  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
+  df_same <- ggSubset[which(ggSubset$mix2_minus_seq0.25==0),]
+  #set these to 1 (they were 0)
+  df_same$mix2_better_seq0.25 <- 1
+  df_same$seq_better_mix20.25 <- 1
+  #rbind them back onto other dataset, probably not needed ??
+  ggSubset <- rbind( ggSubset, df_same)
+
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mix2_not20pc_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
+
+    #i think adding the effectivenesses may give an even better plot
+    #and makes it easier for the reader to interpret
+    #aha! yes that's even better
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.25))+
+
+             geom_point(alpha=0.2)  
+           )
+
+    #facetted 0/1, labelled mix better, seq better
+    #if I get this sorted I could try to have +ve & -ve on the same scale
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.25))+
+             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
+
+             theme(legend.position="none") +
+             geom_point(alpha=1)  
+           )
+
+    #COOLIO !! I really like this figure.
+    library(viridis)
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+
+             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
+
+             #YlGnBu looks really good except that the yellow points are very faint
+             #scale_colour_distiller(palette='YlGnBu', direction=1) +
+             #option="plasma" looks good too
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
+             theme_bw() +
+             #theme_minimal() +
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             #failing to add legend for curtis point
+             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
+             #scale_size_manual(name = "Curtis Fig.2", values="test") +
+             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #add text for Curtis point
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
+           )
+
+  #modify the killer plot to show where mixtures are 20% better ?
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
+
+             facet_grid( ~ mix2_not20pc_better_seq0.25, labeller=as_labeller( c('0'='mix\nresistance >20% slower','1'='mix\nresistance <20% slower'))) +
+
+             theme_bw() +
+             #add point for Curtis inputs
+             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) +
+             #add a lm fitted to the 0 points
+             geom_smooth(data=df_same, method = "lm", se = FALSE) 
+           )
 
@@
 
-    x <- df_diff[,1:num_inputs]
-    y <- df_diff["mixA_minus_seq0.25"]
+    x <- dif_mixA_seq[,1:num_inputs]
+    y <- dif_mixA_seq["mixA_minus_seq0.25"]
 

### 2016-03-11 paper1_results_figs_slimmed.Rmd
tweak figs

@@
 
-  df_diff <- cbind(resistPointsMix_A_T[-indices1], tmp)
+  dif_mixA_seq <- cbind(resistPointsMix_A_T[-indices1], tmp)
 
@@
   #BEWARE would need to increase 1000 if I increase max generations in the runs
-  didntReachThresh <- which( df_diff$gen_cP0.25 > 500 | df_diff$gen_cP0.25seq > 500 )
+  didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )
 
   #subset by runs not making threshold
-  df_diff <- df_diff[-didntReachThresh,]
+  dif_mixA_seq <- dif_mixA_seq[-didntReachThresh,]
 
-  #df_diff$mixA_minus_seq0.25 <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
-  df_diff["mixA_minus_seq0.25"] <- df_diff["gen_cP0.25"] - df_diff["gen_cP0.25seq"]
+  #dif_mixA_seq$mixA_minus_seq0.25 <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
+  dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
 
-  #TODO : later do PRCC on df_diff["mixA_minus_seq0.25"]
+  #TODO : later do PRCC on dif_mixA_seq["mixA_minus_seq0.25"]
 
@@
 
-  ggSubset <- df_diff
+  ggSubset <- dif_mixA_seq
   #if mix-seq >1 resistance took longer for mix so mix is better

### 2016-03-11 resistanceModelNotesAndy.Rmd
tweak figs

@@
 msg from Ian :
-
 Aha! I've understood what he means by setting correct_deploy to 0, to study mosaics.
-
 ~ changed order of strategies in legend
-
 ~ created new killer figure with added effectivenesses on the y & exposure on x.
@@
 ** 8/3/2016 **
-
 ~ add Curtis fig2 params to killer figure
@@
 ** 9/3/2016 **
-
 ~ creating 20% tree & plot
@@
 
+** 10/3/2016 **
+~ looking at the manuscript
+
+I think the manuscript can be about describing development and results of a model that is a more generic version of Curtis. But we do not need to make the model useable by others at this stage.
+
+** 11/3/2016 **
+manuscript
+
+current wordcount 9,300
+
+Details from Plos Computational Biol :
+Software Submissions
+PLOS Computational Biology publishes articles describing outstanding open source software of exceptional importance that has been shown to provide new biological insights, either as a part of the software article, or published elsewhere.
+
+The software must already be widely adopted, or have the promise of wide adoption by a broad community of users.
+Enhancements to existing published open source software will only be considered if those enhancements bring exceptional new capabilities.
+The software must be downloadable anonymously in source code form and licensed under an Open Source Initiative (OSI) compliant license.
+The source code must be accompanied with documentation on building and installing the software from source, as well as for using the software, including instructions on how a user can test the software on supplied test data.
+Software articles require a presubmission inquiry that includes explanations on how the above criteria are met.
+
+Format
+Articles should be concise (less than 3500 words, not including supplementary material)
+
+Sentence on how the decisoin tree level decided.
+
+A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror
+xerror is the cross-validation error
+
+To avoid overfitting the number of levels was restricted to the minimum where the relative error plus the standard error was less than the cross-validation error.
+
+added mix both to paper1_results_figs.Rmd
+repeated all figs for 50% resistance in paper1_results_figs_slimmed_50.Rmd
+
 ### TO DO
+
+In fig x1 sequential & adaptive mixture look very similar indeed ???
+
+Could I write a very short paper describing the shiny app that allows users to take the curtis figure & modify the code ?
+Where might it go ?
+It wouldn't need any of the sensitivity analysis etc.
+?
+It could cite the paper with the snesitivity analysis.
+
+Repeat my killer fig for mixture both (because of slight uncertainty in the adaptive methodology)?
+
+Tables 3 & 4 in the manuscript- can we do these as a figure ?
+These address Curtis statement
+
+Do we want to go with a 50% threshold across all analyses to keep clearer ?
+
+An influential assertion is that mixtures may be counter-productive (made in the abstract and  repeated on page 261) where Curtis refers to his Example (vi) on his Table 1.
+Unfortunately, Curtis Figure 1 only shows the spread of resistance using a mixture and does not quantify the extent to which it is inferior to sequential use.
+However, the calibration used to replicate Curtis Figure 1 allows us to extend the one-generation argument presented in Curtis Table 1(vi) over the numerous generations required to simulate the spread of resistance (see later discussion of our Table 3).
 
@@
 ~ check with Ian what the best resistance threshold to focus on is. In Curtis Fig2 it is 50%. I have been using the 25% in my figs (but I could change). WHO thresholds are :
+
+Potential for me to look at in a 2nd paper :
+Differential exposure of males & females. (I have shown exposure is important so this could have a big effect).
 
@@
 
-~ Ian wants trees. Maybe just do one tree of resistance0.25 on whether seq or maxAdapt is better.
-
-adaptive & sequential
-20% longer
-
 ~ start creating a proposed slimmed down figs doc (maybe as a word doc using pandoc) proposed for the paper
@@
 ~~ see paper1_results_figs_slimmed.Rmd
-
-~ I could try a 2D plot with x=exposure, y=effectiveness & hexbin & colour cells according to proportion of runs where mix or seq is better.
-
-~~ seq is sometimes better than mix both. Where does that occur. The above plot might even help to answer that.
 
@@
 
-## might be good to facet by mixture sequence better and just have points of x=exposure, y=effectiveness_insecticide1
-
 ~ can I get plots to show just a subset of points ?
@@
 
-### My summary of results & thoughts
-
-Resistance to two insecticides arises more slowly if they are used in combination rather than sequentially. See violin plots.
-
-~ I'd like to show prcc on an axis from 0-1 with diff sysmbols for neg & pos. We care about which are most imp. params, not whether they are pos or neg.
-
-~ If I would like to get prcc plots in a similar format to my existing plots. (but may not be space for axis labels)
-~ get data in form : 
-input1, input2, inputN, strategy 
-prcc vals,,             indecticide1
-
 # would it be useful for Ian for me to create an interactive shiny app to view the 1000 sim data. E.g. could just have the ggplot code with options to set the x value & colour by different inputs.
@@
 It is a 2 locus model. One locus for each insecticide. At each locus there are 2 alleles. This gives a total of 9 genotypes if the position of the alleles is ignored.
-
-Main question is : Tree1 is mixture better (0/1), Tree2 is mixture better by >20%.
 

### 2016-03-11 paper1_results_figs_slimmed_50.Rmd
create copy of figs that produces 50% figs

--- 2016-03-11 paper1_results_figs_slimmed.Rmd
@@
 ---
-title: "paper1_results_figs_slimmed"
+title: "paper1_results_figs_slimmed_50"
 output: pdf_document
@@
 
-## Main Figures for paper 1 : version 3 - in progress
-### This file is automagically generated by R from paper1_results_figs_slimmed.Rmd
+## TRIAL VERSION OF ALL FIGS AT 50% resistance threshold
+
+## Main Figures for paper 1 : version 4 - in progress
+### This file is automagically generated by R from paper1_results_figs_slimmed_50.Rmd
 
@@
   #columns to remove from first
-  indices1 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
+  indices1 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.1","gen_cP0.25","strategy"))
   #columns to add to 2nd  
-  indices2 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.25"))  
+  indices2 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.5"))  
   #rename column in 2nd
   tmp <- resistPointsSeq_T[indices2]
-  names(tmp) <- "gen_cP0.25seq"
+  names(tmp) <- "gen_cP0.5seq"
 
@@
   #BEWARE would need to increase 1000 if I increase max generations in the runs
-  didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )
+  didntReachThresh <- which( dif_mixA_seq$gen_cP0.5 > 500 | dif_mixA_seq$gen_cP0.5seq > 500 )
 
@@
 
-  #dif_mixA_seq$mixA_minus_seq0.25 <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
-  dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
+  #dif_mixA_seq$mixA_minus_seq0.5 <- dif_mixA_seq["gen_cP0.5"] - dif_mixA_seq["gen_cP0.5seq"]
+  dif_mixA_seq["mixA_minus_seq0.5"] <- dif_mixA_seq["gen_cP0.5"] - dif_mixA_seq["gen_cP0.5seq"]
 
-  #TODO : later do PRCC on dif_mixA_seq["mixA_minus_seq0.25"]
+  #TODO : later do PRCC on dif_mixA_seq["mixA_minus_seq0.5"]
 
@@
 #names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
-names_results <- c("time_to_resistance0.25")
+names_results <- c("time_to_resistance0.5")
 
@@
 
-## Fig x2 PRCC analysis for time-to-resistance 0.25 for sequential and mixture strategies
+## Fig x2 PRCC analysis for time-to-resistance 0.5 for sequential and mixture strategies
 
@@
     x <- by_strategy[,1:num_inputs]
-    y <- by_strategy['time_to_resistance0.25']
+    y <- by_strategy['time_to_resistance0.5']
 
@@
     #          #ggtitle(paste(strategy,"PRCC")) +
-    #          ggtitle(paste(strategy,"time-to-resistance 0.25")) +
+    #          ggtitle(paste(strategy,"time-to-resistance 0.5")) +
     #          xlab(NULL)
@@
               "dominance_allele1","selection_coef_allele1","maleExposureProp")
-  out_cols <- c("time_to_resistance0.25","strategy")
+  out_cols <- c("time_to_resistance0.5","strategy")
   ggSubset <- ggSubset[, c(in_cols,out_cols) ]
@@
 
-    y <- 'time_to_resistance0.25'
+    y <- 'time_to_resistance0.5'
 
@@
 
-## Fig. x4 PRCC difference in time-to-resistance 0.25 between sequential and adaptive mixture strategies
+## Fig. x4 PRCC difference in time-to-resistance 0.5 between sequential and adaptive mixture strategies
 
@@
     x <- dif_mixA_seq[,1:num_inputs]
-    y <- dif_mixA_seq["mixA_minus_seq0.25"]
+    y <- dif_mixA_seq["mixA_minus_seq0.5"]
 
@@
   #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
-  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
-  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.25)
+  ggSubset$mixA_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 > 0, 1, 0 )
+  ggSubset$seq_better_mixA0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 < 0, 1, 0 )
+  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.5)
 
@@
   #set these to 1 (they were 0)
-  df_same$mixA_better_seq0.25 <- 1
-  df_same$seq_better_mixA0.25 <- 1
+  df_same$mixA_better_seq0.5 <- 1
+  df_same$seq_better_mixA0.5 <- 1
   #rbind them back onto other dataset
@@
   #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
+  ggSubset$mixA_not20pc_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
 
@@
 
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance slower','1'='sequential\nresistance slower'))) +
+             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance slower','1'='sequential\nresistance slower'))) +
 
@@
 
-             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
 

### 2016-03-16 paper1_results_figs_slimmed_50.Rmd
add some mix2 plots and calc of the lm line

@@
   dif_mixA_seq <- cbind(resistPointsMix_A_T[-indices1], tmp)
+  dif_mix2_seq <- cbind(resistPointsMix_2_T[-indices1], tmp)
 
@@
   didntReachThresh <- which( dif_mixA_seq$gen_cP0.5 > 500 | dif_mixA_seq$gen_cP0.5seq > 500 )
-
   #subset by runs not making threshold
@@
 
-  #dif_mixA_seq$mixA_minus_seq0.5 <- dif_mixA_seq["gen_cP0.5"] - dif_mixA_seq["gen_cP0.5seq"]
   dif_mixA_seq["mixA_minus_seq0.5"] <- dif_mixA_seq["gen_cP0.5"] - dif_mixA_seq["gen_cP0.5seq"]
 
-  #TODO : later do PRCC on dif_mixA_seq["mixA_minus_seq0.5"]
+  #for dif_mix2_seq
+  didntReachThresh <- which( dif_mix2_seq$gen_cP0.5 > 500 | dif_mix2_seq$gen_cP0.5seq > 500 )
+  #subset by runs not making threshold
+  dif_mix2_seq <- dif_mix2_seq[-didntReachThresh,]
+
+  dif_mix2_seq["mix2_minus_seq0.5"] <- dif_mix2_seq["gen_cP0.5"] - dif_mix2_seq["gen_cP0.5seq"]
 
@@
       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
-      #scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Sole use", "Mix either"="1st in mix","Sequential"="2nd in sequence", "Mix adaptive"="2nd in mix\nswitch to sole", "Mix both"="2nd in mix" ) )
-      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Sole use", "Mix either"="1st in mix","Sequential"="2nd in sequence", "Mix adaptive"="adaptive mix", "Mix both"="2nd in mix" ) )
+      ylab('time to resistance (gens)') +
+      scale_x_discrete("", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Seq (1st)", "Mix either"="Mix (1st)","Sequential"="Seq (2nd)", "Mix adaptive"="adaptive", "Mix both"="Mix (2nd)" ) )
 
@@
 
-    #draw_plot(p,x=0,y=0,width=.5,height = .5)
-
-    # print( ggdraw() + #panel_border(remove = TRUE) +
-    #          draw_plot( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
-    # #print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
-    #          geom_point(shape=1, colour='red') +
-    #          theme(axis.text.x = axis_text_x) +
-    #          #theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
-    #          geom_hline(yintercept = 0, linetype=3) +
-    #          ylim(-1,1) +
-    #          #ggtitle(paste(strategy,"PRCC")) +
-    #          ggtitle(paste(strategy,"time-to-resistance 0.5")) +
-    #          xlab(NULL)
-    #       , height= height))
-
     plotlist[[strategy_num]] <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) +     
@@
   #subset which output plots to show in the main paper
-  in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
-              "dominance_allele1","selection_coef_allele1","maleExposureProp")
+  # in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
+  #             "dominance_allele1","selection_coef_allele1","maleExposureProp")
+  in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")
+
   out_cols <- c("time_to_resistance0.5","strategy")
@@
 
-    x <- dif_mixA_seq[,1:num_inputs]
-    y <- dif_mixA_seq["mixA_minus_seq0.5"]
+    #mixture2
+    x <- dif_mix2_seq[,1:num_inputs]
+    y <- dif_mix2_seq["mix2_minus_seq0.5"]
 
     pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
 
@@
              ylim(-1,1) +
-             #ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
+             xlab(NULL)
+          )
+
+    ## repeat code for adaptive
+    x <- dif_mixA_seq[,1:num_inputs]
+    y <- dif_mixA_seq["mixA_minus_seq0.5"]
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #results are here I can probably rbind them together into a df that I can ggplot
+    #pcc_res$PRCC
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
+
+    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             geom_hline(yintercept = 0, linetype=3) +
+             #theme_bw() +
+             ylim(-1,1) +
              xlab(NULL)
@@
            )
+
+  # doing the lm outside of ggplot    
+  #df_same = ggSubset[which(ggSubset$mixA_minus_seq0.5==0),]
+  x1=df_same$exposure 
+  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
+
+  l_mod <- lm(y1 ~ x1)
+  l_mod$coefficients
+  #(Intercept)           x 
+  #  0.9790948   0.6095132 
+  #l_mod$coefficients
+  #(Intercept)          x1 
+  #1.019330    0.560508
+
+  #approximation
+  #x2 <- seq(0,1,0.1)
+  #y2 <- 0.6*x2+0.98
+  #plot(x2,y2)
+
 ```
@@
 
+    #add fig label
+    if (treeResponse == treeResponses[1]) main <- "A)" 
+    else  main <- "B)"
+
     #adds in no. scenarios correctly and incorrectly classified
@@
     #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
-    prp(treePruned, extra=1, varlen = 0, main='', box.col=cols, under=TRUE)
+    prp(treePruned, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE)
     }

### 2016-03-16 paper1_results_figs_slimmed_50.Rmd
updating paper figures

@@
 
-## Main Figures for paper 1 : version 4 - in progress
+## Main Figures for paper 1 : version 5 - in progress
 ### This file is automagically generated by R from paper1_results_figs_slimmed_50.Rmd
@@
              #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='black', shape=3) +
              #failing to add legend for curtis point
@@
              #!!!NOW can I add a lm fitted to the 0 points abs_diff
-             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='lightgrey', lwd=0.5) +
+             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
              #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +
+             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
 
              #add text for Curtis point
-             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='red')
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
            )
@@
   # doing the lm outside of ggplot    
-  #df_same = ggSubset[which(ggSubset$mixA_minus_seq0.5==0),]
   x1=df_same$exposure 
@@
   l_mod$coefficients
-  #(Intercept)           x 
-  #  0.9790948   0.6095132 
-  #l_mod$coefficients
   #(Intercept)          x1 
@@
 
+  df_20pc <- ggSubset[which(ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq > 0.19 & 
+                            ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq < 0.21  ),]
+
+  #remove column it is faceted on to get line in both facets
+  col_indices <- colnames(df_20pc) != 'mixA_not20pc_better_seq0.5'
+  df_20pc <- df_20pc[,col_indices]
+
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_diff))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
 
@@
              theme_bw() +
+             theme(legend.position = "bottom") +
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
              #add point for Curtis inputs
@@
              #add a lm fitted to the 0 points abs_diff
-             geom_smooth(data=df_same, method = "lm", se = FALSE) 
+             #geom_smooth(data=df_same, method = "lm", se = FALSE)
+             #todo change to where exactly 20% diff
+             geom_smooth(data=df_20pc, method = "lm", se = FALSE,colour='red', lwd=0.5)+
+             #can I subtley highlight the zero points which are repeated in each plot ?
+             geom_point(data=df_20pc, size=0.8, colour='red', shape=2) #, fill='white') 
            )
+
+  # doing the lm outside of ggplot    
+  x1=df_20pc$exposure 
+  y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
+
+  l_mod <- lm(y1 ~ x1)
+  l_mod$coefficients
+# (Intercept)          x1 
+#   1.3914547   0.2904911
+
 ```

### 2016-03-16 resistanceModelNotesAndy.Rmd
notes

@@
 
-added mix both to paper1_results_figs.Rmd
+added mix both to paper1_results_figs.Rmd (because of my slight uncertainty in mix adaptive method)
 repeated all figs for 50% resistance in paper1_results_figs_slimmed_50.Rmd
 
+** 14/3/16 **
+
+Curtis contentious statement p261 :
+"Example vishows the, at first sight unexpected, point that the use of a mixture where the initial gene frequencies are unequal leads to a more rapid increase in the rarer of the genes."
+
+vi params ?
+exposure 0.9
+strt freq loc1 0.01
+strt freq loc2 0.001
+
+modify fig2 by setting freq1 from 0.01 to 0.001
+
+runcurtis_f2( P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 , correctMixDeployProp = 1 )
+
+when I changejust freq loc1 to 0.001 I get an error :
+Warning in min(which((combmat[, r1col]) > max(criticalPoints))) :
+  no non-missing arguments to min; returning Inf
+
+Trying to work it out with this run :
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.003 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 , correctMixDeployProp = 1 )
+
+in plotcurtis_f2_generic problem is here :
+
+  cutoff_i1_mix <- min(which((combmat[,r1col])>max(criticalPoints)))
+  cutoff_i2_mix <- min(which((combmat[,r2col])>max(criticalPoints)))
+
+  if (cutoff_i1_mix == Inf | cutoff_i2_mix == Inf)
+  {
+    warning("critical point not reached for i1 or i2: cutoff_i1_mix=",cutoff_i1_mix," cutoff_i2_mix=",cutoff_i2_mix)
+    #don't add the combined(adaptive) strategy to the graph
+    addCombinedStrategy <- FALSE
+  }
+
+in plotcurtis_f2_generic this calculation at the start seems to produce diff results than what they appear in the plot :
+  maxGensMix <- findResistancePoints(combmat, locus='both', criticalPoints = criticalPoints)
+
+maxGensMix
+           [,1]
+gen_cP0.1   294
+gen_cP0.25  999
+gen_cP0.5   999
+
+problem partly caused by max_gen default in run_curtis_f2 being 300
+
+runcurtis_f2 <- function( max_gen = 300,
+whereas in sensiAnPaperPart max_gen is set to 500
+
+probably I should set max_gen in run_curtis_f2 to 500 too so it is the same.
+
+yes, now this works, but P_2 = 0.001 still fails to reach threshold
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.002 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 , correctMixDeployProp = 1 )
+
+** 15/3/16 ** on train to liverpool
+
+Looking at the form of the results from modified curtis fig2 plots :
+~ in mixtures resistance in the final insecticide only rises after the 'first' one has become ineffective ~50%
+
+~ in curtis scenario reducing start freq of I2 from 0.01 to 0.002 makes it take longer but doesn't really alter the form of curves
+~ but then changing effectiveness of I2 from 1 to 0.95 makes a big difference.
+~ aha!
+seems that an effectiveness of 1 in one insecticide stops resistance rising in the other one until ~50% resistance.
+Reducing effectiveness to even 0.95 (or just 0.99) seems to remove this effect.
+At effectiveness I2 0.99 resistance to I1 rises slowly until 50% is reached then it accelerates.
+
+~ increasing dominance from 0.0016 to 0.02 doesn't radically alter curtis2 pattern
+
+** 16/3/16 ** meeting Ian in liverpool
+
+paper todo
+~ rename fig x1 y axis to "time in generations" or time (generations)
+~ rename y axes in fig x3 "time in generations" or time (generations)
+~ rename maleExposureProp to male_exposure
+~ rename correctMixDeployProp to correct_mix_deploy
+~ set variable names to be the same as above in table 5 of manuscript
+~ check again that correctMixDeploy is doing what we expect.
+Bit strange that in the plot against time-to-r the curves are completely level when in my curtis simulation it shows that decreasing the correct deploy does decrease time to resistance e.g.
+for mix2 from
+done ~ change fig x3 to just exposure, effectiveness and dominance
+~ add diff between seq and standard mixture to fig x4
+~ Ian will send me the params to rerun the line from table vi (which is just for 1 generation)
+~ find the equation line of zero difference for the 0.5 threshold (and maybe for mix2 vs seq too ? )
+~
+
+~ for the 20% diff plot add the line that exactly 20% diff and keep colour scheme from killer plot
+~ see df_20pc
+~ how to make sure that the same points apear in each plot ?? maybe remove the facet column.
+
+Remember that effectiveness is proportion of the SS genotypes that die when exposed.
+
+Ian said, science is unsure what the effectiveness rates ofr insecticides are under field conditions.
+all sensitive mosquitos ar expected to die in a bottle assay.
+BUt under filed conditions where they briefly land on a net this may not be the case.
+experimental huts.
+
+IVCC are apparently asking if we could model 3 insecticides.
+
+Ian keen to keep sex-linkage & niches in. Say we have developed a general model and say we will explore more behaviour in future ...
+
 ### TO DO
+
+I've done below for the mix-both plot, repeat it for mix-adaptive.
+Ian suggested adding in the maleProp effect to the x axis to see if that improves the split.
+To get the exposure of both M&F : x=exposure+exposure*maleExposureProp
+
+Qs for Ian
+~ is effectiveness of 1 possible/likely ? The pattern in the curtis plot relies quite heavily on it. Reducing it to just 0.98 changes pattern considerably. 
+~ are we confident enough about the mix adaptive strategy to focus the paper on it ? we could look at standard mix instead ?
+
+~ ask Ian where could I submit a paper explaining this resistance thing in simpler terms for dummies.
+If I want to get into anything sciency a first author paper would really help me out.
+It could concentrate just on the mechanism of how the model works rather than on the sensitivity outputs.
+It could have plot combinations, showing how the curves change in response to 1 param changing.
+I could start writing it now.
+It could go to ecological modelling or similar epidemiology journal (maybe ask Jo Lines).
+
+~ would be a good learning tool to have a UI similar to my current curtis fig2 one, but starting with the params equal for each insecticide
+
+Curtis contentious statement p261 :
+"... the use of a mixture where the initial gene frequencies are unequal leads to a more rapid increase in the rarer of the genes."
+Not supported by our results. We need to nail which results to use to show this.
+
+bumped into colin bannister on the train to london.
+he said he really ejoyed my talk.
+he said he would talk to me sometime about keeping the work going in a consultancy, difficult to do.
+He was on his way to london for a shellfish assessment.
 
@@
 It could cite the paper with the snesitivity analysis.
-
-Repeat my killer fig for mixture both (because of slight uncertainty in the adaptive methodology)?
 

### 2016-03-16 paper1_results_figs.Rmd
tweak

@@
 
-    #COOLIO !! I really like this figure.
+    #killer plot mix2 vs seq
     library(viridis)
@@
              xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #add text for Curtis point
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
+           )
+
+    #modify killer plot to add in male exposure prop on x axis
+    library(viridis)
+    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+
+             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
+             theme_bw() +
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides for both sexes') + 
              #main points for the plot

### 2016-03-16 paper1_results_figs_slimmed.Rmd
tweak

@@
              geom_hline(yintercept = 0, linetype=3) +
-             #theme_bw() +
+             #theme_bw() + # to add in gridline but messes up other bits of plot
              #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +

### 2016-03-16 paper2_resistance_mechanisms_mixtures.Rmd
start paper2

---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South, Ian Hastings and Bethany Levick"
date: "`r Sys.Date()`"
---

### version 1

# The development of insecticide resistance with use of insecticide mixtures and sequences : a simple theoretical exploration.

#### (This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
#### Target : Malaria journal

## Abstract

This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence/rotation and mixtures. It is aimed at a readership with limited experience of population genetics and resistance. It uses a model representing two independent loci coding for resistance to two insecticides. It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective adavantage of resistance and the dominance of the alleles. There is uncertainty about the values of these parameters in field situations. We demonstrate how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies. We describe an R package and user-interface that allows further exploration of the development of resistance.

## Introduction

Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.

Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency. We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.

The development of insecticide resistance is a population genetic process.

The development of resistance is expected to be influenced by :

1. starting frequency of resistance alleles
1. selection coefficient
   + selective advatange of resistance in the presence of the insecticide
1. dominance
   + dominance of the resistance allele relative to the susceptible allele
1. exposure
   + proportion of the insect population that are exposed to the insecticide
1. effectiveness
   + proportion of exposed susceptible insects that are killed by the insecticide

We use a 2-locus population genetic model of insecticide resistance (Hastings et al. in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies. Hastings et al. (in press) describe the technical details of the model and explore large parameter ranges. In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the devlopment of resistance.

## Methods

*Maybe start by exploring the influence of the parameters on sole use ?*

## Results

```{r,echo=FALSE,results='hide'}
  library(resistance)
```
### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

#### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

#### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
```{r, eval=TRUE, echo=FALSE, fig.height=4}
runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
```

## Todo

1. look more at GPIRM
1. mention Curtis
1. do I want y axis of resistance frequency plots to be on a log scale ?
1. can I develop explanatory diagrams to describe the input parameters ?

### 2016-03-23 paper1_results_figs_slimmed_50.Rmd
adressing minor issues in check

@@
       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
-      ylab('time to resistance (gens)') +
-      scale_x_discrete("", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Seq (1st)", "Mix either"="Mix (1st)","Sequential"="Seq (2nd)", "Mix adaptive"="adaptive", "Mix both"="Mix (2nd)" ) )
+      ylab('Time to resistance (gens)') +
+      scale_x_discrete("", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Seq 1st", "Mix either"="Mix 1st","Sequential"="Seq both", "Mix adaptive"="Adaptive", "Mix both"="Mix both" ) )
 

### 2016-03-23 resistanceModelNotesAndy.Rmd
notes
adressing minor issues in check

@@
 ~ find the equation line of zero difference for the 0.5 threshold (and maybe for mix2 vs seq too ? )
-~
 
@@
 
-### TO DO
-
-I've done below for the mix-both plot, repeat it for mix-adaptive.
-Ian suggested adding in the maleProp effect to the x axis to see if that improves the split.
-To get the exposure of both M&F : x=exposure+exposure*maleExposureProp
-
-Qs for Ian
-~ is effectiveness of 1 possible/likely ? The pattern in the curtis plot relies quite heavily on it. Reducing it to just 0.98 changes pattern considerably. 
-~ are we confident enough about the mix adaptive strategy to focus the paper on it ? we could look at standard mix instead ?
-
-~ ask Ian where could I submit a paper explaining this resistance thing in simpler terms for dummies.
+~ asked Ian where could I submit a paper explaining this resistance thing in simpler terms for dummies.
 If I want to get into anything sciency a first author paper would really help me out.
@@
 It could go to ecological modelling or similar epidemiology journal (maybe ask Jo Lines).
+~~ concise
+~~ focus on mechanisms
+~~ for non expert audience
+~~ dominance, exposure
+~~ draft something
+~~ ian suggested malaria journal
+~~ paper2_resistance_mechanisms_mixtures.Rmd
+
+** 22/3/16 **
+comments back from Ian on paper1
+
+One other thing for your "to do" list.
+I have made the point that Fig 4 (your x2) analyses the data summarised in Fig 3 (your x1)..
+which helps the flow of the paper.
+But we only present PRCC for 4 of the data shown in Figure 3 so it may worthwhile doing PRCC for the missing one i.e.
+time to resistance to first insecticide in sequential use.
+I assume that would be relatively easy to do and automate in the file?
+
+Fix x1
+I did Time to resistance (gens)~ Y axis label change to  "Time (gens)"
+done ~ X axis labels as in the figure legend i.e.
+"Mix 1st"  "Mix both",  "Seq 1st" and "Seq both".
+I think "both" is better than "second" as it avoids any potential confusion between insecticide labelled 1 and 2.
+done ~ I think we decided to re-labelled "sole use" to "Seq 1st" for clarity
+done ~ You might also like to change "Adaptive mix" to just "Adaptive" to try and save some space which I think will become important later as graphs are scaled down to their published size.
+
+Fig x2..
+can we make the labels identical to those above i.e., in order  "Seq both"  "Mix 1st"  "Adaptive both "   "Mix both"....
+A reviewer will wonder why we don't do all 5 outputs described in Fig x1 i.e.
+whey we omit "Seq 1st"
+
+Fig x3...Y axis label change to "Time (gens)".
+
+Fig x5 and x6....
+Can we change the Y axis label to just "Effectiveness of insecticides"...
+drop the word "added" as it is a bit ugly, makes the label longer and we now say it the sum of effectiveness in the figure legend.
+Can we change the X axis label to just "exposure"... do we need to say it is for "both insecticides"????
+
+** 23/3/16 **
+email from Ian
+
+We extensively look at how best to deploy 2 new insecticides, but not how to deploy one new insecticide when several others are in use and there is resistance to them….
+
+This Curtis argument was not about sequential vs mixtures (i.e.
+where we start off with 2 insecticides and decide how best to deploy them).
+It is simply this questions: “If I have a single new insecticide where resistance is presumably present at a very low frequency, should I deploy it on its own (sole use), or in combination with an existing insecticide whose resistance levels are high”;
+
+Curtis defines allele B as the one encoding resistance to the new insecticide and which therefore has a lower starting frequency and asserted that, under these circumstances, deploying the new insecticide as a mixture would be counterproductive for allele B i.e.
+would drive resistance more quickly than when insecticide B was deployed alone.
+
+Can I show this by modifying the fig2 plot ?
+
+### TO DO
+
+1. add option to rescale x axis to years or months in plotcurtis_f2_generic, generation=1month
+1. tidy these notes
+
+I've done below for the mix-both plot, repeat it for mix-adaptive.
+Ian suggested adding in the maleProp effect to the x axis to see if that improves the split.
+To get the exposure of both M&F : x=exposure+exposure*maleExposureProp
+
+Qs for Ian
+~ is effectiveness of 1 possible/likely ? The pattern in the curtis plot relies quite heavily on it. Reducing it to just 0.98 changes pattern considerably. 
+~ are we confident enough about the mix adaptive strategy to focus the paper on it ? we could look at standard mix instead ?
 
@@
 
+Note that in most functions (e.g.
+runcurtis_f2) there is a single exposure param for both insecticides.
+I could relatively easily allow different exposure for each insecticide.
+It is probably already possible by passing array:a directly.
+
 Tables 3 & 4 in the manuscript- can we do these as a figure ?
@@
 if this first paper is targeted at policy people can we get a 2nd paper that is targetted at biologists and explores mechanisms ?
-
-~ can I facet the smoothed scatter plots to get them to take up less space and avoid repeating legend ?
-~~ or I may be able to use a similar plot_grid() method
-
-~~ if I wanted to facet by inputs might need to restructure the data, put inputs into a single column to replace below :
-  for(i in names_inputs)
-  {
-    print( ggplot(ggSubset, aes_string(x=i
-
-~ start creating a proposed slimmed down figs doc (maybe as a word doc using pandoc) proposed for the paper
-~~ beware that if I send it to Ian as a doc, he will insert text into it that I may not be able to find.
-~~ maybe send it as a pdf first & change to a doc later to allow Ian to copy & paste the figs.
-~~ see paper1_results_figs_slimmed.Rmd
 

### 2016-03-23 R/createArray.r

@@ -13,11 +13,12 @@
 #' 
 #' @examples 
 #' createArray( sex = c('f','m') ) 
-#' #Wloci
-#' createArray( loci=c('SS1','RS1','RR1','SS2','RS2','RR2'), niche1=c('0','a','A'), niche2=c('0','b','B') )
-#' createArray( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2') )
-#' createArray( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2'), niche1 = c('0','a','A') )
-#' createArray( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2'), niche1 = c('0','a','A'), niche2 = c('0','a','A') )
+#' 
+#' createArray( loci=c('SS1','RS1','RR1','SS2','RS2','RR2'), 
+#'              niche1=c('0','a','A'), niche2=c('0','b','B') )
+#' createArray( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), 
+#'              locus2 = c('SS2','RS2','RR2'), 
+#'              niche1 = c('0','a','A'), niche2 = c('0','a','A') )
 #' 
 #' @return matrix
 #' @export

### 2016-03-23 R/createArray2.r

@@ -5,13 +5,13 @@
 #' @param ... named dimensions to create in array e.g. dimname=c('name1','name2') 
 #' 
 #' @examples 
-#' createArray2( sex = c('f','m') ) 
-#' #Wloci
-#' createArray2( loci=c('SS1','RS1','RR1','SS2','RS2','RR2'), niche1=c('0','a','A'), niche2=c('0','b','B') )
-#' createArray2( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2') )
-#' createArray2( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2'), niche1 = c('0','a','A') )
-#' createArray2( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), locus2 = c('SS2','RS2','RR2'), niche1 = c('0','a','A'), niche2 = c('0','a','A') )
+#' createArray( sex = c('f','m') ) 
 #' 
+#' createArray( loci=c('SS1','RS1','RR1','SS2','RS2','RR2'), 
+#'              niche1=c('0','a','A'), niche2=c('0','b','B') )
+#' createArray( sex = c('f','m'), locus1 = c('SS1','RS1','RR1'), 
+#'              locus2 = c('SS2','RS2','RR2'), 
+#'              niche1 = c('0','a','A'), niche2 = c('0','a','A') )#' 
 #' @return array
 #' @export
 

### 2016-03-23 R/plotcurtis_f2_generic.r

@@ -16,9 +16,9 @@
 #' @param strategyLabels = c("seq","mix1","adapt","mix2")
 #' 
 #' @examples
-#' inputAndResultsMix <- sensiAnPaperPart( 2, insecticideUsed = 'mixture' )
-#' inputAndResultsI1 <- sensiAnPaperPart( 2, insecticideUsed = 'insecticide1' )
-#' inputAndResultsI2 <- sensiAnPaperPart( 2, insecticideUsed = 'insecticide2' )
+#' listOutMix <- sensiAnPaperPart( 2, insecticideUsed = 'mixture' )
+#' listOutI1 <- sensiAnPaperPart( 2, insecticideUsed = 'insecticide1' )
+#' listOutI2 <- sensiAnPaperPart( 2, insecticideUsed = 'insecticide2' )
 #' plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )
 #' 
 #' @return list of plot components

### 2016-03-23 R/runModel2.r

@@ -21,7 +21,9 @@
 #' #? I should probably check
 #' #setting calibration & exposures to insecticides as the same for mf
 #' #this doesn't quite seem to work yet, freq of R stays at 0
-#' input <- setInputOneScenario(calibration=1013, a.m_A0 = 0.9, a.m_00 = 0.1, a.m_AB = 0, a.f_A0 = 0.9, a.f_00 = 0.1, a.f_AB = 0)
+#' input <- setInputOneScenario(calibration=1013, 
+#'                              a.m_A0 = 0.9, a.m_00 = 0.1, a.m_AB = 0, 
+#'                              a.f_A0 = 0.9, a.f_00 = 0.1, a.f_AB = 0)
 #' #can either set calibration to 1013 as an arg to setInput or runModel2()
 #' tst <- runModel2(input)
 #' @return a list of 3 lists of one or more scenarios: results, genotype and fitness. e.g. listOut$results[1] gives a results matrix for the first scenario

### 2016-03-23 R/sensiAn1.r

@@ -9,8 +9,9 @@
 #' #inputAndResults <- sensiAn1( h.RS1_A0=c(0.1,1))
 #' #colnames(inputAndResults)
 #' #inputAndResults <- sensiAn1( h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1) )
-#' #inputAndResults <- sensiAn1(500, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
-#' @return todo
+#' #inputAndResults <- sensiAn1(500, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), 
+#' #                              s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
+#' @return dataframe
 #' @export
 #' 
 sensiAn1 <- function( nScenarios = 10,

### 2016-03-23 R/sensiAnPaperPart.r

@@ -14,9 +14,9 @@
 #'    'extended' for adding reduced male exposure & incorrect deploymnet 
 #' 
 #' @examples
-#' inputAndResultsMix <- sensiAnPaperPart( 5, insecticideUsed = 'mixture' )
-#' inputAndResultsI1 <- sensiAnPaperPart( 5, insecticideUsed = 'insecticide1' )
-#' inputAndResultsI2 <- sensiAnPaperPart( 5, insecticideUsed = 'insecticide2' )
+#' listOutMix <- sensiAnPaperPart( 5, insecticideUsed = 'mixture' )
+#' listOutI1 <- sensiAnPaperPart( 5, insecticideUsed = 'insecticide1' )
+#' listOutI2 <- sensiAnPaperPart( 5, insecticideUsed = 'insecticide2' )
 #' plotcurtis_f2_generic( listOutMix$results[[1]], listOutI2$results[[1]], listOutI1$results[[1]] )
 #' #colnames(inputAndResults)
 #' @return a list containing results and inputs

### 2016-03-23 R/sensiAnPaperPartOneAtATime.r

@@ -7,6 +7,7 @@
 #' needs to be run 3 times, once for each insecticide strategy 'insecticide1','insecticide2', & 'mixture'.
 #' 
 #' @param insecticideUsed one of 'insecticide1','insecticide2','mixture'
+#' @param nSamples number of samples
 #' 
 #' @examples
 #' inputAndResultsMix <- sensiAnPaperPartOneAtATime( insecticideUsed = 'mixture' )
@@ -16,7 +17,7 @@
 #' @return a list containing results and inputs
 #' @export
 #' 
-sensiAnPaperPartOneAtATime <- function( insecticideUsed = 'mixture' )
+sensiAnPaperPartOneAtATime <- function( insecticideUsed = 'mixture', nSamples )
 {
   #set random seed
   #if (!is.null(randomSeed)) set.seed(randomSeed)

### 2016-03-23 R/setInputSensiScenarios.r

@@ -12,7 +12,8 @@
 #' @examples
 #' setInputSensiScenarios( nScenarios=2, h.RS1_A0=c(0,1) )
 #' setInputSensiScenarios( nScenarios=3, P_1=c(0,0.5), P_2=c(0,0.5) )
-#' setInputSensiScenarios(20, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
+#' setInputSensiScenarios(20, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), 
+#'                        s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
 #' @return a matrix of input scenarios
 #' @export
 #' 

### 2016-03-23 pap1_figs_slimmed_50_unequal.Rmd
copied from paper1_figures_slimmed50.Rmd to enable looking at new unequal scenarios

--- 2016-03-23 paper1_results_figs_slimmed_50.Rmd
@@
 
-## TRIAL VERSION OF ALL FIGS AT 50% resistance threshold
+## ALL FIGS AT 50% resistance threshold
 
-## Main Figures for paper 1 : version 5 - in progress
+## Main Figures for paper 1 : version 6 - in progress
 ### This file is automagically generated by R from paper1_results_figs_slimmed_50.Rmd
@@
   ## to load previously saved runs
-#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI2_1000.rda'))
 
-  ## trying with the extended experiment _ex100
+  ## trying with the extended experiment
   experiment <- 'extended'
 
-  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
 
-  # load(file=paste0(outFolder,'listOutMix_ex2_1000.rda'))
-  # load(file=paste0(outFolder,'listOutI1_ex2_1000.rda'))
-  # load(file=paste0(outFolder,'listOutI2_ex2_1000.rda'))
+  load(file=paste0(outFolder,'listOutMix_unequal.rda'))
+  load(file=paste0(outFolder,'listOutI1_unequal.rda'))
+  load(file=paste0(outFolder,'listOutI2_unequal.rda'))
 
@@
 
+  #BEWARE eval set to fale becuase new unequal runs caused an error in this chunk
+
   ggSubset <- dif_mixA_seq
@@
 ```
+
+## Fig.
+x8 Introducing a new insecticide when there is already resistance to an existing one.
+This takes Curtis Fig 2 and sets the starting frequency for each allele in turn to a higher level to represent an 'old' insecticide to which there is resistance.
+In both cases resistance to the new insecticide develops slower when the 'old' insecticide continues to be used (mixture) than when it doesn't (sequence).
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5}
+
+#curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
+runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+#curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+```

### 2016-03-23 sensiAnPaper1All.Rmd
add an unequal option for looking at 'old' insecticide scenarios

@@
 
-  nScenarios <- 3  
+  #nScenarios <- 10  
   #nScenarios <- 100
-  #nScenarios <- 1000  
+  nScenarios <- 500  
   #nScenarios <- 10000
@@
   ## do model runs
-  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
-  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment )
-  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment )
+  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
+  # listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment )
+  # listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment )
+
+  ## 23/3/16 expt with an 'old' insecticide
+  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment, old_insecticide=TRUE )
+  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment, old_insecticide=TRUE )
+  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment, old_insecticide=TRUE )
 
   ## save results objects as rda for re-use
-  #save(listOutMix, file=paste0(outFolder,'listOutMix.rda'))
-  #save(listOutI1, file=paste0(outFolder,'listOutI1.rda'))
-  #save(listOutI2, file=paste0(outFolder,'listOutI2.rda'))
+  save(listOutMix, file=paste0(outFolder,'listOutMix_unequal.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1_unequal.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2_unequal.rda'))
 
-  save(listOutMix, file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-  save(listOutI1, file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-  save(listOutI2, file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
-
-#   save(listOutMix, file=paste0(outFolder,'listOutMix_ex2_3.rda'))
-#   save(listOutI1, file=paste0(outFolder,'listOutI1_ex2_3.rda'))
-#   save(listOutI2, file=paste0(outFolder,'listOutI2_ex2_3.rda'))
+  # save(listOutMix, file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  # save(listOutI1, file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  # save(listOutI2, file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
 

### 2016-03-23 R/sensiAnPaperPart.r

@@ -25,7 +25,8 @@
 sensiAnPaperPart <- function( nScenarios = 10,
                            insecticideUsed = 'mixture',
                            randomSeed = 1,
-                           experiment = "curtis1")
+                           experiment = "curtis1",
+                           old_insecticide = FALSE )
 {
   #set random seed
   if (!is.null(randomSeed)) set.seed(randomSeed)
@@ -53,6 +54,9 @@ sensiAnPaperPart <- function( nScenarios = 10,
     #select ‘x’ from uniform -1 to -4 and set to 10^x to give equal weight to each log interval.
     P_1 <- 10^-(runif(1, min=1, max=4))
     P_2 <- 10^-(runif(1, min=1, max=4))  
+    
+    #23/3/16 set freq high in this expt
+    if (old_insecticide) P_2 <- 0.5
     
     ## exposure to insecticide
     exposure <- runif(1, min=0.1, max=0.9)

### 2016-03-23 paper1_results_figs_slimmed_50.Rmd
added fig x8 to look at 'old' insecticide scenarios

@@
   ## to load previously saved runs
-#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI2_1000.rda'))
 
-  ## trying with the extended experiment _ex100
+  ## trying with the extended experiment
   experiment <- 'extended'
@@
 
-  # load(file=paste0(outFolder,'listOutMix_ex2_1000.rda'))
-  # load(file=paste0(outFolder,'listOutI1_ex2_1000.rda'))
-  # load(file=paste0(outFolder,'listOutI2_ex2_1000.rda'))
+  # load(file=paste0(outFolder,'listOutMix_unequal.rda'))
+  # load(file=paste0(outFolder,'listOutI1_unequal.rda'))
+  # load(file=paste0(outFolder,'listOutI2_unequal.rda'))
 

### 2016-03-24 paper1_results_figs.Rmd
look at unequal runs change hi_div_lo to lo_div_hi

@@
   #add an extra predictor, the lower starting freq of resistance divided by the larger
-  resist_start_hi_div_lo <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
-  treeInput <- rbind(treeInput,resist_start_hi_div_lo)
+  resist_start_lo_div_hi <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
+  treeInput <- rbind(treeInput,resist_start_lo_div_hi)
 
@@
                              "correctMixDeployProp"=1,
-                             "resist_start_hi_div_lo"=1,
+                             "resist_start_lo_div_hi"=1,
                              "resist_start_1_div_2"=1)
@@
 
-    #modify killer plot to add in male exposure prop on x axis
+    #modify killer plot mix2 vs seq to add in male exposure prop on x axis
     library(viridis)

### 2016-03-24 paper1_results_figs_slimmed_50.Rmd
look at unequal runs change hi_div_lo to lo_div_hi

@@
 
-Grey triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels).
-The grey line is a linear model through these same points.
+Red triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels).
+The red line is a linear model through these same points.
 
 \pagebreak
+
+## Fig. x5.1 modified for mix2 vs seq
+
+fewer runs where sequential gives slower resistance
+
+```{r, eval=TRUE, echo=FALSE, fig.height=6}
+
+  ggSubset <- dif_mix2_seq
+  #if mix-seq >1 resistance took longer for mix so mix is better
+  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
+  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
+  ggSubset$abs_diff <- abs(ggSubset$mix2_minus_seq0.5)
+
+  #tricky
+  #be a bit careful what to do with zeros where results are same for both strategies
+  #I use these better vars to facet the plots
+  #may want the zeroes in both facets
+  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
+  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
+  #set these to 1 (they were 0)
+  df_same$mix2_better_seq0.5 <- 1
+  df_same$seq_better_mix20.5 <- 1
+  #rbind them back onto other dataset
+  ggSubset <- rbind( ggSubset, df_same)
+
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
+
+    #find unequal runs
+    unequal <- ggSubset[ggSubset$resist_start_lo_div_hi < 0.01,]
+    nrow(ggSubset)
+    nrow(unequal)
+    #2721 out of the 6189 remaining runs
+
+    #in how many runs was seq better than mix
+    sum(unequal$seq_better_mix20.5)
+    #311
+
+    library(viridis)
+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+
+             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mixture\nresistance slower','1'='sequential\nresistance slower'))) +
+
+             #YlGnBu looks really good except that the yellow points are very faint
+             #scale_colour_distiller(palette='YlGnBu', direction=1) +
+             #option="plasma" looks good too
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
+             theme_bw() +
+             theme(legend.position = "bottom") +
+             #theme_minimal() +
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='black', shape=3) +
+             #failing to add legend for curtis point
+             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
+             #scale_size_manual(name = "Curtis Fig.2", values="test") +
+             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #!!!NOW can I add a lm fitted to the 0 points abs_diff
+             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
+             #can I subtley highlight the zero points which are repeated in each plot ?
+             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
+
+             #now add the unequal points
+             #showed that they were on both sides
+             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=2) +
+
+             #add text for Curtis point
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
+           )
+
+  # doing the lm outside of ggplot    
+  x1=df_same$exposure 
+  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
+
+  l_mod <- lm(y1 ~ x1)
+  #l_mod$coefficients
+  #(Intercept)          x1 
+  #1.019330    0.560508
+
+  #approximation
+  #x2 <- seq(0,1,0.1)
+  #y2 <- 0.6*x2+0.98
+  #plot(x2,y2)
+
+```
 

### 2016-03-24 resistanceModelNotesAndy.Rmd
look at unequal runs change hi_div_lo to lo_div_hi

@@
 Can I show this by modifying the fig2 plot ?
+Although probably we would need to rerun a sensitivity type anlysis if we wanted to address this properly ...
+
+curtis f2
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+curtis f2 with i1 set to high starting freq (0.5 breaks plot slightly, but still seems to show that resistance for i2 takes longer in mixture)
+runcurtis_f2( P_1 = 0.5 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
+runcurtis_f2( P_1 = 0.45 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.45 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+similar with P_2 0.3
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.3 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+but what about the params for curtis_tab1_ex_iv ?
+Beths analysis table uses the same params I think.
+
+Hi Ian,
+
+As a first step I've added an extra fig x8 onto the end of the figures document attached and in dropbox (I started to address figure tweaks, but have held fire on that while we sort this).
+Fig x8 addresses whether to keep using an old insecticide when a new one is introduced.
+It just increases the starting frequency of each resistance in turn to just below the threshold (0.49).
+I think this accomplishes the same as Beths table 4.
+
+Would I be able to alter the parameters to recreate Beths table3 and Curtis table 1 example vi in a similar way ? I couldn't quite work out which inputs to change.
+
+How would we investigate this in a sensitivity analysis ?
+Essentially we could repeat what we've done already but with unequal starting frequencies.
+We had the issue before of not wanting one frequency to be dependent on the other.
+So we could simply set P_1 to a low range as in current analysis, and then have P_2 at a higher range (e.g.
+0.3-0.5).
+Or perhaps we don't need a range for the 'old' insecticide we could just set at 0.5 ?
+
+I'm slightly nervous about running more analyses but I could investigate this with a couple of days work. Then we would need to decide how to squeeze into the paper.
+
+Happy to chat,
+
+Andy
+
+### previous current state of play 23/3/16
+1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
+1. sensiAnPaperPart.r   : sets up param values and runs for one treatment (e.g. mixture, insecticide1)
+1. runModel2.r          : runs model for the passed scenarios
+1. paper1_results_figs_slimmed50.Rmd : final figures for paper
+
+What would I need to do ?
+
+1. sensiAnPaperPart.r : set new param ranges for P_2, perhaps add an argument to enable running them 
+1. sensiAnPaper1All.Rmd : rerun calling new version of sensiAnPaperPart, save results to something like : listOutMix_unequal_10000.rda
+1. paper1_results_figs_slimmed50.Rmd : save with new name, just rerun using the new results files
+
+perhaps do for 500 runs first ?
+
+reran new unequal scenarios with start_freq set to constant 0.5.
+first plots OK, but then get a weird thing where mixa_minus_seq0.5 is always -2.
+So mix adapative is always taking 2 generations less than sequential.
+I suspect this may be an artefact of detecting resistance thresholds when the threshold is already reached for i2.
+
+~ created pap1_figs_slimmed_50_unequal.Rmd
+~ all changes marked with #4unequal
+
+Seems that resistance is slower for mixtures in all replicates.
+For the 500 runs at least.
+
+** 23/3/16 **
+email from Ian
+
+I have been racking my brain to think if we can address this question using  data that have already been generated (and hence avoiding having to explain how we generated new data) .
+Rather than trying to explain it informally in an Email, then write it up, I went directly to step 2 and wrote the following as a draft paragraph for the ms….
+Does it make sense???
+
+“We can analyse a subset of the sensitivity analysis to investigate this specific question of whether single  use of a new insecticide is better than deploying it as a mixture with a pre-existing insecticide which already has some resistance present in the population.
+We selected runs where the starting frequency of resistance to insecticide 2 was > 10 fold higher than frequency of resistance to insecticide 1.
+We then compared the time for insecticide 1 resistance to reach 50% when deployed on its own, compared to time for resistance to insecticide 1 to reach 50% when deployed as a mixture with the pre-existing insecticide.
+As in the full analysis, we compute the ratio mixture/single for each run and generate a classification tree to see what circumstances favour what policy and, in particular whether the ratio of resistance starting frequencies has a large impact”
+
+I am assuming insecticide #1 is always deployed first in the sequential runs so the time to its resistance reaching 50% is already recorded.
+I am also assuming that in the mixture runs you have recorded  the time until resistance to each insecticide reaches 50% so you always have the data #1.
+
+this is the important bit from Ian :
+We selected runs where the starting frequency of resistance to insecticide 2 was > 10 fold higher than frequency of resistance to insecticide 1.
+We then compared the time for insecticide 1 resistance to reach 50% when deployed on its own, compared to time for resistance to insecticide 1 to reach 50% when deployed as a mixture with the pre-existing insecticide.
+
+I think it would be easier to do this by new runs, which is close to what what I've done in pap1_figs_slimmed_50_unequal.
+(although I might want to change the freq of the 'old' insecticide from a constant 0.5 to a range e.g.
+0.2 - 0.5, but then would).
+
+Could I practically do what Ian has suggested ? 
+How many runs are there where the start freq of i2 is > 10 * i1 ?
+Where resist_start_1_div_2 < 0.1
+
+see in paper1_results_figs.rmd
+
+note my hi_div_lo was actually lo_div_hi, corrected in the figs_slimmed_50.rmd
+
+resist_start_hi_div_lo <- ifelse( treeInput['P_1',] < treeInput['P_2',], treeInput['P_1',]/treeInput['P_2',], treeInput['P_2',]/treeInput['P_1',])
 
@@
 
-1. add option to rescale x axis to years or months in plotcurtis_f2_generic, generation=1month
+aha! had a thought. Can I highlight those runs in the killer plot where resist_start_1_div_2 < 0.1 ?
+
+probably need to do it for the mix2 graph rather than mixadaptive.
+
 1. tidy these notes
@@
 
-### current state of play 18/12/15
+### current state of play 23/3/16
 1. sensiAnPaper1All.Rmd : code to run whole sensitivity analysis & produce plots 
@@
 1. runModel2.r          : runs model for the passed scenarios
+1. paper1_results_figs_slimmed50.Rmd : final figures for paper
 

### 2016-03-30 paper1_results_figs_slimmed_50.Rmd
tweaks

@@
   #in main paper I want to just display sequential and mixture strategies
-  strategies <- strategies[-1:-2]
+  strategies <- c("Sequential", "Mix adaptive", "Mix both")
+  strat_name <- c("Seq both","Adaptive both","Mix both")
 
@@
              ylim(-1,1) +
-             ggtitle(paste(strategy)) +
+             ggtitle(paste(strat_name[strategy_num])) +
              xlab(NULL) 
@@
 
-  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.8))
 
@@
 
+## Fig. x6 Modifying the previous plot to show where resistance arises over 20% slower in an adaptive mixture strategy versus sequential.
+
+1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the adaptive mixture under these constraints.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=6}
+
+  df_20pc <- ggSubset[which(ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq > 0.19 & 
+                            ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq < 0.21  ),]
+
+  #remove column it is faceted on to get line in both facets
+  col_indices <- colnames(df_20pc) != 'mixA_not20pc_better_seq0.5'
+  df_20pc <- df_20pc[,col_indices]
+
+  #modify the killer plot to show where mixtures are 20% better ?
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+
+             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+
+             theme_bw() +
+             theme(legend.position = "bottom") +
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
+             #add point for Curtis inputs
+             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) +
+             #add a lm fitted to the 0 points abs_diff
+             #geom_smooth(data=df_same, method = "lm", se = FALSE)
+             #todo change to where exactly 20% diff
+             geom_smooth(data=df_20pc, method = "lm", se = FALSE,colour='red', lwd=0.5)+
+             #can I subtley highlight the zero points which are repeated in each plot ?
+             geom_point(data=df_20pc, size=0.8, colour='red', shape=2) #, fill='white') 
+           )
+
+  # doing the lm outside of ggplot    
+  x1=df_20pc$exposure 
+  y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
+
+  l_mod <- lm(y1 ~ x1)
+  l_mod$coefficients
+# (Intercept)          x1 
+#   1.3914547   0.2904911
+
+```
+
+\pagebreak
+
 ## Fig. x5.1 modified for mix2 vs seq
@@
   #plot(x2,y2)
-
-```
-
-## Fig. x6 Modifying the previous plot to show where resistance arises over 20% slower in an adaptive mixture strategy versus sequential.
-
-1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the adaptive mixture under these constraints.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=6}
-
-  df_20pc <- ggSubset[which(ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq > 0.19 & 
-                            ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq < 0.21  ),]
-
-  #remove column it is faceted on to get line in both facets
-  col_indices <- colnames(df_20pc) != 'mixA_not20pc_better_seq0.5'
-  df_20pc <- df_20pc[,col_indices]
-
-  #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
-
-             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
-
-             theme_bw() +
-             theme(legend.position = "bottom") +
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
-             #add point for Curtis inputs
-             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) +
-             #add a lm fitted to the 0 points abs_diff
-             #geom_smooth(data=df_same, method = "lm", se = FALSE)
-             #todo change to where exactly 20% diff
-             geom_smooth(data=df_20pc, method = "lm", se = FALSE,colour='red', lwd=0.5)+
-             #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=df_20pc, size=0.8, colour='red', shape=2) #, fill='white') 
-           )
-
-  # doing the lm outside of ggplot    
-  x1=df_20pc$exposure 
-  y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
-
-  l_mod <- lm(y1 ~ x1)
-  l_mod$coefficients
-# (Intercept)          x1 
-#   1.3914547   0.2904911
 

### 2016-03-30 paper1_results_figs.Rmd
starting new analysis to compare use of new insecticide with old to new on its own

@@
   dif_mix2_seq["mix2_minus_seq0.25"] <- dif_mix2_seq["gen_cP0.25"] - dif_mix2_seq["gen_cP0.25seq"]
+
+```
+
+```{r,echo=FALSE,results='hide'}
+### NEW part to analyse what happens when a new insecticide is added to an existing one
+### either to replace it or to continue in a mixture
+
+  # find the resistance points for I1 & I2 in a mixture 
+  #(diff than Mix_1 & 2 which is for 1st & 2nd threshold to be reached)
+  resistPointsMixI1 <- findResistancePoints(listOutMix, locus=1)
+  resistPointsMixI2 <- findResistancePoints(listOutMix, locus=2)
+
+  resistPointsMixI1_T <- addStrategyColumn(resistPointsMixI1,"I1inMix",ggInput) 
+  resistPointsMixI2_T <- addStrategyColumn(resistPointsMixI2,"I2inMix",ggInput)
+
+  #and use these from above
+  #resistPointsI1_T
+  #resistPointsI2_T
+
+  #OldPlusNew 
+  # find whether I1 or I2 has lower starting freq this is the 'new' insecticide
+  # if I1 < I2 use resistPointsMixI1_T
+  # if I2 < I1 use resistPointsMixI2_T
+
+  #NewOnly
+  # if I1 < I2 use resistPointsI1_T
+  # if I2 < I1 use resistPointsI2_T
+
+  # find whether I1 or I2 has lower starting freq this is the 'new' insecticide
+  # rember the inputs are copied onto the start of all resistPoints objects
+
+  indicesI2_lessthan_I1 <- which( resistPointsI1_T$start_freq_allele2 < resistPointsI1_T$start_freq_allele1 )
+  #this shows about half, I also checked no equals which is good
+  #length(indicesI2_lessthan_I2)
+  #[1] 4983
+
+  #NewOnly first make a copy from existing
+  resistPointsNewOnly <- resistPointsI1_T
+  #replace those values where I2 is the 'new' insecticide with lower resistance
+  resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
+     resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
+
+  #OldPlusNew first make a copy from existing
+  resistPointsOldPlusNew <- resistPointsMixI1_T
+  #replace those values where I2 is the 'new' insecticide with lower resistance
+  resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
+  resistPointsMixI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
+
+  resistPointsNewOnly$strategy <- "new only"
+  resistPointsOldPlusNew$strategy <- "old plus new"
+
+  #BE CAREFUL
+
+  #format data to enable PRCC on difference between the 2 strategies
+  #follow previous naming where
+  #OldPlusNew="gen_cP0.5", NewOnly="gen_cP0.5seq"
+
+  #columns to remove from first
+  indices1 <- which(names(resistPointsOldPlusNew) %in% c("gen_cP0.1","gen_cP0.25","strategy"))
+  #columns to add to 2nd  
+  indices2 <- which(names(resistPointsNewOnly) %in% c("gen_cP0.5")) 
+  #rename column in 2nd
+  tmp <- resistPointsNewOnly[indices2]
+  names(tmp) <- "gen_cP0.5seq"
+
+  dif_oldnew_newonly <- cbind(resistPointsOldPlusNew[-indices1], tmp)
+
+  #remove runs that didn't reach a resistance threshold (999), even if just for one strategy
+  #>1000 excludes sequential strategy that had a 999 in
+  #BEWARE would need to increase 1000 if I increase max generations in the runs
+  didntReachThresh <- which( dif_oldnew_newonly$gen_cP0.5 > 500 | dif_oldnew_newonly$gen_cP0.5seq > 500 )
+  #subset by runs not making threshold
+  dif_oldnew_newonly <- dif_oldnew_newonly[-didntReachThresh,]
+
+  dif_oldnew_newonly["mix_minus_sole0.5"] <- dif_oldnew_newonly["gen_cP0.5"] - dif_oldnew_newonly["gen_cP0.5seq"]
 

### 2016-03-30 resistanceModelNotesAndy.Rmd
starting new analysis to compare use of new insecticide with old to new on its own

@@
 
+Qs for Ian
+~ is effectiveness of 1 possible/likely ? The pattern in the curtis plot relies quite heavily on it. Reducing it to just 0.98 changes pattern considerably. 
+~ are we confident enough about the mix adaptive strategy to focus the paper on it ?
+Ian said yes.
+He has said before that it just takes a couple of generations for frequencies to return to Hardy Weinberg.
+
 Remember that effectiveness is proportion of the SS genotypes that die when exposed.
@@
 
-Fig x2..
+done ~ Fig x2..
 can we make the labels identical to those above i.e., in order  "Seq both"  "Mix 1st"  "Adaptive both "   "Mix both"....
@@
 whey we omit "Seq 1st"
+We can make it clear we are looking at the most likely application strategies so we only show those when resistance reached for both.
+(largely it's a space thing and the others don't show anything different).
 
@@
 
+aha! had a thought. Can I highlight those runs in the killer plot where resist_start_1_div_2 < 0.1 ?
+probably need to do it for the mix2 graph rather than mixadaptive.
+
+I tried and it didn't show the split I expected perhaps because the higher start resist isn't high enough.
+
+** 29/3/16 **
+
+spoke to Ian on the phone. He's still keener on reanalysing the current sensi results.  Because he said its more relevant where the existing insecticide has a resistance level of 5-10% (or even less).
+
+this is the important bit from Ian :
+We selected runs where the starting frequency of resistance to insecticide 2 was > 10 fold higher than frequency of resistance to insecticide 1.
+We then compared the time for insecticide 1 resistance to reach 50% when deployed on its own, compared to time for resistance to insecticide 1 to reach 50% when deployed as a mixture with the pre-existing insecticide.
+
+probably similar to :
+findResistancePointsMixResponsive()
+
+Effectively I think this is just comparing mixAdaptive with mixBoth for the subset of scenarios.
+
+Aha no actually isn't it comparing sole use with mixture 'new' insecticide, the different bit will be that its not quite the same as the current mix1 & mix2.
+Because I don't think we care about the resistance level of the 'old' insecticide.
+
+Maybe
+1 create a function to calc when resistance is reached for new+old
+2 create a 2nd func to calc when resistance reached for new on its own
+
+1) findResistancePointsOldPlusNew <- function( listOutMix,
+                                               listOutI1,
+                                               listOutI2 )
+
+~ find whether I1 or I2 has lower starting freq this is the 'new' insecticide
+~ if new==I1 get from mix time til resistance for I1
+~ if new==I2 get from mix time til resistance for I2
+
+2) findResistancePointsNewOnly <- function( listOutI1,
+                                            listOutI2 )
+
+~ find whether I1 or I2 has lower starting freq this is the 'new' insecticide
+~ if new==I1 get from I1 time til resistance for I1
+~ if new==I2 get from I2 time til resistance for I2
+
+For each get the outputs as
+
+then create a dataframe : dif_mixnew_sole
+
+** 30/3/16 **
+
+~ rather than doing as above I can use the existing files for I1 & I2
+  resistPointsI1 <- findResistancePoints(listOutI1, locus=1)
+  resistPointsI2 <- findResistancePoints(listOutI2, locus=2)
+~ & then calc
+  resistPointsMixI1 <- findResistancePoints(listOutMix, locus=1)
+  resistPointsMixI2 <- findResistancePoints(listOutMix, locus=2)
+
+Then post-process from these files for when starting freq of I2 >10*I1
+
+A future thought.
+If I bit the bullet and allowed the simulation to represent adaptive strategies things might be much easier ?
+Also in future could I allow the effectiveness of insecticides to change over time, similar to how we have done in the game ?
+
+doing this in paper1_results_figs.Rmd
+
+now have :
+dif_oldnew_newonly["mix_minus_sole0.5"]
+
 ### TO DO
 
-aha! had a thought. Can I highlight those runs in the killer plot where resist_start_1_div_2 < 0.1 ?
-
-probably need to do it for the mix2 graph rather than mixadaptive.
+~ check on dif_oldnew_newonly["mix_minus_sole0.5"]
+
+# quick thought check, sequential is calculated by adding together the times for I1 & I2 in isolation, are we sure that is OK ?
+e.g.
+do we expect it to take same time for resistance to 2nd insecticide to reach threshold irrespective of the level of resistance to I1 even when I1 is not present.
 
@@
 To get the exposure of both M&F : x=exposure+exposure*maleExposureProp
-
-Qs for Ian
-~ is effectiveness of 1 possible/likely ? The pattern in the curtis plot relies quite heavily on it. Reducing it to just 0.98 changes pattern considerably. 
-~ are we confident enough about the mix adaptive strategy to focus the paper on it ? we could look at standard mix instead ?
 

### 2016-04-16 paper1_results_figs.Rmd
results for a new insecticide added to an old or on its own

@@
 ```{r,echo=FALSE,results='hide'}
-### NEW part to analyse what happens when a new insecticide is added to an existing one
+### 4/2016 NEW part to analyse what happens when a new insecticide is added to an existing one
 ### either to replace it or to continue in a mixture
@@
   #[1] 4983
+  #check whether this works
+  #check <- resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
+  #aha! problem I was having is that I1 & I2 inputs are not swapped around
+  #if I wanted to swap the inputs for I1 & I2 around, would need to do other coefficients etc.
+  #much potential confusion
+  #might be better just to restrict to the runs in which I2 < I1
+  resistPointsNewOnly <- resistPointsI2_T[indicesI2_lessthan_I1, ]
+  resistPointsOldPlusNew <- resistPointsMixI2_T[indicesI2_lessthan_I1, ]
 
-  #NewOnly first make a copy from existing
-  resistPointsNewOnly <- resistPointsI1_T
-  #replace those values where I2 is the 'new' insecticide with lower resistance
-  resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
-     resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
-
-  #OldPlusNew first make a copy from existing
-  resistPointsOldPlusNew <- resistPointsMixI1_T
-  #replace those values where I2 is the 'new' insecticide with lower resistance
-  resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
-  resistPointsMixI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
+  # #this bit now replaced by the 2 previous lines
+  # #NewOnly first make a copy from existing
+  # resistPointsNewOnly <- resistPointsI1_T
+  # #replace those values where I2 is the 'new' insecticide with lower resistance
+  # resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
+  #    resistPointsI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]  
+  # 
+  # #OldPlusNew first make a copy from existing
+  # resistPointsOldPlusNew <- resistPointsMixI1_T
+  # #replace those values where I2 is the 'new' insecticide with lower resistance
+  # resistPointsNewOnly[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')] <-
+  # resistPointsMixI2_T[indicesI2_lessthan_I1, c('gen_cP0.1','gen_cP0.25','gen_cP0.5')]
 
@@
   dif_oldnew_newonly["mix_minus_sole0.5"] <- dif_oldnew_newonly["gen_cP0.5"] - dif_oldnew_newonly["gen_cP0.5seq"]
+
+  #*****************
+  # to check
+  # are there any cases where startfreqI1 > startfreqI2
+  # by my method above i don't think there should be
+  # i think min should be 1
+  min(dif_oldnew_newonly$resist_start_1_div_2)
+  max(dif_oldnew_newonly$resist_start_1_div_2)
+
+  min( dif_oldnew_newonly$start_freq_allele1 - dif_oldnew_newonly$start_freq_allele2)
+  max( dif_oldnew_newonly$start_freq_allele1 - dif_oldnew_newonly$start_freq_allele2)
+
+  #resistPointsNewOnly
+  min(resistPointsNewOnly$resist_start_1_div_2)
+  max(resistPointsNewOnly$resist_start_1_div_2)
+
+  # because I now just use runs where startfreq12 < I1
+  # resist_start_1_div_2 should be the inverse of resist_start_hi_div_lo
+  # or at least directly correlated
+  # yes they are all on a single curved line
+  # plot( resistPointsNewOnly$resist_start_1_div_2 ~ resistPointsNewOnly$resist_start_lo_div_hi)
+
+  #****************
+  #not sure I need to do this because all results already show mixture better
+  #now I want to subset just those runs where start freq old is > 10 * new
+
+  #proportion of runs in which mix better than sole
+  num_mix_better <- length(which(dif_oldnew_newonly["mix_minus_sole0.5"] > 0))
+  num_all <- nrow(dif_oldnew_newonly)
+  num_mix_better/num_all
+  #1
+  #?now mix always seems to be better
+  #which kind of makes sense, any insecticide better than none
 
@@
 
+```
+
+## Fig. xx3 Using a new insecticide on its own or in combination with old
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
+  ggSubset <- dif_oldnew_newonly
+
+  names_inputs <- names(ggSubset)[1:num_inputs]
+
+  for(i in names_inputs)
+  {
+    #x11()
+
+    y <- "mix_minus_sole0.5"
+
+    print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
+             geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
+             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             geom_smooth(linetype='dashed',size=1, colour="red") +
+             #facet_wrap( ~ strategy) +
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             labs(title = i) +
+             geom_hline(yintercept = 0, colour="white")
+           )
+  }
+
+```
+
+# Fig xxx PRCC difference in time-to-resistance between sole new and mix old+new
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
+
+  library(sensitivity)
+
+    x <- dif_oldnew_newonly[,1:num_inputs]
+    y <- dif_oldnew_newonly["mix_minus_sole0.5"]
+
+    pcc_res <- pcc(x, y, rank=TRUE)
+
+    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
+    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
+    #plot(pcc_res)
+
+    #results are here I can probably rbind them together into a df that I can ggplot
+    #pcc_res$PRCC
+
+    to_plot <- pcc_res$PRCC
+    #rename column 1 from 'original to PRCC
+    names(to_plot)[1] <- 'PRCC'
+    to_plot$inputs <- rownames(to_plot)
+
+    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+             geom_point(shape=1, colour='red') +
+             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             geom_hline(yintercept = 0, linetype=3) +
+             ylim(-1,1) +
+             ggtitle(paste("PRCC benefit of mix old+new over sole new (old=1, new=2)")) +
+             xlab(NULL)
+          )
 ```
@@
           )
-
 ```

### 2016-04-16 resistanceModelNotesAndy.Rmd
results for a new insecticide added to an old or on its own

@@
 
+BUT this is still for all runs (i.e. not just the ones where start freq of the old is > 10 * the new)
+
+** 11/4/2016 **
+
+I think I want to plot this against each input variable to check whats going on, i.e. where is the mix better than sole use.
+
+dif_oldnew_newonly["mix_minus_sole0.5"]
+
+see Fig xx3
+
+In most (70%) runs mixture are better, but that does leave 30% where sole better.
+
+Eyeballing the plots suggests that exposure is still the most influential input. At high exposures the benefit of the mixture is less.
+
+May be something weird going on with start_freq_allele2, noticeable lack of runs where mix-sole < 0 as start_freq_allele2 gets higher ...
+
+Are the inputs relating to allele1 & 2 still meaningfull with the post-processing that I do ?
+
+No.
+aha! problem I was having is that I1 & I2 inputs are not swapped around.
+
+  #might be better just to restrict to the runs in which I2 < I1
+  resistPointsNewOnly <- resistPointsI2_T[indicesI2_lessthan_I1, ]
+  resistPointsOldPlusNew <- resistPointsMixI2_T[indicesI2_lessthan_I1, ]
+
+when I decided to restrict to runs where startfreqI2 < I1 mix seems to be better all the time ??
+
+? am I doing something wrong ?
+seems not
+
+Eyeballing fig xx3
+
+when I1 is the old :
+Most important inputs & their effect on the benefit of mixture are :
+exposure : reduces benefit of mixture
+effectiveness i1,old : increases benefit of mixture
+dominance i2,new : reduces benefit of the mixture
+selection coefficient i2 : reducces benefit of the mixture
+
+My thoughts on mechanisms.
+The mixture does better because the old insecticide kills some mosquitoes that have developed resistance to the new, thus reducing rate of increase in resistance to the new. 
+Exposure. Low exposure reduces the benefits of mixtures. This could be because the beneficial effect of the old insecticide is reduced.
+Effectiveness of the old insecticide. High effectiveness increases the benefit of the mixture because more of the mosquitoes that have developed resistance to the new insecticide are killed.
+Dominance.
+High dominance of the new insecticide allele decreases the benefit of the mixture.
+High dominance in resistance to the new insecticide should allow more mosquitoes to escape the new insecticide therefore I would expect the benefit of the mixture to be greater.
+But its the other way around.
+
+I probably need to get straighter in my head how dominance works.
+Does dominance effect the relative performance of the 2 insecticides where mosquitos are exposed to both.
+i.e.
+Does high dominance of the new allele mean that a mosquito is more likely to survive the presence of the old insecticide ?
+
+To understand the mechanisms more I could look at the effects of the inputs on each strategy on its own.
+
+These results are consistent with previous ones.
+
+~ ran PRCC on the new results (confirmed same important params)
+
+copied new bits into keep_old_insecticide.Rmd and sent the pdf to Ian
+
+** 12/04/2016 **
+
+Ian was surprised that mixtures better in all runs.
+
+He suggested that The Curtis eg where sole better might be strange and not fall within our parameter space.
+
+~ check whether I can recreate some results from the new analysis by putting the params into the UI, or better by calling runcurtis_f2()
+
+e.g. first run in dif_oldnew_newonly
+mix : 180
+sole : 48
+difference : 132
+
+#a previous example
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.3 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+#putting params of 1st run in
+runcurtis_f2( P_1 = 0.0159762192 , P_2 = 0.0076494164 , h.RS1_A0 = 0.660797792 , h.RS2_0B = 0.629114044 , exposure = 0.5582827 , phi.SS1_A0 = 0.9441143 , phi.SS2_0B = 0.9695714 , s.RR1_A0 = 0.1216252 , s.RR2_0B = 0.1720911, addCombinedStrategy=FALSE )
+
+gives warning :
+critical point not reached for i1 or i2: cutoff_i1_mix=Inf cutoff_i2_mix=Inf
+
+is this because ?? :
+1. I've pasted wrong
+no
+2. an indexing issue means that inputs and results in the table are misaligned
+
+3. what i'm running in these new runs is different from whats done in the plot
+don't think so
+4. extra defaults no passed to the plot are different from those used in sensi runs
+***aha! could be maleExposureProp and correctMixDeployProp
+
+Looking at the plot sole use does look like it takes about 48 as in the results, but for I2 in the mix looks like it should take ~600 rather than 180 as suggested in the results.
+
+New corrected version :
+
+runcurtis_f2( P_1 = 0.0159762192 , P_2 = 0.0076494164 , h.RS1_A0 = 0.660797792 , h.RS2_0B = 0.629114044 , exposure = 0.5582827 , phi.SS1_A0 = 0.9441143 , phi.SS2_0B = 0.9695714 , s.RR1_A0 = 0.1216252 , s.RR2_0B = 0.1720911, addCombinedStrategy=FALSE, maleExposureProp = 0.90820779, correctMixDeployProp = 0.6008410 )
+
+YES this does now match. Hurrah!
+(also shows that maleExposureProp and/or correctMixDeployProp in isolation can have a big effect that I could explore in the mechanisms paper).
+
+Should I do for another result too, just to confirm ?
+
+Hi Ian,
+I would like to see if I can recreate Curtis Table 1, (vi).
+
+It makes my head hurt trying to work out what our inputs should be based on his table.
+
+Could you help me out by filling in the below * ? (or is it not possible and I'm missing something?).
+# start frequencies of each resistance allele
+P_1 = *, 
+P_2 = * , 
+# dominances
+h.RS1_A0 =  *, 
+h.RS2_0B =  *, 
+# exposures
+exposure =  *,
+# effectivenesses
+phi.SS1_A0 =  *, 
+phi.SS2_0B =  *, 
+# selection coefficients
+s.RR1_A0 =  *, 
+s.RR2_0B =  *
+
+Here are the equivalent values used to recreate his figure 2.
+# start frequencies of each resistance allele
+P_1 = 0.01 , 
+P_2 = 0.01 , 
+# dominances
+h.RS1_A0 = 0.17 , 
+h.RS2_0B = 0.0016 , 
+# exposures
+exposure = 0.9 ,
+# effectivenesses
+phi.SS1_A0 = 0.73 , 
+phi.SS2_0B = 1 , 
+# selection coefficients
+s.RR1_A0 = 0.23 , 
+s.RR2_0B = 0.43
+
+cheers,
+Andy
+
+Back from Ian :
+
+P_1 = 0.01,
+P_2 = 0.001,
+# dominances
+h.RS1_A0 =  1,
+h.RS2_0B =  1,
+# exposures
+exposure =  0.9,
+# effectivenesses
+phi.SS1_A0 =  1,
+phi.SS2_0B =  1,
+# selection coefficients
+s.RR1_A0 =  1,
+s.RR2_0B =  1 )
+
+runcurtis_f2( P_1 = 0.01,
+P_2 = 0.001,
+# dominances
+h.RS1_A0 =  1,
+h.RS2_0B =  1,
+# exposures
+exposure =  0.9,
+# effectivenesses
+phi.SS1_A0 =  1,
+phi.SS2_0B =  1,
+# selection coefficients
+s.RR1_A0 =  1,
+s.RR2_0B =  1,
+addCombinedStrategy=FALSE )
+
+This seems to generate a very rapid rise in resistance.
+
+and generrates this error in the graph :
+Error in xy.coords(x, y) : 'x' and 'y' lengths differ
+addCombinedStrategy=FALSE fixes the error
+
+i think I should switch i1 & 2 frequencies
+
+runcurtis_f2( P_1 = 0.001,
+P_2 = 0.01,
+# dominances
+h.RS1_A0 =  1,
+h.RS2_0B =  1,
+# exposures
+exposure =  0.9,
+# effectivenesses
+phi.SS1_A0 =  1,
+phi.SS2_0B =  1,
+# selection coefficients
+s.RR1_A0 =  1,
+s.RR2_0B =  1,
+addCombinedStrategy=FALSE,
+addStrategyLabels = FALSE )
+
+This seems to show that resistance to the new insecticide (red) rises faster when alone (dotted) than when in combination (solid) with an old insecticide at a higher frequency.
+This is contrary to what Curtis said.
+
+Also note that resistance arises much faster than in our sensitivity runs. I think  this is because the selection coefficients (1) are much higher than the range we use (0.1-0.45).
+
+** 15/4/16 **
+
+from Ian :
+I have re-read Curtis again and its that paragraph on page 261 that is causing all the problems i.e.
+
+“Example vi shows the, at first unexpected……”
+
+I guess the key phrase  is what he means by “more rapid”…I assume he means more rapid than using it on its own.
+But that doesn’t tally with your figure.
+His example is only a 1-genertion argument so I should be able to do some simple arithmetic.
+*Can you tell me what the allele frequencies were at generation 2 of your plot* and I’ll check against my arithmetic.
+
+to get allele freqs at gen 2 :
+
+tst$results[[2]][1:8,]
+
 ### TO DO
 
-~ check on dif_oldnew_newonly["mix_minus_sole0.5"]
+~ how to recreate Curtis Table 1, (vi)'the use of a mixture when the initial gene frequencies are unequal leads to a more rapid increase in the frequency of the rarer of the genes'.
+
+~ remember this is still for all runs (i.e. not just the ones where start freq of the old is > 10 * the new)
+
+~ **check this from above**
+~ I probably need to get straighter in my head how dominance works.
+Does dominance effect the relative performance of the 2 insecticides where mosquitos are exposed to both.
+i.e.
+Does high dominance of the new allele mean that a mosquito is more likely to survive the presence of the old insecticide ?
+
+To understand the mechanisms more I could look at the effects of the inputs on each strategy on its own.
 

### 2016-04-16 R/runcurtis_f2.r

@@ -180,4 +180,7 @@ runcurtis_f2 <- function( max_gen = 500,
                          strategyLabels = strategyLabels
                          )  
   
+  #add returning the results
+  invisible(listOut)
+  
 }

### 2016-04-28 paper1_results_figs.Rmd
testing that mix always slower than sole

@@
 
-  #****************
   #not sure I need to do this because all results already show mixture better
@@
   #which kind of makes sense, any insecticide better than none
+
+  #28/4/16 testing Ians statement
+  #does resistance always arise faster in sole use than in a mixture (irrespective of starting frequencies)
+  #resistPointsI1_T #"insecticide 1" 
+  #resistPointsI2_T #"insecticide 2"
+  #resistPointsMixI1_T #"I1inMix" 
+  #resistPointsMixI2_T #"I2inMix"
+
+  mix_minus_soleI1 <- resistPointsMixI1_T["gen_cP0.5"] - resistPointsI1_T["gen_cP0.5"]
+  which( mix_minus_soleI1 < 0 ) 
+  #0
+  mix_minus_soleI2 <- resistPointsMixI2_T["gen_cP0.5"] - resistPointsI2_T["gen_cP0.5"]
+  which( mix_minus_soleI2 < 0 ) 
+  #0
+
+  #so yes in our param space adding an extra insecticide always increases resistance time for the first
+
+  #i could plot it against exposure & effectiveness
 

### 2016-04-28 resistanceModelNotesAndy.Rmd
notes

@@
 
+** 18/4/16 **
+
+email from Ian
+Morning Andy,
+
+here is my logic, honed by long walks along the beach
+
+(1)    According to your analysis of a subset of the sensitivity analysis i.e.
+where the “new” insecticide is present at lower frequency…….it always lasts longer under a mixture (presumably this time is defined as that taken for the resistance allele to reach 50%)
+
+(2)    But is some of the runs in the sensitivity analysis, sequential use is favoured as it maximises the time for both resistance alleles to reach 50%.
+Given the result in (1) this can only occur if , in some runs, the “old” insecticide with higher frequency of resistance spreads faster under a mixture.
+
+Does this make sense? Can you check if it is true?
+
+Ian
+
+~ in runs where resistance arises slowest for sequential does resistance to the the old (higher freq) insecticide arise faster in the presence of the new insecticide than without it ?
+
+~ look at an example of a run where sequential is 'best'
+
+Hi Ian,
+Good question. I'm somewhat jet-lagged and a little hungover in a seminar room In Seattle so not thinking at my best but I'll try a quick shot which may be wrong.
+
+The example below is the first one I could find where resistance arises slower for sequential. (it happens to be where the starting frequencies are equal).
+
+In this case it seems that sequential is slower because resistance to insecticide2 (blue) only starts increasing when resistance to the first insecticide has been reached.
+The curve itself is steeper in the absence of the other insecticide (compare blue dashed to solid) but because it starts later it gets to the threshold later.
+
+I'll try to look for another example.
+
+Andy
+
+#sequential best
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+#relative to curtis fig2
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+#sequential best. effectivenesses bothg reduced by 0.2 from curtis fig2
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.53 , phi.SS2_0B = 0.8 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+~ looking at my killer plot gave me a clue how to set the parameter ranges for when sequential should be better. Low effectiveness and high exposure.
+
+#trying to make sequential better, high exposure, low phi
+
+runcurtis_f2( P_1 = 0.05 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 1 , phi.SS1_A0 = 0.1 , phi.SS2_0B = 0.1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
+
+FRom Ian web version gives an error with certain param combinations
+
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 1 , h.RS2_0B = 1 , exposure = 0.9 , phi.SS1_A0 = 1 , phi.SS2_0B = 1 , s.RR1_A0 = 1 , s.RR2_0B = 1, addCombinedStrategy=TRUE )
+Error in xy.coords(x, y) : 'x' and 'y' lengths differ
+
+no error when I set addCombinedStrategy=FALSE, but it doesn't plot a curve for i1 mixture ?? or maybe that i1 mix line is behind i2 mix line. Setting a transparency could solve that.
+
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 1 , h.RS2_0B = 1 , exposure = 0.9 , phi.SS1_A0 = 1 , phi.SS2_0B = 1 , s.RR1_A0 = 1 , s.RR2_0B = 1, addCombinedStrategy=FALSE )
+
+where is the error ?
+
+in plotcurtis_f2_generic.r
+
+lines( gens, comb, col="red", lty=3 )
+
+because comb contains nothing.
+
+** 27/4/16 **
+
+email from Ian :
+I agree entirely with your analysis of why sequential sometimes lasts longer than mixtures (see below in case you have forgotten!).
+I updated the ms to reflect this and tried to get a coherent discussion of the Curtis results.
+See new text on pages 16 (paragraph starting “We next investigated the statement by Curtis  “) and on page 19 (paragraph starting “The secondary objective was to apply”).
+
+As you will see I put my head on the block (page 16 when I said)
+“Finally, for future use, we analysed the other sub-set of the sensitivity data i.e.
+when the first insecticide to be deployed in the sequence was the one with the higher frequency of resistance and determined whether resistance to that insecticide spread faster when it was deployed on its own or as part of a mixture.
+Again, in all cases, resistance spread faster when the insecticide was deployed on its own.
+In summary then, resistance to an insecticide always spreads faster when that insecticide is deployed alone rather than as part of a mixture, irrespective of the relative frequencies of resistance to the two insecticides.“
+
+Can you check if that is true.
+
+Secondly, a minor point, but it may become important when determining number of generation for resistance to reach 50%...
+do you number the initial generation zero so that subsequent generation numbers reflect the elapsed number of generation of selection?
+e.g.
+Figure 10  looks like the initial generation is #1.
+Fig 12 looks like its starts from 0 but the scale is small
+
+I had immense problem understanding Table 3 which came from a very early version.
+I have checked back in my archives but have no previous instnaces of it.
+It makes no sense to me to and it is not compatible with the dynamics shown on Figure 10 (see note on ms).
+For example I cannot understand why time to resistance under a mix is not affected by the starting frequency at locus A.
+If you cannot understand it either, can you re-derive the results and correct it.
+Thanks.
+
+Table 4 looks sensible to me but if you are rechecking table 3, perhaps you could check a couple of values in table 4 just to make sure nothing unexpected has happened.
+
+Like I said, happy to Skype if you want to discuss further.
+Once we get this bit finalised, we can circulate to the Curtis-fans.
+Like I explain on page 19, I think they may have misinterpreted the comparison he was trying to make.
+
+~~ I checked tables 3 & 4, they were Beths so I sent an email to Ian asking him to ask her
+
+** 28/4/16 **
+
+https://www.bayer.co.za/en/bayer-develops-the-first-two-way-insecticide-mixture-for-indoor-residual-spraying.php
+
+Consistent with its mission of delivering ‘Science For A Better Life’, Bayer has recently submitted a dossier to the World Health Organization Pesticide Evaluation Scheme (WHOPES) for the evaluation of a new two-way insecticide mixture which includes a new mode-of-action for indoor residual spraying (IRS) against disease vectors.
+Named Fludora™ Fusion, this first IRS based on two active ingredients is intended to provide an effective solution to help African disease control programs address the challenge of insecticide resistance in malaria-transmitting mosquitoes.
+Field testing of the product has shown excellent results against many different kinds of resistant mosquitoes and strong performance across a wide range of surfaces.
+Bayer foresees the WHOPES evaluation and testing process to take about 2 years and anticipates market availability of the product by the end of 2017.
+
+This has some good stuff about IRM strategies : 
+http://www.slideshare.net/srinivasnaik52643/insecticide-resistance-management-strategy
+
+Curtis et al 1993, potential IRM strategies :
+~ moderation : preserve susceptible genes by limiting selection pressure
+~ saturation : high dose so that heterozugous resistants are killed
+~ multiple attack : independently acting pressures nneither of which is strong enough to lead to resistance
+
+Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
+
+Abstract
+A two-locus model is used to analyze the effectiveness of a mixture of insecticides in delaying resistance, compared to the use of the insecticides singly.
+The effects of factors such as recombination, effective dominance, initial value of allele frequencies and initial value of linkage disequilibrium are considered.
+It is shown that the use of mixtures is always more effective in delaying the onset of resistance, often by many orders of magnitude.
+It is shown that there exists a threshold value of recombination fraction, above which the evolution of resistance is extremely slow.
+Resistance evolves very rapidly for values of recombination fraction below the threshold.
+Finally, the relevance of these results on resistance management is discussed.
+
+~ is mixture always better than sole ?
+Yes see this in paper1_results_figs.Rmd :
+  mix_minus_soleI1 <- resistPointsMixI1_T["gen_cP0.5"] - resistPointsI1_T["gen_cP0.5"]
+  which( mix_minus_soleI1 < 0 ) 
+  #0
+  mix_minus_soleI2 <- resistPointsMixI2_T["gen_cP0.5"] - resistPointsI2_T["gen_cP0.5"]
+  which( mix_minus_soleI2 < 0 )
+
 ### TO DO
+
+~ confirm whether Ians statement is true on all sensi runs : does resistance always spread faster when an insecticide is alone to when it is in a mixture ?
+"resistance to an insecticide always spreads faster when that insecticide is deployed alone rather than as part of a mixture, irrespective of the relative frequencies of resistance to the two insecticides.""
+
+~ do I start figs at gen 0 or 1 ? First generation is 1.
+
+~ read through Ians comments on new version of paper
+
+~ create a user interface to represent Killeen graphical models
 

### 2016-05-16 paper2_resistance_mechanisms_mixtures.Rmd
add todos
adressing minor issues in check
updating paper2

@@
 
+```{r,echo=FALSE,results='hide'}
+  library(resistance)
+```
+
 ### version 1
 
-# The development of insecticide resistance with use of insecticide mixtures and sequences : a simple theoretical exploration.
+# Mechanisms by which insecticide resistance can be influenced by strategies of insectcide use : a simple theoretical exploration.
 
@@
 
-The development of insecticide resistance is a population genetic process.
+The changing frequency of insecticide resistance is a population genetic process that can be influenced by :
 
-The development of resistance is expected to be influenced by :
-
-1. starting frequency of resistance alleles
-1. selection coefficient
-   + selective advatange of resistance in the presence of the insecticide
-1. dominance
-   + dominance of the resistance allele relative to the susceptible allele
-1. exposure
-   + proportion of the insect population that are exposed to the insecticide
-1. effectiveness
-   + proportion of exposed susceptible insects that are killed by the insecticide
+Parameter  | Description
+------------------- | -------------------------------------------------------
+1. Effectiveness | proportion of exposed insects with the susceptible allele that are killed by the insecticide
+2. Exposure | proportion of the insect population that are exposed to the insecticide
+3. Dominance | dominance of the resistance allele relative to the susceptible allele
+4. Selection Coefficient  | selective advatange of resistance in the presence of the insecticide
+5. Frequency  | frequency of resistance alleles within the population
 
@@
 
-*Maybe start by exploring the influence of the parameters on sole use ?*
+*accessible summary of the model
 
@@
 
-```{r,echo=FALSE,results='hide'}
-  library(resistance)
-```
+*Maybe start by exploring the influence of the parameters on sole use *
+*could put here or in the discussion
+
+**single insecticide application effect on increase in resistance of increasing each parameter**
+
+Parameter  | effect on resistance | Mechanism
+------------------ | ------- | ---------------------------------------------------
+1. Effectiveness | faster | increased deaths of susceptibles
+2. Exposure | faster | increased deaths of susceptibles
+3. Dominance | faster | increased survival of resistants
+4. Selection Coefficient | faster | increased survival of resistants
+5. Frequency  | faster | less change needed to reach resistance threshold, and rate of change in resistance tends to increase with frequency
+
+*to help explain I could start from a scenario where the attributes of each insecticide are the same*
+*and then explore changing each in turn*
+*I could then work up to the curtis fig2 plot*
+
+**When all parameters are equal resistance arises slower for sequential than mixture.**
+**Effect of changing parameters on relative performance of mixture versus sequential strategy**
+
+Parameter  | relative performance mix versus seq | Mechanism
+------------------ | ------- | ---------------------------------------------------
+A. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
+1. Effectiveness | faster | ...
+2. Exposure | faster | ...
+3. Dominance | faster | ...
+4. Selection Coefficient | ...
+5. Frequency  | faster | ...
+
 ### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
@@
 
+## Discussion
+
+(@)  In much of parameter space resistance changes more slowly when insecticides are used in a mixture relative to sequential use. The mechanism explaining this is ...
+
+(@) For a mixture of 2 insecticides with differing effectiveness.
+Resistance to the insecticide with a greater effectiveness increases faster.
+The less effective insecticide is 'protected' by the more effective.
+Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
+
+(@) When the effectiveness of 2 insecticides in a mixture is similar, the more effective insecticide provides little 'protection' to the less effective.
+In this case resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence :
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500, P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.17 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.75 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.23 , correctMixDeployProp = 1 , addCombinedStrategy = FALSE )
+```
+
+## References
+1. GPIRM
+1. Curtis 
+1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
+1. Birget 2015
+
 ## Todo
 
-1. look more at GPIRM
-1. mention Curtis
+1. re-read Mani 1985
+1. re-read GPIRM
+1. mention Curtis but don't depend on it
 1. do I want y axis of resistance frequency plots to be on a log scale ?
 1. can I develop explanatory diagrams to describe the input parameters ?
+1. decide on base starting scenarios and how to modify params from that
+1. mention generation time of 1 month, maybe even rescale axes to years ?
+1. look at effect of sex linkage

### 2016-05-16 paper2_resistance_mechanisms_mixtures.Rmd
progress

@@
 *I could then work up to the curtis fig2 plot*
+*could in each case incease value of I1 by 0.2 or some other value*
 
@@
 A. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
-1. Effectiveness | faster | ...
+1. Effectiveness | mix > seq | resistance to the more effective insecticide increases ...
 2. Exposure | faster | ...

### 2016-05-16 resistanceModelNotesAndy.Rmd
updating paper2
progress

@@
 
+Update Report for Ian :
+
+Tsetse : 15 days remaining : defined things to do for Steve
+Resistance : 55 days (from 85) remaining
+
 ### TO DO
+
+~ set up a new UI allowing comparison of 2 model runs side by side (like my curtis one but be able to modify params for both plots)
+~~ quickest would just be to add extra inputs on bottom
 
@@
 https://andysouth.shinyapps.io/shinyFig2Curtis
+
+# time
+To the end of Feb :
+Hours		105:45:00
+Hours rounded		105.00
+Days (/7.5)		14.00
+Days remaining (from 85)		71.00
+
+march, april :
+Hours 124.5
+Days (/7.5)		16.6
+Days previous 71
+Days remaining (from 85)		71-16.6 = 54.4

### 2016-05-17 paper2_resistance_mechanisms_mixtures.Rmd
start from all params equal

@@
 
-*accessible summary of the model
+*accessible summary of the model*
 
@@
 
-**When all parameters are equal resistance arises slower for sequential than mixture.**
+**xA. When all parameters are equal resistance arises slower for sequential than mixture.**
+This is because resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
+When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+```
+
 **Effect of changing parameters on relative performance of mixture versus sequential strategy**
@@
 ------------------ | ------- | ---------------------------------------------------
-A. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
-1. Effectiveness | mix > seq | resistance to the more effective insecticide increases ...
-2. Exposure | faster | ...
-3. Dominance | faster | ...
-4. Selection Coefficient | ...
-5. Frequency  | faster | ...
+xA. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
+x1. Effectiveness | mix > seq | resistance to the more effective insecticide increases ...
+x2. Exposure | faster | ...
+x3. Dominance | faster | ...
+x4. Selection Coefficient | ...
+x5. Frequency  | faster | ...
 

### 2016-05-18 paper2_resistance_mechanisms_mixtures.Rmd
developing mechanisms narrative

@@
 
-*Maybe start by exploring the influence of the parameters on sole use *
-*could put here or in the discussion
-
-**single insecticide application effect on increase in resistance of increasing each parameter**
-
-Parameter  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | faster | increased deaths of susceptibles
-2. Exposure | faster | increased deaths of susceptibles
-3. Dominance | faster | increased survival of resistants
-4. Selection Coefficient | faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold, and rate of change in resistance tends to increase with frequency
-
-*to help explain I could start from a scenario where the attributes of each insecticide are the same*
-*and then explore changing each in turn*
-*I could then work up to the curtis fig2 plot*
-*could in each case incease value of I1 by 0.2 or some other value*
+* explore the influence of the parameters on sole use either here or in the discussion *
 
@@
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 1. All control inputs equal at 0.5")
 ```
 
-**Effect of changing parameters on relative performance of mixture versus sequential strategy**
+**x1. Increasing the effectiveness of insecticides increases time-to-resistance for mixtures**
+A more effective insecticide prompts increased time-to-resistance when used in a mixture.
+This contrasts with the earlier result showing that when used in isolation a single insecticide prompts a shorter time-to-resistance.
 
-Parameter  | relative performance mix versus seq | Mechanism
------------------- | ------- | ---------------------------------------------------
-xA. All equal | seq > mix | resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence
-x1. Effectiveness | mix > seq | resistance to the more effective insecticide increases ...
-x2. Exposure | faster | ...
-x3. Dominance | faster | ...
-x4. Selection Coefficient | ...
-x5. Frequency  | faster | ...
+In a mixture the mechanism is that a more effective insecticide provides greater protection to the other.
+For example taking the previous scenario and increasing the effectiveness of one insecticide from 0.5 to 0.9 increases the time-to-resistance for the 2nd insecticide from around 45 to 75 generations (Fig.
+2).
+In the plot you can see that resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to Fig.2.
+In contrast resistance to insecticide2 increases more slowly initially.
+Once insecticide1 reaches it's resistance threshold (around 40 generations), the curve for insecticide2 becomes steeper.
+This can be explained by the mechanism that once insecticide1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide2 and thus stops 'protecting' it from it's own rise in resistance.
 
-### Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 2. Effectiveness of insecticide1 increased from 0.5 to 0.9")
+```
+
+When the effectiveness of both insecticides are increased to 0.9 the time-to-resistance is increased still further to around 125 generations (Fig.
+3).
+In this case the curves for the rise in resistance to each insecticide are identical.
+The curves are shallower than previous plots because individuals that are resistant to each insecticide are being killed by the other thus reducing the increase in resistance alleles.
+As the resistance frequencies go above 10% around generation 100 both insecticides provide less protection to the other and the slope of the resistance curve increases.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 3. Effectiveness of both insecticides 0.9")
+```
+
+Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
+Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.9 decreases the time-to-resistance to insecticide1 from 40 to 30 generations but leaves the time-to-resistance for the mixture the same (Fig.
+4).
+Setting the dominance for bothe alleles to 0.9 reduces the time-to-resistance for the mixture to 30 generations (Fig.
+5).
+In both of these cases sequential use continues to have a longer time-to-resistance.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 4. Dominace of insecticide1 0.9")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 5. Dominace of both insecticides 0.9")
+```
+
+```{r, eval=FALSE, echo=FALSE, fig.height=4}
+### 3 examples showing how changing effectiveness can alter the ranking of insecticide strategies
+#old modification of curtis values
 #### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
 runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
@@
 
+```{r, eval=FALSE, echo=FALSE, fig.height=4}
+#old modification of curtis values
 #### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
 runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
@@
 
+```{r, eval=FALSE, echo=FALSE, fig.height=4}
+#old modification of curtis values
 #### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
 runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
@@
 
-(@)  In much of parameter space resistance changes more slowly when insecticides are used in a mixture relative to sequential use. The mechanism explaining this is ...
+**single insecticide**
+
+Parameter to increase  | effect on resistance | Mechanism
+------------------ | ------- | ---------------------------------------------------
+1. Effectiveness | faster | increased deaths of susceptibles
+2. Exposure | faster | increased deaths of susceptibles
+3. Dominance | faster | increased survival of heterozygotes
+4. Selection Coefficient | faster | increased survival of resistants
+5. Frequency  | faster | less change needed to reach resistance threshold
+
+**mixture**
+
+Parameter to increase  | effect on resistance | Mechanism
+------------------ | ------- | ---------------------------------------------------
+1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
+2. Exposure | **slower** | increased deaths of individuals resistant to the other insecticide
+3. Dominance | ?faster | increased survival of resistants
+4. Selection Coefficient | ?faster | increased survival of resistants
+5. Frequency  | ?faster | less change needed to reach resistance threshold
+
+**difference between mixture and sequential use**
+
+Parameter to increase  | effect on difference | Mechanism
+------------------ | ------- | ---------------------------------------------------
+1. Effectiveness | improves mixture | faster resistance for single use and slower resistance in mixture
+2. Exposure | improves mixture | faster resistance for single use and slower resistance in mixture
+3. Dominance | ? | ?
+4. Selection Coefficient | ? | ?
+5. Frequency  | ? | ?
+
+or like this ?
+
+**Summary of the effect of changing parameters on relative performance of mixture versus sequential strategy**
+
+Parameter  | relative performance mix versus seq | Mechanism
+------------------ | ------- | ---------------------------------------------------
+x1. High Effectiveness | mix > seq | A highly effective insecticide in a mixture reduces the increase in resistance of the other. 
+x2. High Exposure | mix > seq | High exposure is needed for one insecticide in the mixture to reduce the increase of the other.
+x3. Dominance | ... | ...
+x4. Selection Coefficient | ...
+x5. Frequency  | faster | ...
+
+(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
 
@@
 Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
-
-(@) When the effectiveness of 2 insecticides in a mixture is similar, the more effective insecticide provides little 'protection' to the less effective.
-In this case resistance to the two insecticides in a mixture increases at a similar rate and is reached sooner than if the insecticides are applied individually in sequence :
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500, P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.17 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.75 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.23 , correctMixDeployProp = 1 , addCombinedStrategy = FALSE )
-```
 
@@
 
-1. re-read Mani 1985
+1. create some code to enable plotting 2 scenarios side-by-side
 1. re-read GPIRM
-1. mention Curtis but don't depend on it
-1. do I want y axis of resistance frequency plots to be on a log scale ?
 1. can I develop explanatory diagrams to describe the input parameters ?
-1. decide on base starting scenarios and how to modify params from that
-1. mention generation time of 1 month, maybe even rescale axes to years ?
+1. mention generation time of 1 month
 1. look at effect of sex linkage

### 2016-05-19 paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 We describe an R package and user-interface that allows further exploration of the development of resistance.
+The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.
 
@@
 (in press) describe the technical details of the model and explore large parameter ranges.
-In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the devlopment of resistance.
+In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the development of resistance.
 
@@
 
-* explore the influence of the parameters on sole use either here or in the discussion *
+With all control inputs set equal at 0.5 resistance arises slower for sequential than mixture (Fig.
+1).
+Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
+When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
 
-**xA. When all parameters are equal resistance arises slower for sequential than mixture.**
-This is because resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
-When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
@@
 
-**x1. Increasing the effectiveness of insecticides increases time-to-resistance for mixtures**
+**Effectiveness of the insecticides**
 A more effective insecticide prompts increased time-to-resistance when used in a mixture.
@@
 
+**Dominance of resistance alleles**
 Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
@@
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 4. Dominace of insecticide1 0.9")
+cat("Fig. 4. Dominance of insecticide1 0.9")
 ```
@@
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 5. Dominace of both insecticides 0.9")
+cat("Fig. 5. Dominance of both insecticides 0.9")
+```
+
+**Selective advantage of resistance**
+Decreasing the selective advantage of resistance to insecticide1 from 0.5 to 0.1 increases it's time-to-resistance in the mixture from 40 generations to 150, and has a similar effect on it's use in isolation (Fig.
+6).
+However, it has little effect on resistance to insecticide2.
+Because of this sequential use continues to generate a longer time-to-resistance than the mixture.
+Decreasing the selective advantage of both alleles to 0.1 results in a longer time-to-resistance of 200 generations for the mixture (Fig.
+7), however the relative shape of the resistance curves remains the same as in Fig.
+1, with sequential use continuing to generate longer time-to-resistance.
+
+[Increasing the selective advantage of resistance above 0.5 causes an error about fitness values above 1. How to explain this ?]
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.1 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 6. Selective advantage of insecticide1 0.1")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.1 , s.RR2_0B = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 7. Selective advantage of both insecticides 0.1")
+```
+
+Taking this scenario of lower selective advantage and increasing the effectiveness of both insectides from 0.5 to 0.7 shows again how the change in effectiveness has a greater effect on the mixture than sequential making the mixture preferable (compare Fig.
+8 to Fig.
+7).
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.1 , s.RR2_0B = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 8. Selective advantage of both insecticides 0.1, effectiveness to 0.7")
+```
+
+**Exposure**
+Increasing exposure to 0.9 reduces time-to-resistance for both mixture and sequential scenarios (Fig.
+9).
+The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
+Setting the effectiveness of both insecicides to 0.7 (while keeping all other inputs at 0.5) results in mixture being slightly favoured (Fig.
+10), then making the increase in exposure from 0.5 to 0.9 results in a switch to sequential being the favoured strategy (Fig.
+11).
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 9. Exposure 0.9")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 10. Effectiveness 0.7, sequence favoured")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. 11. Effectiveness 0.7, Exposure 0.9, mixture favoured")
 ```
@@
 3. Dominance | faster | increased survival of heterozygotes
-4. Selection Coefficient | faster | increased survival of resistants
+4. Selective advantage | faster | increased survival of resistants
 5. Frequency  | faster | less change needed to reach resistance threshold
@@
 1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
-2. Exposure | **slower** | increased deaths of individuals resistant to the other insecticide
-3. Dominance | ?faster | increased survival of resistants
-4. Selection Coefficient | ?faster | increased survival of resistants
+2. Exposure | ??faster | ??increased deaths of individuals resistant to the other insecticide
+3. Dominance | faster | increased survival of resistants
+4. Selective advantage | ?faster | increased survival of resistants
 5. Frequency  | ?faster | less change needed to reach resistance threshold
@@
 
-Parameter to increase  | effect on difference | Mechanism
+Parameter to increase  | relative performance mix versus seq | Mechanism
 ------------------ | ------- | ---------------------------------------------------
-1. Effectiveness | improves mixture | faster resistance for single use and slower resistance in mixture
-2. Exposure | improves mixture | faster resistance for single use and slower resistance in mixture
+1. Effectiveness | mixture | faster resistance for single use and slower resistance in mixture
+2. Exposure | sequence | faster resistance for mixture and sequence but has a greater effect on mixture
 3. Dominance | ? | ?
-4. Selection Coefficient | ? | ?
+4. Selective advantage | ? | ?
 5. Frequency  | ? | ?
 
-or like this ?
+(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
 
-**Summary of the effect of changing parameters on relative performance of mixture versus sequential strategy**
-
-Parameter  | relative performance mix versus seq | Mechanism
------------------- | ------- | ---------------------------------------------------
-x1. High Effectiveness | mix > seq | A highly effective insecticide in a mixture reduces the increase in resistance of the other. 
-x2. High Exposure | mix > seq | High exposure is needed for one insecticide in the mixture to reduce the increase of the other.
-x3. Dominance | ... | ...
-x4. Selection Coefficient | ...
-x5. Frequency  | faster | ...
-
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
+(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
+In contrast the dominance and selective advantage of one has less effect on the resistance of the other.
+In this paper we have kept the exposure to each insecticide the same (to represent a mixture in which the 2 insecticides are applied together), ??
+however we predict that if they were applied separately then (similar to effectiveness) a higher exposure to one insecticide would increase the time-to-resistance of the other.
 
@@
 
+1.
+work out what's going on with exposure ...
+why does increasing exposure have a greater effect on mixture than sequential ?
+Is it because the exposure to one insecticide also effects the development of resistance in the other ?
 1. create some code to enable plotting 2 scenarios side-by-side

### 2016-05-20 paper2_resistance_mechanisms_mixtures.Rmd
version sent to Ian

@@
 ## Results
+
+*Before this add figures for effect of inputs on sole use.*
 
@@
 11).
+This demonstrates how higher exposures can favour sequential use.
 
@@
 1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
-2. Exposure | ??faster | ??increased deaths of individuals resistant to the other insecticide
+2. Exposure | faster (but not as fast as for single) | increased deaths of individuals resistant to the other insecticide
 3. Dominance | faster | increased survival of resistants
 4. Selective advantage | ?faster | increased survival of resistants
-5. Frequency  | ?faster | less change needed to reach resistance threshold
+5. Frequency  | faster | less change needed to reach resistance threshold
 
@@
 (@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
-In contrast the dominance and selective advantage of one has less effect on the resistance of the other.
-In this paper we have kept the exposure to each insecticide the same (to represent a mixture in which the 2 insecticides are applied together), ??
-however we predict that if they were applied separately then (similar to effectiveness) a higher exposure to one insecticide would increase the time-to-resistance of the other.
+In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
+For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
+However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
+
+(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
+Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
 
@@
 The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
+Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level at which point it starts to increase faster.
+
+(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
+In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
+The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
+Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
+This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
 

### 2016-05-20 inst/documents/effectOfIndividualInputs.Rmd
moved documents into own folder

--- 2016-02-01 effectOfIndividualInputs.Rmd
@@
 
-  #ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap( ~ cut)
+```
 
-  #cool working!
+# Plot time to resistance for each strategy using all sensitivity analysis values
+```{r, eval=TRUE, echo=FALSE, warning=FALSE}
 
-  #exposure has the most noticeable effect
-  ggplot(ggInsOuts, aes(x=exposure, y=time_to_resistance0.25, color=strategy)) + 
-    geom_point(shape=3) + 
-    facet_wrap( ~ strategy) + 
-    #geom_smooth(aes(x=exposure, y=gen_cP0.5, color=NULL))
-    geom_smooth(colour='red', linetype='dashed',size=0.5)
+names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
 
-  #effectiveness of insecticide1 little noticeable effect
-#   ggplot(ggInsOuts, aes(x=effectiveness_insecticide1, y=time_to_resistance0.25, color=strategy)) + 
-#     geom_point() + 
-#     facet_wrap( ~ strategy) + 
-#     #geom_smooth(aes(x=effectiveness_insecticide1, y=gen_cP0.5, color=NULL)) 
-#     geom_smooth(colour='red', linetype='dashed',size=0.5)
+for(i in names_results)
+{
+  #x11()
+
+  print( ggplot(ggInsOuts, aes_string(x='strategy',y=i, color='strategy')) + 
+      #ylim(0,450) +
+      coord_cartesian(xlim=c(1, 6), ylim=c(0, 350)) +
+      geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE)
+  )#end print
+}
+
+  #can I facet by exposure ?
+  #exposure <- runif(1, min=0.1, max=0.9)
+  #why does it not look uniform ? must be because runs have been removed if too high
+  hist(ggInsOuts$exposure)
+  #create exposure bins
+  ggInsOuts$exposure_cat <- cut(ggInsOuts$exposure,breaks=c(0.1,0.3,0.5,0.7,0.9), labels=c("exposure 0.1-0.3","exposure 0.3-0.5","exposure 0.5-0.7","exposure 0.7-0.9"))
+  #or just 2
+  #ggInsOuts$exposure_cat <- cut(ggInsOuts$exposure,breaks=c(0.1,0.5,0.9), labels=c("exposure 0.1-0.5","exposure 0.5-0.9"))
+  #actually might be better to show just the 2 extremes
+  #ggSubset <- subset(ggInsOuts, exposure_cat==c("exposure 0.1-0.3","exposure 0.7-0.9"))
+
+  print( ggplot(ggInsOuts, aes_string(x='strategy',y=i, color='strategy')) + 
+      #ylim(0,450) +
+      #scale_y_continuous( limits = c( 0,450 ) ) +
+      #to zoom in without changing data  
+      coord_cartesian(xlim=c(1, 6), ylim=c(0, 350)) +
+      facet_wrap( ~ exposure_cat, scales='fixed') +  
+      geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) #, scale='count')
+  )#end print
+
+# or trying as a dotplot, with these options it's very similar to violin
+#   print( ggplot(ggInsOuts, aes_string(x='strategy',y=i, color='strategy')) + 
+#       #ylim(0,450) +
+#       scale_y_continuous( limits = c( 0,450 ) ) +
+#       facet_wrap( ~ exposure_cat, scales='fixed') +  
+#       geom_dotplot(stackdir = 'center', binaxis='y', binwidth = 5 ,dotsize = 0.1, show.legend=FALSE) #, scale='count')
+#   )#end print
+
+#now try the same faceting by effectiveness_insecticide1
+#doesn't really show much ...
+#so why does effectiveness come out so important in the trees ??  
+  #hist(ggInsOuts$effectiveness_insecticide1)  
+  #create effect1 bins
+  ggInsOuts$effect1_cat <- cut(ggInsOuts$effectiveness_insecticide1,breaks=c(0.4,0.6,0.8,1), labels=c("effect1 0.4-0.6","effect1 0.6-0.8","effect1 0.8-1"))
+
+  print( ggplot(ggInsOuts, aes_string(x='strategy',y=i, color='strategy')) + 
+      #ylim(0,450) +
+      #scale_y_continuous( limits = c( 0,450 ) ) +
+      #to zoom in without changing data  
+      coord_cartesian(xlim=c(1, 6), ylim=c(0, 250)) +
+      facet_wrap( ~ effect1_cat, scales='fixed', nrow=1) +  
+      geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) #, scale='count')
+  )#end print
+
+## BEWARE what am I doing with putting mixture better than seq on each repeated input ....
+
+## !!try facetting by whether mixture or sequence better
+## initially just with the time to resistance
+## relies on ggInsOuts being prepared slightly differently from sensiAnPaper1All
+## aha! this is weird it actually shows much shorter times for mixture1, & mix2 but sequential fairly similar  
+#   print( ggplot(ggInsOuts, aes_string(x='strategy',y=i, color='strategy')) + 
+#       #ylim(0,450) +
+#       #scale_y_continuous( limits = c( 0,450 ) ) +
+#       #to zoom in without changing data  
+#       coord_cartesian(xlim=c(1, 6), ylim=c(0, 250)) +
+#       facet_wrap( ~ betterMix2Seq20_cP0.1, scales='fixed', nrow=1) +  
+#       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) #, scale='count')
+#   )#end print
+
+# or I could look at the values for exposure on the y axis
+#   print( ggplot(ggInsOuts, aes_string(x='strategy',y='exposure', color='strategy')) + 
+#       #ylim(0,450) +
+#       #scale_y_continuous( limits = c( 0,450 ) ) +
+#       #to zoom in without changing data  
+#       #coord_cartesian(xlim=c(1, 6), ylim=c(0, 250)) +
+#       facet_wrap( ~ betterMix2Seq20_cP0.1, scales='fixed', nrow=1) +  
+#       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) #, scale='count')
+#   )#end print
 
@@
 
-## In the following plots all inputs are plotted individually, the points are coloured by exposure, which allows us to see that, even as other inputs change, low times to resistance are associated with high exposure.
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+#trying a PRCC analysis using the package sensitivity
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
 
@@
     }
-
-  names_inputs <- names(ggInsOuts)[1:num_inputs]
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- 'time_to_resistance0.25'    
-    #y <- 'time_to_resistance0.5'
-    #color <- 'strategy'
-    #coloring by exposure shows its effect even when other params are varying
-    color <- 'exposure'
-
-    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-             geom_point(shape=3) + 
-             facet_wrap( ~ strategy) + 
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             labs(title = i))
-  }
-
-  #can I facet by particular parameter ranges ?
-  #might need to create a categorical var from the range and use that to facet 
-  #or subset the data for specific parameter ranges
-
-```
-
-## Looking more closely at effect of the ratio between the level of resistance to the first insecticide divided by that of the second one. Restricting the x axis to lower values.
-```{r, eval=TRUE, echo=FALSE, warning=FALSE}
-
-  i <- 'resist_start_1_div_2'
-  color <- "start_freq_allele2"
-
-  print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-           geom_point(shape=3) + 
-           facet_wrap( ~ strategy) + 
-           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-           geom_smooth(colour='red', linetype='dashed',size=0.5) +
-           labs(title = i) +
-           xlim(0,100)
-         )
-
-```
-
-#trying a PRCC analysis using the package sensitivity
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
 
@@
 
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, fig.height=3.5}
+## trying to highlight differences between sequential and mixture strategies in response of time-to-resistance to inputs
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
 
-#trying to find apparent anomaly in PRCC analysis
-#what causes correctMixDeployProp to get a high PRCC score for I1 & 2 when its not used in those runs ??
+  names_inputs <- names(ggInsOuts)[1:num_inputs]
 
-tstI1 <- round(listOutI1$input,3)
-tstMix <- round(listOutMix$input,3)
-tstDif <- round(tstMix - tstI1,3)
+  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
+
+  for(i in names_inputs)
+  {
+    #x11()
+
+    y <- 'time_to_resistance0.25'
+
+    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
+    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
+            #points not wanted if 10000
+             #geom_point(shape=3, show.legend=FALSE) + 
+             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+             geom_smooth(linetype='dashed',size=1.5) +
+             #facet_wrap( ~ strategy) +
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             labs(title = i))
+  }
 
 ```
+
+## response of resistance thresholds to individual inputs faceted by strategy
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
+  names_inputs <- names(ggInsOuts)[1:num_inputs]
+
+  #todo find out to take sample this doesn't work
+  #ggSample <- sample(ggInsOuts,500)
+
+  for(i in names_inputs)
+  {
+    #x11()
+
+    y <- 'time_to_resistance0.25'    
+    #y <- 'time_to_resistance0.5'
+    color <- 'strategy'
+    #coloring by exposure shows its effect even when other params are varying
+    #color <- 'exposure'
+
+    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
+             #points not wanted if 10000
+             #geom_point(shape=3, show.legend=FALSE) + 
+             geom_smooth(colour='red', linetype='dashed',size=0.5, show.legend=FALSE) +
+             facet_wrap( ~ strategy) + #, show_guide = FALSE) + 
+             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+             labs(title = i))
+  }
+
+```
+
+## Looking more closely at ratio between the level of resistance to the first insecticide divided by that of the second one. Restricting the x axis to lower values.
+## ?not very useful
+```{r, eval=FALSE, echo=FALSE, warning=FALSE}
+
+  i <- 'resist_start_1_div_2'
+  color <- "start_freq_allele2"
+
+  print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
+           geom_point(shape=3) + 
+           facet_wrap( ~ strategy) + 
+           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+           geom_smooth(colour='red', linetype='dashed',size=0.5) +
+           labs(title = i) +
+           xlim(0,100)
+         )
+
+```
+
+##

### 2016-05-20 inst/documents/keep_old_insecticide.Rmd
moved documents into own folder

--- 2016-04-16 paper1_results_figs.Rmd
@@
 ---
-  title: "paper1_results_figs"
+  title: "keep on an old insecticide in a mix ?"
   output: pdf_document
@@
 
-# Figures for paper 1 
-# version 5 - in progress
+## new analysis April 2016 for paper1 : v2
 
-## Based on 10000 runs with all inputs varying at once which explains the variability at particular input values.
+Is keeping on an old insecticide in a mix better than just using a new insecticide on its own ?
 
-Comparing these insecticide use strategies:
+Short answer : Yes
 
-1. Single use
-1. Mixture (either AI) : threshold reached for either insecticide
-1. Adaptive mixture : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
-1. Mixture (both AIs) : threshold reached for both insecticides in mixture
-1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold
+(part copied from paper1_results_figs.Rmd)
 
-Renamed from previous naming system :
-Mixture (either AI)   [mixture1]
-Mixture (both AIs)    [mixture3]
-Adaptive mixture      [mixture2]
+This analysis uses previous sensitivity analysis runs :
+1. select out runs where the starting frequency of resistance to I1 > I2
+1. I didn't yet restrict to runs where there was a 10 fold difference in starting frequencies, seemed not to be necessary.
 
-## Notable results
-
-1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
+# Notable results
+1. it is always better to keep on the old insecticide rather than using the new insecticide on its own.
 1.
-In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture switching to sole use.
-The latter takes slightly longer to reach the threshold.
-1. Exposure and effectiveness seem to be the most important inputs.
-1. Exposure has the greatest effect on time-to-resistance.
-1.
-Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) adaptive mixture switching to sole use.
-Below an effectiveness of ~0.55 to 0.6 sequential takes longer, and above this 'mixture switching to sole use' takes longer.
+exposure (to both insecticides) and effectiveness of the old insecticide have the main effect on how much better the mixture is over sole use.
+Exposure has a positive effect, and effectiveness a negative effect on the benefit of the mixture.
+1. dominance and selection coefficient for the new insecticide have a lesser effect (also negative) on the benefit of the mixture
 
@@
 library(resistance)
+library(ggplot2)
 
@@
 dif_oldnew_newonly["mix_minus_sole0.5"] <- dif_oldnew_newonly["gen_cP0.5"] - dif_oldnew_newonly["gen_cP0.5seq"]
+# %add_div
+dif_oldnew_newonly["mix_divby_sole0.5"] <- dif_oldnew_newonly["gen_cP0.5"] / dif_oldnew_newonly["gen_cP0.5seq"]
 
@@
 
----
+\pagebreak
 
-# Fig. x1 Time-to-resistance across all input values
+## Fig. xx1 Recreating analysis from Curtis Table 1 (vi)
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE}
+```{r, eval=TRUE, echo=FALSE, fig.height=4.5}
 
-  library(ggplot2)
-
-#names_results <- c("time_to_resistance0.5","time_to_resistance0.25","time_to_resistance0.1")
-names_results <- c("time_to_resistance0.25")
-
-ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 2"), ]
-
-for(i in names_results)
-{
-  #x11()
-
-  print( ggplot(ggSubset, aes_string(x='strategy',y=i, color='strategy')) + 
-      #ylim(0,450) +
-      coord_cartesian( ylim=c(0, 350)) +
-      geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
-      #scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Sole use", "Mix either"="1st in mix","Sequential"="2nd in sequence", "Mix adaptive"="2nd in mix\nswitch to sole", "Mix both"="2nd in mix" ) )
-      scale_x_discrete("Resistance to which insecticide", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Sole use", "Mix either"="1st in mix","Sequential"="2nd in sequence", "Mix adaptive"="adaptive mix", "Mix both"="2nd in mix" ) )
-
-  ) #end print
-}
-
+runcurtis_f2( P_1 = 0.001,
+P_2 = 0.01,
+# dominances
+h.RS1_A0 =  1,
+h.RS2_0B =  1,
+# exposures
+exposure =  0.9,
+# effectivenesses
+phi.SS1_A0 =  1,
+phi.SS2_0B =  1,
+# selection coefficients
+s.RR1_A0 =  1,
+s.RR2_0B =  1,
+addCombinedStrategy=FALSE,
+addStrategyLabels = FALSE )
 ```
 
-## Fig. x2 How time-to-resistance is effected by input values for each insecticide use strategy
+This seems to show that resistance to the new insecticide (red) rises faster when alone (dotted) than when in combination (solid) with an old insecticide at a higher frequency.
+This is contrary to what Curtis said.
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+Also note that resistance arises much faster than in our sensitivity runs. I think  this is because the selection coefficients (1) are much higher than the range we use (0.1-0.45).
 
-  names_inputs <- names(ggInsOuts)[1:num_inputs]
-
-  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- 'time_to_resistance0.25'
-
-    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
-    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
-            #points not wanted if 10000
-             #geom_point(shape=3, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             geom_smooth(linetype='dashed',size=1) +
-             #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i) +
-             #add line for Curtis inputs
-             geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
-             #tryting to get the vline into the legend fails
-             #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
-
-           )
-  }
-
-```
-
-## Fig. x3 How difference in time-to-resistance between sequential and adaptive mixture is effected by input values
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
-  ggSubset <- dif_mixA_seq
-
-  names_inputs <- names(ggSubset)[1:num_inputs]
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- "mixA_minus_seq0.25"
-
-    print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
-            #points not wanted if 10000
-             geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             geom_smooth(linetype='dashed',size=1, colour="red") +
-             #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i) +
-             geom_hline(yintercept = 0, colour="white")
-           )
-  }
-
-```
+\pagebreak
 
@@
 
+#mix_divby_sole0.5
+for(i in names_inputs)
+{
+  #x11()
+
+  y <- "mix_divby_sole0.5"
+
+  print( ggplot(ggSubset, aes_string(x=i, y=y)) + #, colour="strategy")) + 
+           geom_point(shape=1, colour="darkblue", alpha=0.2, show.legend=FALSE) + 
+           #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+           geom_smooth(linetype='dashed',size=1, colour="red") +
+           #facet_wrap( ~ strategy) +
+           #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
+           labs(title = i) +
+           geom_hline(yintercept = 0, colour="white")
+  )
+}
+
 ```
+
+\pagebreak
 
@@
 x <- dif_oldnew_newonly[,1:num_inputs]
+
 y <- dif_oldnew_newonly["mix_minus_sole0.5"]
@@
 pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
 
@@
          ylim(-1,1) +
-             ggtitle(paste("PRCC benefit of mix old+new over sole new (old=1, new=2)")) +
+         ggtitle(paste("Difference time-to-resistance for mix old+new minus new alone (old=1, new=2)")) +
          xlab(NULL)
 )
-```
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
-  #can I do a 2D hexbin plot with x=exposure, y=effectiveness
-  #but do I just want to use 1 effectiveness, or maybe add both effectivenesses together ?
-  #bin colour : proportion of runs in which sequential better & could have diverging scale about 0.5
-  #I probably still need to tweak data first
-  #currently I have differences in times to resistance : mixA_minus_seq0.25
-  #what I think I want is the proportion of runs in which mix is better
-  #I can get that by setting to 0 & 1 and calc average
-  # tst_runs <- c(1,0,0,0)
-  # tst_runs <- c(1,1,1,0)
-  # mean(tst_runs)
-
-  ##BEWARE toogle this to look at plots for mixboth vs seq
-  ggSubset <- dif_mixA_seq
-  #ggSubset <- dif_mix2_seq
-
-  #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
-  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
-  ggSubset$abs_mixA_minus_seq0.25 <- abs(ggSubset$mixA_minus_seq0.25)
-
-  #tricky
-  #be a bit careful what to do with zeros where results are same for both strategies
-  #I use these better vars to facet the plots
-  #may want the zeroes in both facets
-  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
-  #set these to 1 (they were 0)
-  df_same$mixA_better_seq0.25 <- 1
-  df_same$seq_better_mixA0.25 <- 1
-  #rbind them back onto other dataset, probably not needed ??
-  ggSubset <- rbind( ggSubset, df_same)
-
-  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
-
-    #failed hex plot just gave me counts
-    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
-    #          geom_hex()
-    #        )
-
-    #interesting plot showing where sequence is better is restricted to higher values of exposure and effectiveness 
-    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
-             #geom_hex()
-             geom_point(alpha=0.2)  
-           )
-
-    #can I multiply the two effectivenesses together ?
-    #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_better_seq0.25))+
-             #geom_hex()
-             geom_point(alpha=0.2)  
-           )
-
-    #i think adding the effectivenesses may give an even better plot
-    #and makes it easier for the reader to interpret
-    #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mixA_better_seq0.25))+
-
-             geom_point(alpha=0.2)  
-           )
-
-    #facetted 0/1, labelled mix better, seq better
-    #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mixA0.25))+
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
-
-             theme(legend.position="none") +
-             geom_point(alpha=1)  
-           )
-
-    #what if I do the difference rather than just 0/1
-    #aha! this is close to being the killer plot, just need to get the colour scale right 
-    #but the previous one may be clearer
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_minus_seq0.25))+
-             #geom_hex()
-             geom_point(alpha=1) +
-             #no adding grey outlines obscures the pattern
-             #geom_point(shape=21, colour='grey') +
-             #this one does look good but slight potential to see negatives above even though they aren't there
-             #could I modify the shapes to be different
-             #would a repeated, faceted plot solve that ?
-             #scale looks good but colours could do with being denser
-             #and I'm not sure that making the points that are close to zero white is a good idea ?
-             #although it does mean points that are only a few generations different do not obscure the pattern
-             #maybe I could explicitly do Ians 20% thing & colour those white
-             scale_colour_gradient2()
-
-             #didn't manage to get the colorBrewer scale to diverge from 0
-             #scale_colour_distiller("RdYlGn",type='div')
-            )
-
-    #COOLIO !! I really like this figure.
-    library(viridis)
-    #facet showing generation differences, labelled mix better, seq better
-    #+ve & -ve on the same colour scale
-    #failed attempt at getting Curtis point in legend
-    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
-
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
-
-             #YlGnBu looks really good except that the yellow points are very faint
-             #scale_colour_distiller(palette='YlGnBu', direction=1) +
-             #option="plasma" looks good too
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
-             theme_bw() +
-             #theme_minimal() +
-             #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             #failing to add legend for curtis point
-             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
-             #scale_size_manual(name = "Curtis Fig.2", values="test") +
-             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) + 
-             #add text for Curtis point
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
-           )
-
-    #NOW what if I subset the 0 values & fit a linear model to it
-    #could that be used to decide between the 2 strategies ?
-    #I might even be able to do the lm plot within ggplot
-
-    #trickiness I have is that I have previously put 0 values in the adptive mix better plot 
-    #which is maybe a little misleading anyway
-    #and also means that I only get the lm line in the one plot
-    #maybe what I can do is add the zeros into both plots then  will get lines in both
-    #might even be good to overplot the zero values after
-    #or even remove the zero values ?
-
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
-
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
-
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
-             theme_bw() +
-             #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) + 
-             #!!!NOW can I add a lm fitted to the 0 points mixA_minus_seq0.25
-             geom_smooth(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], method = "lm", se = FALSE) +
-             #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
-             #add text for Curtis point
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
-           )
-
-  # doing the lm outside of ggplot    
-  df_same = ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
-  x1=df_same$exposure 
-  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
-
-  l_mod <- lm(y1 ~ x1)
-  l_mod$coefficients
-  #(Intercept)           x 
-  #  0.9790948   0.6095132
-
-  x2 <- seq(0,1,0.1)
-  #approximation
-  y2 <- 0.6*x2+0.98
-  #plot(x2,y2)
-
-  #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
-
-             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
-
-             theme_bw() +
-             #add point for Curtis inputs
-             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) +
-             #add a lm fitted to the 0 points
-             geom_smooth(data=df_same, method = "lm", se = FALSE) 
-           )
-
-```
-
-# repeating killer plot for mixtures both versus sequential
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
-  #try repeating the killer plot for mixtures both versus sequential
-
-  ##BEWARE toogle this to look at plots for mixboth vs seq
-  #ggSubset <- dif_mixA_seq
-  ggSubset <- dif_mix2_seq
-
-  #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mix2_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 > 0, 1, 0 )
-  ggSubset$seq_better_mix20.25 <- ifelse( ggSubset$mix2_minus_seq0.25 < 0, 1, 0 )
-  ggSubset$abs_mix2_minus_seq0.25 <- abs(ggSubset$mix2_minus_seq0.25)
-
-  #tricky
-  #be a bit careful what to do with zeros where results are same for both strategies
-  #I use these better vars to facet the plots
-  #may want the zeroes in both facets
-  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$mix2_minus_seq0.25==0),]
-  #set these to 1 (they were 0)
-  df_same$mix2_better_seq0.25 <- 1
-  df_same$seq_better_mix20.25 <- 1
-  #rbind them back onto other dataset, probably not needed ??
-  ggSubset <- rbind( ggSubset, df_same)
-
-  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mix2_not20pc_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
-
-    #i think adding the effectivenesses may give an even better plot
-    #and makes it easier for the reader to interpret
-    #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.25))+
-
-             geom_point(alpha=0.2)  
-           )
-
-    #facetted 0/1, labelled mix better, seq better
-    #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.25))+
-             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
-
-             theme(legend.position="none") +
-             geom_point(alpha=1)  
-           )
-
-    #killer plot mix2 vs seq
-    library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
-
-             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
-
-             #YlGnBu looks really good except that the yellow points are very faint
-             #scale_colour_distiller(palette='YlGnBu', direction=1) +
-             #option="plasma" looks good too
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
-             theme_bw() +
-             #theme_minimal() +
-             #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             #failing to add legend for curtis point
-             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
-             #scale_size_manual(name = "Curtis Fig.2", values="test") +
-             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) + 
-             #add text for Curtis point
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
-           )
-
-    #modify killer plot mix2 vs seq to add in male exposure prop on x axis
-    library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
-
-             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold")) + 
-             theme_bw() +
-             #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides for both sexes') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) + 
-             #add text for Curtis point
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
-           )
-
-  #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
-
-             facet_grid( ~ mix2_not20pc_better_seq0.25, labeller=as_labeller( c('0'='mix\nresistance >20% slower','1'='mix\nresistance <20% slower'))) +
-
-             theme_bw() +
-             #add point for Curtis inputs
-             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) +
-             #add a lm fitted to the 0 points
-             geom_smooth(data=df_same, method = "lm", se = FALSE) 
-           )
-
-```
-
-# Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
-### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-# PRCC difference in time-to-resistance between sequential and mixture adaptive
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
-
-  library(sensitivity)
-
-    x <- dif_mixA_seq[,1:num_inputs]
-    y <- dif_mixA_seq["mixA_minus_seq0.25"]
+#mix_divby_sole0.5
+y <- dif_oldnew_newonly["mix_divby_sole0.5"]
 
 pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
 
@@
          ylim(-1,1) +
-             ggtitle(paste("PRCC difference between sequential and mixture adaptive")) +
+         ggtitle(paste("Ratio of time-to-resistance for mix old+new over new alone (old=1, new=2)")) +
          xlab(NULL)
 )
-```
-
-# PRCC analysis for time-to-resistance subsetted by each insecticide use strategy
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
-
-  library(sensitivity)
-
-# pcc(X, y, rank = FALSE, nboot = 0, conf = 0.95)
-# 
-# Arguments
-# X a data frame containing model input variables
-# y a vector containing the responses 
-# rank logical. If TRUE, the analysis is done on the ranks.
-# nboot the number of bootstrap replicates.
-# conf the confidence level of the bootstrap confidence intervals.
-
-  strategies <- unique(ggInsOuts$strategy)
-
-  for(strategy in strategies)
-  {
-
-    by_strategy <- ggInsOuts[ggInsOuts$strategy==strategy,]
-
-    x <- by_strategy[,1:num_inputs]
-    y <- by_strategy['time_to_resistance0.25']
-
-    pcc_res <- pcc(x, y, rank=TRUE)
-
-    #if you add bootstrap, error bars are added to the default plot & extra columns to the PRCC object
-    #pcc_res <- pcc(x, y, rank=TRUE, nboot=100)
-    #plot(pcc_res)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
-
-    to_plot <- pcc_res$PRCC
-    #rename column 1 from 'original to PRCC
-    names(to_plot)[1] <- 'PRCC'
-    to_plot$inputs <- rownames(to_plot)
-
-    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
-             geom_point(shape=1, colour='red') +
-             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
-             geom_hline(yintercept = 0, linetype=3) +
-             ylim(-1,1) +
-             #ggtitle(paste(strategy,"PRCC")) +
-             ggtitle(paste(strategy,"time-to-resistance 0.25")) +
-             xlab(NULL)
-          )
-  }
 
 ```
-
-## response of resistance thresholds to individual inputs faceted by strategy
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
-
-  names_inputs <- names(ggInsOuts)[1:num_inputs]
-
-  #todo find out to take sample this doesn't work
-  #ggSample <- sample(ggInsOuts,500)
-
-  for(i in names_inputs)
-  {
-    #x11()
-
-    y <- 'time_to_resistance0.25'    
-    #y <- 'time_to_resistance0.5'
-    color <- 'strategy'
-    #coloring by exposure shows its effect even when other params are varying
-    #color <- 'exposure'
-
-    print( ggplot(ggInsOuts, aes_string(x=i, y=y, color=color)) + 
-             #points not wanted if 10000
-             #geom_point(shape=3, show.legend=FALSE) + 
-             geom_smooth(colour='red', linetype='dashed',size=0.5, show.legend=FALSE) +
-             facet_wrap( ~ strategy) + #, show_guide = FALSE) + 
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i))
-  }
-
-```

### 2016-05-20 inst/documents/pap1_figs_slimmed_50_unequal.Rmd
moved documents into own folder

--- 2016-03-24 paper1_results_figs_slimmed_50.Rmd
@@
 ---
-title: "paper1_results_figs_slimmed_50"
+title: "pap1_figs_slimmed_50_unequal : all runs with high starting frequency of resistance to 1 insecticide"
 output: pdf_document
@@
 
+# ALL FIGS FOR UNEQUAL SCENARIOS WHERE RESISTANCE FOR INSECTICIDE STARTS AT 0.5
+
+## initially just 500 test runs
+
 ## ALL FIGS AT 50% resistance threshold
-
-## Main Figures for paper 1 : version 6 - in progress
-### This file is automagically generated by R from paper1_results_figs_slimmed_50.Rmd
-
-## Notable results
-
-1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
-1.
-In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture, switching to sole use.
-The latter takes slightly longer to reach the threshold.
-1. Exposure and dominance have the greatest on time-to-resistance as shown in PRCC analyses.
-1.
-Effectiveness of the insecticides is the one input that can change the average relative positioning of the strategies a) sequential and b) adaptive mixture.
-Below an effectiveness of ~0.55 to 0.6 sequential takes longer to reach the resistance threshold, and above this the adaptive mixture takes longer.
-1. Exposure and effectiveness come out as the most important effect on the difference between the sequential and adpative mixture strategies.
-1.
-Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
-1.
-The implication of additional operational costs of adaptive mixtures can be investigated.
-If a requirement of a 20% improvement in resistance is imposed then more runs favour sequential than mixture.
 
@@
 
-  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
 
-  # load(file=paste0(outFolder,'listOutMix_unequal.rda'))
-  # load(file=paste0(outFolder,'listOutI1_unequal.rda'))
-  # load(file=paste0(outFolder,'listOutI2_unequal.rda'))
+  load(file=paste0(outFolder,'listOutMix_unequal.rda'))
+  load(file=paste0(outFolder,'listOutI1_unequal.rda'))
+  load(file=paste0(outFolder,'listOutI2_unequal.rda'))
 
@@
 
-ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 2"), ]
+#ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 2"), ]  
+#4unequal also remove Seq 1st & Mix 1st which are meaningless 
+#unique(ggInsOuts$strategy)
+#"insecticide 1" "insecticide 2" "Sequential"    "Mix either"    "Mix adaptive"  "Mix both" 
+ggSubset <- ggInsOuts[ ggInsOuts$strategy %in% c("Sequential", "Mix adaptive", "Mix both"), ]
 
@@
       #ylim(0,450) +
-      coord_cartesian( ylim=c(0, 350)) +
+      coord_cartesian( ylim=c(0, 450)) +
       geom_violin(draw_quantiles = c(0.25, 0.75), show.legend=FALSE, adjust = .5 ) +
       ylab('Time to resistance (gens)') +
-      scale_x_discrete("", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Seq 1st", "Mix either"="Mix 1st","Sequential"="Seq both", "Mix adaptive"="Adaptive", "Mix both"="Mix both" ) )
+      #4unequal removed 2 strategies
+      scale_x_discrete("", limits=c("Sequential","Mix adaptive","Mix both"), labels = c("Sequential"="Seq both", "Mix adaptive"="Adaptive", "Mix both"="Mix both" ) )
 
@@
 
-  strategies <- unique(ggInsOuts$strategy)
+  #strategies <- unique(ggInsOuts$strategy)
 
   #in main paper I want to just display sequential and mixture strategies
-  strategies <- strategies[-1:-2]
+  #strategies <- strategies[-1:-2]
+
+  #4unequal restrict to these strategies
+  strategies <- c("Sequential", "Mix adaptive", "Mix both")
 
@@
 
-  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))
+  #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
+  #4unequal only 3 strategies now
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.8))
 
@@
 
-  #names_inputs <- names(ggInsOuts)[1:num_inputs]
+  #remove insecticide1 & 2 strategies
+  #ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
+  #4unequal
+  ggSubset <- ggInsOuts[ ggInsOuts$strategy %in% c("Sequential", "Mix adaptive", "Mix both"), ]
 
-  #remove insecticide1 & 2 strategies
-  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
   #subset which output plots to show in the main paper
@@
   #             "dominance_allele1","selection_coef_allele1","maleExposureProp")
-  in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")
+
+  #in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")  
+  #4unequal
+  in_cols <- c("exposure","dominance_allele1","selection_coef_allele1","effectiveness_insecticide1")
 
@@
   #to rearrange order of legend entries
-  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either")) #, labels=c("MBB", "MAA", "MCC"))
+  #ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either"))
+  #4unequal
+  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential"))
 
@@
     ## repeat code for adaptive
-    x <- dif_mixA_seq[,1:num_inputs]
-    y <- dif_mixA_seq["mixA_minus_seq0.5"]
-
-    pcc_res <- pcc(x, y, rank=TRUE)
-
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
-
-    to_plot <- pcc_res$PRCC
-    #rename column 1 from 'original to PRCC
-    names(to_plot)[1] <- 'PRCC'
-    to_plot$inputs <- rownames(to_plot)
-
-    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
-             geom_point(shape=1, colour='red') +
-             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
-             geom_hline(yintercept = 0, linetype=3) +
-             #theme_bw() +
-             ylim(-1,1) +
-             xlab(NULL)
-          )
+    #4unequal comment out adaptive stuff
+    # x <- dif_mixA_seq[,1:num_inputs]
+    # y <- dif_mixA_seq["mixA_minus_seq0.5"]
+    #   
+    # pcc_res <- pcc(x, y, rank=TRUE)
+    # 
+    # #results are here I can probably rbind them together into a df that I can ggplot
+    # #pcc_res$PRCC
+    # 
+    # to_plot <- pcc_res$PRCC
+    # #rename column 1 from 'original to PRCC
+    # names(to_plot)[1] <- 'PRCC'
+    # to_plot$inputs <- rownames(to_plot)  
+    # 
+    # print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+    #          geom_point(shape=1, colour='red') +
+    #          theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+    #          geom_hline(yintercept = 0, linetype=3) +
+    #          #theme_bw() +
+    #          ylim(-1,1) +
+    #          xlab(NULL)
+    #       )
 
@@
 
-## Fig. x5 Effectiveness and exposure together determine most of whether resistance arises slower under adaptive mixture or sequential strategies
+## Fig. x5 with unequal starting resistance this plot works less well, because resistance is always slower for mixture versuis sequential (and maybe other inputs have greater influence)
+(previous text : Effectiveness and exposure together determine most of whether resistance arises slower under mixture or sequential strategies with unequal starting resistance)
 
@@
 
-  ggSubset <- dif_mixA_seq
+  #4unequal change mixA to mix2 (mixA not relevant for unequal scenarios)
+  #in the test 500 runs resistance was always slower in mixtures vs sequential
+  #although not always >20% slower
+
+  ggSubset <- dif_mix2_seq
   #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mixA_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 > 0, 1, 0 )
-  ggSubset$seq_better_mixA0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 < 0, 1, 0 )
-  ggSubset$abs_diff <- abs(ggSubset$mixA_minus_seq0.5)
+  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
+  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
+  ggSubset$abs_diff <- abs(ggSubset$mix2_minus_seq0.5)
 
-  #tricky
-  #be a bit careful what to do with zeros where results are same for both strategies
-  #I use these better vars to facet the plots
-  #may want the zeroes in both facets
-  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
-  #set these to 1 (they were 0)
-  df_same$mixA_better_seq0.5 <- 1
-  df_same$seq_better_mixA0.5 <- 1
-  #rbind them back onto other dataset
-  ggSubset <- rbind( ggSubset, df_same)
+  #4unequal commented out code below
+  # #tricky
+  # #be a bit careful what to do with zeros where results are same for both strategies
+  # #I use these better vars to facet the plots
+  # #may want the zeroes in both facets
+  # #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
+  # df_same <- ggSubset[which(ggSubset$abs_diff==0),]
+  # #set these to 1 (they were 0)
+  # df_same$mix2_better_seq0.5 <- 1
+  # df_same$seq_better_mix20.5 <- 1
+  # #rbind them back onto other dataset
+  # ggSubset <- rbind( ggSubset, df_same)
 
   #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mixA_not20pc_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
+  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
 
@@
     #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
-
-             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance slower','1'='sequential\nresistance slower'))) +
-
-             #YlGnBu looks really good except that the yellow points are very faint
-             #scale_colour_distiller(palette='YlGnBu', direction=1) +
-             #option="plasma" looks good too
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
-             theme_bw() +
-             theme(legend.position = "bottom") +
-             #theme_minimal() +
-             #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='black', shape=3) +
-             #failing to add legend for curtis point
-             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
-             #scale_size_manual(name = "Curtis Fig.2", values="test") +
-             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) + 
-             #!!!NOW can I add a lm fitted to the 0 points abs_diff
-             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
-             #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
-
-             #add text for Curtis point
-             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
-           )
-
-  # doing the lm outside of ggplot    
-  x1=df_same$exposure 
-  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
-
-  l_mod <- lm(y1 ~ x1)
-  l_mod$coefficients
-  #(Intercept)          x1 
-  #1.019330    0.560508
-
-  #approximation
-  #x2 <- seq(0,1,0.1)
-  #y2 <- 0.6*x2+0.98
-  #plot(x2,y2)
-
-```
-
-Red triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels).
-The red line is a linear model through these same points.
-
-\pagebreak
-
-## Fig. x5.1 modified for mix2 vs seq
-
-fewer runs where sequential gives slower resistance
-
-```{r, eval=TRUE, echo=FALSE, fig.height=6}
-
-  ggSubset <- dif_mix2_seq
-  #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
-  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
-  ggSubset$abs_diff <- abs(ggSubset$mix2_minus_seq0.5)
-
-  #tricky
-  #be a bit careful what to do with zeros where results are same for both strategies
-  #I use these better vars to facet the plots
-  #may want the zeroes in both facets
-  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
-  #set these to 1 (they were 0)
-  df_same$mix2_better_seq0.5 <- 1
-  df_same$seq_better_mix20.5 <- 1
-  #rbind them back onto other dataset
-  ggSubset <- rbind( ggSubset, df_same)
-
-  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
-
-    #find unequal runs
-    unequal <- ggSubset[ggSubset$resist_start_lo_div_hi < 0.01,]
-    nrow(ggSubset)
-    nrow(unequal)
-    #2721 out of the 6189 remaining runs
-
-    #in how many runs was seq better than mix
-    sum(unequal$seq_better_mix20.5)
-    #311
-
-    library(viridis)
-
     print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
@@
              geom_point(alpha=1, size=0.5) + 
-             #!!!NOW can I add a lm fitted to the 0 points abs_diff
-             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
-             #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
-
-             #now add the unequal points
-             #showed that they were on both sides
-             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=2) +
+             #4unequal copmment out below
+             #NOW can I add a lm fitted to the 0 points abs_diff
+             # geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
+             # #can I subtley highlight the zero points which are repeated in each plot ?
+             # geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
 
@@
   # doing the lm outside of ggplot    
-  x1=df_same$exposure 
-  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
-
-  l_mod <- lm(y1 ~ x1)
-  #l_mod$coefficients
-  #(Intercept)          x1 
-  #1.019330    0.560508
-
-  #approximation
-  #x2 <- seq(0,1,0.1)
-  #y2 <- 0.6*x2+0.98
-  #plot(x2,y2)
+  # x1=df_same$exposure 
+  # y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
+  # 
+  # l_mod <- lm(y1 ~ x1)
+  # l_mod$coefficients
 
@@
 
-## Fig. x6 Modifying the previous plot to show where resistance arises over 20% slower in an adaptive mixture strategy versus sequential.
+\pagebreak
 
-1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the adaptive mixture under these constraints.
+## Fig. x6 Modifying the previous plot to show where resistance arises over 20% slower in a mixture strategy versus sequential.
+
+1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the mixture under these constraints.
 
@@
 
-  df_20pc <- ggSubset[which(ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq > 0.19 & 
-                            ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq < 0.21  ),]
+#4unequal changing from mixA to mix2
+
+  df_20pc <- ggSubset[which(ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq > 0.19 & 
+                            ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq < 0.21  ),]
 
   #remove column it is faceted on to get line in both facets
-  col_indices <- colnames(df_20pc) != 'mixA_not20pc_better_seq0.5'
+  col_indices <- colnames(df_20pc) != 'mix2_not20pc_better_seq0.5'
   df_20pc <- df_20pc[,col_indices]
@@
 
-             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+             facet_grid( ~ mix2_not20pc_better_seq0.5, labeller=as_labeller( c('0'='mixture\nresistance >20% slower','1'='mixture\nresistance <20% slower'))) +
 
@@
              #main points for the plot
-             geom_point(alpha=1, size=0.5) +
-             #add a lm fitted to the 0 points abs_diff
-             #geom_smooth(data=df_same, method = "lm", se = FALSE)
-             #todo change to where exactly 20% diff
-             geom_smooth(data=df_20pc, method = "lm", se = FALSE,colour='red', lwd=0.5)+
+             geom_point(alpha=1, size=0.5)
+             #add a lm fitted to where exactly 20% diff
+             #geom_smooth(data=df_20pc, method = "lm", se = FALSE,colour='red', lwd=0.5)+
              #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=df_20pc, size=0.8, colour='red', shape=2) #, fill='white') 
+             #geom_point(data=df_20pc, size=0.8, colour='red', shape=2) #, fill='white') 
            )
@@
   # doing the lm outside of ggplot    
-  x1=df_20pc$exposure 
-  y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
-
-  l_mod <- lm(y1 ~ x1)
-  l_mod$coefficients
+  # x1=df_20pc$exposure 
+  # y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
+  # 
+  # l_mod <- lm(y1 ~ x1)
+  # l_mod$coefficients
 # (Intercept)          x1 
@@
 ## Fig.
-x7 Classification Trees for whether resistance arises slower with an adaptive mixture or sequential strategy.
+x7 Classification Trees for whether resistance arises slower with a mixture or sequential strategy.
 The upper plot shows when mixture is slower than sequential, the lower plot is for when the mixture is more than 20% slower.
@@
   #code copied from sensiAnPaper1All
+  #4unequal changed mix_A to mix_2
 
@@
   #T/F
-  resistBetterMix_ASeqBoolean <- resistPointsMix_A > resistPointsSeq
+  resistBetterMix_2SeqBoolean <- resistPointsMix_2 > resistPointsSeq
   #convert to 0/1
-  resistBetterMix_ASeq <- plyr::aaply(resistBetterMix_ASeqBoolean,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_2Seq <- plyr::aaply(resistBetterMix_2SeqBoolean,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix_ASeq) <- gsub('gen','betterMix_ASeq', rownames(resistBetterMix_ASeq))
+  rownames(resistBetterMix_2Seq) <- gsub('gen','betterMix_2Seq', rownames(resistBetterMix_2Seq))
 
   #20% better
-  resistBetterMix_ASeqBoolean20 <- resistPointsMix_A > (1.2 * resistPointsSeq)
+  resistBetterMix_2SeqBoolean20 <- resistPointsMix_2 > (1.2 * resistPointsSeq)
   #convert to 0/1
-  resistBetterMix_ASeq20 <- plyr::aaply(resistBetterMix_ASeqBoolean20,.margins=c(1,2),.fun=as.numeric)
+  resistBetterMix_2Seq20 <- plyr::aaply(resistBetterMix_2SeqBoolean20,.margins=c(1,2),.fun=as.numeric)
   #rename rows ready for binding on to inputs
-  rownames(resistBetterMix_ASeq20) <- gsub('gen','betterMix_ASeq20', rownames(resistBetterMix_ASeq20))
+  rownames(resistBetterMix_2Seq20) <- gsub('gen','betterMix_2Seq20', rownames(resistBetterMix_2Seq20))
 
@@
 
-  resistBetterMix_ASeq <- repMix(resistBetterMix_ASeq)
-  resistBetterMix_ASeq20 <- repMix(resistBetterMix_ASeq20)
+  resistBetterMix_2Seq <- repMix(resistBetterMix_2Seq)
+  resistBetterMix_2Seq20 <- repMix(resistBetterMix_2Seq20)
 
-  treeInput <- rbind( treeInput, resistBetterMix_ASeq, resistBetterMix_ASeq20 )
+  treeInput <- rbind( treeInput, resistBetterMix_2Seq, resistBetterMix_2Seq20 )
 
@@
   #just doing for a few strategies, sensiAnPpaer1All.Rmd does for a bunch
-  treeResponses <- c( rownames(resistBetterMix_ASeq)[2], #2 is 25%
-                      rownames(resistBetterMix_ASeq20)[2]
+  treeResponses <- c( rownames(resistBetterMix_2Seq)[2], #2 is 25%
+                      rownames(resistBetterMix_2Seq20)[2]
                     )

### 2016-05-24 inst/documents/sensiAnPaper1All.Rmd
test scenarios

--- 2016-03-24 sensiAnPaper1All.Rmd
@@
 
-  #nScenarios <- 10  
+  nScenarios <- 10  
   #nScenarios <- 100
-  nScenarios <- 500  
+  #nScenarios <- 500  
   #nScenarios <- 10000
@@
   ## do model runs
-  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
-  # listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment )
-  # listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment )
+  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
+  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment )
+  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment )
 
   ## 23/3/16 expt with an 'old' insecticide
-  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment, old_insecticide=TRUE )
-  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment, old_insecticide=TRUE )
-  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment, old_insecticide=TRUE )
+  # listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment, old_insecticide=TRUE )
+  # listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment, old_insecticide=TRUE )
+  # listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment, old_insecticide=TRUE )
 

### 2016-05-25 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
moved documents into own folder
edit text and start to look at single insecticide effects

--- 2016-05-20 paper2_resistance_mechanisms_mixtures.Rmd
@@
 
-### version 1
+### version 2
 
@@
 (in press) describe the technical details of the model and explore large parameter ranges.
-In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate the mechanisms by which different parameter values are expected to influence the development of resistance.
+In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate mechanisms that can influence the development of resistance.
 
@@
 *Before this add figures for effect of inputs on sole use.*
+* contrasting pairs, or even 3 values, 0.1,0.5 & 0.9*
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+
+  #create empty object to hold input matrix for return
+  #input <- NULL
+
+  #??? but how do I set so just one insecticide ?
+  #??? have I been doing this right in previous runs for this paper ?
+  #??? maybe yes, because setExposure() is called from runcurtis_f2()
+  #this sets exposures according to whether insecticide1,2 or mixture
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.2, 
+                      h.RS2_0B = 0.2,
+                      a = a,
+                      phi.SS1_A0 = 0.2,
+                      phi.SS2_0B = 0.2,
+                      s.RR1_A0 = 0.2,
+                      s.RR2_0B = 0.2 )
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  #input <- cbind(input, inputOneScenario)
+  input <- cbind(i1,i2)
+
+  listOut <- runModel2( input )
+
+  plotallele.freq.andy(listOut, locus="1", sex="mf_mean", add_legend=FALSE)
+
+  cat("Fig. a1. starting to look at single insecticide effects")
+```
 

### 2016-06-02 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
starting single insecticide figs

@@
   library(resistance)
+  library(ggplot2)
 ```
@@
 
-*Before this add figures for effect of inputs on sole use.*
+* add figures for effect of inputs on sole use.*
 * contrasting pairs, or even 3 values, 0.1,0.5 & 0.9*
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-
-  #create empty object to hold input matrix for return
-  #input <- NULL
 
@@
                       h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.2,
+                      h.RS2_0B = 0.5,
                       a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.2,
-                      s.RR1_A0 = 0.2,
-                      s.RR2_0B = 0.2 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
 
@@
 
+  i3 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.8, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
   #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2)
+  input <- cbind(i1,i2,i3)
 
@@
 
-  plotallele.freq.andy(listOut, locus="1", sex="mf_mean", add_legend=FALSE)
+  #plotallele.freq.andy(listOut, locus="1", sex="mf_mean", add_legend=FALSE)
 
-  cat("Fig. a1. starting to look at single insecticide effects")
+  # now doing a better way
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=h.RS1_A0)) + 
+         #geom_line(aes(colour=h.RS1_A0)) )
+         geom_point() #show.legend=FALSE) +
+
+       ) # end print
+
+  cat("Fig. a1. Dominance effect on single insecticide.")
+
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.2,
+                      phi.SS2_0B = 0.5,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.8,
+                      phi.SS2_0B = 0.5,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  #input <- cbind(input, inputOneScenario)
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=h.RS1_A0)) + 
+         #geom_line(aes(colour=h.RS1_A0)) )
+         geom_point() #show.legend=FALSE) +
+
+       ) # end print
+
+  cat("Fig. a2. Effectiveness effect on single insecticide.")
+
 ```

### 2016-06-02 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
add & polish single insecticide scenarios

--- 2016-05-20 paper2_resistance_mechanisms_mixtures.Rmd
@@
   library(resistance)
+  library(ggplot2)
 ```
 
-### version 1
+### version 2
 
@@
 
-*Before this add figures for effect of inputs on sole use.*
+For single insecticide use, increasing the values of either effectiveness, exposure, dominance, or selective advantage, all lead to faster resistance (Figs a1-4).
+(In these four plots the value of the inputs not being varied are set at a constant 0.7 for effectiveness and 0.5 for the other 3).
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.9,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  #input <- cbind(input, inputOneScenario)
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=phi.SS1_A0)) + 
+         labs(colour = "effectiveness") +
+         coord_trans(y = "log10") +    
+         geom_point() #show.legend=FALSE) +
+
+       ) # end print
+
+  cat("Fig. a1. Effectiveness effect on single insecticide.")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+
+  exposure <- 0.3
+  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      exposure = exposure, #used for post-run analysis
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  exposure <- 0.5
+  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      exposure = exposure, #used for post-run analysis
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  exposure <- 0.7
+  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      exposure = exposure, #used for post-run analysis
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  #input <- cbind(input, inputOneScenario)
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=exposure)) + 
+         labs(colour = "exposure") +
+         coord_trans(y = "log10") +    
+         geom_point() #show.legend=FALSE) +
+
+       ) # end print
+
+  cat("Fig. a2. Exposure effect on single insecticide.")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+
+  #this sets exposures according to whether insecticide1, 2 or mixture
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.3, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.7, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  #input <- cbind(input, inputOneScenario)
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  #plotallele.freq.andy(listOut, locus="1", sex="mf_mean", add_legend=FALSE)
+
+  # now doing a better way
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=h.RS1_A0)) + 
+         labs(colour = "dominance") +
+         coord_trans(y = "log10") + 
+         geom_point() #show.legend=FALSE) +
+
+       ) # end print
+
+  cat("Fig. a3. Dominance effect on single insecticide.")
+
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.3,
+                      s.RR2_0B = 0.5 )
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.5,
+                      s.RR2_0B = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.7,
+                      phi.SS2_0B = 0.7,
+                      s.RR1_A0 = 0.7,
+                      s.RR2_0B = 0.5 )
+
+  #input <- cbind(input, inputOneScenario)
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=s.RR1_A0)) + 
+         labs(colour = "selective advantage") +
+         coord_trans(y = "log10") +   
+         geom_point() #show.legend=FALSE) +
+
+       ) # end print
+
+  cat("Fig. a4. Selective advantage effect on single insecticide.")
+
+```
+
+When two insecticides are used the relative performance of a mixture strategy compared to sequential use is determined by the values of these four inputs.
 

### 2016-06-13 R/sensiAnPaperPart.r
started changes to how selection coefficients dealt with from sensitivity analysis

@@ -109,6 +109,15 @@ sensiAnPaperPart <- function( nScenarios = 10,
     s.RR1_A0 <- runif(1, min=0.1, max=0.45)
     s.RR2_0B <- runif(1, min=0.1, max=0.45)     
     
+    # 14/6/16 to add resistance_coeff_I1 & I2
+    # as an alternative to varying s directly from sensi analysis
+    # to allow a wider range (higher values) while avoiding correlation between 
+    # effectiveness and advantage inputs
+    # resistance_coeff_I1 <- runif(1, min=0.2, max=1)
+    # resistance_coeff_I2 <- runif(1, min=0.2, max=1) 
+    # s.RR1_A0 <- resistance_coeff_I1 * phi.SS1_A0
+    # s.RR2_0B <- resistance_coeff_I2 * phi.SS2_0B    
+    
     ## put the generated values into an input matrix, using defaults for non specified parameters
     inputOneScenario <- setInputOneScenario( max_gen = max_gen,
       
@@ -127,7 +136,10 @@ sensiAnPaperPart <- function( nScenarios = 10,
                         s.RR1_A0 = s.RR1_A0,
                         s.RR2_0B = s.RR2_0B,
                         maleExposureProp = maleExposureProp,
-                        correctMixDeployProp = correctMixDeployProp  
+                        correctMixDeployProp = correctMixDeployProp
+                        # 14/6/16
+                        #,resistance_coeff_I1 = resistance_coeff_I1,
+                        #resistance_coeff_I2 = resistance_coeff_I2
                         )   
     
     

### 2016-06-13 R/setInputOneScenario.r

@@ -63,6 +63,8 @@
 #' @param maleExposureProp male exposure as a propoertion of female, default 1 for same, likely <1  
 #' @param correctMixDeployProp proportion of correct deployment of mixtures, 
 #'    if <1 other portion divided between single insecticides
+#' @param resistance_coeff_I1 effect of resistance in overcoming insecticide 1 effectiveness
+#' @param resistance_coeff_I2 effect of resistance in overcoming insecticide 2 effectiveness    
 #' 
 #' @return named vector
 #' @export
@@ -124,7 +126,10 @@ setInputOneScenario <- function( calibration = 100,
                                  sexLinked = 0,
                                  maleExposureProp = 1,
                                  correctMixDeployProp = 1,
-                                 exposure = 0.9
+                                 exposure = 0.9,
+                                 #14/6/16
+                                 resistance_coeff_I1 = 0.5,
+                                 resistance_coeff_I2 = 0.5
                                  )
 {
   
@@ -216,6 +221,10 @@ setInputOneScenario <- function( calibration = 100,
   input[	55	] <-	correctMixDeployProp
   #1/2/16 allowing saving of single exposure param just for use in post-run analyses
   input[	56	] <-	exposure  
+  
+  #14/6/16
+  input[	57	] <-	resistance_coeff_I1  
+  input[	58	] <-	resistance_coeff_I2  
     
   a.m <- sum(a.m_00, a.m_a0, a.m_A0, a.m_0b, a.m_0B, a.m_ab, a.m_AB, a.m_Ab, a.m_aB)
   #if ( a.m != 1 ){

### 2016-06-15 R/sensiAnPaperPart.r
rename to rr_advantage

@@ -109,14 +109,14 @@ sensiAnPaperPart <- function( nScenarios = 10,
     s.RR1_A0 <- runif(1, min=0.1, max=0.45)
     s.RR2_0B <- runif(1, min=0.1, max=0.45)     
     
-    # 14/6/16 to add resistance_coeff_I1 & I2
+    # 14/6/16 to add rr_advantage_I1 & I2
     # as an alternative to varying s directly from sensi analysis
     # to allow a wider range (higher values) while avoiding correlation between 
     # effectiveness and advantage inputs
-    # resistance_coeff_I1 <- runif(1, min=0.2, max=1)
-    # resistance_coeff_I2 <- runif(1, min=0.2, max=1) 
-    # s.RR1_A0 <- resistance_coeff_I1 * phi.SS1_A0
-    # s.RR2_0B <- resistance_coeff_I2 * phi.SS2_0B    
+    # rr_advantage_I1 <- runif(1, min=0.2, max=1)
+    # rr_advantage_I2 <- runif(1, min=0.2, max=1) 
+    # s.RR1_A0 <- rr_advantage_I1 * phi.SS1_A0
+    # s.RR2_0B <- rr_advantage_I2 * phi.SS2_0B    
     
     ## put the generated values into an input matrix, using defaults for non specified parameters
     inputOneScenario <- setInputOneScenario( max_gen = max_gen,
@@ -138,8 +138,8 @@ sensiAnPaperPart <- function( nScenarios = 10,
                         maleExposureProp = maleExposureProp,
                         correctMixDeployProp = correctMixDeployProp
                         # 14/6/16
-                        #,resistance_coeff_I1 = resistance_coeff_I1,
-                        #resistance_coeff_I2 = resistance_coeff_I2
+                        #,rr_advantage_I1 = rr_advantage_I1,
+                        #rr_advantage_I2 = rr_advantage_I2
                         )   
     
     

### 2016-06-15 R/setInputOneScenario.r

@@ -63,8 +63,8 @@
 #' @param maleExposureProp male exposure as a propoertion of female, default 1 for same, likely <1  
 #' @param correctMixDeployProp proportion of correct deployment of mixtures, 
 #'    if <1 other portion divided between single insecticides
-#' @param resistance_coeff_I1 effect of resistance in overcoming insecticide 1 effectiveness
-#' @param resistance_coeff_I2 effect of resistance in overcoming insecticide 2 effectiveness    
+#' @param rr_advantage_I1 effect of resistance in overcoming insecticide 1 effectiveness
+#' @param rr_advantage_I2 effect of resistance in overcoming insecticide 2 effectiveness    
 #' 
 #' @return named vector
 #' @export
@@ -128,8 +128,8 @@ setInputOneScenario <- function( calibration = 100,
                                  correctMixDeployProp = 1,
                                  exposure = 0.9,
                                  #14/6/16
-                                 resistance_coeff_I1 = 0.5,
-                                 resistance_coeff_I2 = 0.5
+                                 rr_advantage_I1 = 0.5,
+                                 rr_advantage_I2 = 0.5
                                  )
 {
   
@@ -223,8 +223,8 @@ setInputOneScenario <- function( calibration = 100,
   input[	56	] <-	exposure  
   
   #14/6/16
-  input[	57	] <-	resistance_coeff_I1  
-  input[	58	] <-	resistance_coeff_I2  
+  input[	57	] <-	rr_advantage_I1  
+  input[	58	] <-	rr_advantage_I2  
     
   a.m <- sum(a.m_00, a.m_a0, a.m_A0, a.m_0b, a.m_0B, a.m_ab, a.m_AB, a.m_Ab, a.m_aB)
   #if ( a.m != 1 ){

### 2016-06-15 inst/documents/paper1_results_figs_rr.Rmd
copied from paper1_results_figs pre changes

--- 2016-04-28 paper1_results_figs.Rmd
@@
 ---
-  title: "paper1_results_figs"
+  title: "paper1_results_figs_rr"
   output: pdf_document
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+```{r, eval=FALSE, echo=FALSE, fig.height=3.5}
+
+  #10/5/2016 seems to crash R, so set to eval=FALSE
 

### 2016-06-15 inst/documents/paper1_results_figs_rr.Rmd
rr_adv changes and putting runs on start - seem to be working

@@
 
-# Figures for paper 1 
-# version 5 - in progress
+# Figures for paper 1 after change to using rr_advantage instead of selection coeff from sensi analysis
+# version 1 - in progress
 
-## Based on 10000 runs with all inputs varying at once which explains the variability at particular input values.
+```{r}
+  #library(devtools)
+  #install_github('AndySouth/resistance')
+  library(resistance)
 
-Comparing these insecticide use strategies:
+  outFolder <- "C:\\Dropbox\\resistanceResults\\"
+```
 
-1. Single use
-1. Mixture (either AI) : threshold reached for either insecticide
-1. Adaptive mixture : threshold reached for either insecticide, switch to sole use of other until it too reaches threshold
-1. Mixture (both AIs) : threshold reached for both insecticides in mixture
-1. Sequential : sole use of one insecticide, switch to other when threshold reached until it too reaches threshold
+<!-- eval=FALSE to stop slow code being executed --> 
+```{r, eval=FALSE}
 
-Renamed from previous naming system :
-Mixture (either AI)   [mixture1]
-Mixture (both AIs)    [mixture3]
-Adaptive mixture      [mixture2]
+  #default setting of the random seed in sensiAnPaperPart() should make the input files mostly the same
+  #the exception is the exposure array a. The single exposure input will be the same, but the array changes
+  #according to setExposure(  )
 
-## Notable results
+  #nScenarios <- 10  
+  #nScenarios <- 100
+  #nScenarios <- 500 
+  nScenarios <- 1000
+  #nScenarios <- 10000
 
-1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
-1.
-In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture switching to sole use.
-The latter takes slightly longer to reach the threshold.
-1. Exposure and effectiveness seem to be the most important inputs.
-1. Exposure has the greatest effect on time-to-resistance.
-1.
-Effectiveness of the insecticides is the one input that has a notably different effect on the strategies of a) sequential and b) adaptive mixture switching to sole use.
-Below an effectiveness of ~0.55 to 0.6 sequential takes longer, and above this 'mixture switching to sole use' takes longer.
+  #to run 'extended' or not
+  experiment <- 'extended' #'curtis'
+
+  ## do model runs
+  listOutMix <- sensiAnPaperPart( nScenarios, insecticideUsed = 'mixture', experiment=experiment )
+  listOutI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment=experiment )
+  listOutI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment=experiment )
+
+  ## save results objects as rda for re-use
+  save(listOutMix, file=paste0(outFolder,'listOutMix_rr_1000.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1_rr_1000.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2_rr_1000.rda'))  
+```
 
@@
 
-  ## to load previously saved runs
-#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI2_1000.rda'))
-
-  ## trying with the extended experiment _ex100
+  ## trying with the extended experiment 
   experiment <- 'extended'
 
-  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
-
-  # load(file=paste0(outFolder,'listOutMix_ex2_1000.rda'))
-  # load(file=paste0(outFolder,'listOutI1_ex2_1000.rda'))
-  # load(file=paste0(outFolder,'listOutI2_ex2_1000.rda'))
+  ## to load previously saved runs  
+  load(file=paste0(outFolder,'listOutMix_rr_1000.rda'))
+  load(file=paste0(outFolder,'listOutI1_rr_1000.rda'))
+  load(file=paste0(outFolder,'listOutI2_rr_1000.rda'))
 
@@
   #hardcode which variables to include in analysis to keep it simple and transparent
-  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #14/6/16 change to rr_advantage
+  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','rr_advantage_I1','rr_advantage_I2')
 
@@
 
-  rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
-  rownames(treeInput)[rownames(treeInput)=="s.RR2_0B"] <- "selection_coef_allele2"
+  #14/6/16 rr_advantage these not to be used anymore
+  #rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
+  #rownames(treeInput)[rownames(treeInput)=="s.RR2_0B"] <- "selection_coef_allele2"
 
@@
                              "dominance_allele2"=0.0016,
-                             "selection_coef_allele1"=0.23,
-                             "selection_coef_allele2"=0.43,
+                             # 14/6/16 rr_advantage = s / effectiveness
+                             #"selection_coef_allele1"=0.23,
+                             #"selection_coef_allele2"=0.43,
+                             "rr_advantage_I1"=0.23/0.73,
+                             "rr_advantage_I2"=0.43/1,                             
                              "maleExposureProp"=1,
@@
 
-# Some examples showing how changing effectiveness can alter the ranking of insecticide strategies
-### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
 # PRCC difference in time-to-resistance between sequential and mixture adaptive

### 2016-06-15 inst/documents/paper1_results_figs_rr.Rmd
to 0.5

@@
   #columns to remove from first
-  indices1 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.1","gen_cP0.5","strategy"))
+  indices1 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.1","gen_cP0.25","strategy"))
   #columns to add to 2nd  
-  indices2 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.25")) 
+  indices2 <- which(names(resistPointsMix_A_T) %in% c("gen_cP0.5")) 
   #rename column in 2nd
   tmp <- resistPointsSeq_T[indices2]
-  names(tmp) <- "gen_cP0.25seq"
+  names(tmp) <- "gen_cP0.5seq"
 
@@
   #BEWARE would need to increase 1000 if I increase max generations in the runs
-  didntReachThresh <- which( dif_mixA_seq$gen_cP0.25 > 500 | dif_mixA_seq$gen_cP0.25seq > 500 )
+  didntReachThresh <- which( dif_mixA_seq$gen_cP0.5 > 500 | dif_mixA_seq$gen_cP0.5seq > 500 )
   #subset by runs not making threshold
@@
 
-  dif_mixA_seq["mixA_minus_seq0.25"] <- dif_mixA_seq["gen_cP0.25"] - dif_mixA_seq["gen_cP0.25seq"]
+  dif_mixA_seq["mixA_minus_seq0.5"] <- dif_mixA_seq["gen_cP0.5"] - dif_mixA_seq["gen_cP0.5seq"]
 
   #for dif_mix2_seq
-  didntReachThresh <- which( dif_mix2_seq$gen_cP0.25 > 500 | dif_mix2_seq$gen_cP0.25seq > 500 )
+  didntReachThresh <- which( dif_mix2_seq$gen_cP0.5 > 500 | dif_mix2_seq$gen_cP0.5seq > 500 )
   #subset by runs not making threshold
@@
 
-  dif_mix2_seq["mix2_minus_seq0.25"] <- dif_mix2_seq["gen_cP0.25"] - dif_mix2_seq["gen_cP0.25seq"]
+  dif_mix2_seq["mix2_minus_seq0.5"] <- dif_mix2_seq["gen_cP0.5"] - dif_mix2_seq["gen_cP0.5seq"]
 
@@
 
-    y <- "mixA_minus_seq0.25"
+    y <- "mixA_minus_seq0.5"
 
@@
   #I probably still need to tweak data first
-  #currently I have differences in times to resistance : mixA_minus_seq0.25
+  #currently I have differences in times to resistance : mixA_minus_seq0.5
   #what I think I want is the proportion of runs in which mix is better
@@
   #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mixA_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 > 0, 1, 0 )
-  ggSubset$seq_better_mixA0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 < 0, 1, 0 )
-  ggSubset$abs_mixA_minus_seq0.25 <- abs(ggSubset$mixA_minus_seq0.25)
+  ggSubset$mixA_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 > 0, 1, 0 )
+  ggSubset$seq_better_mixA0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 < 0, 1, 0 )
+  ggSubset$abs_mixA_minus_seq0.5 <- abs(ggSubset$mixA_minus_seq0.5)
 
@@
   #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
+  df_same <- ggSubset[which(ggSubset$mixA_minus_seq0.5==0),]
   #set these to 1 (they were 0)
-  df_same$mixA_better_seq0.25 <- 1
-  df_same$seq_better_mixA0.25 <- 1
+  df_same$mixA_better_seq0.5 <- 1
+  df_same$seq_better_mixA0.5 <- 1
   #rbind them back onto other dataset, probably not needed ??
@@
   #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mixA_not20pc_better_seq0.25 <- ifelse( ggSubset$mixA_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
+  ggSubset$mixA_not20pc_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
 
     #failed hex plot just gave me counts
-    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.5'))+
     #          geom_hex()
@@
     #interesting plot showing where sequence is better is restricted to higher values of exposure and effectiveness 
-    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.5'))+
              #geom_hex()
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_better_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_better_seq0.5))+
              #geom_hex()
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mixA_better_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mixA_better_seq0.5))+
 
@@
     #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mixA0.25))+
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mixA0.5))+
+             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
 
@@
     #but the previous one may be clearer
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_minus_seq0.5))+
              #geom_hex()
@@
     #failed attempt at getting Curtis point in legend
-    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.5, size="Curtis"))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.5))+
 
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
+             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
 
@@
 
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.5))+
 
-             facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
+             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
 
@@
              geom_point(alpha=1, size=0.5) + 
-             #!!!NOW can I add a lm fitted to the 0 points mixA_minus_seq0.25
-             geom_smooth(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], method = "lm", se = FALSE) +
+             #!!!NOW can I add a lm fitted to the 0 points mixA_minus_seq0.5
+             geom_smooth(data=ggSubset[which(ggSubset$mixA_minus_seq0.5==0),], method = "lm", se = FALSE) +
              #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$mixA_minus_seq0.25==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
+             geom_point(data=ggSubset[which(ggSubset$mixA_minus_seq0.5==0),], size=0.8, colour='lightgrey', shape=2, fill='white') +            
              #add text for Curtis point
@@
   # doing the lm outside of ggplot    
-  df_same = ggSubset[which(ggSubset$mixA_minus_seq0.25==0),]
+  df_same = ggSubset[which(ggSubset$mixA_minus_seq0.5==0),]
   x1=df_same$exposure 
@@
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.5))+
 
-             facet_grid( ~ mixA_not20pc_better_seq0.25, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
 
@@
   #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mix2_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 > 0, 1, 0 )
-  ggSubset$seq_better_mix20.25 <- ifelse( ggSubset$mix2_minus_seq0.25 < 0, 1, 0 )
-  ggSubset$abs_mix2_minus_seq0.25 <- abs(ggSubset$mix2_minus_seq0.25)
+  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
+  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
+  ggSubset$abs_mix2_minus_seq0.5 <- abs(ggSubset$mix2_minus_seq0.5)
 
@@
   #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$mix2_minus_seq0.25==0),]
+  df_same <- ggSubset[which(ggSubset$mix2_minus_seq0.5==0),]
   #set these to 1 (they were 0)
-  df_same$mix2_better_seq0.25 <- 1
-  df_same$seq_better_mix20.25 <- 1
+  df_same$mix2_better_seq0.5 <- 1
+  df_same$seq_better_mix20.5 <- 1
   #rbind them back onto other dataset, probably not needed ??
@@
   #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mix2_not20pc_better_seq0.25 <- ifelse( ggSubset$mix2_minus_seq0.25 / ggSubset$gen_cP0.25seq <= 0.2, 1, 0 )
+  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
 
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.5))+
 
@@
     #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.25))+
-             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.5))+
+             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
 
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.5))+
 
-             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
+             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
 
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.5))+
 
-             facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
+             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
              scale_color_viridis(begin=1, end=0, option="viridis", 
@@
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.5))+
 
-             facet_grid( ~ mix2_not20pc_better_seq0.25, labeller=as_labeller( c('0'='mix\nresistance >20% slower','1'='mix\nresistance <20% slower'))) +
+             facet_grid( ~ mix2_not20pc_better_seq0.5, labeller=as_labeller( c('0'='mix\nresistance >20% slower','1'='mix\nresistance <20% slower'))) +
 
@@
              #ggtitle(paste(strategy,"PRCC")) +
-             ggtitle(paste(strategy,"time-to-resistance 0.25")) +
+             ggtitle(paste(strategy,"time-to-resistance 0.5")) +
              xlab(NULL)
@@
 
-    y <- 'time_to_resistance0.25'    
+    y <- 'time_to_resistance0.5'    
     #y <- 'time_to_resistance0.5'

### 2016-06-15 inst/documents/paper1_results_figs_rr.Rmd
tidy

@@
 
-```{r}
+```{r,echo=FALSE,results='hide'}
   #library(devtools)
@@
 <!-- eval=FALSE to stop slow code being executed --> 
-```{r, eval=FALSE}
+```{r, eval=FALSE,echo=FALSE,results='hide'}
 

### 2016-06-15 inst/documents/sensiAnPaper1All.Rmd
new runs

@@
 The main question is under which scenarios are mixtures better than sequential ?
-
-This breaks down into these subsidiary quetsions :
-1. Under which scenarios is the time to resistance for either insecticide quicker for
 
@@
   ## save results objects as rda for re-use
-  save(listOutMix, file=paste0(outFolder,'listOutMix_unequal.rda'))
-  save(listOutI1, file=paste0(outFolder,'listOutI1_unequal.rda'))
-  save(listOutI2, file=paste0(outFolder,'listOutI2_unequal.rda'))
 
-  # save(listOutMix, file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-  # save(listOutI1, file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-  # save(listOutI2, file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
+  save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1_rr_10.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2_rr_10.rda'))
 

### 2016-06-16 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
bump version
change to rr_advantage and inputs to 0.2,0.5,0.8

@@
 
-### version 2
+### version 3
 
@@
 
-For single insecticide use, increasing the values of either effectiveness, exposure, dominance, or selective advantage, all lead to faster resistance (Figs a1-4).
-(In these four plots the value of the inputs not being varied are set at a constant 0.7 for effectiveness and 0.5 for the other 3).
+For single insecticide use, increasing the values of either effectiveness, exposure, dominance, or the selective advantage of RR, all lead to faster resistance (Figs a1-4).
+(In these four plots the value of the inputs not being varied are set at a constant 0.5).
 
@@
                       a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.2,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
@@
                       a = a,
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
@@
                       a = a,
-                      phi.SS1_A0 = 0.9,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.8,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
@@
 
-  exposure <- 0.3
+  exposure <- 0.2
   a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
@@
                       exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
@@
                       exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
-  exposure <- 0.7
+  exposure <- 0.8
   a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
@@
   i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.3, 
+                      h.RS1_A0 = 0.2, 
                       h.RS2_0B = 0.5,
                       a = a,
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
@@
                       a = a,
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
   i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.7, 
+                      h.RS1_A0 = 0.8, 
                       h.RS2_0B = 0.5,
                       a = a,
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.5,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.5,
+                      rr_advantage_I2 = 0.5 )
 
@@
                       a = a,
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.3,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.2,
+                      rr_advantage_I2 = 0.5 )
 
@@
                       a = a,
-                      phi.SS1_A0 = 0.7,
-                      phi.SS2_0B = 0.7,
-                      s.RR1_A0 = 0.7,
-                      s.RR2_0B = 0.5 )
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_advantage_I1 = 0.8,
+                      rr_advantage_I2 = 0.5 )
 
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=s.RR1_A0)) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=rr_advantage_I1)) + 
          labs(colour = "selective advantage") +
@@
 
-  cat("Fig. a4. Selective advantage effect on single insecticide.")
+  cat("Fig. a4. Selective advantage of RR effect on single insecticide.")
 
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 1. All control inputs equal at 0.5")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 2. Effectiveness of insecticide1 increased from 0.5 to 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 3. Effectiveness of both insecticides 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 4. Dominance of insecticide1 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 5. Dominance of both insecticides 0.9")
@@
 **Selective advantage of resistance**
+#TODO replace Figs 6,7,8 with AA,BB,CC below and modify this text
 Decreasing the selective advantage of resistance to insecticide1 from 0.5 to 0.1 increases it's time-to-resistance in the mixture from 40 generations to 150, and has a similar effect on it's use in isolation (Fig.
@@
 
-[Increasing the selective advantage of resistance above 0.5 causes an error about fitness values above 1. How to explain this ?]
-
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.1 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.1 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 6. Selective advantage of insecticide1 0.1")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.1 , s.RR2_0B = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.1 , rr_advantage_I2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 7. Selective advantage of both insecticides 0.1")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.1 , s.RR2_0B = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.1 , rr_advantage_I2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 8. Selective advantage of both insecticides 0.1, effectiveness to 0.7")
+```
+
+#cool contrast between these scenarios modifying effectiveness and rr_advantage in which all defaults set to 0.5 except
+TODO incorporate this into text above
+
+#AA effectiveness I1 to 0.8 causes mixture to be slower, although I1 is faster it protects I2 in mixture, making mixture slower
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. AA. effectiveness to 0.8")
+```
+
+#BB rr advantage I1 to 0.8 sequential remains slower, although I1 is faster as in previous it provides less protection to I2 in the mixture
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.8 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. BB rr_advantage I1 0.8")
+```
+
+#CC rr advantage I2 to 0.2 sequential remains slower (just), I2 slower than I1 in mixture similar to AA but receives less protection in mixture
+```{r, eval=TRUE, echo=FALSE, fig.height=4}
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.2 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. CC. rr_advantage I2 0.2")
 ```
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 9. Exposure 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 10. Effectiveness 0.7, sequence favoured")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , s.RR1_A0 = 0.5 , s.RR2_0B = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 11. Effectiveness 0.7, Exposure 0.9, mixture favoured")
-```
-
-```{r, eval=FALSE, echo=FALSE, fig.height=4}
-### 3 examples showing how changing effectiveness can alter the ranking of insecticide strategies
-#old modification of curtis values
-#### Curtis' Fig 2. inputs : mix2 > adaptive > mix1 > seq
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-```{r, eval=FALSE, echo=FALSE, fig.height=4}
-#old modification of curtis values
-#### decrease effectiveness of I2 from 1 to 0.85 : mix2 > adaptive > seq > mix1
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
-```
-
-```{r, eval=FALSE, echo=FALSE, fig.height=4}
-#old modification of curtis values
-#### also decrease effectiveness of I1 from 0.73 to 0.63 : seq > mix2 > adaptive > mix1
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.63 , phi.SS2_0B = 0.85 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43 )
 ```

### 2016-06-17 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
copied from paper1_results_figs_slimmed_50

--- 2016-03-30 paper1_results_figs_slimmed_50.Rmd
@@
 ```
+
+## Fig.
+x12 How sequential can be better than mixture.
+When the time taken for resistance to the first insecticide to be reached is greater than the difference in time taken for the second resistance to be reached in the presence and absence of the first insecticide.
+
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5}
+
+#sequential best. effectivenesses bothg reduced by 0.2 from curtis fig2
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.53 , phi.SS2_0B = 0.8 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+
+```

### 2016-06-17 inst/documents/paper1_results_figs_rr.Rmd
rr_advantage

@@
   #nScenarios <- 500 
-  nScenarios <- 1000
-  #nScenarios <- 10000
+  #nScenarios <- 1000
+  nScenarios <- 10000
 
@@
   ## save results objects as rda for re-use
-  save(listOutMix, file=paste0(outFolder,'listOutMix_rr_1000.rda'))
-  save(listOutI1, file=paste0(outFolder,'listOutI1_rr_1000.rda'))
-  save(listOutI2, file=paste0(outFolder,'listOutI2_rr_1000.rda'))  
+  save(listOutMix, file=paste0(outFolder,'listOutMix_rr_10000.rda'))
+  save(listOutI1, file=paste0(outFolder,'listOutI1_rr_10000.rda'))
+  save(listOutI2, file=paste0(outFolder,'listOutI2_rr_10000.rda'))  
 ```
@@
   ## to load previously saved runs  
-  load(file=paste0(outFolder,'listOutMix_rr_1000.rda'))
-  load(file=paste0(outFolder,'listOutI1_rr_1000.rda'))
-  load(file=paste0(outFolder,'listOutI2_rr_1000.rda'))
+  load(file=paste0(outFolder,'listOutMix_rr_10000.rda'))
+  load(file=paste0(outFolder,'listOutI1_rr_10000.rda'))
+  load(file=paste0(outFolder,'listOutI2_rr_10000.rda'))
 

### 2016-06-17 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
rr_advantage

@@
 ---
-title: "paper1_results_figs_slimmed_50"
+title: "paper1_results_figs_slimmed_50_rr"
 output: pdf_document
@@
 
-## ALL FIGS AT 50% resistance threshold
+## Figures for paper 1 : version 7
 
-## Main Figures for paper 1 : version 6 - in progress
-### This file is automagically generated by R from paper1_results_figs_slimmed_50.Rmd
+## using rr_advantage instead of selection coefficient
+
+### automagically generated by R from paper1_results_figs_slimmed_50_rr.Rmd
 
@@
 
-  load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-  load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
+  ## to load previously saved runs  
+  load(file=paste0(outFolder,'listOutMix_rr_10000.rda'))
+  load(file=paste0(outFolder,'listOutI1_rr_10000.rda'))
+  load(file=paste0(outFolder,'listOutI2_rr_10000.rda'))
 
-  # load(file=paste0(outFolder,'listOutMix_unequal.rda'))
-  # load(file=paste0(outFolder,'listOutI1_unequal.rda'))
-  # load(file=paste0(outFolder,'listOutI2_unequal.rda'))
+  # load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
+  # load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
 
@@
 
-  rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
-  rownames(treeInput)[rownames(treeInput)=="s.RR2_0B"] <- "selection_coef_allele2"
+  # 14/6/16
+  # rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
+  # rownames(treeInput)[rownames(treeInput)=="s.RR2_0B"] <- "selection_coef_allele2"
 
@@
                              "dominance_allele2"=0.0016,
-                             "selection_coef_allele1"=0.23,
-                             "selection_coef_allele2"=0.43,
+                             # 14/6/16 rr_advantage = s / effectiveness
+                             #"selection_coef_allele1"=0.23,
+                             #"selection_coef_allele2"=0.43,
+                             "rr_advantage_I1"=0.23/0.73,
+                             "rr_advantage_I2"=0.43/1, 
                              "maleExposureProp"=1,
@@
 #curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
-runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_advantage_I1=0.23/0.73, rr_advantage_I2 = 0.43/1, addCombinedStrategy=FALSE )
 
 #curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , s.RR1_A0 = 0.23 , s.RR2_0B = 0.43, addCombinedStrategy=FALSE )
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_advantage_I1=0.23/0.73, rr_advantage_I2 = 0.43/1, addCombinedStrategy=FALSE )
 

### 2016-06-20 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
added specificity calcs, removed mix either from one fig and added gridlines back into PRCC plots

@@
              geom_hline(yintercept = 0, linetype=3) +
-             #theme_bw() +
              #theme(panel.grid.major.y = element_line(color = "grey", size = 1)) +
@@
 
+  #to put back in some ref gridlines that are removed by cowplot
+  #theme_set(theme_gray())
+  theme_set(theme_bw())
+
   #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
-  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.8))
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6))
 
@@
   #remove insecticide1 & 2 strategies
-  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
+  #ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
+  #20/6/16 try also removing "Mix either" 
+  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2","Mix either"), ]
+
   #subset which output plots to show in the main paper
@@
   #to rearrange order of legend entries
-  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either")) #, labels=c("MBB", "MAA", "MCC"))
+  #ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either")) 
+  #20/6/16 try removing "Mix either" 
+  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential"))
 
@@
              #geom_smooth(colour='red', linetype='dashed',size=0.5) +
-             geom_smooth(linetype='dashed',size=1) +
+             geom_smooth(linetype='dashed',size=0.5) +
              #facet_wrap( ~ strategy) +
@@
 ## Fig. x5 Effectiveness and exposure together determine most of whether resistance arises slower under adaptive mixture or sequential strategies
-
 1.
 Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
-
 ```{r, eval=TRUE, echo=FALSE, fig.height=6}
@@
   ggSubset <- rbind( ggSubset, df_same)
-
-  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mixA_not20pc_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
 
@@
   #(Intercept)          x1 
-  #1.019330    0.560508
+  ## 1.0746184 0.5590287
+
+  #20/6/16 calculated T/F positives & negatives, according to the linear model calculation
+  #this is a very heath-robinson way of doing, must be a better way
+  #so I want to find whether a point is above or below the line (and probably ignore zeros where mix & sequence the same)
+  #the eq is approx y=0.56*x+1.07
+  #so I could create a new column that has the expected lm val of added effectivenesses
+  #where +ve is mix better
+  #true+ve : effectiveness1+2 > lm & mixA_minus_seq0.5>0
+  #true-ve : effectiveness1+2 < lm & mixA_minus_seq0.5<0
+  #false+ve : effectiveness1+2 > lm & mixA_minus_seq0.5<0
+  #false-ve : effectiveness1+2 < lm & mixA_minus_seq0.5>0  
+  ggSubset$l_mod_eff <- l_mod$coefficients[2] * ggSubset$exposure + l_mod$coefficients[1]
+
+  #BEWARE that this sets 0s (where mix & sequence are equal) to NA
+  ggSubset$l_mod_confusion <- 
+    ifelse ( (ggSubset$effectiveness_insecticide1+ggSubset$effectiveness_insecticide2) > ggSubset$l_mod_eff,
+               ifelse(ggSubset$mixA_minus_seq0.5>0,"TP",ifelse(ggSubset$mixA_minus_seq0.5<0,"FP",NA))
+
+            ,ifelse(ggSubset$mixA_minus_seq0.5<0,"TN",ifelse(ggSubset$mixA_minus_seq0.5>0,"FN",NA))
+            )
+
+  table(ggSubset$l_mod_confusion)
+
+  #  FN   FP   TN   TP 
+  #602  219 3834 2787
+
+  #sensitivity or true positive rate TPR=TP/(TP+FN)
+  TPR = length(which(ggSubset$l_mod_confusion=="TP")) / (length(which(ggSubset$l_mod_confusion=="TP"))+length(which(ggSubset$l_mod_confusion=="FN")))
+  #specificity (SPC) or true negative rate SPC=TN/(TN+FP)
+  SPC = length(which(ggSubset$l_mod_confusion=="TN")) / (length(which(ggSubset$l_mod_confusion=="TN"))+length(which(ggSubset$l_mod_confusion=="FP")))
+
+  cat("sensitivity(TPR) : ",TPR)#,"\n")
+  cat("specificity(SPC) : ",SPC)#,"\n")
 
@@
 ```
-
 Red triangles are those runs in which resistance reached the threshold at the same time in the two strategies (and are repeated in both panels).
 The red line is a linear model through these same points.
-
 \pagebreak
 
-## Fig. x6 Modifying the previous plot to show where resistance arises over 20% slower in an adaptive mixture strategy versus sequential.
+## Fig. x6 Modifying previous plot to show where resistance arises over 20% slower for adaptive mixture versus sequential.
 
 1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the adaptive mixture under these constraints.
+```{r, eval=TRUE, echo=FALSE, fig.height=6}
 
-```{r, eval=TRUE, echo=FALSE, fig.height=6}
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mixA_not20pc_better_seq0.5 <- ifelse( ggSubset$mixA_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
 
@@
 
+#20/6/16 calculated T/F positives & negatives, according to the linear model calculation
+  #this is a very heath-robinson way of doing, must be a better way
+  #so I want to find whether a point is above or below the line (and probably ignore zeros where mix & sequence the same)
+  #the eq is approx y=0.56*x+1.07
+  #so I could create a new column that has the expected lm val of added effectivenesses
+  #where +ve is mix better
+  #true+ve : effectiveness1+2 > lm & mixA_minus_seq0.5>0
+  #true-ve : effectiveness1+2 < lm & mixA_minus_seq0.5<0
+  #false+ve : effectiveness1+2 > lm & mixA_minus_seq0.5<0
+  #false-ve : effectiveness1+2 < lm & mixA_minus_seq0.5>0  
+  ggSubset$l_mod_eff <- l_mod$coefficients[2] * ggSubset$exposure + l_mod$coefficients[1]
+
+  #BEWARE that this sets 0s (where mix & sequence are equal) to NA
+  ggSubset$l_mod_confusion <- 
+    ifelse ( (ggSubset$effectiveness_insecticide1+ggSubset$effectiveness_insecticide2) > ggSubset$l_mod_eff,
+               ifelse(ggSubset$mixA_not20pc_better_seq0.5==0,"TP",ifelse(ggSubset$mixA_not20pc_better_seq0.5==1,"FP",NA))
+
+            ,ifelse(ggSubset$mixA_not20pc_better_seq0.5==1,"TN",ifelse(ggSubset$mixA_not20pc_better_seq0.5==0,"FN",NA))
+            )
+
+  table(ggSubset$l_mod_confusion)
+
+  #  FN   FP   TN   TP 
+  ## 251 468 6178 915
+
+  #sensitivity or true positive rate TPR=TP/(TP+FN)
+  TPR = length(which(ggSubset$l_mod_confusion=="TP")) / (length(which(ggSubset$l_mod_confusion=="TP"))+length(which(ggSubset$l_mod_confusion=="FN")))
+  #specificity (SPC) or true negative rate SPC=TN/(TN+FP)
+  SPC = length(which(ggSubset$l_mod_confusion=="TN")) / (length(which(ggSubset$l_mod_confusion=="TN"))+length(which(ggSubset$l_mod_confusion=="FP")))
+
+  cat("sensitivity(TPR) : ",TPR)#,"\n")
+  cat("specificity(SPC) : ",SPC)#,"\n")
+
 ```
@@
     unequal <- ggSubset[ggSubset$resist_start_lo_div_hi < 0.01,]
-    nrow(ggSubset)
-    nrow(unequal)
+    #nrow(ggSubset)
+    #nrow(unequal)
     #2721 out of the 6189 remaining runs
@@
     #in how many runs was seq better than mix
-    sum(unequal$seq_better_mix20.5)
+    #sum(unequal$seq_better_mix20.5)
     #311
@@
 
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5} 
+#eval=FALSE so this one not displayed
+
 ## Fig.
@@
 In both cases resistance to the new insecticide develops slower when the 'old' insecticide continues to be used (mixture) than when it doesn't (sequence).
-
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5}
 

### 2016-06-24 inst/documents/paper1_results_figs_rr.Rmd
adding checks of number runs not reaching resistance thresholds, changing figs to effectiveness_added and dominance_added

@@
   experiment <- 'extended'
+
+  #doing; for all inputs
+  if (experiment=='extended') {
+    num_inputs <- 13
+    } else {
+    num_inputs <- 11  
+    }
 
@@
                              "effectiveness_insecticide2"=1,
+                             "effectiveness_added"=1.73,                             
                              "start_freq_allele1"=0.01,
@@
                              "dominance_allele2"=0.0016,
+                             "dominance_added"=0.1716,                             
                              # 14/6/16 rr_advantage = s / effectiveness
@@
 # getting data into format for ggplot
+  library(ggplot2)
 
@@
 
-  #subset by runs not making threshold
-  ggInsOuts <- ggInsOuts[-didntReachThresh,]
-
   #prettify output names
@@
 
-  #doing; for all inputs
-  if (experiment=='extended') {
-    num_inputs <- 13
-    } else {
-    num_inputs <- 11  
-    }
+  # #23/6/2016 add a check for which runs are being excluded (note that some are excluded in a different place too) 
+  # #just those runs that didn't make threshold
+  ggSubset <- ggInsOuts[didntReachThresh,]
+  names_inputs <- names(ggInsOuts)[1:num_inputs]
+  # 
+  # ggSubset <- ggSubset[ !ggSubset$strategy %in% c("insecticide 1","insecticide 2"), ]
+  # 
+  # for(i in names_inputs)
+  # {
+  #   y <- 'time_to_resistance0.5'    
+  #   
+  #   print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) + 
+  #            
+  #            geom_point(shape=3) #, show.legend=FALSE) + 
+  #            #geom_smooth(linetype='dashed',size=1)
+  #          )
+  # }
+
+  # #seems that most failed runs are at low start freqs
+  # hist(ggSubset$start_freq_allele1)
+  # #compare to all runs
+  # hist(ggInsOuts$start_freq_allele1)  
+  # #and for start freqs combined
+  # hist(ggSubset$start_freq_allele1+ggSubset$start_freq_allele2)
+  # #compare to all runs
+  # hist(ggInsOuts$start_freq_allele1+ggInsOuts$start_freq_allele2)
+  # #also low exposure is a large contributor to failed runs
+  # hist(ggSubset$exposure)  
+  # #compared to all runs
+  # hist(ggInsOuts$exposure)
+
+  #quick look at low start freq runs. Can they make it to thresholds ? Yes
+  # for(i in names_inputs)
+  # {
+  #   y <- 'time_to_resistance0.5'    
+  #   
+  #   print( ggplot(ggInsOuts[ggInsOuts$start_freq_allele1 < 0.001 & ggInsOuts$start_freq_allele2 < 0.001,], aes_string(x=i, y=y, colour="strategy")) + 
+  #            
+  #            geom_point(shape=3) + #, show.legend=FALSE) + 
+  #            geom_smooth(linetype='dashed',size=1)
+  #          )
+  # }
+
+  ####### end of new check
+
+  #subset by runs not making threshold
+  ggInsOuts <- ggInsOuts[-didntReachThresh,]
 
@@
 
-```{r, eval=FALSE, echo=FALSE, fig.height=3.5}
-
-  #10/5/2016 seems to crash R, so set to eval=FALSE
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
 
@@
 
+  ggSubset$effectiveness_added <- ggSubset$effectiveness_insecticide1 + ggSubset$effectiveness_insecticide2
+  ggSubset$dominance_added <- ggSubset$dominance_allele1 + ggSubset$dominance_allele2
+  names_inputs <- c(names_inputs, "effectiveness_added", "dominance_added")
+
   for(i in names_inputs)
   {
-    #x11()
-
     y <- 'time_to_resistance0.5'
 
-    #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
     print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) +
-            #points not wanted if 10000
-             #geom_point(shape=3, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+
+             geom_point(shape=3, size=0.2, alpha=0.2, show.legend=FALSE) + 
              geom_smooth(linetype='dashed',size=1) +
-             #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
-             labs(title = i) +
+
+             #labs(title = i) +
              #add line for Curtis inputs
              geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
-             #tryting to get the vline into the legend fails
-             #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
              geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
+           )
+  }
 
+```
+
+## Fig. xx2 Quick check whether difference between effectiveness 1 & 2 dissapears when minimising excluded runs
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.5}
+
+  #names_inputs <- names(ggInsOuts)[1:num_inputs]
+  names_inputs <- c("effectiveness_insecticide1", "effectiveness_insecticide2", "effectiveness_added")
+
+  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2","Mix either","Sequential"), ]
+
+  ggSubset$effectiveness_added <- ggSubset$effectiveness_insecticide1 + ggSubset$effectiveness_insecticide2
+
+  # ggSubset <- ggSubset[ ggSubset$start_freq_allele1 > 0.005 & 
+  #                       ggSubset$start_freq_allele2 > 0.005 &
+  #                       ggSubset$dominance_allele1 > 0.1 &
+  #                       ggSubset$dominance_allele2 > 0.1 &
+  #                       ggSubset$exposure > 0.3  , ]
+
+  for(i in names_inputs)
+  {
+    y <- 'time_to_resistance0.5'
+
+    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) +
+
+             geom_point(shape=1, size=1, alpha=0.2, show.legend=FALSE) + 
+             geom_smooth(linetype='dashed',size=1)
+
+             #labs(title = i) +
+             #add line for Curtis inputs
+             #geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
+             #geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
            )

### 2016-06-24 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
adding checks of number runs not reaching resistance thresholds, changing figs to effectiveness_added and dominance_added

@@
                              "dominance_allele2"=0.0016,
+                             "dominance_added"=0.1716,  
                              # 14/6/16 rr_advantage = s / effectiveness
@@
 
+  #22/6/2016 to check % of runs excluded 
+  #its complicated given that there are runs for I1,I2 & mix 
+  #this is an approximation
+  #TODO, are any others excluded later ?
+  percentExcluded <- 100*length(didntReachThresh)/nrow(ggInsOuts)
+  #16.5
+
   #subset by runs not making threshold
@@
   didntReachThresh <- which( dif_mixA_seq$gen_cP0.5 > 500 | dif_mixA_seq$gen_cP0.5seq > 500 )
+
+  #TODO 22/6/16 RECHECK that nothing dodgy going on here, seems that more runs being excluded
+  #but this is because its going back to the raw files
+  percentExcluded2 <- 100*length(didntReachThresh)/nrow(dif_mixA_seq)
+  #24%
+
   #subset by runs not making threshold
@@
   didntReachThresh <- which( dif_mix2_seq$gen_cP0.5 > 500 | dif_mix2_seq$gen_cP0.5seq > 500 )
+
+  percentExcluded3 <- 100*length(didntReachThresh)/nrow(dif_mix2_seq)
+  #30%
+
+  ####################
+  #23/6/2016 experimenting to see if can reduce the % excluded by tweaking input ranges
+  #20%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$start_freq_allele1 > 0.005, ]
+  #16%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$start_freq_allele1 > 0.005 & dif_mix2_seq$start_freq_allele2 > 0.005, ]
+  #18%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$exposure > 0.25, ]
+  #16%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$exposure > 0.3, ]
+  #7%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$start_freq_allele1 > 0.005 & dif_mix2_seq$start_freq_allele2 > 0.005 & dif_mix2_seq$exposure > 0.25, ]   
+  #5%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$start_freq_allele1 > 0.005 & dif_mix2_seq$start_freq_allele2 > 0.005 & dif_mix2_seq$exposure > 0.3, ] 
+  #26%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_insecticide1 > 0.4 & dif_mix2_seq$effectiveness_insecticide2 > 0.4, ]
+  #13%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_insecticide1 > 0.4 & dif_mix2_seq$effectiveness_insecticide2 > 0.4 & dif_mix2_seq$exposure > 0.3, ]
+  #10%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$start_freq_allele1 > 0.001 & dif_mix2_seq$start_freq_allele2 > 0.001 & dif_mix2_seq$exposure > 0.3, ] 
+  #8%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_insecticide1 > 0.4 & dif_mix2_seq$effectiveness_insecticide2 > 0.4 & dif_mix2_seq$start_freq_allele1 > 0.001 & dif_mix2_seq$start_freq_allele2 > 0.001 & dif_mix2_seq$exposure > 0.3, ]
+  #22%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$dominance_allele1 > 0.1 & dif_mix2_seq$dominance_allele2 > 0.1, ] 
+  #5%
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$start_freq_allele1 > 0.001 & dif_mix2_seq$start_freq_allele2 > 0.001 & dif_mix2_seq$exposure > 0.3 & dif_mix2_seq$dominance_allele1 > 0.1 & dif_mix2_seq$dominance_allele2 > 0.1, ]
+
+  # didntReachThresh <- which( dif_mix2_seq2$gen_cP0.5 > 500 | dif_mix2_seq2$gen_cP0.5seq > 500 )
+  # 
+  # percentExcluded4 <- 100*length(didntReachThresh)/nrow(dif_mix2_seq2)
+  # 
+  # percentExcluded4
+
+  #### end of expt 23/6/2016
+
   #subset by runs not making threshold
@@
       ylab('Time to resistance (gens)') +
-      scale_x_discrete("", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Seq 1st", "Mix either"="Mix 1st","Sequential"="Seq both", "Mix adaptive"="Adaptive", "Mix both"="Mix both" ) )
+      scale_x_discrete("", limits=c("insecticide 1","Mix either","Sequential","Mix adaptive","Mix both"), labels = c("insecticide 1"="Seq 1st", "Mix either"="Mix 1st","Sequential"="Seq both", "Mix adaptive"="Adaptive both", "Mix both"="Mix both" ) )
 
@@
   strategies <- c("Sequential", "Mix adaptive", "Mix both")
-  strat_name <- c("Seq both","Adaptive both","Mix both")
+  strat_name <- c("Sequential both","Adaptive both","Mix both")
 
@@
 
-  #names_inputs <- names(ggInsOuts)[1:num_inputs]
+  #remove some strategies for plots
+  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2","Mix either"), ]
 
-  #remove insecticide1 & 2 strategies
-  #ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2"), ]
-  #20/6/16 try also removing "Mix either" 
-  ggSubset <- ggInsOuts[ !ggInsOuts$strategy %in% c("insecticide 1","insecticide 2","Mix either"), ]
+  # create added columns for effectiveness and dominance
+  ggSubset$effectiveness_added <- ggSubset$effectiveness_insecticide1 + ggSubset$effectiveness_insecticide2
+  ggSubset$dominance_added <- ggSubset$dominance_allele1 + ggSubset$dominance_allele2
 
@@
   #             "dominance_allele1","selection_coef_allele1","maleExposureProp")
-  in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")
+  #in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")  
+  in_cols <- c("exposure","effectiveness_added","dominance_added")
 
@@
 
-  #num_inputs <- length(in_cols)
-  #names_inputs <- in_cols
-
   #to rearrange order of legend entries
-  #ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either")) 
+  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential", "Mix either")) 
   #20/6/16 try removing "Mix either" 
-  ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential"))
-
-  #TODO could I facet the plots below rather than using the loop
-  #probably need to put all inputs into a single column and then facet by that
+  #ggSubset$strategy <- factor(ggSubset$strategy, levels=c("Mix both", "Mix adaptive", "Sequential"))
 
@@
     {
-    #x11()
 
@@
     print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) +
-            #points not wanted if 10000
-             #geom_point(shape=3, show.legend=FALSE) + 
-             #geom_smooth(colour='red', linetype='dashed',size=0.5) +
+
+             #ideally would include points but they obscure means
+             #geom_point(shape=1, size=1, alpha=0.2, show.legend=FALSE) + 
              geom_smooth(linetype='dashed',size=0.5) +
              #facet_wrap( ~ strategy) +
-             #geom_smooth(aes_string(x=i, y=y, color=NULL)) )
              #labs(title = i) +
@@
              geom_vline(data = curtisInputs, aes_string(xintercept = i), linetype='dotted', colour='red') +
-             #tryting to get the vline into the legend fails
+             #trying to get the vline into the legend fails
              #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
@@
 
-1. Exposure and effectiveness come out as the most important effect on the difference between the sequential and adpative mixture strategies.
+1. Exposure and effectiveness come out as the most important effect on the difference between A) sequential and mixture, and B) difference between sequential and adaptive mixture strategies.
 
@@
 
-    #results are here I can probably rbind them together into a df that I can ggplot
-    #pcc_res$PRCC
-
     to_plot <- pcc_res$PRCC
@@
     to_plot$inputs <- rownames(to_plot)
+
+    cat("A)")
 
@@
     to_plot$inputs <- rownames(to_plot)
+
+    cat("B)")
 

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
change fig labels

@@
 
-             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance slower','1'='sequential\nresistance slower'))) +
+             #facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance slower','1'='sequential\nresistance slower'))) +
+
+             facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix\ntime to resistance\n longer','1'='sequential\ntime to resistance\n longer'))) +
 
@@
 
-             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\nresistance >20% slower','1'='adaptive mix\nresistance <20% slower'))) +
+             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\ntime to resistance\n >20% longer','1'='adaptive mix\ntime to resistance\n >20% longer'))) +
 
@@
 
-             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mixture\nresistance slower','1'='sequential\nresistance slower'))) +
+             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mixture\ntime to resistance\nlonger','1'='sequential\ntime to resistance\nlonger'))) +
 

### 2016-06-27 inst/documents/keep_old_insecticide.Rmd
rename correct mix & male exposure inputs

@@
 if (experiment=='extended')
-  treePredictors <- c(treePredictors,'maleExposureProp','correctMixDeployProp')
+  treePredictors <- c(treePredictors,'male_exposure_prop','correct_mix_deploy')
 
@@
                            "selection_coef_allele2"=0.43,
-                           "maleExposureProp"=1,
-                           "correctMixDeployProp"=1,
+                           "male_exposure_prop"=1,
+                           "correct_mix_deploy"=1,
                            "resist_start_lo_div_hi"=1,

### 2016-06-27 inst/documents/pap1_figs_slimmed_50_unequal.Rmd
rename correct mix & male exposure inputs

@@
   if (experiment=='extended')
-    treePredictors <- c(treePredictors,'maleExposureProp','correctMixDeployProp')
+    treePredictors <- c(treePredictors,'male_exposure_prop','correct_mix_deploy')
 
@@
                              "selection_coef_allele2"=0.43,
-                             "maleExposureProp"=1,
-                             "correctMixDeployProp"=1,
+                             "male_exposure_prop"=1,
+                             "correct_mix_deploy"=1,
                              "resist_start_lo_div_hi"=1,
@@
   ggSubset <- ggInsOuts[ ggInsOuts$strategy %in% c("Sequential", "Mix adaptive", "Mix both"), ]
-
-  #subset which output plots to show in the main paper
-  # in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
-  #             "dominance_allele1","selection_coef_allele1","maleExposureProp")
 

### 2016-06-27 inst/documents/paper1_results_figs.Rmd
moved documents into own folder
rename correct mix & male exposure inputs

--- 2016-04-28 paper1_results_figs.Rmd
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
 
@@
              #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             geom_point(data = curtisInputs, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
              ylab('effectiveness of insecticides added') +

### 2016-06-27 inst/documents/paper1_results_figs_rr.Rmd
rename correct mix & male exposure inputs

@@
                              "rr_advantage_I2"=0.43/1,                             
-                             "maleExposureProp"=1,
-                             "correctMixDeployProp"=1,
+                             "male_exposure_prop"=1,
+                             "correct_mix_deploy"=1,
                              "resist_start_lo_div_hi"=1,
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure+exposure*maleExposureProp, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.5))+
 

### 2016-06-27 inst/documents/paper1_results_figs_slimmed.Rmd
rename correct mix & male exposure inputs

--- 2016-03-24 paper1_results_figs_slimmed.Rmd
@@
   in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
-              "dominance_allele1","selection_coef_allele1","maleExposureProp")
+              "dominance_allele1","selection_coef_allele1","male_exposure_prop")
   out_cols <- c("time_to_resistance0.25","strategy")

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50.Rmd
rename correct mix & male exposure inputs

--- 2016-03-30 paper1_results_figs_slimmed_50.Rmd
@@
   # in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
-  #             "dominance_allele1","selection_coef_allele1","maleExposureProp")
+  #             "dominance_allele1","selection_coef_allele1","male_exposure_prop")
   in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
rename correct mix & male exposure inputs

@@
                              "rr_advantage_I2"=0.43/1, 
-                             "maleExposureProp"=1,
-                             "correctMixDeployProp"=1,
+                             "male_exposure_prop"=1,
+                             "correct_mix_deploy"=1,
                              "resist_start_lo_div_hi"=1,
@@
   # in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
-  #             "dominance_allele1","selection_coef_allele1","maleExposureProp")
+  #             "dominance_allele1","selection_coef_allele1","male_exposure_prop")
   #in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")  

### 2016-06-27 R/runcurtis_f2.r

@@ -28,8 +28,8 @@
 #' @param z.RR1_00 fitness cost of resistance allele 1 in insecticide free environment
 #' @param z.RR2_00 fitness cost of resistance allele 2 in insecticide free environment
 #' @param sexLinked whether resistance is sex linked, default=0(FALSE)
-#' @param maleExposureProp male exposure as a propoertion of female, default 1 for same, likely <1  
-#' @param correctMixDeployProp proportion of correct deployment of mixtures, 
+#' @param male_exposure_prop male exposure as a propoertion of female, default 1 for same, likely <1  
+#' @param  proportion of correct deployment of mixtures, 
 #'    if <1 other portion divided between single insecticides
 #' @param addCombinedStrategy whether to add a line for the combined mixture/individual strategy, makes plot confusing so set to FALSE by default
 #' @param addStrategyLabels add labels to plot for where strategies cross 0.5 line TRUE gives seq,mix1,mix2,mix3
@@ -68,8 +68,8 @@ runcurtis_f2 <- function( max_gen = 500,
                           z.RR1_00	=	0	,
                           z.RR2_00	=	0	,
                           sexLinked = 0 ,
-                          maleExposureProp = 1,
-                          correctMixDeployProp = 1,
+                          male_exposure_prop = 1,
+                           = 1,
                           addCombinedStrategy = TRUE,
                           addStrategyLabels = TRUE,
                           strategyLabels = c("seq","mix1","adapt","mix2"),
@@ -80,8 +80,8 @@ runcurtis_f2 <- function( max_gen = 500,
   
   #fill exposure array I1
   a <- setExposure(exposure=exposure, insecticideUsed = "insecticide1",
-                   maleExposureProp = maleExposureProp,
-                   correctMixDeployProp = correctMixDeployProp )
+                   male_exposure_prop = male_exposure_prop,
+                    =  )
   
   inputI1 <- setInputOneScenario(a=a,
                                  max_gen = max_gen,
@@ -111,13 +111,13 @@ runcurtis_f2 <- function( max_gen = 500,
                                  z.RR1_00	=	z.RR1_00	,
                                  z.RR2_00	=	z.RR2_00	,
                                  sexLinked = sexLinked ,
-                                 maleExposureProp = maleExposureProp,
-                                 correctMixDeployProp = correctMixDeployProp )
+                                 male_exposure_prop = male_exposure_prop,
+                                  =  )
   
   #fill exposure array I2
[…]

### 2016-06-27 R/sensiAnPaperPart.r

@@ -69,18 +69,18 @@ sensiAnPaperPart <- function( nScenarios = 10,
     
     if (experiment=='extended')
     {
-      maleExposureProp <- runif(1, min=0, max=1)
-      correctMixDeployProp <- runif(1, min=0.5, max=1)          
+      male_exposure_prop <- runif(1, min=0, max=1)
+      correct_mix_deploy <- runif(1, min=0.5, max=1)          
     } else
     {
-      maleExposureProp <- 1
-      correctMixDeployProp <- 1     
+      male_exposure_prop <- 1
+      correct_mix_deploy <- 1     
     }
     
     
     #this sets exposures according to whether insecticide1,2 or mixture
     a <- setExposure(exposure=exposure, insecticideUsed=insecticideUsed,
-                     maleExposureProp=maleExposureProp, correctMixDeployProp=correctMixDeployProp) 
+                     male_exposure_prop=male_exposure_prop, correct_mix_deploy=correct_mix_deploy) 
   
     
     ## effectiveness of insecticides, selection against SS in presence of that insecticide
@@ -134,8 +134,8 @@ sensiAnPaperPart <- function( nScenarios = 10,
                         # 14/6/16 now these calc from rr_ inside the func
                         # s.RR1_A0 = s.RR1_A0,
                         # s.RR2_0B = s.RR2_0B,
-                        maleExposureProp = maleExposureProp,
-                        correctMixDeployProp = correctMixDeployProp,
+                        male_exposure_prop = male_exposure_prop,
+                        correct_mix_deploy = correct_mix_deploy,
                         # 14/6/16
                         rr_advantage_I1 = rr_advantage_I1,
                         rr_advantage_I2 = rr_advantage_I2

### 2016-06-27 R/setExposure.r

@@ -6,8 +6,8 @@
 #' 
 #' @param exposure exposure to the insecticide(s)
 #' @param insecticideUsed one of 'insecticide1','insecticide2','mixture'
-#' @param maleExposureProp proportion tht males are exposed relative to f, default 1, likely to be <1
-#' @param correctMixDeployProp proportion of times that mixture is deployed correctly, 
+#' @param male_exposure_prop proportion tht males are exposed relative to f, default 1, likely to be <1
+#' @param correct_mix_deploy proportion of times that mixture is deployed correctly, 
 #'    assumes that when not deployed correctly the single insecticides are used instead
 #'    TODO check my understanding of this
 #' 
@@ -19,8 +19,8 @@
 #' 
 setExposure <- function( exposure = 0.9,
                          insecticideUsed = 'mixture',
-                         maleExposureProp = 1,
-                         correctMixDeployProp = 1 )
+                         male_exposure_prop = 1,
+                         correct_mix_deploy = 1 )
 {
 
   #exposure to insecticide
@@ -33,7 +33,7 @@ setExposure <- function( exposure = 0.9,
     #f
     a['f','A','0'] <- exposure
     #m
-    a['m','A','0'] <- exposure * maleExposureProp    
+    a['m','A','0'] <- exposure * male_exposure_prop    
     #for both m&f
     a[,'0','0'] <- 1 - a[,'A','0']  
     
@@ -42,22 +42,22 @@ setExposure <- function( exposure = 0.9,
     #f
     a['f','0','B'] <- exposure
     #m
-    a['m','0','B'] <- exposure * maleExposureProp     
+    a['m','0','B'] <- exposure * male_exposure_prop     
     #for both m&f
     a[,'0','0'] <- 1 - a[,'0','B']  
     
   } else if (insecticideUsed == "mixture")
   {
     #f
-    a['f','A','B'] <- exposure * correctMixDeployProp
+    a['f','A','B'] <- exposure * correct_mix_deploy
     #m
-    a['m','A','B'] <- a['f','A','B'] * maleExposureProp
+    a['m','A','B'] <- a['f','A','B'] * male_exposure_prop
     
-    if ( correctMixDeployProp < 1 )
+    if ( correct_mix_deploy < 1 )
     {
-      a['f','A','0'] <- (1 - correctMixDeployProp)/2 * exposure
[…]

### 2016-06-27 R/setInputOneScenario.r

@@ -60,8 +60,8 @@
 #' @param niche_Ab insecticide niche toggle hi1 lo2 0=off 1=on
 #' @param niche_aB insecticide niche toggle lo1 hi2 0=off 1=on
 #' @param sexLinked whether resistance is sex linked, default=0(FALSE)
-#' @param maleExposureProp male exposure as a propoertion of female, default 1 for same, likely <1  
-#' @param correctMixDeployProp proportion of correct deployment of mixtures, 
+#' @param male_exposure_prop male exposure as a propoertion of female, default 1 for same, likely <1  
+#' @param correct_mix_deploy proportion of correct deployment of mixtures, 
 #'    if <1 other portion divided between single insecticides
 #' @param rr_advantage_I1 effect of resistance in overcoming insecticide 1 effectiveness in RR
 #' @param rr_advantage_I2 effect of resistance in overcoming insecticide 2 effectiveness in RR    
@@ -124,8 +124,8 @@ setInputOneScenario <- function( calibration = 100,
                                  niche_Ab	=	1	,
                                  niche_aB	=	1 ,
                                  sexLinked = 0,
-                                 maleExposureProp = 1,
-                                 correctMixDeployProp = 1,
+                                 male_exposure_prop = 1,
+                                 correct_mix_deploy = 1,
                                  exposure = 0.9,
                                  #14/6/16
                                  rr_advantage_I1 = 1,
@@ -224,8 +224,8 @@ setInputOneScenario <- function( calibration = 100,
   
   #22/1/16 new variables for extended experiment
   #they aren't used in runModel2() but are needed for post run analyses
-  input[	54	] <-	maleExposureProp
-  input[	55	] <-	correctMixDeployProp
+  input[	54	] <-	male_exposure_prop
+  input[	55	] <-	correct_mix_deploy
   #1/2/16 allowing saving of single exposure param just for use in post-run analyses
   input[	56	] <-	exposure  
   

### 2016-06-27 inst/shiny/MixSeqResist1/ui.r

@@ -108,7 +108,7 @@ shinyUI(fixedPage(
     )
     # column(2, offset = 0,
     #        h5("Extra params not included in Curtis"),
-    #        numericInput("correctMixDeployProp", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
+    #        numericInput("correct_mix_deploy", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
     #        checkboxInput("addCombinedStrategy", "plot adaptive strategy",FALSE)
     # )
   ) #end fixed row

### 2016-06-27 inst/shiny/resistmob1/server.r

@@ -43,7 +43,7 @@ shinyServer(function(input, output) {
                       s.RR2_0B = input$advantage_A2,
                       addCombinedStrategy = FALSE,
                       strategyLabels = c('seq','','adapt','mix2') )
-                      #correctMixDeployProp = input$correctMixDeployProp,
+                      #correct_mix_deploy = input$correct_mix_deploy,
                       #addCombinedStrategy = input$addCombinedStrategy )
         
         #a hack to output the inputs
@@ -62,7 +62,7 @@ shinyServer(function(input, output) {
                       "strategyLabels = c('seq','','adapt','mix2')",
                       ")\n" )
         
-        #"correctMixDeployProp =",input$correctMixDeployProp,",",
+        #"correct_mix_deploy =",input$correct_mix_deploy,",",
         #"addCombinedStrategy =",input$addCombinedStrategy, 
                
       }) #end isolate  
@@ -89,7 +89,7 @@ shinyServer(function(input, output) {
                       s.RR2_0B = input$advantage_B2,
                       addCombinedStrategy = FALSE,
                       strategyLabels = c('seq','','adapt','mix2') )
-        #correctMixDeployProp = input$correctMixDeployProp,
+        #correct_mix_deploy = input$correct_mix_deploy,
         #addCombinedStrategy = input$addCombinedStrategy )
         
         #a hack to output the inputs

### 2016-06-27 inst/shiny/resistmob2/server.r

@@ -47,7 +47,7 @@ shinyServer(function(input, output) {
                       
                       addCombinedStrategy = FALSE,
                       strategyLabels = c('seq','','adapt','mix2') )
-                      #correctMixDeployProp = input$correctMixDeployProp,
+                      #correct_mix_deploy = input$correct_mix_deploy,
                       #addCombinedStrategy = input$addCombinedStrategy )
         
         #a hack to output the inputs
@@ -66,7 +66,7 @@ shinyServer(function(input, output) {
                       "strategyLabels = c('seq','','adapt','mix2')",
                       ")\n" )
         
-        #"correctMixDeployProp =",input$correctMixDeployProp,",",
+        #"correct_mix_deploy =",input$correct_mix_deploy,",",
         #"addCombinedStrategy =",input$addCombinedStrategy, 
                
       }) #end isolate  
@@ -97,7 +97,7 @@ shinyServer(function(input, output) {
                       
                       addCombinedStrategy = FALSE,
                       strategyLabels = c('seq','','adapt','mix2') )
-        #correctMixDeployProp = input$correctMixDeployProp,
+        #correct_mix_deploy = input$correct_mix_deploy,
         #addCombinedStrategy = input$addCombinedStrategy )
         
         #a hack to output the inputs

### 2016-06-27 inst/shiny/resistmob2/ui.r

@@ -139,7 +139,7 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
     )
     # column(2, offset = 0,
     #        h5("Extra params not included in Curtis"),
-    #        numericInput("correctMixDeployProp", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
+    #        numericInput("correct_mix_deploy", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
     #        checkboxInput("addCombinedStrategy", "plot adaptive strategy",FALSE)
     # )
   ) #end fixed row

### 2016-06-27 inst/shiny/shinyFig2Curtis/server.r

@@ -38,7 +38,7 @@ shinyServer(function(input, output) {
                       phi.SS2_0B = input$phi.SS2_0B,
                       s.RR1_A0 = input$s.RR1_A0,
                       s.RR2_0B = input$s.RR2_0B,
-                      correctMixDeployProp = input$correctMixDeployProp,
+                      correct_mix_deploy = input$correct_mix_deploy,
                       addCombinedStrategy = input$addCombinedStrategy )
         
         #a hack to output the inputs
@@ -52,7 +52,7 @@ shinyServer(function(input, output) {
                       "phi.SS2_0B =",input$phi.SS2_0B,",",
                       "s.RR1_A0 =",input$s.RR1_A0,",",
                       "s.RR2_0B =",input$s.RR2_0B,",",
-                      "correctMixDeployProp =",input$correctMixDeployProp,",",
+                      "correct_mix_deploy =",input$correct_mix_deploy,",",
                       "addCombinedStrategy =",input$addCombinedStrategy,
                       ")\n" )
         

### 2016-06-27 inst/shiny/shinyFig2Curtis/ui.r

@@ -54,7 +54,7 @@ shinyUI(fluidPage(
     ),
     column(2, offset = 0,
            h5("Extra params not included in Curtis"),
-           numericInput("correctMixDeployProp", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
+           numericInput("correct_mix_deploy", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1),
            checkboxInput("addCombinedStrategy", "plot adaptive strategy",FALSE)
     )
   ), #end fluid row

### 2016-06-27 inst/shiny/shinyNiche/server.r

@@ -15,8 +15,8 @@ shinyServer(function(input, output) {
     #isolate({
       a <- setExposure(exposure = input$exposure,
                        insecticideUsed = input$insecticideUsed,
-                       maleExposureProp = input$maleExposureProp, 
-                       correctMixDeployProp = input$correctMixDeployProp)
+                       male_exposure_prop = input$male_exposure_prop, 
+                       correct_mix_deploy = input$correct_mix_deploy)
       
       #to replace 0s with NAs to look better in table
       a[which(a[,,]==0)] <- NA
@@ -29,8 +29,8 @@ shinyServer(function(input, output) {
     
     a <- setExposure(exposure = input$exposure,
                      insecticideUsed = input$insecticideUsed,
-                     maleExposureProp = input$maleExposureProp, 
-                     correctMixDeployProp = input$correctMixDeployProp)
+                     male_exposure_prop = input$male_exposure_prop, 
+                     correct_mix_deploy = input$correct_mix_deploy)
     
     #to replace 0s with NAs to look better in table
     a[which(a[,,]==0)] <- NA

### 2016-06-27 inst/shiny/shinyNiche/ui.r

@@ -28,10 +28,10 @@ shinyUI(fluidPage(
            #numericInput("a.m_B", "insecticide2:", 0.5, min = 0.1, max = 0.9, step = 0.1)
     ),
     column(2, offset = 0,
-           numericInput("correctMixDeployProp", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1)
+           numericInput("correct_mix_deploy", "Mixture correct deployment", 1, min = 0, max = 1, step = 0.1)
     ),
     column(2, offset = 0,
-           numericInput("maleExposureProp", "Male exposure proportion", 0.9, min = 0, max = 1, step = 0.1)
+           numericInput("male_exposure_prop", "Male exposure proportion", 0.9, min = 0, max = 1, step = 0.1)
     )
   ), #end fluid row
 

### 2016-06-27 inst/documents/paper1_results_figs_rr.Rmd
renamed rr_restoration

@@
 
-# Figures for paper 1 after change to using rr_advantage instead of selection coeff from sensi analysis
+# Figures for paper 1 after change to using rr_restoration instead of selection coeff from sensi analysis
 # version 1 - in progress
@@
   #treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
-  #14/6/16 change to rr_advantage
-  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','rr_advantage_I1','rr_advantage_I2')
+  #14/6/16 change to rr_restoration
+  treePredictors <- c('P_1','P_2','exposure','phi.SS1_A0','phi.SS2_0B','h.RS1_A0','h.RS2_0B','rr_restoration_ins1','rr_restoration_ins2')
 
@@
 
-  #14/6/16 rr_advantage these not to be used anymore
+  #14/6/16 rr_restoration these not to be used anymore
   #rownames(treeInput)[rownames(treeInput)=="s.RR1_A0"] <- "selection_coef_allele1"
@@
                              "dominance_added"=0.1716,                             
-                             # 14/6/16 rr_advantage = s / effectiveness
+                             # 14/6/16 rr_restoration = s / effectiveness
                              #"selection_coef_allele1"=0.23,
                              #"selection_coef_allele2"=0.43,
-                             "rr_advantage_I1"=0.23/0.73,
-                             "rr_advantage_I2"=0.43/1,                             
+                             "rr_restoration_ins1"=0.23/0.73,
+                             "rr_restoration_ins2"=0.43/1,                             
                              "male_exposure_prop"=1,

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
renamed rr_restoration

@@
 
-## using rr_advantage instead of selection coefficient
+## using rr_restoration instead of selection coefficient
 
@@
                              "dominance_added"=0.1716,  
-                             # 14/6/16 rr_advantage = s / effectiveness
+                             # 14/6/16 rr_restoration = s / effectiveness
                              #"selection_coef_allele1"=0.23,
                              #"selection_coef_allele2"=0.43,
-                             "rr_advantage_I1"=0.23/0.73,
-                             "rr_advantage_I2"=0.43/1, 
+                             "rr_restoration_ins1"=0.23/0.73,
+                             "rr_restoration_ins2"=0.43/1, 
                              "male_exposure_prop"=1,
@@
 #curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
-runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_advantage_I1=0.23/0.73, rr_advantage_I2 = 0.43/1, addCombinedStrategy=FALSE )
+runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_restoration_ins1=0.23/0.73, rr_restoration_ins2 = 0.43/1, addCombinedStrategy=FALSE )
 
 #curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_advantage_I1=0.23/0.73, rr_advantage_I2 = 0.43/1, addCombinedStrategy=FALSE )
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_restoration_ins1=0.23/0.73, rr_restoration_ins2 = 0.43/1, addCombinedStrategy=FALSE )
 

### 2016-06-27 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
renamed rr_restoration

@@
                       phi.SS2_0B = 0.5,
-                      rr_advantage_I1 = 0.5,
-                      rr_advantage_I2 = 0.5 )
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5 )
 
@@
                       phi.SS2_0B = 0.5,
-                      rr_advantage_I1 = 0.2,
-                      rr_advantage_I2 = 0.5 )
+                      rr_restoration_ins1 = 0.2,
+                      rr_restoration_ins2 = 0.5 )
 
@@
                       phi.SS2_0B = 0.5,
-                      rr_advantage_I1 = 0.8,
-                      rr_advantage_I2 = 0.5 )
+                      rr_restoration_ins1 = 0.8,
+                      rr_restoration_ins2 = 0.5 )
 
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=rr_advantage_I1)) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=rr_restoration_ins1)) + 
          labs(colour = "selective advantage") +
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 1. All control inputs equal at 0.5")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 2. Effectiveness of insecticide1 increased from 0.5 to 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 3. Effectiveness of both insecticides 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 4. Dominance of insecticide1 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 5. Dominance of both insecticides 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.1 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 6. Selective advantage of insecticide1 0.1")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.1 , rr_advantage_I2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 7. Selective advantage of both insecticides 0.1")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.1 , rr_advantage_I2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 8. Selective advantage of both insecticides 0.1, effectiveness to 0.7")
@@
 
-#cool contrast between these scenarios modifying effectiveness and rr_advantage in which all defaults set to 0.5 except
+#cool contrast between these scenarios modifying effectiveness and rr_restoration in which all defaults set to 0.5 except
 TODO incorporate this into text above
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. AA. effectiveness to 0.8")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.8 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. BB rr_advantage I1 0.8")
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. BB rr_restoration I1 0.8")
 ```
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.2 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. CC. rr_advantage I2 0.2")
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.2 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+cat("Fig. CC. rr_restoration I2 0.2")
 ```
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 9. Exposure 0.9")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 10. Effectiveness 0.7, sequence favoured")
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_advantage_I1 = 0.5 , rr_advantage_I2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 cat("Fig. 11. Effectiveness 0.7, Exposure 0.9, mixture favoured")

### 2016-06-27 R/runcurtis_f2.r

@@ -23,8 +23,8 @@
 #' @param s.RR1_A0 Selection coefficient locus1 in A
 #' @param s.RR2_0b Selection coefficient locus2 in b
 #' @param s.RR2_0B Selection coefficient locus2 in B
-#' @param rr_advantage_I1 effect of resistance in overcoming insecticide 1 effectiveness in RR
-#' @param rr_advantage_I2 effect of resistance in overcoming insecticide 2 effectiveness in RR  
+#' @param rr_restoration_ins1 effect of resistance in overcoming insecticide 1 effectiveness in RR
+#' @param rr_restoration_ins2 effect of resistance in overcoming insecticide 2 effectiveness in RR  
 #' @param z.RR1_00 fitness cost of resistance allele 1 in insecticide free environment
 #' @param z.RR2_00 fitness cost of resistance allele 2 in insecticide free environment
 #' @param sexLinked whether resistance is sex linked, default=0(FALSE)
@@ -61,9 +61,9 @@ runcurtis_f2 <- function( max_gen = 500,
                           s.RR1_A0	=	NULL	,
                           s.RR2_0b	=	0	,
                           s.RR2_0B	=	NULL	,
-                          # 14/6/16 rr_advantage = s / effectiveness
-                          rr_advantage_I1 = 0.23/0.73,
-                          rr_advantage_I2 = 0.43/1,    
+                          # 14/6/16 rr_restoration = s / effectiveness
+                          rr_restoration_ins1 = 0.23/0.73,
+                          rr_restoration_ins2 = 0.43/1,    
 
                           z.RR1_00	=	0	,
                           z.RR2_00	=	0	,
@@ -105,8 +105,8 @@ runcurtis_f2 <- function( max_gen = 500,
                                  s.RR2_0b	=	s.RR2_0b	,
                                  s.RR2_0B	=	s.RR2_0B	,
                                  # 14/6/16
-                                 rr_advantage_I1 = rr_advantage_I1,
-                                 rr_advantage_I2 = rr_advantage_I2,    
+                                 rr_restoration_ins1 = rr_restoration_ins1,
+                                 rr_restoration_ins2 = rr_restoration_ins2,    
                                  
                                  z.RR1_00	=	z.RR1_00	,
                                  z.RR2_00	=	z.RR2_00	,
@@ -141,8 +141,8 @@ runcurtis_f2 <- function( max_gen = 500,
                                  s.RR2_0b	=	s.RR2_0b	,
                                  s.RR2_0B	=	s.RR2_0B	,
[…]

### 2016-06-27 R/sensiAnPaperPart.r

@@ -86,7 +86,7 @@ sensiAnPaperPart <- function( nScenarios = 10,
     ## effectiveness of insecticides, selection against SS in presence of that insecticide
     # phi.SS1_A0 <- runif(1, min=0.45, max=1)
     # phi.SS2_0B <- runif(1, min=0.45, max=1)
-    # 14/6/16 increased range now that we are using rr_advantage to avoid fitness > 1
+    # 14/6/16 increased range now that we are using rr_restoration to avoid fitness > 1
     phi.SS1_A0 <- runif(1, min=0.3, max=1)
     phi.SS2_0B <- runif(1, min=0.3, max=1)      
     
@@ -106,15 +106,15 @@ sensiAnPaperPart <- function( nScenarios = 10,
     # s.RR1_A0 <- runif(1, min=0.1, max=0.45)
     # s.RR2_0B <- runif(1, min=0.1, max=0.45)     
     
-    # 14/6/16 to add rr_advantage_I1 & I2
+    # 14/6/16 to add rr_restoration_ins1 & I2
     # as an alternative to varying s directly from sensi analysis
     # to allow a wider range (higher values) while avoiding correlation between 
     # effectiveness and advantage inputs
-    rr_advantage_I1 <- runif(1, min=0.2, max=1)
-    rr_advantage_I2 <- runif(1, min=0.2, max=1)
+    rr_restoration_ins1 <- runif(1, min=0.2, max=1)
+    rr_restoration_ins2 <- runif(1, min=0.2, max=1)
     # now have moved s calculation into setInputOneScenario()
-    # s.RR1_A0 <- rr_advantage_I1 * phi.SS1_A0
-    # s.RR2_0B <- rr_advantage_I2 * phi.SS2_0B
+    # s.RR1_A0 <- rr_restoration_ins1 * phi.SS1_A0
+    # s.RR2_0B <- rr_restoration_ins2 * phi.SS2_0B
     
     ## put the generated values into an input matrix, using defaults for non specified parameters
     inputOneScenario <- setInputOneScenario( max_gen = max_gen,
@@ -137,8 +137,8 @@ sensiAnPaperPart <- function( nScenarios = 10,
                         male_exposure_prop = male_exposure_prop,
                         correct_mix_deploy = correct_mix_deploy,
                         # 14/6/16
-                        rr_advantage_I1 = rr_advantage_I1,
-                        rr_advantage_I2 = rr_advantage_I2
[…]

### 2016-06-27 R/setInputOneScenario.r

@@ -63,8 +63,8 @@
 #' @param male_exposure_prop male exposure as a propoertion of female, default 1 for same, likely <1  
 #' @param correct_mix_deploy proportion of correct deployment of mixtures, 
 #'    if <1 other portion divided between single insecticides
-#' @param rr_advantage_I1 effect of resistance in overcoming insecticide 1 effectiveness in RR
-#' @param rr_advantage_I2 effect of resistance in overcoming insecticide 2 effectiveness in RR    
+#' @param rr_restoration_ins1 effect of resistance in overcoming insecticide 1 effectiveness in RR
+#' @param rr_restoration_ins2 effect of resistance in overcoming insecticide 2 effectiveness in RR    
 #' 
 #' @return named vector
 #' @export
@@ -109,9 +109,9 @@ setInputOneScenario <- function( calibration = 100,
                                  h.RS2_0b	=	0	,
                                  h.RS2_0B	=	1	,
                                  s.RR1_a0	=	0	,
-                                 s.RR1_A0	=	NULL	, #14/6/16 so rr_advantage is used as default
+                                 s.RR1_A0	=	NULL	, #14/6/16 so rr_restoration is used as default
                                  s.RR2_0b	=	0	,
-                                 s.RR2_0B	=	NULL	, #14/6/16 so rr_advantage is used as default
+                                 s.RR2_0B	=	NULL	, #14/6/16 so rr_restoration is used as default
                                  z.RR1_00	=	0	,
                                  z.RR2_00	=	0	,
                                  niche_00	=	1	,
@@ -128,8 +128,8 @@ setInputOneScenario <- function( calibration = 100,
                                  correct_mix_deploy = 1,
                                  exposure = 0.9,
                                  #14/6/16
-                                 rr_advantage_I1 = 1,
-                                 rr_advantage_I2 = 1
+                                 rr_restoration_ins1 = 1,
[…]

### 2016-06-27 inst/shiny/resistmob1/ui.r

@@ -1,7 +1,7 @@
 #resistance/shiny/resistmob1/ui.r
 #andy south 3/6/16
 #a mobile compatible app
-#uses OLD system where selection coefficient rather than rr_advantage used as input
+#uses OLD system where selection coefficient rather than rr_restoration used as input
 
 library(shiny)
 

### 2016-06-27 inst/shiny/resistmob2/server.r

@@ -42,8 +42,8 @@ shinyServer(function(input, output) {
                       # 14/6/16
                       #s.RR1_A0 = input$advantage_A1,
                       #s.RR2_0B = input$advantage_A2,
-                      rr_advantage_I1 = input$advantage_A1,
-                      rr_advantage_I2 = input$advantage_A2,
+                      rr_restoration_ins1 = input$advantage_A1,
+                      rr_restoration_ins2 = input$advantage_A2,
                       
                       addCombinedStrategy = FALSE,
                       strategyLabels = c('seq','','adapt','mix2') )
@@ -60,8 +60,8 @@ shinyServer(function(input, output) {
                       "exposure =",input$exposure_A,",",
                       "phi.SS1_A0 =",input$effectiveness_A1,",",
                       "phi.SS2_0B =",input$effectiveness_A2,",",
-                      "rr_advantage_I1 =",input$advantage_A1,",",
-                      "rr_advantage_I2 =",input$advantage_A2,",",
+                      "rr_restoration_ins1 =",input$advantage_A1,",",
+                      "rr_restoration_ins2 =",input$advantage_A2,",",
                       "addCombinedStrategy = FALSE,", 
                       "strategyLabels = c('seq','','adapt','mix2')",
                       ")\n" )
@@ -92,8 +92,8 @@ shinyServer(function(input, output) {
                       # 14/6/16
                       #s.RR1_A0 = input$advantage_B1,
                       #s.RR2_0B = input$advantage_B2,
-                      rr_advantage_I1 = input$advantage_B1,
-                      rr_advantage_I2 = input$advantage_B2,
+                      rr_restoration_ins1 = input$advantage_B1,
+                      rr_restoration_ins2 = input$advantage_B2,
                       
                       addCombinedStrategy = FALSE,
                       strategyLabels = c('seq','','adapt','mix2') )
@@ -110,8 +110,8 @@ shinyServer(function(input, output) {
             "exposure =",input$exposure_B,",",
             "phi.SS1_A0 =",input$effectiveness_B1,",",
             "phi.SS2_0B =",input$effectiveness_B2,",",
-            "rr_advantage_I1 =",input$advantage_B1,",",
-            "rr_advantage_I2 =",input$advantage_B2,",",
+            "rr_restoration_ins1 =",input$advantage_B1,",",
+            "rr_restoration_ins2 =",input$advantage_B2,",",
             "addCombinedStrategy = FALSE,", 
             "strategyLabels = c('seq','','adapt','mix2')",
             ")\n" )

### 2016-06-27 inst/shiny/resistmob2/ui.r

@@ -1,7 +1,7 @@
 #resistance/shiny/resistmob2/ui.r
 #andy south 15/6/16
 #a mobile compatible app
-#uses NEW system where rr_advantage used as input instead of selection coefficient
+#uses NEW system where rr_restoration used as input instead of selection coefficient
 
 library(shiny)
 

### 2016-06-27 inst/documents/effectOfIndividualInputs.Rmd
rename correct mix & male exposure inputs
shorten effectiveness var names

@@
   #renaming other rownames to make nicer plots
-  rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_insecticide1"
-  rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_insecticide2"
+  rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_ins1"
+  rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_ins2"
 
@@
 
-#now try the same faceting by effectiveness_insecticide1
+#now try the same faceting by effectiveness_ins1
 #doesn't really show much ...
 #so why does effectiveness come out so important in the trees ??  
-  #hist(ggInsOuts$effectiveness_insecticide1)  
+  #hist(ggInsOuts$effectiveness_ins1)  
   #create effect1 bins
-  ggInsOuts$effect1_cat <- cut(ggInsOuts$effectiveness_insecticide1,breaks=c(0.4,0.6,0.8,1), labels=c("effect1 0.4-0.6","effect1 0.6-0.8","effect1 0.8-1"))
+  ggInsOuts$effect1_cat <- cut(ggInsOuts$effectiveness_ins1,breaks=c(0.4,0.6,0.8,1), labels=c("effect1 0.4-0.6","effect1 0.6-0.8","effect1 0.8-1"))
 

### 2016-06-27 inst/documents/keep_old_insecticide.Rmd
shorten effectiveness var names

@@
 #renaming other rownames to make nicer plots
-rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_insecticide1"
-rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_insecticide2"
+rownames(treeInput)[rownames(treeInput)=="phi.SS1_A0"] <- "effectiveness_ins1"
+rownames(treeInput)[rownames(treeInput)=="phi.SS2_0B"] <- "effectiveness_ins2"
 
@@
 curtisInputs <- data.frame("exposure"=0.9,
-                           "effectiveness_insecticide1"=0.73,
-                           "effectiveness_insecticide2"=1,
+                           "effectiveness_ins1"=0.73,
+                           "effectiveness_ins2"=1,
                            "start_freq_allele1"=0.01,

### 2016-06-27 inst/documents/paper1_results_figs_rr.Rmd
shorten effectiveness var names

@@
   curtisInputs <- data.frame("exposure"=0.9,
-                             "effectiveness_insecticide1"=0.73,
-                             "effectiveness_insecticide2"=1,
+                             "effectiveness_ins1"=0.73,
+                             "effectiveness_ins2"=1,
                              "effectiveness_added"=1.73,                             
@@
 
-  ggSubset$effectiveness_added <- ggSubset$effectiveness_insecticide1 + ggSubset$effectiveness_insecticide2
+  ggSubset$effectiveness_added <- ggSubset$effectiveness_ins1 + ggSubset$effectiveness_ins2
   ggSubset$dominance_added <- ggSubset$dominance_allele1 + ggSubset$dominance_allele2
@@
   #names_inputs <- names(ggInsOuts)[1:num_inputs]
-  names_inputs <- c("effectiveness_insecticide1", "effectiveness_insecticide2", "effectiveness_added")
+  names_inputs <- c("effectiveness_ins1", "effectiveness_ins2", "effectiveness_added")
 
@@
 
-  ggSubset$effectiveness_added <- ggSubset$effectiveness_insecticide1 + ggSubset$effectiveness_insecticide2
+  ggSubset$effectiveness_added <- ggSubset$effectiveness_ins1 + ggSubset$effectiveness_ins2
 
@@
     #failed hex plot just gave me counts
-    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.5'))+
+    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_ins1', colour='mixA_better_seq0.5'))+
     #          geom_hex()
@@
     #interesting plot showing where sequence is better is restricted to higher values of exposure and effectiveness 
-    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.5'))+
+    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_ins1', colour='mixA_better_seq0.5'))+
              #geom_hex()
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_better_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 * effectiveness_ins2, colour=mixA_better_seq0.5))+
              #geom_hex()
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mixA_better_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=mixA_better_seq0.5))+
 
@@
     #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mixA0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=seq_better_mixA0.5))+
              facet_grid( ~ seq_better_mixA0.5, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
@@
     #but the previous one may be clearer
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_minus_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 * effectiveness_ins2, colour=mixA_minus_seq0.5))+
              #geom_hex()
@@
     #failed attempt at getting Curtis point in legend
-    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.5, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.5))+
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mixA_minus_seq0.5, size="Curtis"))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mixA_minus_seq0.5))+
 
@@
              #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=3) +
              #failing to add legend for curtis point
@@
              #add text for Curtis point
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
+             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_ins1 + effectiveness_ins2', label='"Curtis Fig.2"'), nudge_x=-0.05, size=3, colour='red')
            )
@@
 
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mixA_minus_seq0.5))+
 
@@
              #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=3) +
              ylab('effectiveness of insecticides added') +
@@
   x1=df_same$exposure 
-  y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
+  y1=df_same$effectiveness_ins1 + df_same$effectiveness_ins2
 
@@
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.5))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2))+ #, colour=abs_mixA_minus_seq0.5))+
 
@@
              #add point for Curtis inputs
-             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             #geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=3) +
              ylab('effectiveness of insecticides added') +
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=mix2_better_seq0.5))+
 
@@
     #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=seq_better_mix20.5))+
              facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mix2_minus_seq0.5))+
 
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.5))+
+    print( ggplot(ggSubset, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mix2_minus_seq0.5))+
 
@@
              #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=3) +
+             geom_point(data = curtisInputs, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=3) +
              ylab('effectiveness of insecticides added') +

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50.Rmd
shorten effectiveness var names

@@
   curtisInputs <- data.frame("exposure"=0.9,
-                             "effectiveness_insecticide1"=0.73,
-                             "effectiveness_insecticide2"=1,
+                             "effectiveness_ins1"=0.73,
+                             "effectiveness_ins2"=1,
                              "start_freq_allele1"=0.01,
@@
   #subset which output plots to show in the main paper
-  # in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
+  # in_cols <- c("start_freq_allele1", "exposure","effectiveness_ins1",
   #             "dominance_allele1","selection_coef_allele1","male_exposure_prop")
-  in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")
+  in_cols <- c("exposure","effectiveness_ins1","dominance_allele1")
 
@@
     #failed attempt at getting Curtis point in legend
-    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_diff, size="Curtis"))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_diff))+
 
@@
              #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='black', shape=3) +
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='black', shape=3) +
              #failing to add legend for curtis point
@@
              #add text for Curtis point
-             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_ins1 + effectiveness_ins2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
            )
@@
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_diff))+
 
@@
   x1=df_20pc$exposure 
-  y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
+  y1=df_20pc$effectiveness_ins1 + df_20pc$effectiveness_ins2
 
@@
 
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_diff))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_diff))+
 
@@
              #showed that they were on both sides
-             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2), colour='red', shape=2) +
+             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=2) +
 
              #add text for Curtis point
-             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_ins1 + effectiveness_ins2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
            )

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
shorten effectiveness var names

@@
   #26%
-  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_insecticide1 > 0.4 & dif_mix2_seq$effectiveness_insecticide2 > 0.4, ]
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_ins1 > 0.4 & dif_mix2_seq$effectiveness_ins2 > 0.4, ]
   #13%
-  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_insecticide1 > 0.4 & dif_mix2_seq$effectiveness_insecticide2 > 0.4 & dif_mix2_seq$exposure > 0.3, ]
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_ins1 > 0.4 & dif_mix2_seq$effectiveness_ins2 > 0.4 & dif_mix2_seq$exposure > 0.3, ]
   #10%
@@
   #8%
-  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_insecticide1 > 0.4 & dif_mix2_seq$effectiveness_insecticide2 > 0.4 & dif_mix2_seq$start_freq_allele1 > 0.001 & dif_mix2_seq$start_freq_allele2 > 0.001 & dif_mix2_seq$exposure > 0.3, ]
+  #dif_mix2_seq2 <- dif_mix2_seq[ dif_mix2_seq$effectiveness_ins1 > 0.4 & dif_mix2_seq$effectiveness_ins2 > 0.4 & dif_mix2_seq$start_freq_allele1 > 0.001 & dif_mix2_seq$start_freq_allele2 > 0.001 & dif_mix2_seq$exposure > 0.3, ]
   #22%
@@
   # create added columns for effectiveness and dominance
-  ggSubset$effectiveness_added <- ggSubset$effectiveness_insecticide1 + ggSubset$effectiveness_insecticide2
+  ggSubset$effectiveness_added <- ggSubset$effectiveness_ins1 + ggSubset$effectiveness_ins2
   ggSubset$dominance_added <- ggSubset$dominance_allele1 + ggSubset$dominance_allele2
@@
   #subset which output plots to show in the main paper
-  # in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
+  # in_cols <- c("start_freq_allele1", "exposure","effectiveness_ins1",
   #             "dominance_allele1","selection_coef_allele1","male_exposure_prop")
-  #in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")  
+  #in_cols <- c("exposure","effectiveness_ins1","dominance_allele1")  
   in_cols <- c("exposure","effectiveness_added","dominance_added")
@@
   ggSubset$l_mod_confusion <- 
-    ifelse ( (ggSubset$effectiveness_insecticide1+ggSubset$effectiveness_insecticide2) > ggSubset$l_mod_eff,
+    ifelse ( (ggSubset$effectiveness_ins1+ggSubset$effectiveness_ins2) > ggSubset$l_mod_eff,
                ifelse(ggSubset$mixA_minus_seq0.5>0,"TP",ifelse(ggSubset$mixA_minus_seq0.5<0,"FP",NA))
@@
   ggSubset$l_mod_confusion <- 
-    ifelse ( (ggSubset$effectiveness_insecticide1+ggSubset$effectiveness_insecticide2) > ggSubset$l_mod_eff,
+    ifelse ( (ggSubset$effectiveness_ins1+ggSubset$effectiveness_ins2) > ggSubset$l_mod_eff,
                ifelse(ggSubset$mixA_not20pc_better_seq0.5==0,"TP",ifelse(ggSubset$mixA_not20pc_better_seq0.5==1,"FP",NA))

### 2016-06-27 inst/documents/sensiAnPaper1All.Rmd
shorten effectiveness var names

@@
   #subsetting now done above before names prettified
-  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_insecticide1','effectiveness_insecticide2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
+  #treePredictors <- c('P_1','P_2','startResistanceRatio','exposure','effectiveness_ins1','effectiveness_ins2','h.RS1_A0','h.RS2_0B','s.RR1_A0','s.RR2_0B')
   #treeInput <- treeInput[ treePredictors, ]

### 2016-06-27 inst/documents/paper1_results_figs_rr.Rmd
convert varnames from old output files

@@
   #rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+  #27/6/16 to convert any varnames in old output files
+  rownames(treeInput)[ "maleExposureProp" == rownames(treeInput) ] <- "male_exposure_prop"
+  rownames(treeInput)[ "correctMixDeployProp" == rownames(treeInput) ] <- "correct_mix_deploy"
+  rownames(treeInput)[ "rr_advantage_I1" == rownames(treeInput) ] <- "rr_restoration_ins1"
+  rownames(treeInput)[ "rr_advantage_I2" == rownames(treeInput) ] <- "rr_restoration_ins2"
 

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
convert varnames from old output files

@@
   #except that exposure will be in a.f_AB, a.f_A0 and a.f_B0 respectively (& a.m* too)
-  #I could just rename one to exposure
-  #BEWARE risk if future changes
   #1/2/16 don't need this now because I've saved exposure from the single original random value
   #rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+  #27/6/16 to convert any varnames in old output files
+  rownames(treeInput)[ "maleExposureProp" == rownames(treeInput) ] <- "male_exposure_prop"
+  rownames(treeInput)[ "correctMixDeployProp" == rownames(treeInput) ] <- "correct_mix_deploy"
+  rownames(treeInput)[ "rr_advantage_I1" == rownames(treeInput) ] <- "rr_restoration_ins1"
+  rownames(treeInput)[ "rr_advantage_I2" == rownames(treeInput) ] <- "rr_restoration_ins2"
 

### 2016-06-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
improve plot

@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=3.5}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=4.5}
 
   library(sensitivity)
+
+  #list to put plots in for cowplot
+  plotlist <- list(2)
 
@@
 
-    cat("A)")
+    #cat("A)")
 
-    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+    plotlist[[1]] <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
              geom_point(shape=1, colour='red') +
-             theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
+             theme(axis.text.x = element_blank()) +
              geom_hline(yintercept = 0, linetype=3) +
-             #theme_bw() +
+             ggtitle(paste("A) Mixture minus sequential")) +
              ylim(-1,1) +
              xlab(NULL)
-          )
 
@@
 
-    cat("B)")
+    #cat("B)")
 
-    print( ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
+    plotlist[[2]] <- ggplot( to_plot, aes_string(x='inputs',y='PRCC') ) + 
              geom_point(shape=1, colour='red') +
@@
              geom_hline(yintercept = 0, linetype=3) +
-             #theme_bw() +
+             ggtitle(paste("B) Adaptive minus sequential")) +
              ylim(-1,1) +
              xlab(NULL)
-          )
+
+  #to put back in some ref gridlines that are removed by cowplot
+  #theme_set(theme_gray())
+  theme_set(theme_bw())
+
+  plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(1,1.6))
 

### 2016-06-27 inst/documents/keep_old_insecticide.Rmd
update for new varnames

@@
 
-## to load previously saved runs
-#   load(file=paste0(outFolder,'listOutMix_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI1_1000.rda'))
-#   load(file=paste0(outFolder,'listOutI2_1000.rda'))
-
 ## trying with the extended experiment _ex100
@@
 
-load(file=paste0(outFolder,'listOutMix_ex2_10000.rda'))
-load(file=paste0(outFolder,'listOutI1_ex2_10000.rda'))
-load(file=paste0(outFolder,'listOutI2_ex2_10000.rda'))
-
-# load(file=paste0(outFolder,'listOutMix_ex2_1000.rda'))
-# load(file=paste0(outFolder,'listOutI1_ex2_1000.rda'))
-# load(file=paste0(outFolder,'listOutI2_ex2_1000.rda'))
+  ## to load previously saved runs  
+  load(file=paste0(outFolder,'listOutMix_rr_10000.rda'))
+  load(file=paste0(outFolder,'listOutI1_rr_10000.rda'))
+  load(file=paste0(outFolder,'listOutI2_rr_10000.rda'))
 
@@
 #rownames(treeInput)[rownames(treeInput)=="a.f_AB"] <- "exposure"
+
+  #27/6/16 to convert any varnames in old output files
+  rownames(treeInput)[ "maleExposureProp" == rownames(treeInput) ] <- "male_exposure_prop"
+  rownames(treeInput)[ "correctMixDeployProp" == rownames(treeInput) ] <- "correct_mix_deploy"
+  rownames(treeInput)[ "rr_advantage_I1" == rownames(treeInput) ] <- "rr_restoration_ins1"
+  rownames(treeInput)[ "rr_advantage_I2" == rownames(treeInput) ] <- "rr_restoration_ins2"
 

### 2016-06-28 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
putting figs side by side

@@
 
-#### (This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
-#### Target : Malaria journal
+(This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
+Target : Malaria journal
+
+#### TO DO
+1. decide whether to use selective advantage or rr restoration coefficient
 
@@
 It uses a model representing two independent loci coding for resistance to two insecticides.
-It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective adavantage of resistance and the dominance of the alleles.
+It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective advantage of resistance and the dominance of the alleles.
 There is uncertainty about the values of these parameters in field situations.
@@
   print( ggplot(df_resist, aes(x=generation, y=resistance, colour=exposure)) + 
+         theme_bw() +
          labs(colour = "exposure") +
@@
   print( ggplot(df_resist, aes(x=generation, y=resistance, colour=h.RS1_A0)) + 
+         theme_bw() +
          labs(colour = "dominance") +
@@
   print( ggplot(df_resist, aes(x=generation, y=resistance, colour=rr_restoration_ins1)) + 
+         theme_bw() +
          labs(colour = "selective advantage") +
@@
 With all control inputs set equal at 0.5 resistance arises slower for sequential than mixture (Fig.
-1).
+1 A).
 Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
 When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 1. All control inputs equal at 0.5")
-```
 
@@
 
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
+All control inputs equal at 0.5, B.
+Effectiveness of insecticide1 increased from 0.5 to 0.9"}
+#fig.cap seems necessary to keep figs side by side
+
+#plot margins
+oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+
+#plot margins
+par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", main='B.' )
+
+par(oldpar)
+
+```
+
 In a mixture the mechanism is that a more effective insecticide provides greater protection to the other.
 For example taking the previous scenario and increasing the effectiveness of one insecticide from 0.5 to 0.9 increases the time-to-resistance for the 2nd insecticide from around 45 to 75 generations (Fig.
-2).
-In the plot you can see that resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to Fig.2.
+1 B).
+In the plot you can see that resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to Fig.
+1A.
 In contrast resistance to insecticide2 increases more slowly initially.
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 2. Effectiveness of insecticide1 increased from 0.5 to 0.9")
-```
-
 When the effectiveness of both insecticides are increased to 0.9 the time-to-resistance is increased still further to around 125 generations (Fig.
-3).
+2).
 In this case the curves for the rise in resistance to each insecticide are identical.
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="Effectiveness of both insecticides 0.9"}
+
+#plot margins
+oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 3. Effectiveness of both insecticides 0.9")
+
 ```
@@
 Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.9 decreases the time-to-resistance to insecticide1 from 40 to 30 generations but leaves the time-to-resistance for the mixture the same (Fig.
-4).
-Setting the dominance for bothe alleles to 0.9 reduces the time-to-resistance for the mixture to 30 generations (Fig.
-5).
+3A).
+Setting the dominance for both alleles to 0.9 reduces the time-to-resistance for the mixture to 30 generations (Fig.
+3B).
 In both of these cases sequential use continues to have a longer time-to-resistance.
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 4. Dominance of insecticide1 0.9")
-```
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A. Dominance of insecticide1 0.9, B. Dominance of both insecticides 0.9"}
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 5. Dominance of both insecticides 0.9")
+#plot margins
+oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+#cat("Dominance of insecticide1 0.9")
+
+#plot margins
+par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+#cat("Dominance of both insecticides 0.9")
+
+par(oldpar)
+
 ```

### 2016-06-28 inst/documents/testfigs.Rmd
for testing fig positioning

--- 2016-06-28 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
@@
 ---
-title: "paper2_resistance_mechanisms_mixtures"
+title: "testfigs"
 output: pdf_document
 #output: word_document
-author: "Andy South, Ian Hastings and Bethany Levick"
+author: "Andy South"
 date: "`r Sys.Date()`"
@@
 
+#testing getting figures side-by-side
+
 ```{r,echo=FALSE,results='hide'}
   library(resistance)
-  library(ggplot2)
+  #library(ggplot2)
 ```
 
-### version 3
-
-# Mechanisms by which insecticide resistance can be influenced by strategies of insectcide use : a simple theoretical exploration.
-
-(This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
-Target : Malaria journal
-
-#### TO DO
-1. decide whether to use selective advantage or rr restoration coefficient
-
-## Abstract
-
-This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence/rotation and mixtures.
-It is aimed at a readership with limited experience of population genetics and resistance.
-It uses a model representing two independent loci coding for resistance to two insecticides.
-It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective advantage of resistance and the dominance of the alleles.
-There is uncertainty about the values of these parameters in field situations.
-We demonstrate how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies.
-We describe an R package and user-interface that allows further exploration of the development of resistance.
-The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.
-
-## Introduction
-
-Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.
-
-Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency.
-We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
-
-The changing frequency of insecticide resistance is a population genetic process that can be influenced by :
-
-Parameter  | Description
-------------------- | -------------------------------------------------------
-1. Effectiveness | proportion of exposed insects with the susceptible allele that are killed by the insecticide
-2. Exposure | proportion of the insect population that are exposed to the insecticide
-3. Dominance | dominance of the resistance allele relative to the susceptible allele
-4. Selection Coefficient  | selective advatange of resistance in the presence of the insecticide
-5. Frequency  | frequency of resistance alleles within the population
-
-We use a 2-locus population genetic model of insecticide resistance (Hastings et al.
-in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
-Hastings et al.
-(in press) describe the technical details of the model and explore large parameter ranges.
-In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate mechanisms that can influence the development of resistance.
-
-## Methods
-
-*accessible summary of the model*
-
-## Results
-
-For single insecticide use, increasing the values of either effectiveness, exposure, dominance, or the selective advantage of RR, all lead to faster resistance (Figs a1-4).
-(In these four plots the value of the inputs not being varied are set at a constant 0.5).
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.8,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=phi.SS1_A0)) + 
-         labs(colour = "effectiveness") +
-         coord_trans(y = "log10") +    
-         geom_point() #show.legend=FALSE) +
-
-       ) # end print
-
-  cat("Fig. a1. Effectiveness effect on single insecticide.")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-
-  exposure <- 0.2
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.5
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.8
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=exposure)) + 
-         theme_bw() +
-         labs(colour = "exposure") +
-         coord_trans(y = "log10") +    
-         geom_point() #show.legend=FALSE) +
-
-       ) # end print
-
-  cat("Fig. a2. Exposure effect on single insecticide.")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-
-  #this sets exposures according to whether insecticide1, 2 or mixture
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.8, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  #plotallele.freq.andy(listOut, locus="1", sex="mf_mean", add_legend=FALSE)
-
-  # now doing a better way
-  df_resist <- get_resistance(locus=1, listOut)
-
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=h.RS1_A0)) + 
-         theme_bw() +
-         labs(colour = "dominance") +
-         coord_trans(y = "log10") + 
-         geom_point() #show.legend=FALSE) +
-
-       ) # end print
-
-  cat("Fig. a3. Dominance effect on single insecticide.")
-
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.2,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.8,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=rr_restoration_ins1)) + 
-         theme_bw() +
-         labs(colour = "selective advantage") +
-         coord_trans(y = "log10") +   
-         geom_point() #show.legend=FALSE) +
-
-       ) # end print
-
-  cat("Fig. a4. Selective advantage of RR effect on single insecticide.")
-
-```
-
-When two insecticides are used the relative performance of a mixture strategy compared to sequential use is determined by the values of these four inputs.
-
-With all control inputs set equal at 0.5 resistance arises slower for sequential than mixture (Fig.
-1 A).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
-When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
-
-**Effectiveness of the insecticides**
-A more effective insecticide prompts increased time-to-resistance when used in a mixture.
-This contrasts with the earlier result showing that when used in isolation a single insecticide prompts a shorter time-to-resistance.
+why are the figs being pushed down the page ....
 
@@
 ```
-
-In a mixture the mechanism is that a more effective insecticide provides greater protection to the other.
-For example taking the previous scenario and increasing the effectiveness of one insecticide from 0.5 to 0.9 increases the time-to-resistance for the 2nd insecticide from around 45 to 75 generations (Fig.
-1 B).
-In the plot you can see that resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to Fig.
-1A.
-In contrast resistance to insecticide2 increases more slowly initially.
-Once insecticide1 reaches it's resistance threshold (around 40 generations), the curve for insecticide2 becomes steeper.
-This can be explained by the mechanism that once insecticide1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide2 and thus stops 'protecting' it from it's own rise in resistance.
-
-When the effectiveness of both insecticides are increased to 0.9 the time-to-resistance is increased still further to around 125 generations (Fig.
-2).
-In this case the curves for the rise in resistance to each insecticide are identical.
-The curves are shallower than previous plots because individuals that are resistant to each insecticide are being killed by the other thus reducing the increase in resistance alleles.
-As the resistance frequencies go above 10% around generation 100 both insecticides provide less protection to the other and the slope of the resistance curve increases.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="Effectiveness of both insecticides 0.9"}
-
-#plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-
-```
-
-**Dominance of resistance alleles**
-Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
-Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.9 decreases the time-to-resistance to insecticide1 from 40 to 30 generations but leaves the time-to-resistance for the mixture the same (Fig.
-3A).
-Setting the dominance for both alleles to 0.9 reduces the time-to-resistance for the mixture to 30 generations (Fig.
-3B).
-In both of these cases sequential use continues to have a longer time-to-resistance.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A. Dominance of insecticide1 0.9, B. Dominance of both insecticides 0.9"}
-
-#plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
-#cat("Dominance of insecticide1 0.9")
-
-#plot margins
-par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
-#cat("Dominance of both insecticides 0.9")
-
-par(oldpar)
-
-```
-
-**Selective advantage of resistance**
-#TODO replace Figs 6,7,8 with AA,BB,CC below and modify this text
-Decreasing the selective advantage of resistance to insecticide1 from 0.5 to 0.1 increases it's time-to-resistance in the mixture from 40 generations to 150, and has a similar effect on it's use in isolation (Fig.
-6).
-However, it has little effect on resistance to insecticide2.
-Because of this sequential use continues to generate a longer time-to-resistance than the mixture.
-Decreasing the selective advantage of both alleles to 0.1 results in a longer time-to-resistance of 200 generations for the mixture (Fig.
-7), however the relative shape of the resistance curves remains the same as in Fig.
-1, with sequential use continuing to generate longer time-to-resistance.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 6. Selective advantage of insecticide1 0.1")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 7. Selective advantage of both insecticides 0.1")
-```
-
-Taking this scenario of lower selective advantage and increasing the effectiveness of both insectides from 0.5 to 0.7 shows again how the change in effectiveness has a greater effect on the mixture than sequential making the mixture preferable (compare Fig.
-8 to Fig.
-7).
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 8. Selective advantage of both insecticides 0.1, effectiveness to 0.7")
-```
-
-#cool contrast between these scenarios modifying effectiveness and rr_restoration in which all defaults set to 0.5 except
-TODO incorporate this into text above
-
-#AA effectiveness I1 to 0.8 causes mixture to be slower, although I1 is faster it protects I2 in mixture, making mixture slower
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. AA. effectiveness to 0.8")
-```
-
-#BB rr advantage I1 to 0.8 sequential remains slower, although I1 is faster as in previous it provides less protection to I2 in the mixture
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. BB rr_restoration I1 0.8")
-```
-
-#CC rr advantage I2 to 0.2 sequential remains slower (just), I2 slower than I1 in mixture similar to AA but receives less protection in mixture
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.2 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. CC. rr_restoration I2 0.2")
-```
-
-**Exposure**
-Increasing exposure to 0.9 reduces time-to-resistance for both mixture and sequential scenarios (Fig.
-9).
-The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
-Setting the effectiveness of both insecicides to 0.7 (while keeping all other inputs at 0.5) results in mixture being slightly favoured (Fig.
-10), then making the increase in exposure from 0.5 to 0.9 results in a switch to sequential being the favoured strategy (Fig.
-11).
-This demonstrates how higher exposures can favour sequential use.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 9. Exposure 0.9")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 10. Effectiveness 0.7, sequence favoured")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 11. Effectiveness 0.7, Exposure 0.9, mixture favoured")
-```
-
-## Discussion
-
-**single insecticide**
-
-Parameter to increase  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | faster | increased deaths of susceptibles
-2. Exposure | faster | increased deaths of susceptibles
-3. Dominance | faster | increased survival of heterozygotes
-4. Selective advantage | faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-
-**mixture**
-
-Parameter to increase  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
-2. Exposure | faster (but not as fast as for single) | increased deaths of individuals resistant to the other insecticide
-3. Dominance | faster | increased survival of resistants
-4. Selective advantage | ?faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-
-**difference between mixture and sequential use**
-
-Parameter to increase  | relative performance mix versus seq | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | mixture | faster resistance for single use and slower resistance in mixture
-2. Exposure | sequence | faster resistance for mixture and sequence but has a greater effect on mixture
-3. Dominance | ? | ?
-4. Selective advantage | ? | ?
-5. Frequency  | ? | ?
-
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
-In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
-For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
-However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
-
-(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
-Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
-
-(@) For a mixture of 2 insecticides with differing effectiveness.
-Resistance to the insecticide with a greater effectiveness increases faster.
-The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level at which point it starts to increase faster.
-
-(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
-In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
-The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
-Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
-This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
-
-## References
-1. GPIRM
-1. Curtis 
-1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
-1. Birget 2015
-
-## Todo
-
-1.
-work out what's going on with exposure ...
-why does increasing exposure have a greater effect on mixture than sequential ?
-Is it because the exposure to one insecticide also effects the development of resistance in the other ?
-1. create some code to enable plotting 2 scenarios side-by-side
-1. re-read GPIRM
-1. can I develop explanatory diagrams to describe the input parameters ?
-1. mention generation time of 1 month
-1. look at effect of sex linkage

### 2016-06-29 inst/documents/testfigs_simpler.Rmd
failed attempt at getting fig.subcap to work

---
title: "This is my minimal example"
output:
  pdf_document:
    fig_caption: yes
  graphics: yes
  keep_tex: yes  
  header-includes:
    -\usepackage{subfig}

#includes:
#  in_header: styles.tex
---

My working plot.

```{r, results = 'asis', echo=FALSE, fig.width=3, fig.height=3, fig.cap="Figure 1: A smiley face because I work."}
plot(-5:5, (-5:5)^2)
```

My failing plot.

```{r, results = 'asis', echo=FALSE, fig.show='hold', fig.width=3, fig.height=3, fig.cap="Figure 2: I do not work :(", fig.subcap = c("Neither do I", "or me!")}
plot(1:10)
plot(rnorm(10))
```

### 2016-06-30 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
tweak figs, move to end, add text

@@
 
-The changing frequency of insecticide resistance is a population genetic process that can be influenced by :
+The changing frequency of insecticide resistance is a population genetic process that can be influenced by the parameters outlined in Table 1.
+
+**Table 1. Parameters inflencing the development of insecticide resistance**
 
@@
 3. Dominance | dominance of the resistance allele relative to the susceptible allele
-4. Selection Coefficient  | selective advatange of resistance in the presence of the insecticide
+4. Resistance restoration  | ability of resistance to restore fitness in the presence of the insecticide
 5. Frequency  | frequency of resistance alleles within the population
@@
 
-For single insecticide use, increasing the values of either effectiveness, exposure, dominance, or the selective advantage of RR, all lead to faster resistance (Figs a1-4).
-(In these four plots the value of the inputs not being varied are set at a constant 0.5).
+For single insecticide use, faster resistance resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 1-3).
+In these plots each input was varied in isolation with the others set at a constant 0.5, or 0.01 for starting frequency of resistance.
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
+When two insecticides are used the relative performance of a mixture strategy compared to sequential use is determined by the values of these inputs.
+To help us understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs indivividually.
+For our base scenario we set the values for the first four inputs in Table 1 to 0.5 and the starting frequencies of resistance to 0.1.
+For this base scenario resistance arises slower for sequential than mixture (Fig 4A).
+This plot, and those that follow, compares mixture and sequence for a single input scenario (where the previous plots showed 3 input scenarios).
+Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
+When used in sequence the curve (dashed) for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
+
+### Insecticide effectiveness
+A more effective insecticide prompts increased time-to-resistance when used in a mixture.
+This contrasts with the earlier result showing that when used in isolation a single insecticide prompts a shorter time-to-resistance.
+
+For example taking the base scenario and increasing the effectiveness of one insecticide from 0.5 to 0.8 increases the time-to-resistance for the mixture and decreases that for the sequence (Fig 4B).
+In the mixture resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to the base scenario (Fig 4A).
+
+When the effectiveness of both insecticides are increased to 0.8 the time-to-resistance is increased still further to around 125 generations (Fig 5).
+In this case the curves for the rise in resistance to each insecticide are identical.
+The curves are shallower than previous plots because individuals that are resistant to each insecticide are being killed by the other thus reducing the increase in resistance alleles.
+As the resistance frequencies go above 10% around generation 100 both insecticides provide less protection to the other and the slope of the resistance curve increases.
+
+### Exposure
+Increasing exposure to 0.8 reduces time-to-resistance for both mixture and sequential scenarios (Fig 6A relative to 4A).
+The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
+Remembering that earlier increasing one effectiveness to 0.8 resulted in mixtures being favoured over sequence (Fig 4B).
+Increasing exposure from 0.5 to 0.8 switches back to sequential being the favoured strategy (Fig 6B).
+This demonstrates how higher exposures can favour sequential use.
+
+### Dominance of resistance alleles
+Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
+Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance to insecticide1 within the mixture but leaves the time-to-resistance for the second insecticide in the mixture the same (Fig 7A).
+Setting the dominance for both alleles to 0.8 reduces the time-to-resistance for both the mixture and the sequence by 20 generations from the base scenario (compare Fig 7B to 4A).
+In both of these cases sequential use continues to have a longer time-to-resistance under the higher dominance.
+
+### Resistance restoration coefficient
+Increasing the resistance restoration coefficient to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance for insecticide1 in the mixture, and has a similar effect on it's use in the sequence (Fig 8A).
+However, it has little effect on resistance to insecticide2.
+Because of this sequential use continues to generate a longer time-to-resistance than the mixture.
+Increasing the resistance restoration coefficient of both alleles to 0.8 results in a shorter time-to-resistance  for the mixture (Fig 8B), however the relative shape of the resistance curves remains similar to the base scenario  (Fig 4A), with sequential use continuing to generate longer time-to-resistance.
+
+### Starting frequencies of resistance
+Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
+For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 9A to Fig 4A).
+Similarly taking the base scenario with effectiveness of the first insecticide increased to 0.8 (Fig 4B) and decreasing starting frequency of that resistance allele did not result in the mixture being favoured (Fig 9B).
+
+## Discussion
+
+Higher values of all of the inputs lead to faster resistance when the insecticide was used singly. This can be explained by different mechanisms (see Table 2 for summary). [write more about this here]
+
+[explain the effectivenes result, referring to Fig 4B versus 4A]
+In contrast, resistance to insecticide2 increases more slowly initially.
+Once insecticide1 reaches it's resistance threshold, the curve for insecticide2 becomes steeper.
+This can be explained by the mechanism that once insecticide1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide2 and thus stops 'protecting' it from it's own rise in resistance.
+
+[Summary tables to refer to from discussion.]
+
+**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
+
+Parameter to increase  | effect on resistance | Mechanism
+------------------ | ------- | ---------------------------------------------------
+1. Effectiveness | faster | increased deaths of susceptibles
+2. Exposure | faster | increased deaths of susceptibles
+3. Dominance | faster | increased survival of heterozygotes
+4. Resistance restoration | faster | increased survival of resistants
+5. Frequency  | faster | less change needed to reach resistance threshold
+
+**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
+
+Parameter to increase  | effect on resistance | Mechanism
+------------------ | --------- | ---------------------------------------------------
+1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
+2.
+Exposure | faster (but less than for single) | increased deaths of individuals susceptible to one insecticide increases selection pressure for that resistance.
+However at the same time selection pressure is reduced by higher deaths of resistant individuals caused by the other insecticide.
+3. Dominance | faster | increased survival of resistants
+4. Resistance restoration | faster | increased survival of resistants
+5. Frequency  | faster | less change needed to reach resistance threshold
+
+**Table 4. Effect of inputs on the difference between mixture and sequential use**
+
+Parameter to increase  | increase favours mix or sequence | Mechanism
+------------------ | ------- | ---------------------------------------------------
+1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
+2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
+3. Dominance | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
+4. Resistance restoration | neither | As for dominance. Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
+5.
+Frequency  | neither | As for dominance and resistance restoration.
+Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
+
+(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
+
+(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
+In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
+For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
+However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
+
+(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
+Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
+
+(@) For a mixture of 2 insecticides with differing effectiveness.
+Resistance to the insecticide with a greater effectiveness increases faster.
+The less effective insecticide is 'protected' by the more effective.
+Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level at which point it starts to increase faster.
+
+(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
+In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
+The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
+Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
+This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
+
+## References
+1. GPIRM
+1. Curtis 
+1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
+1. Birget 2015
+
+## Todo
+
+1. what plots might I want to show for starting frequency ?
+1.
+work out what's going on with exposure ...
+why does increasing exposure have a greater effect on mixture than sequential ?
+Is it because the exposure to one insecticide also effects the development of resistance in the other ?
+1. In discussion and maybe elewhere address potential concern that the relationships shown here may not hold for all of paramter space.
+1. re-read GPIRM
+1. can I develop explanatory diagrams to describe the input parameters ?
+1. mention generation time of 1 month
+1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Effectiveness, B. Exposure"}
 
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=phi.SS1_A0)) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
+         theme_bw() +  
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) +  
          labs(colour = "effectiveness") +
          coord_trans(y = "log10") +    
-         geom_point() #show.legend=FALSE) +
+         geom_line()
 
-       ) # end print
-
-  cat("Fig. a1. Effectiveness effect on single insecticide.")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
+       )
 
@@
 
-  #input <- cbind(input, inputOneScenario)
   input <- cbind(i1,i2,i3)
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=exposure)) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
          theme_bw() +
-         labs(colour = "exposure") +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) +
+         labs(colour = "exposure", y=NULL) +
          coord_trans(y = "log10") +   
-         geom_point() #show.legend=FALSE) +
+         geom_line()
 
-       ) # end print
+       )
 
-  cat("Fig. a2. Exposure effect on single insecticide.")
 ```
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Dominance, B. Resistance restoration"}
 
-  #this sets exposures according to whether insecticide1, 2 or mixture
+  #single insecticide dominance fig
+
   a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
@@
 
-  #plotallele.freq.andy(listOut, locus="1", sex="mf_mean", add_legend=FALSE)
-
-  # now doing a better way
   df_resist <- get_resistance(locus=1, listOut)
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=h.RS1_A0)) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
          theme_bw() +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) + 
          labs(colour = "dominance") +
          coord_trans(y = "log10") + 
-         geom_point() #show.legend=FALSE) +
+         geom_line()
 
-       ) # end print
+       )
 
-  cat("Fig. a3. Dominance effect on single insecticide.")
-
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
+  # single insecticide resistance restoration fig.
 
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=rr_restoration_ins1)) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
          theme_bw() +
-         labs(colour = "selective advantage") +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) +
+         labs(colour = "rr restoration", y=NULL) +
          coord_trans(y = "log10") +   
-         geom_point() #show.legend=FALSE) +
-
-       ) # end print
-
-  cat("Fig. a4. Selective advantage of RR effect on single insecticide.")
+         geom_line()   
+       )
 
@@
 
-When two insecticides are used the relative performance of a mixture strategy compared to sequential use is determined by the values of these four inputs.
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of resistance starting frequency on single insecticide use."}
 
-With all control inputs set equal at 0.5 resistance arises slower for sequential than mixture (Fig.
-1 A).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
-When used in sequence the curve for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
+  #single insecticide frequency fig
 
-**Effectiveness of the insecticides**
-A more effective insecticide prompts increased time-to-resistance when used in a mixture.
-This contrasts with the earlier result showing that when used in isolation a single insecticide prompts a shorter time-to-resistance.
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.001,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5 )
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.1,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5 )
+
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
+         theme_bw() +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) + 
+         labs(colour = "start\nfrequency") +
+         coord_trans(y = "log10") + 
+         geom_line()
+
+       )
+```
 
@@
 All control inputs equal at 0.5, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.9"}
-#fig.cap seems necessary to keep figs side by side
+Effectiveness of insecticide1 increased from 0.5 to 0.8"}
+#effectiveness on mixture
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", main='B.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", main='B.' )
 
@@
 
-In a mixture the mechanism is that a more effective insecticide provides greater protection to the other.
-For example taking the previous scenario and increasing the effectiveness of one insecticide from 0.5 to 0.9 increases the time-to-resistance for the 2nd insecticide from around 45 to 75 generations (Fig.
-1 B).
-In the plot you can see that resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to Fig.
-1A.
-In contrast resistance to insecticide2 increases more slowly initially.
-Once insecticide1 reaches it's resistance threshold (around 40 generations), the curve for insecticide2 becomes steeper.
-This can be explained by the mechanism that once insecticide1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide2 and thus stops 'protecting' it from it's own rise in resistance.
-
-When the effectiveness of both insecticides are increased to 0.9 the time-to-resistance is increased still further to around 125 generations (Fig.
-2).
-In this case the curves for the rise in resistance to each insecticide are identical.
-The curves are shallower than previous plots because individuals that are resistant to each insecticide are being killed by the other thus reducing the increase in resistance alleles.
-As the resistance frequencies go above 10% around generation 100 both insecticides provide less protection to the other and the slope of the resistance curve increases.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="Effectiveness of both insecticides 0.9"}
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="Effectiveness of both insecticides 0.8"}
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.9 , phi.SS2_0B = 0.9 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
 
@@
 
-**Dominance of resistance alleles**
-Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
-Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.9 decreases the time-to-resistance to insecticide1 from 40 to 30 generations but leaves the time-to-resistance for the mixture the same (Fig.
-3A).
-Setting the dominance for both alleles to 0.9 reduces the time-to-resistance for the mixture to 30 generations (Fig.
-3B).
-In both of these cases sequential use continues to have a longer time-to-resistance.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A. Dominance of insecticide1 0.9, B. Dominance of both insecticides 0.9"}
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
+Exposure increased to 0.8, B.
+Exposure and Effectiveness of insecticide1 increased from 0.5 to 0.8"}
+#fig.cap seems necessary to keep figs side by side
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
-#cat("Dominance of insecticide1 0.9")
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+#cat("Fig. Exposure 0.8 benefit of sequence over mixture increases")
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.9 , h.RS2_0B = 0.9 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
-#cat("Dominance of both insecticides 0.9")
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), main='B.' )
+#cat("Fig. Effectiveness 0.8, Exposure 0.8, increased exposure nullifies the effect of increased effectiveness, sequence remains better")
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A. Dominance of insecticide1 0.8, B. Dominance of both insecticides 0.8"}
+
+#plot margins
+oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+#cat("Dominance of insecticide1 0.8")
+
+#plot margins
+par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.8 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+#cat("Dominance of both insecticides 0.8")
 
@@
 
-**Selective advantage of resistance**
-#TODO replace Figs 6,7,8 with AA,BB,CC below and modify this text
-Decreasing the selective advantage of resistance to insecticide1 from 0.5 to 0.1 increases it's time-to-resistance in the mixture from 40 generations to 150, and has a similar effect on it's use in isolation (Fig.
-6).
-However, it has little effect on resistance to insecticide2.
-Because of this sequential use continues to generate a longer time-to-resistance than the mixture.
-Decreasing the selective advantage of both alleles to 0.1 results in a longer time-to-resistance of 200 generations for the mixture (Fig.
-7), however the relative shape of the resistance curves remains the same as in Fig.
-1, with sequential use continuing to generate longer time-to-resistance.
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
+Resistance restoration coefficient allele1 0.8, B.
+Resistance restoration coefficient both alleles 0.8"}
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 6. Selective advantage of insecticide1 0.1")
+#plot margins
+oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+#cat("Fig. rr_restoration I1 0.8")
+
+#plot margins
+par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.8 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+#cat("Fig. CC. rr_restoration I2 0.2")
+
+par(oldpar)
+
 ```
 
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 7. Selective advantage of both insecticides 0.1")
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
+Start frequency allele1 reduced from 0.001, B.
+as A plus effectiveness insecticide1 increased to 0.8"}
+
+#plot margins
+oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#P_1 to 0.001 seq remains better
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+
+#plot margins
+par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#P_1 to 0.001 & effect1 to 0.8 mix remains better
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+
+par(oldpar)
+
 ```
-
-Taking this scenario of lower selective advantage and increasing the effectiveness of both insectides from 0.5 to 0.7 shows again how the change in effectiveness has a greater effect on the mixture than sequential making the mixture preferable (compare Fig.
-8 to Fig.
-7).
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.1 , rr_restoration_ins2 = 0.1 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 8. Selective advantage of both insecticides 0.1, effectiveness to 0.7")
-```
-
-#cool contrast between these scenarios modifying effectiveness and rr_restoration in which all defaults set to 0.5 except
-TODO incorporate this into text above
-
-#AA effectiveness I1 to 0.8 causes mixture to be slower, although I1 is faster it protects I2 in mixture, making mixture slower
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. AA. effectiveness to 0.8")
-```
-
-#BB rr advantage I1 to 0.8 sequential remains slower, although I1 is faster as in previous it provides less protection to I2 in the mixture
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. BB rr_restoration I1 0.8")
-```
-
-#CC rr advantage I2 to 0.2 sequential remains slower (just), I2 slower than I1 in mixture similar to AA but receives less protection in mixture
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.2 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. CC. rr_restoration I2 0.2")
-```
-
-**Exposure**
-Increasing exposure to 0.9 reduces time-to-resistance for both mixture and sequential scenarios (Fig.
-9).
-The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
-Setting the effectiveness of both insecicides to 0.7 (while keeping all other inputs at 0.5) results in mixture being slightly favoured (Fig.
-10), then making the increase in exposure from 0.5 to 0.9 results in a switch to sequential being the favoured strategy (Fig.
-11).
-This demonstrates how higher exposures can favour sequential use.
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 9. Exposure 0.9")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 10. Effectiveness 0.7, sequence favoured")
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=4}
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.9 , phi.SS1_A0 = 0.7 , phi.SS2_0B = 0.7 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
-cat("Fig. 11. Effectiveness 0.7, Exposure 0.9, mixture favoured")
-```
-
-## Discussion
-
-**single insecticide**
-
-Parameter to increase  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | faster | increased deaths of susceptibles
-2. Exposure | faster | increased deaths of susceptibles
-3. Dominance | faster | increased survival of heterozygotes
-4. Selective advantage | faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-
-**mixture**
-
-Parameter to increase  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
-2. Exposure | faster (but not as fast as for single) | increased deaths of individuals resistant to the other insecticide
-3. Dominance | faster | increased survival of resistants
-4. Selective advantage | ?faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-
-**difference between mixture and sequential use**
-
-Parameter to increase  | relative performance mix versus seq | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | mixture | faster resistance for single use and slower resistance in mixture
-2. Exposure | sequence | faster resistance for mixture and sequence but has a greater effect on mixture
-3. Dominance | ? | ?
-4. Selective advantage | ? | ?
-5. Frequency  | ? | ?
-
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
-In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
-For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
-However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
-
-(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
-Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
-
-(@) For a mixture of 2 insecticides with differing effectiveness.
-Resistance to the insecticide with a greater effectiveness increases faster.
-The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level at which point it starts to increase faster.
-
-(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
-In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
-The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
-Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
-This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
-
-## References
-1. GPIRM
-1. Curtis 
-1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
-1. Birget 2015
-
-## Todo
-
-1.
-work out what's going on with exposure ...
-why does increasing exposure have a greater effect on mixture than sequential ?
-Is it because the exposure to one insecticide also effects the development of resistance in the other ?
-1. create some code to enable plotting 2 scenarios side-by-side
-1. re-read GPIRM
-1. can I develop explanatory diagrams to describe the input parameters ?
-1. mention generation time of 1 month
-1. look at effect of sex linkage

### 2016-06-30 inst/shiny/resistmob2/server.r
tweak doc

@@ -1,4 +1,4 @@
-#resistance/shiny/resistmob1/server.r
+#resistance/shiny/resistmob2/server.r
 #andy south 3/6/16
 #unmodified from MixSeqResist1, only changes in UI
 

### 2016-07-04 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
add ref
tweak

@@
 1. Birget 2015
+1. Curtis et al 1993, potential IRM strategies :
+~ moderation : preserve susceptible genes by limiting selection pressure
+~ saturation : high dose so that heterozygous resistants are killed
+~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
 
@@
 
-1. what plots might I want to show for starting frequency ?
 1.

### 2016-07-04 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
fig bigger
correct bug in classification tree figures

@@
 
-## Figures for paper 1 : version 7
+## Figures for paper 1 : version 8
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=4.5}
+```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=5.5}
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5}
+```{r, eval=TRUE, echo=FALSE, warning=TRUE, message=FALSE, fig.height=3.5}
 
@@
   require('plyr')
+
+  #4/7/16 need to remove runs where gens >= 500
+  length(which(resistPointsSeq['gen_cP0.5',] >= 500 | resistPointsMix_A['gen_cP0.5',] >= 500))
+  #2376
+  length(which(resistPointsSeq['gen_cP0.5',] >= 500 ))
+  #2151
+  length(which(resistPointsMix_A['gen_cP0.5',] >= 500))
+  #2306
+
+  #BEWARE1 of subtracting if diff num in each ...
+  #this removes runs that fail either
+
+  failed_run_indices <- (resistPointsSeq['gen_cP0.5',] >= 500 | resistPointsMix_A['gen_cP0.5',] >= 500)
+
+  resistPointsMix_A <- resistPointsMix_A[,!failed_run_indices]
+  resistPointsSeq <- resistPointsSeq[,!failed_run_indices]
+  #BEWARE2 I need to remove these indices from the predictors too
+  treeInput <- treeInput[,!failed_run_indices]
 
@@
   #just doing for a few strategies, sensiAnPpaer1All.Rmd does for a bunch
-  treeResponses <- c( rownames(resistBetterMix_ASeq)[2], #2 is 25%
-                      rownames(resistBetterMix_ASeq20)[2]
-                    )
+  # treeResponses <- c( rownames(resistBetterMix_ASeq)[2], #2 is 25%
+  #                     rownames(resistBetterMix_ASeq20)[2]
+  #                   )
+  #4/7/16 eek above before was a bug showing 25% rather than 50%
+  treeResponses <- c( "betterMix_ASeq_cP0.5","betterMix_ASeq20_cP0.5")
 
@@
     cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
-
     treePruned <- prune(tree, cp=cpOptimal)
@@
     prp(treePruned, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE)
+  }
+
+  #4/7/2016 testing the unpruned trees, repeats previous loop with pruning removed
+  for( treeResponse in treeResponses )
+  {
+    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')
+
+    #to set box colours
+    cols <- ifelse(tree$frame$yval == 1, "red", "green3")
+
+    #add fig label
+    if (treeResponse == treeResponses[1]) main <- "A)"
+    else  main <- "B)"
+
+    #adds in no. scenarios correctly and incorrectly classified
+    prp(tree, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE)
   }

### 2016-09-07 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
add graphics & to discussion

@@
 ------------------- | -------------------------------------------------------
-1. Effectiveness | proportion of exposed insects with the susceptible allele that are killed by the insecticide
+1. Effectiveness | proportion of exposed insects with both susceptible alleles that are killed by the insecticide
 2. Exposure | proportion of the insect population that are exposed to the insecticide
@@
 
-We use a 2-locus population genetic model of insecticide resistance (Hastings et al.
+We use a 2-locus population genetic model of insecticide resistance (Levick et al.
 in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
-Hastings et al.
+Levick et al.
 (in press) describe the technical details of the model and explore large parameter ranges.
@@
 
-*accessible summary of the model*
+*start with the framework, then after describe specifics about how many loci etc*
+*try hard to start simple, explain all terms and make few assumptions*
+
+The simulation represents a population of randomly mixing individuals using standard population genetic approaches so that it does not need to follow every individual.
+One locus is represented for each insecticide.
+There are two potential alleles per locus.
+Each allele confers either resistance or susceptibility to the insecticide.
+The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model.
+These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
+
+The simulation proceeds through generations.
+In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
+This acts to make the fitter alleles more common over time.
+Sexual reproduction is represented first by creating gametes by splitting the adult genomes in two, and secound by forming offspring genotypes by adding gametes with recombination.
+
+So far this description has just considered a single insecticide and associated resistance allele (Fig 1A).
+In the simulation a secound insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
+
+In addition the model structure allows for 3 different levels of exposure to be set for each insecticide (Fig 2).
+So in the two insecticide case that gives the potential for nine (3*3) different exposure niches.
+In the runs shown here we have just used two levels of exposure for each insecticide (effectively no or yes with the value for yes set from the exposure input in Table 1).
+The model structure allows males and females to experience different exposures.
 
@@
 
-For single insecticide use, faster resistance resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 1-3).
-In these plots each input was varied in isolation with the others set at a constant 0.5, or 0.01 for starting frequency of resistance.
+For single insecticide use, faster resistance resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 3-5).
+In the plots each input was varied in isolation with the remaining set at a constant 0.5 (except for starting frequency of resistance which was set to 0.01).
 
@@
 For our base scenario we set the values for the first four inputs in Table 1 to 0.5 and the starting frequencies of resistance to 0.1.
-For this base scenario resistance arises slower for sequential than mixture (Fig 4A).
+For this base scenario resistance arises slower for sequential than mixture (Fig 6A).
 This plot, and those that follow, compares mixture and sequence for a single input scenario (where the previous plots showed 3 input scenarios).
@@
 
-For example taking the base scenario and increasing the effectiveness of one insecticide from 0.5 to 0.8 increases the time-to-resistance for the mixture and decreases that for the sequence (Fig 4B).
-In the mixture resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to the base scenario (Fig 4A).
+For example taking the base scenario and increasing the effectiveness of one insecticide from 0.5 to 0.8 increases the time-to-resistance for the mixture and decreases that for the sequence (Fig 6B).
+In the mixture resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to the base scenario (Fig 6A).
 
-When the effectiveness of both insecticides are increased to 0.8 the time-to-resistance is increased still further to around 125 generations (Fig 5).
+When the effectiveness of both insecticides are increased to 0.8 the time-to-resistance is increased still further to around 125 generations (Fig 7).
 In this case the curves for the rise in resistance to each insecticide are identical.
@@
 ### Exposure
-Increasing exposure to 0.8 reduces time-to-resistance for both mixture and sequential scenarios (Fig 6A relative to 4A).
+Increasing exposure to 0.8 reduces time-to-resistance for both mixture and sequential scenarios (Fig 8A relative to 4A).
 The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
-Remembering that earlier increasing one effectiveness to 0.8 resulted in mixtures being favoured over sequence (Fig 4B).
-Increasing exposure from 0.5 to 0.8 switches back to sequential being the favoured strategy (Fig 6B).
+Remembering that earlier increasing one effectiveness to 0.8 resulted in mixtures being favoured over sequence (Fig 6B).
+Increasing exposure from 0.5 to 0.8 switches back to sequential being the favoured strategy (Fig 8B).
 This demonstrates how higher exposures can favour sequential use.
@@
 Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
-Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance to insecticide1 within the mixture but leaves the time-to-resistance for the second insecticide in the mixture the same (Fig 7A).
-Setting the dominance for both alleles to 0.8 reduces the time-to-resistance for both the mixture and the sequence by 20 generations from the base scenario (compare Fig 7B to 4A).
+Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance to insecticide1 within the mixture but leaves the time-to-resistance for the second insecticide in the mixture the same (Fig 9A).
+Setting the dominance for both alleles to 0.8 reduces the time-to-resistance for both the mixture and the sequence by 20 generations from the base scenario (compare Fig 9B to 6A).
 In both of these cases sequential use continues to have a longer time-to-resistance under the higher dominance.
@@
 ### Resistance restoration coefficient
-Increasing the resistance restoration coefficient to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance for insecticide1 in the mixture, and has a similar effect on it's use in the sequence (Fig 8A).
+Increasing the resistance restoration coefficient to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance for insecticide1 in the mixture, and has a similar effect on it's use in the sequence (Fig 10A).
 However, it has little effect on resistance to insecticide2.
 Because of this sequential use continues to generate a longer time-to-resistance than the mixture.
-Increasing the resistance restoration coefficient of both alleles to 0.8 results in a shorter time-to-resistance  for the mixture (Fig 8B), however the relative shape of the resistance curves remains similar to the base scenario  (Fig 4A), with sequential use continuing to generate longer time-to-resistance.
+Increasing the resistance restoration coefficient of both alleles to 0.8 results in a shorter time-to-resistance  for the mixture (Fig 10B), however the relative shape of the resistance curves remains similar to the base scenario  (Fig 6A), with sequential use continuing to generate longer time-to-resistance.
 
@@
 Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 9A to Fig 4A).
-Similarly taking the base scenario with effectiveness of the first insecticide increased to 0.8 (Fig 4B) and decreasing starting frequency of that resistance allele did not result in the mixture being favoured (Fig 9B).
+For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 11A to Fig 6A).
+Similarly taking the base scenario with effectiveness of the first insecticide increased to 0.8 (Fig 6B) and decreasing starting frequency of that resistance allele did not result in the mixture being favoured (Fig 11B).
 
@@
 
-Higher values of all of the inputs lead to faster resistance when the insecticide was used singly. This can be explained by different mechanisms (see Table 2 for summary). [write more about this here]
+The response of resistance to insecticide use in the model can be explained by a number of mechanisms that can help us develop a more robust understanding of how we would expect resistance to respond in the field.
+
+When single insecticide use was represented (Figs 3-5) higher values of all of the inputs led to faster resistance. 
+[write more about single insecticide responses here]
+
+The response of resistance to changing exposure to, and effectiveness of, the insecticide was identical (compare Fig 3A to 3B).
+This makes sense given, for example, that exposing half of a vector population to an insecticide which kills a quarter of the exposed would be expected to kill the same number as exposing quarter of the population to an insecticide that kills a half of the exposed.
+This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance.
+In both cases the increased deaths of susceptible vectors cause a higher selection pressure that can explain the faster increase.
+
+These mechanisms explaining the response of resistance to single insecticide use are summarised in Table 2.
+
+In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
 
@@
 1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
+
+\pagebreak
+
+## Figures
+
+![Building up the model structure.](mod_graphic_a_to_c.png)
+
+![Exposure to insecticides in the model.](mod_graphic_d.png)
 

### 2016-09-12 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-(This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
 Target : Malaria journal
 
-#### TO DO
-1. decide whether to use selective advantage or rr restoration coefficient
+THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. All comments and & suggestions welcome.
+
+(This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
+
+## ANDY TO DO
+1.
+work out what's going on with exposure ...
+why does increasing exposure have a greater effect on mixture than sequential ?
+Is it because the exposure to one insecticide also effects the development of resistance in the other ?
+1. check & add to discussion text
+1. add to intro about results in levick et al. 
+1. explain resistance-restoration and its relation to selection coefficient
 
@@
 
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches so that it does not need to follow every individual.
+The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
 One locus is represented for each insecticide.
@@
 In the simulation a secound insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
+Thus the simulation can represent populations exposed to two insecticides together in a mixture.
 
@@
 In the runs shown here we have just used two levels of exposure for each insecticide (effectively no or yes with the value for yes set from the exposure input in Table 1).
-The model structure allows males and females to experience different exposures.
+The model structure allows males and females to experience different exposures, we explore that in Levick et al.
+(in press) but the runs presented here use equal exposures.
+
+Here we principally look at model outputs of the change in allele frequencies over time (generations). We investigate three insecticide use strategies.
+1. single insecticide
+1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches 50%
+1. two insecticides used in a mixture (concurrently)
 
@@
 
+### Single insecticide
 For single insecticide use, faster resistance resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 3-5).
@@
 
-When two insecticides are used the relative performance of a mixture strategy compared to sequential use is determined by the values of these inputs.
-To help us understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs indivividually.
+### Two insecticides
+When two insecticides are used the relative performance of a mixture strategy compared to sequential use changes dependent on input values.
+To help us understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs individually.
 For our base scenario we set the values for the first four inputs in Table 1 to 0.5 and the starting frequencies of resistance to 0.1.
@@
 This plot, and those that follow, compares mixture and sequence for a single input scenario (where the previous plots showed 3 input scenarios).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time.
-When used in sequence the curve (dashed) for each insecticide individually is steeper but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
+Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
+When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
 
-### Insecticide effectiveness
+#### results: Insecticide effectiveness
 A more effective insecticide prompts increased time-to-resistance when used in a mixture.
-This contrasts with the earlier result showing that when used in isolation a single insecticide prompts a shorter time-to-resistance.
+This is opposite to what happens when used in sequence and in isolation, both prompting a shorter times-to-resistance.
 
-For example taking the base scenario and increasing the effectiveness of one insecticide from 0.5 to 0.8 increases the time-to-resistance for the mixture and decreases that for the sequence (Fig 6B).
-In the mixture resistance to insecticide1 (with the increased effectiveness) rises at a similar speed to the base scenario (Fig 6A).
+For example, taking the base scenario and increasing the effectiveness of one insecticide from 0.5 to 0.8 increases the time-to-resistance for the mixture (from 90 to 115 generations) and decreases that for the sequence (from 120 to 90 generations) (Fig 6).
+Thus at low effectiveness a sequential strategy is favoured while at higher effectiveness a mixture is favoured.
+
+*? it would be nice to add grey lines to fig6 & others for the 'other' scenario to facilitate comparison, but I'd need to do some tricky rewriting of Beths code.*
 
 When the effectiveness of both insecticides are increased to 0.8 the time-to-resistance is increased still further to around 125 generations (Fig 7).
-In this case the curves for the rise in resistance to each insecticide are identical.
-The curves are shallower than previous plots because individuals that are resistant to each insecticide are being killed by the other thus reducing the increase in resistance alleles.
-As the resistance frequencies go above 10% around generation 100 both insecticides provide less protection to the other and the slope of the resistance curve increases.
 
-### Exposure
+#### results: Exposure
 Increasing exposure to 0.8 reduces time-to-resistance for both mixture and sequential scenarios (Fig 8A relative to 4A).
@@
 
-### Dominance of resistance alleles
+#### results: Dominance of resistance alleles
 Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
@@
 
-### Resistance restoration coefficient
+#### results: Resistance restoration coefficient
 Increasing the resistance restoration coefficient to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance for insecticide1 in the mixture, and has a similar effect on it's use in the sequence (Fig 10A).
@@
 
-### Starting frequencies of resistance
+#### results: Starting frequencies of resistance
 Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
@@
 
-When single insecticide use was represented (Figs 3-5) higher values of all of the inputs led to faster resistance. 
-[write more about single insecticide responses here]
+When single insecticide use was represented (Figs 3-5) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
+Higher values of the five inputs invetsigated all led to more rapid increases in resistance.
 
-The response of resistance to changing exposure to, and effectiveness of, the insecticide was identical (compare Fig 3A to 3B).
-This makes sense given, for example, that exposing half of a vector population to an insecticide which kills a quarter of the exposed would be expected to kill the same number as exposing quarter of the population to an insecticide that kills a half of the exposed.
+The response of resistance to changing exposure and effectiveness was identical (compare Fig 3A to 3B).
+This makes sense as vector kill is effectively a product of exposure times effectiveness.
+For example exposing 50% of a  population to an insecticide which is 75% effective would be expecetd to have the same result as exposing 75% of the population to one which is 50% effective.
 This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance.
@@
 
-These mechanisms explaining the response of resistance to single insecticide use are summarised in Table 2.
+*can we explain why at low dominance 0.2 the rate of increase in resistance is slower at the start ?*
+
+The pattern of more rapid increase in resistance at higher dominance levels (Fig 4A) can be explained by increased survival of heterozygotes in the presence of the insecticide.
+Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival.
+Thus selection pressure for the resistance allele will be increased because it confers more of an advantage when only present on one chromosome.
+
+The faster development of resistance under higher resistance-restoration (Fig 4B) can be explained by it's effect on the survival of the resistant genotypes.
+Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.
+
+The effect of the starting frequency of resistance (Fig 5) is the most different from the other inputs and perhaps the easiest to explain.
+Simply when starting from a higher frequency of resistance there is a smaller change to make to reach the higher resistance thresholds.
+
+These mechanisms explaining the response of resistance to single insecticide use under different input values are summarised in Table 2 and Fig 12.
+Exposure and effectiveness act by decreasing the survival of all genotypes, dominance acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
+
+#### discussion: Two insecticides
 
 In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
+Looking at our base scenario (Fig 6A)
 
-[explain the effectivenes result, referring to Fig 4B versus 4A]
-In contrast, resistance to insecticide2 increases more slowly initially.
-Once insecticide1 reaches it's resistance threshold, the curve for insecticide2 becomes steeper.
-This can be explained by the mechanism that once insecticide1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide2 and thus stops 'protecting' it from it's own rise in resistance.
+#### discussion: Insecticide effectiveness
+
+The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (Fig 6B).
+In contrast in both the sequential (Fig 6B) and single insecticide (Fig 3A) experiments resistance was speeded up by increasing effectiveness.
+This can be explained by the effect that in a mixture increased effectiveness of one insecticide increases killing of vectors that are resistant to the other and thus decreases selection pressure.
+
+This effect can be seen by comparing Fig 6B, where the effectiveness of insecticide 1 has been increased, to 6A.
+In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
+However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
+Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
+This can be explained by the mechanism that once insecticide 1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide 2 and thus stops 'protecting' it from it's own rise in resistance.
+A similar effect is visible when the effectiveness of both insecticides are increased (Fig 7).
+The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 6A) because more individuals resistant to each insecticide are being killed by the other insecticide.
+In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 6B.
+
+Fig 6B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal resistance will be expected to increase faster to the more effective insecticide.
+The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
+
+#### discussion: Exposure
+
+*may want to copy or move some text from summary points below into discussion here*
+
+### Discussion summary points
+
+(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
+
+(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
+In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
+For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
+However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
+
+(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
+Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
+
+(@) For a mixture of 2 insecticides with differing effectiveness.
+Resistance to the insecticide with a greater effectiveness increases faster.
+The less effective insecticide is 'protected' by the more effective.
+Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
+
+(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
+In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
+The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
+Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
+This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
 
@@
 
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
-In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
-For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
-However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
-
-(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
-Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
-
-(@) For a mixture of 2 insecticides with differing effectiveness.
-Resistance to the insecticide with a greater effectiveness increases faster.
-The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level at which point it starts to increase faster.
-
-(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
-In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
-The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
-Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
-This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
-
 ## References
@@
 1. Birget 2015
-1. Curtis et al 1993, potential IRM strategies :
+1. Curtis et al 1993, 
+1. Georghiou 1983 Management of resistance in Arthropods.
+potential IRM strategies :
 ~ moderation : preserve susceptible genes by limiting selection pressure
@@
 
-1.
-work out what's going on with exposure ...
-why does increasing exposure have a greater effect on mixture than sequential ?
-Is it because the exposure to one insecticide also effects the development of resistance in the other ?
 1. In discussion and maybe elewhere address potential concern that the relationships shown here may not hold for all of paramter space.
-1. re-read GPIRM
 1. can I develop explanatory diagrams to describe the input parameters ?
@@
 1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
+1. add to discussion about relevance of these mechanisms in the field
+1.
+e.g.
+exposure can be reduced if either insecticide has a repellant effect.
+(Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
 
@@
 
-![Building up the model structure.](mod_graphic_a_to_c.png)
+![Building up the model structure.](paper2_mod_graphic_a_to_c.png)
 
-![Exposure to insecticides in the model.](mod_graphic_d.png)
+![Exposure to insecticides in the model.](paper2_mod_graphic_d.png)
 
@@
 ```
+
+![Summarising mechanisms of single insecticide effects.](paper2_input_effects_single.png)

### 2016-09-12 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
added pruning test
relabel
minor legend tweak

@@
 
+  #renaming levels just for fig labels
+  levels(ggSubset$strategy) <- c("Mix both","Adaptive both","Sequential both","Mix either")
+
   #for(i in names_inputs)
@@
 
-             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\ntime to resistance\n >20% longer','1'='adaptive mix\ntime to resistance\n >20% longer'))) +
+             facet_grid( ~ mixA_not20pc_better_seq0.5, labeller=as_labeller( c('0'='adaptive mix\ntime to resistance\n >=20% longer','1'='adaptive mix\ntime to resistance\n <=20% longer'))) +
 
@@
 
+    #test tree pruning, seemed to suggest we are pruning more than needed
+    # plotcp(tree)  
+    # 
+    # cat("optimal cp used in pruning=",cpOptimal,"\n")
+
     #to set box colours

### 2016-09-14 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-### version 3
+### version 5
 
-# Mechanisms by which insecticide resistance can be influenced by strategies of insectcide use : a simple theoretical exploration.
+# Mechanisms by which insecticide resistance can be influenced by strategies of insecticide use : exploring by simulation.
 
@@
 ## ANDY TO DO
-1.
-work out what's going on with exposure ...
-why does increasing exposure have a greater effect on mixture than sequential ?
-Is it because the exposure to one insecticide also effects the development of resistance in the other ?
-1. check & add to discussion text
-1. add to intro about results in levick et al. 
+1. work out what's going on with exposure ... why does increasing exposure have a greater effect on mixture than sequential ? 
+1. add to intro including about about results in levick et al. 
 1. explain resistance-restoration and its relation to selection coefficient
+1. expand discussion to include relevance to real world
 
@@
 
-This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence/rotation and mixtures.
+This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence and mixtures.
 It is aimed at a readership with limited experience of population genetics and resistance.
 It uses a model representing two independent loci coding for resistance to two insecticides.
-It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, the selective advantage of resistance and the dominance of the alleles.
-There is uncertainty about the values of these parameters in field situations.
-We demonstrate how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies.
+It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, and the selective advantage of resistance.
+We use the model to develop a mechanistic understanding of how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies.
+We expect this mechanistic understanding to be more robust to uncertainty in parameter values in the field.
 We describe an R package and user-interface that allows further exploration of the development of resistance.
@@
 
+We use a 2-locus population genetic model of insecticide resistance (Levick et al.
+in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
+Levick et al.
+(in press) describe the technical details of the model and explore large parameter ranges.
+We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance.
+This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide use strategies.
+
 **Table 1. Parameters inflencing the development of insecticide resistance**
@@
 
-We use a 2-locus population genetic model of insecticide resistance (Levick et al.
-in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
-Levick et al.
-(in press) describe the technical details of the model and explore large parameter ranges.
-In contrast we provide an accessible summary of the model and use selected parameter values to demonstrate mechanisms that can influence the development of resistance.
-
 ## Methods
-
-*start with the framework, then after describe specifics about how many loci etc*
-*try hard to start simple, explain all terms and make few assumptions*
 
@@
 This acts to make the fitter alleles more common over time.
+Thus a summary of the selection stage is :
+
+> selection coefficient = resistance restoration * effectiveness  
+>
+> Locus fitness : 	
+> SS = 1-effectiveness  
+> RS = 1-effectiveness + (dominance*selection coeff.)  
+> RR = 1-effectiveness + selection coeff.  
+> 
+> Allele fitness :   
+> SS1SS2 = SS * SS  
+> SS1RS2 = SS * RS  
+> … and for other combinations  
+> RR1RR2 = RR * RR  
+> 
+> Individual fitness = Allele fitness * exposure to insecticide  
+> 
+> In each generation :   
+> Allele frequency[t+1] = Allele frequency[t] * Individual fitness
+
 Sexual reproduction is represented first by creating gametes by splitting the adult genomes in two, and secound by forming offspring genotypes by adding gametes with recombination.
@@
 Here we principally look at model outputs of the change in allele frequencies over time (generations). We investigate three insecticide use strategies.
+
 1. single insecticide
@@
 
-*? it would be nice to add grey lines to fig6 & others for the 'other' scenario to facilitate comparison, but I'd need to do some tricky rewriting of Beths code.*
-
 When the effectiveness of both insecticides are increased to 0.8 the time-to-resistance is increased still further to around 125 generations (Fig 7).
@@
 #### results: Exposure
-Increasing exposure to 0.8 reduces time-to-resistance for both mixture and sequential scenarios (Fig 8A relative to 4A).
+Increasing exposure to 0.8 from the base scenario reduces time-to-resistance for both mixture and sequential scenarios (Fig 8A relative to 6A).
 The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
-Remembering that earlier increasing one effectiveness to 0.8 resulted in mixtures being favoured over sequence (Fig 6B).
-Increasing exposure from 0.5 to 0.8 switches back to sequential being the favoured strategy (Fig 8B).
-This demonstrates how higher exposures can favour sequential use.
+
+Similarly increasing exposure reduces time-to-resistance more for the mixture relative to sequence in the scenario in which one insecticide effectiveness has been increased (compare Fig 8B to 6B).
+The result is that the advantage of mixtures over sequence is removed by the increase in exposure.
 
@@
 When single insecticide use was represented (Figs 3-5) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
-Higher values of the five inputs invetsigated all led to more rapid increases in resistance.
+Higher values of the five inputs investigated all led to more rapid increases in resistance.
 
-The response of resistance to changing exposure and effectiveness was identical (compare Fig 3A to 3B).
+For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
 This makes sense as vector kill is effectively a product of exposure times effectiveness.
@@
 In both cases the increased deaths of susceptible vectors cause a higher selection pressure that can explain the faster increase.
-
-*can we explain why at low dominance 0.2 the rate of increase in resistance is slower at the start ?*
 
@@
 In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
-Looking at our base scenario (Fig 6A)
+Looking at our base scenario (Fig 6A) the the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
+The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
+Thus each insecticide reduces the selection pressure for the increase in resistance to the other and could be said to 'protect' the other.
+However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
+As shown in our base scenario, resistance can arise more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
 
@@
 
-*may want to copy or move some text from summary points below into discussion here*
+**Ian, Matt this is still the bit I can't quite explain. Any suggestions ?**
+
+The biggest difference I can't explain is why that in mixtures, increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 in Fig 6A to 40 in Fig 8A) and yet increasing the effectiveness of both insecticides by the same amount results in a slight increase in time-to-resistance (from 90 in Fig 6A to 100 in Fig 7).
+This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (compare Figs 7 & 8A).
+
+either
+
+1. Why doesn't increased exposure provide 'protection' to the other insecticides in a mixture ?
+1. Why doesn't increased effectiveness lead to a greater selection for resistance in each insecticide ?
+1. Why do exposure & effectivness have the same effect on a sequence but different effects on a mixture ?
+
+*some previous guesswork I'm now not sure about: Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
+In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
+The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
+Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
+This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.*
+
+Mechanisms for the response of resistance in mixtures are summarised in Table 3. Table 4 summarises mechanisms explaining the difference between mixture and sequential strategies.
 
@@
 (@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
-In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
-For exposure we have kept values the same for each insecticide the same (to represent a mixture in which the 2 insecticides are applied together).
-However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
 
@@
 
-(@) Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
-In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
-The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
-Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
-This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.
+(@) exposure ...
+
+(@) ?
+In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
+In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
+For exposure we kept values the same for each insecticide (to represent a mixture in which the 2 insecticides are applied together).
+However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
 
@@
 e.g.
-exposure can be reduced if either insecticide has a repellant effect.
+exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
 (Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)

### 2016-09-29 R/runModel2.r
rename k to gen_num

@@ -184,26 +184,26 @@ runModel2 <- function(input,
            
     ##################
     ## generation loop
-    for (k in 1:max_gen){
+    for (gen_num in 1:max_gen){
       
       # In calibration 1011, selection relaxed for a set time
-      if( calibration == 1011 & scen_num==2 ) Windiv <- relaxSelection(Windiv, k)      
+      if( calibration == 1011 & scen_num==2 ) Windiv <- relaxSelection(Windiv, gen_num)      
       
       # save record of genotype proportions each generation
-      genotype[k,1] <- k	
-      #genotype[k,2:11] <- fgenotypes['m',]
+      genotype[gen_num,1] <- gen_num	
+      #genotype[gen_num,2:11] <- fgenotypes['m',]
       # 27/9/16 changing this to output mean of m&f
-      genotype[k,2:11] <- sum( 0.5*(fgenotypes['f',] + fgenotypes['m',]))
+      genotype[gen_num,2:11] <- sum( 0.5*(fgenotypes['f',] + fgenotypes['m',]))
       
       # saving results 
-      results[k,1] <- k #generation number
-      # 
-      results <- resistance_freq_count( fgenotypes=fgenotypes, gen_num=k, results=results )
+      results[gen_num,1] <- gen_num #generation number
+      
+      results <- resistance_freq_count( fgenotypes=fgenotypes, gen_num=gen_num, results=results )
       
       # previous sequential insecticide code that was here now done post-processing
       
       ## linkage calculations
-      results <- linkage_calc( fgenotypes=fgenotypes, recomb_rate=recomb_rate, gen_num=k, results=results )
+      results <- linkage_calc( fgenotypes=fgenotypes, recomb_rate=recomb_rate, gen_num=gen_num, results=results )
 
       ## selection
       fgenotypes_s <- selection( fgenotypes=fgenotypes, Windiv=Windiv, calibration=calibration)

### 2016-10-04 inst/shiny/simpfitvis/server.r
single locus & genotype fitness for one scenario

@@ -22,9 +22,19 @@ shinyServer(function(input, output, session) {
       #isolate reactivity of other objects
     #  isolate({
 
-        plotsimpfit( effectiveness1 = input$effectiveness_A1,
-                     dominance = input$dominance_A1,
-                     rr_restoration1 = input$advantage_A1 )
+        # plotsimpfit( effectiveness1 = input$effectiveness_A1,
+        #              dominance = input$dominance_A1,
+        #              rr_restoration1 = input$advantage_A1 )
+    
+        #trying new way
+        fitnessSingleLocus( eff1 = input$effectiveness_A1,
+                            eff2 = input$effectiveness_A2,
+                            dom1 = input$dominance_A1,
+                            dom2 = input$dominance_A2,
+                            rr_1 = input$advantage_A1,
+                            rr_2 = input$advantage_A2,
+                            plot = TRUE)
+    
 
       #}) #end isolate
     #} #end if ( input$aButtonRun > 0 )
@@ -39,9 +49,23 @@ shinyServer(function(input, output, session) {
     #isolate reactivity of other objects
     #  isolate({
     
-    plotsimpfit( effectiveness1 = input$effectiveness_B1,
-                 dominance = input$dominance_B1,
-                 rr_restoration1 = input$advantage_B1 )
+    # plotsimpfit( eff1 = input$effectiveness_A1,
+    #              eff2 = input$effectiveness_A2,
+    #              dom1 = input$dominance_A1,
+    #              dom2 = input$dominance_A2,
+    #              rr_1 = input$advantage_A1,
+    #              rr_2 = input$advantage_A2 )
+    
+    #trying new way
+    fitnessGenotype( eff1 = input$effectiveness_A1,
+                        eff2 = input$effectiveness_A2,
+                        dom1 = input$dominance_A1,
+                        dom2 = input$dominance_A2,
+                        rr_1 = input$advantage_A1,
+                        rr_2 = input$advantage_A2,
+                        exposure = input$exposure_A,
+                        plot = TRUE)
+    
     
     #}) #end isolate
     #} #end if ( input$aButtonRun > 0 )

### 2016-10-04 inst/shiny/simpfitvis/ui.r

@@ -55,15 +55,15 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
                      #column(1)    
                    ), #end fixed row
                    
-                   fluidRow(
-                     #column(4, h5("takes a few seconds :")),
-                     column(2, p()),
-                     column(2, actionButton('aButtonRunA', 'Run Scenario A')),
-                     column(1, p()),   
-                     column(3, h6("(takes a few seconds)")),
-                     column(2, actionButton('aButtonRunB', 'Run Scenario B')),
-                     column(2, p())
-                   ), #end fixed row
+                   # fluidRow(
+                   #   #column(4, h5("takes a few seconds :")),
+                   #   column(2, p()),
+                   #   column(2, actionButton('aButtonRunA', 'Run Scenario A')),
+                   #   column(1, p()),   
+                   #   column(3, h6("(takes a few seconds)")),
+                   #   column(2, actionButton('aButtonRunB', 'Run Scenario B')),
+                   #   column(2, p())
+                   # ), #end fixed row
                    
                    
                    hr(),
@@ -77,38 +77,38 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
                             h2("B1"),
                             h2("B2")
                      ),
-                     column(2,
-                            #h5("Starting frequency of resistance"),
-                            h5("Start Freq."),
-                            #numericInput("P_1", "locus1: 0.01", 0.01, min = 0.0001, max = 0.1, step = 0.001),
-                            #numericInput("P_2", "locus2: 0.01", 0.01, min = 0.0001, max = 0.1, step = 0.001)
-                            sliderInput("frequency_A1", NULL, val=0.01, min = 0.0001, max = 0.1, step = 0.001, ticks=FALSE),
-                            sliderInput("frequency_A2", NULL, val=0.01, min = 0.0001, max = 0.1, step = 0.001, ticks=FALSE),
-                            hr(),
-                            sliderInput("frequency_B1", NULL, val=0.01, min = 0.0001, max = 0.1, step = 0.001, ticks=FALSE),
-                            sliderInput("frequency_B2", NULL, val=0.01, min = 0.0001, max = 0.1, step = 0.001, ticks=FALSE)
-                            
-                     ),    
+                     # column(2,
[…]

### 2016-10-05 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
Ians suggestions

@@
 
-### version 5
+### version 6
 
@@
 It is aimed at a readership with limited experience of population genetics and resistance.
-It uses a model representing two independent loci coding for resistance to two insecticides.
-It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, and the selective advantage of resistance.
+It uses a model representing two independent genes coding for resistance to two insecticides.
+It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, and how resistance restores fitness in the presence of the insecticides.
 We use the model to develop a mechanistic understanding of how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies.
@@
 > 
-> Individual fitness = Allele fitness * exposure to insecticide  
+> Population fitness = Allele fitness * exposure to insecticide  
 > 
 > In each generation :   
-> Allele frequency[t+1] = Allele frequency[t] * Individual fitness
+> Allele frequency[t+1] = Allele frequency[t] * Population fitness
 
-Sexual reproduction is represented first by creating gametes by splitting the adult genomes in two, and secound by forming offspring genotypes by adding gametes with recombination.
+Sexual reproduction is represented first by creating gametes by splitting the adult genomes in two with recombination, and second by forming offspring genotypes by adding gametes.
 
 So far this description has just considered a single insecticide and associated resistance allele (Fig 1A).
-In the simulation a secound insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
+In the simulation a second insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
 Thus the simulation can represent populations exposed to two insecticides together in a mixture.
@@
 ### Single insecticide
-For single insecticide use, faster resistance resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 3-5).
-In the plots each input was varied in isolation with the remaining set at a constant 0.5 (except for starting frequency of resistance which was set to 0.01).
+For single insecticide use, faster resistance spread resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 3-5).
+In the plots each input was varied in isolation with the remaining set to 0.5 (except for starting frequency of resistance which was set to 0.01 in all bar Fig 5).
 
@@
 To help us understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs individually.
-For our base scenario we set the values for the first four inputs in Table 1 to 0.5 and the starting frequencies of resistance to 0.1.
+The base scenario was similar to that used for the single insecticide.
+Values for the first four inputs in Table 1 were set to 0.5 and the starting frequencies of resistance to 0.01.
 For this base scenario resistance arises slower for sequential than mixture (Fig 6A).
-This plot, and those that follow, compares mixture and sequence for a single input scenario (where the previous plots showed 3 input scenarios).
+This plot, and those that follow, compares mixture and sequence for a single combination of inputs.
 Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer to reach the resistance threshold for both.
+When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold of 50%.
 
 #### results: Insecticide effectiveness
-A more effective insecticide prompts increased time-to-resistance when used in a mixture.
-This is opposite to what happens when used in sequence and in isolation, both prompting a shorter times-to-resistance.
+A more effective insecticide increases time-to-resistance when used in a mixture. This is opposite to what happens when used in sequence and in isolation, when both shortened times-to-resistance.
 
@@
 Increasing exposure to 0.8 from the base scenario reduces time-to-resistance for both mixture and sequential scenarios (Fig 8A relative to 6A).
-The change in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
+The increase in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
 
@@
 In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
-Looking at our base scenario (Fig 6A) the the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
+Looking at our base scenario (Fig 6A) the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
 The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
@@
          guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "rr restoration", y=NULL) +
+         labs(colour = "resistance restoration", y=NULL) +
          coord_trans(y = "log10") +   

### 2016-10-06 R/fitnessGenotype.r
to pass R CMD check

@@ -5,12 +5,23 @@
 #' @param a_fitnic array of niche fitnesses
 #' @param a_expos array of exposure in each niche
 #' @param a_fitgen array of individual fitnesses to fill
+#' @param eff1 effectiveness1
+#' @param eff2 effectiveness2
+#' @param dom1 dominance1
+#' @param dom2 dominance2
+#' @param rr_1 resistance restoration 1 selection coefficient = resistance restoration * effectiveness
+#' @param rr_2 resistance restoration 2
+#' @param cost1 fitness cost of R in no insecticide
+#' @param cost2 fitness cost of R in no insecticide
+#' @param fitSS1 fitness of SS1 if no insecticide
+#' @param fitSS2 fitness of SS2 if no insecticide
+#' @param exposure exposure 
 #' @param plot whether to plot fitness results
 #' 
 #' @examples 
 #' fitnessGenotype()
 #' 
-#' a_fitloc <- fitnessSingleLocus( effectiveness = 0.8 )
+#' a_fitloc <- fitnessSingleLocus( eff1 = 0.8, eff2 = 0.8 )
 #' a_fitnic <- fitnessNiche( a_fitloc = a_fitloc )
 #' a_expos <- setExposure( exposure=0.5, insecticideUsed = 'mixture' )
 #' a_fitgen <- fitnessGenotype( a_fitnic = a_fitnic, a_expos = a_expos )
@@ -87,9 +98,9 @@ fitnessGenotype <- function ( a_fitnic = NULL,
   
   #error check for fitnesses > 1 or < 0
   if ( any(a_fitgen > 1 ) ) 
-    warning( sum(a_fitgen > 1  ), " individual fitness values (Wloci) are >1")
+    warning( sum(a_fitgen > 1  ), " individual fitness values (a_fitloc) are >1")
   if ( any( a_fitgen < 0 ) ) 
-    warning( sum( a_fitgen < 0 ), " individual fitness values (Wloci) are <0")
+    warning( sum( a_fitgen < 0 ), " individual fitness values (a_fitloc) are <0")
   
   #testing
   #cat("in fitnessGenotype\n")

### 2016-10-06 R/fitnessNiche.r

@@ -9,10 +9,10 @@
 #' @examples 
 #' fitnessNiche()
 #' 
-#' a_fitloc <- fitnessSingleLocus( effectiveness = 0.8 )
+#' a_fitloc <- fitnessSingleLocus( eff1 = 0.8 )
 #' a_fitnic <- fitnessNiche( a_fitloc = a_fitloc )
-#' a <- setExposure( exposure=0.5, insecticideUsed = 'mixture' )
-#' Windiv <- fitnessIndiv( a_fitnic = a_fitnic, a = a )
+#' a_expos <- setExposure( exposure=0.5, insecticideUsed = 'mixture' )
+#' a_fitgen <- fitnessGenotype( a_fitnic = a_fitnic, a_expos = a_expos )
 
 #' @return fitness values in an array
 #' @export

### 2016-10-06 R/fitnessPrint.r

@@ -2,69 +2,82 @@
 #' 
 #' for model testing & understanding
 
-#' @param effectiveness1 effectiveness
-#' @param effectiveness2 effectiveness
-#' @param dominance dominance
-#' @param selection_co selection_coefficient
-#' @param cost fitness cost of R in no insecticide
-#' @param SS fitness of SS if no insecticide
-#' @param h dominance array
-#' @param s selection coefficient array
-#' @param phi effectiveness array
-#' @param z cost array
-#' @param Wloci array of single locus fitnesses to fill
+#' @param eff1 effectiveness1
+#' @param eff2 effectiveness2
+#' @param dom1 dominance1
+#' @param dom2 dominance2
+#' @param rr_1 resistance restoration 1 selection coefficient = resistance restoration * effectiveness
+#' @param rr_2 resistance restoration 2
+#' @param cost1 fitness cost of R in no insecticide
+#' @param cost2 fitness cost of R in no insecticide
+#' @param fitSS1 fitness of SS1 if no insecticide
+#' @param fitSS2 fitness of SS2 if no insecticide
+#' @param a_dom dominance array
+#' @param a_sel selection coefficient array
+#' @param a_effect effectiveness array
+#' @param a_cost cost array
+#' @param a_fitloc array of single locus fitnesses to fill
+#' @param exposure exposure
 #' @param insecticideUsed one of 'insecticide1','insecticide2','mixture'
 #' 
 #' @examples 
 #' fitnessPrint()
-#' fitnessPrint(effectiveness1=0.8, effectiveness2=0.8, exposure=0.5, insecticideUsed='insecticide1')
-#' fitnessPrint(effectiveness1=0.8, effectiveness2=0.8, exposure=0.5, insecticideUsed='insecticide2')
-#' fitnessPrint(effectiveness1=0.8, effectiveness2=0.8, exposure=0.5, insecticideUsed='mixture')
+#' fitnessPrint(eff1=0.8, eff2=0.8, exposure=0.5, insecticideUsed='insecticide1')
+#' fitnessPrint(eff1=0.8, eff2=0.8, exposure=0.5, insecticideUsed='insecticide2')
+#' fitnessPrint(eff1=0.8, eff2=0.8, exposure=0.5, insecticideUsed='mixture')
 
 #' @return fitness values
 #' @export
 
-fitnessPrint <- function ( effectiveness1 = 0.5,
-                           effectiveness2 = 0.5,
-                                 dominance = 0.5,
-                                 selection_co = 0.5,
[…]

### 2016-10-06 R/linkage_calc.r

@@ -8,7 +8,10 @@
 #' @param results matrix
 #' 
 #' @examples 
-#' linkage_calc()
+#' results <- matrix ( nrow = 1, ncol = 12 )
+#' genotype.freq <- make.genotypemat ( 0.1, 0.5 )
+#' a_gtypes <- createArray2( sex=c("m","f"), loci=rownames( genotype.freq ) )
+#' linkage_calc(a_gtypes, 0.5, 1, results)
 
 #' @return results matrix with values for this generation added
 #' @export

### 2016-10-06 R/plot_fit_rs.r

@@ -16,8 +16,8 @@
 #' 
 #'  
 #' #niche
-#' library(reshape2)
-#' df_fit2 <- melt(fitnessNiche()[,,'A','B'])
+#' #library(reshape2)
+#' df_fit2 <- reshape2::melt(fitnessNiche()[,,'A','B'])
 #' rownames(df_fit2) <- paste(df_fit2$locus1, df_fit2$locus2)
 #' plot_fit_rs(df_fit2, 'value')
 #' 
@@ -27,24 +27,23 @@
 #' df_fit3 <- as.data.frame(t(as.data.frame(fitnessGenotype())))
 #' plot_fit_rs(df_fit3,'f')
 #' 
-#' @return 
+#' @return dataframe of plotted fitness values and colours
 #' @export
 
 plot_fit_rs <- function ( df_fit, column, column_facet = NULL, title = NULL ){
  
-  library(ggrepel)
-  
-  library(stringr)
+  #library(ggrepel)
+  #library(stringr)
 
   #counting S & R, and making a first test colour scheme with red for resist & blue for susceptible 
-  df_fit$num_s <- str_count(rownames(df_fit),'S')
-  df_fit$num_r <- str_count(rownames(df_fit),'R')  
+  df_fit$num_s <- stringr::str_count(rownames(df_fit),'S')
+  df_fit$num_r <- stringr::str_count(rownames(df_fit),'R')  
   df_fit$col <- rgb(df_fit$num_r/max(df_fit$num_r),0,df_fit$num_s/max(df_fit$num_s), alpha=0.5)
   
   gg <- ggplot(df_fit, aes_string(x=1, y=column)) +
         ylim(0,1) +
         geom_point(size=3, colour=df_fit$col) +
-        geom_text_repel( aes(label=rownames(df_fit)), nudge_x=0.1) +
+        ggrepel::geom_text_repel( aes(label=rownames(df_fit)), nudge_x=0.1) +
         ylab('fitness') +
         ggtitle(title) + 
         theme_bw() +

### 2016-10-06 R/plotlinkage.r

@@ -1,8 +1,8 @@
 #' plot Linkage Disequilibrium over generations
 #' 
 #' @param mat results matrix
-#' @param plot_d
-#' @param plot_dprime
+#' @param plot_d whether to plot d
+#' @param plot_dprime whether to plot d_prime
 #' @param main plot title
 #' @param xlab xlab
 #' @param ylab ylab

### 2016-10-06 R/plotsimpfit.r

@@ -42,7 +42,7 @@ par(mar = c(0,0,0,0),oma = c(0, 0, 0, 0))
 plot(c(-0.2,1.5),c(0,1), type='n', axes=FALSE, xlab='', ylab='', asp=1)
 
 
-rect(xleft = 0, xright = 0.05, ybottom = 0, ytop = 1, col=rgb(r=0, g=0, b=0.5, a=0.5))
+rect(xleft = 0, xright = 0.05, ybottom = 0, ytop = 1, col=rgb(0, 0, 0.5, alpha=0.5))
 
 text(x=c(0,0), y=c(0,1), labels=c(0,1), pos=2) #pos=1234 bltr
 

### 2016-10-06 R/randomMating.r

@@ -16,12 +16,11 @@
 #'                 "RS1SS2", "RS1RS2_cis", "RS1RS2_trans", "RS1RR2",
 #'                 "RR1SS2", "RR1RS2", "RR1RR2")
 #' sex2 <- c("m","f")
-#' #f  = genotype frequencies before selection
-#' f <- createArray2( sex=sex2, loci=namesLoci )
+#' a_gtypes <- createArray2( sex=sex2, loci=namesLoci )
 #' #assign all of both parents to one genotype for convenience
-#' f['m','RS1RS2_trans'] <- 1
-#' f['f','RS1RS2_trans'] <- 1
-#' G <- createGametes( f = f, recomb_rate = 0.5 )
+#' a_gtypes['m','RS1RS2_trans'] <- 1
+#' a_gtypes['f','RS1RS2_trans'] <- 1
+#' G <- createGametes( a_gtypes = a_gtypes, recomb_rate = 0.5 )
 #' fGenotypeExpanded <- randomMating(G) 
 #' #and for sex linked
 #' fGenotypeExpanded2 <- randomMating(G, sexLinked=TRUE, isMale=TRUE) 

### 2016-10-06 R/relaxSelection.r

@@ -1,36 +1,36 @@
 #' relax selection by modifying indiv fitnesses
 #' 
 #' 
-#' @param Windiv array of individual fitnesses
+#' @param a_fitgen array of individual fitnesses
 #' @param k generation number
 #' 
 #' @return array of individual fitnesses modified
 
 #don't @export
 
-relaxSelection <- function(Windiv, k){
+relaxSelection <- function(a_fitgen, k){
   
 
   if( (2<k) & (k<12)){
     
     # relaxed selection fitnesses
-    Windiv[] <- 0.1
+    a_fitgen[] <- 0.1
     
   } else if( (k>11) & (k<15)){
     
     # non relaxed fitnesses
-    Windiv[] <- 0.1
+    a_fitgen[] <- 0.1
     #setting m,f at same time
-    Windiv[,'RS1','RS2'] <- 1
-    Windiv[,'RS1','RR2'] <- 1
-    Windiv[,'RR1','RS2'] <- 1
-    Windiv[,'RR1','RR2'] <- 1
+    a_fitgen[,'RS1','RS2'] <- 1
+    a_fitgen[,'RS1','RR2'] <- 1
+    a_fitgen[,'RR1','RS2'] <- 1
+    a_fitgen[,'RR1','RR2'] <- 1
     
   } else if( k>14 ){
     
     # relaxed selection fitnesses
-    Windiv[] <- 0.1
+    a_fitgen[] <- 0.1
   }
   
-return(Windiv)
+return(a_fitgen)
 }

### 2016-10-06 R/resistance-package.r

@@ -1,0 +1,15 @@
+#' insecticide resistance modelling
+#' 
+#' submitted model of the development of insecticide resistance.
+#' 
+#' @name resistance-package
+#' @aliases resistance-package resistance
+#' @docType package
+#' @author Andy South
+#' 
+#' @keywords package
+#' @import grDevices graphics ggplot2
+#' @importFrom stats as.formula runif
+#' @importFrom utils read.csv write.csv
+
+NULL

### 2016-10-06 R/runcurtis_f2.r

@@ -29,8 +29,8 @@
 #' @param z.RR2_00 fitness cost of resistance allele 2 in insecticide free environment
 #' @param sexLinked whether resistance is sex linked, default=0(FALSE)
 #' @param male_exposure_prop male exposure as a propoertion of female, default 1 for same, likely <1  
-#' @param  proportion of correct deployment of mixtures, 
-#'    if <1 other portion divided between single insecticides
+#' @param correct_mix_deploy proportion of times that mixture is deployed correctly, 
+#'    assumes that when not deployed correctly the single insecticides are used instead
 #' @param addCombinedStrategy whether to add a line for the combined mixture/individual strategy, makes plot confusing so set to FALSE by default
 #' @param addStrategyLabels add labels to plot for where strategies cross 0.5 line TRUE gives seq,mix1,mix2,mix3
 #' @param strategyLabels = c("seq","mix1","adapt","mix2")

### 2016-10-06 R/setExposure.r

@@ -11,7 +11,6 @@
 #' @param male_exposure_prop proportion tht males are exposed relative to f, default 1, likely to be <1
 #' @param correct_mix_deploy proportion of times that mixture is deployed correctly, 
 #'    assumes that when not deployed correctly the single insecticides are used instead
-#'    TODO check my understanding of this
 #' 
 #' @examples
 #' a <- setExposure( exposure=0.9, insecticideUsed = 'mixture' )

### 2016-10-06 inst/shiny/simpfitvis/server.r

@@ -5,10 +5,14 @@
 
 library(shiny)
 #library(devtools)
-#install_github('AndySouth/coverage')
-library(coverage)
+#install_github('AndySouth/resistance')
 library(resistance)
 library(png)
+
+#these may be needed: I should really import them to resistance
+library(ggrepel)
+library(stringr)
+
 
 shinyServer(function(input, output, session) {
 

### 2016-10-06 tests/testthat/testRefactoring.r

@@ -4,6 +4,9 @@ context("Refactoring")
 
 test_that("refactoring named variables with arrays doesn't change results", {
   
+  
+  skip("not doing refactoring tests anymore")
+  
   input <- setInputOneScenario()
   
   #expect refactored runModel2() to generate same results as runModel() for single default run
@@ -12,12 +15,14 @@ test_that("refactoring named variables with arrays doesn't change results", {
   expect_equal( runModel2(input, produce.plots=FALSE),
                     runModel(input, produce.plots=FALSE) )  
   
+  # removed this because they started to differ due to a valid change
+  
   #set random seed (may want to move option to setInputSensiScenarios() later )
-  set.seed(1)
-  input <- setInputSensiScenarios(20, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
-
-  #expect refactored runModel2() to generate same results as runModel() for a range of scenarios
-  listOut2 <- runModel2(input, produce.plots=FALSE)
-  listOut <- runModel(input, produce.plots=FALSE)
-  expect_equal( listOut2, listOut )
+  # set.seed(1)
+  # input <- setInputSensiScenarios(20, h.RS1_A0=c(0.1,1), h.RS2_0B=c(0.1,1), s.RR1_A0=c(0.2,1), s.RR2_0B=c(0.2,1))
+  # 
+  # #expect refactored runModel2() to generate same results as runModel() for a range of scenarios
+  # listOut2 <- runModel2(input, produce.plots=FALSE)
+  # listOut <- runModel(input, produce.plots=FALSE)
+  # expect_equal( listOut2, listOut )
 })

### 2016-10-06 tests/testthat/test_fitnessNiche.r

@@ -3,22 +3,19 @@ library(resistance)
 context("fitness")
 
 test_that("niche fitness ...", {
-  
-  # Wniche <- fitnessNiche( Wloci = Wloci,
-  #                         niche = niche,
-  #                         Wniche = Wniche )
+
   
   # to do add a loop here to do tests for a range of input values
   
-  expect_silent( Wniche <- fitnessNiche() )
+  expect_silent( a_fitnic <- fitnessNiche() )
   
   #these warnings are already in fitnessNiche()
   #so I could put expect_silent() around the above call and it would already be covered
   #error check for fitnesses > 1 or < 0
-  # if ( any(Wniche > 1  ) ) 
-  #   warning( sum(Wniche > 1  ), " niche fitness values (Wniche) are >1 ")
-  # if ( any( Wniche < 0 ) ) 
-  #   warning( sum( Wniche < 0 ), " niche fitness values (Wniche) are <0")   
+  # if ( any(a_fitnic > 1  ) ) 
+  #   warning( sum(a_fitnic > 1  ), " niche fitness values (a_fitnic) are >1 ")
+  # if ( any( a_fitnic < 0 ) ) 
+  #   warning( sum( a_fitnic < 0 ), " niche fitness values (a_fitnic) are <0")   
 
   
   # ..$ locus1: chr [1:3] "SS1" "RS1" "RR1"
@@ -28,7 +25,7 @@ test_that("niche fitness ...", {
   
   #expect ...
   #just an initial test
-  expect_equal( Wniche['RR1','RR2','A','B'], 1 )
+  #expect_equal( a_fitnic['RR1','RR2','A','B'], 1 )
   
   
 })

### 2016-10-06 tests/testthat/test_fitnessSingleLocus.r

@@ -4,28 +4,17 @@ context("fitness")
 
 test_that("single locus fitness ...", {
   
-  # Wloci <- fitnessSingleLocus(h = h,
-  #                             s = s,
-  #                             phi = phi,
-  #                             z = z)
-  # can be called with arrays or
-  # effectiveness1 effectiveness
-  # effectiveness2 effectiveness
-  # dominance dominance
-  # selection_co selection_coefficient
-  # cost fitness cost of R in no insecticide
-  # SS fitness of SS if no insecticide
   
   # to do add a loop here to do tests for a range of input values
   
-  expect_silent( Wloci <- fitnessSingleLocus() )
+  expect_silent( a_fitloc <- fitnessSingleLocus() )
   
   #these warnings are already in fitnessSingleLocus()
   #so I could put expect_silent() around the above call and it would already be covered
-  # if ( any( Wloci > 1  ) ) 
-  #   warning( sum(Wloci > 1 ), " locus fitness values (Wloci) are >1 : ", Wloci[Wloci>1])
-  # if ( any( Wloci < 0 ) ) 
-  #   warning( sum( Wloci < 0 ), " locus fitness values (Wloci) are <0")   
+  # if ( any( a_fitloc > 1  ) ) 
+  #   warning( sum(a_fitloc > 1 ), " locus fitness values (a_fitloc) are >1 : ", a_fitloc[a_fitloc>1])
+  # if ( any( a_fitloc < 0 ) ) 
+  #   warning( sum( a_fitloc < 0 ), " locus fitness values (a_fitloc) are <0")   
 
   # todo be careful that it is still possible to genereate fitness > 1
   # because selection_coefficient is used
@@ -33,18 +22,18 @@ test_that("single locus fitness ...", {
   # then i think it should be impossible to have fitness > 1
   
   # test like this
-  # > Wloci['SS1',]
+  # > a_fitloc['SS1',]
   # no  lo  hi 
   # 1.0 1.0 0.5 
-  # > Wloci['SS1','hi']
+  # > a_fitloc['SS1','hi']
   # [1] 0.5
-  # > Wloci[,'hi']
+  # > a_fitloc[,'hi']
   # SS1  RS1  RR1  SS2  RS2  RR2 
   # 0.50 0.75 1.00 0.50 0.75 1.00   
   
   #expect ...
   #just an initial test
-  expect_equal( Wloci['RR1','hi'], 1 )
+  #expect_equal( a_fitloc['RR1','hi'], 1 )
   
   
 })

### 2016-10-10 Matt/git-instructions-for-matt.Rmd
instructions for Matt

---
title: "git-instructions-for-matt"
author: "Andy South"
date: "10 October 2016"
output: html_document
---

Instructions for creating a fork of the resistance repository, being able to add your own updates, and to receive updates from me.

https://github.com/AndySouth/resistance

1) click fork
will create : https://github.com/Mapowney/resistance

2) click clone to get the URL to use in git clone
#from new local folder 
git clone git@github.com:Mapowney/resistance.git

3) set up syncing
cd resistance
git remote add upstream git@github.com:AndySouth/resistance.git

#to view remote repos
git remote -v

4) SYNC FORK from upstream, to get updates from me
git fetch upstream
git checkout master
git merge upstream/master

5) COMMIT & PUSH local changes to the fork, to put your own updates on github
(this step can be done from RStudio UI)
git commit -a -m "message" 
git push

6)

### 2016-10-14 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
tweak figs

@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", main='B.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", ylabs = FALSE, main='B.', maxX = 120 )
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2') )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), maxX = 120 )
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
 #cat("Fig. Exposure 0.8 benefit of sequence over mixture increases")
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), main='B.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylabs = FALSE, main='B.', maxX = 120 )
 #cat("Fig. Effectiveness 0.8, Exposure 0.8, increased exposure nullifies the effect of increased effectiveness, sequence remains better")
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
 #cat("Dominance of insecticide1 0.8")
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.8 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.8 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', ylabs = FALSE, main='B.', maxX = 120 )
 #cat("Dominance of both insecticides 0.8")
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
 #cat("Fig. rr_restoration I1 0.8")
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.8 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.8 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', ylabs = FALSE, main='B.', maxX = 120 )
 #cat("Fig. CC. rr_restoration I2 0.2")
@@
 #P_1 to 0.001 seq remains better
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 150 )
 
@@
 #P_1 to 0.001 & effect1 to 0.8 mix remains better
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', main='B.' )
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', ylabs = FALSE, main='B.', maxX = 150 )
 

### 2016-10-14 inst/documents/effectiveness_exposure_difference.Rmd
start doc on diff between effectiveness & exposure

---
title: "Why is time-to-resistance in mixtures influenced differently by effectiveness and exposure ?"
output: pdf_document
#output: word_document
author: "Andy South"
date: "`r Sys.Date()`"
---

```{r,echo=FALSE,results='hide'}
library(resistance)
library(ggplot2)
```

Increasing the exposure to one insecticide in a mixture has liitle influence 
on the time-to-resistance of the other. In contrast increasing the effectiveness 
of one insecticide in a mixture increases the time-to-resistance of the other.

Why is there this difference between the effects of effectiviness and exposure in mixtures ?

For single insecticide application increasing effectiveness or exposure have the same effect
of decreasing time-to-resistance by increasing selection pressure.

```{r, eval=TRUE, echo=FALSE}

#setting up the data

df_base_ <- as.data.frame(fitnessGenotype(exp1=0.5, exp2=0.5, plot=FALSE))[2,]
df_base <- df_base_/max(df_base_)

df_ef_ <- as.data.frame(fitnessGenotype(exp1=0.5, exp2=0.5, eff1=0.8,  plot=FALSE))[2,]
df_ef <- df_ef_/max(df_ef_)
#df_ef

df_ex_ <- as.data.frame(fitnessGenotype(exp1=0.8, exp2=0.5, plot=FALSE))[2,]
df_ex <- df_ex_/max(df_ex_)
#df_ex

#to allow me to plot them

#normalised ones
df_base <- data.frame(t(df_base),row.names=row.names(t(df_base)))
df_ef <- data.frame(t(df_ef),row.names=row.names(t(df_ef)))
df_ex <- data.frame(t(df_ex),row.names=row.names(t(df_ex)))
#not normalised
df_base_ <- data.frame(t(df_base_),row.names=row.names(t(df_base_)))
df_ef_ <- data.frame(t(df_ef_),row.names=row.names(t(df_ef_)))
df_ex_ <- data.frame(t(df_ex_),row.names=row.names(t(df_ex_)))

```

```{r, eval=TRUE, echo=FALSE}

plot_fit_rs(df_base, column='f', title='base')

plot_fit_rs(df_ef, column='f', title='effectiveness')

plot_fit_rs(df_ex, column='f', title='exposure')

```

Seems that in the case where effectiveness is increased the fitnesses are closer together so does that mean that there is less selection for resistance ?

What is happening with the 2nd insecticide ?

#Normalised
```{r, eval=TRUE, echo=FALSE}

calc_fit_rs(df_base,'f',round_places=3)
#R2-S2  R2-S2 0.066

calc_fit_rs(df_ef,'f',round_places=3)
#R2-S2  R2-S2 0.045

calc_fit_rs(df_ex,'f',round_places=3)
#R2-S2  R2-S2 0.073
```

So yes, increasing effectiveness I1 results in a decrease in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.045), this can explain the slower selection for resistance.

In contrast increasing exposure to I1 results in an increase in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.073), this can explain the faster selection for resistance.

Be a bit careful, does my normalising do anything weird ? I might be able to run calc_fit_rs() on the non normalised versions.
not normalised
calc_fit_rs(df_base_,'f')
      allele    meanfit
R1        R1 0.721
S1        S1 0.669
R2        R2 0.721
S2        S2 0.669
R1-S1  R1-S1 0.052
#R2-S2  R2-S2 0.052

calc_fit_rs(df_ef_,'f')
      allele    meanfit
R1        R1 0.666
S1        S1 0.583
R2        R2 0.641
S2        S2 0.608
R1-S1  R1-S1 0.083
#R2-S2  R2-S2 0.033

calc_fit_rs(df_ex_,'f')
      allele    meanfit
R1        R1 0.633
S1        S1 0.531
R2        R2 0.608
S2        S2 0.556
R1-S1  R1-S1 0.102
#R2-S2  R2-S2 0.052

So in the non normalised version the difference in fitness for S&R I2 stays the same when I1 exposure is increased but decreases when I1 effectiveness is increased.

In text form :
In a mixture, increasing the exposure to one insecticide has very little effect on the selection pressure for resistance to the other. In contrast increasing the effectiveness of one insecticide in a mixture decreases the slection pressure for resistance to the other.

### 2016-10-17 inst/documents/effectiveness_exposure_difference.Rmd
adding examples & finessing case

@@
 
-Increasing the exposure to one insecticide in a mixture has liitle influence 
-on the time-to-resistance of the other. In contrast increasing the effectiveness 
+Increasing the exposure to one insecticide in a mixture has little influence 
+on the time-to-resistance of the other. In contrast, increasing the effectiveness 
 of one insecticide in a mixture increases the time-to-resistance of the other.
@@
 of decreasing time-to-resistance by increasing selection pressure.
+
+scaneario exploration :
+
+1. base : all inputs set to 0.5
+1. effectiveness : effectiveness set to 0.8
+1. exposure : exposure set to 0.8
 
@@
 
-```{r, eval=TRUE, echo=FALSE}
+# before normalisation
+```{r, eval=TRUE, echo=FALSE, results='hide', fig.width=2}
 
-plot_fit_rs(df_base, column='f', title='base')
+plot_fit_rs(df_base_, column='f', title='base', ylim=c(0.4,0.8))
 
-plot_fit_rs(df_ef, column='f', title='effectiveness')
+plot_fit_rs(df_ef_, column='f', title='effectiveness', ylim=c(0.4,0.8), yblank=TRUE)
 
-plot_fit_rs(df_ex, column='f', title='exposure')
+plot_fit_rs(df_ex_, column='f', title='exposure', ylim=c(0.4,0.8), yblank=TRUE)
+
+```
+
+# normalised
+```{r, eval=TRUE, echo=FALSE, results='hide', fig.width=2}
+
+plot_fit_rs(df_base, column='f', title='base', ylim=c(0.5,1))
+
+plot_fit_rs(df_ef, column='f', title='effectiveness', ylim=c(0.5,1), yblank=TRUE)
+
+plot_fit_rs(df_ex, column='f', title='exposure', ylim=c(0.5,1), yblank=TRUE)
 
@@
 What is happening with the 2nd insecticide ?
+
+To look at this I calculated the mean fitness of the R & S alleles for each insecticide across all genotypes.
+(and I repeated this both for normalised and non-normalised fitnesses).
+This was done by multiplying the fitness of each genotype by the number of each allele and then calculating a mean.
+
+```{r, eval=TRUE, echo=FALSE, results='markup'}
+#just to give the example of the calculation
+calc_fit_rs(df_base,'f',round_places=2, verbose=TRUE)
+```
+
+The difference between the resistant and susceptible allele for each insecticide (e.g. R1-S1) was then calculated.
 
@@
 
-So yes, increasing effectiveness I1 results in a decrease in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.045), this can explain the slower selection for resistance.
+So yes, increasing effectiveness I1 results in a decrease in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.045). This could explain the slower selection for resistance.
 
-In contrast increasing exposure to I1 results in an increase in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.073), this can explain the faster selection for resistance.
+In contrast increasing exposure to I1 results in an increase in the difference between the relative fitness of S & R to I2 (from 0.066 to 0.073).
+This could explain the faster selection for resistance.
 
-Be a bit careful, does my normalising do anything weird ? I might be able to run calc_fit_rs() on the non normalised versions.
-not normalised
+Similar results are obtained without normalising the fitnesses first.
+
+#not normalised
+```{r, eval=TRUE, echo=FALSE}
+
 calc_fit_rs(df_base_,'f')
-      allele    meanfit
-R1        R1 0.721
-S1        S1 0.669
-R2        R2 0.721
-S2        S2 0.669
-R1-S1  R1-S1 0.052
 #R2-S2  R2-S2 0.052
@@
 calc_fit_rs(df_ef_,'f')
-      allele    meanfit
-R1        R1 0.666
-S1        S1 0.583
-R2        R2 0.641
-S2        S2 0.608
-R1-S1  R1-S1 0.083
 #R2-S2  R2-S2 0.033
@@
 calc_fit_rs(df_ex_,'f')
-      allele    meanfit
-R1        R1 0.633
-S1        S1 0.531
-R2        R2 0.608
-S2        S2 0.556
-R1-S1  R1-S1 0.102
 #R2-S2  R2-S2 0.052
+
+```
 

### 2016-10-18 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
adding figs for final submission

@@
 
-## Figures for paper 1 : version 8
+## Figures for paper 1 : version 9 : looking at producing Tiffs for submission
 
@@
 
+```{r,include=FALSE}
+
+library(knitr)
+
+# options to create final publication quality figures
+# causes the pdf document to fail, but the figs do get stored in pap1figs/
+opts_chunk$set(dev="tiff",
+               dev.args=list(compression="lzw"),
+               dpi=300,
+               cache=FALSE,
+               fig.path='pap1figs/')
+# opts_chunk$set(dev='postscript')
+```
+
 ```{r,echo=FALSE,results='hide'}
@@
   # load(file=paste0(outFolder,'listOutI2_ex2_3.rda'))
+
+```
+
+#Fig1 
+```{r Fig1, eval=TRUE, echo=FALSE, fig.height=7.5, fig.width=5, fig.show='hold', fig.align='left'}
+
+library(ggplot2)
+library(cowplot)
+
+#list to put plots in for cowplot
+plotlist <- list(2)
+
+#Fig 1A
+
+dfd1 <- selection_simple( freq=0.01, selection_co = 1, dominance = 1, num_gens = 150, plot=FALSE)
+dfd05 <- selection_simple( freq=0.01, selection_co = 1, dominance = 0.5, num_gens = 150, plot=FALSE)
+dfd0 <- selection_simple( freq=0.01, selection_co = 1, dominance = 0, num_gens = 150, plot=FALSE)
+
+df_to_plot <- rbind(dfd1, dfd05, dfd0)
+
+#old way of getting in wide format  
+# df_to_plot <- data.frame(generation=dfd1$generation,
+#                          dominance1=dfd1$freq,
+#                          dominance0.5=dfd05$freq,
+#                          dominance0=dfd0$freq)
+
+plotlist[[1]] <- ggplot(df_to_plot, aes(x=generation, y=freq, 
+                              colour=factor(dominance,labels=c("0   recessive",
+                                                               "0.5 partial",
+                                                               "1   dominant")))) + 
+       theme_bw() +  
+       #theme(legend.position = "bottom", legend.key = element_blank()) + 
+       #guides(colour = guide_legend(reverse=TRUE)) +  
+       labs(colour = "dominance") +
+       ylab("Resistance Allele Frequency") +   
+       #coord_trans(y = "log10") +    
+       geom_line()
+
+#Fig 1B
+
+df_labs <- data.frame(lab=rep(c('RR','RS','SS'),4),
+                      x=c(10,10,10,30,30,30,50,50,50,70,70,70),
+                      y=c(50,46,42,38,34,30,26,22,18,14,10,6),
+                      #col=c(2,1,1,2,1.5,1,2,2,1,2,2,2))
+                      col=c(1,0,0,1,0.5,0,1,1,0,1,1,1))
+
+df_boxes <- data.frame(lab=c('R recessive\n(slow spread)',
+                             'R semi-dominant\n(rapid spread)',
+                             'R dominant\n(rapid spread)',
+                             'No selection for R'),
+                      x=c(10,30,50,80),
+                      y=c(90,75,60,35))
+
+plotlist[[2]] <- ggplot(data.frame(x=c(0,100),y=c(80,0)), aes(x=x, y=y)) + 
+       theme_bw() + 
+       ylim(0,100) +   
+       ylab("Insecticide concentration") +
+       xlab("Time since treatment") +         
+       geom_line(size=2, colour=rgb(0.5,0,0.5)) +
+       geom_vline(xintercept = c(20,40,60), linetype=2) +
+       geom_text(data=df_labs,aes(x=x, y=y, label=lab, colour=factor(col)),size=4) +
+       scale_colour_manual(values = c("0" = "red","0.5" = "orange","1" = "green"), guide=FALSE) +  
+       geom_label(data=df_boxes,aes(x=x, y=y, label=lab)) +
+       theme(axis.text = element_blank(),
+             axis.ticks = element_blank(),
+             panel.grid = element_blank())
+
+plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(0.7,1))
+
+```
+
+# Fig2 replicating Curtis
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, fig.keep='last'}
+
+  #Beths code for these figures taken from resistanceMaster
+
+  #set layout for 2 plots
+  layout(t(1:2))
+
+  #2A
+  calibration <- 1011
+  input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
+  listOut <- runModel2( input, calibration )
+  #we submitted the LD one but should have done the Fig1
+  #plotcurtis_ld( listOut$results[[1]], listOut$results[[2]], 1, 4 )
+  plotcurtis_f1( listOut$results[[1]], listOut$results[[2]], 1, 2 )
+
+  #2B
+  calibration <- 1012
+  input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
+  listOut <- runModel2( input, calibration )  
+  plotcurtis_f2( listOut$results[[3]], listOut$results[[1]], listOut$results[[2]], 1, 2, 3 )
 
@@
 
-## Fig. x1 Time-to-resistance across all input values
+## Fig 3
+
+```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=4.5}
+
+#get input values from Ian, then I can re-generate, might want to change mix2 to mix
+#90% exposure, equal for female and male
+#Starting freq. at locus A =0.01 and at locus B is 0.001
+#Fitness =1 for RR genotypes at 0 for SR and SS at both loci
+
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.001, h.RS1_A0 = 1, h.RS2_0B = 1, exposure = 0.9,                        phi.SS1_A0 = 1 , phi.SS2_0B = 1 , s.RR1_A0 = 1 , s.RR2_0B = 1,
+              addCombinedStrategy=FALSE, addStrategyLabels=FALSE )
+
+```
+
+## Fig4 Time-to-resistance across all input values
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=4.5}
+```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=4.5}
 
@@
 
-## Fig x2 PRCC analysis for time-to-resistance 0.5 for sequential and mixture strategies
+## Fig5 PRCC analysis for time-to-resistance 0.5 for sequential and mixture strategies
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8}
+```{r Fig5, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8}
 
@@
 
-## Fig. x3 How time-to-resistance is effected by input values for each insecticide use strategy
+## Fig6 How time-to-resistance is effected by input values for each insecticide use strategy
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=2.5}
+```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7.5, fig.keep='last'}
+
+  #layout for 3 panel plot (t=horiz, c=vertical)
+  #doesn't work for ggplot
+  #layout(c(1:3))
+
+  #list to put plots in for cowplot
+  plotlist <- list(3)
 
@@
   #for(i in names_inputs)
+
+  plotnum <- 0
+
   for(i in in_cols)
@@
 
+    plotnum <- plotnum + 1
+
     #print( ggplot(ggInsOuts, aes_string(x=i, y=y, colour="strategy")) + 
-    print( ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) +
+    plotlist[[plotnum]] <- ggplot(ggSubset, aes_string(x=i, y=y, colour="strategy")) +
 
@@
              #geom_vline(data = curtisInputs, aes_string(xintercept = i, colour='"Curtis Fig.2"' ), linetype='dotted', show.legend=TRUE) +
-             geom_text(data = curtisInputs, inherit.aes = FALSE, aes_string(x = i, y=210, label='"Curtis Fig.2"'), size=3, colour='red')
-
-           )
+             geom_label(data = curtisInputs, inherit.aes=FALSE, aes_string(x=i, y=210, label='"Curtis\nFig.2"'), size=3, colour='red', nudge_x = -0.05)
+
   }
 
+  #to put back in some ref gridlines that are removed by cowplot
+  #theme_set(theme_gray())
+  theme_set(theme_bw())
+
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1))
+
 ```
@@
 
-## Fig. x4 PRCC difference in time-to-resistance 0.5 between sequential and adaptive mixture strategies
+## Fig7 PRCC difference in time-to-resistance 0.5 between sequential and adaptive mixture strategies
 
@@
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=5.5}
+```{r Fig7, eval=TRUE, echo=FALSE, warning=FALSE, fig.height=5.5}
 
@@
 
-## Fig. x5 Effectiveness and exposure together determine most of whether resistance arises slower under adaptive mixture or sequential strategies
+## Fig9 Effectiveness and exposure together determine most of whether resistance arises slower under adaptive mixture or sequential strategies
 1.
 Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
-```{r, eval=TRUE, echo=FALSE, fig.height=6}
+```{r Fig9, eval=TRUE, echo=FALSE, fig.height=6}
 
@@
 
-## Fig. x6 Modifying previous plot to show where resistance arises over 20% slower for adaptive mixture versus sequential.
+## Fig10 Modifying previous plot to show where resistance arises over 20% slower for adaptive mixture versus sequential.
 
 1. The line fitted to runs where resistance is reached at the same time for mixture and sequential is retained. Less of the parameter space favours the adaptive mixture under these constraints.
-```{r, eval=TRUE, echo=FALSE, fig.height=6}
+```{r Fig10, eval=TRUE, echo=FALSE, fig.height=6}
 
@@
 
-## Fig. x5.1 modified for mix2 vs seq
-
-fewer runs where sequential gives slower resistance
-
-```{r, eval=TRUE, echo=FALSE, fig.height=6}
-
-  ggSubset <- dif_mix2_seq
-  #if mix-seq >1 resistance took longer for mix so mix is better
-  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
-  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
-  ggSubset$abs_diff <- abs(ggSubset$mix2_minus_seq0.5)
-
-  #tricky
-  #be a bit careful what to do with zeros where results are same for both strategies
-  #I use these better vars to facet the plots
-  #may want the zeroes in both facets
-  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
-  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
-  #set these to 1 (they were 0)
-  df_same$mix2_better_seq0.5 <- 1
-  df_same$seq_better_mix20.5 <- 1
-  #rbind them back onto other dataset
-  ggSubset <- rbind( ggSubset, df_same)
-
-  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
-  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
-
-    #find unequal runs
-    unequal <- ggSubset[ggSubset$resist_start_lo_div_hi < 0.01,]
-    #nrow(ggSubset)
-    #nrow(unequal)
-    #2721 out of the 6189 remaining runs
-
-    #in how many runs was seq better than mix
-    #sum(unequal$seq_better_mix20.5)
-    #311
-
-    library(viridis)
-
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_diff))+
-
-             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mixture\ntime to resistance\nlonger','1'='sequential\ntime to resistance\nlonger'))) +
-
-             #YlGnBu looks really good except that the yellow points are very faint
-             #scale_colour_distiller(palette='YlGnBu', direction=1) +
-             #option="plasma" looks good too
-             scale_color_viridis(begin=1, end=0, option="viridis", 
-                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
-             theme_bw() +
-             theme(legend.position = "bottom") +
-             #theme_minimal() +
-             #add point for Curtis inputs
-             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='black', shape=3) +
-             #failing to add legend for curtis point
-             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
-             #scale_size_manual(name = "Curtis Fig.2", values="test") +
-             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
-             ylab('effectiveness of insecticides added') +
-             xlab('exposure to both insecticides') + 
-             #main points for the plot
-             geom_point(alpha=1, size=0.5) + 
-             #!!!NOW can I add a lm fitted to the 0 points abs_diff
-             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
-             #can I subtley highlight the zero points which are repeated in each plot ?
-             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
-
-             #now add the unequal points
-             #showed that they were on both sides
-             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=2) +
-
-             #add text for Curtis point
-             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_ins1 + effectiveness_ins2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
-           )
-
-  # doing the lm outside of ggplot    
-  x1=df_same$exposure 
-  y1=df_same$effectiveness_ins1 + df_same$effectiveness_ins2
-
-  l_mod <- lm(y1 ~ x1)
-  #l_mod$coefficients
-  #(Intercept)          x1 
-  #1.019330    0.560508
-
-  #approximation
-  #x2 <- seq(0,1,0.1)
-  #y2 <- 0.6*x2+0.98
-  #plot(x2,y2)
-
-```
-
-## Fig.
-x7 Classification Trees for whether resistance arises slower with an adaptive mixture or sequential strategy.
+## Fig8 Classification Trees for whether resistance arises slower with an adaptive mixture or sequential strategy.
 The upper plot shows when mixture is slower than sequential, the lower plot is for when the mixture is more than 20% slower.
 
-```{r, eval=TRUE, echo=FALSE, warning=TRUE, message=FALSE, fig.height=3.5}
-
-  #code copied from sensiAnPaper1All
+```{r Fig8, eval=TRUE, echo=FALSE, warning=TRUE, message=FALSE, fig.height=7, fig.keep='last'}
+
+  #layout for 2 panel plot (t=horiz, c=vertical)
+  layout(c(1:2))
 
@@
 
-  #just doing for a few strategies, sensiAnPpaer1All.Rmd does for a bunch
-  # treeResponses <- c( rownames(resistBetterMix_ASeq)[2], #2 is 25%
-  #                     rownames(resistBetterMix_ASeq20)[2]
-  #                   )
-  #4/7/16 eek above before was a bug showing 25% rather than 50%
   treeResponses <- c( "betterMix_ASeq_cP0.5","betterMix_ASeq20_cP0.5")
 
-  #to do trees & plots for all response variables
+  #pruned trees commented out, not used in paper1
+  # for( treeResponse in treeResponses )
+  # {
+  #   tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class') 
+  #   #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
+  #   #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
+  #   
+  #   cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
+  #   treePruned <- prune(tree, cp=cpOptimal)
+  #   
+  #   #test tree pruning, seemed to suggest we are pruning more than needed
+  #   # plotcp(tree)  
+  #   # cat("optimal cp used in pruning=",cpOptimal,"\n")
+  #   
+  #   #to set box colours
+  #   cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
+  #   
+  #   #add fig label
+  #   if (treeResponse == treeResponses[1]) main <- "A)" 
+  #   else  main <- "B)"
+  # 
+  #   #adds in no. scenarios correctly and incorrectly classified
+  #   prp(treePruned, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE)
+  # }
+
+  #4/7/2016 unpruned trees, repeats previous loop with pruning removed
   for( treeResponse in treeResponses )
@@
     tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')
-    #http://stackoverflow.com/questions/29197213/what-is-the-difference-between-rel-error-and-x-error-in-a-rpart-decision-tree
-    #A rule of thumb is to choose the lowest level where the rel_error + xstd < xerror.
-
-    cpOptimal <- tree$cptable[ tree$cptable[,"rel error"] + tree$cptable[,"xstd"]  < tree$cptable[,"xerror"],"CP"][1]
-    treePruned <- prune(tree, cp=cpOptimal)
-
-    #test tree pruning, seemed to suggest we are pruning more than needed
-    # plotcp(tree)  
-    # 
-    # cat("optimal cp used in pruning=",cpOptimal,"\n")
 
     #to set box colours
-    cols <- ifelse(treePruned$frame$yval == 1, "red", "green3")
+    cols <- ifelse(tree$frame$yval == 1, "red", "green3")
 
@@
     #adds in no. scenarios correctly and incorrectly classified
-    #prp(tree, extra=1, varlen = 10, main=treeResponse)
-    #prp(treePruned, extra=1, varlen = 10, main=treeResponse, col=cols, border.col=cols)
-    #prp(treePruned, extra=1, varlen = 0, main=treeResponse, box.col=cols, under=TRUE)
-    prp(treePruned, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE)
+    #cex to set textsize (0.7 all text doesn't quite fit in), good that A&B are now same size
+    #round=0 to make boxes square
+    #split.font=1 to stop bold text in labels
+    prp(tree, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE, 
+        cex=0.4, round=0, split.font=1, ycompress = TRUE )
   }
 
-  #4/7/2016 testing the unpruned trees, repeats previous loop with pruning removed
-  for( treeResponse in treeResponses )
-  {
-    tree <- rpart::rpart(as.formula(paste(treeResponse,"~",treePredictorString)), data = treeInput, method = 'class')
-
-    #to set box colours
-    cols <- ifelse(tree$frame$yval == 1, "red", "green3")
-
-    #add fig label
-    if (treeResponse == treeResponses[1]) main <- "A)"
-    else  main <- "B)"
-
-    #adds in no. scenarios correctly and incorrectly classified
-    prp(tree, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE)
-  }
-
-```
-
-```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5} 
-#eval=FALSE so this one not displayed
-
-## Fig.
-x8 Introducing a new insecticide when there is already resistance to an existing one.
-This takes Curtis Fig 2 and sets the starting frequency for each allele in turn to a higher level to represent an 'old' insecticide to which there is resistance.
-In both cases resistance to the new insecticide develops slower when the 'old' insecticide continues to be used (mixture) than when it doesn't (sequence).
-
-#curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
-runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_restoration_ins1=0.23/0.73, rr_restoration_ins2 = 0.43/1, addCombinedStrategy=FALSE )
-
-#curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
-runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_restoration_ins1=0.23/0.73, rr_restoration_ins2 = 0.43/1, addCombinedStrategy=FALSE )
-
-```
-
-## Fig.
-x12 How sequential can be better than mixture.
+```
+
+## Fig11  How sequential can be better than mixture.
 When the time taken for resistance to the first insecticide to be reached is greater than the difference in time taken for the second resistance to be reached in the presence and absence of the first insecticide.
 
-```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5}
+```{r Fig11, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5}
 

### 2016-10-18 inst/documents/paper1_results_figs_rr.Rmd
copying a fig from slimmed not used in the paper

@@
 ```
+
+```{r, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5} 
+#eval=FALSE so this one not displayed
+
+## Fig.
+x8 Introducing a new insecticide when there is already resistance to an existing one.
+This takes Curtis Fig 2 and sets the starting frequency for each allele in turn to a higher level to represent an 'old' insecticide to which there is resistance.
+In both cases resistance to the new insecticide develops slower when the 'old' insecticide continues to be used (mixture) than when it doesn't (sequence).
+
+#curtis_f2 with P_1 changed from 0.01 to 0.45 : mix slower than seq
+runcurtis_f2( P_1 = 0.49 , P_2 = 0.01 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_restoration_ins1=0.23/0.73, rr_restoration_ins2 = 0.43/1, addCombinedStrategy=FALSE )
+
+#curtis_f2 with P_2 changed from 0.01 to 0.45 : mix slower than seq
+runcurtis_f2( P_1 = 0.01 , P_2 = 0.49 , h.RS1_A0 = 0.17 , h.RS2_0B = 0.0016 , exposure = 0.9 , phi.SS1_A0 = 0.73 , phi.SS2_0B = 1 , rr_restoration_ins1=0.23/0.73, rr_restoration_ins2 = 0.43/1, addCombinedStrategy=FALSE )
+
+```
+
+## Fig. x5.1 modified for mix2 vs seq 
+NOT USED IN PAPER1
+fewer runs where sequential gives slower resistance
+
+```{r, eval=TRUE, echo=FALSE, fig.height=6}
+
+  ggSubset <- dif_mix2_seq
+  #if mix-seq >1 resistance took longer for mix so mix is better
+  ggSubset$mix2_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 > 0, 1, 0 )
+  ggSubset$seq_better_mix20.5 <- ifelse( ggSubset$mix2_minus_seq0.5 < 0, 1, 0 )
+  ggSubset$abs_diff <- abs(ggSubset$mix2_minus_seq0.5)
+
+  #tricky
+  #be a bit careful what to do with zeros where results are same for both strategies
+  #I use these better vars to facet the plots
+  #may want the zeroes in both facets
+  #to get that I can rbind the zeros from one onto the other BEWARE this may screw up existing plots
+  df_same <- ggSubset[which(ggSubset$abs_diff==0),]
+  #set these to 1 (they were 0)
+  df_same$mix2_better_seq0.5 <- 1
+  df_same$seq_better_mix20.5 <- 1
+  #rbind them back onto other dataset
+  ggSubset <- rbind( ggSubset, df_same)
+
+  #add an option for 20% better, put the not in to get facets in correct order in ggplot later
+  ggSubset$mix2_not20pc_better_seq0.5 <- ifelse( ggSubset$mix2_minus_seq0.5 / ggSubset$gen_cP0.5seq <= 0.2, 1, 0 )
+
+    #find unequal runs
+    unequal <- ggSubset[ggSubset$resist_start_lo_div_hi < 0.01,]
+    #nrow(ggSubset)
+    #nrow(unequal)
+    #2721 out of the 6189 remaining runs
+
+    #in how many runs was seq better than mix
+    #sum(unequal$seq_better_mix20.5)
+    #311
+
+    library(viridis)
+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_diff))+
+
+             facet_grid( ~ seq_better_mix20.5, labeller=as_labeller( c('0'='mixture\ntime to resistance\nlonger','1'='sequential\ntime to resistance\nlonger'))) +
+
+             #YlGnBu looks really good except that the yellow points are very faint
+             #scale_colour_distiller(palette='YlGnBu', direction=1) +
+             #option="plasma" looks good too
+             scale_color_viridis(begin=1, end=0, option="viridis", 
+                                 guide=guide_colourbar(title="difference in \ngenerations to reach\nresistance threshold", direction='horizontal')) + 
+             theme_bw() +
+             theme(legend.position = "bottom") +
+             #theme_minimal() +
+             #add point for Curtis inputs
+             geom_point(data = curtisInputs, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='black', shape=3) +
+             #failing to add legend for curtis point
+             #scale_fill_manual(name = "Curtis Fig.2", values = 'red', labels = 'test') +
+             #scale_size_manual(name = "Curtis Fig.2", values="test") +
+             #scale_shape_manual(name = "Curtis Fig.2", values="test") +             
+             ylab('effectiveness of insecticides added') +
+             xlab('exposure to both insecticides') + 
+             #main points for the plot
+             geom_point(alpha=1, size=0.5) + 
+             #!!!NOW can I add a lm fitted to the 0 points abs_diff
+             geom_smooth(data=ggSubset[which(ggSubset$abs_diff==0),], method = "lm", se = FALSE, colour='red', lwd=0.5) +
+             #can I subtley highlight the zero points which are repeated in each plot ?
+             geom_point(data=ggSubset[which(ggSubset$abs_diff==0),], size=0.8, colour='red', shape=2, fill='white') +
+
+             #now add the unequal points
+             #showed that they were on both sides
+             #geom_point(data = unequal, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2), colour='red', shape=2) +
+
+             #add text for Curtis point
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_ins1 + effectiveness_ins2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='black')
+           )
+
+  # doing the lm outside of ggplot    
+  x1=df_same$exposure 
+  y1=df_same$effectiveness_ins1 + df_same$effectiveness_ins2
+
+  l_mod <- lm(y1 ~ x1)
+  #l_mod$coefficients
+  #(Intercept)          x1 
+  #1.019330    0.560508
+
+  #approximation
+  #x2 <- seq(0,1,0.1)
+  #y2 <- 0.6*x2+0.98
+  #plot(x2,y2)
+
+```

### 2016-10-24 inst/documents/effectiveness_exposure_difference.Rmd
tweaking figure options for paper1 & 2

@@
 
-plot_fit_rs(df_base_, column='f', title='base', ylim=c(0.4,0.8))
+plot_fit_rs(df_base_, column='f', title='base 0.5', ylim=c(0.4,0.8))
 
-plot_fit_rs(df_ef_, column='f', title='effectiveness', ylim=c(0.4,0.8), yblank=TRUE)
+plot_fit_rs(df_ef_, column='f', title='effectivenessI1 0.8', ylim=c(0.4,0.8), yblank=TRUE)
 
-plot_fit_rs(df_ex_, column='f', title='exposure', ylim=c(0.4,0.8), yblank=TRUE)
+plot_fit_rs(df_ex_, column='f', title='exposureI1 0.8', ylim=c(0.4,0.8), yblank=TRUE)
 
@@
 plot_fit_rs(df_ex, column='f', title='exposure', ylim=c(0.5,1), yblank=TRUE)
+
+#can use cowplot to size 1st plot with y axis differently
 
@@
 In contrast increasing the effectiveness of one insecticide in a mixture decreases the slection pressure for resistance to the other.
+
+```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=5, fig.show='hold', fig.align='left', fig.cap="test A.  B. C. D."}
+
+#todo to get grid fig working.
+#think about implementing in ggplot2 to make my life easier later ?
+#although I'm not sure it would ?
+#allow x axis labels and title to be removed
+#add axis titles outside of the plot region
+#remove curve legend on plots and put it somewhere else maybe on bottom
+#? allow curves to be thicker
+
+#test of rearranging figures on a grid
+#eff1  0.5  0.8
+#exp
+#0.5    1   2
+#0.8    3   4
+
+layout(matrix(1:4,2,2, byrow = TRUE))
+#effectiveness on mixture
+
+#plot margins
+oldpar <- par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#base scenario
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#effectiveness1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+#plot margins
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#exposure 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1 & expos 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+par(oldpar)
+
+```

### 2016-10-24 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
tweaking figure options for paper1 & 2

@@
 #Fig1 
-```{r Fig1, eval=TRUE, echo=FALSE, fig.height=7.5, fig.width=5, fig.show='hold', fig.align='left'}
+```{r Fig1, eval=TRUE, echo=FALSE, fig.height=7.5, fig.width=6, fig.show='hold'}
+#, fig.align='left'}
+
+#PLOS fig dimensions
+#Maximum width 7.5 or 5.2 to fit in text column.
+#Maximum height 8.75
 
@@
        geom_line()
+       #annotate('text',label='A)',x=-10, y=1.1)
 
@@
              panel.grid = element_blank())
-
-plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(0.7,1))
+       #draw_plot_label('B)') +
+
+plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(0.7,1), labels='AUTO')
+
+#draw_figure_label('B)', pos='top.left')
 
@@
 # Fig2 replicating Curtis
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5.5, fig.keep='last'}
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6, fig.keep='last'}
 
@@
   layout(t(1:2))
+  par(mar=c(0.4,2.9,0,0.4)) #bltr default is c(5, 4, 4, 2)
 
@@
   plotcurtis_f1( listOut$results[[1]], listOut$results[[2]], 1, 2 )
+  #side=1=bottom, 2=left, 3=top, 4=right, adj=0 left justify, lines move outwards
+  mtext('A)',side=3, adj=0, line=1)
 
@@
   plotcurtis_f2( listOut$results[[3]], listOut$results[[1]], listOut$results[[2]], 1, 2, 3 )
+  #side=1=bottom, 2=left, 3=top, 4=right, adj=0 left justify, lines move outwards
+  mtext('B)',side=3, adj=0, line=1)
 
@@
   #plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]],ncol=1, rel_heights=c(1,1,1,1.8))  
-  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6))
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1.6), labels='AUTO')
 
@@
 
-  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1))
+  plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],ncol=1, rel_heights=c(1,1,1), labels='AUTO', vjust=1)
 
@@
 
-  plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(1,1.6))
+  plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(1,1.6), labels='AUTO')
 

### 2016-10-24 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
tweaking figure options for paper1 & 2

@@
 
+#base scenario
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
@@
 
+#effectiveness1 0.8
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", ylabs = FALSE, main='B.', maxX = 120 )

### 2016-10-24 R/plotcurtis_f1.r

@@ -23,10 +23,17 @@ plotcurtis_f1 <- function ( nrelaxmat, relaxmat, gencol, r1col ){
   nrelax <- log10( 100 * nrelaxmat[,r1col] )			# only prints results from males, m = f, A[R] = B[R], so only prints one col
   gens <- nrelaxmat[,gencol]	
   
-  par(pty="s") 			
+  #tcl for smaller ticks (def=-0.5)
+  #mgp margin line for axis title, labels and line.
+  #mgp[1] affects title whereas mgp[2:3] affect axis. default is c(3, 1, 0).
+  par(pty="s", tcl=-0.2, mgp=c(2.5, 0.5, 0) ) #, plt=c(0.1,0.8,0.1,0.8)) 
+  
   plot( 0, 0, type="n", axes=F,						## Blank square 1,1 plot
         xlim=c(1,(max(nrelaxmat[,gencol]))), ylim=c((min(fl)),(max(fl))),
-        xlab="Generation", ylab="Allele Frequency - (A[R]= B[R])", main="")
+        xlab="", ylab="", main="")
+  
+  #to put titles closer to axes
+  title(xlab="Generation", ylab="Allele Frequency",line=2)
   
   axis( side=1, at=c(0,2,4,6,8,10,12,14,16,18), labels=c(0,2,4,6,8,10,12,14,16,18), tick=T )
   axis( side=2, at=fl, labels=labs, tick=T )
@@ -46,7 +53,7 @@ plotcurtis_f1 <- function ( nrelaxmat, relaxmat, gencol, r1col ){
   
   pos <- log10( 4.5 )
   legend( 7, pos , legend=c("selection by mixture", "selection relaxed"), 
-          col=c("black", "darkviolet"), pch=c(16,16), bty="n" )
+          col=c("black", "darkviolet"),lty=c(1,0), pch=c(NA,16), bty="n", cex=0.7 )
   
   box()
   

### 2016-10-24 R/plotcurtis_f2.r

@@ -17,12 +17,20 @@
 plotcurtis_f2 <- function( combmat, bmat, amat, gencol, r1col, r2col ){
   
   f <- c(1:100)				## for y axis 
-  fl <- log10(f)				
-  par(pty="s") 
+  fl <- log10(f)
+  
+  #tcl for smaller ticks (def=-0.5)
+  #mgp margin line for axis title, labels and line.
+  #mgp[1] affects title whereas mgp[2:3] affect axis. default is c(3, 1, 0).
+  par(pty="s", tcl=-0.2, mgp=c(2.5, 0.5, 0) ) #, plt=c(0.1,0.8,0.1,0.8)) 
+  
   # blank plot with x axis max as max_gen for combination scenario, and y axis max as log10(100) (for 100%)			
   plot( 0, 0, type="n", axes=F,						## Blank square 1,1 plot
         xlim=c(1,(max(combmat[,gencol]))), ylim=c(0,(max(fl))),
-        xlab="Generation", ylab="Allele Frequency", main="")
+        xlab="", ylab="", main="")
+  
+  #to put titles closer to axes
+  title(xlab="Generation", ylab="Allele Frequency",line=2)
   
   # plotting combination scenario (in combmat)
   # set frequencies in vectors as log10 percentages from frequencies
@@ -63,12 +71,15 @@ plotcurtis_f2 <- function( combmat, bmat, amat, gencol, r1col, r2col ){
   
   abline( v = 31, col="black" )
   
-  yval <- log10( 2 )
-  legend( 105, yval, legend=c("Sequential", "Combination","R at Locus 1 (DDT)", "R at Locus 2(HCH)"), 
-          col=c("black","black","red", "darkblue"), lty=c(2,1,1,1), bty="n", cex=0.7 )
+  yval <- log10( 3 )
+  # legend( 105, yval, legend=c("Sequence", "Mixture","DDT", "HCH"), 
+  #         col=c("black","black","red", "darkblue"), lty=c(2,1,1,1), bty="n", cex=0.6 )
+  legend( 105, yval, legend=c("HCH alone", "HCH in mix", "DDT alone", "DDT in mix"), 
+          col=c("blue","blue","red","red"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.6 )    
+  
   
   yval <- log10( 3 )			
-  text(27, yval, "Switch to DDT", srt=90, cex=0.75) 			
[…]

### 2016-10-24 R/plotcurtis_f2_generic.r

@@ -17,6 +17,7 @@
 #' @param xlab xlab
 #' @param ylab ylab
 #' @param ylabs whether to add tick labels
+#' @param xlabs whether to add tick labels
 #' @param yticks tick positions
 #' @param cex.axis scaling for x&y tick labs 
 #' @param addLegend whether to add a legend inside plot
@@ -43,6 +44,7 @@ plotcurtis_f2_generic <- function( combmat, bmat, amat, gencol=1, r1col=2, r2col
                                    ylabs = TRUE,
                                    yticks = c(0.1,1,5,10,50,100),
                                    xlab = "Generation",
+                                   xlabs = TRUE,
                                    cex.axis = 0.8,
                                    addLegend = TRUE,
                                    vlines = 'grey95',
@@ -75,22 +77,27 @@ plotcurtis_f2_generic <- function( combmat, bmat, amat, gencol=1, r1col=2, r2col
   #f <- c(1:100)				## for y axis 
   fl <- log10(f)
   
-  #28/6/16 disabled this, then put it back
-  par(pty="s") #to create a square plot region
+  #pty="s" #to create a square plot region
+  #tcl for smaller ticks (def=-0.5)
+  #mgp margin line for axis title, labels and line.
+  #mgp[1] affects title whereas mgp[2:3] affect axis. default is c(3, 1, 0).
+  par(pty="s", tcl=-0.2, mgp=c(2.5, 0.5, 0) ) #, plt=c(0.1,0.8,0.1,0.8)) 
   
   # blank plot with x axis max as max_gen for combination scenario, and y axis max as log10(100) (for 100%)			
   plot( 0, 0, type="n", axes=F,						## Blank square 1,1 plot
         #xlim=c(1,maxGens), ylim=c(0,(max(fl))),
         xlim=c(1,maxX), ylim=c(min(fl),(max(fl))),
         #xlim=c(1,(max(combmat[,gencol]))), ylim=c(0,(max(fl))),
-        xlab=xlab, ylab=ylab ) #, main=main)
+        xlab="", ylab="" ) #, main=main)
   
   #to left justify title
   title(main,adj=0)
+  #to put titles closer to axes
+  title(xlab=xlab, ylab=ylab,line=2)
   
   # add axis labels
   # default x axis
[…]

### 2016-10-24 R/runcurtis_f2.r

@@ -39,6 +39,7 @@
 #' @param xlab xlab
 #' @param ylab ylab
 #' @param ylabs whether to add tick labels
+#' @param xlabs whether to add tick labels
 #' @param yticks tick positions
 #' @param cex.axis scaling for x&y tick labs 
 #' @param addLegend whether to add a legend inside plot
@@ -88,6 +89,7 @@ runcurtis_f2 <- function( max_gen = 500,
                           ylabs = TRUE,
                           yticks = c(0.1,1,5,10,50,100),
                           xlab = "Generation",
+                          xlabs = TRUE,
                           cex.axis = 0.8,
                           addLegend = TRUE,
                           main = "",
@@ -223,6 +225,7 @@ runcurtis_f2 <- function( max_gen = 500,
                          addStrategyLabels = addStrategyLabels,
                          strategyLabels = strategyLabels,
                          xlab = xlab,
+                         xlabs = xlabs,
                          ylab = ylab,
                          ylabs = ylabs,
                          yticks = yticks,

### 2016-10-26 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
tweaking pap1 figs

@@
 
-# Fig2 replicating Curtis
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6, fig.keep='last'}
+![curtis fig1](pap1_curtis1.png)
+
+# Fig2 replicating Curtis, and now adding in copies of Curtis plots
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8.5, fig.width=6, fig.keep='last'}
+
+  #Beths code for these figures taken from resistanceMaster
+
+  #set layout for 4 plots, A&C Curtis, B&D new ones
+  layout(matrix(1:4,2,2,byrow=FALSE))
+
+  #add in images of curtis plots here
+  #to get into the layout I need to do from R
+  #may be ugly
+  par(mar=c(0.4,2.9,0,0.4)) #bltr default is c(5, 4, 4, 2)
+
+  image(system.file("documents","pap1_curtis1.png", package="resistance"))
+
+  image(system.file("documents","pap1_curtis2.png", package="resistance"))
+
+  par(mar=c(0.4,2.9,0,0.4)) #bltr default is c(5, 4, 4, 2)
+
+  #2B
+  calibration <- 1011
+  input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
+  listOut <- runModel2( input, calibration )
+  #we submitted the LD one but should have done the Fig1
+  #plotcurtis_ld( listOut$results[[1]], listOut$results[[2]], 1, 4 )
+  plotcurtis_f1( listOut$results[[1]], listOut$results[[2]], 1, 2 )
+  #side=1=bottom, 2=left, 3=top, 4=right, adj=0 left justify, lines move outwards
+  mtext('B)',side=3, adj=0, line=1)
+
+  #2D
+  calibration <- 1012
+  input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
+  listOut <- runModel2( input, calibration )  
+  plotcurtis_f2( listOut$results[[3]], listOut$results[[1]], listOut$results[[2]], 1, 2, 3 )
+  #side=1=bottom, 2=left, 3=top, 4=right, adj=0 left justify, lines move outwards
+  mtext('D)',side=3, adj=0, line=1)
+
+```
+
+```{r Fig2_justNew, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6, fig.keep='last'}
 

### 2016-10-26 R/fit_time_genotype.r
starting fitness variance calc

@@ -17,7 +17,7 @@
 
 fit_time_genotype <- function ( genotype = NULL,
                                 a_fitgen = NULL,
-                                plot = FALSE )
+                                plot = TRUE )
 {
   #I'm not sure if these default args will work
   # to allow this function to be called with no args
@@ -55,7 +55,7 @@ fit_time_genotype <- function ( genotype = NULL,
                                    sex=c('m','f'), 
                                    genotype = c("SS1SS2", "SS1RS2", "SS1RR2", 
                                                 "RS1SS2", "RS1RS2", "RS1RR2",
-                                                "RR1SS2", "RR1RS2", "RR1RR2") )
+                                                "RR1SS2", "RR1RS2", "RR1RR2", "variance") )
   
   #for each generation 
   for (gen_num in 1:max_gen)
@@ -70,6 +70,8 @@ fit_time_genotype <- function ( genotype = NULL,
           # have to do cis/trans specially
           if ( locus1=='RS1' & locus2=='RS2' )
           {
+            #TODO check that it's ok to ad this and not to calc the mean
+            #ok adding the freqs but not sure about the fitness component
             a_fit_time_gen[gen_num, sex, paste0(locus1,locus2)] <- 
               genotype[gen_num, 'RS1RS2_cis'] * a_fitgen[sex,locus1,locus2] +
               genotype[gen_num, 'RS1RS2_trans'] * a_fitgen[sex,locus1,locus2]      
@@ -82,6 +84,66 @@ fit_time_genotype <- function ( genotype = NULL,
       }
     }
   } #end gen_num loop
+
+  # to calculate variance in genotype fitness over time
+  # using a loop because I'm tired !
+  # BUT ian says the variance wants to be calcualted in a different way, not just from the 9 numbers
+  # instead it needs to take into account that there are very few of the low frequency ones 
+  # for (gen_num in 1:max_gen)
+  # {
+  #   for( sex in dimnames(a_fitgen)$sex)
+  #   {
+  #     a_fit_time_gen[gen_num,sex,'variance'] <- var(a_fit_time_gen[gen_num,sex,]) 
+  #   }
+  # }
+ 
+  # 25/10/16 Ians new variance calc
+  # similar (but not same) to what hapens in selection function
[…]

### 2016-10-27 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
tweak figs

--- 2016-10-24 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
@@
 
-# Fig2 replicating Curtis
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6, fig.keep='last'}
+![curtis fig1](pap1_curtis1.png)
+
+# Fig2 replicating Curtis, and now adding in copies of Curtis plots
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8.5, fig.width=6, fig.keep='last'}
+
+  #Beths code for these figures taken from resistanceMaster
+
+  #set layout for 4 plots, A&C Curtis, B&D new ones
+  layout(matrix(1:4,2,2,byrow=FALSE))
+
+  #add in images of curtis plots here
+  #to get into the layout I need to do from R
+  #may be ugly
+  #par(mar=c(0,0,0,0)) #bltr default is c(5, 4, 4, 2)
+  #pty="s" #to create a square plot region
+  par(pty="s",mar=c(0.4,2.9,0,0.4)) #bltr default is c(5, 4, 4, 2)
+
+  library(png)
+  f1png <- readPNG(system.file("documents","pap1_curtis1.png", package="resistance"))
+  plot( 0, 0, type="n", axes=F, xlim=c(0,1), ylim=c(0,1), xlab="", ylab="")
+  rasterImage(f1png, xleft=0, ybottom=0, xright=1, ytop=1, interpolate = FALSE)
+  mtext('A)',side=3, adj=0, line=1)
+
+  f2png <- readPNG(system.file("documents","pap1_curtis2.png", package="resistance"))
+  plot( 0, 0, type="n", axes=F, xlim=c(0,1), ylim=c(0,1), xlab="", ylab="")
+  rasterImage(f2png, xleft=0, ybottom=0, xright=1, ytop=1, interpolate = FALSE)
+  mtext('C)',side=3, adj=0, line=1)
+
+  par(mar=c(0.4,2.9,0,0.4)) #bltr default is c(5, 4, 4, 2)
+
+  #2B
+  calibration <- 1011
+  input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
+  listOut <- runModel2( input=input )
+  #we submitted the LD one but should have done the Fig1
+  #plotcurtis_ld( listOut$results[[1]], listOut$results[[2]], 1, 4 )
+  plotcurtis_f1( listOut$results[[1]], listOut$results[[2]], 1, 2 )
+  #side=1=bottom, 2=left, 3=top, 4=right, adj=0 left justify, lines move outwards
+  mtext('B)',side=3, adj=0, line=1)
+
+  #2D
+  calibration <- 1012
+  input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
+  listOut <- runModel2( input=input )  
+  plotcurtis_f2( listOut$results[[3]], listOut$results[[1]], listOut$results[[2]], 1, 2, 3 )
+  #side=1=bottom, 2=left, 3=top, 4=right, adj=0 left justify, lines move outwards
+  mtext('D)',side=3, adj=0, line=1)
+
+```
+
+```{r Fig2_justNew, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=3.5, fig.width=6, fig.keep='last'}
 
@@
   input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
-  listOut <- runModel2( input, calibration )
+  listOut <- runModel2( input=input )
   #we submitted the LD one but should have done the Fig1
@@
   input <- createInputMatrix( params.csv=FALSE, calibration=calibration )
-  listOut <- runModel2( input, calibration )  
+  listOut <- runModel2( input=input )  
   plotcurtis_f2( listOut$results[[3]], listOut$results[[1]], listOut$results[[2]], 1, 2, 3 )

### 2016-10-27 R/plot_ld_gpirm.r
started function to recreate gpirm LD plots from p122

@@ -1,0 +1,119 @@
+#' try to plot genotype frequencies to show linkage disequilibrium
+#' 
+#' as done in GPIRM p122
+
+#' @param genotype genotype frequencies by generation
+#' @param gen_num which generation to plot for
+#' 
+#' @examples 
+#' input <- setInputOneScenario()
+#' listOut <- runModel2(input)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=5)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=6)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=7)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=8)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=9)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=10)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=11)
+#' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen_num=12)
+#' 
+#' 
+#' @return dataframe of plotted fitness values and colours
+#' @export
+
+plot_ld_gpirm <- function ( genotype, gen_num = 1  ){
+ 
+  
+  #could use listOut$genotype but it has cis & trans in ...
+  #AHA 2nd problem was that genotype includes cis & trans
+  #quick fix
+  #TODO make a safer version of this !!
+  gen2 <- genotype
+  gen2[,'RS1RS2_cis'] <- gen2[,'RS1RS2_cis'] + gen2[,'RS1RS2_trans']
+  colnames(gen2)[ which(colnames(gen2)=='RS1RS2_cis')] <- 'RS1RS2'
+  gen2 <- gen2[,- which(colnames(gen2)=='RS1RS2_trans')]
+  
+  tp <- gen2[gen_num,]
+  
+  #rectangle  boundaries (see notebook)
+  #first assuming complete dominance
+  #TODO see notebook for way of including dominance 
+  p1 <- tp['RR1SS2'] + tp['RS1SS2']
+  p2 <- tp['RR1RR2'] + tp['RS1RR2'] +
+        tp['RR1RS2'] + tp['RS1RS2']
+  p3 <- tp['SS1RR2'] + tp['SS1RS2']
+  p4 <- p2
+  
+  #remove blank borders
[…]

### 2016-11-11 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
tweaking paper1 figs

@@
 
-## Figures for paper 1 : version 9 : looking at producing Tiffs for submission
-
+## Figures for paper 1 : version 9 : produces Tiffs for submission (fails to produce pdf due to tiff options set in first chunk)
 ## using rr_restoration instead of selection coefficient
-
-### automagically generated by R from paper1_results_figs_slimmed_50_rr.Rmd
-
-## Notable results
-
-1. Time-to-resistance is lowest for single use and for the first insecticide in a mixture. Time-to-resistance is highest for the 2nd insecticide in a mixture.
-1.
-In between these extremes, there is less difference between the alternative strategies of a) sequential and b) adaptive mixture, switching to sole use.
-The latter takes slightly longer to reach the threshold.
-1. Exposure and dominance have the greatest on time-to-resistance as shown in PRCC analyses.
-1.
-Effectiveness of the insecticides is the one input that can change the average relative positioning of the strategies a) sequential and b) adaptive mixture.
-Below an effectiveness of ~0.55 to 0.6 sequential takes longer to reach the resistance threshold, and above this the adaptive mixture takes longer.
-1. Exposure and effectiveness come out as the most important effect on the difference between the sequential and adpative mixture strategies.
-1.
-Exposure and effectiveness can be used to predict most of whether resistance will arise slower under the sequential or adaptive mixture strategy, either by a simple plot or using classification trees.
-1.
-The implication of additional operational costs of adaptive mixtures can be investigated.
-If a requirement of a 20% improvement in resistance is imposed then more runs favour sequential than mixture.
 
@@
        ylim(0,100) +
+       coord_cartesian(expand = FALSE) + #so that axes fit data exactly
        ylab("Insecticide concentration") +
@@
   f1png <- readPNG(system.file("documents","pap1_curtis1.png", package="resistance"))
+  #both the ld and the rfreq plots (ld on its own means little with the way axes are labelled)
+  #f1png <- readPNG(system.file("documents","pap1_curtis1_both.png", package="resistance"))
   plot( 0, 0, type="n", axes=F, xlim=c(0,1), ylim=c(0,1), xlab="", ylab="")

### 2016-11-11 R/plotcurtis_ld.r

@@ -25,7 +25,7 @@ plotcurtis_ld <- function(nrelaxmat, relaxmat, gencol, ldcol){
   
   plot( 0, 0, type="n", axes=F,						## Blank square 1,1 plot
         xlim=c(1,(max(nrelaxmat[,gencol]))), ylim=c(ymin,ymax),
-        xlab="Generation", ylab="Index of Linkage Disequilibrium - D", main="Linkage Disequilibrium")
+        xlab="Generation", ylab="Index of Linkage Disequilibrium - D", main="")#"Linkage Disequilibrium")
   
   lines( gens, res )		#plot log10 LD against generation
   lines( gens, rel )
@@ -33,7 +33,7 @@ plotcurtis_ld <- function(nrelaxmat, relaxmat, gencol, ldcol){
   points( c(3:12), rel[3:12], col="darkviolet", pch=16 )		# add purple points on relaxed line to indicate relaxed generations
   points( c(15:18), rel[15:18], col="darkviolet", pch=16 )
   
-  axis( side=1, at=c(0,2,4,6,8,10,12,14,16,18), labels=c(0,2,4,6,8,10,12,14,16,18), tick=T )
+  axis( side=1, at=c(0,2,4,6,8,10,12,14,16,18), labels=c(0,2,4,6,8,10,12,14,16,18), tick=T, cex=0.7 )
   
   labs <- c(0.000001,0.00001,0.0001,0.001,0.01)
   labpos <- log10(labs)
@@ -41,8 +41,8 @@ plotcurtis_ld <- function(nrelaxmat, relaxmat, gencol, ldcol){
   axis( side=2, at=labpos, labels=labs, tick=T )
   
   pos <- log10( 0.01 )
-  legend( 9, pos , legend=c("Linkage disequilbrium","With relaxed selection"), 
-          col=c("black", "darkviolet"), pch=c(16,16), bty="n" )
+  legend( 7, pos , legend=c("Linkage disequilbrium","With relaxed selection"), 
+          col=c("black", "darkviolet"),lty=c(1,0), pch=c(NA,16), bty="n", cex=0.7 )
   
   box()
   

### 2016-11-23 Matt/AttemptPlottingRotations.Rmd

---
title: "AttemptPlottingRotations"
output: pdf_document
author: "Matt Powney"
date: "`r Sys.Date()`"
---

#The following code generates and plots the rotation resistances for Curtis figure 2.

```{r,echo=FALSE,results='hide'}
findResistancePointsRotation <- function( rotation_generations = 10,
                                          listOutI1,
                                          listOutI2,
                                          locus,
                                          criticalPoints = c(0.1,0.25,0.5)
){

  #set the number of scenarios
  numScenarios <- length(listOutI1$results)

  #create matrix to receive the results, name rows including gen_cP for generations to reach criticalPoints.
  resistGens <- matrix( nrow=length(criticalPoints), ncol=numScenarios, dimnames=list(paste0('gen_cP',criticalPoints) ) )

  #for each scenario
  for(scenarioNum in 1:numScenarios)
  {
    #for each critical point
    for(criticalPointNum in seq_along(criticalPoints) )
    {
      criticalPoint <- criticalPoints[criticalPointNum]
      #cat("criticalPoint ", criticalPoint,"\n")

      #getting results matrix out of the list for the separate insecticides
      resultsI1 <- listOutI1$results[[scenarioNum]]
      resultsI2 <- listOutI2$results[[scenarioNum]]

      #extracting the resistances for insecticides 1 and 2
      resistancesI1 <-    rowMeans(cbind(resultsI1[,'m.R1'],
                                         resultsI1[,'f.R1'] ))
      resistancesI2 <-    rowMeans(cbind(resultsI2[,'m.R2'],
                                         resultsI2[,'f.R2'] ))

      #Isolating generations where resistance hasn't been met for I1 and I2.
      gensI1 <- which( resistancesI1 < criticalPoint)
      gensI2 <- which( resistancesI2 < criticalPoint)

      #Connecting first detectable resistance generations.
      gensI1=c(gensI1, max(gensI1)+1)
      gensI2=c(gensI2, max(gensI2)+1)

      #Calculate total number of generations until resistance for rotation.
      overallgens = length(gensI1) + length(gensI2)

      #The following code calculates the final number of generations before resistance for each insecticide individually.

      roundsI1=(length(gensI1)%/%rotation_generations)
      roundsI2=(length(gensI2)%/%rotation_generations)

      leftover=(length(gensI2)%%rotation_generations)

      if (roundsI1 > roundsI2){

        I1gen=overallgens

        if (leftover==0){

          I2gen=(2*roundsI2)*rotation_generations + leftover
        }

        if (leftover!=0){

          I2gen=(2*roundsI2 + 1)*rotation_generations + leftover
        }

      }

      else 
      {
        I2gen=overallgens

        I1gen=(2*(roundsI1)*rotation_generations + leftover)

      }

      #Calculate total number of generations until resistance for rotation.
      if (locus == "both")
      {
        if (I1gen>I2gen){
          resistGens[criticalPointNum,scenarioNum] <- I1gen
        }

        if (I1gen<I2gen){
          resistGens[criticalPointNum,scenarioNum] <- I2gen
        }

      }

      #Report number of generations for insecticide 1 and insecticide 2.
      else if (locus ==1)
      {
        resistGens[criticalPointNum,scenarioNum] <- I1gen

      }

      else if (locus == 2)
      {
        resistGens[criticalPointNum,scenarioNum] <- I2gen

      }

    }
  }

  #replace any NAs with 999 to show that resistance not reached
  resistGens[is.na(resistGens)] <- 999

  return(resistGens)

}

```

```{r,echo=FALSE,results='hide'}
#To mirror the inputs for Curtis Figure 2.

exp1 = NULL
exp2 = NULL
max_gen = 500
P_1 = 0.01
P_2 = 0.01
recomb_rate = 0.5
exposure = 0.9
phi.SS1_a0 = 0
phi.SS1_A0 = 0.73
phi.SS2_0b = 0
phi.SS2_0B = 1
W.SS1_00 = 1 
W.SS2_00 = 1
h.RS1_00 = 0
h.RS1_a0 = 0
h.RS1_A0 = 0.17
h.RS2_00 = 0
h.RS2_0b = 0
h.RS2_0B = 0.0016
s.RR1_a0 = 0 
s.RR1_A0 = NULL
s.RR2_0b = 0
s.RR2_0B = NULL
rr_restoration_ins1 = 0.23/0.73
rr_restoration_ins2 = 0.43/1
z.RR1_00 = 0
z.RR2_00 = 0
sexLinked = 0
male_exposure_prop = 1
correct_mix_deploy = 1

#Set exposure and inputs for insecticide 1
a <- setExposure(exposure = exposure, exp1 = exp1, exp2 = exp2, 
                 insecticideUsed = "insecticide1", male_exposure_prop = male_exposure_prop, 
                 correct_mix_deploy = correct_mix_deploy)

inputI1 <- setInputOneScenario(a = a, max_gen = max_gen, 
                               P_1 = P_1, P_2 = P_2, recomb_rate = recomb_rate, phi.SS1_a0 = phi.SS1_a0, 
                               phi.SS1_A0 = phi.SS1_A0, phi.SS2_0b = phi.SS2_0b, phi.SS2_0B = phi.SS2_0B, 
                               W.SS1_00 = W.SS1_00, W.SS2_00 = W.SS2_00, h.RS1_00 = h.RS1_00, 
                               h.RS1_a0 = h.RS1_a0, h.RS1_A0 = h.RS1_A0, h.RS2_00 = h.RS2_00, 
                               h.RS2_0b = h.RS2_0b, h.RS2_0B = h.RS2_0B, s.RR1_a0 = s.RR1_a0, 
                               s.RR1_A0 = s.RR1_A0, s.RR2_0b = s.RR2_0b, s.RR2_0B = s.RR2_0B, 
                               rr_restoration_ins1 = rr_restoration_ins1, rr_restoration_ins2 = rr_restoration_ins2, 
                               z.RR1_00 = z.RR1_00, z.RR2_00 = z.RR2_00, sexLinked = sexLinked, 
                               male_exposure_prop = male_exposure_prop, correct_mix_deploy = correct_mix_deploy)

#Set exposure and inputs for insecticide 1

a <- setExposure(exposure = exposure, exp1 = exp1, exp2 = exp2, 
                 insecticideUsed = "insecticide2", male_exposure_prop = male_exposure_prop, 
                 correct_mix_deploy = correct_mix_deploy)

inputI2 <- setInputOneScenario(a = a, max_gen = max_gen, 
                               P_1 = P_1, P_2 = P_2, recomb_rate = recomb_rate, phi.SS1_a0 = phi.SS1_a0, 
                               phi.SS1_A0 = phi.SS1_A0, phi.SS2_0b = phi.SS2_0b, phi.SS2_0B = phi.SS2_0B, 
                               W.SS1_00 = W.SS1_00, W.SS2_00 = W.SS2_00, h.RS1_00 = h.RS1_00, 
                               h.RS1_a0 = h.RS1_a0, h.RS1_A0 = h.RS1_A0, h.RS2_00 = h.RS2_00, 
                               h.RS2_0b = h.RS2_0b, h.RS2_0B = h.RS2_0B, s.RR1_a0 = s.RR1_a0, 
                               s.RR1_A0 = s.RR1_A0, s.RR2_0b = s.RR2_0b, s.RR2_0B = s.RR2_0B, 
                               rr_restoration_ins1 = rr_restoration_ins1, rr_restoration_ins2 = rr_restoration_ins2, 
                               z.RR1_00 = z.RR1_00, z.RR2_00 = z.RR2_00, sexLinked = sexLinked, 
                               male_exposure_prop = male_exposure_prop, correct_mix_deploy = correct_mix_deploy)

```

```{r,echo=FALSE,results='hide'}
input <- cbind(inputI1, inputI2)

#Run the code for sequential inputs as the population dynamics are required to calculate the rotation resistance levels.
listOut <- runModel2(input = input, produce.plots = FALSE)

#Set the critical point 0.5.
criticalPoint=0.5

#Come back to these for test runs
#listOutCurtI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment='curtis')
#listOutCurtI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment='curtis')

#listOutCurtI1 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide1', experiment='curtis')
#listOutCurtI2 <- sensiAnPaperPart( nScenarios, insecticideUsed = 'insecticide2', experiment='curtis')
#resultsI1 <- listOutCurtI1$results[[1]]
#resultsI2 <- listOutCurtI2$results[[1]]

#Extract the results for insecticides 1 and 2.
resultsI1 <- listOut$results[[1]]
resultsI2 <- listOut$results[[2]]

#extracting the resistances for insecticides 1 and 2
resistancesI1 <-    rowMeans(cbind(resultsI1[,'m.R1'],
                                   resultsI1[,'f.R1'] ))
resistancesI2 <-    rowMeans(cbind(resultsI2[,'m.R2'],
                                   resultsI2[,'f.R2'] ))

#Isolating generations where resistance hasn't been met for I1 and I2.
gensI1 <- which( resistancesI1 < criticalPoint)
gensI2 <- which( resistancesI2 < criticalPoint)

gensI1=c(gensI1,max(gensI1)+1)
gensI2=c(gensI2,max(gensI2)+1)

ins1=resistancesI1[gensI1]
ins2=resistancesI2[gensI2]

ins1
ins2

#Come back to the following to generate a function

#m=2

#rotation_generations=5

#for(i in 1:m){

#for (k in 1:rotation_generations)

#{
#fl[(2*(i-1)*rotation_generations)+k]=ins1[k]

#fl[rotation_generations*i+k]=ins1[rotation_generations*i]

#}
#}

#fl=rep(0,)

gens=seq(1,65, by=1)

i1fiverot=c(ins1[1:5],rep(ins1[6],5),ins1[7:11],rep(ins1[11],5),ins1[12:16],rep(ins1[16],5),ins1[17:21],rep(ins1[21],5),ins1[22:26],rep(ins1[26],5),ins1[27:31],rep(ins1[31],5),ins1[32:36])

i2fiverot=c(rep(ins2[1],5),ins2[2:6],rep(ins2[6],5),ins2[7:11],rep(ins2[11],5),ins2[12:16],rep(ins2[16],5),ins2[17:21],rep(ins2[21],5),ins2[22:26],rep(ins2[26],5),ins2[27:30],rep(ins2[30],6))

i1tenrot=c(ins1[1:10],rep(ins1[11],10),ins1[12:21],rep(ins1[21],10),ins1[22:31],rep(ins1[31],10),ins1[32:36])

i2tenrot=c(rep(ins2[1],10),ins2[2:11],rep(ins2[11],10),ins2[12:21],rep(ins2[21],10),ins2[22:30],rep(ins2[30],6))

i1twentyrot=c(ins1[1:20],rep(ins1[21],20),ins1[22:36],rep(ins1[36],10))

i2twentyrot=c(rep(ins2[1],20),ins2[2:21],rep(ins2[21],16),ins2[22:30])

```

## The following code makes a plot of the Curtis resistances along with rotations.

```{r, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=8}

#Use the run Curtis function to generate the plots for mixtures and sequential strategies.

runcurtis_f2()

#The following code crudely generates the resistance levels for rotations required for the ASTMH presentation dependent on the resistance_generations. IMPROVE THIS.

#For 5 generation rotations
lines( gens, log10(i1fiverot*100), col=adjustcolor("red"), lty=4,lwd=3 )
lines( gens, log10(i2fiverot*100), col=adjustcolor("blue"), lty=4,lwd=3 )

#For 10 generation rotations
lines( gens, log10(i1tenrot*100), col=adjustcolor("red"), lty=4, lwd=3 )
lines( gens, log10(i2tenrot*100), col=adjustcolor("blue"), lty=4, lwd=3 )

#For 20 generation rotations
lines( gens, log10(i1twentyrot*100), col=adjustcolor("red"), lty=4, lwd=3 )
lines( gens, log10(i2twentyrot*100), col=adjustcolor("blue"), lty=4, lwd=3 )

```

### 2016-11-25 Matt/DefaultValuesVarianceInFitness.Rmd

---
title: "DefaultValuesVarianceInFitness"
output: pdf_document
author: "Matt Powney"
date: "`r Sys.Date()`"
---

#The code for the Variance in Fitness when LD does not equal zero for the default inputs of linkage disequilibrium. These are for low, medium and high levels of "insect killing" using both mosaic and mixture strategies.

```{r}

#The code for setting the default inputs of low, medium and high insecticide killing.
male_exposure_prop=0.5
exposure=0.8
strategy="mixture"

a <- setExposure(exposure=exposure, insecticideUsed=strategy,
                 male_exposure_prop=male_exposure_prop, correct_mix_deploy=1)

defaultlowmix=setInputOneScenario(P_1=0.001, P_2=0.01, a=a, exposure = 0.8, male_exposure_prop=1, phi.SS1_A0=0.3, phi.SS2_0B=0.5, h.RS1_A0	=0.4, h.RS2_0B = 0.8,rr_restoration_ins1 = 0.3, rr_restoration_ins2=0.5,correct_mix_deploy = 0)

a <- setExposure(exposure=exposure, insecticideUsed=strategy,
                 male_exposure_prop=male_exposure_prop, correct_mix_deploy=1)

defaultmediummix=setInputOneScenario(P_1=0.001, P_2=0.01,a=a,  exposure = 0.8, male_exposure_prop=1, phi.SS1_A0=0.6, phi.SS2_0B=0.8, h.RS1_A0	=0.4, h.RS2_0B = 0.8, rr_restoration_ins1 = 0.6, rr_restoration_ins2=0.8,correct_mix_deploy = 0)

a <- setExposure(exposure=exposure, insecticideUsed=strategy,
                 male_exposure_prop=male_exposure_prop, correct_mix_deploy=1)

defaulthighmix=setInputOneScenario(P_1=0.001, P_2=0.01,a=a,  exposure = 0.8, male_exposure_prop=1, phi.SS1_A0=1, phi.SS2_0B=1, h.RS1_A0	=0.4, h.RS2_0B = 0.8, rr_restoration_ins1 = 1, rr_restoration_ins2=1,correct_mix_deploy = 1)

a <- setExposure(exposure=exposure, insecticideUsed=strategy,
                 male_exposure_prop=male_exposure_prop, correct_mix_deploy=0)

defaultlowmos=setInputOneScenario(P_1=0.001, P_2=0.01,a=a,  exposure = 0.8, male_exposure_prop=1, phi.SS1_A0=0.3, phi.SS2_0B=0.5, h.RS1_A0	=0.4, h.RS2_0B = 0.8,rr_restoration_ins1 = 0.3, rr_restoration_ins2=0.5,correct_mix_deploy = 0)

a <- setExposure(exposure=exposure, insecticideUsed=strategy,
                 male_exposure_prop=male_exposure_prop, correct_mix_deploy=0)

defaultmediummos=setInputOneScenario(P_1=0.001, P_2=0.01,a=a,  exposure = 0.8, male_exposure_prop=1, phi.SS1_A0=0.6, phi.SS2_0B=0.8, h.RS1_A0	=0.4, h.RS2_0B = 0.8, rr_restoration_ins1 = 0.6, rr_restoration_ins2=0.8,correct_mix_deploy = 0)

a <- setExposure(exposure=exposure, insecticideUsed=strategy,
                 male_exposure_prop=male_exposure_prop, correct_mix_deploy=0)

defaulthighmos=setInputOneScenario(P_1=0.001, P_2=0.01,a=a,  exposure = 0.8, male_exposure_prop=1, phi.SS1_A0=1, phi.SS2_0B=1, h.RS1_A0	=0.4, h.RS2_0B = 0.8, rr_restoration_ins1 = 1, rr_restoration_ins2=1,correct_mix_deploy = 0)

#The following code generates the resistance for each scenario.

listOutLMix <- runModel2(defaultlowmix, produce.plots = FALSE)

listOutLMos <- runModel2(defaultlowmos, produce.plots = FALSE)

listOutMMix <- runModel2(defaultmediummix, produce.plots = FALSE)

listOutMMos <- runModel2(defaultmediummos, produce.plots = FALSE)

listOutHMix <- runModel2(defaulthighmix, produce.plots = FALSE)

listOutHMos <- runModel2(defaulthighmos, produce.plots = FALSE)
```

```{r, eval=TRUE} 
#To plot the variance in fitness of LOW killing for a MIXTURE
plot_fit_variance(listOutLMix)
#To plot the variance in fitness of LOW killing for a MOSAIC
plot_fit_variance(listOutLMos)
#To plot the variance in fitness of MEDIUm killing for a MIXTURE
plot_fit_variance(listOutMMix)
#To plot the variance in fitness of MEDIUM killing for a MOSAIC
plot_fit_variance(listOutMMos)
#To plot the variance in fitness of HIGH killing for a MIXTURE
plot_fit_variance(listOutHMix)
#To plot the variance in fitness of HIGH killing for a MOSAIC
plot_fit_variance(listOutHMos)
```

### 2016-11-25 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
bump version

@@
 #output: word_document
-author: "Andy South, Ian Hastings and Bethany Levick"
+author: "Andy South and Ian Hastings"
 date: "`r Sys.Date()`"
@@
 
-### version 6
+### version 7
 

### 2016-12-09 inst/documents/effectiveness_exposure_difference.Rmd
typo
increase 2x2 plot size
nice 4 panel fig with labels working

@@
 
-scaneario exploration :
+scenario exploration :
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=5, fig.show='hold', fig.align='left', fig.cap="test A.  B. C. D."}
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="test A.  B. C. D."}
 
@@
 
-layout(matrix(1:4,2,2, byrow = TRUE))
+#layout(matrix(1:4,2,2, byrow = TRUE))
+#add lower panel for legend
+#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
+
+# layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
+# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
+
+#layout.show(n=5)
+
+# layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.2,0.2), 
+# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
+
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
 #effectiveness on mixture
 
-#plot margins
-oldpar <- par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+#testing
+# plot(1)
+# plot(1)
+# plot(1)
+# plot(1)
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('exposure: 0.5', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
 
 #base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
@@
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('exposure: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
 
@@
 
-par(oldpar)
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
 
 ```
+
+<!-- par(mar=c(0, 0, 0, 0)) -->
+<!-- layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.3),  -->
+<!-- respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) ) -->
+
+<!-- #layout.show(n=5) -->
+
+<!-- plot(1) -->
+<!-- plot(1) -->
+<!-- plot(1) -->
+<!-- plot(1) -->
+
+<!-- #dev.new(height=5,width=10) -->
+<!-- plot( 0, type="n", axes=TRUE, ann=FALSE, pty='m', fin=c(5,1)) #par("din"))  -->
+
+<!-- legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE ) -->
+
+<!-- par(oldpar) -->

### 2016-12-09 inst/documents/effectiveness_exposure_difference.Rmd
added abcd labs

@@
 #although I'm not sure it would ?
-#allow x axis labels and title to be removed
-#add axis titles outside of the plot region
-#remove curve legend on plots and put it somewhere else maybe on bottom
+
 #? allow curves to be thicker
@@
 #layout(matrix(1:4,2,2, byrow = TRUE))
+
 #add lower panel for legend
@@
 
-# layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
+#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
 # respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-
 #layout.show(n=5)
 
-# layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.2,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-
+# add side panels for labels
 layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
@@
 
+mtext('A)',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
 #plot margins
@@
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('B)',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
+mtext('C)',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
 #plot margins
@@
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('D)',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 

### 2016-12-12 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
develop abstract

@@
 
+or
+
+#Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
+
 Target : Malaria journal
@@
 ## ANDY TO DO
+1. add to intro including about about results in levick et al.
 1. work out what's going on with exposure ... why does increasing exposure have a greater effect on mixture than sequential ? 
-1. add to intro including about about results in levick et al. 
 1. explain resistance-restoration and its relation to selection coefficient
 1. expand discussion to include relevance to real world 
+1.
+in discussion address potential criticism that these results are just for a limited region of paramter space (a.
+point to Levick et al.
+that explored more of space, b.
+point to UI that allows readers to explore spce for themselves and identify whether there are places where this does not hold).
+1. ? briefly refer to linkage disequilibrium and say we are investigating.
+
+1. can I develop explanatory diagrams to describe the input parameters ?
+1. mention generation time of 1 month
+1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
+1.
+add to discussion about relevance of these mechanisms in the field e.g.
+exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
+(Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
 
@@
 
-This paper explores the mechanisms by which insectide resistance is expected to increase with the use of insecticides in isolation, sequence and mixtures.
-It is aimed at a readership with limited experience of population genetics and resistance.
-It uses a model representing two independent genes coding for resistance to two insecticides.
-It explores factors that are expected to influence the development of resitance, notably the starting frequency of resistance alleles, their dominance, the exposure of insects to the insecticides, the effectiveness of the insecticides, and how resistance restores fitness in the presence of the insecticides.
-We use the model to develop a mechanistic understanding of how the values of these parameters are expected to influence the development of resistance and the relative benefits of alternative insecticide application strategies.
-We expect this mechanistic understanding to be more robust to uncertainty in parameter values in the field.
+Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes.
+To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones.
+There are different potential strategies for using new insecticides and some recomendations for which may be best.
+There is however limited recent, accessible modelling work to support suggestions of which strategies may be best in different situations.
+Previous work suggests that preferred strategies will be situation specific.
+This points to a need for an accessible, mechanistic understanding of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
+Such a mechanistic understanding would be more robust to uncertainty in parameter values in field situations.
+
+Here we use a relatively simple model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures.
+We target a readership with limited experience of population genetics and resistance.
+The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies.
+We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide,  and how much the resistance counteracts the effect of the insecticide.
+
+We show firstly that using a second insecticide, at the same time as a first, always slows evolution of resistance to the first.
+Secondly we show that the ability of an insecticide to kill the susceptible mosquitoes (effectiveness), is the most important factor influencing whether using insecticides in a sequence or a mixture is likely to lead to faster resistance to both of them.
+These results are explained in terms of the mechanisms that cause them.
+We discuss how these results can inform intervention strategies in the field.
+
 We describe an R package and user-interface that allows further exploration of the development of resistance.
@@
 
+http://www.sciencemag.org/news/2016/10/after-40-years-most-important-weapon-against-mosquitoes-may-be-failing
+
 Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.
@@
 We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
+_is their research indicating that mnost mutations already there?
+something from ASTMH?_
 
@@
 ~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
-
-## Todo
-
-1. In discussion and maybe elewhere address potential concern that the relationships shown here may not hold for all of paramter space.
-1. can I develop explanatory diagrams to describe the input parameters ?
-1. mention generation time of 1 month
-1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
-1. add to discussion about relevance of these mechanisms in the field
-1.
-e.g.
-exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
-(Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
 
@@
 
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+All control inputs equal at 0.5, B.
+Effectiveness of insecticide1 increased from 0.5 to 0.8, C.
+Exposure increased to 0.8, D.
+Effectiveness of insecticide1 and exposure increased to 0.8"}
+
+#4 panel figure
+
+#think about implementing in ggplot2 to make my life easier later ?
+#although I'm not sure it would ?
+
+#? allow curves to be thicker
+
+#test of rearranging figures on a grid
+#eff1  0.5  0.8
+#exp
+#0.5    1   2
+#0.8    3   4
+
+#layout(matrix(1:4,2,2, byrow = TRUE))
+
+#add lower panel for legend
+#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
+
+#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
+# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
+#layout.show(n=5)
+
+# add side panels for labels
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
+#effectiveness on mixture
+
+#testing
+# plot(1)
+# plot(1)
+# plot(1)
+# plot(1)
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('exposure: 0.5', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+
+#base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#effectiveness1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('exposure: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#exposure 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1 & expos 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+
+```
+
 ```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.

### 2016-12-13 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
adding referencing and putting in own folder

--- 2016-12-12 inst/documents/paper2_resistance_mechanisms_mixtures.Rmd
@@
 date: "`r Sys.Date()`"
+bibliography: library_ir.bib
+csl: biomed-central.csl
 ---
@@
 
-# Mechanisms by which insecticide resistance can be influenced by strategies of insecticide use : exploring by simulation.
+#Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
 
-or
-
-#Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
+Then I can focus on how effectiveness, unlike other parameters has a different effect on mixtures & sequences.
+(I can refer to the poor application, dominance as a side issue)
 
@@
 
-Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes.
+Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potential huge consequences for malaria burden.
 To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones.
-There are different potential strategies for using new insecticides and some recomendations for which may be best.
+There are different potential strategies for using new insecticides and recomendations for how best to preserve their utility for controlling mosquitoes and malaria.
 There is however limited recent, accessible modelling work to support suggestions of which strategies may be best in different situations.
@@
 
-## Introduction
+## Background
 
-http://www.sciencemag.org/news/2016/10/after-40-years-most-important-weapon-against-mosquitoes-may-be-failing
-
-Insecticide resistance is a problem for malaria (GPIRM), other vector borne diseases and agriculture.
+Insecticide resistance is a problem for malaria (GPIRM)[@Who2012][@Ranson2016] other vector borne diseases and agriculture.
 
@@
 
-We use a 2-locus population genetic model of insecticide resistance (Levick et al.
-in press) to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
+We use a 2-locus population genetic model of insecticide resistance [@Levick2017] to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
 Levick et al.
@@
 This makes sense as vector kill is effectively a product of exposure times effectiveness.
-For example exposing 50% of a  population to an insecticide which is 75% effective would be expecetd to have the same result as exposing 75% of the population to one which is 50% effective.
+For example exposing 50% of a  population to an insecticide which is 75% effective would be expected to have the same result as exposing 75% of the population to one which is 50% effective.
 This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance.
@@
 Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
-
-## References
-1. GPIRM
-1. Curtis 
-1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
-1. Birget 2015
-1. Curtis et al 1993, 
-1. Georghiou 1983 Management of resistance in Arthropods.
-potential IRM strategies :
-~ moderation : preserve susceptible genes by limiting selection pressure
-~ saturation : high dose so that heterozygous resistants are killed
-~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
 
@@
 ![Summarising mechanisms of single insecticide effects.](paper2_input_effects_single.png)
+
+\pagebreak
+
+## ReferencesOld
+
+1. Barbosa S, Black WC IV, Hastings I (2011) Challenges in Estimating Insecticide Selection Pressures from Mosquito Field Data. PLoS Negl Trop Dis 5(11):
+e1387. doi:10.1371/journal.pntd.0001387
+_suggests incomplete dominance in the field_
+
+1. Birget PLG, Koella JC. 2015 A genetic model of the effects of insecticide-treated bed nets on the evolution of insecticide-resistance. Evol Med Public Health. (1):205-15. doi: 10.1093/emph/eov019.
+
+1. Churcher et al. 2016 The impact of pyrethroid resistance on the efficacy and effectiveness of bednets for malaria control in Africa. eLife
+_Transmission dynamics models indicate that even low levels of resistance would increase the incidence of malaria due to reduced mosquito mortality and lower overall community protection over the life-time of the net._
+_+ may have some useful stuff about expected degradation of nets over time_
+
+1. Curtis CF. 1985 Theoretical models of the use of insecticide mixtures for the management of resistance. Bull Entomol Res. 75: 259-65.
+
+1. Curtis et al 1993,
+
+1. Georghiou 1983 Management of resistance in Arthropods.
+potential IRM strategies :
+~ moderation : preserve susceptible genes by limiting selection pressure
+~ saturation : high dose so that heterozygous resistants are killed
+~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
+
+1. Levick et al (in press) A two-locus model of the evolution of insecticide resistance to inform and optimise public health insecticide deployment strategies. PLOS Computational Biology.
+
+1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
+
+1. R Core Team. 2016 R: A language and environment for statistical computing.
+
+1.
+Ranson, H.
+and Lissenden, N.
+2016 Insecticide resistance in African anopheles mosquitoes: A worsening situation that needs urgent action to maintain malaria control.
+Trends in Parasitology, 32, 187-196, http://dx.doi.org/10.1016/j.pt.2015.11.010.
+
+1. Roush RT. 1993 Occurrence, Genetics and Management of Insecticide Resistance. Parasitol Today. 9: 174-9.
+
+1. World Health Organization. 2012. Global plan for insecticide resistance management in malaria vectors (GPIRM). Geneva.
+
+## References

### 2016-12-14 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
tweak text, add to legend

@@
 
-Then I can focus on how effectiveness, unlike other parameters has a different effect on mixtures & sequences.
-(I can refer to the poor application, dominance as a side issue)
+Then I can focus on how effectiveness, unlike other parameters has a different effect on mixtures & sequences. (I can refer to the poor application, dominance as a side issue)
 
@@
 ## ANDY TO DO
+1. should I include male exposure prop ?
 1. add to intro including about about results in levick et al.
@@
 1. can I develop explanatory diagrams to describe the input parameters ?
-1. mention generation time of 1 month
 1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
@@
 (Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
+1. mention in discussion that repellency would lead to decrease in exposure
 
@@
 ------------------- | -------------------------------------------------------
-1. Effectiveness | proportion of exposed insects with both susceptible alleles that are killed by the insecticide
-2. Exposure | proportion of the insect population that are exposed to the insecticide
-3. Dominance | dominance of the resistance allele relative to the susceptible allele
-4. Resistance restoration  | ability of resistance to restore fitness in the presence of the insecticide
+1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
+2. Exposure | proportion of insects exposed to the insecticide
+3. Dominance | resistance of an insect with only 1 resistance allele (SR)
+4. Resistance restoration  | ability of resistance (RR) to restore fitness when insect exposed to the insecticide
 5. Frequency  | frequency of resistance alleles within the population
@@
 
-**Ian, Matt this is still the bit I can't quite explain. Any suggestions ?**
+**work on explaining this without getting too carried away**
 
@@
 
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#add for when insecticides same
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
 

### 2016-12-14 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
replacing old figures with grid of 4 figures

@@
 ## ANDY TO DO
-1. should I include male exposure prop ?
 1. add to intro including about about results in levick et al.
@@
 1. mention in discussion that repellency would lead to decrease in exposure
+1. is there a refernce to say that the effectiveness of insecticides can be modified at the design stage ?
 
@@
 When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold of 50%.
+
+_TODO rewrite this & probably put effectiveness and exposure in together_
+_starting like this_
+In the baseline scenario where the effectiveness of both insecticides is set the same at 0.5 the resistance frequencies at each locus for the mixture remain identical (the single solid line in Fig 6A).
+For this scenario time-to-resistance is longer when the insecticides are used in sequence (dashed lines).
+Increasing exposure to the insecticides from 0.5 to 0.8 (Fig 6C) decreases the time-to-resistance for both the sequential and mixture strategy and so does not change their relative performance; time-to-resistance remains longer for the sequence.
 
@@
 ```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
-All control inputs equal at 0.5, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.8, C.
-Exposure increased to 0.8, D.
-Effectiveness of insecticide1 and exposure increased to 0.8"}
+All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B.
+Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C.
+Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D.
+Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence"}
 
-#4 panel figure
-
-#think about implementing in ggplot2 to make my life easier later ?
-#although I'm not sure it would ?
-
-#? allow curves to be thicker
+#4 panel figure looking at effectiveness and exposure
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.8"}
-#effectiveness on mixture
+Resistance restoration for insecticide1 increased from 0.5 to 0.8, C.
+Dominance for insecticide1 increased to 0.8, D.
+Resistance restoration and dominance for insecticide1 increased to 0.8"}
+
+#4 panel figure looking at dominance and resistance restoration
+
+#rr1  0.5  0.8
+#dom1
+#0.5    1   2
+#0.8    3   4
+
+# add side panels for labels
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
+#testing
+# plot(1)
+# plot(1)
+# plot(1)
+# plot(1)
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('dominance1: 0.5', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+
+#base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 
 #plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-#base scenario
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
+#rr1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('dominance1: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
 
 #plot margins
-par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab="", ylabs = FALSE, main='B.', maxX = 120 )
+#dominance1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
 
-par(oldpar)
+mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#dom1 & rr1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+
+mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#add for when insecticides same
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('resistance restoration1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('resistance restoration1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="Effectiveness of both insecticides 0.8"}
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
+Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
+Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D.
+Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.001 to 0.0001"}
+
+#4 panel figure looking at effectiveness and starting freq.
+
+#eff1  0.5  0.8
+#strt freq1
+#0.01    1   2
+#0.001    3   4
+
+# add side panels for labels
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('frequency1: 1%', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+
+#base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 
 #plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), maxX = 120 )
+#eff1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('frequency1: 0.1%', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#freq1 0.001
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#freq1 0.001, eff1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#add for when insecticides same
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
-Exposure increased to 0.8, B.
-Exposure and Effectiveness of insecticide1 increased from 0.5 to 0.8"}
-#fig.cap seems necessary to keep figs side by side
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+All control inputs equal at 0.5, male exposure 0.8, B.
+Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
+Male exposure decreased from 0.8 to 0.2, D.
+Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
+
+#4 panel figure looking at male exposure prop and starting freq.
+
+#eff1  0.5  0.8
+#maleprop
+#0.8    1   2
+#0.2    3   4
+
+# add side panels for labels
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('male exposure: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+
+#base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
+
+mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 
 #plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
-#cat("Fig. Exposure 0.8 benefit of sequence over mixture increases")
+#eff1 0.8
+runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
+
+mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('male exposure:0.2', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
 
 #plot margins
-par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylabs = FALSE, main='B.', maxX = 120 )
-#cat("Fig. Effectiveness 0.8, Exposure 0.8, increased exposure nullifies the effect of increased effectiveness, sequence remains better")
-```
+#freq1 0.001
+runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A. Dominance of insecticide1 0.8, B. Dominance of both insecticides 0.8"}
+mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 
 #plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
-#cat("Dominance of insecticide1 0.8")
+#freq1 0.001, eff1 0.8
+runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
 
+mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#adding legend in lower panel
 #plot margins
-par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.8 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', ylabs = FALSE, main='B.', maxX = 120 )
-#cat("Dominance of both insecticides 0.8")
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
 
-par(oldpar)
+# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
 
-```
+#add for when insecticides same
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
-Resistance restoration coefficient allele1 0.8, B.
-Resistance restoration coefficient both alleles 0.8"}
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
 
-#plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 120 )
-#cat("Fig. rr_restoration I1 0.8")
-
-#plot margins
-par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.8 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', ylabs = FALSE, main='B.', maxX = 120 )
-#cat("Fig. CC. rr_restoration I2 0.2")
-
-par(oldpar)
-
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3.7, fig.show='hold', fig.align='left', fig.cap="A.
-Start frequency allele1 reduced from 0.001, B.
-as A plus effectiveness insecticide1 increased to 0.8"}
-
-#plot margins
-oldpar <- par(mar=c(4, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#P_1 to 0.001 seq remains better
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), addLegend=FALSE, main='A.', maxX = 150 )
-
-#plot margins
-par(mar=c(4, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#P_1 to 0.001 & effect1 to 0.8 mix remains better
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('seq','','adapt','mix2'), ylab='', ylabs = FALSE, main='B.', maxX = 150 )
-
-par(oldpar)
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
 
@@
 ## References
+<!-- now auto created here-->

### 2016-12-15 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
add effectiveness fig, tweak label sizes

@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure: 0.5', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('exposure: 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 #effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('exposure: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 #eff1 & expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+```
+
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Higher effectivenesses for both insecticides, times-to-resistance longer for the mixture in all scenarios.
+A.
+Effectiveness of insecticide1 increased from 0.5 to 0.8, B.
+Effectiveness of insecticide1 increased from 0.5 to 1, C.
+Effectiveness of insecticides 1 and 2 increased to 0.8, D.
+Effectiveness of insecticide2 0.8 and of insecticide1 1"}
+
+#4 panel figure looking at effectiveness changes alone
+
+#test of rearranging figures on a grid
+#eff1  0.8  1
+#eff2
+#0.5    1   2
+#0.8    3   4
+
+# add side panels for labels
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness2: 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+
+#base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1 1 eff2 0.5
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness2: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1 0.8 eff2 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1 1 eff2 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+
+mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#add for when insecticides same
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1: 0.5', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('dominance1: 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('dominance1: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 #dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+mtext('resistance restoration1: 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('resistance restoration1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1: 1%', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('frequency1: 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
 
@@
 #eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1: 0.1%', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('frequency1: 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
 
@@
 #freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.8) #side=1b,2l,3t,4r
+mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure: 0.8', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('male exposure: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
 
@@
 #eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure:0.2', line= -2, side=2, cex=0.8) #side=1b,2l,3t,4r
+mtext('male exposure:0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
 
@@
 #freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.6, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
 

### 2016-12-15 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
rewriting results

@@
 date: "`r Sys.Date()`"
-bibliography: library_ir.bib
+bibliography: IR.bib
 csl: biomed-central.csl
@@
 
-```{r,echo=FALSE,results='hide'}
+```{r,echo=FALSE,results='hide',warning=FALSE}
   library(resistance)
@@
 ## ANDY TO DO
+1. add a screenshot of the UI
 1. add to intro including about about results in levick et al.
@@
 
-Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potential huge consequences for malaria burden.
+Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potentially huge consequences for malaria burden.
 To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones.
@@
 We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
-_is their research indicating that mnost mutations already there?
+_is their research indicating that most mutations already there?
 something from ASTMH?_
@@
 
-So far this description has just considered a single insecticide and associated resistance allele (Fig 1A).
-In the simulation a second insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
+So far this description just considers a single insecticide and associated resistance allele, with equal behaviour for males and females (Fig 1A).
+To allow males and females to be exposed differently separate genders are also represented (Fig 1B).
+In addition a second insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
 Thus the simulation can represent populations exposed to two insecticides together in a mixture.
-
-In addition the model structure allows for 3 different levels of exposure to be set for each insecticide (Fig 2).
-So in the two insecticide case that gives the potential for nine (3*3) different exposure niches.
-In the runs shown here we have just used two levels of exposure for each insecticide (effectively no or yes with the value for yes set from the exposure input in Table 1).
-The model structure allows males and females to experience different exposures, we explore that in Levick et al.
-(in press) but the runs presented here use equal exposures.
 
@@
 ### Single insecticide
-For single insecticide use, faster resistance spread resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 3-5).
-In the plots each input was varied in isolation with the remaining set to 0.5 (except for starting frequency of resistance which was set to 0.01 in all bar Fig 5).
+For single insecticide use, faster resistance spread resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 2-4).
+In the plots each input was varied in isolation with the remaining set to 0.5 (except for starting frequency of resistance which was set to 0.01 in all bar Fig 4).
 
@@
 Values for the first four inputs in Table 1 were set to 0.5 and the starting frequencies of resistance to 0.01.
-For this base scenario resistance arises slower for sequential than mixture (Fig 6A).
+For this base scenario resistance arises slower for sequential than mixture (Fig 5A).
 This plot, and those that follow, compares mixture and sequence for a single combination of inputs.
@@
 
-_TODO rewrite this & probably put effectiveness and exposure in together_
-_starting like this_
-In the baseline scenario where the effectiveness of both insecticides is set the same at 0.5 the resistance frequencies at each locus for the mixture remain identical (the single solid line in Fig 6A).
-For this scenario time-to-resistance is longer when the insecticides are used in sequence (dashed lines).
-Increasing exposure to the insecticides from 0.5 to 0.8 (Fig 6C) decreases the time-to-resistance for both the sequential and mixture strategy and so does not change their relative performance; time-to-resistance remains longer for the sequence.
-
-#### results: Insecticide effectiveness
-A more effective insecticide increases time-to-resistance when used in a mixture. This is opposite to what happens when used in sequence and in isolation, when both shortened times-to-resistance.
-
-For example, taking the base scenario and increasing the effectiveness of one insecticide from 0.5 to 0.8 increases the time-to-resistance for the mixture (from 90 to 115 generations) and decreases that for the sequence (from 120 to 90 generations) (Fig 6).
-Thus at low effectiveness a sequential strategy is favoured while at higher effectiveness a mixture is favoured.
-
-When the effectiveness of both insecticides are increased to 0.8 the time-to-resistance is increased still further to around 125 generations (Fig 7).
-
-#### results: Exposure
-Increasing exposure to 0.8 from the base scenario reduces time-to-resistance for both mixture and sequential scenarios (Fig 8A relative to 6A).
-The increase in exposure has a greater effect on the mixture relative to the sequential strategy and this results in a greater relative benefit for sequence versus mixture.
-
-Similarly increasing exposure reduces time-to-resistance more for the mixture relative to sequence in the scenario in which one insecticide effectiveness has been increased (compare Fig 8B to 6B).
-The result is that the advantage of mixtures over sequence is removed by the increase in exposure.
-
-#### results: Dominance of resistance alleles
-Increasing the dominance of the alleles within a mixture has the same direction of effect as it does for sole use.
-Increasing the dominance of the allele corresponding to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance to insecticide1 within the mixture but leaves the time-to-resistance for the second insecticide in the mixture the same (Fig 9A).
-Setting the dominance for both alleles to 0.8 reduces the time-to-resistance for both the mixture and the sequence by 20 generations from the base scenario (compare Fig 9B to 6A).
-In both of these cases sequential use continues to have a longer time-to-resistance under the higher dominance.
-
-#### results: Resistance restoration coefficient
-Increasing the resistance restoration coefficient to insecticide1 from 0.5 to 0.8 decreases the time-to-resistance for insecticide1 in the mixture, and has a similar effect on it's use in the sequence (Fig 10A).
-However, it has little effect on resistance to insecticide2.
-Because of this sequential use continues to generate a longer time-to-resistance than the mixture.
-Increasing the resistance restoration coefficient of both alleles to 0.8 results in a shorter time-to-resistance  for the mixture (Fig 10B), however the relative shape of the resistance curves remains similar to the base scenario  (Fig 6A), with sequential use continuing to generate longer time-to-resistance.
+#### results: Insecticide exposure and effectiveness
+
+Increasing exposure to the insecticides (the proportion of insects that come into contact) from 0.5 to 0.8 (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy and so does not change their relative performance; time-to-resistance remains longer for the sequence.
+In contrast keeping the exposure constant at 0.5 and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides to 0.8 (Fig 5D) results in a longer time-to-resistance for the mixture relative to the sequence.
+Resistance for the more effective insecticide (in red) increases faster than the lesser.
+Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effecttive insecticide has reached the resistance threshold (~50 generations).
+
+Thus a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
+This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
+
+Figure 5D has the same inputs as Fig 5B except that the exposure is increased from 0.5 to 0.8.
+The increase in exposure reduces time-to-resistance more for the mixture relative to the sequence.
+The result is that the advantage of mixtures over sequence is removed by the increase in exposure, leading to approximately equal times-to-resistance for the two strategies (Fig 5D).
+
+Figure 6 shows how increasing the effectiveness of either or both insecticides improves the performance of mixtures over sequences in terms of time-to-resistance.
+Figure 6A uses the same inputs as Fig 5B.
+From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
+
+#### results: Dominance and resistance restoration
+Increasing dominance or resistance restoration decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
+The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
 
@@
 Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 11A to Fig 6A).
-Similarly taking the base scenario with effectiveness of the first insecticide increased to 0.8 (Fig 6B) and decreasing starting frequency of that resistance allele did not result in the mixture being favoured (Fig 11B).
+For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
+Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
+
+TODO Fig 9 male exposure proportion
 
@@
 
-When single insecticide use was represented (Figs 3-5) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
+When single insecticide use was represented (Figs 2-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
 Higher values of the five inputs investigated all led to more rapid increases in resistance.
 
-For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
+For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 2A to 2B).
 This makes sense as vector kill is effectively a product of exposure times effectiveness.
@@
 
-The pattern of more rapid increase in resistance at higher dominance levels (Fig 4A) can be explained by increased survival of heterozygotes in the presence of the insecticide.
+The pattern of more rapid increase in resistance at higher dominance levels (Fig 3A) can be explained by increased survival of heterozygotes in the presence of the insecticide.
 Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival.
@@
 
-The faster development of resistance under higher resistance-restoration (Fig 4B) can be explained by it's effect on the survival of the resistant genotypes.
+The faster development of resistance under higher resistance-restoration (Fig 3B) can be explained by it's effect on the survival of the resistant genotypes.
 Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.
 
-The effect of the starting frequency of resistance (Fig 5) is the most different from the other inputs and perhaps the easiest to explain.
+The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs and perhaps the easiest to explain.
 Simply when starting from a higher frequency of resistance there is a smaller change to make to reach the higher resistance thresholds.
 
-These mechanisms explaining the response of resistance to single insecticide use under different input values are summarised in Table 2 and Fig 12.
+These mechanisms explaining the response of resistance to single insecticide use under different input values are summarised in Table 2.
 Exposure and effectiveness act by decreasing the survival of all genotypes, dominance acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
@@
 In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
-Looking at our base scenario (Fig 6A) the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
+Looking at our base scenario (Fig 5A) the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
 The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
@@
 
+TODO check figure numbering in next section
 #### discussion: Insecticide effectiveness
 
-The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (Fig 6B).
-In contrast in both the sequential (Fig 6B) and single insecticide (Fig 3A) experiments resistance was speeded up by increasing effectiveness.
+The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (Fig 5B).
+In contrast in both the sequential (Fig 5B) and single insecticide (Fig 2A) experiments resistance was speeded up by increasing effectiveness.
 This can be explained by the effect that in a mixture increased effectiveness of one insecticide increases killing of vectors that are resistant to the other and thus decreases selection pressure.
 
-This effect can be seen by comparing Fig 6B, where the effectiveness of insecticide 1 has been increased, to 6A.
+This effect can be seen by comparing Fig 5B, where the effectiveness of insecticide 1 has been increased, to 5A.
 In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
@@
 This can be explained by the mechanism that once insecticide 1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide 2 and thus stops 'protecting' it from it's own rise in resistance.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 7).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 6A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 6B.
-
-Fig 6B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal resistance will be expected to increase faster to the more effective insecticide.
+A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
+The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
+In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
+
+Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal resistance will be expected to increase faster to the more effective insecticide.
 The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
@@
 
-![Exposure to insecticides in the model.](paper2_mod_graphic_d.png)
-
 ```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Effectiveness, B. Exposure"}
@@
 
-![Summarising mechanisms of single insecticide effects.](paper2_input_effects_single.png)
-
 \pagebreak

### 2016-12-16 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-### version 7
+### version 8
 
 #Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
-
-Then I can focus on how effectiveness, unlike other parameters has a different effect on mixtures & sequences. (I can refer to the poor application, dominance as a side issue)
 
@@
 1. add to intro including about about results in levick et al.
-1. work out what's going on with exposure ... why does increasing exposure have a greater effect on mixture than sequential ? 
 1. explain resistance-restoration and its relation to selection coefficient
@@
 1. can I develop explanatory diagrams to describe the input parameters ?
-1. mention other potential effects e.g. differential exposure of M&F, sex linkage, ...
+1. mention other potential effects e.g. costs of resistance, sex linkage, ...
 1.
@@
 
-Insecticide resistance is a problem for malaria (GPIRM)[@Who2012][@Ranson2016] other vector borne diseases and agriculture.
+Insecticide resistance is a problem for malaria [@Who2012][@Ranson2016] other vector borne diseases and agriculture.
 
@@
 We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
-_is their research indicating that most mutations already there?
-something from ASTMH?_
+There are suggestions [e.g.
+ref from ASTMH] that mutations are present in populations even prior to exposure to a novel insecticide thus leading to the potential for selection.
 
@@
 
+The model is implemented in R [RCoreTeam2016], the code is hosted on Github.
+
 ## Results
@@
 
-TODO Fig 9 male exposure proportion
+#### results: reduced male exposure
+_? maybe cut this section and fig 9_
+Reduced male exposure with respect to female (Fig 9) has a similar effect to changing overall exposure.
+Reducing male exposure (Figs 9C & D) increases times-to-resistance and favours mixtures over sequences.
 
@@
 
-The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs and perhaps the easiest to explain.
-Simply when starting from a higher frequency of resistance there is a smaller change to make to reach the higher resistance thresholds.
+The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs.
+Simply, when starting from a higher frequency of resistance there is a smaller change to make to reach the higher resistance thresholds.
 
@@
 
-TODO check figure numbering in next section
 #### discussion: Insecticide effectiveness
 
-The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (Fig 5B).
-In contrast in both the sequential (Fig 5B) and single insecticide (Fig 2A) experiments resistance was speeded up by increasing effectiveness.
-This can be explained by the effect that in a mixture increased effectiveness of one insecticide increases killing of vectors that are resistant to the other and thus decreases selection pressure.
-
-This effect can be seen by comparing Fig 5B, where the effectiveness of insecticide 1 has been increased, to 5A.
+The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (compare Fig 5B to 5A).
+In contrast in both the sequential (Fig 5B) and single insecticide (Fig 2A) scenarios the development of resistance was faster when effectiveness was higher.
+This can be explained by the increased effectiveness of one insecticide in a mixture increasing killing of vectors that are resistant to the other and thus decreasing selection pressure.
+
+This effect can be seen by comparing Fig 5B, where the effectiveness of insecticide 1 has been increased, to Fig 5A.
 In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
@@
 Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-This can be explained by the mechanism that once insecticide 1 has become ineffective due to resistance it stops killing indivduals that are resistant to insecticide 2 and thus stops 'protecting' it from it's own rise in resistance.
+Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide and thus stops limiting the rise in resistance to the second insecticide.
 A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
@@
 
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal resistance will be expected to increase faster to the more effective insecticide.
+Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide.
 The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
@@
 
-**work on explaining this without getting too carried away**
-
-The biggest difference I can't explain is why that in mixtures, increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 in Fig 6A to 40 in Fig 8A) and yet increasing the effectiveness of both insecticides by the same amount results in a slight increase in time-to-resistance (from 90 in Fig 6A to 100 in Fig 7).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (compare Figs 7 & 8A).
-
-either
-
-1. Why doesn't increased exposure provide 'protection' to the other insecticides in a mixture ?
-1. Why doesn't increased effectiveness lead to a greater selection for resistance in each insecticide ?
-1. Why do exposure & effectivness have the same effect on a sequence but different effects on a mixture ?
-
-*some previous guesswork I'm now not sure about: Increasing exposure gives shorter time-to-resistance for both mixtures and sequences due to increased deaths of susceptibles creating a higher selection pressure.
-In a mixture the effect of exposure is greater and this means that increasing exposure can make sequences preferable to mixtures.
-The reason for the greater effect of exposure on mixtures could be that exposure is having both a direct effect on each insecticide and an indirect effect on the other.
-Thus higher exposures lead to faster resistance in each insecticide (direct), in turn increased resistance to each insecticide reduces the 'protection' that it gives the other, leading to faster resistance in the other (indirect).
-This could act as a postive feedback loop where faster resistance in one contributes to faster resistance in the other.*
+It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
+Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 in Fig 5A to 40 in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 110 in Fig 5B).
+This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 2 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
+
+_todo : develop a concise explanation of this difference between effectiveness and exposure_
 
@@
 
-(@) ?
-In a mixture the effectiveness of one insecticide has major effects on the resistance of the other.
-In contrast the dominance and selective advantage of one insecticide has less effect on the resistance of the other.
-For exposure we kept values the same for each insecticide (to represent a mixture in which the 2 insecticides are applied together).
-However we predict that if they were applied separately then (similar to effectiveness) the level of exposure to one insecticide would influence the time-to-resistance of the other.
+### TODO ADD TO DISCUSSION : link these refs to the results here
+[@Mani1985]
+"use of mixtures is always more effective in delaying the onset of resistance, often by many orders of magnitude"
+~ assumed that SS always killed i.e. effectiveness=1
+~ exposure fixed at 0.9
+
+[@Gould1986]
+~ insecticidal toxins genetically engineered into plants
+~ additive rtaher than multiplicative fitness
+~ effectiveness ~0.2-0.5 (might be expected to favour sequences by our results)
+~ no clearly superior strategy
+
+[@Roush1998]
+~ identified mortality of SS homozygotes (our “insecticide effectiveness” parameter) as critical (his Fig 3) and concluded (page 1784) :
+“As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective”
+
+[@Georghiou1983]
+potential IRM strategies :
+~ moderation : preserve susceptible genes by limiting selection pressure
+~ saturation : high dose so that heterozygous resistants are killed
+~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
+
+[@Barbosa2011] 
+_suggests incomplete dominance in the field_
+
+[@Birget2015] 
+We show that indoor use of insecticides leads to less selection pressure than their use as larvicides. 
+Reasons for relatively low selection pressure by adulticides
+(i) males are not affected by the ITNs
+(ii) insecticides are also repellents, keeping mosquitoes at bay from contacting the insecticide but also driving them to bite either people who do not use the insecticide or alternative hosts.
+
+[@Churcher2016] 
+The impact of pyrethroid resistance on the efficacy and effectiveness of bednets for malaria control in Africa.
+Transmission dynamics models indicate that even low levels of resistance would increase the incidence of malaria due to reduced mosquito mortality and lower overall community protection over the life-time of the net.
+_+ may have some useful stuff about expected degradation of nets over time_
+
+[@Curtis1998]
+can anything be done to maintain effectiveness of nets ?
+
+[@Thomas2012]
+Lessons from agriculture for malaria vector management
+~ suggests that partially effective tools used in combination as a part of Integrated Vector Managament is a more sustainable solution in the light of evolution
+~ this modelling approach could be used to investigate implications of such combined strategies
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="*KEY FIG OF PAPER* Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
+A.
 All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B.
@@
 
-## ReferencesOld
-
-1. Barbosa S, Black WC IV, Hastings I (2011) Challenges in Estimating Insecticide Selection Pressures from Mosquito Field Data. PLoS Negl Trop Dis 5(11):
-e1387. doi:10.1371/journal.pntd.0001387
-_suggests incomplete dominance in the field_
-
-1. Birget PLG, Koella JC. 2015 A genetic model of the effects of insecticide-treated bed nets on the evolution of insecticide-resistance. Evol Med Public Health. (1):205-15. doi: 10.1093/emph/eov019.
-
-1. Churcher et al. 2016 The impact of pyrethroid resistance on the efficacy and effectiveness of bednets for malaria control in Africa. eLife
-_Transmission dynamics models indicate that even low levels of resistance would increase the incidence of malaria due to reduced mosquito mortality and lower overall community protection over the life-time of the net._
-_+ may have some useful stuff about expected degradation of nets over time_
-
-1. Curtis CF. 1985 Theoretical models of the use of insecticide mixtures for the management of resistance. Bull Entomol Res. 75: 259-65.
-
-1. Curtis et al 1993,
-
-1. Georghiou 1983 Management of resistance in Arthropods.
-potential IRM strategies :
-~ moderation : preserve susceptible genes by limiting selection pressure
-~ saturation : high dose so that heterozygous resistants are killed
-~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
-
-1. Levick et al (in press) A two-locus model of the evolution of insecticide resistance to inform and optimise public health insecticide deployment strategies. PLOS Computational Biology.
-
-1. Mani GS. Evolution of resistance in the presence of two insecticides. Genetics 1985;109(4):761–83.
-
-1. R Core Team. 2016 R: A language and environment for statistical computing.
-
-1.
-Ranson, H.
-and Lissenden, N.
-2016 Insecticide resistance in African anopheles mosquitoes: A worsening situation that needs urgent action to maintain malaria control.
-Trends in Parasitology, 32, 187-196, http://dx.doi.org/10.1016/j.pt.2015.11.010.
-
-1. Roush RT. 1993 Occurrence, Genetics and Management of Insecticide Resistance. Parasitol Today. 9: 174-9.
-
-1. World Health Organization. 2012. Global plan for insecticide resistance management in malaria vectors (GPIRM). Geneva.
-
 ## References

### 2017-01-03 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
pap2 writing

@@
 
-Insecticide resistance is a problem for malaria [@Who2012][@Ranson2016] other vector borne diseases and agriculture.
+Insecticide resistance is a problem for malaria [@Who2012][@Ranson2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
+
+TODO add more stuff here or after following para about protocols for resistance management.
 

### 2017-01-03 inst/documents/paper1/old/pap1_figs_slimmed_50_unequal.Rmd
tidy pap1 files into own folder

--- 2016-06-27 inst/documents/pap1_figs_slimmed_50_unequal.Rmd
@@
 
-  #in_cols <- c("exposure","effectiveness_insecticide1","dominance_allele1")  
+  #in_cols <- c("exposure","effectiveness_ins1","dominance_allele1")  
   #4unequal
-  in_cols <- c("exposure","dominance_allele1","selection_coef_allele1","effectiveness_insecticide1")
+  in_cols <- c("exposure","dominance_allele1","selection_coef_allele1","effectiveness_ins1")
 
@@
   # x1=df_same$exposure 
-  # y1=df_same$effectiveness_insecticide1 + df_same$effectiveness_insecticide2
+  # y1=df_same$effectiveness_ins1 + df_same$effectiveness_ins2
   # 
@@
   # x1=df_20pc$exposure 
-  # y1=df_20pc$effectiveness_insecticide1 + df_20pc$effectiveness_insecticide2
+  # y1=df_20pc$effectiveness_ins1 + df_20pc$effectiveness_ins2
   # 

### 2017-01-03 inst/documents/paper1/old/paper1_results_figs.Rmd
tidy pap1 files into own folder

--- 2016-06-27 inst/documents/paper1_results_figs.Rmd
@@
     #failed hex plot just gave me counts
-    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+    # print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_ins1', colour='mixA_better_seq0.25'))+
     #          geom_hex()
@@
     #interesting plot showing where sequence is better is restricted to higher values of exposure and effectiveness 
-    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_insecticide1', colour='mixA_better_seq0.25'))+
+    print( ggplot(ggSubset, aes_string(x='exposure', y='effectiveness_ins1', colour='mixA_better_seq0.25'))+
              #geom_hex()
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_better_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 * effectiveness_ins2, colour=mixA_better_seq0.25))+
              #geom_hex()
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mixA_better_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=mixA_better_seq0.25))+
 
@@
     #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mixA0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=seq_better_mixA0.25))+
              facet_grid( ~ seq_better_mixA0.25, labeller=as_labeller( c('0'='adaptive mix better','1'='sequential better'))) +
@@
     #but the previous one may be clearer
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 * effectiveness_insecticide2, colour=mixA_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 * effectiveness_ins2, colour=mixA_minus_seq0.25))+
              #geom_hex()
@@
     #failed attempt at getting Curtis point in legend
-    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25, size="Curtis"))+
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
+    #print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mixA_minus_seq0.25, size="Curtis"))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mixA_minus_seq0.25))+
 
@@
 
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mixA_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mixA_minus_seq0.25))+
 
@@
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_mixA_minus_seq0.25))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2))+ #, colour=abs_mixA_minus_seq0.25))+
 
@@
     #aha! yes that's even better
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=mix2_better_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=mix2_better_seq0.25))+
 
@@
     #if I get this sorted I could try to have +ve & -ve on the same scale
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=seq_better_mix20.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=seq_better_mix20.25))+
              facet_grid( ~ seq_better_mix20.25, labeller=as_labeller( c('0'='mix better','1'='sequential better'))) +
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mix2_minus_seq0.25))+
 
@@
     library(viridis)
-    print( ggplot(ggSubset, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_insecticide1 + effectiveness_insecticide2, colour=abs_mix2_minus_seq0.25))+
+    print( ggplot(ggSubset, aes(x=exposure+exposure*male_exposure_prop, y=effectiveness_ins1 + effectiveness_ins2, colour=abs_mix2_minus_seq0.25))+
 

### 2017-01-03 inst/documents/paper1/old/paper1_results_figs_slimmed.Rmd
tidy pap1 files into own folder

--- 2016-06-27 inst/documents/paper1_results_figs_slimmed.Rmd
@@
   #subset which output plots to show in the main paper
-  in_cols <- c("start_freq_allele1", "exposure","effectiveness_insecticide1",
+  in_cols <- c("start_freq_allele1", "exposure","effectiveness_ins1",
               "dominance_allele1","selection_coef_allele1","male_exposure_prop")
@@
              #add text for Curtis point
-             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_insecticide1 + effectiveness_insecticide2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='red')
+             geom_label(data = curtisInputs, inherit.aes = FALSE, aes_string(x = 'exposure', y='effectiveness_ins1 + effectiveness_ins2', label='"Curtis Fig.2"', hjust='"right"'), nudge_x=-0.02, size=3, colour='red')
            )       
@@
   #modify the killer plot to show where mixtures are 20% better ?
-  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_insecticide1 + effectiveness_insecticide2))+ #, colour=abs_diff))+
+  print( ggplot(ggSubset, aes(x=exposure, y=effectiveness_ins1 + effectiveness_ins2))+ #, colour=abs_diff))+
 

### 2017-01-03 inst/documents/paper1/paper1_results_figs_slimmed_50_rr.Rmd
tidy pap1 files into own folder

--- 2016-11-11 inst/documents/paper1_results_figs_slimmed_50_rr.Rmd
@@
                               colour=factor(dominance,labels=c("0   recessive",
-                                                               "0.5 partial",
+                                                               "0.5 semi-dominant",
                                                                "1   dominant")))) + 
@@
        #coord_trans(y = "log10") +    
-       geom_line()
+       geom_line() +
+       theme(axis.text.x = element_blank(),
+             axis.ticks = element_blank(),
+             panel.grid = element_blank())  
        #annotate('text',label='A)',x=-10, y=1.1)
@@
 
-df_labs <- data.frame(lab=rep(c('RR','RS','SS'),4),
-                      x=c(10,10,10,30,30,30,50,50,50,70,70,70),
-                      y=c(50,46,42,38,34,30,26,22,18,14,10,6),
-                      #col=c(2,1,1,2,1.5,1,2,2,1,2,2,2))
-                      col=c(1,0,0,1,0.5,0,1,1,0,1,1,1))
-
-df_boxes <- data.frame(lab=c('R recessive\n(slow spread)',
+#22/11/16 adding extra section on left where all killed following request from Ian
+#first try adding on left @-10 and modify x extent so I don't have to fiddle with all coords
+df_labs <- data.frame(lab=rep(c('RR','RS','SS'),5),
+                      x=c(-10,-10,-10,10,10,10,30,30,30,50,50,50,70,70,70),
+                      y=seq(62,6,-4),
+                      #y=c(50,46,42,38,34,30,26,22,18,14,10,6),
+                      col=c(0,0,0,1,0,0,1,0.5,0,1,1,0,1,1,1))
+
+df_boxes <- data.frame(lab=c('All killed\n(no spread)',
+                             'R recessive\n(slow spread)',
                              'R semi-dominant\n(rapid spread)',
                              'R dominant\n(rapid spread)',
-                             'No selection for R'),
-                      x=c(10,30,50,80),
-                      y=c(90,75,60,35))
-
-plotlist[[2]] <- ggplot(data.frame(x=c(0,100),y=c(80,0)), aes(x=x, y=y)) + 
+                             'All survive\n(no spread)'),
+                      x=c(-10,10,30,50,70),
+                      y=c(105,90,75,60,45))
+
+# df_labs <- data.frame(lab=rep(c('RR','RS','SS'),4),
+#                       x=c(10,10,10,30,30,30,50,50,50,70,70,70),
+#                       y=c(50,46,42,38,34,30,26,22,18,14,10,6),
+#                       #col=c(2,1,1,2,1.5,1,2,2,1,2,2,2))
+#                       col=c(1,0,0,1,0.5,0,1,1,0,1,1,1))
+# 
+# df_boxes <- data.frame(lab=c('R recessive\n(slow spread)',
+#                              'R semi-dominant\n(rapid spread)',
+#                              'R dominant\n(rapid spread)',
+#                              'No selection for R'),
+#                       x=c(10,30,50,80),
+#                       y=c(90,75,60,35))
+
+plotlist[[2]] <- ggplot(data.frame(x=c(-20,85),y=c(90,15)), aes(x=x, y=y)) + 
        theme_bw() + 
-       ylim(0,100) +
+       ylim(0,115) +
+       xlim(-20,85) +
        coord_cartesian(expand = FALSE) + #so that axes fit data exactly
@@
        geom_line(size=2, colour=rgb(0.5,0,0.5)) +
-       geom_vline(xintercept = c(20,40,60), linetype=2) +
+       geom_vline(xintercept = c(0,20,40,60), linetype=2) +
        geom_text(data=df_labs,aes(x=x, y=y, label=lab, colour=factor(col)),size=4) +
        scale_colour_manual(values = c("0" = "red","0.5" = "orange","1" = "green"), guide=FALSE) +  
-       geom_label(data=df_boxes,aes(x=x, y=y, label=lab)) +
+       geom_label(data=df_boxes,aes(x=x, y=y, label=lab),size=3) +
        theme(axis.text = element_blank(),

### 2017-01-03 inst/documents/paper1/paper1_results_figs_slimmed_50_rr.Rmd
increase font size in fig 8

@@
     #adds in no. scenarios correctly and incorrectly classified
-    #cex to set textsize (0.7 all text doesn't quite fit in), good that A&B are now same size
+    #cex to set textsize, good that A&B are now same size
     #round=0 to make boxes square
@@
     prp(tree, extra=1, varlen = 0, main=main, box.col=cols, under=TRUE, 
-        cex=0.4, round=0, split.font=1, ycompress = TRUE )
+        cex=0.6, round=0, split.font=1, ycompress = TRUE,
+        eq="\n= ", lt="\n< ", ge="\n>= " )
   }

### 2017-01-06 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

--- 2016-12-16 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
 date: "`r Sys.Date()`"
-bibliography: IR.bib
+bibliography: library.bib
 csl: biomed-central.csl
+#header-includes:
+#    - \usepackage{setspace}
+#    - \doublespacing
+#    - \usepackage{lineno}
+#    - \linenumbers
 ---
+
+```{r, eval=FALSE, include=FALSE}
+#!!!! paper protocol
+#!!!! might be easier to put text & figs into separate Rmd files after draft stage
+#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
+#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
+#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
+#!!!! submission figs, set this chunk to eval=TRUE
+
+# options to create final publication quality figures
+# causes the document to fail, but the figs do get stored in pap2figs/
+library(knitr)
+opts_chunk$set(dev="tiff",
+               dev.args=list(compression="lzw"),
+               dpi=300,
+               cache=FALSE,
+               fig.path='pap2figs/')
+```
 
@@
 
-(This file is automagically generated by R from paper2_resistance_mechanisms_mixtures.Rmd)
-
 ## ANDY TO DO
-1. add a screenshot of the UI
-1. add to intro including about about results in levick et al.
-1. explain resistance-restoration and its relation to selection coefficient
-1. expand discussion to include relevance to real world 
-1.
-in discussion address potential criticism that these results are just for a limited region of paramter space (a.
-point to Levick et al.
-that explored more of space, b.
-point to UI that allows readers to explore spce for themselves and identify whether there are places where this does not hold).
-1. ? briefly refer to linkage disequilibrium and say we are investigating.
-
+1. mention single gene vs additive effects (with ref.)
+1. ? briefly refer to linkage disequilibrium and say we are investigating OR include a figure OR refer to one of my existing figures
 1. can I develop explanatory diagrams to describe the input parameters ?
-1. mention other potential effects e.g. costs of resistance, sex linkage, ...
-1.
-add to discussion about relevance of these mechanisms in the field e.g.
-exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
-(Birget & Koella 2015 mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
-1. mention in discussion that repellency would lead to decrease in exposure
-1. is there a refernce to say that the effectiveness of insecticides can be modified at the design stage ?
+1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage ?
 
@@
 
+_must be 350 words, currently 375 with subheadings Background, Results & Conclusions_
+
+##### Background
 Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potentially huge consequences for malaria burden.
@@
 Here we use a relatively simple model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures.
-We target a readership with limited experience of population genetics and resistance.
 The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies.
@@
 
+##### Results
 We show firstly that using a second insecticide, at the same time as a first, always slows evolution of resistance to the first.
@@
 
+##### Conclusions
 We describe an R package and user-interface that allows further exploration of the development of resistance.
@@
 
+## Keywords
+insecticide resistance; public health; mosquitoes; vector-borne diseases; infectious diseases; malaria; population genetics
+
 ## Background
 
-Insecticide resistance is a problem for malaria [@Who2012][@Ranson2016] other vector borne diseases and agriculture.
+Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
+
+Public health efforts in controlling malaria have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
+Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million malaria cases between 2000 and 2015 [@Bhatt2016].
+
+A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
+
+The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
+Several new insecticides are now in development [@IVCC2016] and very soon decisions will need to be made about how best to use them.
+The model presented here can help inform strategies to minimise the developm,ent of resiostance in new and existing insecticides.
 
@@
 
+_TODO add in here about theory of IR & protocols for resistance management._
+
 We use a 2-locus population genetic model of insecticide resistance [@Levick2017] to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
-Levick et al.
-(in press) describe the technical details of the model and explore large parameter ranges.
+In an earlier paper [@Levick2017] we described the technical details of the model and explore large parameter ranges.
 We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance.
-This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide use strategies.
+This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Rousch1989][@IRAC2011][@FAO2012][@GPIRM2012].
 
@@
 Each allele confers either resistance or susceptibility to the insecticide.
+Fitness is the main currency of the model, representing how much each genotype survives and reproduces, with a maximum value of 1.
 The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model.
 These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
+Each of these inputs can be varied as a proportion between 0 and 1.
+We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
+The more conventional 'selection coefficient' is calculated by multiplying Resistance restoration by insecticide effectiveness.
+Doing it this way allows us to vary all inputs between 0 and 1 without fitness going above 1 in the model.
 
@@
 
-Here we principally look at model outputs of the change in allele frequencies over time (generations). We investigate three insecticide use strategies.
+Here we principally look at model outputs of the change in resistance allele frequencies over time (generations).
+Resistance allele frequencies are calculated from the frequencies of each genotype as follows.
+
+> Resistance allele frequency = 2*RR + RS
+
+We investigate three insecticide use strategies.
 
 1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches 50%
+1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches a threshold of 50%
 1. two insecticides used in a mixture (concurrently)
 
-The model is implemented in R [RCoreTeam2016], the code is hosted on Github.
+The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an online user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
 
@@
 ### Single insecticide
-For single insecticide use, faster resistance spread resulted from higher values of any of the inputs outlined in Table 1, namely effectiveness, exposure, dominance, the resistance restoration coefficient, and starting frequencies of resistance (Figs 2-4).
-In the plots each input was varied in isolation with the remaining set to 0.5 (except for starting frequency of resistance which was set to 0.01 in all bar Fig 4).
+For single insecticide use, higher values of insecticide effectiveness, exposure, dominance or resistance restoration resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach the resistance threshold of either 0.5 or 0.75.
+
+Similarly, higher values of starting resistance frequency also resulted in faster spread of resistance (Fig 4) and shorter times to resistance thresholds.
 
@@
 
-When single insecticide use was represented (Figs 2-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
+When single insecticide use was represented (Figs 3-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
 Higher values of the five inputs investigated all led to more rapid increases in resistance.
 
-For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 2A to 2B).
+For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
 This makes sense as vector kill is effectively a product of exposure times effectiveness.
@@
 
-The pattern of more rapid increase in resistance at higher dominance levels (Fig 3A) can be explained by increased survival of heterozygotes in the presence of the insecticide.
+The pattern of more rapid increase in resistance at higher dominance levels (Fig 3C) can be explained by increased survival of heterozygotes in the presence of the insecticide.
 Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival.
@@
 
-The faster development of resistance under higher resistance-restoration (Fig 3B) can be explained by it's effect on the survival of the resistant genotypes.
+The faster development of resistance under higher resistance-restoration (Fig 3D) can be explained by it's effect on the survival of the resistant genotypes.
 Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.
@@
 
-### Discussion summary points
-
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
-Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
-
-(@) For a mixture of 2 insecticides with differing effectiveness.
-Resistance to the insecticide with a greater effectiveness increases faster.
-The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
-
-(@) exposure ...
-
-### TODO ADD TO DISCUSSION : link these refs to the results here
-[@Mani1985]
-"use of mixtures is always more effective in delaying the onset of resistance, often by many orders of magnitude"
-~ assumed that SS always killed i.e. effectiveness=1
-~ exposure fixed at 0.9
-
-[@Gould1986]
-~ insecticidal toxins genetically engineered into plants
-~ additive rtaher than multiplicative fitness
-~ effectiveness ~0.2-0.5 (might be expected to favour sequences by our results)
-~ no clearly superior strategy
-
-[@Roush1998]
+It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria.
+As such our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
+Using no insecticide could have serious consequences for mosquito and malaria abundance.
+The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while and sustaining that in the longer term by restricting the development of resistance.
+Others are looking at combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
+
+The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ransom2016].
+The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of paramter space (e.g.
+holding most inputs constant at intermediate values while modifying single inputs in isolation).
+Thus there is the potential that the system may behave differently in the field than we describe here.
+However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
+In addition we provide an online user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown here.
+
+#### discussion: putting results into context (some parts may want to go earlier)
+I will now put these predicted responses of resistance into the context of previous work and recomendations.
+
+Most of these predicted reponses of resistance are consistent with recomendations that have been produced by WHO[@WHO2012], IRAC[@IRAC2011] and FAO[@FAO2012].
+
+[Next bit might want to go in intro, but not if I refer to our results]
+Earlier modelling work mostly from the 1980s has made sometimes contrasting predictions and these can be explained by the details of their structure and the particular situations they represent.
+One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
+That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
+This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
+As described earlier we particular predict that the very high effectiveness is likely to favour mixtures over sequences.
+
+[@Roush1989]
+
+The model we describe is more flexible and as such can reproduce similar results to these earlier approaches and be used to represent more detailed, operationally relevant strategies.
+
+[@Thomas2012]
+Lessons from agriculture for malaria vector management
+~ suggests that partially effective tools used in combination as a part of Integrated Vector Managament is a more sustainable solution in the light of evolution
+~ this modelling approach could be used to investigate implications of such combined strategies
+
+Relevance of these mechanisms in the field e.g.
+exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
+[@Birget2015a] mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
+_TODO check which of the Birget papers mention this_
+
+One consistent aspect of the results presented here can at first seem to go against current advice on good practice.
+We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5).
+This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance.
+In contrast the widely held view seems to be that low effectiveness or exposure associated with poor implementation can promote the development of resistance.
+e.g.
+from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
+These include: ...
+• the use of application rates that are below or above those recommended on the label; • poor coverage of the area being treated ...".
+The explanation for this apparent difference is in the effect of dominance.
+Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?].
+Our model does show that increasing dominance leads to faster spread of resaistance (Fig 3C, 7).
+So, the effect of poor implementation of an insecticide intervention on resistance could be eiether positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
+Our model cobined with laboratory and field results could help to assess these trade-offs.
+
+[@Taillebois2016]
+Laboratory experiments on aphids showing that in low dose mixtures insecticides most often had additive effects as represented in our model, but that also commonly mortality was higher than expected (synergistic) which we can also represent within the model.
+
+[@Roush1989]
 ~ identified mortality of SS homozygotes (our “insecticide effectiveness” parameter) as critical (his Fig 3) and concluded (page 1784) :
@@
 
-[@Georghiou1983]
+[@Georghiou1994]
 potential IRM strategies :
@@
 
-[@Thomas2012]
-Lessons from agriculture for malaria vector management
-~ suggests that partially effective tools used in combination as a part of Integrated Vector Managament is a more sustainable solution in the light of evolution
-~ this modelling approach could be used to investigate implications of such combined strategies
+### Discussion summary points
+
+(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
+
+(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
+Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
+
+(@) For a mixture of 2 insecticides with differing effectiveness.
+Resistance to the insecticide with a greater effectiveness increases faster.
+The less effective insecticide is 'protected' by the more effective.
+Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
+
+(@) exposure ...
+
+_todo mention other potential effects e.g. costs of resistance, sex linkage, ..._
+
+_TODO : Should I try to cite in press Killeen about combining interventions ?_
+
+TODO FINAL CONCLUDING PARAGRAPH
+~ Effectiveness and exposure are key and can be modified.
+~ The modelling approach can be used to investigate more detailed strategies.
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Effectiveness, B. Exposure"}
-
+![Screenshot of one online model user interface, accessible at : https://andysouth.shinyapps.io/resistmob2/](paper2_UI1.png)
+
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on changing resistance frequency over time with single insecticide use.
+A.
+Effectiveness, B.
+Exposure, C.
+Dominance, D.
+Resistance restoration.
+For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
+"}
+
+  library(cowplot)
+
+  #list to put plots in for cowplot
+  plotlist <- list(4)
+
+  #A effectiveness  
   a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
+  #A  
+  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
          theme_bw() +  
@@
 
-       )
-
+  #B exposure
   exposure <- 0.2
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
+  #B  
+  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
          theme_bw() +
@@
 
-       )
-
-```
-
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on single insecticide use. A. Dominance, B. Resistance restoration"}
-
-  #single insecticide dominance fig
-
+  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')
+
+  #C single insecticide dominance
   a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
+  #C 
+  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
          theme_bw() +
@@
 
-       )
-
-  # single insecticide resistance restoration fig.
+  #D single insecticide resistance restoration
 
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
+  #D  
+  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
          theme_bw() +
@@
          geom_line()
-       )
+
+  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of resistance starting frequency on single insecticide use."}
+```{r Fig3, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of changing resistance starting frequency on changing resistance frequency over time with single insecticide use."}
 
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="*KEY FIG OF PAPER* Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
+```{r Fig4, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="*KEY FIG OF PAPER* Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
 A.
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Higher effectivenesses for both insecticides, times-to-resistance longer for the mixture in all scenarios.
+```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Higher effectivenesses for both insecticides, times-to-resistance longer for the mixture in all scenarios.
 A.
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, B.
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
@@
 
-```{r, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, male exposure 0.8, B.

### 2017-01-11 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

@@
 
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. All comments and & suggestions welcome.
-
-## ANDY TO DO
-1. mention single gene vs additive effects (with ref.)
-1. ? briefly refer to linkage disequilibrium and say we are investigating OR include a figure OR refer to one of my existing figures
-1. can I develop explanatory diagrams to describe the input parameters ?
-1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage ?
+THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and & suggestions welcome
 
@@
 
-Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
-
-Public health efforts in controlling malaria have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
-Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million malaria cases between 2000 and 2015 [@Bhatt2016].
-
+Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
+Focusing on malaria which still results in hundreds of thousands of deaths per year, recent control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
+Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million malaria cases between 2000 and 2015 [@Bhatt2015].
 A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
 
+There have been modelling studies investigating the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
+[@Curtis1985][@Mani1985][@Roush1989] but much of the modelling work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
+For those who (wisely in some cases) distrust models, work in a moth and genetically modified brocolli system has shown that toxins used in combination can delay the onset of resistance in a way that is consistent with theoretical models [@Zhao2003].
+
 The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
-Several new insecticides are now in development [@IVCC2016] and very soon decisions will need to be made about how best to use them.
-The model presented here can help inform strategies to minimise the developm,ent of resiostance in new and existing insecticides.
+Three new insecticides are now in development [@IVCC2016] and likely to be available around 2020 [@Ranson2016].
+It is important that decisions about how best to use the new insecticides to delay the onset of resistance are made before insecticides are released [@Hemingway2016].
+The model presented here can help inform strategies to minimise the development of resistance in new and existing insecticides.
 
@@
 We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
-There are suggestions [e.g.
-ref from ASTMH] that mutations are present in populations even prior to exposure to a novel insecticide thus leading to the potential for selection.
-
-The changing frequency of insecticide resistance is a population genetic process that can be influenced by the parameters outlined in Table 1.
+It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
+The changing frequency of insecticide resistance is a population genetic process that can be influenced by (among others) the parameters outlined in Table 1.
 
@@
 We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance.
-This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Rousch1989][@IRAC2011][@FAO2012][@GPIRM2012].
+This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
 
@@
 This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-
 Figure 5D has the same inputs as Fig 5B except that the exposure is increased from 0.5 to 0.8.
@@
 From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
+
+Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
+Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
 
@@
 
-#### results: reduced male exposure
-_? maybe cut this section and fig 9_
-Reduced male exposure with respect to female (Fig 9) has a similar effect to changing overall exposure.
-Reducing male exposure (Figs 9C & D) increases times-to-resistance and favours mixtures over sequences.
-
 ## Discussion
 
-The response of resistance to insecticide use in the model can be explained by a number of mechanisms that can help us develop a more robust understanding of how we would expect resistance to respond in the field.
+The response of resistance to insecticide use in the model can be explained by mechanisms that can help us develop and communicate a more robust understanding of how we would expect resistance to respond in the field.
 
@@
 Higher values of the five inputs investigated all led to more rapid increases in resistance.
-
 For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
@@
 Thus selection pressure for the resistance allele will be increased because it confers more of an advantage when only present on one chromosome.
-
 The faster development of resistance under higher resistance-restoration (Fig 3D) can be explained by it's effect on the survival of the resistant genotypes.
 Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.
-
 The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs.
@@
 
-These mechanisms explaining the response of resistance to single insecticide use under different input values are summarised in Table 2.
+These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
 Exposure and effectiveness act by decreasing the survival of all genotypes, dominance acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
@@
 
-#### discussion: Exposure
-
+#### discussion: Exposure vs effectiveness
 It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
 Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 in Fig 5A to 40 in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 110 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 2 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-
-_todo : develop a concise explanation of this difference between effectiveness and exposure_
-
-Mechanisms for the response of resistance in mixtures are summarised in Table 3. Table 4 summarises mechanisms explaining the difference between mixture and sequential strategies.
-
+This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
+This difference between effectiveness and exposure was unexpected for us.
+_todo could this be because exposure results doesn't alter the relative fitness of genotypes because all genotypes are exposed equally ? doesn't quite make sense to me_
+
+The mechanisms involved in the response of resistance in mixtures are summarised in Table3.
+Whether a mixture or sequence is favoured for reducing the spread of resistance will depend on these mechanisms and is summarised in Table 4.
+
+#### discussion: caveats
 It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria.
@@
 
-#### discussion: putting results into context (some parts may want to go earlier)
+#### discussion: putting results into context
+
 I will now put these predicted responses of resistance into the context of previous work and recomendations.
-
-Most of these predicted reponses of resistance are consistent with recomendations that have been produced by WHO[@WHO2012], IRAC[@IRAC2011] and FAO[@FAO2012].
-
-[Next bit might want to go in intro, but not if I refer to our results]
-Earlier modelling work mostly from the 1980s has made sometimes contrasting predictions and these can be explained by the details of their structure and the particular situations they represent.
+Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and the particular situations they represent.
 One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
@@
 This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-As described earlier we particular predict that the very high effectiveness is likely to favour mixtures over sequences.
-
-[@Roush1989]
-
-The model we describe is more flexible and as such can reproduce similar results to these earlier approaches and be used to represent more detailed, operationally relevant strategies.
-
-[@Thomas2012]
-Lessons from agriculture for malaria vector management
-~ suggests that partially effective tools used in combination as a part of Integrated Vector Managament is a more sustainable solution in the light of evolution
-~ this modelling approach could be used to investigate implications of such combined strategies
+This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
+see Fig 6D).
+
+A more recent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
+This "mortality of susceptible homozygotes" is equivalent to the parameter we have termed insecticide effectiveness, so lends further support to our main conclusion of the importance of effectiveness.
+It's also shown [@Roush1998 Fig 3] that decreasing exposure from 90% to 80% also increases the benefit of mixtures over sequences in the same way that we have shown.
+
+Modelling and theory have been used to generalise Insecticide Resistance Management into three potential strategies [@Georghiou1994].
+These are A) moderation : preserving susceptible genes by limiting selection pressure, B) saturation : high dose so that heterozygous resistants are killed and C) multiple attack : independently acting pressures neither of which is strong enough to lead to resistance.
+Our work here principally focuses on multiple attack but the other strategies are relevant too.
+_todo expand this paragraph_
+
+One consistent aspect of the results presented here can at first seem to go against current advice on good practice.
+We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5).
+This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
+@Roush1998].
+In contrast the widely held view seems to be that low effectiveness or exposure associated with poor implementation can promote the development of resistance.
+e.g.
+from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
+These include: ...
+- the use of application rates that are below or above those recommended on the label; - poor coverage of the area being treated ...".
+The explanation for this apparent difference is in the effect of dominance.
+Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?ref].
+Our model does show that increasing dominance leads to faster spread of resaistance (Fig 3C, 7).
+So, the effect of poor implementation of an insecticide intervention on resistance could be either positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
+Our model cobined with laboratory and field results could help to assess these trade-offs.
+
+Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
+It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
+A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
+Mosquitoes with different behaviours e.g.
+feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
+This lower exposure could favour mixtures over sequences as we have indicated.
+In both cases our modelling approach could be used to investigate implications for the devlopment of resistance.
+
+## FINAL CONCLUDING PARAGRAPH
+
+~ Effectiveness and exposure are key and can be modified.
+~ The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies.
+
+### other papers to incorporate into discussion above
+
+Roush older paper : [@Roush1989]
 
@@
 
-One consistent aspect of the results presented here can at first seem to go against current advice on good practice.
-We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance.
-In contrast the widely held view seems to be that low effectiveness or exposure associated with poor implementation can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-• the use of application rates that are below or above those recommended on the label; • poor coverage of the area being treated ...".
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?].
-Our model does show that increasing dominance leads to faster spread of resaistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be eiether positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Our model cobined with laboratory and field results could help to assess these trade-offs.
+How do the rest of our results relate to recomendations that have been produced by WHO[@WHO2012], IRAC[@IRAC2011] and FAO[@FAO2012] ?
 
@@
 Laboratory experiments on aphids showing that in low dose mixtures insecticides most often had additive effects as represented in our model, but that also commonly mortality was higher than expected (synergistic) which we can also represent within the model.
-
-[@Roush1989]
-~ identified mortality of SS homozygotes (our “insecticide effectiveness” parameter) as critical (his Fig 3) and concluded (page 1784) :
-“As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective”
-
-[@Georghiou1994]
-potential IRM strategies :
-~ moderation : preserve susceptible genes by limiting selection pressure
-~ saturation : high dose so that heterozygous resistants are killed
-~ multiple attack : independently acting pressures neither of which is strong enough to lead to resistance
 
@@
 
-[@Churcher2016] 
-The impact of pyrethroid resistance on the efficacy and effectiveness of bednets for malaria control in Africa.
-Transmission dynamics models indicate that even low levels of resistance would increase the incidence of malaria due to reduced mosquito mortality and lower overall community protection over the life-time of the net.
-_+ may have some useful stuff about expected degradation of nets over time_
-
 [@Curtis1998]
@@
 
-### Discussion summary points
+[@Kliot2012]
+fitness costs associated with insecticide resistance
+
+### Discussion some summary points probably already covered ...
 
@@
 (@) exposure ...
-
-_todo mention other potential effects e.g. costs of resistance, sex linkage, ..._
-
-_TODO : Should I try to cite in press Killeen about combining interventions ?_
-
-TODO FINAL CONCLUDING PARAGRAPH
-~ Effectiveness and exposure are key and can be modified.
-~ The modelling approach can be used to investigate more detailed strategies.
-
-[Summary tables to refer to from discussion.]
 
@@
 
+## ANDY TO DO
+1. mention single gene vs additive effects (with ref.)
+1. can I add a diagram to help with methods, e.g. perhaps fitness under difference inputs, similar to fitnessvis ?
+1. ? briefly refer to linkage disequilibrium and say we are investigating OR include a figure OR refer to one of my existing figures
+1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage, or by the dose ?
+1. mention other potential effects e.g. costs of resistance, sex linkage, ..._
+1. cite in press Killeen about combining interventions
+
 \pagebreak
@@
 
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig8, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, male exposure 0.8, B.
@@
 
+# not evaluated at the moment
 #4 panel figure looking at male exposure prop and starting freq.

### 2017-01-23 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
replace : with = in fig legends

@@
 
-### version 8
+### version 9
+
+#An accessible model to help understand the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
+
+#A tutorial, primer ...
 
@@
 
+Good review of tactics for IRM inlcuding mention of mixtures and new AIs [@Denholm1992]
+"Once developed and marketed, new products introduce a fresh challenge how to exploit their characteristics while restricting exposure to preserve their effectiveness."
+
 ### Discussion some summary points probably already covered ...
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure: 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('exposure = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2: 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('effectiveness2 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('effectiveness2 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1: 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1 = 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1: 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('dominance1 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('dominance1 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1: 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1: 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+mtext('resistance restoration1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('resistance restoration1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1: 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('frequency1 = 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1: 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('frequency1 = 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure: 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('male exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure:0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('male exposure =0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 

### 2017-02-01 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 This points to a need for an accessible, mechanistic understanding of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-Such a mechanistic understanding would be more robust to uncertainty in parameter values in field situations.
+Such a mechanistic understanding could allow a more robust response to uncertainty in parameter values in field situations.*finesse that sentence or move to discussion*
 
@@
 The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide,  and how much the resistance counteracts the effect of the insecticide.
+We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide, and how much the resistance alleles counteract the effect of the insecticide.
 
 ##### Results
-We show firstly that using a second insecticide, at the same time as a first, always slows evolution of resistance to the first.
+We show firstly that using a second insecticide, in a mixture with a first, always slows evolution of resistance to the first.
 Secondly we show that the ability of an insecticide to kill the susceptible mosquitoes (effectiveness), is the most important factor influencing whether using insecticides in a sequence or a mixture is likely to lead to faster resistance to both of them.
-These results are explained in terms of the mechanisms that cause them.
+The mechanisms behind these results are explained.
 We discuss how these results can inform intervention strategies in the field.
@@
 Each allele confers either resistance or susceptibility to the insecticide.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces, with a maximum value of 1.
+Fitness is the main currency of the model, representing how much each genotype survives and reproduces, with a maximum value of 1 defined as the fitness of the SS genotype in absence of the insecticide.
 The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model.
@@
 > 
-> Allele fitness :   
+> Genotype fitness :   
 > SS1SS2 = SS * SS  
@@
 > 
-> Population fitness = Allele fitness * exposure to insecticide  
+> Population fitness = Genotype fitness * exposure to insecticide  
 > 
@@
 
-> Resistance allele frequency = 2*RR + RS
+> Resistance allele frequency = (2*RR + RS) / 2
 
@@
 When two insecticides are used the relative performance of a mixture strategy compared to sequential use changes dependent on input values.
-To help us understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs individually.
+To help understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs individually.
 The base scenario was similar to that used for the single insecticide.
@@
 In contrast keeping the exposure constant at 0.5 and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides to 0.8 (Fig 5D) results in a longer time-to-resistance for the mixture relative to the sequence.
-Resistance for the more effective insecticide (in red) increases faster than the lesser.
+Resistance for the more effective insecticide (in red) increases faster than the less effective.
 Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effecttive insecticide has reached the resistance threshold (~50 generations).
@@
 Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
+
+It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
+Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
+In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
 
@@
 It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 in Fig 5A to 40 in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 110 in Fig 5B).
+Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 110 in Fig 5B).
 This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
@@
 Using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while and sustaining that in the longer term by restricting the development of resistance.
+The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining that in the longer term by restricting the development of resistance.
 Others are looking at combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
@@
 
-A more recent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
+A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
 This "mortality of susceptible homozygotes" is equivalent to the parameter we have termed insecticide effectiveness, so lends further support to our main conclusion of the importance of effectiveness.
@@
 ## ANDY TO DO
+1. remove first person for Malaria journal
+1. explain resistance thresholds and time-to-resistance in methods
+1. refer to ratios of mix:seq in plots
+1. refer to PRCC from pap1 in ref. to Table4
 1. mention single gene vs additive effects (with ref.)
 1. can I add a diagram to help with methods, e.g. perhaps fitness under difference inputs, similar to fitnessvis ?
-1. ? briefly refer to linkage disequilibrium and say we are investigating OR include a figure OR refer to one of my existing figures
+1. in discussion refer to linkage disequilibrium about Fig 8D
 1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage, or by the dose ?
@@
 
-![Building up the model structure.](paper2_mod_graphic_a_to_c.png)
-
-![Screenshot of one online model user interface, accessible at : https://andysouth.shinyapps.io/resistmob2/](paper2_UI1.png)
-
-```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Effect of inputs on changing resistance frequency over time with single insecticide use.
+![New figure showing how fitness determined by inputs](paper2_mod_graphic_a_to_c.png)
+
+![Screenshot of one online model user interface, accessible at : https://andysouth.shinyapps.io/resistmob2/.
+The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
+Two scenarios (A&B) can be run and the results viewed side by side.
+This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)
+
+```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
 A.
@@
 Resistance restoration.
+Increasing any of the inputs (going from red to blue) results in shorter times-to-resistance.
 For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Effect of changing resistance starting frequency on changing resistance frequency over time with single insecticide use."}
+```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of the starting frequency of resistance on how resistance frequency changes over time.
+Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds."}
 
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="*KEY FIG OF PAPER* Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
+```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
+Exposure to the insecticides is increased from row 1 (A&B) to row 2 (C&D).
+The effectiveness of one insecticide is increased from column 1 (A&C) to column2 (B&D).
+On the upper X axis s&m indicate where the 50% resistance threshold is reached for the sequence and the mixture.
+In the lower right of each panel the the ratio of time-to-resistance for mixture/sequence is shown rounded to 1 decimal place to give an indication of the relative performance of mixtures and sequences.
 A.
@@
 Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D.
-Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence"}
+Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence.
+Increasing effectiveness increases times-to-resistance for mixtures and improves their performance relative to sequences.
+Increasing exposure decreases times-to-resistance for mixtures and reduces their performance relative to sequences.
+"}
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 #effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 #exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 #eff1 & expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Higher effectivenesses for both insecticides, times-to-resistance longer for the mixture in all scenarios.
+```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
+Effectiveness of insecticide 2 is increased from row 1 (A&B) to row 2 (C&D).
+The effectiveness of insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
+s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
 A.
@@
 Effectiveness of insecticides 1 and 2 increased to 0.8, D.
-Effectiveness of insecticide2 0.8 and of insecticide1 1"}
+Effectiveness of insecticide2 0.8 and of insecticide1 1.
+With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios."}
 
@@
 #eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #eff1 1 eff2 0.5
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #eff1 0.8 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #eff1 1 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
+Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A&B) to row 2 (C&D).
+Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
+s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
+A.
 All control inputs equal at 0.5, B.
@@
 Dominance for insecticide1 increased to 0.8, D.
-Resistance restoration and dominance for insecticide1 increased to 0.8"}
+Resistance restoration and dominance for insecticide1 increased to 0.8.
+Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios."}
 
@@
 #rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 #dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 #dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
 
@@
 
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
+Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A&B) to row 2 (C&D).
+Effectiveness of insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
+s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
+A.
 All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
@@
 Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.001 to 0.0001"}
+Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.01 to 0.001.
+In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness."}
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150 )
+runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
 
@@
 #eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
 
@@
 #freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
 
@@
 #freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','m','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200 )
+runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
 

### 2017-02-12 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-#An accessible model to help understand the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
+#Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
+
+#An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
 
@@
 
+### Declarations
+TODO see https://malariajournal.biomedcentral.com/submission-guidelines/preparing-your-manuscript/research-article
+
+##### Ethics approval and consent to participate
+Not applicable
+##### Consent for publication
+Not applicable
+##### Availability of data and material
+
+##### Competing interests
+
+##### Funding
+
+##### Authors' contributions
+AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
+IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
+
+##### Acknowledgements
+Thankyou to Bethany Levick who developed the first version of the model code.
+
 **Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**

### 2017-02-21 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
add fig 1

@@
 
-![New figure showing how fitness determined by inputs](paper2_mod_graphic_a_to_c.png)
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Diagrammatic representation of the effect of model inputs on the fitness of genotypes for a single insecticide.
+Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
+Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance determines how the fitness for SR sits between that of SS and RR.
+For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance determines how the fitness for SR sits between that of SS and RR.
+In this example effectiveness=0.8, dominance=0.5, resistance restoration=0.5 and resistance cost=0.3."}
+
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance=0.5, cost=0.3 )
+
+```
 
@@
 
-```{r Fig8, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig9, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, male exposure 0.8, B.

### 2017-02-21 R/plot_fit_calc.r

@@ -9,6 +9,8 @@
 # @param yblank whether to remove y axis title & labels
 # @param size text size for RS labels, default 4
 #' 
+#' @import ggplot2 gridExtra
+#' 
 #' @examples 
 #' plot_fit_calc(effectiveness=0.5, resistance_restoration=0.5, dominance=0.5)
 #' 
@@ -16,12 +18,17 @@
 #' @return 
 #' @export
 
-plot_fit_calc <- function ( effectiveness=0.5, resistance_restoration=0.5, dominance=0.5 ){
+plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, dominance=0.5, cost=0.3 ){
  
+  
+  #library(gridExtra)
+  #library(grid)
+
   
   #x axis SS=1, SR=2, RR=3
   #y axis fitness 0 to 1
   
+  # In presence of insecticide
   ss <- 1-effectiveness
   rr <- 1-effectiveness + (effectiveness*resistance_restoration)
   sr <- 1-effectiveness + (effectiveness*resistance_restoration*dominance)
@@ -31,24 +38,32 @@ plot_fit_calc <- function ( effectiveness=0.5, resistance_restoration=0.5, domin
                     y1 = c(1,  rr, ss), 
                     y2 = c(ss, sr, rr))
 
-  gg <- ggplot(dfg, aes(x = x2, y = y2)) +    
+  gg1 <- ggplot(dfg, aes(x = x2, y = y2)) +    
         ylim(0,1) +
         #points
         geom_point(size=3, colour='blue') +
         #arrows
         geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2),size=1, arrow = arrow(length = unit(3,"mm"))) +
-        #dotted line at SS fitness, need to sort how to subset data
-        #geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2))
+        #dotted horizontal lines showing calculations
+        annotate("segment", x='SS', xend='RR', y=ss, yend=ss, linetype="dotted", colour = "blue") +
+        annotate("segment", x='SR', xend='RR', y=rr, yend=rr, linetype="dotted", colour = "blue") +
         ylab('fitness') +
-        #ggtitle(title) + 
-        #scale_x_discrete(labels = c("1"="SS","2"="SR","3"="RR")) +
+        #theme(axis.title.y = element_text(size = rel(3))) +
+        ggtitle("exposed to insecticide") +
         #to set ordering of x axis
[…]

### 2017-02-22 inst/documents/paper2/cost_of_resistance.Rmd
cost of resistance graphs maybe to go in paper2

--- 2017-02-01 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
 ---
-title: "paper2_resistance_mechanisms_mixtures"
+title: "coct_of_resistance"
 output: pdf_document
@@
 date: "`r Sys.Date()`"
-bibliography: library.bib
-csl: biomed-central.csl
-#header-includes:
-#    - \usepackage{setspace}
-#    - \doublespacing
-#    - \usepackage{lineno}
-#    - \linenumbers
 ---
-
-```{r, eval=FALSE, include=FALSE}
-#!!!! paper protocol
-#!!!! might be easier to put text & figs into separate Rmd files after draft stage
-#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
-#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
-#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
-#!!!! submission figs, set this chunk to eval=TRUE
-
-# options to create final publication quality figures
-# causes the document to fail, but the figs do get stored in pap2figs/
-library(knitr)
-opts_chunk$set(dev="tiff",
-               dev.args=list(compression="lzw"),
-               dpi=300,
-               cache=FALSE,
-               fig.path='pap2figs/')
-```
 
@@
 
-### version 9
+```{r Fig cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
 
-#An accessible model to help understand the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
+  #resistance cost
 
-#A tutorial, primer ...
-
-#Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
-
-Target : Malaria journal
-
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and & suggestions welcome
-
-## Abstract
-
-_must be 350 words, currently 375 with subheadings Background, Results & Conclusions_
-
-##### Background
-Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potentially huge consequences for malaria burden.
-To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones.
-There are different potential strategies for using new insecticides and recomendations for how best to preserve their utility for controlling mosquitoes and malaria.
-There is however limited recent, accessible modelling work to support suggestions of which strategies may be best in different situations.
-Previous work suggests that preferred strategies will be situation specific.
-This points to a need for an accessible, mechanistic understanding of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-Such a mechanistic understanding could allow a more robust response to uncertainty in parameter values in field situations.*finesse that sentence or move to discussion*
-
-Here we use a relatively simple model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures.
-The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide, and how much the resistance alleles counteract the effect of the insecticide.
-
-##### Results
-We show firstly that using a second insecticide, in a mixture with a first, always slows evolution of resistance to the first.
-Secondly we show that the ability of an insecticide to kill the susceptible mosquitoes (effectiveness), is the most important factor influencing whether using insecticides in a sequence or a mixture is likely to lead to faster resistance to both of them.
-The mechanisms behind these results are explained.
-We discuss how these results can inform intervention strategies in the field.
-
-##### Conclusions
-We describe an R package and user-interface that allows further exploration of the development of resistance.
-The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.
-
-## Keywords
-insecticide resistance; public health; mosquitoes; vector-borne diseases; infectious diseases; malaria; population genetics
-
-## Background
-
-Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
-Focusing on malaria which still results in hundreds of thousands of deaths per year, recent control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
-Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million malaria cases between 2000 and 2015 [@Bhatt2015].
-A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
-
-There have been modelling studies investigating the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
-[@Curtis1985][@Mani1985][@Roush1989] but much of the modelling work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
-For those who (wisely in some cases) distrust models, work in a moth and genetically modified brocolli system has shown that toxins used in combination can delay the onset of resistance in a way that is consistent with theoretical models [@Zhao2003].
-
-The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
-Three new insecticides are now in development [@IVCC2016] and likely to be available around 2020 [@Ranson2016].
-It is important that decisions about how best to use the new insecticides to delay the onset of resistance are made before insecticides are released [@Hemingway2016].
-The model presented here can help inform strategies to minimise the development of resistance in new and existing insecticides.
-
-Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency.
-We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
-It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
-The changing frequency of insecticide resistance is a population genetic process that can be influenced by (among others) the parameters outlined in Table 1.
-
-_TODO add in here about theory of IR & protocols for resistance management._
-
-We use a 2-locus population genetic model of insecticide resistance [@Levick2017] to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
-In an earlier paper [@Levick2017] we described the technical details of the model and explore large parameter ranges.
-We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance.
-This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
-
-**Table 1. Parameters inflencing the development of insecticide resistance**
-
-Parameter  | Description
-------------------- | -------------------------------------------------------
-1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
-2. Exposure | proportion of insects exposed to the insecticide
-3. Dominance | resistance of an insect with only 1 resistance allele (SR)
-4. Resistance restoration  | ability of resistance (RR) to restore fitness when insect exposed to the insecticide
-5. Frequency  | frequency of resistance alleles within the population
-
-## Methods
-
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
-One locus is represented for each insecticide.
-There are two potential alleles per locus.
-Each allele confers either resistance or susceptibility to the insecticide.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces, with a maximum value of 1 defined as the fitness of the SS genotype in absence of the insecticide.
-The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
-Each of these inputs can be varied as a proportion between 0 and 1.
-We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The more conventional 'selection coefficient' is calculated by multiplying Resistance restoration by insecticide effectiveness.
-Doing it this way allows us to vary all inputs between 0 and 1 without fitness going above 1 in the model.
-
-The simulation proceeds through generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
-This acts to make the fitter alleles more common over time.
-Thus a summary of the selection stage is :
-
-> selection coefficient = resistance restoration * effectiveness  
->
-> Locus fitness : 	
-> SS = 1-effectiveness  
-> RS = 1-effectiveness + (dominance*selection coeff.)  
-> RR = 1-effectiveness + selection coeff.  
-> 
-> Genotype fitness :   
-> SS1SS2 = SS * SS  
-> SS1RS2 = SS * RS  
-> … and for other combinations  
-> RR1RR2 = RR * RR  
-> 
-> Population fitness = Genotype fitness * exposure to insecticide  
-> 
-> In each generation :   
-> Allele frequency[t+1] = Allele frequency[t] * Population fitness
-
-Sexual reproduction is represented first by creating gametes by splitting the adult genomes in two with recombination, and second by forming offspring genotypes by adding gametes.
-
-So far this description just considers a single insecticide and associated resistance allele, with equal behaviour for males and females (Fig 1A).
-To allow males and females to be exposed differently separate genders are also represented (Fig 1B).
-In addition a second insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
-Thus the simulation can represent populations exposed to two insecticides together in a mixture.
-
-Here we principally look at model outputs of the change in resistance allele frequencies over time (generations).
-Resistance allele frequencies are calculated from the frequencies of each genotype as follows.
-
-> Resistance allele frequency = (2*RR + RS) / 2
-
-We investigate three insecticide use strategies.
-
-1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches a threshold of 50%
-1. two insecticides used in a mixture (concurrently)
-
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an online user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
-
-## Results
-
-### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, dominance or resistance restoration resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach the resistance threshold of either 0.5 or 0.75.
-
-Similarly, higher values of starting resistance frequency also resulted in faster spread of resistance (Fig 4) and shorter times to resistance thresholds.
-
-### Two insecticides
-When two insecticides are used the relative performance of a mixture strategy compared to sequential use changes dependent on input values.
-To help understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs individually.
-The base scenario was similar to that used for the single insecticide.
-Values for the first four inputs in Table 1 were set to 0.5 and the starting frequencies of resistance to 0.01.
-For this base scenario resistance arises slower for sequential than mixture (Fig 5A).
-This plot, and those that follow, compares mixture and sequence for a single combination of inputs.
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold of 50%.
-
-#### results: Insecticide exposure and effectiveness
-
-Increasing exposure to the insecticides (the proportion of insects that come into contact) from 0.5 to 0.8 (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy and so does not change their relative performance; time-to-resistance remains longer for the sequence.
-In contrast keeping the exposure constant at 0.5 and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides to 0.8 (Fig 5D) results in a longer time-to-resistance for the mixture relative to the sequence.
-Resistance for the more effective insecticide (in red) increases faster than the less effective.
-Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effecttive insecticide has reached the resistance threshold (~50 generations).
-
-Thus a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
-This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-Figure 5D has the same inputs as Fig 5B except that the exposure is increased from 0.5 to 0.8.
-The increase in exposure reduces time-to-resistance more for the mixture relative to the sequence.
-The result is that the advantage of mixtures over sequence is removed by the increase in exposure, leading to approximately equal times-to-resistance for the two strategies (Fig 5D).
-
-Figure 6 shows how increasing the effectiveness of either or both insecticides improves the performance of mixtures over sequences in terms of time-to-resistance.
-Figure 6A uses the same inputs as Fig 5B.
-From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
-
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
-
-#### results: Dominance and resistance restoration
-Increasing dominance or resistance restoration decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
-The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
-
-#### results: Starting frequencies of resistance
-Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
-Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
-
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
-
-## Discussion
-
-The response of resistance to insecticide use in the model can be explained by mechanisms that can help us develop and communicate a more robust understanding of how we would expect resistance to respond in the field.
-
-When single insecticide use was represented (Figs 3-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
-Higher values of the five inputs investigated all led to more rapid increases in resistance.
-For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
-This makes sense as vector kill is effectively a product of exposure times effectiveness.
-For example exposing 50% of a  population to an insecticide which is 75% effective would be expected to have the same result as exposing 75% of the population to one which is 50% effective.
-This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance.
-In both cases the increased deaths of susceptible vectors cause a higher selection pressure that can explain the faster increase.
-
-The pattern of more rapid increase in resistance at higher dominance levels (Fig 3C) can be explained by increased survival of heterozygotes in the presence of the insecticide.
-Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival.
-Thus selection pressure for the resistance allele will be increased because it confers more of an advantage when only present on one chromosome.
-The faster development of resistance under higher resistance-restoration (Fig 3D) can be explained by it's effect on the survival of the resistant genotypes.
-Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.
-The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs.
-Simply, when starting from a higher frequency of resistance there is a smaller change to make to reach the higher resistance thresholds.
-
-These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Exposure and effectiveness act by decreasing the survival of all genotypes, dominance acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
-
-#### discussion: Two insecticides
-
-In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
-Looking at our base scenario (Fig 5A) the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
-The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Thus each insecticide reduces the selection pressure for the increase in resistance to the other and could be said to 'protect' the other.
-However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
-As shown in our base scenario, resistance can arise more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-
-#### discussion: Insecticide effectiveness
-
-The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (compare Fig 5B to 5A).
-In contrast in both the sequential (Fig 5B) and single insecticide (Fig 2A) scenarios the development of resistance was faster when effectiveness was higher.
-This can be explained by the increased effectiveness of one insecticide in a mixture increasing killing of vectors that are resistant to the other and thus decreasing selection pressure.
-
-This effect can be seen by comparing Fig 5B, where the effectiveness of insecticide 1 has been increased, to Fig 5A.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide and thus stops limiting the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide.
-The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-
-#### discussion: Exposure vs effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 110 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference between effectiveness and exposure was unexpected for us.
-_todo could this be because exposure results doesn't alter the relative fitness of genotypes because all genotypes are exposed equally ? doesn't quite make sense to me_
-
-The mechanisms involved in the response of resistance in mixtures are summarised in Table3.
-Whether a mixture or sequence is favoured for reducing the spread of resistance will depend on these mechanisms and is summarised in Table 4.
-
-#### discussion: caveats
-It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria.
-As such our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining that in the longer term by restricting the development of resistance.
-Others are looking at combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
-
-The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ransom2016].
-The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of paramter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-Thus there is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-In addition we provide an online user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown here.
-
-#### discussion: putting results into context
-
-I will now put these predicted responses of resistance into the context of previous work and recomendations.
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and the particular situations they represent.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D).
-
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to the parameter we have termed insecticide effectiveness, so lends further support to our main conclusion of the importance of effectiveness.
-It's also shown [@Roush1998 Fig 3] that decreasing exposure from 90% to 80% also increases the benefit of mixtures over sequences in the same way that we have shown.
-
-Modelling and theory have been used to generalise Insecticide Resistance Management into three potential strategies [@Georghiou1994].
-These are A) moderation : preserving susceptible genes by limiting selection pressure, B) saturation : high dose so that heterozygous resistants are killed and C) multiple attack : independently acting pressures neither of which is strong enough to lead to resistance.
-Our work here principally focuses on multiple attack but the other strategies are relevant too.
-_todo expand this paragraph_
-
-One consistent aspect of the results presented here can at first seem to go against current advice on good practice.
-We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-In contrast the widely held view seems to be that low effectiveness or exposure associated with poor implementation can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label; - poor coverage of the area being treated ...".
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?ref].
-Our model does show that increasing dominance leads to faster spread of resaistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Our model cobined with laboratory and field results could help to assess these trade-offs.
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the devlopment of resistance.
-
-## FINAL CONCLUDING PARAGRAPH
-
-~ Effectiveness and exposure are key and can be modified.
-~ The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies.
-
-### other papers to incorporate into discussion above
-
-Roush older paper : [@Roush1989]
-
-Relevance of these mechanisms in the field e.g.
-exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
-[@Birget2015a] mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
-_TODO check which of the Birget papers mention this_
-
-How do the rest of our results relate to recomendations that have been produced by WHO[@WHO2012], IRAC[@IRAC2011] and FAO[@FAO2012] ?
-
-[@Taillebois2016]
-Laboratory experiments on aphids showing that in low dose mixtures insecticides most often had additive effects as represented in our model, but that also commonly mortality was higher than expected (synergistic) which we can also represent within the model.
-
-[@Barbosa2011] 
-_suggests incomplete dominance in the field_
-
-[@Birget2015] 
-We show that indoor use of insecticides leads to less selection pressure than their use as larvicides. 
-Reasons for relatively low selection pressure by adulticides
-(i) males are not affected by the ITNs
-(ii) insecticides are also repellents, keeping mosquitoes at bay from contacting the insecticide but also driving them to bite either people who do not use the insecticide or alternative hosts.
-
-[@Curtis1998]
-can anything be done to maintain effectiveness of nets ?
-
-[@Kliot2012]
-fitness costs associated with insecticide resistance
-
-Good review of tactics for IRM inlcuding mention of mixtures and new AIs [@Denholm1992]
-"Once developed and marketed, new products introduce a fresh challenge how to exploit their characteristics while restricting exposure to preserve their effectiveness."
-
-### Discussion some summary points probably already covered ...
-
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
-Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
-
-(@) For a mixture of 2 insecticides with differing effectiveness.
-Resistance to the insecticide with a greater effectiveness increases faster.
-The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
-
-(@) exposure ...
-
-**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
-
-Parameter to increase  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | faster | increased deaths of susceptibles
-2. Exposure | faster | increased deaths of susceptibles
-3. Dominance | faster | increased survival of heterozygotes
-4. Resistance restoration | faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-
-**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
-
-Parameter to increase  | effect on resistance | Mechanism
------------------- | --------- | ---------------------------------------------------
-1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
-2.
-Exposure | faster (but less than for single) | increased deaths of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by higher deaths of resistant individuals caused by the other insecticide.
-3. Dominance | faster | increased survival of resistants
-4. Resistance restoration | faster | increased survival of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-
-**Table 4. Effect of inputs on the difference between mixture and sequential use**
-
-Parameter to increase  | increase favours mix or sequence | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
-2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4. Resistance restoration | neither | As for dominance. Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
-5.
-Frequency  | neither | As for dominance and resistance restoration.
-Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
-
-## ANDY TO DO
-1. remove first person for Malaria journal
-1. explain resistance thresholds and time-to-resistance in methods
-1. refer to ratios of mix:seq in plots
-1. refer to PRCC from pap1 in ref. to Table4
-1. mention single gene vs additive effects (with ref.)
-1. can I add a diagram to help with methods, e.g. perhaps fitness under difference inputs, similar to fitnessvis ?
-1. in discussion refer to linkage disequilibrium about Fig 8D
-1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage, or by the dose ?
-1. mention other potential effects e.g. costs of resistance, sex linkage, ..._
-1. cite in press Killeen about combining interventions
-
-\pagebreak
-
-## Figures
-
-![New figure showing how fitness determined by inputs](paper2_mod_graphic_a_to_c.png)
-
-![Screenshot of one online model user interface, accessible at : https://andysouth.shinyapps.io/resistmob2/.
-The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
-Two scenarios (A&B) can be run and the results viewed side by side.
-This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)
-
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
-A.
-Effectiveness, B.
-Exposure, C.
-Dominance, D.
-Resistance restoration.
-Increasing any of the inputs (going from red to blue) results in shorter times-to-resistance.
-For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
-"}
-
-  library(cowplot)
-
-  #list to put plots in for cowplot
-  plotlist <- list(4)
-
-  #A effectiveness  
   a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
@@
   i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
                       h.RS1_A0 = 0.5, 
@@
                       rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.8,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
-         theme_bw() +  
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +  
-         labs(colour = "effectiveness") +
-         coord_trans(y = "log10") +    
-         geom_line()
-
-  #B exposure
-  exposure <- 0.2
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.5
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.8
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "exposure", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-  #C single insecticide dominance
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.8, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #C 
-  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #D single insecticide resistance restoration
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.2,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.8,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #D  
-  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "resistance restoration", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of the starting frequency of resistance on how resistance frequency changes over time.
-Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds."}
-
-  #single insecticide frequency fig
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.001,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0,
+                      h.RS1_00 = 0.1)
 
@@
                       rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.25,
+                      h.RS1_00 = 0.1 )
 
   i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.1,       
+                      P_1 = 0.01,       
                       h.RS1_A0 = 0.5, 
@@
                       rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.75,
+                      h.RS1_00 = 0.1 )
 
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
          theme_bw() +
@@
          guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "start\nfrequency") +
+         labs(colour = "cost of\nresistance") +
          coord_trans(y = "log10") + 
          geom_line()   
+       )
 
-       )
 ```
 
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
-Exposure to the insecticides is increased from row 1 (A&B) to row 2 (C&D).
-The effectiveness of one insecticide is increased from column 1 (A&C) to column2 (B&D).
-On the upper X axis s&m indicate where the 50% resistance threshold is reached for the sequence and the mixture.
-In the lower right of each panel the the ratio of time-to-resistance for mixture/sequence is shown rounded to 1 decimal place to give an indication of the relative performance of mixtures and sequences.
-A.
-All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C.
-Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D.
-Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence.
-Increasing effectiveness increases times-to-resistance for mixtures and improves their performance relative to sequences.
-Increasing exposure decreases times-to-resistance for mixtures and reduces their performance relative to sequences.
-"}
+Oooo cost-of-resistance stops resistance reaching fixation, I hadn't expected that !
+e.g.
+even with cost set fairly low at 0.25 and other params at 0.5 the resistance frequency stabilises at 50%.
+This is with dominance at 0.1.
+Seems like low dominance around 0.1 or lower may be necessary to maintain the resistance frequency below 1.
+So this might be quite an unusual result.
+
+What happens if I modify dominance ?
+
+```{r Fig dominance of cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+
+  #dominance cost
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.25,
+                      h.RS1_00 = 0)
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.25,
+                      h.RS1_00 = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.25,
+                      h.RS1_00 = 1 )
+
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_00))) + 
+         theme_bw() +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) + 
+         labs(colour = "dominance\nof cost") +
+         coord_trans(y = "log10") + 
+         geom_line()   
+       )
+
+```
+
+If dominance of cost is set > 0.5 with cost at 0.25 then the frequency of resistance declines.
+
+Try changing dominance when cost is lower :
+
+```{r Fig dominance of cost at low cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+
+  #dominance cost at low cost
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.1,
+                      h.RS1_00 = 0)
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.1,
+                      h.RS1_00 = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.1,
+                      h.RS1_00 = 1 )
+
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_00))) + 
+         theme_bw() +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) + 
+         labs(colour = "dominance\nof cost") +
+         coord_trans(y = "log10") + 
+         geom_line()   
+       )
+
+```
+
+Ah good, this now does what I would expect.
+So maybe it was working correctly before too.
+Here higher dominance of cost gives longer times-to-resistance.
+Higher dominance of cost makes the resistance costlier therefore reducing the selection pressure for it and increasing time-to-resistance.
+
+And in all these with cost set to 0.1, resistance does proceed to fixation, where when set to 0.25 resistance equilibriated at 50%.
+
+What happens if I fix dominance of cost at 0.5 and try costs of 0, 0.1, 0.2
+
+```{r Fig low costs of resistance, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+
+  #low costs at dominance 0.5
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0,
+                      h.RS1_00 = 0.5)
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.2,
+                      h.RS1_00 = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.25,
+                      h.RS1_00 = 0.5 )
+
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
+         theme_bw() +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) + 
+         labs(colour = "cost of\nresistance") +
+         coord_trans(y = "log10") + 
+         geom_line()   
+       )
+
+```
+
+```{r Fig cost on effectiveness, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="look at cost & see if it disrupts effect of effectiveness, keep A & B same as before, copy to C&D and add cost"}
 
@@
 
-#testing
-# plot(1)
-# plot(1)
-# plot(1)
-# plot(1)
-
 #label
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
 
@@
 #effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('cost 1 = 0.1', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
 
@@
 
-#eff1 & expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
+#eff1  and  expos 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
 
@@
 
+Aha! note here how cost of just 0.15 nullifies the benefit of the mixture in C.
+
+Note that cost effects happen later when the frequency of resistance has built up.
+
 ```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
-Effectiveness of insecticide 2 is increased from row 1 (A&B) to row 2 (C&D).
-The effectiveness of insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
-s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
+Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
+The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
+The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
 A.
@@
 ```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
-Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A&B) to row 2 (C&D).
-Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
-s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
+Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
+Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
+The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
 A.
@@
 ```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
-Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A&B) to row 2 (C&D).
-Effectiveness of insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
-s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
+Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
+Effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
+The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
 A.

### 2017-02-22 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

--- 2017-02-12 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
 #!!!! paper protocol
-#!!!! might be easier to put text & figs into separate Rmd files after draft stage
+#!!!! might be easier to put text and figs into separate Rmd files after draft stage
 #!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
@@
 
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and & suggestions welcome
+THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and suggestions welcome
 
@@
 
-_must be 350 words, currently 375 with subheadings Background, Results & Conclusions_
+_must be 350 words, currently 375 with subheadings Background, Results and Conclusions_
 
@@
 
-_TODO add in here about theory of IR & protocols for resistance management._
+_TODO add in here about theory of IR and protocols for resistance management._
 
@@
 These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
+
+The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 1.
+In the model itself this is repeated for both insecticides and the fitnesses multiplied.
+Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS and resistance restoration 'restores' a portion of the fitness for RR.
+A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
+Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
+For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
+Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Figure 1).
+
 Each of these inputs can be varied as a proportion between 0 and 1.
 We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The more conventional 'selection coefficient' is calculated by multiplying Resistance restoration by insecticide effectiveness.
-Doing it this way allows us to vary all inputs between 0 and 1 without fitness going above 1 in the model.
+The more conventional 'selection coefficient' is calculated by multiplying resistance restoration by insecticide effectiveness.
 
@@
 This acts to make the fitter alleles more common over time.
-Thus a summary of the selection stage is :
+Thus a summary of the selection stage is as follows with 1 & 2 representing the two insecticides :
 
@@
 >
-> Locus fitness : 	
+> Single gene, insecticide fitness : 	
 > SS = 1-effectiveness  
@@
 > 
-> Genotype fitness :   
-> SS1SS2 = SS * SS  
-> SS1RS2 = SS * RS  
+> Double gene, insecticide fitness :   
+> SS1SS2 = SS1 * SS2  
+> SS1RS2 = SS1 * RS2  
 > … and for other combinations  
-> RR1RR2 = RR * RR  
+> RR1RR2 = RR1 * RR2  
 > 
-> Population fitness = Genotype fitness * exposure to insecticide  
+> Population fitness = Double fitness * exposure to insecticides  
 > 
@@
 
-The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ransom2016].
+The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ranson2016].
 The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of paramter space (e.g.
@@
 
-![New figure showing how fitness determined by inputs](paper2_mod_graphic_a_to_c.png)
-
-![Screenshot of one online model user interface, accessible at : https://andysouth.shinyapps.io/resistmob2/.
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Diagrammatic representation of the effect of model inputs on the fitness of genotypes for a single insecticide.
+Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
+Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of restoration determines how the fitness for SR sits between that of SS and RR.
+For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
+In this example effectiveness=0.8, dominance of restoration=0.4, resistance restoration=0.5, dominance of cost=0.4 and resistance cost=0.3."}
+
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.4, dominance_cost=0.4, cost=0.3 )
+
+```
+
+![Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
 The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
-Two scenarios (A&B) can be run and the results viewed side by side.
+Two scenarios (A and B) can be run and the results viewed side by side.
 This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)
@@
          geom_line()   
-
        )
+
 ```
@@
 ```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
-Exposure to the insecticides is increased from row 1 (A&B) to row 2 (C&D).
-The effectiveness of one insecticide is increased from column 1 (A&C) to column2 (B&D).
-On the upper X axis s&m indicate where the 50% resistance threshold is reached for the sequence and the mixture.
+Exposure to the insecticides is increased from row 1 (A and B) to row 2 (C and D).
+The effectiveness of one insecticide is increased from column 1 (A and C) to column2 (B and D).
+On the upper X axis 's' and 'm' indicate where the 50 percent resistance threshold is reached for the sequence and the mixture.
 In the lower right of each panel the the ratio of time-to-resistance for mixture/sequence is shown rounded to 1 decimal place to give an indication of the relative performance of mixtures and sequences.
@@
 
-#eff1 & expos 0.8
+#eff1  and  expos 0.8
 runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
@@
 ```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
-Effectiveness of insecticide 2 is increased from row 1 (A&B) to row 2 (C&D).
-The effectiveness of insecticide 1 is increased from column 1 (A&C) to column2 (B&D).
-s&m on the upper X axis and mix/seq in th lower right of each panel are explained in the legend to Fig 5.
+Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
+The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
+The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
 A.

### 2017-02-22 R/plot_fit_calc.r

@@ -2,9 +2,11 @@
 #' 
 #' accepts inputs
 
-#' @param effectiveness
+#' @param effectiveness insecticide effectiveness
 #' @param resistance_restoration
-#' @param dominance
+#' @param dominance_restoration
+#' @param dominance_cost
+#' @param cost cost of resistance
 #' @param title optional title for the plot
 # @param yblank whether to remove y axis title & labels
 # @param size text size for RS labels, default 4
@@ -18,7 +20,7 @@
 #' @return 
 #' @export
 
-plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, dominance=0.5, cost=0.3 ){
+plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.5, dominance_cost=0.5, cost=0.3 ){
  
   
   #library(gridExtra)
@@ -31,7 +33,7 @@ plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, domin
   # In presence of insecticide
   ss <- 1-effectiveness
   rr <- 1-effectiveness + (effectiveness*resistance_restoration)
-  sr <- 1-effectiveness + (effectiveness*resistance_restoration*dominance)
+  sr <- 1-effectiveness + (effectiveness*resistance_restoration*dominance_restoration)
   
   dfg <- data.frame(x1 = c('SS',  'SR',  'RR'), 
                     x2 = c('SS',  'SR',  'RR'),                    
@@ -43,7 +45,7 @@ plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, domin
         #points
         geom_point(size=3, colour='blue') +
         #arrows
-        geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2),size=1, arrow = arrow(length = unit(3,"mm"))) +
+        geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2),size=0.9, arrow = arrow(length = unit(3,"mm"))) +
         #dotted horizontal lines showing calculations
         annotate("segment", x='SS', xend='RR', y=ss, yend=ss, linetype="dotted", colour = "blue") +
         annotate("segment", x='SR', xend='RR', y=rr, yend=rr, linetype="dotted", colour = "blue") +
[…]

### 2017-03-07 inst/documents/paper2/cost_of_resistance.Rmd
paper2 and cost

@@
 ---
-title: "coct_of_resistance"
+title: "cost_of_resistance"
 output: pdf_document
@@
 
-```{r Fig cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+```{r Fig cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost.
+Cost-of-resistance can stop resistance reaching fixation, I hadn't expected that, e.g.
+even with cost set fairly low at 0.25 and other params at 0.5 the resistance frequency stabilises at 0.5.
+This is with dominance of cost at 0.1.
+Seems like low dominance around 0.1 or lower may be necessary to maintain the resistance frequency below 1, otherwise it results in decline in resistance.
+All other inputs set to 0.5."}
 
@@
 
-Oooo cost-of-resistance stops resistance reaching fixation, I hadn't expected that !
-e.g.
-even with cost set fairly low at 0.25 and other params at 0.5 the resistance frequency stabilises at 50%.
-This is with dominance at 0.1.
-Seems like low dominance around 0.1 or lower may be necessary to maintain the resistance frequency below 1.
-So this might be quite an unusual result.
-
-What happens if I modify dominance ?
-
-```{r Fig dominance of cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+```{r Fig dominance of cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap=" If dominance of cost is set greater than 0.5 with cost at 0.25 then the frequency of resistance declines."}
 
@@
 
-If dominance of cost is set > 0.5 with cost at 0.25 then the frequency of resistance declines.
-
-Try changing dominance when cost is lower :
-
-```{r Fig dominance of cost at low cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+```{r Fig dominance of cost at low cost, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of dominance of cost.
+Cost set to 0.1.
+Higher dominance of cost gives longer times-to-resistance.
+Higher dominance of cost makes the resistance costlier therefore reducing the selection pressure for it and increasing time-to-resistance.
+And in all these with cost set to 0.1, resistance does proceed to fixation, where when set to 0.25 resistance equilibriated at 0.5."}
 
@@
 
-Ah good, this now does what I would expect.
-So maybe it was working correctly before too.
-Here higher dominance of cost gives longer times-to-resistance.
-Higher dominance of cost makes the resistance costlier therefore reducing the selection pressure for it and increasing time-to-resistance.
-
-And in all these with cost set to 0.1, resistance does proceed to fixation, where when set to 0.25 resistance equilibriated at 50%.
-
-What happens if I fix dominance of cost at 0.5 and try costs of 0, 0.1, 0.2
-
-```{r Fig low costs of resistance, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of resistance cost."}
+```{r Fig low costs of resistance, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Dominance of cost 0.5."}
 
@@
                       rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.2,
+                      z.RR1_00 = 0.05,
                       h.RS1_00 = 0.5 )
@@
                       rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.25,
+                      z.RR1_00 = 0.15,
                       h.RS1_00 = 0.5 )
@@
 
-```{r Fig cost on effectiveness, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="look at cost & see if it disrupts effect of effectiveness, keep A & B same as before, copy to C&D and add cost"}
+```{r Fig cost of 1 resistance on effectiveness, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Seeing if cost disrupts effect of effectiveness, A and B same as Fig5 in paper2.
+Cost for one insecticide set to 0.15 in C and D.
+Note here how cost nullifies the benefit of the sequence in C.
+Note also that cost effects happen later when the frequency of resistance has built up.
+Costs seem to favour mixtures in these plots.
+This makes sense, costs reduce selection pressure on resistance for one insecticide, keeping its resistance down and allowing it to provide more protection for the other in a mixture.
+BUT also be aware that costs would probably lead to a decline in resistance for the first insecticide in a sequence when it stops being used."}
 
@@
 #exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.1 )
 
@@
 #eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.1  )
 
@@
 
-Aha! note here how cost of just 0.15 nullifies the benefit of the mixture in C.
+```{r Fig cost of both resistances on effectiveness, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="As above but costs for resistances to both insecticides set to 0.15."}
 
-Note that cost effects happen later when the frequency of resistance has built up.
-
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
-Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-Effectiveness of insecticide1 increased from 0.5 to 0.8, B.
-Effectiveness of insecticide1 increased from 0.5 to 1, C.
-Effectiveness of insecticides 1 and 2 increased to 0.8, D.
-Effectiveness of insecticide2 0.8 and of insecticide1 1.
-With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios."}
-
-#4 panel figure looking at effectiveness changes alone
+#4 panel figure looking at effectiveness and exposure
 
 #test of rearranging figures on a grid
-#eff1  0.8  1
-#eff2
+#eff1  0.5  0.8
+#exp
 #0.5    1   2
 #0.8    3   4
+
+#layout(matrix(1:4,2,2, byrow = TRUE))
+
+#add lower panel for legend
+#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
+
+#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
+# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
+#layout.show(n=5)
 
@@
 
+#effectiveness on mixture
+
 #label
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 350, labelMixSeqRatio = 1 )
 
@@
 
-#eff1 1 eff2 0.5
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
+#effectiveness1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 350, labelMixSeqRatio = 1 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('cost 1&2 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-#eff1 0.8 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
+#exposure 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 350, labelMixSeqRatio = 1, z.RR1_00 = 0.15, z.RR2_00 = 0.15 )
 
@@
 
-#eff1 1 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
-Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
-Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, B.
-Resistance restoration for insecticide1 increased from 0.5 to 0.8, C.
-Dominance for insecticide1 increased to 0.8, D.
-Resistance restoration and dominance for insecticide1 increased to 0.8.
-Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios."}
-
-#4 panel figure looking at dominance and resistance restoration
-
-#rr1  0.5  0.8
-#dom1
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#testing
-# plot(1)
-# plot(1)
-# plot(1)
-# plot(1)
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
-Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
-Effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.01 to 0.001.
-In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness."}
-
-#4 panel figure looking at effectiveness and starting freq.
-
-#eff1  0.5  0.8
-#strt freq1
-#0.01    1   2
-#0.001    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
+#eff1  and  expos 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 350, labelMixSeqRatio = 1, z.RR1_00 = 0.15, z.RR2_00 = 0.15  )
 
@@
 ```
-
-```{r Fig9, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
-All control inputs equal at 0.5, male exposure 0.8, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Male exposure decreased from 0.8 to 0.2, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
-
-# not evaluated at the moment
-#4 panel figure looking at male exposure prop and starting freq.
-
-#eff1  0.5  0.8
-#maleprop
-#0.8    1   2
-#0.2    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure =0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-\pagebreak
-
-## References
-<!-- now auto created here-->

### 2017-03-07 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 and cost

@@
 2. Exposure | proportion of insects exposed to the insecticide
-3. Dominance | resistance of an insect with only 1 resistance allele (SR)
-4. Resistance restoration  | ability of resistance (RR) to restore fitness when insect exposed to the insecticide
+3. Resistance restoration  | ability of resistance (RR) to restore fitness when insect exposed to the insecticide
+4. Dominance of resistance | multiplier of resistance restoratio for heterozygous (SR) insects
 5. Frequency  | frequency of resistance alleles within the population
+6. Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide
+7. Dominance of cost  | multiplier of cost for heterozygous (SR) insects in absence of insecticide
 
@@
 It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria.
-As such our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Using no insecticide could have serious consequences for mosquito and malaria abundance.
+Different insecticide use strategies, as well as effecting the evolution of resistance, are also likely to influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
+Insecticide resistance could reduce vector transmission of disease (e.g.
+by reduced lifespan of the vector) or increase transmission (e.g.
+by reduced immunity of the vector to the disease) [@Rivero2010].
+Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
+Of course, using no insecticide could have serious consequences for mosquito and malaria abundance.
 The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining that in the longer term by restricting the development of resistance.
@@
 ~ The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies. 
+~ This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
+Other factors also need to be considered.
+On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
+On the negative side a mixture is more expensive.
 
@@
 Exposure, C.
-Dominance, D.
+Dominance of resistance, D.
 Resistance restoration.
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=3, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of the starting frequency of resistance on how resistance frequency changes over time.
-Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds."}
+```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
+A.
+Starting frequency, B.
+Cost of resistance.
+Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds.
+Increasing costs of resistance increases time-to-resistance."}
 
@@
 
+  #A starting frequency
   a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
@@
 
-  print( ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
+  #A  
+  plotlist[[1]] <-  ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
          theme_bw() +
@@
          geom_line()
-       )
+
+  #B changing costs at dominance 0.5
+
+  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
+
+  i1 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0,
+                      h.RS1_00 = 0.5)
+
+  i2 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.1,
+                      h.RS1_00 = 0.5 )
+
+  i3 <- setInputOneScenario( max_gen = 500,
+                      P_1 = 0.01,       
+                      h.RS1_A0 = 0.5, 
+                      h.RS2_0B = 0.5,
+                      a = a,
+                      phi.SS1_A0 = 0.5,
+                      phi.SS2_0B = 0.5,
+                      rr_restoration_ins1 = 0.5,
+                      rr_restoration_ins2 = 0.5,
+                      z.RR1_00 = 0.2,
+                      h.RS1_00 = 0.5 )
+
+  input <- cbind(i1,i2,i3)
+
+  listOut <- runModel2( input )
+
+  df_resist <- get_resistance(locus=1, listOut)
+
+  #B  
+  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
+         theme_bw() +
+         theme(legend.position = "bottom", legend.key = element_blank()) + 
+         guides(colour = guide_legend(reverse=TRUE)) + 
+         labs(colour = "cost of\nresistance") +
+         coord_trans(y = "log10") + 
+         geom_line()
+
+  cowplot::plot_grid( plotlist[[1]],plotlist[[2]], ncol=2, rel_widths=c(1,1), labels='AUTO')  
+  #cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
 

### 2017-03-07 R/plot_fit_calc.r

@@ -2,10 +2,10 @@
 #' 
 #' accepts inputs
 
-#' @param effectiveness insecticide effectiveness
-#' @param resistance_restoration
-#' @param dominance_restoration
-#' @param dominance_cost
+#' @param effectiveness insecticide effectiveness 0-1
+#' @param resistance_restoration resistance restoration 0-1
+#' @param dominance_restoration dominance of resistance restoration 0-1
+#' @param dominance_cost dominance of resistance cost 0-1
 #' @param cost cost of resistance
 #' @param title optional title for the plot
 # @param yblank whether to remove y axis title & labels
@@ -50,6 +50,7 @@ plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, domin
         annotate("segment", x='SS', xend='RR', y=ss, yend=ss, linetype="dotted", colour = "blue") +
         annotate("segment", x='SR', xend='RR', y=rr, yend=rr, linetype="dotted", colour = "blue") +
         ylab('fitness') +
+        xlab('genotype') +
         #theme(axis.title.y = element_text(size = rel(3))) +
         ggtitle("exposed to insecticide") +
         #to set ordering of x axis
@@ -59,14 +60,14 @@ plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, domin
         # annotate("text", x = 'SR', y = sr+(rr-sr)/2, hjust=0, label = " dominance") + 
         # annotate("text", x = 'RR', y = ss+(rr-ss)/2, hjust=0, label = " resistance\n  restoration") + 
         annotate("label", x = 'SS', y = ss+(1-ss)/2, size=3, label = " effectiveness") +
-        annotate("label", x = 'SR', y = sr+(rr-sr)/2, size=3, label = " dominance of\n  restoration") + 
+        annotate("label", x = 'SR', y = sr+(rr-sr)/2, size=3, label = " 1-dominance of\n  restoration") + 
         annotate("label", x = 'RR', y = ss+(rr-ss)/2, size=3, label = " resistance\n  restoration") + 
     
         theme_bw() +
         theme(
               axis.title.y = element_text(size = rel(1.5)),
               axis.text.x = element_text(size = rel(1.5)),
-              axis.title.x = element_blank(),
+              #axis.title.x = element_blank(),
               #axis.line.x = element_blank(),
[…]

### 2017-03-08 inst/documents/paper2/cost_of_resistance.Rmd
paper2 writing

@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 #effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
 
@@
 plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0.1', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+mtext('cost 1 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
 
@@
 #exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
 
@@
 #eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1, z.RR1_00 = 0.1  )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
 

### 2017-03-08 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

@@
 
-### version 9
-
-#Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
-
-#An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
-
-#A tutorial, primer ...
-
-#Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
+### version 10
+
+#Insecticide resistance evolution with mixtures and sequences : a graphical primer.
+
+###Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
+
+###An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
+
+###A tutorial, primer ...
+
+###Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
 
@@
 Parameter  | Description
-------------------- | -------------------------------------------------------
+---------------------- | ----------------------------------------------------
 1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
@@
 3. Resistance restoration  | ability of resistance (RR) to restore fitness when insect exposed to the insecticide
-4. Dominance of resistance | multiplier of resistance restoratio for heterozygous (SR) insects
+4. Dominance of restoration | multiplier of resistance restoration for heterozygous (SR) insects
 5. Frequency  | frequency of resistance alleles within the population
@@
 ### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, dominance or resistance restoration resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach the resistance threshold of either 0.5 or 0.75.
-
-Similarly, higher values of starting resistance frequency also resulted in faster spread of resistance (Fig 4) and shorter times to resistance thresholds.
+For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
+
+Similarly, higher values of starting resistance frequency also resulted in faster spread of resistance (Fig 4A) and shorter times to resistance thresholds.
+Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect with higher values leading to a slower development of resistance.
 
@@
 
+#### results: Costs of resistance
+Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
+The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 9C).
+The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C.
+With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 9A and 1 in Fig 9C.
+At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
+
 It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
@@
 
-The pattern of more rapid increase in resistance at higher dominance levels (Fig 3C) can be explained by increased survival of heterozygotes in the presence of the insecticide.
+The pattern of more rapid increase in resistance at higher dominance of restoration levels (Fig 3C) can be explained by increased survival of heterozygotes in the presence of the insecticide.
 Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival.
@@
 These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Exposure and effectiveness act by decreasing the survival of all genotypes, dominance acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
+Exposure and effectiveness act by decreasing the survival of all genotypes, dominance of restoration acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
 
@@
 Whether a mixture or sequence is favoured for reducing the spread of resistance will depend on these mechanisms and is summarised in Table 4.
+
+#### discussion costs of resistance
+Costs of resistance favour mixtures in these analyses.
+This is because the evolution of a resistance with an associated cost is slowed more in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
+Cost of resistance to one insecticide reduces selection pressure for that gene and this effect becomes more as the frequency of resistance increases.
+Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
+Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
+Also note that if the costs are set even higher than we set them here (e.g.
+above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
+We do not expect such high costs to occur in operational settings.
+
+However, also be aware that in our analyses we have used a single resistance threshold.
+When there are costs of resistance, the frequency of resistance for insectides that are not being used is likely to decline.
+Thus the frequency of resistance for the first insecticide in a sequence may decline to a level where it becomes operationally useful again.
+As such a repeated sequential strategy could offer benefits over mixtures.
 
@@
 Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?ref].
-Our model does show that increasing dominance leads to faster spread of resaistance (Fig 3C, 7).
+Our model does show that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
 So, the effect of poor implementation of an insecticide intervention on resistance could be either positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Our model cobined with laboratory and field results could help to assess these trade-offs.
+Our model combined with laboratory and field results could help to assess these trade-offs.
 
@@
 
-~ Effectiveness and exposure are key and can be modified.
+~ Effectiveness and exposure (and cost?) are key and can be modified.
 ~ The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies. 
@@
 
-Parameter to increase  | effect on resistance | Mechanism
------------------- | ------- | ---------------------------------------------------
-1. Effectiveness | faster | increased deaths of susceptibles
-2. Exposure | faster | increased deaths of susceptibles
-3. Dominance | faster | increased survival of heterozygotes
-4. Resistance restoration | faster | increased survival of resistants
+Parameter to increase  | effect on resistance evolution | Mechanism
+------------------ | -------- | --------------------------------------------------
+1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
+2. Exposure | faster | reduced fitness of SS overall
+3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
+4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
 5. Frequency  | faster | less change needed to reach resistance threshold
+6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
 
@@
 
-Parameter to increase  | effect on resistance | Mechanism
------------------- | --------- | ---------------------------------------------------
-1. Effectiveness | **slower** | increased deaths of individuals resistant to the other insecticide
+Parameter to increase  | effect on resistance evolution | Mechanism
+------------------ | ---------- | --------------------------------------------------
+1. Effectiveness | **slower** | reduced fitness of individuals resistant to the other insecticide
 2.
-Exposure | faster (but less than for single) | increased deaths of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by higher deaths of resistant individuals caused by the other insecticide.
-3. Dominance | faster | increased survival of resistants
-4. Resistance restoration | faster | increased survival of resistants
+Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
+However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide.
+3. Dominance of restoration | faster | increased fitness of heterozygotes
+4. Resistance restoration | faster | increased fitness of resistants
 5. Frequency  | faster | less change needed to reach resistance threshold
+6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
 
@@
 2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4. Resistance restoration | neither | As for dominance. Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
+3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
+4.
+Resistance restoration | neither | As for dominance of restoration.
+Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
 5.
@@
 Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
+6.
+Cost of resistance | mixture | Higher costs slow the evolution of resistance more in a mixture than when used in sequence (but with higher costs in a sequence there is a greater chance for resistance levels to decline for the insecticide not being used).
 
 ## ANDY TO DO
-1. remove first person for Malaria journal
 1. explain resistance thresholds and time-to-resistance in methods
@@
 1. mention single gene vs additive effects (with ref.)
-1. can I add a diagram to help with methods, e.g. perhaps fitness under difference inputs, similar to fitnessvis ?
 1. in discussion refer to linkage disequilibrium about Fig 8D
 1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage, or by the dose ?
-1. mention other potential effects e.g. costs of resistance, sex linkage, ..._
+1. mention other potential effects e.g. sex linkage
 1. cite in press Killeen about combining interventions
@@
 Exposure, C.
-Dominance of resistance, D.
+Dominance of restoration, D.
 Resistance restoration.
@@
          guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance") +
+         labs(colour = "dominance of restoration") +
          coord_trans(y = "log10") + 
@@
 
-```{r Fig9, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
+```{r Fig 9 cost of 1 resistance, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of cost of resistance on time-to-resistance for mixtures and sequences.
+Row 1 (A and B) same as Fig.
+5 with no cost of resistance, row 2 (C and D) cost of resistance set to 0.15 with dominance of cost set to 0.5.
+The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
+The increase in cost of resistance from A.
+to C.
+removes the advantage of sequence relative to mixture.
+The ratio of time-to-resistance for the mixture divided by that for the sequence goes from 0.
+to 1.
+Costs seem to favour mixtures in these plots.
+Also be aware that costs would lead to a decline in the frequency of resistance for the first insecticide in a sequence when it stops being used."}
+
+#4 panel figure looking at effectiveness and exposure
+
+#test of rearranging figures on a grid
+#eff1  0.5  0.8
+#exp
+#0.5    1   2
+#0.8    3   4
+
+#layout(matrix(1:4,2,2, byrow = TRUE))
+
+#add lower panel for legend
+#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
+
+#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
+# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
+#layout.show(n=5)
+
+# add side panels for labels
+layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+
+#effectiveness on mixture
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+
+#base scenario
+
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
+
+mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#effectiveness1 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
+
+mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('cost 1 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#exposure 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
+
+mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+#eff1  and  expos 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
+
+mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
+
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#add for when insecticides same
+legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+#label
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+
+```
+
+```{r Figxx not eval, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
 All control inputs equal at 0.5, male exposure 0.8, B.

### 2017-03-08 R/plot_fit_calc.r

@@ -17,7 +17,7 @@
 #' plot_fit_calc(effectiveness=0.5, resistance_restoration=0.5, dominance=0.5)
 #' 
 #' 
-#' @return 
+#' @return gtable plot object
 #' @export
 
 plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.5, dominance_cost=0.5, cost=0.3 ){

### 2017-03-17 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-### version 10
-
-#Insecticide resistance evolution with mixtures and sequences : a graphical primer.
+### version 11
+
+#Insecticide resistance evolution with mixtures and sequences : a graphical introduction.
 
@@
 1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
-2. Exposure | proportion of insects exposed to the insecticide
-3. Resistance restoration  | ability of resistance (RR) to restore fitness when insect exposed to the insecticide
-4. Dominance of restoration | multiplier of resistance restoration for heterozygous (SR) insects
+2. Exposure | proportion of insects exposed to insecticide
+3. Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide
+4. Dominance of restoration | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
 5. Frequency  | frequency of resistance alleles within the population
 6. Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide
-7. Dominance of cost  | multiplier of cost for heterozygous (SR) insects in absence of insecticide
+7. Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide
 
@@
 We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The more conventional 'selection coefficient' is calculated by multiplying resistance restoration by insecticide effectiveness.
+The more 'selection coefficient' can be calculated by multiplying resistance restoration by insecticide effectiveness.
 
@@
 This acts to make the fitter alleles more common over time.
-Thus a summary of the selection stage is as follows with 1 & 2 representing the two insecticides :
-
-> selection coefficient = resistance restoration * effectiveness  
->
-> Single gene, insecticide fitness : 	
-> SS = 1-effectiveness  
-> RS = 1-effectiveness + (dominance*selection coeff.)  
-> RR = 1-effectiveness + selection coeff.  
-> 
-> Double gene, insecticide fitness :   
-> SS1SS2 = SS1 * SS2  
-> SS1RS2 = SS1 * RS2  
-> … and for other combinations  
-> RR1RR2 = RR1 * RR2  
-> 
-> Population fitness = Double fitness * exposure to insecticides  
-> 
-> In each generation :   
-> Allele frequency[t+1] = Allele frequency[t] * Population fitness
-
-Sexual reproduction is represented first by creating gametes by splitting the adult genomes in two with recombination, and second by forming offspring genotypes by adding gametes.
-
-So far this description just considers a single insecticide and associated resistance allele, with equal behaviour for males and females (Fig 1A).
-To allow males and females to be exposed differently separate genders are also represented (Fig 1B).
-In addition a second insecticide and resistance allele are included and fitness is calculated by multiplying these with those for the first insecticide described above (Fig 1C).
-Thus the simulation can represent populations exposed to two insecticides together in a mixture.
+Separate sexes and standard sexual reproduction with recombination is included.
+There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that here.
+
+So far this description just considers a single insecticide and associated resistance allele.
+In the model itself a second insecticide and resistance allele are included and overall fitness is calculated by multiplying the two results.
+This allows the model to represent populations exposed to two insecticides together in a mixture.
 
 Here we principally look at model outputs of the change in resistance allele frequencies over time (generations).
-Resistance allele frequencies are calculated from the frequencies of each genotype as follows.
-
-> Resistance allele frequency = (2*RR + RS) / 2
 

### 2017-03-21 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

@@
 
-_must be 350 words, currently 375 with subheadings Background, Results and Conclusions_
+_must be 350 words, currently 360_
 
 ##### Background
-Insecticide resistance has been identified as a major problem for the control of malarial mosquitoes with potentially huge consequences for malaria burden.
-To address this problem effort is being made to develop new insecticide tools including new active ingredients and combinations with existing ones.
-There are different potential strategies for using new insecticides and recomendations for how best to preserve their utility for controlling mosquitoes and malaria.
-There is however limited recent, accessible modelling work to support suggestions of which strategies may be best in different situations.
-Previous work suggests that preferred strategies will be situation specific.
-This points to a need for an accessible, mechanistic understanding of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-Such a mechanistic understanding could allow a more robust response to uncertainty in parameter values in field situations.*finesse that sentence or move to discussion*
-
-Here we use a relatively simple model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures.
-The model flexibly represents two independent genes coding for resistance to two insecticides allowing us to represent the different intervention strategies.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, the proportion of mosquitoes that come into contact with the insecticide, and how much the resistance alleles counteract the effect of the insecticide.
+Insecticide resistance threatens the control of vectors and particularly malaria.
+New insecticide tools are being developed to address this including new active ingredients and combinations.
+Potential strategies for using new insecticides include mixtures, sequences and rotations and there are recommendations for how best to implement these to manage resistance.
+There is, however, limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
+Previous work concludes that preferred strategies will be situation specific.
+There appears to be a low level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
+
+This paper develops an accessible, mechanistic picture of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
+The aim is to reach an audience that might not read a more detailed modelling paper.
+We use our model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures.
+The model flexibly represents two independent genes coding for resistance to two insecticides.
+We look principally at the ability of the insecticides to kill susceptible mosquitoes, how much resistance counteracts this, the proportion of mosquitoes that are exposed to insecticides and costs associated with resistance.
 
 ##### Results
-We show firstly that using a second insecticide, in a mixture with a first, always slows evolution of resistance to the first.
-Secondly we show that the ability of an insecticide to kill the susceptible mosquitoes (effectiveness), is the most important factor influencing whether using insecticides in a sequence or a mixture is likely to lead to faster resistance to both of them.
-The mechanisms behind these results are explained.
-We discuss how these results can inform intervention strategies in the field.
+We use the model to demonstrate the evolution of resistance under different parameter scenarios and how this fits with intuitive reasoning about the selection pressure exerted by insecticides.
+We show that an insecticide used in a mixture, relative to alone, always prompts slower evolution of resistance, but resistance to the two insecticides may evolve more slowly when used in sequence.
+We show that the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on whether resistance to two insecticides is likely to arise faster in a mixture or sequence.
 
 ##### Conclusions
-We describe an R package and user-interface that allows further exploration of the development of resistance.
+We describe an R package and user-interface allowing further exploration of resistance evolution.
 The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.
@@
 Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
-Focusing on malaria which still results in hundreds of thousands of deaths per year, recent control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
-Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million malaria cases between 2000 and 2015 [@Bhatt2015].
+Malaria alone still results in hundreds of thousands of deaths per year.
+Recent malaria control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
+Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million falciparum malaria cases between 2000 and 2015 [@Bhatt2015].
 A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
 
-There have been modelling studies investigating the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
-[@Curtis1985][@Mani1985][@Roush1989] but much of the modelling work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
-For those who (wisely in some cases) distrust models, work in a moth and genetically modified brocolli system has shown that toxins used in combination can delay the onset of resistance in a way that is consistent with theoretical models [@Zhao2003].
-
+The WHO produced a Global Plan for Insecticide Resistance Managment in malaria vectors (GPIRM)[@WHO2012] which includes recomendations on operational strategies for managing resistance including the use of insecticide mixtures when they become available.
+Efforts are underway to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
 The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
@@
 It is important that decisions about how best to use the new insecticides to delay the onset of resistance are made before insecticides are released [@Hemingway2016].
-The model presented here can help inform strategies to minimise the development of resistance in new and existing insecticides.
-
-Insecticide resistance becomes a problem when genes coding for resistance firstly arise in a population and secondly increase in frequency.
-We concern ourselves with this second process of how insecticide resistance increases in frequency within a population.
+
+Modelling studies have investigated the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
+[@Curtis1985][@Mani1985][@Roush1989] but much of the work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
+In an earlier paper [@Levick2017] we described the technical details of a flexible model used to investigate the relative benefits of mixtures and sequences.
+Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance is influenced by different inputs.
+This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
+
+Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
 It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
-The changing frequency of insecticide resistance is a population genetic process that can be influenced by (among others) the parameters outlined in Table 1.
-
-_TODO add in here about theory of IR and protocols for resistance management._
-
-We use a 2-locus population genetic model of insecticide resistance [@Levick2017] to illustrate the mechanisms by which these parameters are expected to influence the development of insectide resistance under alternative insecticide use strategies.
-In an earlier paper [@Levick2017] we described the technical details of the model and explore large parameter ranges.
-We provide an accessible summary of the model and use selected parameter values to develop a mechanistic understanding of how inputs influence the development of resistance.
-This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
+
+## Methods
+
+The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
+One gene or locus is represented for each insecticide.
+Each individual has two copies of the gene and there are two potential alleles per locus.
+Each allele confers either resistance or susceptibility to the insecticide.
+Thus individuals can either have both resistance genes (RR homozygous), both susceptible genes (SS) or one of each (SR, heterozygous).
+The combination of the two genes is termed the genotype.
+Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
+The fitness of the SS genotype in absence of the insecticide is set at a maximum value of 1.
+All other fitnesses are less, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
+These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
 
@@
 7. Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide
-
-## Methods
-
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
-One locus is represented for each insecticide.
-There are two potential alleles per locus.
-Each allele confers either resistance or susceptibility to the insecticide.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces, with a maximum value of 1 defined as the fitness of the SS genotype in absence of the insecticide.
-The fitness of an individual depends on it's alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
 
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Diagrammatic representation of the effect of model inputs on the fitness of genotypes for a single insecticide.
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="The effect of model inputs on the fitness of genotypes for a single insecticide.
 Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
 Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of restoration determines how the fitness for SR sits between that of SS and RR.
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of restoration determines how the fitness for SR lies between that of SS and RR.
 For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, dominance of restoration=0.4, resistance restoration=0.5, dominance of cost=0.4 and resistance cost=0.3."}
-
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.4, dominance_cost=0.4, cost=0.3 )
+In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3."}
+
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
 

### 2017-03-21 R/plot_fit_calc.r

@@ -21,7 +21,7 @@
 #' @export
 
 #plot_fit_calc <- function ( effectiveness=0.8, resistance_restoration=0.6, dominance_restoration=0.6, dominance_cost=0.5, cost=0.3 ){
-plot_fit_calc <- function ( effectiveness=0.75, resistance_restoration=0.75, dominance_restoration=0.75, dominance_cost=0.75, cost=0.2 ){
+plot_fit_calc <- function ( effectiveness=0.75, resistance_restoration=0.75, dominance_restoration=0.4, dominance_cost=0.6, cost=0.2 ){
     
   
   #library(gridExtra)
@@ -38,7 +38,7 @@ plot_fit_calc <- function ( effectiveness=0.75, resistance_restoration=0.75, dom
   
   dfg <- data.frame(x1 = c('SS',  'SR',  'RR'), 
                     x2 = c('SS',  'SR',  'RR'),                    
-                    y1 = c(1,  rr, ss), 
+                    y1 = c(1,  ss, ss), 
                     y2 = c(ss, sr, rr))
 
   gg1 <- ggplot(dfg, aes(x = x2, y = y2)) +    
@@ -51,7 +51,7 @@ plot_fit_calc <- function ( effectiveness=0.75, resistance_restoration=0.75, dom
         annotate("segment", x='SS', xend='RR', y=ss, yend=ss, linetype="dotted", colour = "blue") +
         annotate("segment", x='SR', xend='RR', y=rr, yend=rr, linetype="dotted", colour = "blue") +
         #dotted vertical for dominance
-        annotate("segment", x='SR', xend='SR', y=sr, yend=ss, linetype="dotted", colour = "blue") +    
+        annotate("segment", x='SR', xend='SR', y=sr, yend=rr, linetype="dotted", colour = "blue") +    
         ylab('fitness') +
         xlab('genotype') +
         #theme(axis.title.y = element_text(size = rel(3))) +
@@ -92,7 +92,7 @@ plot_fit_calc <- function ( effectiveness=0.75, resistance_restoration=0.75, dom
 
   dfg <- data.frame(x1 = c('SS',  'SR',  'RR'), 
                     x2 = c('SS',  'SR',  'RR'),                    
-                    y1 = c(1,  rr, ss), 
+                    y1 = c(1,  ss, ss), 
                     y2 = c(ss, sr, rr))
     
   gg2 <- ggplot(dfg, aes(x = x2, y = y2)) +    
[…]

### 2017-03-22 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-#Insecticide resistance evolution with mixtures and sequences : a graphical introduction.
-
+#Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary.
+
+###alternative titles:
 ###Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
-
 ###An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
-
 ###A tutorial, primer ...
-
 ###Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
@@
 
-_must be 350 words, currently 360_
+_must be 350 words, currently 351_
 
@@
 Insecticide resistance threatens the control of vectors and particularly malaria.
-New insecticide tools are being developed to address this including new active ingredients and combinations.
+New insecticides are being developed to address this.
 Potential strategies for using new insecticides include mixtures, sequences and rotations and there are recommendations for how best to implement these to manage resistance.
@@
 Previous work concludes that preferred strategies will be situation specific.
-There appears to be a low level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
+There is potential to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
 
 This paper develops an accessible, mechanistic picture of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-The aim is to reach an audience that might not read a more detailed modelling paper.
-We use our model to develop a mechanistic understanding of how insecticide resistance is expected to increase under different intervention strategies, namely the use of insecticides in isolation, sequence and mixtures.
+The aim is to reach an audience unlikely to read a more detailed modelling paper.
+We use our model to develop a mechanistic understanding of how insecticide resistance is expected to increase with the use of insecticides in isolation, sequence and mixtures.
 The model flexibly represents two independent genes coding for resistance to two insecticides.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, how much resistance counteracts this, the proportion of mosquitoes that are exposed to insecticides and costs associated with resistance.
+We look principally at the ability of the insecticides to kill susceptible mosquitoes, how much resistance counteracts this, the proportion of mosquitoes that are exposed to insecticides and costs of resistance.
 
 ##### Results
-We use the model to demonstrate the evolution of resistance under different parameter scenarios and how this fits with intuitive reasoning about the selection pressure exerted by insecticides.
+We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about the selection pressure exerted by insecticides.
 We show that an insecticide used in a mixture, relative to alone, always prompts slower evolution of resistance, but resistance to the two insecticides may evolve more slowly when used in sequence.
@@
 ##### Conclusions
-We describe an R package and user-interface allowing further exploration of resistance evolution.
-The user-interface (https://andysouth.shinyapps.io/MixSeqResist1) allows replication of our results and exploration of what happens when inputs are changed outside of the ranges we have used.
+Our model makes more open the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
+We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/MixSeqResist1).
+These tools can contribute to operational decisions in insecticide resistance management.
 
@@
 A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
+We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
+The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
+
 Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
@@
 Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Figure 1).
-
-Each of these inputs can be varied as a proportion between 0 and 1.
-We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The more 'selection coefficient' can be calculated by multiplying resistance restoration by insecticide effectiveness.
 
@@
 For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
-
-Similarly, higher values of starting resistance frequency also resulted in faster spread of resistance (Fig 4A) and shorter times to resistance thresholds.
-Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect with higher values leading to a slower development of resistance.
+This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
+Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that less change is required to reach the threshold.
+
+Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect.
+Higher values led to a slower development of resistance.
+Again this is intuitive because costs reduce the advantage of resistance and decrease selection.
+These effects of inputs on single insecticide use are summarised in Table 2.
 
 ### Two insecticides
-When two insecticides are used the relative performance of a mixture strategy compared to sequential use changes dependent on input values.
-To help understand the mechanisms influencing this relative performance we started with a single base scenario and investigated changing inputs individually.
-The base scenario was similar to that used for the single insecticide.
-Values for the first four inputs in Table 1 were set to 0.5 and the starting frequencies of resistance to 0.01.
-For this base scenario resistance arises slower for sequential than mixture (Fig 5A).
-This plot, and those that follow, compares mixture and sequence for a single combination of inputs.
+When two insecticides are used we can compare the relative performance of a mixture with sequential use.
+Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
+To help understand the mechanisms influencing this relative performance we start with a single base scenario and demonstrate changing inputs individually.
+In this base scenario we set effectiveness, exposure, resistance restoration, and dominance at 0.5, the starting frequencies of resistance to 0.01 and cost of resistance to 0.
+Resistance arises slower for sequential than mixture for this base scenario (Fig 5A).
 Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold of 50%.
-
-#### results: Insecticide exposure and effectiveness
-
-Increasing exposure to the insecticides (the proportion of insects that come into contact) from 0.5 to 0.8 (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy and so does not change their relative performance; time-to-resistance remains longer for the sequence.
-In contrast keeping the exposure constant at 0.5 and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides to 0.8 (Fig 5D) results in a longer time-to-resistance for the mixture relative to the sequence.
-Resistance for the more effective insecticide (in red) increases faster than the less effective.
-Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effecttive insecticide has reached the resistance threshold (~50 generations).
-
-Thus a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
+When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 5A).
+Thus the time-to-resistance for the mixture divided by the sequence is less than 1 at 0.8.
+
+#### Insecticide exposure and effectiveness
+
+Increasing exposure (the proportion of insects that come into contact with the insecticides) (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy.
+The effect on the mixture is greater so time-to-resistance remains longer for the sequence (the ratio decreasing from 0.8 to 0.6).
+In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 5B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
+The longer time-to-resistance for the mixture results from a slower increase in resistance for the less effective insecticide.
+Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
+This points a mechanism whereby the more effective insecticide initially slows the rate o evolution of resistance to the less effective one.
+
+The results is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
 This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-Figure 5D has the same inputs as Fig 5B except that the exposure is increased from 0.5 to 0.8.
-The increase in exposure reduces time-to-resistance more for the mixture relative to the sequence.
-The result is that the advantage of mixtures over sequence is removed by the increase in exposure, leading to approximately equal times-to-resistance for the two strategies (Fig 5D).
-
-Figure 6 shows how increasing the effectiveness of either or both insecticides improves the performance of mixtures over sequences in terms of time-to-resistance.
+
+In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 5, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 5.
+Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
+Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
+
+Increasing the effectiveness further of either or both insecticides increases the performance of mixtures over sequences in terms of time-to-resistance.
 Figure 6A uses the same inputs as Fig 5B.
 From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
-
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
-
-#### results: Dominance and resistance restoration
-Increasing dominance or resistance restoration decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
+This again points to how more effective insecticides can slow resistance to another when used in a mixture.
+
+#### Resistance restoration and dominance
+Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
 The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
-
-#### results: Starting frequencies of resistance
+So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
+
+#### Starting frequencies of resistance
 Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
@@
 
-#### results: Costs of resistance
+#### Costs of resistance
 Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
@@
 
+The effects of inputs on the evolution of resistance for mixtures is summarised in Table 3 and for the difference between mixtures and sequences is summarised in Table 4.
+
+#### Linkage disequilibrium
 It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
@@
 
-The response of resistance to insecticide use in the model can be explained by mechanisms that can help us develop and communicate a more robust understanding of how we would expect resistance to respond in the field.
-
-When single insecticide use was represented (Figs 3-4) in all cases the initial increase in resistance was most rapid with straight lines on the log plots up to around 50% followed by a slowing rate of increase up to 100%.
-Higher values of the five inputs investigated all led to more rapid increases in resistance.
+Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
+This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
+
+When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resisitance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
+The latter is as expected because it reduces the selective advantage of resistance.
+
 For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
@@
 The faster development of resistance under higher resistance-restoration (Fig 3D) can be explained by it's effect on the survival of the resistant genotypes.
-Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selection pressure of the resistance allele.
+Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selective advantage of the resistance allele.
 The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs.
@@
 
-#### discussion: Two insecticides
-
-In comparison with the use of a single insecticide, using two insecticides either in sequence or in mixtures differs in some ways and where it differs that can again help us to understand the mechanisms.
-Looking at our base scenario (Fig 5A) the curves for resistance to both insecticides in the mixture are shallower than those for sequential use and in this particular example are identical because they have set the same parameter values for each insecticide to be the same.
+#### Two insecticides
+Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide and, where it differs can again help us to understand the dynamics of resistance evolution.
+Our base scenario (Fig 5A) shows curves for resistance to both insecticides in the mixture that are shallower than those for sequential use.
 The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
@@
 However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
-As shown in our base scenario, resistance can arise more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-
-#### discussion: Insecticide effectiveness
+Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
+In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture.
+
+#### Insecticide effectiveness
 
@@
 In contrast in both the sequential (Fig 5B) and single insecticide (Fig 2A) scenarios the development of resistance was faster when effectiveness was higher.
-This can be explained by the increased effectiveness of one insecticide in a mixture increasing killing of vectors that are resistant to the other and thus decreasing selection pressure.
+This can be explained by the increased effectiveness of one insecticide in a mixture increasing killing of vectors that are resistant to the other and thus decreasing selection pressure in a protective effect.
 
@@
 Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide and thus stops limiting the rise in resistance to the second insecticide.
+Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide, loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
 A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
@@
 
-#### discussion: Exposure vs effectiveness
+#### Difference between exposure and effectiveness
 It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
@@
 This difference between effectiveness and exposure was unexpected for us.
+
 _todo could this be because exposure results doesn't alter the relative fitness of genotypes because all genotypes are exposed equally ? doesn't quite make sense to me_
@@
 
-#### discussion costs of resistance
+#### Costs of resistance
 Costs of resistance favour mixtures in these analyses.
@@
 
-#### discussion: caveats
+#### Caveats
 It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria.
@@
 
-#### discussion: putting results into context
+#### Putting results into context
 
@@
 
+\pagebreak
+
 **Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
@@
 Parameter to increase  | effect on resistance evolution | Mechanism
------------------- | -------- | --------------------------------------------------
+----------------------- | -------- | ---------------------------------------------
 1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
@@
 Parameter to increase  | effect on resistance evolution | Mechanism
------------------- | ---------- | --------------------------------------------------
-1. Effectiveness | **slower** | reduced fitness of individuals resistant to the other insecticide
+----------------------- | ---------- | ---------------------------------------------
+1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure from the other
 2.
 Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide.
+However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
 3. Dominance of restoration | faster | increased fitness of heterozygotes
@@
 Parameter to increase  | increase favours mix or sequence | Mechanism
------------------- | ------- | ---------------------------------------------------
+----------------------- | -------- | ---------------------------------------------
 1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
@@
 1. mention other potential effects e.g. sex linkage
-1. cite in press Killeen about combining interventions
 

### 2017-03-23 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
-The WHO produced a Global Plan for Insecticide Resistance Managment in malaria vectors (GPIRM)[@WHO2012] which includes recomendations on operational strategies for managing resistance including the use of insecticide mixtures when they become available.
-Efforts are underway to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
+The WHO produced a Global Plan for Insecticide Resistance Management in malaria vectors (GPIRM)[@WHO2012] which includes recommendations on operational strategies for managing resistance including the use of insecticide mixtures when they become available.
+Efforts are under way to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
 The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
@@
 Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance is influenced by different inputs.
-This mechanistic understanding can provide a more robust framework for the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
+This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
 
@@
 The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
+Cost of resistance determines the fitness of RR in the portion of the population not exposed to the insecticide.
 
@@
 
-Here we principally look at model outputs of the change in resistance allele frequencies over time (generations).
+The model outputs the change in resistance allele frequencies over time (generations).
+We use this to calculate a time-to-resistance using a resistance threshold of 50%.
+This is consistent with thersholds used in previous modelling analyses [todo refs] and represents a level where the resistance alleles are close to fixation.
 
@@
 
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an online user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
+The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an on-line user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
 
@@
 Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
-This points a mechanism whereby the more effective insecticide initially slows the rate o evolution of resistance to the less effective one.
+This points a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
 
@@
 
-The effects of inputs on the evolution of resistance for mixtures is summarised in Table 3 and for the difference between mixtures and sequences is summarised in Table 4.
+A summary of the effects of inputs on the evolution of resistance for mixtures is given in Table 3 and for the difference between mixtures and sequences in Table 4.
+The mechanisms identified in the tables are covered further in the discussion.
 
@@
 
-When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resisitance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
+When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resistance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
 The latter is as expected because it reduces the selective advantage of resistance.
@@
 This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference between effectiveness and exposure was unexpected for us.
-
-_todo could this be because exposure results doesn't alter the relative fitness of genotypes because all genotypes are exposed equally ? doesn't quite make sense to me_
-
-The mechanisms involved in the response of resistance in mixtures are summarised in Table3.
-Whether a mixture or sequence is favoured for reducing the spread of resistance will depend on these mechanisms and is summarised in Table 4.
+This difference between effectiveness and exposure, that effectiveness increases times-to-resistance for a mixture and exposure reduces it, was unexpected for us.
+
+Effectiveness and exposure were also found to be the most important inputs in our earlier analysis where we conducted a more comprehensive sensiitivty analysis [@Levick2017].
+If these predicted effects are confirmed then in designs of insecticide mixtures it would be best in terms of evolution to have high effectiveness and low exposure.
+This low exposure could be achieved by having active ingredients with some repellency.
 
@@
 This is because the evolution of a resistance with an associated cost is slowed more in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-Cost of resistance to one insecticide reduces selection pressure for that gene and this effect becomes more as the frequency of resistance increases.
+In a mixture the protective effect of the other insecticide identified above interacts with the cost of resistance to reduce selection pressure further.
+
+It can also be seen that the effects of cost of resistance become more as the frequency of resistance increases.
+The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
+This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
 Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
@@
 above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-We do not expect such high costs to occur in operational settings.
+Unfortunately, we do not expect such high costs to occur in operational settings.
 
 However, also be aware that in our analyses we have used a single resistance threshold.
-When there are costs of resistance, the frequency of resistance for insectides that are not being used is likely to decline.
+When there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
 Thus the frequency of resistance for the first insecticide in a sequence may decline to a level where it becomes operationally useful again.
@@
 #### Caveats
-It is worth re-iterating that we are looking solely at the evolution of resistance and not at the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as effecting the evolution of resistance, are also likely to influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
+The analysis presented here examines solely the evolution of resistance and not the control of mosquitoes or malaria.
+Different insecticide use strategies, as well as effecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
 Insecticide resistance could reduce vector transmission of disease (e.g.
@@
 Of course, using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining that in the longer term by restricting the development of resistance.
-Others are looking at combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
+The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
+Others are combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
+
+In our model we represented insecticide interaction as being additive within the mixture.
+This is consistent with recent work on aphids [@Taillebois2016], although that work did also indicate some synergistic effects.
+Our model has the potential to incorporate either synergistic or antagonistic effects too.
 
 The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ranson2016].
-The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of paramter space (e.g.
+The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
 holding most inputs constant at intermediate values while modifying single inputs in isolation).
@@
 However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-In addition we provide an online user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown here.
+This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
+In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown here.
 
@@
 
-I will now put these predicted responses of resistance into the context of previous work and recomendations.
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and the particular situations they represent.
+The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
+Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
 One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
@@
 This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D).
+see Fig 6D and Table 4).
 
@@
 
-Modelling and theory have been used to generalise Insecticide Resistance Management into three potential strategies [@Georghiou1994].
-These are A) moderation : preserving susceptible genes by limiting selection pressure, B) saturation : high dose so that heterozygous resistants are killed and C) multiple attack : independently acting pressures neither of which is strong enough to lead to resistance.
-Our work here principally focuses on multiple attack but the other strategies are relevant too.
-_todo expand this paragraph_
-
-One consistent aspect of the results presented here can at first seem to go against current advice on good practice.
-We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5).
+One consistent aspect of the results presented here can at first seem to contradict published advice.
+We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
 This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
 @Roush1998].
-In contrast the widely held view seems to be that low effectiveness or exposure associated with poor implementation can promote the development of resistance.
+In contrast the widely held view seems to be that poor implementation can promote the development of resistance.
 e.g.
@@
 These include: ...
-- the use of application rates that are below or above those recommended on the label; - poor coverage of the area being treated ...".
+- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
 The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of spread in the population [?ref].
-Our model does show that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive and negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Our model combined with laboratory and field results could help to assess these trade-offs.
+Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Barbosa2011][@FAO2012].
+Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
+So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
+Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that effectivenes favours mixtures, exposure favours sequences and higher dominance favours neither (Table 4).
 
@@
 This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the devlopment of resistance.
-
-## FINAL CONCLUDING PARAGRAPH
-
-~ Effectiveness and exposure (and cost?) are key and can be modified.
-~ The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies. 
-~ This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture is more expensive.
-
-### other papers to incorporate into discussion above
-
-Roush older paper : [@Roush1989]
-
-Relevance of these mechanisms in the field e.g.
-exposure can be reduced if either insecticide has a repellant effect, which can be set in formulation.
-[@Birget2015a] mention this not for resistance but for ITNs with repellancy requiring higher coverage for elimination)
-_TODO check which of the Birget papers mention this_
-
-How do the rest of our results relate to recomendations that have been produced by WHO[@WHO2012], IRAC[@IRAC2011] and FAO[@FAO2012] ?
-
-[@Taillebois2016]
-Laboratory experiments on aphids showing that in low dose mixtures insecticides most often had additive effects as represented in our model, but that also commonly mortality was higher than expected (synergistic) which we can also represent within the model.
-
-[@Barbosa2011] 
-_suggests incomplete dominance in the field_
-
-[@Birget2015] 
-We show that indoor use of insecticides leads to less selection pressure than their use as larvicides. 
-Reasons for relatively low selection pressure by adulticides
-(i) males are not affected by the ITNs
-(ii) insecticides are also repellents, keeping mosquitoes at bay from contacting the insecticide but also driving them to bite either people who do not use the insecticide or alternative hosts.
-
-[@Curtis1998]
-can anything be done to maintain effectiveness of nets ?
-
-[@Kliot2012]
-fitness costs associated with insecticide resistance
-
-Good review of tactics for IRM inlcuding mention of mixtures and new AIs [@Denholm1992]
-"Once developed and marketed, new products introduce a fresh challenge how to exploit their characteristics while restricting exposure to preserve their effectiveness."
-
-### Discussion some summary points probably already covered ...
+In both cases our modelling approach could be used to investigate implications for the development of resistance.
+
+#### Future work
+
+This model has been designed to be flexible to address further issues in Insecticide Resistance Management.
+With a little restructuring our model can represent other insecticide use strategies such as rotations and mosaics.
+
+#### FINAL CONCLUDING PARAGRAPH
+
+(@) Effectiveness and exposure are key and can be modified. Cost also has an effect and future work may want to look more at it's interaction with these other important inputs.
 
@@
 
-(@) exposure ...
+(@) The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies.
+
+(@) This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
+Other factors also need to be considered.
+On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
+On the negative side a mixture is more expensive.
 
@@
 ##### Acknowledgements
-Thankyou to Bethany Levick who developed the first version of the model code.
+Thank you to Bethany Levick who developed the first version of the model code.
 
@@
 
-## ANDY TO DO
-1. explain resistance thresholds and time-to-resistance in methods
-1. refer to ratios of mix:seq in plots
-1. refer to PRCC from pap1 in ref. to Table4
-1. mention single gene vs additive effects (with ref.)
-1. in discussion refer to linkage disequilibrium about Fig 8D
-1. is there a reference to say that the effectiveness of insecticides can be modified at the design stage, or by the dose ?
-1. mention other potential effects e.g. sex linkage
-
 \pagebreak

### 2017-03-24 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

@@
 
-The model outputs the change in resistance allele frequencies over time (generations).
-We use this to calculate a time-to-resistance using a resistance threshold of 50%.
-This is consistent with thersholds used in previous modelling analyses [todo refs] and represents a level where the resistance alleles are close to fixation.
+The model outputs the change in resistance allele frequencies over time measured in generations.
+We extract the number of generations at which a resistance allele frequency of 50% is reached and term this 'time-to-resistance'.
+Using a 50% threshold is consistent with previous modelling analyses e.g.
+[@Curtis1985][@Birget2015].
+Preliminary investigations showed that using other potential thresholds of 25 or 75% gave very similar results.
 
@@
 1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the proportion of resistance alleles reaches a threshold of 50%
+1. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
 1. two insecticides used in a mixture (concurrently)
@@
 When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resistance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
-The latter is as expected because it reduces the selective advantage of resistance.
-
-For single insecticide use resistance responded identically to changing exposure and effectiveness (compare Fig 3A to 3B).
-This makes sense as vector kill is effectively a product of exposure times effectiveness.
-For example exposing 50% of a  population to an insecticide which is 75% effective would be expected to have the same result as exposing 75% of the population to one which is 50% effective.
-This observation points to the mechanism by which increasing both exposure and effectiveness lead to a faster increase in resistance.
-In both cases the increased deaths of susceptible vectors cause a higher selection pressure that can explain the faster increase.
-
-The pattern of more rapid increase in resistance at higher dominance of restoration levels (Fig 3C) can be explained by increased survival of heterozygotes in the presence of the insecticide.
-Higher dominance of the resistant allele causes it to contribute more to the phenotype of the heterozygotes leading to higher survival.
-Thus selection pressure for the resistance allele will be increased because it confers more of an advantage when only present on one chromosome.
-The faster development of resistance under higher resistance-restoration (Fig 3D) can be explained by it's effect on the survival of the resistant genotypes.
-Resistance-restoration restores the survival of resistant genotypes in the presence of the insecticide back towards what it would be in the absence of the insecticide, thus increasing the selective advantage of the resistance allele.
-The effect of the starting frequency of resistance (Fig 4) is the most different from the other inputs.
-Simply, when starting from a higher frequency of resistance there is a smaller change to make to reach the higher resistance thresholds.
 
 These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Exposure and effectiveness act by decreasing the survival of all genotypes, dominance of restoration acts by increasing survival of RS and resistance-restoration acts by increasing survival of RR & RS.
+Referring to figure 1 can help explain these.
+Exposure and effectiveness act by decreasing the fitness of all genotypes, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
+Cost acts by decreasing the fitness of RR and SR for those not exposed.
 
@@
 Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture.
-
-#### Insecticide effectiveness
-
-The main difference between mixtures and the other strategies is that in mixtures increasing the effectiveness of either insecticide slows the development of resistance to both (compare Fig 5B to 5A).
-In contrast in both the sequential (Fig 5B) and single insecticide (Fig 2A) scenarios the development of resistance was faster when effectiveness was higher.
-This can be explained by the increased effectiveness of one insecticide in a mixture increasing killing of vectors that are resistant to the other and thus decreasing selection pressure in a protective effect.
-
-This effect can be seen by comparing Fig 5B, where the effectiveness of insecticide 1 has been increased, to Fig 5A.
+In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
+Incraesing the effectiveness of on insecticide to Fig 5B decreases the added times for the sequence to 90 and increases those for the mixture to 115 thus changing the order and favouring the mixture.
+
+The mechanism for this change can be sen by comparing the shape of the curves in Figs 5A & B.
 In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
@@
 It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 110 in Fig 5B).
+Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
 This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference between effectiveness and exposure, that effectiveness increases times-to-resistance for a mixture and exposure reduces it, was unexpected for us.
-
-Effectiveness and exposure were also found to be the most important inputs in our earlier analysis where we conducted a more comprehensive sensiitivty analysis [@Levick2017].
-If these predicted effects are confirmed then in designs of insecticide mixtures it would be best in terms of evolution to have high effectiveness and low exposure.
-This low exposure could be achieved by having active ingredients with some repellency.
+This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us.
 
 #### Costs of resistance
-Costs of resistance favour mixtures in these analyses.
-This is because the evolution of a resistance with an associated cost is slowed more in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-In a mixture the protective effect of the other insecticide identified above interacts with the cost of resistance to reduce selection pressure further.
-
-It can also be seen that the effects of cost of resistance become more as the frequency of resistance increases.
+Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
+It seems that the protective effect of the other insecticide in a mixture interacts with the cost of resistance to reduce selection pressure more than in a sequence.
+
+The effects of cost of resistance become more as the frequency of resistance increases.
 The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
@@
 
-However, also be aware that in our analyses we have used a single resistance threshold.
-When there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
-Thus the frequency of resistance for the first insecticide in a sequence may decline to a level where it becomes operationally useful again.
-As such a repeated sequential strategy could offer benefits over mixtures.
+Relatedly, when there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
+Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where it becomes operationally useful again.
+As such a repeated sequential strategy could offer benefits over mixtures not considered here.
+To incorporat this one would need to use a more detailed metric than time-to-resistance.
 
 #### Caveats
-The analysis presented here examines solely the evolution of resistance and not the control of mosquitoes or malaria.
+Our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
 Different insecticide use strategies, as well as effecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
@@
 This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown here.
-
-#### Putting results into context
-
+In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown.
+
+#### Consistency with previous work 
 The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
@@
 A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to the parameter we have termed insecticide effectiveness, so lends further support to our main conclusion of the importance of effectiveness.
-It's also shown [@Roush1998 Fig 3] that decreasing exposure from 90% to 80% also increases the benefit of mixtures over sequences in the same way that we have shown.
-
-One consistent aspect of the results presented here can at first seem to contradict published advice.
-We show that lower insecticide effectiveness or exposure are expected to lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
+This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
+Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
+Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
+
+One consistent aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
+Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
 This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
 @Roush1998].
-In contrast the widely held view seems to be that poor implementation can promote the development of resistance.
+This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
 e.g.
@@
 - the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
+
+Poor implementation would likely cause lower effectiveness of and exposure to insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
 The explanation for this apparent difference is in the effect of dominance.
@@
 Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that effectivenes favours mixtures, exposure favours sequences and higher dominance favours neither (Table 4).
+
+#### Implications
+_todo get more of [@Roush1989] in here_
 
@@
 
-#### Future work
-
-This model has been designed to be flexible to address further issues in Insecticide Resistance Management.
-With a little restructuring our model can represent other insecticide use strategies such as rotations and mosaics.
-
-#### FINAL CONCLUDING PARAGRAPH
-
-(@) Effectiveness and exposure are key and can be modified. Cost also has an effect and future work may want to look more at it's interaction with these other important inputs.
-
-(@) In a mixture each insecticide reduces the rate of increase in resistance to the other by killing individuals that are resistant to the other.
-
-(@) In a mixture higher effectiveness of either insecticide kills more individuals resistant to the other and thus increases time-to-resistance for the other.
-Thus although increasing effectiveness of an insecticide decreases time-to-resistance when used alone, when used in a mixture time-to-resistance for both insecticides is increased.
-
-(@) For a mixture of 2 insecticides with differing effectiveness.
-Resistance to the insecticide with a greater effectiveness increases faster.
-The less effective insecticide is 'protected' by the more effective.
-Resistance to the less effective insecticide increases slowly until resistance to the more effective insecticide reaches a high level.
-
-(@) The model we describe is flexible and as such can reproduce similar results to earlier models and be used to represent more detailed, operationally relevant strategies.
-
-(@) This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
+_low exposure could be achieved by having active ingredients with some repellency._
+
+_todo get Birget and Koella in & link to exposure_
+'whereas repellency is beneficial to delay the evolution of resistance, other models have shown that it decreases the population-level protection of the insecticide.'
+
+_todo, where to put this, earlier ?_
+This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
 Other factors also need to be considered.
@@
 On the negative side a mixture is more expensive.
+
+_todo, somewhere earlier say good thing about effectiveness & exposure is that they can be modified_
+
+#### Conclusions
+
+We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured.
+For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance.
+High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other.
+High exposure favours sequences through reducing the otherwise high selection pressures on both resistances together.
+High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced.
+Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them.
+Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.
 
@@
 6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
+
+\pagebreak
 

### 2017-04-12 inst/documents/paper2/test_linenumbers.Rmd
test line numbering and double spacing

--- 2017-03-24 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
 ---
-title: "paper2_resistance_mechanisms_mixtures"
-output: pdf_document
+title: "trying to get line numbers working"
+# output:
+#   pdf_document:
+#     fig_caption: yes
+#   graphics: yes
+#   keep_tex: yes  
+#   latex_engine: pdflatex
+#   template: bmc_template/bmc_article.tex
+
+output: 
+  pdf_document:
+    #citation_package: natbib
+    keep_tex: true
+    fig_caption: true
+    latex_engine: pdflatex
 #output: word_document
@@
 csl: biomed-central.csl
-#header-includes:
-#    - \usepackage{setspace}
-#    - \doublespacing
-#    - \usepackage{lineno}
-#    - \linenumbers
+fontsize: 11pt
+spacing: double
+
+header-includes:
+   - \usepackage{setspace}
+   - \doublespacing
+   - \usepackage{lineno}
+   - \linenumbers
+
 ---
 
-```{r, eval=FALSE, include=FALSE}
-#!!!! paper protocol
-#!!!! might be easier to put text and figs into separate Rmd files after draft stage
-#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
-#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
-#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
-#!!!! submission figs, set this chunk to eval=TRUE
-
-# options to create final publication quality figures
-# causes the document to fail, but the figs do get stored in pap2figs/
-library(knitr)
-opts_chunk$set(dev="tiff",
-               dev.args=list(compression="lzw"),
-               dpi=300,
-               cache=FALSE,
-               fig.path='pap2figs/')
-```
-
-```{r,echo=FALSE,results='hide',warning=FALSE}
-  library(resistance)
-  library(ggplot2)
-```
-
-### version 11
-
-#Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary.
-
-###alternative titles:
-###Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
-###An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
-###A tutorial, primer ...
-###Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
-
-Target : Malaria journal
-
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and suggestions welcome
-
-## Abstract
-
-_must be 350 words, currently 351_
+Seems that the header-includes bit above does work to add line numbers and double space (although it does generate a few warning messages)
 
@@
 
-Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
-It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
+My working plot.
 
-## Methods
-
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
-One gene or locus is represented for each insecticide.
-Each individual has two copies of the gene and there are two potential alleles per locus.
-Each allele confers either resistance or susceptibility to the insecticide.
-Thus individuals can either have both resistance genes (RR homozygous), both susceptible genes (SS) or one of each (SR, heterozygous).
-The combination of the two genes is termed the genotype.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
-The fitness of the SS genotype in absence of the insecticide is set at a maximum value of 1.
-All other fitnesses are less, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
-
-**Table 1. Parameters inflencing the development of insecticide resistance**
-
-Parameter  | Description
----------------------- | ----------------------------------------------------
-1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
-2. Exposure | proportion of insects exposed to insecticide
-3. Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide
-4. Dominance of restoration | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
-5. Frequency  | frequency of resistance alleles within the population
-6. Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide
-7. Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide
-
-The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 1.
-In the model itself this is repeated for both insecticides and the fitnesses multiplied.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS and resistance restoration 'restores' a portion of the fitness for RR.
-A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
-We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
-Cost of resistance determines the fitness of RR in the portion of the population not exposed to the insecticide.
-
-Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
-For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
-Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Figure 1).
-
-The simulation proceeds through generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
-This acts to make the fitter alleles more common over time.
-Separate sexes and standard sexual reproduction with recombination is included.
-There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that here.
-
-So far this description just considers a single insecticide and associated resistance allele.
-In the model itself a second insecticide and resistance allele are included and overall fitness is calculated by multiplying the two results.
-This allows the model to represent populations exposed to two insecticides together in a mixture.
-
-The model outputs the change in resistance allele frequencies over time measured in generations.
-We extract the number of generations at which a resistance allele frequency of 50% is reached and term this 'time-to-resistance'.
-Using a 50% threshold is consistent with previous modelling analyses e.g.
-[@Curtis1985][@Birget2015].
-Preliminary investigations showed that using other potential thresholds of 25 or 75% gave very similar results.
-
-We investigate three insecticide use strategies.
-
-1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
-1. two insecticides used in a mixture (concurrently)
-
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an on-line user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
-
-## Results
-
-### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
-This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
-Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that less change is required to reach the threshold.
-
-Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect.
-Higher values led to a slower development of resistance.
-Again this is intuitive because costs reduce the advantage of resistance and decrease selection.
-These effects of inputs on single insecticide use are summarised in Table 2.
-
-### Two insecticides
-When two insecticides are used we can compare the relative performance of a mixture with sequential use.
-Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
-To help understand the mechanisms influencing this relative performance we start with a single base scenario and demonstrate changing inputs individually.
-In this base scenario we set effectiveness, exposure, resistance restoration, and dominance at 0.5, the starting frequencies of resistance to 0.01 and cost of resistance to 0.
-Resistance arises slower for sequential than mixture for this base scenario (Fig 5A).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 5A).
-Thus the time-to-resistance for the mixture divided by the sequence is less than 1 at 0.8.
-
-#### Insecticide exposure and effectiveness
-
-Increasing exposure (the proportion of insects that come into contact with the insecticides) (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy.
-The effect on the mixture is greater so time-to-resistance remains longer for the sequence (the ratio decreasing from 0.8 to 0.6).
-In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 5B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
-The longer time-to-resistance for the mixture results from a slower increase in resistance for the less effective insecticide.
-Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
-This points a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
-
-The results is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
-This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-
-In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 5, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 5.
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
-
-Increasing the effectiveness further of either or both insecticides increases the performance of mixtures over sequences in terms of time-to-resistance.
-Figure 6A uses the same inputs as Fig 5B.
-From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
-This again points to how more effective insecticides can slow resistance to another when used in a mixture.
-
-#### Resistance restoration and dominance
-Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
-The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
-So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
-
-#### Starting frequencies of resistance
-Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
-Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
-
-#### Costs of resistance
-Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
-The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 9C).
-The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C.
-With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 9A and 1 in Fig 9C.
-At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
-
-A summary of the effects of inputs on the evolution of resistance for mixtures is given in Table 3 and for the difference between mixtures and sequences in Table 4.
-The mechanisms identified in the tables are covered further in the discussion.
-
-#### Linkage disequilibrium
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
-
-## Discussion
-
-Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
-This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
-
-When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resistance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
-
-These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Referring to figure 1 can help explain these.
-Exposure and effectiveness act by decreasing the fitness of all genotypes, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
-Cost acts by decreasing the fitness of RR and SR for those not exposed.
-
-#### Two insecticides
-Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide and, where it differs can again help us to understand the dynamics of resistance evolution.
-Our base scenario (Fig 5A) shows curves for resistance to both insecticides in the mixture that are shallower than those for sequential use.
-The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Thus each insecticide reduces the selection pressure for the increase in resistance to the other and could be said to 'protect' the other.
-However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
-Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
-Incraesing the effectiveness of on insecticide to Fig 5B decreases the added times for the sequence to 90 and increases those for the mixture to 115 thus changing the order and favouring the mixture.
-
-The mechanism for this change can be sen by comparing the shape of the curves in Figs 5A & B.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide, loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide.
-The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-
-#### Difference between exposure and effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us.
-
-#### Costs of resistance
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-It seems that the protective effect of the other insecticide in a mixture interacts with the cost of resistance to reduce selection pressure more than in a sequence.
-
-The effects of cost of resistance become more as the frequency of resistance increases.
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
-Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
-Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
-Also note that if the costs are set even higher than we set them here (e.g.
-above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-Unfortunately, we do not expect such high costs to occur in operational settings.
-
-Relatedly, when there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
-Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where it becomes operationally useful again.
-As such a repeated sequential strategy could offer benefits over mixtures not considered here.
-To incorporat this one would need to use a more detailed metric than time-to-resistance.
-
-#### Caveats
-Our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as effecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
-Insecticide resistance could reduce vector transmission of disease (e.g.
-by reduced lifespan of the vector) or increase transmission (e.g.
-by reduced immunity of the vector to the disease) [@Rivero2010].
-Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Of course, using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
-Others are combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
-
-In our model we represented insecticide interaction as being additive within the mixture.
-This is consistent with recent work on aphids [@Taillebois2016], although that work did also indicate some synergistic effects.
-Our model has the potential to incorporate either synergistic or antagonistic effects too.
-
-The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ranson2016].
-The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-Thus there is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown.
-
-#### Consistency with previous work 
-The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
-
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
-Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
-
-One consistent aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
-
-Poor implementation would likely cause lower effectiveness of and exposure to insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Barbosa2011][@FAO2012].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that effectivenes favours mixtures, exposure favours sequences and higher dominance favours neither (Table 4).
-
-#### Implications
-_todo get more of [@Roush1989] in here_
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the development of resistance.
-
-_low exposure could be achieved by having active ingredients with some repellency._
-
-_todo get Birget and Koella in & link to exposure_
-'whereas repellency is beneficial to delay the evolution of resistance, other models have shown that it decreases the population-level protection of the insecticide.'
-
-_todo, where to put this, earlier ?_
-This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture is more expensive.
-
-_todo, somewhere earlier say good thing about effectiveness & exposure is that they can be modified_
-
-#### Conclusions
-
-We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured.
-For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance.
-High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other.
-High exposure favours sequences through reducing the otherwise high selection pressures on both resistances together.
-High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced.
-Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them.
-Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.
-
-### Declarations
-TODO see https://malariajournal.biomedcentral.com/submission-guidelines/preparing-your-manuscript/research-article
-
-##### Ethics approval and consent to participate
-Not applicable
-##### Consent for publication
-Not applicable
-##### Availability of data and material
-
-##### Competing interests
-
-##### Funding
-
-##### Authors' contributions
-AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
-IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
-
-##### Acknowledgements
-Thank you to Bethany Levick who developed the first version of the model code.
-
-\pagebreak
-
-**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
-2. Exposure | faster | reduced fitness of SS overall
-3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
-4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
-
-**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | ---------- | ---------------------------------------------
-1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure from the other
-2.
-Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
-3. Dominance of restoration | faster | increased fitness of heterozygotes
-4. Resistance restoration | faster | increased fitness of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
-
-\pagebreak
-
-**Table 4. Effect of inputs on the difference between mixture and sequential use**
-
-Parameter to increase  | increase favours mix or sequence | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
-2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4.
-Resistance restoration | neither | As for dominance of restoration.
-Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
-5.
-Frequency  | neither | As for dominance and resistance restoration.
-Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
-6.
-Cost of resistance | mixture | Higher costs slow the evolution of resistance more in a mixture than when used in sequence (but with higher costs in a sequence there is a greater chance for resistance levels to decline for the insecticide not being used).
-
-\pagebreak
-
-## Figures
-
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="The effect of model inputs on the fitness of genotypes for a single insecticide.
-Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of restoration determines how the fitness for SR lies between that of SS and RR.
-For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3."}
-
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
-
+```{r, results = 'asis', echo=FALSE, fig.width=3, fig.height=3, fig.cap="Figure 1: A smiley face because I work."}
+plot(-5:5, (-5:5)^2)
 ```
-
-![Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
-The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
-Two scenarios (A and B) can be run and the results viewed side by side.
-This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)
-
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
-A.
-Effectiveness, B.
-Exposure, C.
-Dominance of restoration, D.
-Resistance restoration.
-Increasing any of the inputs (going from red to blue) results in shorter times-to-resistance.
-For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
-"}
-
-  library(cowplot)
-
-  #list to put plots in for cowplot
-  plotlist <- list(4)
-
-  #A effectiveness  
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.8,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
-         theme_bw() +  
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +  
-         labs(colour = "effectiveness") +
-         coord_trans(y = "log10") +    
-         geom_line()
-
-  #B exposure
-  exposure <- 0.2
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.5
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.8
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "exposure", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-  #C single insecticide dominance
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.8, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #C 
-  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance of restoration") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #D single insecticide resistance restoration
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.2,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.8,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #D  
-  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "resistance restoration", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
-A.
-Starting frequency, B.
-Cost of resistance.
-Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds.
-Increasing costs of resistance increases time-to-resistance."}
-
-  #single insecticide frequency fig
-
-  #A starting frequency
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.001,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.1,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <-  ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "start\nfrequency") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #B changing costs at dominance 0.5
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0,
-                      h.RS1_00 = 0.5)
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.1,
-                      h.RS1_00 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.2,
-                      h.RS1_00 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "cost of\nresistance") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]], ncol=2, rel_widths=c(1,1), labels='AUTO')  
-  #cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
-Exposure to the insecticides is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of one insecticide is increased from column 1 (A and C) to column2 (B and D).
-On the upper X axis 's' and 'm' indicate where the 50 percent resistance threshold is reached for the sequence and the mixture.
-In the lower right of each panel the the ratio of time-to-resistance for mixture/sequence is shown rounded to 1 decimal place to give an indication of the relative performance of mixtures and sequences.
-A.
-All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C.
-Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D.
-Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence.
-Increasing effectiveness increases times-to-resistance for mixtures and improves their performance relative to sequences.
-Increasing exposure decreases times-to-resistance for mixtures and reduces their performance relative to sequences.
-"}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
-Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-Effectiveness of insecticide1 increased from 0.5 to 0.8, B.
-Effectiveness of insecticide1 increased from 0.5 to 1, C.
-Effectiveness of insecticides 1 and 2 increased to 0.8, D.
-Effectiveness of insecticide2 0.8 and of insecticide1 1.
-With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios."}
-
-#4 panel figure looking at effectiveness changes alone
-
-#test of rearranging figures on a grid
-#eff1  0.8  1
-#eff2
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.5
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
-Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
-Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, B.
-Resistance restoration for insecticide1 increased from 0.5 to 0.8, C.
-Dominance for insecticide1 increased to 0.8, D.
-Resistance restoration and dominance for insecticide1 increased to 0.8.
-Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios."}
-
-#4 panel figure looking at dominance and resistance restoration
-
-#rr1  0.5  0.8
-#dom1
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#testing
-# plot(1)
-# plot(1)
-# plot(1)
-# plot(1)
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
-Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
-Effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.01 to 0.001.
-In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness."}
-
-#4 panel figure looking at effectiveness and starting freq.
-
-#eff1  0.5  0.8
-#strt freq1
-#0.01    1   2
-#0.001    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig 9 cost of 1 resistance, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of cost of resistance on time-to-resistance for mixtures and sequences.
-Row 1 (A and B) same as Fig.
-5 with no cost of resistance, row 2 (C and D) cost of resistance set to 0.15 with dominance of cost set to 0.5.
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-The increase in cost of resistance from A.
-to C.
-removes the advantage of sequence relative to mixture.
-The ratio of time-to-resistance for the mixture divided by that for the sequence goes from 0.
-to 1.
-Costs seem to favour mixtures in these plots.
-Also be aware that costs would lead to a decline in the frequency of resistance for the first insecticide in a sequence when it stops being used."}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Figxx not eval, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
-All control inputs equal at 0.5, male exposure 0.8, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Male exposure decreased from 0.8 to 0.2, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
-
-# not evaluated at the moment
-#4 panel figure looking at male exposure prop and starting freq.
-
-#eff1  0.5  0.8
-#maleprop
-#0.8    1   2
-#0.2    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure =0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-\pagebreak
-
-## References
-<!-- now auto created here-->

### 2017-04-12 inst/documents/paper2/test_linenumbers.tex
test line numbering and double spacing

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={trying to get line numbers working},
            pdfauthor={Andy South and Ian Hastings},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{trying to get line numbers working}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Andy South and Ian Hastings}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{2017-04-11}

\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers

\begin{document}
\maketitle

\subparagraph{Background}\label{background}

Insecticide resistance threatens the control of vectors and particularly
malaria. New insecticides are being developed to address this. Potential
strategies for using new insecticides include mixtures, sequences and
rotations and there are recommendations for how best to implement these
to manage resistance. There is, however, limited recent, accessible
modelling work addressing the evolution of resistance under different
operational strategies. Previous work concludes that preferred
strategies will be situation specific. There is potential to improve the
level of mechanistic understanding within the operational community of
how insecticide resistance can be expected to evolve in response to
different strategies.

This paper develops an accessible, mechanistic picture of how the
evolution of insecticide resistance is likely to respond to potential
intervention strategies to help guide both management and policy. The
aim is to reach an audience unlikely to read a more detailed modelling
paper. We use our model to develop a mechanistic understanding of how
insecticide resistance is expected to increase with the use of
insecticides in isolation, sequence and mixtures. The model flexibly
represents two independent genes coding for resistance to two
insecticides. We look principally at the ability of the insecticides to
kill susceptible mosquitoes, how much resistance counteracts this, the
proportion of mosquitoes that are exposed to insecticides and costs of
resistance.

\subparagraph{Results}\label{results}

We use the model to demonstrate the evolution of resistance under
different inputs and how this fits with intuitive reasoning about the
selection pressure exerted by insecticides. We show that an insecticide
used in a mixture, relative to alone, always prompts slower evolution of
resistance, but resistance to the two insecticides may evolve more
slowly when used in sequence. We show that the ability of insecticides
to kill susceptible mosquitoes (effectiveness), has the most influence
on whether resistance to two insecticides is likely to arise faster in a
mixture or sequence.

\subparagraph{Conclusions}\label{conclusions}

Our model makes more open the process of insecticide resistance
evolution and how it is likely to respond to insecticide use. We provide
a simple user-interface allowing further exploration
(\url{https://andysouth.shinyapps.io/MixSeqResist1}). These tools can
contribute to operational decisions in insecticide resistance
management.

\subsection{Keywords}\label{keywords}

insecticide resistance; public health; mosquitoes; vector-borne
diseases; infectious diseases; malaria; population genetics

\subsection{Background}\label{background-1}

Insecticide resistance is a problem for malaria
\citep{WHO2012}\citep{Ranson2016}\citep{Hemingway2016} other vector
borne diseases \citep{IRAC2011} and agriculture \citep{FAO2012}. Malaria
alone still results in hundreds of thousands of deaths per year. Recent
malaria control efforts have centred on treated bed nets and indoor
residual spraying, both reliant on insecticides. Treated nets were
recently estimated to contribute 68\% and indoor residual spraying 13\%
to averting more than 500 million falciparum malaria cases between 2000
and 2015 \citep{Bhatt2015}. A recent malaria transmission model
\citep{Churcher2016} predicts that even low pyrethroid resistance is
likely to increase malaria incidence in Africa by reducing the
performance of bed nets.

The WHO produced a Global Plan for Insecticide Resistance Management in
malaria vectors (GPIRM)\citep{WHO2012} which includes recommendations on
operational strategies for managing resistance including the use of
insecticide mixtures when they become available. Efforts are under way
to develop new insecticides that will be effective in the light of
existing resistance and allow additional options within insecticide
resistance management. The Innovative Vector Control Consortium (IVCC)
was set up in 2005 to develop new vector control tools and particularly
new insecticides to address insecticide resistance in malaria
transmitting mosquitoes \citep{Hemingway2006}\citep{IVCC2016}. Three new
insecticides are now in development \citep{IVCC2016} and likely to be
available around 2020 \citep{Ranson2016}. It is important that decisions
about how best to use the new insecticides to delay the onset of
resistance are made before insecticides are released
\citep{Hemingway2016}.

Modelling studies have investigated the evolution of insecticide
resistance in insecticide mixtures including in a public health context
e.g. \citep{Curtis1985}\citep{Mani1985}\citep{Roush1989} but much of the
work was done more than 20 years ago and there remained some confusion
about the results \citep{Levick2017}. In an earlier paper
\citep{Levick2017} we described the technical details of a flexible
model used to investigate the relative benefits of mixtures and
sequences. Here we provide an accessible summary of the model and use
selected parameter values to describe mechanistically how the evolution
of resistance is influenced by different inputs. This mechanistic
understanding can contribute to the debate on the relative merits of
different insecticide strategies extending existing frameworks
\citep{Roush1989}\citep{IRAC2011}\citep{FAO2012}\citep{WHO2012}.

My working plot.

\begin{figure}[htbp]
\centering
\includegraphics{test_linenumbers_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{Figure 1: A smiley face because I work.}
\end{figure}

\bibliography{library}

\end{document}

### 2017-04-12 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 ---
-title: "paper2_resistance_mechanisms_mixtures"
-output: pdf_document
+title: "Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary"
+output: 
+  pdf_document:
+    #citation_package: natbib
+    keep_tex: true
+    fig_caption: true
+    latex_engine: pdflatex
+
 #output: word_document
@@
 csl: biomed-central.csl
-#header-includes:
-#    - \usepackage{setspace}
-#    - \doublespacing
-#    - \usepackage{lineno}
-#    - \linenumbers
+fontsize: 11pt
+spacing: double
+# below worked in test_linenumbers.Rmd
+header-includes:
+   - \usepackage{setspace}
+   - \doublespacing
+   - \usepackage{lineno}
+   - \linenumbers
 ---
@@
 
-### version 11
-
-#Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary.
+### version 12
 
@@
 
-Target : Malaria journal
-
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and suggestions welcome
+Target : Malaria journal (other suggestions welcome)
+
+THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome
 
@@
 #### Implications
-_todo get more of [@Roush1989] in here_
 
@@
 
-_low exposure could be achieved by having active ingredients with some repellency._
-
-_todo get Birget and Koella in & link to exposure_
-'whereas repellency is beneficial to delay the evolution of resistance, other models have shown that it decreases the population-level protection of the insecticide.'
-
-_todo, where to put this, earlier ?_
-This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
+_TODO bits to incorporate into discussion :_
+
+_todo more reference to [@Roush1989] who made many of these points 25 years ago_
+
+_Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mictures over sequences.
+This would need to be balanced with a likely loss in protection to the population as a whole provided by the insecticide [@Birget2015]._
+
+_This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
 Other factors also need to be considered.
 On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture is more expensive.
+On the negative side a mixture is more expensive._
 
@@
 High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other.
-High exposure favours sequences through reducing the otherwise high selection pressures on both resistances together.
+High exposure favours sequences because of the higher selection pressures on both resistances together in a mixture.
 High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced.
@@
 Not applicable
+
 ##### Consent for publication
 Not applicable
+
 ##### Availability of data and material

### 2017-04-12 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.tex
paper writing

--- 2017-04-12 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
----
-title: "Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary"
-output: 
-  pdf_document:
-    #citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-
-#output: word_document
-author: "Andy South and Ian Hastings"
-date: "`r Sys.Date()`"
-bibliography: library.bib
-csl: biomed-central.csl
-fontsize: 11pt
-spacing: double
-# below worked in test_linenumbers.Rmd
-header-includes:
-   - \usepackage{setspace}
-   - \doublespacing
-   - \usepackage{lineno}
-   - \linenumbers
----
-
-```{r, eval=FALSE, include=FALSE}
-#!!!! paper protocol
-#!!!! might be easier to put text and figs into separate Rmd files after draft stage
-#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
-#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
-#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
-#!!!! submission figs, set this chunk to eval=TRUE
-
-# options to create final publication quality figures
-# causes the document to fail, but the figs do get stored in pap2figs/
-library(knitr)
-opts_chunk$set(dev="tiff",
-               dev.args=list(compression="lzw"),
-               dpi=300,
-               cache=FALSE,
-               fig.path='pap2figs/')
-```
-
-```{r,echo=FALSE,results='hide',warning=FALSE}
-  library(resistance)
-  library(ggplot2)
-```
-
-### version 12
-
-###alternative titles:
-###Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
-###An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
-###A tutorial, primer ...
-###Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
+\documentclass[11pt,]{article}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary},
+            pdfauthor={Andy South and Ian Hastings},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+  \title{Insecticide resistance evolution with mixtures and sequences : a
+model-based graphical introduction and summary}
+  \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+  \author{Andy South and Ian Hastings}
+  \preauthor{\centering\large\emph}
+  \postauthor{\par}
+  \predate{\centering\large\emph}
+  \postdate{\par}
+  \date{2017-04-11}
+
+\usepackage{setspace}
+\doublespacing
+\usepackage{lineno}
+\linenumbers
+
+\begin{document}
+\maketitle
+
+\subsubsection{version 12}\label{version-12}
+
+\subsubsection{alternative titles:}\label{alternative-titles}
+
+\subsubsection{Evolution of insecticide resistance with insecticide
+mixtures and sequences : an accessible model to promote
+understanding.}\label{evolution-of-insecticide-resistance-with-insecticide-mixtures-and-sequences-an-accessible-model-to-promote-understanding.}
+
+\subsubsection{An accessible model to promote understanding of the
+evolution of insecticide resistance : insecticide effectiveness
+predicted to be key to the relative performance of
+mixtures.}\label{an-accessible-model-to-promote-understanding-of-the-evolution-of-insecticide-resistance-insecticide-effectiveness-predicted-to-be-key-to-the-relative-performance-of-mixtures.}
+
+\subsubsection{A tutorial, primer \ldots{}}\label{a-tutorial-primer}
+
+\subsubsection{Insecticide effectiveness predicted to be most important
+factor determining whether insecticide mixtures can slow evolution of
+resistance.}\label{insecticide-effectiveness-predicted-to-be-most-important-factor-determining-whether-insecticide-mixtures-can-slow-evolution-of-resistance.}
 
@@
 
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome
-
-## Abstract
-
-_must be 350 words, currently 351_
-
-##### Background
-Insecticide resistance threatens the control of vectors and particularly malaria.
-New insecticides are being developed to address this.
-Potential strategies for using new insecticides include mixtures, sequences and rotations and there are recommendations for how best to implement these to manage resistance.
-There is, however, limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
-Previous work concludes that preferred strategies will be situation specific.
-There is potential to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
-
-This paper develops an accessible, mechanistic picture of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-The aim is to reach an audience unlikely to read a more detailed modelling paper.
-We use our model to develop a mechanistic understanding of how insecticide resistance is expected to increase with the use of insecticides in isolation, sequence and mixtures.
-The model flexibly represents two independent genes coding for resistance to two insecticides.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, how much resistance counteracts this, the proportion of mosquitoes that are exposed to insecticides and costs of resistance.
-
-##### Results
-We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about the selection pressure exerted by insecticides.
-We show that an insecticide used in a mixture, relative to alone, always prompts slower evolution of resistance, but resistance to the two insecticides may evolve more slowly when used in sequence.
-We show that the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on whether resistance to two insecticides is likely to arise faster in a mixture or sequence.
-
-##### Conclusions
-Our model makes more open the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
-We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/MixSeqResist1).
-These tools can contribute to operational decisions in insecticide resistance management.
-
-## Keywords
-insecticide resistance; public health; mosquitoes; vector-borne diseases; infectious diseases; malaria; population genetics
-
-## Background
-
-Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
-Malaria alone still results in hundreds of thousands of deaths per year.
-Recent malaria control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
-Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million falciparum malaria cases between 2000 and 2015 [@Bhatt2015].
-A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
-
-The WHO produced a Global Plan for Insecticide Resistance Management in malaria vectors (GPIRM)[@WHO2012] which includes recommendations on operational strategies for managing resistance including the use of insecticide mixtures when they become available.
-Efforts are under way to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
-The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
-Three new insecticides are now in development [@IVCC2016] and likely to be available around 2020 [@Ranson2016].
-It is important that decisions about how best to use the new insecticides to delay the onset of resistance are made before insecticides are released [@Hemingway2016].
-
-Modelling studies have investigated the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
-[@Curtis1985][@Mani1985][@Roush1989] but much of the work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
-In an earlier paper [@Levick2017] we described the technical details of a flexible model used to investigate the relative benefits of mixtures and sequences.
-Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance is influenced by different inputs.
-This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
-
-Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
-It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
-
-## Methods
-
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
-One gene or locus is represented for each insecticide.
-Each individual has two copies of the gene and there are two potential alleles per locus.
-Each allele confers either resistance or susceptibility to the insecticide.
-Thus individuals can either have both resistance genes (RR homozygous), both susceptible genes (SS) or one of each (SR, heterozygous).
-The combination of the two genes is termed the genotype.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
-The fitness of the SS genotype in absence of the insecticide is set at a maximum value of 1.
-All other fitnesses are less, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
-
-**Table 1. Parameters inflencing the development of insecticide resistance**
-
-Parameter  | Description
----------------------- | ----------------------------------------------------
-1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
-2. Exposure | proportion of insects exposed to insecticide
-3. Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide
-4. Dominance of restoration | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
-5. Frequency  | frequency of resistance alleles within the population
-6. Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide
-7. Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide
-
-The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 1.
-In the model itself this is repeated for both insecticides and the fitnesses multiplied.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS and resistance restoration 'restores' a portion of the fitness for RR.
-A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
-We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
-Cost of resistance determines the fitness of RR in the portion of the population not exposed to the insecticide.
-
-Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
-For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
-Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Figure 1).
-
-The simulation proceeds through generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
-This acts to make the fitter alleles more common over time.
-Separate sexes and standard sexual reproduction with recombination is included.
-There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that here.
-
-So far this description just considers a single insecticide and associated resistance allele.
-In the model itself a second insecticide and resistance allele are included and overall fitness is calculated by multiplying the two results.
-This allows the model to represent populations exposed to two insecticides together in a mixture.
-
-The model outputs the change in resistance allele frequencies over time measured in generations.
-We extract the number of generations at which a resistance allele frequency of 50% is reached and term this 'time-to-resistance'.
-Using a 50% threshold is consistent with previous modelling analyses e.g.
-[@Curtis1985][@Birget2015].
-Preliminary investigations showed that using other potential thresholds of 25 or 75% gave very similar results.
+THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without
+the permission of one of the authors. Comments and suggestions welcome
+
+\subsection{Abstract}\label{abstract}
+
+\emph{must be 350 words, currently 351}
+
+\subparagraph{Background}\label{background}
+
+Insecticide resistance threatens the control of vectors and particularly
+malaria. New insecticides are being developed to address this. Potential
+strategies for using new insecticides include mixtures, sequences and
+rotations and there are recommendations for how best to implement these
+to manage resistance. There is, however, limited recent, accessible
+modelling work addressing the evolution of resistance under different
+operational strategies. Previous work concludes that preferred
+strategies will be situation specific. There is potential to improve the
+level of mechanistic understanding within the operational community of
+how insecticide resistance can be expected to evolve in response to
+different strategies.
+
+This paper develops an accessible, mechanistic picture of how the
+evolution of insecticide resistance is likely to respond to potential
+intervention strategies to help guide both management and policy. The
+aim is to reach an audience unlikely to read a more detailed modelling
+paper. We use our model to develop a mechanistic understanding of how
+insecticide resistance is expected to increase with the use of
+insecticides in isolation, sequence and mixtures. The model flexibly
+represents two independent genes coding for resistance to two
+insecticides. We look principally at the ability of the insecticides to
+kill susceptible mosquitoes, how much resistance counteracts this, the
+proportion of mosquitoes that are exposed to insecticides and costs of
+resistance.
+
+\subparagraph{Results}\label{results}
+
+We use the model to demonstrate the evolution of resistance under
+different inputs and how this fits with intuitive reasoning about the
+selection pressure exerted by insecticides. We show that an insecticide
+used in a mixture, relative to alone, always prompts slower evolution of
+resistance, but resistance to the two insecticides may evolve more
+slowly when used in sequence. We show that the ability of insecticides
+to kill susceptible mosquitoes (effectiveness), has the most influence
+on whether resistance to two insecticides is likely to arise faster in a
+mixture or sequence.
+
+\subparagraph{Conclusions}\label{conclusions}
+
+Our model makes more open the process of insecticide resistance
+evolution and how it is likely to respond to insecticide use. We provide
+a simple user-interface allowing further exploration
+(\url{https://andysouth.shinyapps.io/MixSeqResist1}). These tools can
+contribute to operational decisions in insecticide resistance
+management.
+
+\subsection{Keywords}\label{keywords}
+
+insecticide resistance; public health; mosquitoes; vector-borne
+diseases; infectious diseases; malaria; population genetics
+
+\subsection{Background}\label{background-1}
+
+Insecticide resistance is a problem for malaria {[}1{]}{[}2{]}{[}3{]}
+other vector borne diseases {[}4{]} and agriculture {[}5{]}. Malaria
+alone still results in hundreds of thousands of deaths per year. Recent
+malaria control efforts have centred on treated bed nets and indoor
+residual spraying, both reliant on insecticides. Treated nets were
+recently estimated to contribute 68\% and indoor residual spraying 13\%
+to averting more than 500 million falciparum malaria cases between 2000
+and 2015 {[}6{]}. A recent malaria transmission model {[}7{]} predicts
+that even low pyrethroid resistance is likely to increase malaria
+incidence in Africa by reducing the performance of bed nets.
+
+The WHO produced a Global Plan for Insecticide Resistance Management in
+malaria vectors (GPIRM){[}1{]} which includes recommendations on
+operational strategies for managing resistance including the use of
+insecticide mixtures when they become available. Efforts are under way
+to develop new insecticides that will be effective in the light of
+existing resistance and allow additional options within insecticide
+resistance management. The Innovative Vector Control Consortium (IVCC)
+was set up in 2005 to develop new vector control tools and particularly
+new insecticides to address insecticide resistance in malaria
+transmitting mosquitoes {[}8{]}{[}9{]}. Three new insecticides are now
+in development {[}9{]} and likely to be available around 2020 {[}2{]}.
+It is important that decisions about how best to use the new
+insecticides to delay the onset of resistance are made before
+insecticides are released {[}3{]}.
+
+Modelling studies have investigated the evolution of insecticide
+resistance in insecticide mixtures including in a public health context
+e.g. {[}10{]}{[}11{]}{[}12{]} but much of the work was done more than 20
+years ago and there remained some confusion about the results {[}13{]}.
+In an earlier paper {[}13{]} we described the technical details of a
+flexible model used to investigate the relative benefits of mixtures and
+sequences. Here we provide an accessible summary of the model and use
+selected parameter values to describe mechanistically how the evolution
+of resistance is influenced by different inputs. This mechanistic
+understanding can contribute to the debate on the relative merits of
+different insecticide strategies extending existing frameworks
+{[}12{]}{[}4{]}{[}5{]}{[}1{]}.
+
+Our modelling approach focuses on the change in frequency of a single
+resistance gene for each insecticide. It is likely that genes conferring
+resistance are present in populations even prior to exposure to novel
+insecticides thus leading to the potential for selection {[}14{]}.
+
+\subsection{Methods}\label{methods}
+
+The simulation represents a population of randomly mixing individuals
+using standard population genetic approaches to avoid the need to follow
+every individual. One gene or locus is represented for each insecticide.
+Each individual has two copies of the gene and there are two potential
+alleles per locus. Each allele confers either resistance or
+susceptibility to the insecticide. Thus individuals can either have both
+resistance genes (RR homozygous), both susceptible genes (SS) or one of
+each (SR, heterozygous). The combination of the two genes is termed the
+genotype. Fitness is the main currency of the model, representing how
+much each genotype survives and reproduces. The fitness of the SS
+genotype in absence of the insecticide is set at a maximum value of 1.
+All other fitnesses are less, determined by alleles, exposure to the
+insecticide and other inputs that can be set in the model. These inputs
+include the effectiveness of the insecticide, the dominance of the
+resistance allele and the ability of resistance to restore fitness in
+the presence of the insecticide (see Table 1).
+
+\textbf{Table 1. Parameters inflencing the development of insecticide
+resistance}
+
+\begin{longtable}[]{@{}ll@{}}
+\toprule
+\begin{minipage}[b]{0.28\columnwidth}\raggedright\strut
+Parameter\strut
+\end{minipage} & \begin{minipage}[b]{0.66\columnwidth}\raggedright\strut
+Description\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+proportion of susceptible (SS) insects killed by exposure to
+insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+proportion of insects exposed to insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+3. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+ability of resistance (RR) to restore fitness of insects exposed to
+insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+4. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+sets fitness of heterozygous (SR) insects between that of SS \& RR in
+presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+frequency of resistance alleles within the population\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+fitness of resistant (RR) insects in absence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+7. Dominance of cost\strut
+\end{minipage} & \begin{minipage}[t]{0.66\columnwidth}\raggedright\strut
+sets fitness of heterozygous (SR) insects between that of SS \& RR in
+absence of insecticide\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+The calculation of fitness for each genotype when considering just a
+single insecticide is illustrated in Figure 1. In the model itself this
+is repeated for both insecticides and the fitnesses multiplied. Firstly
+the exposure input determines the proportion of the population in the
+left and right panels (exposed and not exposed). For those that are
+exposed (left panel) insecticide effectiveness sets the fitness for SS
+and resistance restoration `restores' a portion of the fitness for RR. A
+value of 1 for resistance restoration would bring the fitness of RR back
+to 1 (equivalent to the unexposed SS), a value of 0 for resistance
+restoration would lead to the RR having the same fitness as SS in the
+presence of the insecticide and effectively there would be no
+resistance. We use `Resistance restoration' to represent how much the RR
+genotype restores fitness in the presence of the insecticide. The
+`selection coefficient' of resistance, that may be seen referred to
+elsewhere {[}10{]}, can be calculated by multiplying resistance
+restoration by insecticide effectiveness. Cost of resistance determines
+the fitness of RR in the portion of the population not exposed to the
+insecticide.
+
+Dominance of restoration determines how the fitness for SR sits between
+that of SS and RR for those that are exposed to the insecticide. For
+those that are not exposed, fitness of SS is set to 1 by definition and
+resistance cost determines the fitness of RR. Dominance of cost
+determines how the fitness for SR sits between that of SS and RR in the
+absence of the insecticide (Figure 1).
+
+The simulation proceeds through generations. In each generation
+selection is represented by multiplying genotype frequencies in the
+population by their relative fitness. This acts to make the fitter
+alleles more common over time. Separate sexes and standard sexual
+reproduction with recombination is included. There is also the potential
+to set exposure rates to be different for males and females to represent
+their different behaviours but we do not explore that here.
+
+So far this description just considers a single insecticide and
+associated resistance allele. In the model itself a second insecticide
+and resistance allele are included and overall fitness is calculated by
+multiplying the two results. This allows the model to represent
+populations exposed to two insecticides together in a mixture.
+
+The model outputs the change in resistance allele frequencies over time
+measured in generations. We extract the number of generations at which a
+resistance allele frequency of 50\% is reached and term this
+`time-to-resistance'. Using a 50\% threshold is consistent with previous
+modelling analyses e.g. {[}10{]}{[}15{]}. Preliminary investigations
+showed that using other potential thresholds of 25 or 75\% gave very
+similar results.
 
@@
 
-1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
-1. two insecticides used in a mixture (concurrently)
-
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an on-line user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
-
-## Results
-
-### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
-This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
-Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that less change is required to reach the threshold.
-
-Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect.
-Higher values led to a slower development of resistance.
-Again this is intuitive because costs reduce the advantage of resistance and decrease selection.
-These effects of inputs on single insecticide use are summarised in Table 2.
-
-### Two insecticides
-When two insecticides are used we can compare the relative performance of a mixture with sequential use.
-Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
-To help understand the mechanisms influencing this relative performance we start with a single base scenario and demonstrate changing inputs individually.
-In this base scenario we set effectiveness, exposure, resistance restoration, and dominance at 0.5, the starting frequencies of resistance to 0.01 and cost of resistance to 0.
-Resistance arises slower for sequential than mixture for this base scenario (Fig 5A).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 5A).
-Thus the time-to-resistance for the mixture divided by the sequence is less than 1 at 0.8.
-
-#### Insecticide exposure and effectiveness
-
-Increasing exposure (the proportion of insects that come into contact with the insecticides) (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy.
-The effect on the mixture is greater so time-to-resistance remains longer for the sequence (the ratio decreasing from 0.8 to 0.6).
-In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 5B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
-The longer time-to-resistance for the mixture results from a slower increase in resistance for the less effective insecticide.
-Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
-This points a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
-
-The results is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
-This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-
-In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 5, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 5.
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
-
-Increasing the effectiveness further of either or both insecticides increases the performance of mixtures over sequences in terms of time-to-resistance.
-Figure 6A uses the same inputs as Fig 5B.
-From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
-This again points to how more effective insecticides can slow resistance to another when used in a mixture.
-
-#### Resistance restoration and dominance
-Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
-The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
-So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
-
-#### Starting frequencies of resistance
-Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
-Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
-
-#### Costs of resistance
-Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
-The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 9C).
-The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C.
-With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 9A and 1 in Fig 9C.
-At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
-
-A summary of the effects of inputs on the evolution of resistance for mixtures is given in Table 3 and for the difference between mixtures and sequences in Table 4.
-The mechanisms identified in the tables are covered further in the discussion.
-
-#### Linkage disequilibrium
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
-
-## Discussion
-
-Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
-This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
-
-When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resistance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
-
-These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Referring to figure 1 can help explain these.
-Exposure and effectiveness act by decreasing the fitness of all genotypes, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
-Cost acts by decreasing the fitness of RR and SR for those not exposed.
-
-#### Two insecticides
-Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide and, where it differs can again help us to understand the dynamics of resistance evolution.
-Our base scenario (Fig 5A) shows curves for resistance to both insecticides in the mixture that are shallower than those for sequential use.
-The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Thus each insecticide reduces the selection pressure for the increase in resistance to the other and could be said to 'protect' the other.
-However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
-Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
-Incraesing the effectiveness of on insecticide to Fig 5B decreases the added times for the sequence to 90 and increases those for the mixture to 115 thus changing the order and favouring the mixture.
-
-The mechanism for this change can be sen by comparing the shape of the curves in Figs 5A & B.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide, loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide.
-The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-
-#### Difference between exposure and effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us.
-
-#### Costs of resistance
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-It seems that the protective effect of the other insecticide in a mixture interacts with the cost of resistance to reduce selection pressure more than in a sequence.
-
-The effects of cost of resistance become more as the frequency of resistance increases.
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
-Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
-Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
-Also note that if the costs are set even higher than we set them here (e.g.
-above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-Unfortunately, we do not expect such high costs to occur in operational settings.
-
-Relatedly, when there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
-Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where it becomes operationally useful again.
-As such a repeated sequential strategy could offer benefits over mixtures not considered here.
-To incorporat this one would need to use a more detailed metric than time-to-resistance.
-
-#### Caveats
-Our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as effecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
-Insecticide resistance could reduce vector transmission of disease (e.g.
-by reduced lifespan of the vector) or increase transmission (e.g.
-by reduced immunity of the vector to the disease) [@Rivero2010].
-Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Of course, using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
-Others are combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
-
-In our model we represented insecticide interaction as being additive within the mixture.
-This is consistent with recent work on aphids [@Taillebois2016], although that work did also indicate some synergistic effects.
-Our model has the potential to incorporate either synergistic or antagonistic effects too.
-
-The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ranson2016].
-The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-Thus there is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown.
-
-#### Consistency with previous work 
-The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
-
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
-Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
-
-One consistent aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
-
-Poor implementation would likely cause lower effectiveness of and exposure to insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Barbosa2011][@FAO2012].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that effectivenes favours mixtures, exposure favours sequences and higher dominance favours neither (Table 4).
-
-#### Implications
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the development of resistance.
-
-_TODO bits to incorporate into discussion :_
-
-_todo more reference to [@Roush1989] who made many of these points 25 years ago_
-
-_Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mictures over sequences.
-This would need to be balanced with a likely loss in protection to the population as a whole provided by the insecticide [@Birget2015]._
-
-_This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture is more expensive._
-
-_todo, somewhere earlier say good thing about effectiveness & exposure is that they can be modified_
-
-#### Conclusions
-
-We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured.
-For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance.
-High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other.
-High exposure favours sequences because of the higher selection pressures on both resistances together in a mixture.
-High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced.
-Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them.
-Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.
-
-### Declarations
-TODO see https://malariajournal.biomedcentral.com/submission-guidelines/preparing-your-manuscript/research-article
-
-##### Ethics approval and consent to participate
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  single insecticide
+\item
+  two insecticides used in sequence, replacing the first with a second
+  once the frequency of resistance alleles reaches a threshold of 50\%
+\item
+  two insecticides used in a mixture (concurrently)
+\end{enumerate}
+
+The model is implemented in R {[}16{]}, the code is hosted on Github
+{[}17{]} and we provide an on-line user interface {[}18{]} (Fig 2)
+enabling the reader to change inputs and run the model themselves.
+
+\subsection{Results}\label{results-1}
+
+\subsubsection{Single insecticide}\label{single-insecticide}
+
+For single insecticide use, higher values of insecticide effectiveness,
+exposure, resistance restoration and its dominance all resulted in
+faster resistance spread (Fig 3 A-D) and thus fewer generations to reach
+a resistance threshold of 0.5. This makes intuitive sense as all
+increase the impact of the insecticide and thus increase selection for
+resistance. Similarly, higher values of starting resistance frequency
+also resulted in shorter times to resistance thresholds (Fig 4A) through
+the even simpler mechanism that less change is required to reach the
+threshold.
+
+Cost of resistance (Fig 4B) and its dominance (not shown) had the
+opposite effect. Higher values led to a slower development of
+resistance. Again this is intuitive because costs reduce the advantage
+of resistance and decrease selection. These effects of inputs on single
+insecticide use are summarised in Table 2.
+
+\subsubsection{Two insecticides}\label{two-insecticides}
+
+When two insecticides are used we can compare the relative performance
+of a mixture with sequential use. Each panel in Figures 5-9 compares
+mixture and sequence for a single combination of inputs. To help
+understand the mechanisms influencing this relative performance we start
+with a single base scenario and demonstrate changing inputs
+individually. In this base scenario we set effectiveness, exposure,
+resistance restoration, and dominance at 0.5, the starting frequencies
+of resistance to 0.01 and cost of resistance to 0. Resistance arises
+slower for sequential than mixture for this base scenario (Fig 5A).
+Resistance to both insecticides in the mixture follow the same path and
+reach the threshold at the same time (as would be expected given that
+they have identical input parameters). When used in sequence the curve
+(dashed) for each insecticide individually is steeper than in the
+mixture but, because they happen one after the other, it takes longer
+for both to reach the resistance threshold (Fig 5A). Thus the
+time-to-resistance for the mixture divided by the sequence is less than
+1 at 0.8.
+
+\paragraph{Insecticide exposure and
+effectiveness}\label{insecticide-exposure-and-effectiveness}
+
+Increasing exposure (the proportion of insects that come into contact
+with the insecticides) (Fig 5C) decreases the time-to-resistance for
+both the sequential and mixture strategy. The effect on the mixture is
+greater so time-to-resistance remains longer for the sequence (the ratio
+decreasing from 0.8 to 0.6). In contrast keeping the exposure constant
+and increasing effectiveness (the proportion of SS insects that are
+killed by contact) of one of the insecticides (Fig 5B) results in a
+longer time-to-resistance for the mixture relative to the sequence (and
+the ratio increasing from 0.8 to 1.2). The longer time-to-resistance for
+the mixture results from a slower increase in resistance for the less
+effective insecticide. Resistance to the less effective insecticide in
+the mixture increases slowly initially and then speeds up after the more
+effective insecticide has reached the resistance threshold. This points
+a mechanism whereby the more effective insecticide initially slows the
+rate of evolution of resistance to the less effective one.
+
+The results is that a more effective insecticide increases
+time-to-resistance when used in a mixture (compare the solid lines in
+Fig 5A \& B). This is opposite to what happens when used in sequence
+(compare the red dashed line in Fig 5A \& B) and in isolation (Fig 2A),
+when more effective insecticides shortened times-to-resistance.
+
+In summary increasing effectiveness can favour mixtures moving from
+column 1 to 2 in Figure 5, and increasing exposure can favour sequences
+moving from row 1 to 2 in Figure 5. Reduced male exposure with respect
+to female, as might be expected for male mosquitoes not seeking blood
+meals, had a similar effect to changing overall exposure. Reducing male
+exposure (not shown) increased times-to-resistance and favoured mixtures
+over sequences.
+
+Increasing the effectiveness further of either or both insecticides
+increases the performance of mixtures over sequences in terms of
+time-to-resistance. Figure 6A uses the same inputs as Fig 5B. From this
+scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide
+1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference
+in time-to-resistance for mixture over sequence. This again points to
+how more effective insecticides can slow resistance to another when used
+in a mixture.
+
+\paragraph{Resistance restoration and
+dominance}\label{resistance-restoration-and-dominance}
+
+Increasing resistance restoration or its dominance decreases
+time-to-resistance for both sequences and mixtures (Fig 7) in the same
+way that it did for sole use (Fig 3). The result is that increasing
+either or both of dominance and resistance restoration (Figs 7B,C,D)
+doesn't change the relative time-to-resistance for mixtures and
+sequences from that in the base scenario (Fig 7A) where it takes
+resistance longer to develop for the sequence. So, unlike effectiveness
+and exposure, the levels of resistance restoration and dominance do not
+effect whether a mixture or a sequence are likely to be favoured.
+
+\paragraph{Starting frequencies of
+resistance}\label{starting-frequencies-of-resistance}
+
+Changing the starting frequency of resistance had a similar effect on
+time to resistance for both sequences and mixtures and thus had little
+effect on their relative performance. For example taking the base
+scenario and reducing the starting frequency of one resistance allele
+did not change from sequence being favoured (compare Fig 8C to 8A).
+Similarly taking a scenario in which time-to-resistance is longer for a
+mixture (Fig 8B) and decreasing the starting frequency of that
+resistance allele did not change the fact that the mixture was favoured
+(Fig 8D).
+
+\paragraph{Costs of resistance}\label{costs-of-resistance}
+
+Increasing the cost of resistance to one insecticide in a mixture leads
+resistance to that insecticide to evolve more slowly (compare the solid
+red line to the solid blue line in Figure 9C, the red line is for the
+resistance with the higher cost). The resistance with the higher cost
+also increases more slowly in the mixture relative to when it is used in
+sequence (compare the solid red line to the dotted red line in Fig 9C).
+The greater effect of cost on mixtures means the advantage of sequences
+over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C. With costs
+set to 0 time-to-resistance for mixture divided by sequence is 0.8 in
+Fig 9A and 1 in Fig 9C. At higher insecticide effectiveness the benefit
+of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
+
+A summary of the effects of inputs on the evolution of resistance for
+mixtures is given in Table 3 and for the difference between mixtures and
+sequences in Table 4. The mechanisms identified in the tables are
+covered further in the discussion.
+
+\paragraph{Linkage disequilibrium}\label{linkage-disequilibrium}
+
+It has previously been suggested {[}10{]}{[}1{]} that using an
+insecticide to which there was existing resistance in a mixture with a
+new insecticide with little resistance could lead to more rapid
+evolution of resistance to the latter through a process termed linkage
+disequilibrium. Figure 8 C\&D show how our model predicts that linkage
+disequilibrium does not lead to more rapid evolution of resistance to
+the newer insecticide. In Figures 8 C\&D the newer insecticide with
+lower starting resistance (in red) leads to faster resistance when used
+alone (red dotted) than when used in a mixture (red solid).
+
+\subsection{Discussion}\label{discussion}
+
+Resistance responds to insecticide use in the model in ways that we can
+explain mechanistically through the process of selection. This
+mechanistic explanation can help us develop a more robust understanding
+of how we would expect resistance to evolve in the field.
+
+When single insecticide use was investigated (Figs 3-4) higher values of
+inputs that increased the selective advantage of resistance all led to
+more rapid evolution resistance, and higher values for cost led to
+slower evolution of (Fig 4B).
+
+These mechanisms, explaining the response of resistance to single
+insecticide use under different input values, are summarised in Table 2.
+Referring to figure 1 can help explain these. Exposure and effectiveness
+act by decreasing the fitness of all genotypes, dominance of restoration
+acts by increasing fitness of SR and resistance-restoration acts by
+increasing fitness of RR \& SR. Cost acts by decreasing the fitness of
+RR and SR for those not exposed.
+
+\paragraph{Two insecticides}\label{two-insecticides-1}
+
+Using two insecticides either in sequence or in mixtures differs from
+the use of a single insecticide and, where it differs can again help us
+to understand the dynamics of resistance evolution. Our base scenario
+(Fig 5A) shows curves for resistance to both insecticides in the mixture
+that are shallower than those for sequential use. The shallower curves
+for the mixture can be explained by each insecticide killing individuals
+that are resistant to the other insecticide. Thus each insecticide
+reduces the selection pressure for the increase in resistance to the
+other and could be said to `protect' the other. However this protection
+is not necessarily enough to ensure that resistance arises more slowly
+in a mixture relative to a sequential strategy. Resistance can evolve
+more slowly for both insecticides in sequential use, despite being
+faster for each, because they occur one after the other. In Figure 5A it
+takes 60 generations for resistance to reach the threshold for each
+insecticide in the sequence, and 90 when they are applied together in
+the mixture, so the sequence is still favoured because 2x60 is greater
+than 90. Incraesing the effectiveness of on insecticide to Fig 5B
+decreases the added times for the sequence to 90 and increases those for
+the mixture to 115 thus changing the order and favouring the mixture.
+
+The mechanism for this change can be sen by comparing the shape of the
+curves in Figs 5A \& B. In the mixture, resistance to insecticide 1
+(with the increased effectiveness) rises at a similar speed in both
+figures. However, the increased effectiveness of insecticide 1 causes
+resistance to insecticide 2 to increase more slowly initially in the
+mixture. Once insecticide 1 reaches resistance of around 50\% at around
+50 generations, the curve for insecticide 2 becomes steeper. Once the
+first insecticide has become ineffective due to resistance it stops
+killing individuals that are resistant to the second insecticide, loses
+its protective effect and thus stops slowing the rise in resistance to
+the second insecticide. A similar effect is visible when the
+effectiveness of both insecticides are increased (Fig 6C). The identical
+resistance curves for each insecticide in the mixture are shallower than
+at the lower effectiveness (Fig 5A) because more individuals resistant
+to each insecticide are being killed by the other insecticide. In this
+case the `protection' given to both insecticides by the other declines
+at the same rate so we don't see the change in slope seen in Fig 5B.
+
+Fig 5B also demonstrates that when two insecticides have a different
+effectiveness in a mixture, other parameters being equal, resistance
+will be expected to increase faster to the more effective insecticide.
+The more effective insecticide prompts both a) higher selection pressure
+for it's own resistance and b) greater `protection' reducing the rise in
+resistance to the other insecticide.
+
+\paragraph{Difference between exposure and
+effectiveness}\label{difference-between-exposure-and-effectiveness}
+
+It seems initially counter-intuitive that whereas increasing
+effectiveness slows the development of resistance for mixtures,
+increasing exposure speeds it up. Increasing the exposure to both
+insecticides results in a decrease in time-to-resistance (from 90
+generations in Fig 5A to 40 generations in Fig 5C) where increasing the
+effectiveness of both insecticides by the same amount results in a
+slight increase (from 90 in Fig 5A to 115 in Fig 5B). This contrasts
+with the identical effects that exposure and effectiveness have on a
+single insecticide (Figs 3 A,B) and on insecticides in sequence (the red
+dashed lines are the same in Figs 5B \& 5C). This difference, that
+exposure reduces times-to-resistance for a mixture and effectiveness
+increases it, was unexpected for us.
+
+\paragraph{Costs of resistance}\label{costs-of-resistance-1}
+
+Costs of resistance favour mixtures in these analyses because the
+evolution of resistance is slowed more by cost in a mixture than in a
+sequence (compare the red solid lines for a mixture to the red dotted
+line for a sequence in Fig 9C and D). It seems that the protective
+effect of the other insecticide in a mixture interacts with the cost of
+resistance to reduce selection pressure more than in a sequence.
+
+The effects of cost of resistance become more as the frequency of
+resistance increases. The resistance curves for two insecticides in a
+mixture that differ only in their cost of resistance, diverge more as
+the frequency of resistance increases (Fig 9C). This makes intuitive
+sense, as the frequency of resistance genes increases the impact of
+fitness costs associated with those genes becomes greater. Thus
+differences due to cost are less noticeable at low resistance
+frequencies and become more noticeable at higher resistance frequencies.
+Thus if a lower resistance threshold was used cost would have less of an
+effect and if a higher resistance threshold was used cost would have
+more of an effect. Also note that if the costs are set even higher than
+we set them here (e.g.~above 0.2) they can maintain resistance
+frequencies below 1 or lead to their decline. Unfortunately, we do not
+expect such high costs to occur in operational settings.
+
+Relatedly, when there are costs of resistance, the frequency of
+resistance for insecticides that are not being used is likely to
+decline. Thus the frequency of resistance for the first insecticide in a
+sequence, after it has been replaced, may decline to a level where it
+becomes operationally useful again. As such a repeated sequential
+strategy could offer benefits over mixtures not considered here. To
+incorporat this one would need to use a more detailed metric than
+time-to-resistance.
+
+\paragraph{Caveats}\label{caveats}
+
+Our model examines solely the evolution of resistance and not the
+control of mosquitoes or malaria. Different insecticide use strategies,
+as well as effecting the evolution of resistance, will also influence
+both the numbers of vectors and their propensity to transmit the disease
+{[}19{]}. Insecticide resistance could reduce vector transmission of
+disease (e.g.~by reduced lifespan of the vector) or increase
+transmission (e.g.~by reduced immunity of the vector to the disease)
+{[}19{]}. Our model would predict that the best strategy for limiting
+the development of resistance in the short term would be to use no
+insecticide. Of course, using no insecticide could have serious
+consequences for mosquito and malaria abundance. The challenge is to
+develop good strategies for delivering mosquito and malaria control in
+the short term while sustaining their operational capacity in the longer
+term by restricting the development of resistance. Others are combining
+models of population genetics and population dynamics to investigate
+these trade-offs {[}20{]}.
+
+In our model we represented insecticide interaction as being additive
+within the mixture. This is consistent with recent work on aphids
+{[}21{]}, although that work did also indicate some synergistic effects.
+Our model has the potential to incorporate either synergistic or
+antagonistic effects too.
+
+The development of insecticide resistance in operational field settings
+is subject to considerable uncertainty {[}2{]}. The model we present
+here is able to cover the range of that uncertainty, yet in this paper,
+to develop an understanding of the key mechanisms we restrict our
+explorations to a limited region of parameter space (e.g.~holding most
+inputs constant at intermediate values while modifying single inputs in
+isolation). Thus there is the potential that the system may behave
+differently in the field than we describe here. However, the results
+presented here are supported by an earlier analysis where we ran 10,000
+unique scenarios varying more inputs {[}13{]}. This previous analysis
+also highlighted the relative importance of insecticide effectiveness
+and exposure in determining the relative performance of mixtures and
+sequences. In addition we provide an on-line user interface to the model
+{[}18{]} and enabling readers to investigate model behaviour in areas of
+parameter space we have not shown.
+
+\paragraph{Consistency with previous
+work}\label{consistency-with-previous-work}
+
+The predicted responses of resistance presented here can be put into the
+context of previous work and recommendations. Earlier modelling work has
+made sometimes contrasting predictions and these can be explained by the
+details of their structure and input choices. One model predicted that
+``the use of mixtures is always more effective in delaying the onset of
+resistance often by many orders of magnitude''{[}11{]}. That model
+assumed that SS genotypes were always killed, RR always survived, and
+that a constant 10\% of mosquitoes escaped the insecticide. This would
+be represented in our model by a very high `insecticide effectiveness'
+of 1, a `resistance restoration' of 1 and an `exposure' of 0.9. This
+result is consistent with our prediction that the very high
+effectiveness is likely to favour mixtures over sequences (e.g.~see Fig
+6D and Table 4).
+
+A subsequent model {[}22{]} lead to the following more pessimistic
+conclusion about the potential of mixtures : ``As a result of incomplete
+coverage and residue decay, the mortality of susceptible homozygotes is
+rarely consistently high enough for pesticide mixtures to be effective''
+{[}22{]}. This ``mortality of susceptible homozygotes'' is equivalent to
+our insecticide effectiveness. Again this points to the importance of
+effectiveness and the question of whether the effectiveness levels
+needed to make mixtures better are attainable operationally. Our result
+showing that decreasing exposure favours mixtures over sequences is also
+consistent with figure 3 from that paper {[}22{]}.
+
+One consistent aspect of the results presented here can seem to
+contradict published advice and has been questioned when we have
+presented model results. Our model predicts that lower insecticide
+effectiveness or exposure will lead to slower spread of resistance in
+sole use and sequences (Figs 3 \& 5, Table 2). This is consistent with
+expectations that lower effectiveness and exposure would lead to lower
+mortality of susceptibles and thus lower selection pressure for
+resistance {[}22{]}. This can seem to contradict recommendations that
+poor implementation of insecticide interventions can promote the
+development of resistance. e.g.~from {[}5{]} ``Certain pest control
+practices have consistently been shown to exacerbate the loss of
+susceptible pest populations and the development of resistance. These
+include: \ldots{} - the use of application rates that are below or above
+those recommended on the label, - poor coverage of the area being
+treated \ldots{}''.
+
+Poor implementation would likely cause lower effectiveness of and
+exposure to insecticides, so how can this increase the potential for
+resistance when our model suggests it would be decreased ? The
+explanation for this apparent difference is in the effect of dominance.
+Poor application can reduce the mortality of heterozygotes (RS) thus
+effectively increasing the dominance of the resistant allele and
+increasing its rate of evolution {[}23{]}{[}5{]}. Our model results are
+consistent with this, showing that increasing dominance of restoration
+leads to faster spread of resistance (Fig 3C, 7). So, the effect of poor
+implementation of an insecticide intervention on resistance could be
+either positive or negative depending on the relative effects on
+dominance of the resistance gene, insecticide effectiveness and
+exposure. Similarly, the effects of poor application on the relative
+benefits of mixtures and sequences will depend on these trade-offs
+because of our result that effectivenes favours mixtures, exposure
+favours sequences and higher dominance favours neither (Table 4).
+
+\paragraph{Implications}\label{implications}
+
+Experience from agriculture where single interventions have been
+associated with a rapid development of resistance have led for calls for
+a more Integrated Vector Management composed of a series of partially
+effective tools {[}24{]}. It is suggested that such an integrated
+approach is more sustainable and `evolution-proof' {[}24{]}. A similar
+combining of interventions has been advocated to cover mosquitoes
+exhibiting different behaviours {[}25{]}. Mosquitoes with different
+behaviours e.g.~feeding outdoors and/or on animals are likely to result
+in lower exposures to insecticides in nets or sprayed on walls. This
+lower exposure could favour mixtures over sequences as we have
+indicated. In both cases our modelling approach could be used to
+investigate implications for the development of resistance.
+
+\emph{TODO bits to incorporate into discussion :}
+
+\emph{todo more reference to {[}12{]} who made many of these points 25
+years ago}
+
+\emph{Insecticides with some repellency will reduce exposure to the
+insecticide and thus, as shown here, are likely to slow the evolution of
+resistance and favour mictures over sequences. This would need to be
+balanced with a likely loss in protection to the population as a whole
+provided by the insecticide {[}15{]}.}
+
+\emph{This model can explore when a mixture strategy may be favoured
+over a sequence purely for the evolution of resistance. Other factors
+also need to be considered. On the positive side a mixture is likely to
+kill more vectors and may also limit their ability to transmit the
+disease. On the negative side a mixture is more expensive.}
+
+\emph{todo, somewhere earlier say good thing about effectiveness \&
+exposure is that they can be modified}
+
+\paragraph{Conclusions}\label{conclusions-1}
+
+We have developed flexible, modern model that can replicate previous
+modelling work and outputs that help us understand the mechanisms by
+which different insecticide application strategies may be favoured. For
+assessing the relative benefits of mixtures and sequences we have
+demonstrated the likely importance of insecticide effectiveness and
+exposure and costs of resistance. High insecticide effectiveness favours
+mixtures through the protective effect of one insecticide on the other.
+High exposure favours sequences because of the higher selection
+pressures on both resistances together in a mixture. High costs of
+resistance can favour mixtures because the selection pressures on both
+resistances are reduced. Similar processes will occur in other
+strategies such as rotations and mosaics and we are starting to
+investigate them. Predictions from the model can be tested against data
+from trials of new insectides to help advance guidelines on good
+practice.
+
+\subsubsection{Declarations}\label{declarations}
+
+TODO see
+\url{https://malariajournal.biomedcentral.com/submission-guidelines/preparing-your-manuscript/research-article}
+
+\subparagraph{Ethics approval and consent to
+participate}\label{ethics-approval-and-consent-to-participate}
+
 Not applicable
 
-##### Consent for publication
+\subparagraph{Consent for publication}\label{consent-for-publication}
+
 Not applicable
 
-##### Availability of data and material
-
-##### Competing interests
-
-##### Funding
-
-##### Authors' contributions
-AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
-IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
-
-##### Acknowledgements
-Thank you to Bethany Levick who developed the first version of the model code.
+\subparagraph{Availability of data and
+material}\label{availability-of-data-and-material}
+
+\subparagraph{Competing interests}\label{competing-interests}
+
+\subparagraph{Funding}\label{funding}
+
+\subparagraph{Authors' contributions}\label{authors-contributions}
+
+AS conceived the paper, re-factored the model code, performed the
+analyses and wrote the first and subsequent versions of the paper. IH
+conceived the modelling approach, developed the first version of the
+model in collaboration with Bethany Levick, explained concepts to AS and
+reviewed and commented on all manuscript versions.
+
+\subparagraph{Acknowledgements}\label{acknowledgements}
+
+Thank you to Bethany Levick who developed the first version of the model
+code.
 
@@
 
-**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
-2. Exposure | faster | reduced fitness of SS overall
-3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
-4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
-
-**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | ---------- | ---------------------------------------------
-1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure from the other
-2.
-Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
-3. Dominance of restoration | faster | increased fitness of heterozygotes
-4. Resistance restoration | faster | increased fitness of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
+\textbf{Table 2. Effect of inputs on resistance when insecticides used
+singly or in sequence}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.28\columnwidth}\raggedright\strut
+Parameter to increase\strut
+\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+effect on resistance evolution\strut
+\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright\strut
+Mechanism\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+reduced fitness of SS in presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+reduced fitness of SS overall\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+3. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+increased fitness of SR in presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+4. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+increased fitness of RR in presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+less change needed to reach resistance threshold\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+slower\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+reduced fitness of RR in absence of insecticide\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Table 3. Effect of inputs on resistance when insecticides used
+in a mixture}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
+Parameter to increase\strut
+\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
+effect on resistance evolution\strut
+\end{minipage} & \begin{minipage}[b]{0.52\columnwidth}\raggedright\strut
+Mechanism\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\textbf{slower}\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+one insecticide reduces the fitness of individuals resistant to the
+other thereby reducing selection pressure from the other\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster (but less than for single)\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+reduced fitness of individuals susceptible to one insecticide increases
+selection pressure for that resistance. However at the same time
+selection pressure is reduced by reduced fitness of resistant
+individuals caused by the other insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+3. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+increased fitness of heterozygotes\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+4. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+increased fitness of resistants\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+less change needed to reach resistance threshold\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+slower\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+reduced fitness of resistants in absence of insecticide\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-**Table 4. Effect of inputs on the difference between mixture and sequential use**
-
-Parameter to increase  | increase favours mix or sequence | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
-2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4.
-Resistance restoration | neither | As for dominance of restoration.
-Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
-5.
-Frequency  | neither | As for dominance and resistance restoration.
-Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
-6.
-Cost of resistance | mixture | Higher costs slow the evolution of resistance more in a mixture than when used in sequence (but with higher costs in a sequence there is a greater chance for resistance levels to decline for the insecticide not being used).
+\textbf{Table 4. Effect of inputs on the difference between mixture and
+sequential use}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.28\columnwidth}\raggedright\strut
+Parameter to increase\strut
+\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+increase favours mix or sequence\strut
+\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright\strut
+Mechanism\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+mixture\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher effectiveness gives faster resistance for sequence and slower
+resistance in mixture\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+sequence\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher exposure gives faster resistance for sequence and mixture but the
+greater effect on mixture favours sequence.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+3. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+neither\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher dominance gives faster resistance in both sequences and mixtures
+such that the difference between them is not changed.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+4. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+neither\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+As for dominance of restoration. Higher resistance restoration gives
+faster resistance in both sequences and mixtures such that the
+difference between them is not changed.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+neither\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+As for dominance and resistance restoration. Higher starting frequencies
+give faster resistance in both sequences and mixtures such that the
+difference between them is not changed.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+mixture\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher costs slow the evolution of resistance more in a mixture than
+when used in sequence (but with higher costs in a sequence there is a
+greater chance for resistance levels to decline for the insecticide not
+being used).\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
 
@@
 
-## Figures
-
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="The effect of model inputs on the fitness of genotypes for a single insecticide.
+\subsection{Figures}\label{figures}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig1-1} \hfill{}
+
+\caption{The effect of model inputs on the fitness of genotypes for a single insecticide.
 Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
@@
 For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3."}
-
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
-
-```
-
-![Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
-The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
-Two scenarios (A and B) can be run and the results viewed side by side.
-This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)
-
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
+In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3.}\label{fig:Fig1}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{paper2_UI1.png}
+\caption{Screenshot of one online model user interface, accessible at :
+\url{https://andysouth.shinyapps.io/resistmob2/}. The user can modify
+values of the input parameters considered in this paper using simple
+sliders and run the model to get graphs of resulting resistance
+frequency over time. Two scenarios (A and B) can be run and the results
+viewed side by side. This makes it easy to explore the effect of
+changing individual inputs.}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig3-1} \hfill{}
+
+\caption{Single insecticide use and the effect of inputs on changing resistance frequency over time.
 A.
@@
 For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
-"}
-
-  library(cowplot)
-
-  #list to put plots in for cowplot
-  plotlist <- list(4)
-
-  #A effectiveness  
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.8,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
-         theme_bw() +  
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +  
-         labs(colour = "effectiveness") +
-         coord_trans(y = "log10") +    
-         geom_line()
-
-  #B exposure
-  exposure <- 0.2
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.5
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.8
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "exposure", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-  #C single insecticide dominance
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.8, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #C 
-  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance of restoration") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #D single insecticide resistance restoration
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.2,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.8,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #D  
-  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "resistance restoration", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
+}\label{fig:Fig3}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig4-1} \hfill{}
+
+\caption{Single insecticide use and the effect of inputs on changing resistance frequency over time.
 A.
@@
 Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds.
-Increasing costs of resistance increases time-to-resistance."}
-
-  #single insecticide frequency fig
-
-  #A starting frequency
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.001,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.1,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <-  ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "start\nfrequency") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #B changing costs at dominance 0.5
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0,
-                      h.RS1_00 = 0.5)
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.1,
-                      h.RS1_00 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.2,
-                      h.RS1_00 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "cost of\nresistance") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]], ncol=2, rel_widths=c(1,1), labels='AUTO')  
-  #cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
+Increasing costs of resistance increases time-to-resistance.}\label{fig:Fig4}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig5-1} \hfill{}
+
+\caption{Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
 Exposure to the insecticides is increased from row 1 (A and B) to row 2 (C and D).
@@
 Increasing exposure decreases times-to-resistance for mixtures and reduces their performance relative to sequences.
-"}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
+}\label{fig:Fig5}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig6-1} \hfill{}
+
+\caption{Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
 Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
@@
 Effectiveness of insecticide2 0.8 and of insecticide1 1.
-With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios."}
-
-#4 panel figure looking at effectiveness changes alone
-
-#test of rearranging figures on a grid
-#eff1  0.8  1
-#eff2
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.5
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
+With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios.}\label{fig:Fig6}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig7-1} \hfill{}
+
+\caption{Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
 Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
@@
 Resistance restoration and dominance for insecticide1 increased to 0.8.
-Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios."}
-
-#4 panel figure looking at dominance and resistance restoration
-
-#rr1  0.5  0.8
-#dom1
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#testing
-# plot(1)
-# plot(1)
-# plot(1)
-# plot(1)
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
+Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios.}\label{fig:Fig7}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig8-1} \hfill{}
+
+\caption{Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
 Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
@@
 Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.01 to 0.001.
-In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness."}
-
-#4 panel figure looking at effectiveness and starting freq.
-
-#eff1  0.5  0.8
-#strt freq1
-#0.01    1   2
-#0.001    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig 9 cost of 1 resistance, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of cost of resistance on time-to-resistance for mixtures and sequences.
+In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness.}\label{fig:Fig8}
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{paper2_resistance_mechanisms_mixtures_files/figure-latex/Fig 9 cost of 1 resistance-1} \hfill{}
+
+\caption{Influence of cost of resistance on time-to-resistance for mixtures and sequences.
 Row 1 (A and B) same as Fig.
@@
 Costs seem to favour mixtures in these plots.
-Also be aware that costs would lead to a decline in the frequency of resistance for the first insecticide in a sequence when it stops being used."}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Figxx not eval, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
-All control inputs equal at 0.5, male exposure 0.8, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Male exposure decreased from 0.8 to 0.2, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
-
-# not evaluated at the moment
-#4 panel figure looking at male exposure prop and starting freq.
-
-#eff1  0.5  0.8
-#maleprop
-#0.8    1   2
-#0.2    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure =0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
+Also be aware that costs would lead to a decline in the frequency of resistance for the first insecticide in a sequence when it stops being used.}\label{fig:Fig 9 cost of 1 resistance}
+\end{figure}
 
@@
 
-## References
-<!-- now auto created here-->
+\subsection{References}\label{references}
+
+\hypertarget{refs}{}
+\hypertarget{ref-WHO2012}{}
+1. WHO: \emph{Global plan for insecticide resistance management in
+malaria vectors (GPIRM).} Geneva. 2012:130.
+
+\hypertarget{ref-Ranson2016}{}
+2. Ranson H, Lissenden N: \textbf{Insecticide Resistance in African
+Anopheles Mosquitoes: A Worsening Situation that Needs Urgent Action to
+Maintain Malaria Control}. \emph{Trends in Parasitology} 2016,
+\textbf{32}:187--196.
+
+\hypertarget{ref-Hemingway2016}{}
+3. Hemingway J, Ranson H, Magill A, Kolaczinski J, Fornadel C, Gimnig J,
+Coetzee M, Simard F, Roch DK, Hinzoumbe CK, Pickett J, Schellenberg D,
+Gething P, Hoppé M, Hamon N: \textbf{Averting a malaria disaster: Will
+insecticide resistance derail malaria control?} \emph{The Lancet} 2016,
+\textbf{387}:1785--1788.
+
+\hypertarget{ref-IRAC2011}{}
+4. IRAC: \emph{Prevention and Management of Insecticide Resistance in
+Vectors of Public Health Importance}. 2011:71.
+
+\hypertarget{ref-FAO2012}{}
+5. FAO: \emph{Guidelines on Prevention and Management of Pesticide
+Resistance}. Rome; 2012:55.
+
+\hypertarget{ref-Bhatt2015}{}
+6. Bhatt S, Weiss DJ, Cameron E, Bisanzio D, Mappin B, Dalrymple U,
+Battle KE, Moyes CL, Henry A, Eckhoff PA, Wenger EA, Briët O, Penny MA,
+Smith TA, Bennett A, Yukich J, Eisele TP, Griffin JT, Fergus CA, Lynch
+M, Lindgren F, Cohen JM, Murray CLJ, Smith DL, Hay SI, Cibulskis RE,
+Gething PW: \textbf{The effect of malaria control on Plasmodium
+falciparum in Africa between 2000 and 2015.} \emph{Nature} 2015,
+\textbf{526}:207--11.
+
+\hypertarget{ref-Churcher2016}{}
+7. Churcher TS, Lissenden N, Griffin JT, Worrall E, Ranson H:
+\textbf{The impact of pyrethroid resistance on the efficacy and
+effectiveness of bednets for malaria control in Africa}. \emph{eLife}
+2016, \textbf{5}:e16090.
+
+\hypertarget{ref-Hemingway2006}{}
+8. Hemingway J, Beaty BJ, Rowland M, Scott TW, Sharp BL: \textbf{The
+Innovative Vector Control Consortium: improved control of mosquito-borne
+diseases}. \emph{Trends in Parasitology} 2006, \textbf{22}:308--312.
+
+\hypertarget{ref-IVCC2016}{}
+9. IVCC: \textbf{Annual Report 2015-16}. 2016.
+
+\hypertarget{ref-Curtis1985}{}
+10. Curtis CF: \textbf{Theoretical models of the use of insecticide
+mixtures for the management of resistance}. \emph{Bulletin of
+Entomological Research} 1985, \textbf{75}:259.
+
+\hypertarget{ref-Mani1985}{}
+11. Mani GS: \textbf{Evolution of resistance in the presence of two
+insecticides.} \emph{Genetics} 1985, \textbf{109}:761--783.
+
+\hypertarget{ref-Roush1989}{}
+12. Roush RT: \textbf{Designing resistance management programs: How can
+you choose?} \emph{Pesticide Science} 1989, \textbf{26}:423--441.
+
+\hypertarget{ref-Levick2017}{}
+13. Levick B, South A, Hastings IM: \textbf{A Two-locus Model of The
+Evolution of Insecticide Resistance to Inform and Optimise Public Health
+Insecticide Deployment Strategies.} \emph{PLOS Computational Biology}
+2017.
+
+\hypertarget{ref-Liu2015}{}
+14. Liu N: \textbf{Insecticide Resistance in Mosquitoes: Impact,
+Mechanisms, and Research Directions}. \emph{Annual Review of Entomology}
+2015, \textbf{60}:537--559.
+
+\hypertarget{ref-Birget2015}{}
+15. Birget PLG, Koella JC: \textbf{A genetic model of the effects of
+insecticide-treated bed nets on the evolution of
+insecticide-resistance}. \emph{Evolution, medicine, and public health}
+2015, \textbf{2015}:205--215.
+
+\hypertarget{ref-RCoreTeam2016}{}
+16. R Core Team: \textbf{R: A language and environment for statistical
+computing.} 2016.
+
+\hypertarget{ref-South2017}{}
+17. South A: \textbf{https://github.com/AndySouth/resistance}.
+
+\hypertarget{ref-South2017a}{}
+18. South A: \textbf{Online model user interface :
+https://andysouth.shinyapps.io/resistmob2/}. 2017.
+
+\hypertarget{ref-Rivero2010}{}
+19. Rivero A, Vézilier J, Weill M, Read AF, Gandon S:
+\textbf{Insecticide control of vector-borne diseases: When is
+insecticide resistance a problem?} \emph{PLoS Pathogens} 2010.
+
+\hypertarget{ref-Barbosa2017}{}
+20. Barbosa S, Kay K, Chitnis N, Hastings IM, Tropical S:
+\textbf{Modelling the public health impact of insecticide deployment on
+malaria-transmitting mosquito populations and its role in driving
+insecticide resistance .} 2017:1--58.
+
+\hypertarget{ref-Taillebois2016}{}
+21. Taillebois E, Thany SH: \textbf{The differential effect of low-dose
+mixtures of four pesticides on the pea aphid Acyrthosiphon pisum}.
+\emph{Insects} 2016, \textbf{7}:1--7.
+
+\hypertarget{ref-Roush1998}{}
+22. Roush RT: \textbf{Two-toxin strategies for management of
+insecticidal transgenic crops: can pyramiding succeed where pesticide
+mixtures have not?} \emph{Philosophical Transactions of the Royal
+Society B: Biological Sciences} 1998, \textbf{353}:1777--1786.
+
+\hypertarget{ref-Barbosa2011}{}
+23. Barbosa S, Black IV WC, Hastings I: \textbf{Challenges in estimating
+insecticide selection pressures from mosquito field data}. \emph{PLoS
+Neglected Tropical Diseases} 2011, \textbf{5}.
+
+\hypertarget{ref-Thomas2012}{}
+24. Thomas MB, Godfray HCJ, Read AF, Berg H van den, Tabashnik BE,
+Lenteren JC van, Waage JK, Takken W: \textbf{Lessons from agriculture
+for the sustainable management of Malaria vectors}. \emph{PLoS Medicine}
+2012, \textbf{9}:7--10.
+
+\hypertarget{ref-Killeen2017}{}
+25. Killeen GF, Marshall JM, Kiware SS, Andy S, Chaki PP, Govella NJ:
+\textbf{Measuring, manipulating and exploiting behaviours of adult
+mosquitoes to optimize malaria vector control impact}. \emph{BMJ Global
+Health} 2017.
+
+\end{document}

### 2017-04-21 inst/documents/pub_health_j/pub_health_j_figs.Rmd
figs for pub health jour article

--- 2017-03-24 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
 ---
-title: "paper2_resistance_mechanisms_mixtures"
-output: pdf_document
+title: "Figs for public health journal article April 2017"
+output: 
+  pdf_document:
+
 #output: word_document
@@
 date: "`r Sys.Date()`"
-bibliography: library.bib
-csl: biomed-central.csl
-#header-includes:
-#    - \usepackage{setspace}
-#    - \doublespacing
-#    - \usepackage{lineno}
-#    - \linenumbers
 ---
@@
 ```{r, eval=FALSE, include=FALSE}
-#!!!! paper protocol
-#!!!! might be easier to put text and figs into separate Rmd files after draft stage
-#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
-#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
-#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
-#!!!! submission figs, set this chunk to eval=TRUE
-
 # options to create final publication quality figures
-# causes the document to fail, but the figs do get stored in pap2figs/
+# causes the document to fail, but the figs do get stored in figs/
 library(knitr)
@@
                cache=FALSE,
-               fig.path='pap2figs/')
+               fig.path='figs/')
 ```
@@
 ```
-
-### version 11
-
-#Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary.
-
-###alternative titles:
-###Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
-###An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
-###A tutorial, primer ...
-###Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
-
-Target : Malaria journal
-
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and suggestions welcome
-
-## Abstract
-
-_must be 350 words, currently 351_
-
-##### Background
-Insecticide resistance threatens the control of vectors and particularly malaria.
-New insecticides are being developed to address this.
-Potential strategies for using new insecticides include mixtures, sequences and rotations and there are recommendations for how best to implement these to manage resistance.
-There is, however, limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
-Previous work concludes that preferred strategies will be situation specific.
-There is potential to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
-
-This paper develops an accessible, mechanistic picture of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-The aim is to reach an audience unlikely to read a more detailed modelling paper.
-We use our model to develop a mechanistic understanding of how insecticide resistance is expected to increase with the use of insecticides in isolation, sequence and mixtures.
-The model flexibly represents two independent genes coding for resistance to two insecticides.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, how much resistance counteracts this, the proportion of mosquitoes that are exposed to insecticides and costs of resistance.
-
-##### Results
-We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about the selection pressure exerted by insecticides.
-We show that an insecticide used in a mixture, relative to alone, always prompts slower evolution of resistance, but resistance to the two insecticides may evolve more slowly when used in sequence.
-We show that the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on whether resistance to two insecticides is likely to arise faster in a mixture or sequence.
-
-##### Conclusions
-Our model makes more open the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
-We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/MixSeqResist1).
-These tools can contribute to operational decisions in insecticide resistance management.
-
-## Keywords
-insecticide resistance; public health; mosquitoes; vector-borne diseases; infectious diseases; malaria; population genetics
-
-## Background
-
-Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
-Malaria alone still results in hundreds of thousands of deaths per year.
-Recent malaria control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
-Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million falciparum malaria cases between 2000 and 2015 [@Bhatt2015].
-A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
-
-The WHO produced a Global Plan for Insecticide Resistance Management in malaria vectors (GPIRM)[@WHO2012] which includes recommendations on operational strategies for managing resistance including the use of insecticide mixtures when they become available.
-Efforts are under way to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
-The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
-Three new insecticides are now in development [@IVCC2016] and likely to be available around 2020 [@Ranson2016].
-It is important that decisions about how best to use the new insecticides to delay the onset of resistance are made before insecticides are released [@Hemingway2016].
-
-Modelling studies have investigated the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
-[@Curtis1985][@Mani1985][@Roush1989] but much of the work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
-In an earlier paper [@Levick2017] we described the technical details of a flexible model used to investigate the relative benefits of mixtures and sequences.
-Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance is influenced by different inputs.
-This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
-
-Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
-It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
-
-## Methods
-
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
-One gene or locus is represented for each insecticide.
-Each individual has two copies of the gene and there are two potential alleles per locus.
-Each allele confers either resistance or susceptibility to the insecticide.
-Thus individuals can either have both resistance genes (RR homozygous), both susceptible genes (SS) or one of each (SR, heterozygous).
-The combination of the two genes is termed the genotype.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
-The fitness of the SS genotype in absence of the insecticide is set at a maximum value of 1.
-All other fitnesses are less, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
-
-**Table 1. Parameters inflencing the development of insecticide resistance**
-
-Parameter  | Description
----------------------- | ----------------------------------------------------
-1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
-2. Exposure | proportion of insects exposed to insecticide
-3. Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide
-4. Dominance of restoration | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
-5. Frequency  | frequency of resistance alleles within the population
-6. Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide
-7. Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide
-
-The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 1.
-In the model itself this is repeated for both insecticides and the fitnesses multiplied.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS and resistance restoration 'restores' a portion of the fitness for RR.
-A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
-We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
-Cost of resistance determines the fitness of RR in the portion of the population not exposed to the insecticide.
-
-Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
-For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
-Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Figure 1).
-
-The simulation proceeds through generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
-This acts to make the fitter alleles more common over time.
-Separate sexes and standard sexual reproduction with recombination is included.
-There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that here.
-
-So far this description just considers a single insecticide and associated resistance allele.
-In the model itself a second insecticide and resistance allele are included and overall fitness is calculated by multiplying the two results.
-This allows the model to represent populations exposed to two insecticides together in a mixture.
-
-The model outputs the change in resistance allele frequencies over time measured in generations.
-We extract the number of generations at which a resistance allele frequency of 50% is reached and term this 'time-to-resistance'.
-Using a 50% threshold is consistent with previous modelling analyses e.g.
-[@Curtis1985][@Birget2015].
-Preliminary investigations showed that using other potential thresholds of 25 or 75% gave very similar results.
-
-We investigate three insecticide use strategies.
-
-1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
-1. two insecticides used in a mixture (concurrently)
-
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an on-line user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
-
-## Results
-
-### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
-This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
-Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that less change is required to reach the threshold.
-
-Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect.
-Higher values led to a slower development of resistance.
-Again this is intuitive because costs reduce the advantage of resistance and decrease selection.
-These effects of inputs on single insecticide use are summarised in Table 2.
-
-### Two insecticides
-When two insecticides are used we can compare the relative performance of a mixture with sequential use.
-Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
-To help understand the mechanisms influencing this relative performance we start with a single base scenario and demonstrate changing inputs individually.
-In this base scenario we set effectiveness, exposure, resistance restoration, and dominance at 0.5, the starting frequencies of resistance to 0.01 and cost of resistance to 0.
-Resistance arises slower for sequential than mixture for this base scenario (Fig 5A).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 5A).
-Thus the time-to-resistance for the mixture divided by the sequence is less than 1 at 0.8.
-
-#### Insecticide exposure and effectiveness
-
-Increasing exposure (the proportion of insects that come into contact with the insecticides) (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy.
-The effect on the mixture is greater so time-to-resistance remains longer for the sequence (the ratio decreasing from 0.8 to 0.6).
-In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 5B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
-The longer time-to-resistance for the mixture results from a slower increase in resistance for the less effective insecticide.
-Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
-This points a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
-
-The results is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
-This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-
-In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 5, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 5.
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
-
-Increasing the effectiveness further of either or both insecticides increases the performance of mixtures over sequences in terms of time-to-resistance.
-Figure 6A uses the same inputs as Fig 5B.
-From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
-This again points to how more effective insecticides can slow resistance to another when used in a mixture.
-
-#### Resistance restoration and dominance
-Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
-The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
-So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
-
-#### Starting frequencies of resistance
-Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
-Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
-
-#### Costs of resistance
-Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
-The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 9C).
-The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C.
-With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 9A and 1 in Fig 9C.
-At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
-
-A summary of the effects of inputs on the evolution of resistance for mixtures is given in Table 3 and for the difference between mixtures and sequences in Table 4.
-The mechanisms identified in the tables are covered further in the discussion.
-
-#### Linkage disequilibrium
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
-
-## Discussion
-
-Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
-This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
-
-When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resistance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
-
-These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Referring to figure 1 can help explain these.
-Exposure and effectiveness act by decreasing the fitness of all genotypes, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
-Cost acts by decreasing the fitness of RR and SR for those not exposed.
-
-#### Two insecticides
-Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide and, where it differs can again help us to understand the dynamics of resistance evolution.
-Our base scenario (Fig 5A) shows curves for resistance to both insecticides in the mixture that are shallower than those for sequential use.
-The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Thus each insecticide reduces the selection pressure for the increase in resistance to the other and could be said to 'protect' the other.
-However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
-Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
-Incraesing the effectiveness of on insecticide to Fig 5B decreases the added times for the sequence to 90 and increases those for the mixture to 115 thus changing the order and favouring the mixture.
-
-The mechanism for this change can be sen by comparing the shape of the curves in Figs 5A & B.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide, loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide.
-The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-
-#### Difference between exposure and effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us.
-
-#### Costs of resistance
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-It seems that the protective effect of the other insecticide in a mixture interacts with the cost of resistance to reduce selection pressure more than in a sequence.
-
-The effects of cost of resistance become more as the frequency of resistance increases.
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
-Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
-Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
-Also note that if the costs are set even higher than we set them here (e.g.
-above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-Unfortunately, we do not expect such high costs to occur in operational settings.
-
-Relatedly, when there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
-Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where it becomes operationally useful again.
-As such a repeated sequential strategy could offer benefits over mixtures not considered here.
-To incorporat this one would need to use a more detailed metric than time-to-resistance.
-
-#### Caveats
-Our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as effecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
-Insecticide resistance could reduce vector transmission of disease (e.g.
-by reduced lifespan of the vector) or increase transmission (e.g.
-by reduced immunity of the vector to the disease) [@Rivero2010].
-Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Of course, using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
-Others are combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
-
-In our model we represented insecticide interaction as being additive within the mixture.
-This is consistent with recent work on aphids [@Taillebois2016], although that work did also indicate some synergistic effects.
-Our model has the potential to incorporate either synergistic or antagonistic effects too.
-
-The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ranson2016].
-The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-Thus there is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown.
-
-#### Consistency with previous work 
-The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
-
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
-Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
-
-One consistent aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
-
-Poor implementation would likely cause lower effectiveness of and exposure to insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Barbosa2011][@FAO2012].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that effectivenes favours mixtures, exposure favours sequences and higher dominance favours neither (Table 4).
-
-#### Implications
-_todo get more of [@Roush1989] in here_
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the development of resistance.
-
-_low exposure could be achieved by having active ingredients with some repellency._
-
-_todo get Birget and Koella in & link to exposure_
-'whereas repellency is beneficial to delay the evolution of resistance, other models have shown that it decreases the population-level protection of the insecticide.'
-
-_todo, where to put this, earlier ?_
-This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture is more expensive.
-
-_todo, somewhere earlier say good thing about effectiveness & exposure is that they can be modified_
-
-#### Conclusions
-
-We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured.
-For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance.
-High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other.
-High exposure favours sequences through reducing the otherwise high selection pressures on both resistances together.
-High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced.
-Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them.
-Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.
-
-### Declarations
-TODO see https://malariajournal.biomedcentral.com/submission-guidelines/preparing-your-manuscript/research-article
-
-##### Ethics approval and consent to participate
-Not applicable
-##### Consent for publication
-Not applicable
-##### Availability of data and material
-
-##### Competing interests
-
-##### Funding
-
-##### Authors' contributions
-AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
-IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
-
-##### Acknowledgements
-Thank you to Bethany Levick who developed the first version of the model code.
-
-\pagebreak
-
-**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
-2. Exposure | faster | reduced fitness of SS overall
-3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
-4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
-
-**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | ---------- | ---------------------------------------------
-1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure from the other
-2.
-Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
-3. Dominance of restoration | faster | increased fitness of heterozygotes
-4. Resistance restoration | faster | increased fitness of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
-
-\pagebreak
-
-**Table 4. Effect of inputs on the difference between mixture and sequential use**
-
-Parameter to increase  | increase favours mix or sequence | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
-2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4.
-Resistance restoration | neither | As for dominance of restoration.
-Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
-5.
-Frequency  | neither | As for dominance and resistance restoration.
-Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
-6.
-Cost of resistance | mixture | Higher costs slow the evolution of resistance more in a mixture than when used in sequence (but with higher costs in a sequence there is a greater chance for resistance levels to decline for the insecticide not being used).
-
-\pagebreak
-
-## Figures
 
@@
 
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
+plot_fit_calc( simple=TRUE, effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
 
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
-A.
-Effectiveness, B.
-Exposure, C.
-Dominance of restoration, D.
-Resistance restoration.
-Increasing any of the inputs (going from red to blue) results in shorter times-to-resistance.
-For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
-"}
-
-  library(cowplot)
-
-  #list to put plots in for cowplot
-  plotlist <- list(4)
-
-  #A effectiveness  
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.8,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
-         theme_bw() +  
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +  
-         labs(colour = "effectiveness") +
-         coord_trans(y = "log10") +    
-         geom_line()
-
-  #B exposure
-  exposure <- 0.2
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.5
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.8
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "exposure", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-  #C single insecticide dominance
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.8, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #C 
-  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance of restoration") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #D single insecticide resistance restoration
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.2,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.8,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #D  
-  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "resistance restoration", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
-A.
-Starting frequency, B.
-Cost of resistance.
-Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds.
-Increasing costs of resistance increases time-to-resistance."}
-
-  #single insecticide frequency fig
-
-  #A starting frequency
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.001,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.1,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <-  ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "start\nfrequency") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #B changing costs at dominance 0.5
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0,
-                      h.RS1_00 = 0.5)
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.1,
-                      h.RS1_00 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.2,
-                      h.RS1_00 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "cost of\nresistance") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]], ncol=2, rel_widths=c(1,1), labels='AUTO')  
-  #cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
+#TODO decide which of later figures to modify for pubhealth J paper
 

### 2017-04-24 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 
+_comments from Hilary_
+_This would seem to be a critical point to stress - with examples of how this might be acheived._
+
+_Limitations that haven't been addressed is the possiblility that resistance is poylgenic, and the possibility of cross resistance (e.g a P450 may cause partial resistance to both)._
+
+_Something further to guide readers on how they might determine dominance would also be useful given how crucial this is._
+
 #### Conclusions
@@
 layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+#layout.show(n=9)
 

### 2017-04-24 inst/documents/pub_health_j/pub_health_j_figs.Rmd
paper writing

@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
-Exposure to the insecticides is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of one insecticide is increased from column 1 (A and C) to column2 (B and D).
-On the upper X axis 's' and 'm' indicate where the 50 percent resistance threshold is reached for the sequence and the mixture.
-In the lower right of each panel the the ratio of time-to-resistance for mixture/sequence is shown rounded to 1 decimal place to give an indication of the relative performance of mixtures and sequences.
-A.
-All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C.
-Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D.
-Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence.
-Increasing effectiveness increases times-to-resistance for mixtures and improves their performance relative to sequences.
-Increasing exposure decreases times-to-resistance for mixtures and reduces their performance relative to sequences.
-"}
+```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Comparing time-to-resistance for mixtures and sequences  A.
+Resistance arises slower for the mixture, B.
+Resistance arises slower for the sequence.
+All inputs set to 0.5 except for starting frequency of resistance set to 0.01, and effectiveness of insecticide 1 set to 0.4 in A and 0.8 in B."}
 
-#4 panel figure looking at effectiveness and exposure
+#2 panel fig A. sequence better, B. mix better
 
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
+#eff1  0.4  0.8
 
 # add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+#layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
 
-#effectiveness on mixture
+layout( matrix(c(1:2,3,3),2,2, byrow = TRUE), heights=c(1,0.15), widths=c(1,1) )
+#layout.show(n=3)
 
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
+#eff1 0.4 seq better
 
-#base scenario
+#plot margins, hight t just a bodge to get graphs down closer to legend
+par(mar=c(1.5, 4, 15, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.4 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = 1 )
 
@@
 
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+#plot margins, hight t just a bodge to get graphs down closer to legend
+par(mar=c(1.5, 4, 15, 0)) #b,l,t,r default c(5, 4, 4, 2)
 
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
+#eff1 0.8 mix better
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = 1 )
 
 mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
 
@@
 
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
+legend( 'bottom', inset=0.1, legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   "), col=c("red","blue","red","blue"), lty=c(2,2,1,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.9, horiz =TRUE )
 
 ```
-
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
-Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-Effectiveness of insecticide1 increased from 0.5 to 0.8, B.
-Effectiveness of insecticide1 increased from 0.5 to 1, C.
-Effectiveness of insecticides 1 and 2 increased to 0.8, D.
-Effectiveness of insecticide2 0.8 and of insecticide1 1.
-With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios."}
-
-#4 panel figure looking at effectiveness changes alone
-
-#test of rearranging figures on a grid
-#eff1  0.8  1
-#eff2
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.5
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
-Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
-Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, B.
-Resistance restoration for insecticide1 increased from 0.5 to 0.8, C.
-Dominance for insecticide1 increased to 0.8, D.
-Resistance restoration and dominance for insecticide1 increased to 0.8.
-Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios."}
-
-#4 panel figure looking at dominance and resistance restoration
-
-#rr1  0.5  0.8
-#dom1
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#testing
-# plot(1)
-# plot(1)
-# plot(1)
-# plot(1)
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
-Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
-Effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.01 to 0.001.
-In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness."}
-
-#4 panel figure looking at effectiveness and starting freq.
-
-#eff1  0.5  0.8
-#strt freq1
-#0.01    1   2
-#0.001    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig 9 cost of 1 resistance, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of cost of resistance on time-to-resistance for mixtures and sequences.
-Row 1 (A and B) same as Fig.
-5 with no cost of resistance, row 2 (C and D) cost of resistance set to 0.15 with dominance of cost set to 0.5.
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-The increase in cost of resistance from A.
-to C.
-removes the advantage of sequence relative to mixture.
-The ratio of time-to-resistance for the mixture divided by that for the sequence goes from 0.
-to 1.
-Costs seem to favour mixtures in these plots.
-Also be aware that costs would lead to a decline in the frequency of resistance for the first insecticide in a sequence when it stops being used."}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Figxx not eval, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
-All control inputs equal at 0.5, male exposure 0.8, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Male exposure decreased from 0.8 to 0.2, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
-
-# not evaluated at the moment
-#4 panel figure looking at male exposure prop and starting freq.
-
-#eff1  0.5  0.8
-#maleprop
-#0.8    1   2
-#0.2    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure =0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-\pagebreak
-
-## References
-<!-- now auto created here-->

### 2017-08-08 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
comments from Hilary
paper2 writing

@@
 In an earlier paper [@Levick2017] we described the technical details of a flexible model used to investigate the relative benefits of mixtures and sequences.
-Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance is influenced by different inputs.
+Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance can be expected to change in different scenarios.
 This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
@@
 Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
-It is likely that genes conferring resistance are present in populations even prior to exposure to novel insecticides thus leading to the potential for selection [@Liu2015].
+We assume both that genes conferring resistance are already present in the population and that resistance to each insecticide is coded by a single gene.
+These assumptions are consistent with previous modelling work.
+In support of the first assumption insecticide resistance has been termed 'pre-adaptive' meaning that resistance alleles are present in very low numbers even prior to exposure to novel insecticides [@Ffrench-Constant2013][@Liu2015] as has been documented in blowflies [Hartley2006].
+In support of using a single gene representation of resistance evidence suggests that whilst polygenic resistance is common, control failures, particularly those due to target site resistance, are mostly due to single major genes [@Ffrench-Constant2013].
 
@@
 As such a repeated sequential strategy could offer benefits over mixtures not considered here.
-To incorporat this one would need to use a more detailed metric than time-to-resistance.
+To incorporate this one would need to use a more detailed metric than time-to-resistance.
 
@@
 
-The development of insecticide resistance in operational field settings is subject to considerable uncertainty [@Ranson2016].
-The model we present here is able to cover the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
+There is much evidence of the involvement of multiple genes in resistance mechanisms.
+Our model, in common with most previous ones, just follows a single gene per insecticide.
+In most cases we believe this is reasonable.
+Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes[].
+Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
+To represent polygenic resistance a different 'quantitative genetic' approach would be required.
+A quantitive genetic model would likely lead to a more rapid initial evolution of resistance, a slower middle phase and a continuing increase.
+If polygenic resistance is shown in future to be more of a problem then these results could be compared to those of quantitative genetic models.
+
+Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
+The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
 holding most inputs constant at intermediate values while modifying single inputs in isolation).
@@
 _Something further to guide readers on how they might determine dominance would also be useful given how crucial this is._
+
+_Can you give some indications of the type of data that should be collected to help paramertise the model?_
 

### 2017-08-11 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 writing

@@
 ---
-title: "Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary"
+title: "Insecticide resistance evolution with mixtures and sequences : a model-based introduction."
 output: 
@@
 ###A tutorial, primer ...
-###Insecticide effectiveness predicted to be most important factor determining whether insecticide mixtures can slow evolution of resistance.
-
-Target : Malaria journal (other suggestions welcome)
+
+Target : Malaria journal or Parasites & Vectors
 
@@
 
-_must be 350 words, currently 351_
+_must be 350 words, currently 349_
 
 ##### Background
-Insecticide resistance threatens the control of vectors and particularly malaria.
-New insecticides are being developed to address this.
-Potential strategies for using new insecticides include mixtures, sequences and rotations and there are recommendations for how best to implement these to manage resistance.
-There is, however, limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
-Previous work concludes that preferred strategies will be situation specific.
-There is potential to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
-
-This paper develops an accessible, mechanistic picture of how the evolution of insecticide resistance is likely to respond to potential intervention strategies to help guide both management and policy.
-The aim is to reach an audience unlikely to read a more detailed modelling paper.
-We use our model to develop a mechanistic understanding of how insecticide resistance is expected to increase with the use of insecticides in isolation, sequence and mixtures.
-The model flexibly represents two independent genes coding for resistance to two insecticides.
-We look principally at the ability of the insecticides to kill susceptible mosquitoes, how much resistance counteracts this, the proportion of mosquitoes that are exposed to insecticides and costs of resistance.
+Insecticide resistance threatens effective vector control, especially for mosquitoes and malaria.
+Insecticide use strategies that have been recommended to manage resistance include mixtures, sequences and rotations.
+New insecticides are being developed to address resistance and there is an opportunity to develop use strategies that limit the evolution of further resistance in the short term.
+There is limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
+Previous work concludes that preferred strategies will be situation specific and there appears to be little recent consensus e.g.
+on the merits of mixtures.
+There is an opportunity to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
+
+We provide a concise, accessible description of the use of a flexible model to develop a mechanistic picture the evolution of insecticide resistance and how it is likely to respond to potential insecticide use strategies.
+The aim is to reach an audience unlikely to read a more detailed modelling paper by assuming little knowledge of models.
+The model itself, as described here, represents two independent genes coding for resistance to two insecticides.
+This allows us to represent the use of insecticides in isolation, sequence and mixtures.
 
 ##### Results
-We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about the selection pressure exerted by insecticides.
-We show that an insecticide used in a mixture, relative to alone, always prompts slower evolution of resistance, but resistance to the two insecticides may evolve more slowly when used in sequence.
-We show that the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on whether resistance to two insecticides is likely to arise faster in a mixture or sequence.
+We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about selection pressure.
+We show that using an insecticide in a mixture, relative to alone, always prompts slower evolution of resistance to that insecticide.
+However, when the resistance to both insecticides is considered, resistance thresholds may be reached later for a sequence relative to a mixture.
+Increasing the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on favouring a mixture over a sequence.
 
 ##### Conclusions
-Our model makes more open the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
+Our model provides an accessible description of the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
 We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/MixSeqResist1).
-These tools can contribute to operational decisions in insecticide resistance management.
+These tools can contribute to an improved discussion about operational decisions in insecticide resistance management.
 
@@
 Efforts are under way to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
-The Innovative Vector Control Consortium (IVCC) was set up in 2005 to develop new vector control tools and particularly new insecticides to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
-Three new insecticides are now in development [@IVCC2016] and likely to be available around 2020 [@Ranson2016].
-It is important that decisions about how best to use the new insecticides to delay the onset of resistance are made before insecticides are released [@Hemingway2016].
+The Innovative Vector Control Consortium (IVCC) was set up in 2005 to work with industry in the development of new vector control tools and particularly new insecticides with new modes of action to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
+Three new insecticides are now in development [@IVCC2016] and likely to be available after 2020 [@Ranson2016].
+It is important that decisions about how best to use the new insecticides, with a clear mindset of delaying the evolution of resistance, are made before insecticides are released [@Hemingway2016].
 
@@
 
+We start by describing the results of model runs with a single insecticide which are more straightforward and move on to describing results of using two insecticides in sequences and mixtures.
+
 ### Single insecticide
@@
 ### Two insecticides
-When two insecticides are used we can compare the relative performance of a mixture with sequential use.
+When two insecticides are used we can compare the relative performance of concurrent use in a mixture with sequential use.
 Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
@@
 
-The results is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
+The result is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
 This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
@@
 This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
-
-When single insecticide use was investigated (Figs 3-4) higher values of inputs that increased the selective advantage of resistance all led to more rapid evolution resistance, and higher values for cost led to slower evolution of (Fig 4B).
-
-These mechanisms, explaining the response of resistance to single insecticide use under different input values, are summarised in Table 2.
-Referring to figure 1 can help explain these.
+We start by explaining the evolution of resistance to a single insecticide, which is relatively straightforward, and move on to explaining the effect of two insecticides in sequences and mixtures.
+
+When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance.
+Higher values for cost led to slower evolution of (Fig 4B).
+
+These mechanisms, explaining the  of resistance to single insecticide use under different input values, are summarised in Table 2.
+Referring to figure 1 can help explain these responses.
 Exposure and effectiveness act by decreasing the fitness of all genotypes, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
-Cost acts by decreasing the fitness of RR and SR for those not exposed.
+Cost acts by decreasing the fitness of RR and SR among those mosquitoes  not exposed.
 
 #### Two insecticides
-Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide and, where it differs can again help us to understand the dynamics of resistance evolution.
+Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide.
+Where it differs   helps us to understand the dynamics of resistance evolution.
 Our base scenario (Fig 5A) shows curves for resistance to both insecticides in the mixture that are shallower than those for sequential use.
 The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Thus each insecticide reduces the selection pressure for the increase in resistance to the other and could be said to 'protect' the other.
-However this protection is not necessarily enough to ensure that resistance arises more slowly in a mixture relative to a sequential strategy.
-Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
-Incraesing the effectiveness of on insecticide to Fig 5B decreases the added times for the sequence to 90 and increases those for the mixture to 115 thus changing the order and favouring the mixture.
-
-The mechanism for this change can be sen by comparing the shape of the curves in Figs 5A & B.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide has become ineffective due to resistance it stops killing individuals that are resistant to the second insecticide, loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide.
-The more effective insecticide prompts both a) higher selection pressure for it's own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-
-#### Difference between exposure and effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us.
-
-#### Costs of resistance
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-It seems that the protective effect of the other insecticide in a mixture interacts with the cost of resistance to reduce selection pressure more than in a sequence.
-
-The effects of cost of resistance become more as the frequency of resistance increases.
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
-Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
-Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
-Also note that if the costs are set even higher than we set them here (e.g.
-above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-Unfortunately, we do not expect such high costs to occur in operational settings.
-
-Relatedly, when there are costs of resistance, the frequency of resistance for insecticides that are not being used is likely to decline.
-Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where it becomes operationally useful again.
-As such a repeated sequential strategy could offer benefits over mixtures not considered here.
-To incorporate this one would need to use a more detailed metric than time-to-resistance.
-
-#### Caveats
-Our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as effecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
-Insecticide resistance could reduce vector transmission of disease (e.g.
-by reduced lifespan of the vector) or increase transmission (e.g.
-by reduced immunity of the vector to the disease) [@Rivero2010].
-Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Of course, using no insecticide could have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
-Others are combining models of population genetics and population dynamics to investigate these trade-offs [@Barbosa2017].
-
-In our model we represented insecticide interaction as being additive within the mixture.
-This is consistent with recent work on aphids [@Taillebois2016], although that work did also indicate some synergistic effects.
-Our model has the potential to incorporate either synergistic or antagonistic effects too.
-
-There is much evidence of the involvement of multiple genes in resistance mechanisms.
-Our model, in common with most previous ones, just follows a single gene per insecticide.
-In most cases we believe this is reasonable.
-Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes[].
-Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
-To represent polygenic resistance a different 'quantitative genetic' approach would be required.
-A quantitive genetic model would likely lead to a more rapid initial evolution of resistance, a slower middle phase and a continuing increase.
-If polygenic resistance is shown in future to be more of a problem then these results could be compared to those of quantitative genetic models.
-
-Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
-The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-Thus there is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-In addition we provide an on-line user interface to the model [@South2017a] and enabling readers to investigate model behaviour in areas of parameter space we have not shown.
-
-#### Consistency with previous work 
-The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
-
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
-Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
-
-One consistent aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
-
-Poor implementation would likely cause lower effectiveness of and exposure to insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Barbosa2011][@FAO2012].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that effectivenes favours mixtures, exposure favours sequences and higher dominance favours neither (Table 4).
-
-#### Implications
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the development of resistance.
+Because resistant individuals are being killed, the selection pressure for that resistance is lower.
+Thus each insecticide reduces the selection pressure for resistance to the other and could be said to 'protect' the other.
+This protection means that evolution of resistance to one insecticide is always slower when it is used in a mixture than when used alone.
+This slower evolution of resistance to each insecticide in the mixture is shown in Figs.
+5-9 by the solid lines (mixture) always being shallower than the dotted lines (sequence) for the same insecticide in the same panel.
 
@@
 
+_add a paragraph, maybe not even in discussion, walking through how the online UI can be used to see the effect of changing params_
+
+_add para see notebook 11/8/17 saying about inputs that effect time-to-r, but don't effect whether mix or seq is preferred_
+
 _todo more reference to [@Roush1989] who made many of these points 25 years ago_
-
-_Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mictures over sequences.
-This would need to be balanced with a likely loss in protection to the population as a whole provided by the insecticide [@Birget2015]._
-
-_This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture is more expensive._
 
@@
 _This would seem to be a critical point to stress - with examples of how this might be acheived._
+_relating to how to maintain control & minimise resistance_
 
@@
 _Can you give some indications of the type of data that should be collected to help paramertise the model?_
-
-#### Conclusions
-
-We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured.
-For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance.
-High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other.
-High exposure favours sequences because of the higher selection pressures on both resistances together in a mixture.
-High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced.
-Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them.
-Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.
 

### 2017-08-11 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.tex
paper2 writing

@@
 \hypersetup{unicode=true,
-            pdftitle={Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary},
+            pdftitle={Insecticide resistance evolution with mixtures and sequences : a model-based introduction.},
             pdfauthor={Andy South and Ian Hastings},
@@
   \title{Insecticide resistance evolution with mixtures and sequences : a
-model-based graphical introduction and summary}
+model-based introduction.}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
   \postdate{\par}
-  \date{2017-04-11}
+  \date{2017-08-11}
 
@@
 
-\subsubsection{Insecticide effectiveness predicted to be most important
-factor determining whether insecticide mixtures can slow evolution of
-resistance.}\label{insecticide-effectiveness-predicted-to-be-most-important-factor-determining-whether-insecticide-mixtures-can-slow-evolution-of-resistance.}
-
-Target : Malaria journal (other suggestions welcome)
+Target : Malaria journal or Parasites \& Vectors
 
@@
 
-\emph{must be 350 words, currently 351}
+\emph{must be 350 words, currently 349}
 
@@
 
-Insecticide resistance threatens the control of vectors and particularly
-malaria. New insecticides are being developed to address this. Potential
-strategies for using new insecticides include mixtures, sequences and
-rotations and there are recommendations for how best to implement these
-to manage resistance. There is, however, limited recent, accessible
-modelling work addressing the evolution of resistance under different
-operational strategies. Previous work concludes that preferred
-strategies will be situation specific. There is potential to improve the
-level of mechanistic understanding within the operational community of
-how insecticide resistance can be expected to evolve in response to
-different strategies.
-
-This paper develops an accessible, mechanistic picture of how the
-evolution of insecticide resistance is likely to respond to potential
-intervention strategies to help guide both management and policy. The
-aim is to reach an audience unlikely to read a more detailed modelling
-paper. We use our model to develop a mechanistic understanding of how
-insecticide resistance is expected to increase with the use of
-insecticides in isolation, sequence and mixtures. The model flexibly
-represents two independent genes coding for resistance to two
-insecticides. We look principally at the ability of the insecticides to
-kill susceptible mosquitoes, how much resistance counteracts this, the
-proportion of mosquitoes that are exposed to insecticides and costs of
-resistance.
+Insecticide resistance threatens effective vector control, especially
+for mosquitoes and malaria. Insecticide use strategies that have been
+recommended to manage resistance include mixtures, sequences and
+rotations. New insecticides are being developed to address resistance
+and there is an opportunity to develop use strategies that limit the
+evolution of further resistance in the short term. There is limited
+recent, accessible modelling work addressing the evolution of resistance
+under different operational strategies. Previous work concludes that
+preferred strategies will be situation specific and there appears to be
+little recent consensus e.g.~on the merits of mixtures. There is an
+opportunity to improve the level of mechanistic understanding within the
+operational community of how insecticide resistance can be expected to
+evolve in response to different strategies.
+
+We provide a concise, accessible description of the use of a flexible
+model to develop a mechanistic picture the evolution of insecticide
+resistance and how it is likely to respond to potential insecticide use
+strategies. The aim is to reach an audience unlikely to read a more
+detailed modelling paper by assuming little knowledge of models. The
+model itself, as described here, represents two independent genes coding
+for resistance to two insecticides. This allows us to represent the use
+of insecticides in isolation, sequence and mixtures.
 
@@
 We use the model to demonstrate the evolution of resistance under
-different inputs and how this fits with intuitive reasoning about the
-selection pressure exerted by insecticides. We show that an insecticide
-used in a mixture, relative to alone, always prompts slower evolution of
-resistance, but resistance to the two insecticides may evolve more
-slowly when used in sequence. We show that the ability of insecticides
-to kill susceptible mosquitoes (effectiveness), has the most influence
-on whether resistance to two insecticides is likely to arise faster in a
-mixture or sequence.
+different inputs and how this fits with intuitive reasoning about
+selection pressure. We show that using an insecticide in a mixture,
+relative to alone, always prompts slower evolution of resistance to that
+insecticide. However, when the resistance to both insecticides is
+considered, resistance thresholds may be reached later for a sequence
+relative to a mixture. Increasing the ability of insecticides to kill
+susceptible mosquitoes (effectiveness), has the most influence on
+favouring a mixture over a sequence.
 
@@
 
-Our model makes more open the process of insecticide resistance
-evolution and how it is likely to respond to insecticide use. We provide
-a simple user-interface allowing further exploration
-(\url{https://andysouth.shinyapps.io/MixSeqResist1}). These tools can
-contribute to operational decisions in insecticide resistance
-management.
+Our model provides an accessible description of the process of
+insecticide resistance evolution and how it is likely to respond to
+insecticide use. We provide a simple user-interface allowing further
+exploration (\url{https://andysouth.shinyapps.io/MixSeqResist1}). These
+tools can contribute to an improved discussion about operational
+decisions in insecticide resistance management.
 
@@
 resistance management. The Innovative Vector Control Consortium (IVCC)
-was set up in 2005 to develop new vector control tools and particularly
-new insecticides to address insecticide resistance in malaria
-transmitting mosquitoes {[}8{]}{[}9{]}. Three new insecticides are now
-in development {[}9{]} and likely to be available around 2020 {[}2{]}.
-It is important that decisions about how best to use the new
-insecticides to delay the onset of resistance are made before
+was set up in 2005 to work with industry in the development of new
+vector control tools and particularly new insecticides with new modes of
+action to address insecticide resistance in malaria transmitting
+mosquitoes {[}8{]}{[}9{]}. Three new insecticides are now in development
+{[}9{]} and likely to be available after 2020 {[}2{]}. It is important
+that decisions about how best to use the new insecticides, with a clear
+mindset of delaying the evolution of resistance, are made before
 insecticides are released {[}3{]}.
@@
 selected parameter values to describe mechanistically how the evolution
-of resistance is influenced by different inputs. This mechanistic
-understanding can contribute to the debate on the relative merits of
-different insecticide strategies extending existing frameworks
+of resistance can be expected to change in different scenarios. This
+mechanistic understanding can contribute to the debate on the relative
+merits of different insecticide strategies extending existing frameworks
 {[}12{]}{[}4{]}{[}5{]}{[}1{]}.
@@
 Our modelling approach focuses on the change in frequency of a single
-resistance gene for each insecticide. It is likely that genes conferring
-resistance are present in populations even prior to exposure to novel
-insecticides thus leading to the potential for selection {[}14{]}.
+resistance gene for each insecticide. We assume both that genes
+conferring resistance are already present in the population and that
+resistance to each insecticide is coded by a single gene. These
+assumptions are consistent with previous modelling work. In support of
+the first assumption insecticide resistance has been termed
+`pre-adaptive' meaning that resistance alleles are present in very low
+numbers even prior to exposure to novel insecticides {[}14{]}{[}15{]} as
+has been documented in blowflies {[}Hartley2006{]}. In support of using
+a single gene representation of resistance evidence suggests that whilst
+polygenic resistance is common, control failures, particularly those due
+to target site resistance, are mostly due to single major genes
+{[}14{]}.
 
@@
 `time-to-resistance'. Using a 50\% threshold is consistent with previous
-modelling analyses e.g. {[}10{]}{[}15{]}. Preliminary investigations
+modelling analyses e.g. {[}10{]}{[}16{]}. Preliminary investigations
 showed that using other potential thresholds of 25 or 75\% gave very
@@
 
-The model is implemented in R {[}16{]}, the code is hosted on Github
-{[}17{]} and we provide an on-line user interface {[}18{]} (Fig 2)
+The model is implemented in R {[}17{]}, the code is hosted on Github
+{[}18{]} and we provide an on-line user interface {[}19{]} (Fig 2)
 enabling the reader to change inputs and run the model themselves.
@@
 \subsection{Results}\label{results-1}
+
+We start by describing the results of model runs with a single
+insecticide which are more straightforward and move on to describing
+results of using two insecticides in sequences and mixtures.
 
@@
 When two insecticides are used we can compare the relative performance
-of a mixture with sequential use. Each panel in Figures 5-9 compares
-mixture and sequence for a single combination of inputs. To help
-understand the mechanisms influencing this relative performance we start
-with a single base scenario and demonstrate changing inputs
-individually. In this base scenario we set effectiveness, exposure,
-resistance restoration, and dominance at 0.5, the starting frequencies
-of resistance to 0.01 and cost of resistance to 0. Resistance arises
-slower for sequential than mixture for this base scenario (Fig 5A).
-Resistance to both insecticides in the mixture follow the same path and
-reach the threshold at the same time (as would be expected given that
-they have identical input parameters). When used in sequence the curve
-(dashed) for each insecticide individually is steeper than in the
-mixture but, because they happen one after the other, it takes longer
-for both to reach the resistance threshold (Fig 5A). Thus the
-time-to-resistance for the mixture divided by the sequence is less than
-1 at 0.8.
+of concurrent use in a mixture with sequential use. Each panel in
+Figures 5-9 compares mixture and sequence for a single combination of
+inputs. To help understand the mechanisms influencing this relative
+performance we start with a single base scenario and demonstrate
+changing inputs individually. In this base scenario we set
+effectiveness, exposure, resistance restoration, and dominance at 0.5,
+the starting frequencies of resistance to 0.01 and cost of resistance to
+0. Resistance arises slower for sequential than mixture for this base
+scenario (Fig 5A). Resistance to both insecticides in the mixture follow
+the same path and reach the threshold at the same time (as would be
+expected given that they have identical input parameters). When used in
+sequence the curve (dashed) for each insecticide individually is steeper
+than in the mixture but, because they happen one after the other, it
+takes longer for both to reach the resistance threshold (Fig 5A). Thus
+the time-to-resistance for the mixture divided by the sequence is less
+than 1 at 0.8.
 
@@
 
-The results is that a more effective insecticide increases
+The result is that a more effective insecticide increases
 time-to-resistance when used in a mixture (compare the solid lines in
@@
 mechanistic explanation can help us develop a more robust understanding
-of how we would expect resistance to evolve in the field.
+of how we would expect resistance to evolve in the field. We start by
+explaining evolution of resistance to a single insecticide, which is
+relatively straightforward, and move on to explaining the effect of two
+insecticides in sequences and mixtures.
 
@@
 inputs that increased the selective advantage of resistance all led to
-more rapid evolution resistance, and higher values for cost led to
-slower evolution of (Fig 4B).
+more rapid evolution resistance. Higher values for cost led to slower
+evolution of (Fig 4B).
 
@@
 for the mixture can be explained by each insecticide killing individuals
-that are resistant to the other insecticide. Thus each insecticide
-reduces the selection pressure for the increase in resistance to the
-other and could be said to `protect' the other. However this protection
-is not necessarily enough to ensure that resistance arises more slowly
-in a mixture relative to a sequential strategy. Resistance can evolve
-more slowly for both insecticides in sequential use, despite being
-faster for each, because they occur one after the other. In Figure 5A it
-takes 60 generations for resistance to reach the threshold for each
-insecticide in the sequence, and 90 when they are applied together in
-the mixture, so the sequence is still favoured because 2x60 is greater
-than 90. Incraesing the effectiveness of on insecticide to Fig 5B
-decreases the added times for the sequence to 90 and increases those for
-the mixture to 115 thus changing the order and favouring the mixture.
-
-The mechanism for this change can be sen by comparing the shape of the
-curves in Figs 5A \& B. In the mixture, resistance to insecticide 1
-(with the increased effectiveness) rises at a similar speed in both
-figures. However, the increased effectiveness of insecticide 1 causes
-resistance to insecticide 2 to increase more slowly initially in the
-mixture. Once insecticide 1 reaches resistance of around 50\% at around
-50 generations, the curve for insecticide 2 becomes steeper. Once the
-first insecticide has become ineffective due to resistance it stops
-killing individuals that are resistant to the second insecticide, loses
-its protective effect and thus stops slowing the rise in resistance to
-the second insecticide. A similar effect is visible when the
-effectiveness of both insecticides are increased (Fig 6C). The identical
-resistance curves for each insecticide in the mixture are shallower than
-at the lower effectiveness (Fig 5A) because more individuals resistant
-to each insecticide are being killed by the other insecticide. In this
-case the `protection' given to both insecticides by the other declines
-at the same rate so we don't see the change in slope seen in Fig 5B.
+that are resistant to the other insecticide. Because resistant
+individuals are being killed, the selection pressure for that resistance
+is lower. Thus each insecticide reduces the selection pressure for the
+increase in resistance to the other and could be said to `protect' the
+other. This protection means that evolution of resistance to one
+insecticide is always slower when it is used in a mixture than when used
+alone. This slower evolution of resistance to each insecticide in the
+mixture is shown in Figs. 5-9 by the solid lines (mixture) always being
+shallower than the dotted lines (sequence) for the same insecticide in
+the same panel.
+
+However whilst this protection ensures that resistance to each
+insecticide evolves more slowly in a mixture it does not guarantee that
+a mixture strategy is preferable to a sequence once the resistance to
+both insecticides is taken into account. In a mixture selection pressure
+is applied by both insecticides at the same time, whereas in a sequence
+there is no selection for the insecticide not being used. Resistance can
+evolve more slowly for both insecticides in sequential use, despite
+being faster for each, because they occur one after the other. In Figure
+5A it takes 60 generations for resistance to reach the threshold for
+each insecticide in the sequence, and 90 when they are applied together
+in the mixture, so the sequence is still favoured because 2x60 is
+greater than 90.
+
+This advantage of sequence over mixtures can be removed by increasing
+the effectiveness of one of the insecticides. Increasing the
+effectiveness of one insecticide (Fig 5B) decreases the added times for
+the sequence from 120 to 90 and increases those for the mixture from 90
+to 115 thus changing the order and favouring the mixture. The mechanism
+for this change can be seen by comparing the shape of the curves in Figs
+5A \& B. In the mixture, resistance to insecticide 1 (with the increased
+effectiveness) rises at a similar speed in both figures. However, the
+increased effectiveness of insecticide 1 causes resistance to
+insecticide 2 to increase more slowly initially in the mixture. Once
+insecticide 1 reaches resistance of around 50\% at around 50
+generations, the curve for insecticide 2 becomes steeper. Once the first
+insecticide reaches the resistance threshold of 50\% it kills fewer
+individuals that are resistant to the second insecticide, loses its
+protective effect and thus stops slowing the rise in resistance to the
+second insecticide. A similar effect is visible when the effectiveness
+of both insecticides are increased (Fig 6C). The identical resistance
+curves for each insecticide in the mixture are shallower than at the
+lower effectiveness (Fig 5A) because more individuals resistant to each
+insecticide are being killed by the other insecticide. In this case the
+`protection' given to both insecticides by the other declines at the
+same rate so we don't see the change in slope seen in Fig 5B. In both
+cases increasing the effectiveness of insecticides increases their
+protective effect when used in mixtures and favours mixtures over
+sequences.
 
@@
 for it's own resistance and b) greater `protection' reducing the rise in
-resistance to the other insecticide.
+resistance to the other insecticide. The more effective insecticide has
+a faster evolution of resistance but this is not caused by the presence
+of the less effective one.
 
@@
 exposure reduces times-to-resistance for a mixture and effectiveness
-increases it, was unexpected for us.
+increases it, was unexpected for us and we are keen to explore it in
+more detail.
 
 \paragraph{Costs of resistance}\label{costs-of-resistance-1}
+
+Fitness costs of resistance have been widely reported(review in
+{[}20{]}) but it has been suggested that few of these are directly
+relevant to control failure in the field {[}{\textbf{???}}{]}. For this
+reason most of our analyses (with the exception of Fig 9) do not include
+costs of resistance. It has been suggested {[}{\textbf{???}}{]} that
+strategies using more than one insecticide rely on the assumption of
+costs for resistance to one insecticide to decline when it is not being
+used. Our model allows us to compare strategies with or without the
+assumption of costs. Without costs resistance levels do stay constant
+when an insecticide is not in use.
 
@@
 sequence (compare the red solid lines for a mixture to the red dotted
-line for a sequence in Fig 9C and D). It seems that the protective
-effect of the other insecticide in a mixture interacts with the cost of
+line for a sequence in Fig 9C and D). It makes sense that the protective
+effect of the other insecticide in a mixture combines with the cost of
 resistance to reduce selection pressure more than in a sequence.
@@
 the frequency of resistance increases (Fig 9C). This makes intuitive
-sense, as the frequency of resistance genes increases the impact of
+sense too, as the frequency of resistance genes increases the impact of
 fitness costs associated with those genes becomes greater. Thus
@@
 we set them here (e.g.~above 0.2) they can maintain resistance
-frequencies below 1 or lead to their decline. Unfortunately, we do not
-expect such high costs to occur in operational settings.
+frequencies below 1 or lead to their decline. However, we do not expect
+such high costs to occur in operational settings.
 
 Relatedly, when there are costs of resistance, the frequency of
-resistance for insecticides that are not being used is likely to
-decline. Thus the frequency of resistance for the first insecticide in a
-sequence, after it has been replaced, may decline to a level where it
-becomes operationally useful again. As such a repeated sequential
-strategy could offer benefits over mixtures not considered here. To
-incorporat this one would need to use a more detailed metric than
-time-to-resistance.
-
-\paragraph{Caveats}\label{caveats}
-
-Our model examines solely the evolution of resistance and not the
-control of mosquitoes or malaria. Different insecticide use strategies,
-as well as effecting the evolution of resistance, will also influence
-both the numbers of vectors and their propensity to transmit the disease
-{[}19{]}. Insecticide resistance could reduce vector transmission of
-disease (e.g.~by reduced lifespan of the vector) or increase
-transmission (e.g.~by reduced immunity of the vector to the disease)
-{[}19{]}. Our model would predict that the best strategy for limiting
-the development of resistance in the short term would be to use no
-insecticide. Of course, using no insecticide could have serious
-consequences for mosquito and malaria abundance. The challenge is to
-develop good strategies for delivering mosquito and malaria control in
-the short term while sustaining their operational capacity in the longer
-term by restricting the development of resistance. Others are combining
-models of population genetics and population dynamics to investigate
-these trade-offs {[}20{]}.
-
-In our model we represented insecticide interaction as being additive
-within the mixture. This is consistent with recent work on aphids
-{[}21{]}, although that work did also indicate some synergistic effects.
-Our model has the potential to incorporate either synergistic or
-antagonistic effects too.
-
-The development of insecticide resistance in operational field settings
-is subject to considerable uncertainty {[}2{]}. The model we present
-here is able to cover the range of that uncertainty, yet in this paper,
-to develop an understanding of the key mechanisms we restrict our
-explorations to a limited region of parameter space (e.g.~holding most
-inputs constant at intermediate values while modifying single inputs in
-isolation). Thus there is the potential that the system may behave
-differently in the field than we describe here. However, the results
-presented here are supported by an earlier analysis where we ran 10,000
-unique scenarios varying more inputs {[}13{]}. This previous analysis
-also highlighted the relative importance of insecticide effectiveness
-and exposure in determining the relative performance of mixtures and
-sequences. In addition we provide an on-line user interface to the model
-{[}18{]} and enabling readers to investigate model behaviour in areas of
-parameter space we have not shown.
+resistance for insecticides in a sequence that are not being used is
+likely to decline. Thus the frequency of resistance for the first
+insecticide in a sequence, after it has been replaced, may decline to a
+level where it becomes operationally useful again. As such a repeated
+sequential strategy could offer benefits over mixtures not considered
+here. To quantify these benefits one would need to use a more detailed
+metric than the time-to-resistance we have used. We quantified the
+number of generations until resistance thresholds are first reached.
+Other measures for quantifying resistance over time can be imagined.
 
@@
 
-A subsequent model {[}22{]} lead to the following more pessimistic
+A subsequent model {[}21{]} lead to the following more pessimistic
 conclusion about the potential of mixtures : ``As a result of incomplete
@@
 rarely consistently high enough for pesticide mixtures to be effective''
-{[}22{]}. This ``mortality of susceptible homozygotes'' is equivalent to
+{[}21{]}. This ``mortality of susceptible homozygotes'' is equivalent to
 our insecticide effectiveness. Again this points to the importance of
@@
 showing that decreasing exposure favours mixtures over sequences is also
-consistent with figure 3 from that paper {[}22{]}.
-
-One consistent aspect of the results presented here can seem to
-contradict published advice and has been questioned when we have
-presented model results. Our model predicts that lower insecticide
-effectiveness or exposure will lead to slower spread of resistance in
-sole use and sequences (Figs 3 \& 5, Table 2). This is consistent with
-expectations that lower effectiveness and exposure would lead to lower
-mortality of susceptibles and thus lower selection pressure for
-resistance {[}22{]}. This can seem to contradict recommendations that
-poor implementation of insecticide interventions can promote the
-development of resistance. e.g.~from {[}5{]} ``Certain pest control
-practices have consistently been shown to exacerbate the loss of
-susceptible pest populations and the development of resistance. These
-include: \ldots{} - the use of application rates that are below or above
-those recommended on the label, - poor coverage of the area being
-treated \ldots{}''.
+consistent with figure 3 from that paper {[}21{]}.
+
+One aspect of the results presented here can seem to contradict
+published advice and has been questioned when we have presented model
+results. Our model predicts that lower insecticide effectiveness or
+exposure will lead to slower spread of resistance in sole use and
+sequences (Figs 3 \& 5, Table 2). This is consistent with expectations
+that lower effectiveness and exposure would lead to lower mortality of
+susceptibles and thus lower selection pressure for resistance {[}21{]}.
+This can seem to contradict recommendations that poor implementation of
+insecticide interventions can promote the development of resistance.
+e.g.~from {[}5{]} ``Certain pest control practices have consistently
+been shown to exacerbate the loss of susceptible pest populations and
+the development of resistance. These include: \ldots{} - the use of
+application rates that are below or above those recommended on the
+label, - poor coverage of the area being treated \ldots{}''.
 
@@
 effectively increasing the dominance of the resistant allele and
-increasing its rate of evolution {[}23{]}{[}5{]}. Our model results are
+increasing its rate of evolution {[}22{]} {[}5{]}. Our model results are
 consistent with this, showing that increasing dominance of restoration
@@
 benefits of mixtures and sequences will depend on these trade-offs
-because of our result that effectivenes favours mixtures, exposure
-favours sequences and higher dominance favours neither (Table 4).
-
-\paragraph{Implications}\label{implications}
+because of our result that effectiveness favours mixtures, exposure
+favours sequences and higher dominance favours neither (Table 4). More
+research into how the mortality of different genotypes (susceptible,
+resistant and heterozygous) is affected by poor application rates or
+insecticide degradation over time would help make this clearer.
 
@@
 a more Integrated Vector Management composed of a series of partially
-effective tools {[}24{]}. It is suggested that such an integrated
-approach is more sustainable and `evolution-proof' {[}24{]}. A similar
+effective tools {[}23{]}. It is suggested that such an integrated
+approach is more sustainable and `evolution-proof' {[}23{]}. A similar
 combining of interventions has been advocated to cover mosquitoes
-exhibiting different behaviours {[}25{]}. Mosquitoes with different
+exhibiting different behaviours {[}24{]}. Mosquitoes with different
 behaviours e.g.~feeding outdoors and/or on animals are likely to result
@@
 
-\emph{TODO bits to incorporate into discussion :}
-
-\emph{todo more reference to {[}12{]} who made many of these points 25
-years ago}
-
-\emph{Insecticides with some repellency will reduce exposure to the
-insecticide and thus, as shown here, are likely to slow the evolution of
-resistance and favour mictures over sequences. This would need to be
-balanced with a likely loss in protection to the population as a whole
-provided by the insecticide {[}15{]}.}
-
-\emph{This model can explore when a mixture strategy may be favoured
-over a sequence purely for the evolution of resistance. Other factors
-also need to be considered. On the positive side a mixture is likely to
-kill more vectors and may also limit their ability to transmit the
-disease. On the negative side a mixture is more expensive.}
-
-\emph{todo, somewhere earlier say good thing about effectiveness \&
-exposure is that they can be modified}
+\paragraph{Caveats}\label{caveats}
+
+There is much evidence of the involvement of multiple genes in
+resistance mechanisms. Our model, in common with most previous ones,
+just follows a single gene per insecticide. In most cases we believe
+this is reasonable. Target-site resistance has been shown to be mostly
+determined by single genes, and control failures in general have mostly
+been attributed to the effect of single major genes {[}{\textbf{???}}{]}
+{[}14{]}. Control doses of insecticides in the field are likely to lead
+to the selection of single major genes whereas the lower doses used to
+select laboratory strains are more likely to lead to polygenic
+resistance {[}14{]}. To represent polygenic resistance a different
+`quantitative genetic' approach would be required. A quantitative
+genetic model would likely lead to a more rapid initial evolution of
+resistance, a slower middle phase and a continuing increase. If
+polygenic resistance is shown in future to be more of a problem then
+these results could be compared to those of quantitative genetic models.
+
+Even if the assumption that single genes are responsible for control
+failures is correct there is still considerable other uncertainty around
+the development of insecticide resistance in operational field settings
+{[}2{]}. The model we present here is able to cover much of the range of
+that uncertainty, yet in this paper, to develop an understanding of the
+key mechanisms we restrict our explorations to a limited region of
+parameter space (e.g.~holding most inputs constant at intermediate
+values while modifying single inputs in isolation). There is the
+potential that the system may behave differently in the field than we
+describe here. However, the results presented here are supported by an
+earlier analysis where we ran 10,000 unique scenarios varying more
+inputs {[}13{]}. This previous analysis also highlighted the relative
+importance of insecticide effectiveness and exposure in determining the
+relative performance of mixtures and sequences. In addition we provide
+an on-line user interface to the model {[}19{]} enabling readers to
+investigate model behaviour in areas of parameter space we have not
+shown.
+
+There are other model behaviours that can be changed by modifying input
+files. For example, here we assume that insecticide interaction in a
+mixture is multiplicative, e.g.~if the probability of surviving exposure
+to insecticide A alone is 0.3 and of surviving insecticide B alone is
+0.2, then the probability of surviving exposure to a mixture of A and B
+is 0.3*0.2 = 0.06. This is consistent with recent work on aphids
+{[}25{]}, although that did also indicate some synergistic effects.
+Modifying input files would allow us to incorporate such synergistic or
+antagonistic effects of mixtures.
+
+In addition to these uncertainties regarding resistance it is important
+to remember that our model examines solely the evolution of resistance
+and not the control of mosquitoes or malaria. Different insecticide use
+strategies, as well as affecting the evolution of resistance, will also
+influence both the numbers of vectors and their propensity to transmit
+the disease {[}26{]}. There are also potential knock-on effects of
+resistance itself; reduced vector transmission of disease (e.g.~by
+reduced lifespan of the vector) or increased transmission (e.g.~by
+reduced immunity of the vector to the disease) {[}26{]}. This model can
+explore when a mixture strategy may be favoured over a sequence purely
+for the evolution of resistance. Other operational factors also need to
+be considered. On the positive side a mixture is likely to kill more
+vectors and may also limit their ability to transmit the disease. On the
+negative side a mixture may be more expensive. Other attributes of
+insecticides such as their repellency are also relevant. Insecticides
+with some repellency will reduce exposure to the insecticide and thus,
+as shown here, are likely to slow the evolution of resistance and favour
+mixtures over sequences. However the reduced exposure resulting from
+repellency also reduces the population protection provided by the
+insecticide {[}16{]}.
+
+Our model would predict that the best strategy for limiting the
+development of resistance in the short term would be to use no
+insecticide. Of course, using no insecticide could have serious
+consequences for mosquito and malaria abundance. The challenge is to
+develop good strategies for delivering mosquito and malaria control in
+the short term while sustaining their operational capacity in the longer
+term by restricting the development of resistance.
 
@@
 strategies such as rotations and mosaics and we are starting to
-investigate them. Predictions from the model can be tested against data
-from trials of new insectides to help advance guidelines on good
-practice.
+investigate them. Predictions from the model will also be tested against
+data from insecticide trials to help understand better the process of
+resistance evolution.
+
+\emph{TODO bits to incorporate into discussion :}
+
+\emph{todo more reference to {[}12{]} who made many of these points 25
+years ago}
+
+\emph{todo, somewhere earlier say good thing about effectiveness \&
+exposure is that they can be modified}
+
+\emph{comments from Hilary} \emph{This would seem to be a critical point
+to stress - with examples of how this might be acheived.} \emph{relating
+to how to maintain control \& minimise resistance}
+
+\emph{Limitations that haven't been addressed is the possiblility that
+resistance is poylgenic, and the possibility of cross resistance (e.g a
+P450 may cause partial resistance to both).}
+
+\emph{Something further to guide readers on how they might determine
+dominance would also be useful given how crucial this is.}
+
+\emph{Can you give some indications of the type of data that should be
+collected to help paramertise the model?}
 
@@
 
+\hypertarget{ref-Ffrench-Constant2013}{}
+14. Ffrench-Constant RH: \textbf{The molecular genetics of insecticide
+resistance}. 2013:807--815.
+
 \hypertarget{ref-Liu2015}{}
-14. Liu N: \textbf{Insecticide Resistance in Mosquitoes: Impact,
+15. Liu N: \textbf{Insecticide Resistance in Mosquitoes: Impact,
 Mechanisms, and Research Directions}. \emph{Annual Review of Entomology}
@@
 \hypertarget{ref-Birget2015}{}
-15. Birget PLG, Koella JC: \textbf{A genetic model of the effects of
+16. Birget PLG, Koella JC: \textbf{A genetic model of the effects of
 insecticide-treated bed nets on the evolution of
@@
 \hypertarget{ref-RCoreTeam2016}{}
-16. R Core Team: \textbf{R: A language and environment for statistical
+17. R Core Team: \textbf{R: A language and environment for statistical
 computing.} 2016.
@@
 \hypertarget{ref-South2017}{}
-17. South A: \textbf{https://github.com/AndySouth/resistance}.
+18. South A: \textbf{https://github.com/AndySouth/resistance}.
 
 \hypertarget{ref-South2017a}{}
-18. South A: \textbf{Online model user interface :
+19. South A: \textbf{Online model user interface :
 https://andysouth.shinyapps.io/resistmob2/}. 2017.
 
-\hypertarget{ref-Rivero2010}{}
-19. Rivero A, Vézilier J, Weill M, Read AF, Gandon S:
-\textbf{Insecticide control of vector-borne diseases: When is
-insecticide resistance a problem?} \emph{PLoS Pathogens} 2010.
-
-\hypertarget{ref-Barbosa2017}{}
-20. Barbosa S, Kay K, Chitnis N, Hastings IM, Tropical S:
-\textbf{Modelling the public health impact of insecticide deployment on
-malaria-transmitting mosquito populations and its role in driving
-insecticide resistance .} 2017:1--58.
-
-\hypertarget{ref-Taillebois2016}{}
-21. Taillebois E, Thany SH: \textbf{The differential effect of low-dose
-mixtures of four pesticides on the pea aphid Acyrthosiphon pisum}.
-\emph{Insects} 2016, \textbf{7}:1--7.
+\hypertarget{ref-Kliot2012}{}
+20. Kliot A, Ghanim M: \textbf{Fitness costs associated with insecticide
+resistance}. \emph{Pest Management Science} 2012,
+\textbf{68}:1431--1437.
 
 \hypertarget{ref-Roush1998}{}
-22. Roush RT: \textbf{Two-toxin strategies for management of
+21. Roush RT: \textbf{Two-toxin strategies for management of
 insecticidal transgenic crops: can pyramiding succeed where pesticide
@@
 \hypertarget{ref-Barbosa2011}{}
-23. Barbosa S, Black IV WC, Hastings I: \textbf{Challenges in estimating
+22. Barbosa S, Black IV WC, Hastings I: \textbf{Challenges in estimating
 insecticide selection pressures from mosquito field data}. \emph{PLoS
@@
 \hypertarget{ref-Thomas2012}{}
-24. Thomas MB, Godfray HCJ, Read AF, Berg H van den, Tabashnik BE,
+23. Thomas MB, Godfray HCJ, Read AF, Berg H van den, Tabashnik BE,
 Lenteren JC van, Waage JK, Takken W: \textbf{Lessons from agriculture
@@
 \hypertarget{ref-Killeen2017}{}
-25. Killeen GF, Marshall JM, Kiware SS, Andy S, Chaki PP, Govella NJ:
+24. Killeen GF, Marshall JM, Kiware SS, Andy S, Chaki PP, Govella NJ:
 \textbf{Measuring, manipulating and exploiting behaviours of adult
@@
 
+\hypertarget{ref-Taillebois2016}{}
+25. Taillebois E, Thany SH: \textbf{The differential effect of low-dose
+mixtures of four pesticides on the pea aphid Acyrthosiphon pisum}.
+\emph{Insects} 2016, \textbf{7}:1--7.
+
+\hypertarget{ref-Rivero2010}{}
+26. Rivero A, Vézilier J, Weill M, Read AF, Gandon S:
+\textbf{Insecticide control of vector-borne diseases: When is
+insecticide resistance a problem?} \emph{PLoS Pathogens} 2010.
+
 \end{document}

### 2017-09-08 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
add couple sentences to intro

@@
 ---
-title: "Insecticide resistance evolution with mixtures and sequences : a model-based introduction."
+title: "Insecticide resistance evolution with mixtures and sequences : a model-based explanation."
 output: 
@@
 
-### version 12
-
-###alternative titles:
-###Evolution of insecticide resistance with insecticide mixtures and sequences : an accessible model to promote understanding.
-###An accessible model to promote understanding of the evolution of insecticide resistance : insecticide effectiveness predicted to be key to the relative performance of mixtures.
-###A tutorial, primer ...
+### version 13
 
@@
 
+A recent comprehensive review of strategies to avoid resistance evolution across pesticides and drugs [@Consortium2013] concluded that mixtures (combination of molecules) are usually the best resistance management strategy.
+This was based on both empirical and modelling work.
 Modelling studies have investigated the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
@@
 Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance can be expected to change in different scenarios.
-This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012].
+This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012][@Consortium2013].
 
@@
 _add para see notebook 11/8/17 saying about inputs that effect time-to-r, but don't effect whether mix or seq is preferred_
-
-_todo more reference to [@Roush1989] who made many of these points 25 years ago_
 

### 2017-09-11 inst/documents/paper2/test_linenumbers.Rmd
testing

@@
 
-output: 
-  pdf_document:
-    #citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-#output: word_document
+# output: 
+#   pdf_document:
+#     keep_tex: true
+#     fig_caption: true
+#     latex_engine: pdflatex
+output: word_document
 author: "Andy South and Ian Hastings"
@@
 
-Seems that the header-includes bit above does work to add line numbers and double space (although it does generate a few warning messages)
+Seems that the header-includes bit above works to add line numbers and double space to pdf, but not to word doc.
+
+Tables, figs and ref list seem to work both for pdf & word doc.
 
@@
 ```
+
+**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
+
+Parameter to increase  | effect on resistance evolution | Mechanism
+----------------------- | -------- | ---------------------------------------------
+1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
+2. Exposure | faster | reduced fitness of SS overall
+3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
+4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
+5. Frequency  | faster | less change needed to reach resistance threshold
+6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
+
+**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
+
+Parameter to increase  | effect on resistance evolution | Mechanism
+----------------------- | ---------- | ---------------------------------------------
+1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure from the other
+2.
+Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
+However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
+3. Dominance of restoration | faster | increased fitness of heterozygotes
+4. Resistance restoration | faster | increased fitness of resistants
+5. Frequency  | faster | less change needed to reach resistance threshold
+6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
+
+\pagebreak
+
+**Table 4. Effect of inputs on the difference between mixture and sequential use**
+
+Parameter to increase  | increase favours mix or sequence | Mechanism
+----------------------- | -------- | ---------------------------------------------
+1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
+2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
+3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
+4.
+Resistance restoration | neither | As for dominance of restoration.
+Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
+5.
+Frequency  | neither | As for dominance and resistance restoration.
+Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
+6.
+Cost of resistance | mixture | Higher costs slow the evolution of resistance more in a mixture than when used in sequence (but with higher costs in a sequence there is a greater chance for resistance levels to decline for the insecticide not being used).

### 2017-09-11 inst/documents/paper2/test_linenumbers.tex
testing

@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{plainnat}
+\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
   \postdate{\par}
-  \date{2017-04-11}
+  \date{2017-09-11}
 
@@
 \maketitle
+
+Seems that the header-includes bit above does work to add line numbers
+and double space (although it does generate a few warning messages)
 
@@
 
-Insecticide resistance is a problem for malaria
-\citep{WHO2012}\citep{Ranson2016}\citep{Hemingway2016} other vector
-borne diseases \citep{IRAC2011} and agriculture \citep{FAO2012}. Malaria
+Insecticide resistance is a problem for malaria {[}1{]}{[}2{]}{[}3{]}
+other vector borne diseases {[}4{]} and agriculture {[}5{]}. Malaria
 alone still results in hundreds of thousands of deaths per year. Recent
@@
 to averting more than 500 million falciparum malaria cases between 2000
-and 2015 \citep{Bhatt2015}. A recent malaria transmission model
-\citep{Churcher2016} predicts that even low pyrethroid resistance is
-likely to increase malaria incidence in Africa by reducing the
-performance of bed nets.
+and 2015 {[}6{]}. A recent malaria transmission model {[}7{]} predicts
+that even low pyrethroid resistance is likely to increase malaria
+incidence in Africa by reducing the performance of bed nets.
 
 The WHO produced a Global Plan for Insecticide Resistance Management in
-malaria vectors (GPIRM)\citep{WHO2012} which includes recommendations on
+malaria vectors (GPIRM){[}1{]} which includes recommendations on
 operational strategies for managing resistance including the use of
@@
 new insecticides to address insecticide resistance in malaria
-transmitting mosquitoes \citep{Hemingway2006}\citep{IVCC2016}. Three new
-insecticides are now in development \citep{IVCC2016} and likely to be
-available around 2020 \citep{Ranson2016}. It is important that decisions
-about how best to use the new insecticides to delay the onset of
-resistance are made before insecticides are released
-\citep{Hemingway2016}.
+transmitting mosquitoes {[}8{]}{[}9{]}. Three new insecticides are now
+in development {[}9{]} and likely to be available around 2020 {[}2{]}.
+It is important that decisions about how best to use the new
+insecticides to delay the onset of resistance are made before
+insecticides are released {[}3{]}.
 
@@
 resistance in insecticide mixtures including in a public health context
-e.g. \citep{Curtis1985}\citep{Mani1985}\citep{Roush1989} but much of the
-work was done more than 20 years ago and there remained some confusion
-about the results \citep{Levick2017}. In an earlier paper
-\citep{Levick2017} we described the technical details of a flexible
-model used to investigate the relative benefits of mixtures and
+e.g. {[}10{]}{[}11{]}{[}12{]} but much of the work was done more than 20
+years ago and there remained some confusion about the results {[}13{]}.
+In an earlier paper {[}13{]} we described the technical details of a
+flexible model used to investigate the relative benefits of mixtures and
 sequences. Here we provide an accessible summary of the model and use
@@
 different insecticide strategies extending existing frameworks
-\citep{Roush1989}\citep{IRAC2011}\citep{FAO2012}\citep{WHO2012}.
+{[}12{]}{[}4{]}{[}5{]}{[}1{]}.
 
@@
 
-\bibliography{library}
+\textbf{Table 2. Effect of inputs on resistance when insecticides used
+singly or in sequence}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.28\columnwidth}\raggedright\strut
+Parameter to increase\strut
+\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+effect on resistance evolution\strut
+\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright\strut
+Mechanism\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+reduced fitness of SS in presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+reduced fitness of SS overall\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+3. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+increased fitness of SR in presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+4. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+increased fitness of RR in presence of insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+less change needed to reach resistance threshold\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+slower\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+reduced fitness of RR in absence of insecticide\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Table 3. Effect of inputs on resistance when insecticides used
+in a mixture}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.27\columnwidth}\raggedright\strut
+Parameter to increase\strut
+\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
+effect on resistance evolution\strut
+\end{minipage} & \begin{minipage}[b]{0.52\columnwidth}\raggedright\strut
+Mechanism\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+\textbf{slower}\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+one insecticide reduces the fitness of individuals resistant to the
+other thereby reducing selection pressure from the other\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster (but less than for single)\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+reduced fitness of individuals susceptible to one insecticide increases
+selection pressure for that resistance. However at the same time
+selection pressure is reduced by reduced fitness of resistant
+individuals caused by the other insecticide\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+3. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+increased fitness of heterozygotes\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+4. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+increased fitness of resistants\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+faster\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+less change needed to reach resistance threshold\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.27\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
+slower\strut
+\end{minipage} & \begin{minipage}[t]{0.52\columnwidth}\raggedright\strut
+reduced fitness of resistants in absence of insecticide\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\pagebreak
+
+\textbf{Table 4. Effect of inputs on the difference between mixture and
+sequential use}
+
+\begin{longtable}[]{@{}lll@{}}
+\toprule
+\begin{minipage}[b]{0.28\columnwidth}\raggedright\strut
+Parameter to increase\strut
+\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
+increase favours mix or sequence\strut
+\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright\strut
+Mechanism\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+1. Effectiveness\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+mixture\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher effectiveness gives faster resistance for sequence and slower
+resistance in mixture\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+2. Exposure\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+sequence\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher exposure gives faster resistance for sequence and mixture but the
+greater effect on mixture favours sequence.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+3. Dominance of restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+neither\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher dominance gives faster resistance in both sequences and mixtures
+such that the difference between them is not changed.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+4. Resistance restoration\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+neither\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+As for dominance of restoration. Higher resistance restoration gives
+faster resistance in both sequences and mixtures such that the
+difference between them is not changed.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+5. Frequency\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+neither\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+As for dominance and resistance restoration. Higher starting frequencies
+give faster resistance in both sequences and mixtures such that the
+difference between them is not changed.\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.28\columnwidth}\raggedright\strut
+6. Cost of resistance\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+mixture\strut
+\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright\strut
+Higher costs slow the evolution of resistance more in a mixture than
+when used in sequence (but with higher costs in a sequence there is a
+greater chance for resistance levels to decline for the insecticide not
+being used).\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{refs}{}
+\hypertarget{ref-WHO2012}{}
+1. WHO: \emph{Global plan for insecticide resistance management in
+malaria vectors (GPIRM).} Geneva. 2012:130.
+
+\hypertarget{ref-Ranson2016}{}
+2. Ranson H, Lissenden N: \textbf{Insecticide Resistance in African
+Anopheles Mosquitoes: A Worsening Situation that Needs Urgent Action to
+Maintain Malaria Control}. \emph{Trends in Parasitology} 2016,
+\textbf{32}:187--196.
+
+\hypertarget{ref-Hemingway2016}{}
+3. Hemingway J, Ranson H, Magill A, Kolaczinski J, Fornadel C, Gimnig J,
+Coetzee M, Simard F, Roch DK, Hinzoumbe CK, Pickett J, Schellenberg D,
+Gething P, Hoppé M, Hamon N: \textbf{Averting a malaria disaster: Will
+insecticide resistance derail malaria control?} \emph{The Lancet} 2016,
+\textbf{387}:1785--1788.
+
+\hypertarget{ref-IRAC2011}{}
+4. IRAC: \emph{Prevention and Management of Insecticide Resistance in
+Vectors of Public Health Importance}. 2011:71.
+
+\hypertarget{ref-FAO2012}{}
+5. FAO: \emph{Guidelines on Prevention and Management of Pesticide
+Resistance}. Rome; 2012:55.
+
+\hypertarget{ref-Bhatt2015}{}
+6. Bhatt S, Weiss DJ, Cameron E, Bisanzio D, Mappin B, Dalrymple U,
+Battle KE, Moyes CL, Henry A, Eckhoff PA, Wenger EA, Briët O, Penny MA,
+Smith TA, Bennett A, Yukich J, Eisele TP, Griffin JT, Fergus CA, Lynch
+M, Lindgren F, Cohen JM, Murray CLJ, Smith DL, Hay SI, Cibulskis RE,
+Gething PW: \textbf{The effect of malaria control on Plasmodium
+falciparum in Africa between 2000 and 2015.} \emph{Nature} 2015,
+\textbf{526}:207--11.
+
+\hypertarget{ref-Churcher2016}{}
+7. Churcher TS, Lissenden N, Griffin JT, Worrall E, Ranson H:
+\textbf{The impact of pyrethroid resistance on the efficacy and
+effectiveness of bednets for malaria control in Africa}. \emph{eLife}
+2016, \textbf{5}:e16090.
+
+\hypertarget{ref-Hemingway2006}{}
+8. Hemingway J, Beaty BJ, Rowland M, Scott TW, Sharp BL: \textbf{The
+Innovative Vector Control Consortium: improved control of mosquito-borne
+diseases}. \emph{Trends in Parasitology} 2006, \textbf{22}:308--312.
+
+\hypertarget{ref-IVCC2016}{}
+9. IVCC: \textbf{Annual Report 2015-16}. 2016.
+
+\hypertarget{ref-Curtis1985}{}
+10. Curtis CF: \textbf{Theoretical models of the use of insecticide
+mixtures for the management of resistance}. \emph{Bulletin of
+Entomological Research} 1985, \textbf{75}:259.
+
+\hypertarget{ref-Mani1985}{}
+11. Mani GS: \textbf{Evolution of resistance in the presence of two
+insecticides.} \emph{Genetics} 1985, \textbf{109}:761--783.
+
+\hypertarget{ref-Roush1989}{}
+12. Roush RT: \textbf{Designing resistance management programs: How can
+you choose?} \emph{Pesticide Science} 1989, \textbf{26}:423--441.
+
+\hypertarget{ref-Levick2017}{}
+13. Levick B, South A, Hastings IM: \textbf{A Two-locus Model of The
+Evolution of Insecticide Resistance to Inform and Optimise Public Health
+Insecticide Deployment Strategies.} \emph{PLOS Computational Biology}
+2017.
 

### 2017-09-11 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

@@
 title: "Insecticide resistance evolution with mixtures and sequences : a model-based explanation."
-output: 
-  pdf_document:
-    #citation_package: natbib
-    keep_tex: true
-    fig_caption: true
-    latex_engine: pdflatex
-
-#output: word_document
+# output: 
+#   pdf_document:
+#     keep_tex: true
+#     fig_caption: true
+#     latex_engine: pdflatex
+
+output: word_document
 author: "Andy South and Ian Hastings"
@@
 #!!!! paper protocol
-#!!!! might be easier to put text and figs into separate Rmd files after draft stage
-#!!!! drafts knit to pdf with pdf_document in YAML and this chunk eval=FALSE
 #!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
@@
 #!!!! submission figs, set this chunk to eval=TRUE
+#!!!! previous drafts knitted to pdf with pdf_document in YAML and this chunk eval=FALSE
 

### 2017-09-11 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
abstract and discussion reworked in : paper2 discussion reworking2.docx

@@
 
-### version 13
-
-Target : Malaria journal or Parasites & Vectors
+### version 14
+
+Target : Malaria journal
 
@@
 
-_must be 350 words, currently 349_
+_must be 350 words, currently 348_
 
@@
 Insecticide resistance threatens effective vector control, especially for mosquitoes and malaria.
-Insecticide use strategies that have been recommended to manage resistance include mixtures, sequences and rotations.
-New insecticides are being developed to address resistance and there is an opportunity to develop use strategies that limit the evolution of further resistance in the short term.
-There is limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
-Previous work concludes that preferred strategies will be situation specific and there appears to be little recent consensus e.g.
-on the merits of mixtures.
-There is an opportunity to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
-
-We provide a concise, accessible description of the use of a flexible model to develop a mechanistic picture the evolution of insecticide resistance and how it is likely to respond to potential insecticide use strategies.
-The aim is to reach an audience unlikely to read a more detailed modelling paper by assuming little knowledge of models.
+To manage resistance, recommended insecticide use strategies include mixtures, sequences and rotations.
+New insecticides are being developed and there is an opportunity to develop use strategies that limit the evolution of further resistance in the short term.
+A 2013 review of modelling and empirical studies of resistance points to the advantages of mixtures.
+However, there is limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
+We see an opportunity to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
+
+We provide a concise, accessible description of a flexible model of the evolution of insecticide resistance.
+We use the model to develop a mechanistic picture the evolution of insecticide resistance and how it is likely to respond to potential insecticide use strategies.
+The aim is to reach an audience unlikely to read a more detailed modelling paper.
 The model itself, as described here, represents two independent genes coding for resistance to two insecticides.
@@
 However, when the resistance to both insecticides is considered, resistance thresholds may be reached later for a sequence relative to a mixture.
-Increasing the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on favouring a mixture over a sequence.
+Increasing the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on favouring a mixture over a sequence because one highly effective insecticide provides more protection to another in a mixture.
 
@@
 ## Discussion
-
 Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
@@
 
-When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance.
-Higher values for cost led to slower evolution of (Fig 4B).
-
-These mechanisms, explaining the  of resistance to single insecticide use under different input values, are summarised in Table 2.
+When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 4B).
+
+These mechanisms, explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
 Referring to figure 1 can help explain these responses.
-Exposure and effectiveness act by decreasing the fitness of all genotypes, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
+Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
 Cost acts by decreasing the fitness of RR and SR among those mosquitoes  not exposed.
@@
 #### Two insecticides
-Using two insecticides either in sequence or in mixtures differs from the use of a single insecticide.
-Where it differs   helps us to understand the dynamics of resistance evolution.
-Our base scenario (Fig 5A) shows curves for resistance to both insecticides in the mixture that are shallower than those for sequential use.
+Responses to changing input parameters are different when two insecticides are used either in sequence or in mixtures as compared to when used alone.
+These differences can help us to understand the dynamics of resistance evolution.
+In each panel of Figures 5-9 the solid lines, indicating that an insecticide is being used at the same time as another in a mixture, are always shallower than the same colour dashed line indicating the insecticide is being used in isolation as part of a sequence.
 The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
@@
 This protection means that evolution of resistance to one insecticide is always slower when it is used in a mixture than when used alone.
-This slower evolution of resistance to each insecticide in the mixture is shown in Figs.
-5-9 by the solid lines (mixture) always being shallower than the dotted lines (sequence) for the same insecticide in the same panel.
-
-_TODO bits to incorporate into discussion :_
-
-_add a paragraph, maybe not even in discussion, walking through how the online UI can be used to see the effect of changing params_
-
-_add para see notebook 11/8/17 saying about inputs that effect time-to-r, but don't effect whether mix or seq is preferred_
-
-_todo, somewhere earlier say good thing about effectiveness & exposure is that they can be modified_
-
-_comments from Hilary_
-_This would seem to be a critical point to stress - with examples of how this might be acheived._
-_relating to how to maintain control & minimise resistance_
-
-_Limitations that haven't been addressed is the possiblility that resistance is poylgenic, and the possibility of cross resistance (e.g a P450 may cause partial resistance to both)._
-
-_Something further to guide readers on how they might determine dominance would also be useful given how crucial this is._
-
-_Can you give some indications of the type of data that should be collected to help paramertise the model?_
+This protection has been termed multiple intra-generational killing [@Consortium2013] because individuals have the potential to be killed by more than one insecticide.
+
+Whilst this protection ensures that resistance to each insecticide evolves more slowly in a mixture it does not guarantee that a mixture strategy is preferable  to a sequence once the resistance to both insecticides is taken into account.
+In a mixture, selection pressure is applied by both insecticides at the same time, whereas in a sequence there is no selection for the insecticide not being used.
+Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
+In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
+
+This advantage of sequence over mixtures can be removed by increasing the effectiveness of one of the insecticides.
+Increasing the effectiveness of one insecticide (Fig 5B) decreases the total time for the sequence from 120 to 90 and increases that for the mixture from 90 to 115 thus changing the order and favouring the mixture.
+The mechanism for this change can be seen by comparing the shape of the curves in Figs 5A & B.
+In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
+However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
+Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
+Once the first insecticide reaches the resistance threshold of 50% it kills fewer individuals that are resistant to the second insecticide, it loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
+A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
+The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
+In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
+In both cases increasing the effectiveness of insecticides increases their protective effect when used in mixtures and favours mixtures over sequences.
+
+Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide (the red curve in this case).
+The more effective insecticide prompts both a) higher selection pressure for its own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
+The more effective insecticide has a faster evolution of resistance than the less effective one but this is not caused by the presence of the less effective one.
+(Indeed the presence of the less effective one is still slowing the evolution of resistance to the more effective one, as indicated by the red solid curve for use in the mixture being shallower than the red dashed curve for use alone as part of the sequence).
+
+#### Difference between exposure and effectiveness
+It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
+Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
+This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
+This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us and we are keen to explore it in more detail.
+The most likely explanation is that increasing exposure increases overall selection pressure and lowers time to resistance, whereas increasing effectiveness increases mutual protection of the insecticides in a mixture and lowers the speed at which resistance evolves.
+
+#### Costs of resistance
+Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@ffrench-Constant2017].
+Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 9).
+
+Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
+It makes sense that the protective effect of the other insecticide in a mixture combines with the cost of resistance to reduce selection pressure more than in a sequence.
+
+The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
+This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
+Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
+Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
+Also note that if the costs are set even higher than we set them here (e.g.
+above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
+However, we do not expect such high costs to occur in operational settings.
+
+Declines in resistance frequencies are likely to occur when there are costs of resistance and the insecticide is not being used [@ffrench-Constant2017].
+Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where the first insecticide becomes operationally useful again.
+Consequently, a repeated sequential strategy could offer benefits over mixtures not considered here.
+To quantify these benefits one would need to use a more detailed metric than the time-to-resistance we have used.
+We quantified the number of generations until resistance thresholds are first reached.
+Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a threshold.
+
+#### Consistency with previous work 
+The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
+A recent review [@Consortium2013] found that 8 out of 10 empirical studies and 11 out of 14 modelling studies favoured mixtures over sequences.
+In the 3 other modelling studies the relative performance of mixtures and sequences were dependent on the values of other inputs as we also have shown.
+
+The review concluded that the advantage of mixtures is greatest when seven other conditions are satisfied [@Consortium2013] : i) rare starting resistance , ii) independent loci (i.e.
+no cross resistance), iii) high recombination, iv) high susceptible mortality, v) resistance is recessive, vi) similar persistence of insecticides, and vii) some of population remains untreated.
+Their points iv) and vii) agree with our results that mixtures are favoured by high effectiveness and low exposure.
+Some of the other conditions, notably i) the starting frequency of resistance and v) the dominance of resistance did not alter whether mixtures or sequences were favoured in our model [@Levick2017].
+
+Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
+One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
+That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
+This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
+This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
+see Fig 6D and Table 4).
+
+A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
+This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
+Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
+Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
+
+One aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
+Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
+This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
+@Roush1998].
+This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
+e.g.
+from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
+These include: ...
+- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
+
+Poor implementation would likely cause lower effectiveness of, and exposure to, insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
+The explanation for this apparent difference is in the effect of dominance.
+Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [Bourguet2000][@Barbosa2011] [@FAO2012] see Fig 1 of [@Levick2017].
+Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
+So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
+Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that higher effectiveness favours mixtures, higher exposure favours sequences and higher dominance favours neither (Table 4).
+More research into how the mortality of different genotypes (susceptible, resistant and heterozygous) is affected by poor application rates or insecticide degradation over time would help make this clearer.
+
+Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
+It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
+A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
+Mosquitoes with different behaviours e.g.
+feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
+This lower exposure could favour mixtures over sequences as we have indicated.
+In both cases our modelling approach could be used to investigate implications for the development of resistance.
+
+#### Caveats
+There is much evidence of the involvement of multiple genes in resistance mechanisms.
+Our model, in common with most previous ones, assumes a single ‘resistance’ gene per insecticide.
+In most cases we believe this is reasonable.
+Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Ffrench-Constant2004] [@Ffrench-Constant2013].
+Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
+To represent polygenic resistance a different 'quantitative genetic' approach would be required.
+A quantitative genetic model would likely lead to a more rapid initial evolution of resistance, a slower middle phase and a continuing increase.
+If polygenic resistance is shown in future to be more of a problem then these results could be compared to those of quantitative genetic models.
+
+Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
+The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
+holding most inputs constant at intermediate values while modifying single inputs in isolation).
+There is the potential that the system may behave differently in the field than we describe here.
+However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
+This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
+
+In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers to investigate model behaviour in areas of parameter space we have not shown.
+The user-interface allows inputs to be set for two scenarios and results to be viewed side by side.
+For example you can set a higher insecticide effectivenesses in one scenario to see how this favours mixtures over sequences (Fig 2) or increase the exposure levels to see how this favours sequences.
+
+There are other model behaviours that can be changed by modifying input files.
+For example, here we assume that insecticide interaction in a mixture is multiplicative, e.g.
+if the probability of surviving exposure to insecticide A alone is 0.3 and of surviving insecticide B alone is 0.2, then the probability of surviving exposure to a mixture of A and B is 0.3*0.2 = 0.06.
+This is consistent with recent work on aphids [@Taillebois2016], although that did also indicate some synergistic effects.
+Modifying input files would allow us to incorporate such synergistic or antagonistic effects of mixtures, i.e.
+that insecticides used together perform better or worse than when used separately.
+
+In addition to these uncertainties regarding resistance it is important to remember that our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
+Different insecticide use strategies, as well as affecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
+There are also potential knock-on effects of resistance itself; reduced vector transmission of disease (e.g.
+by reduced lifespan of the vector) or increased transmission (e.g.
+by increasing the susceptibility  of the vector to the disease) [@Rivero2010].
+This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
+Other operational factors also need to be considered.
+On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
+On the negative side a mixture may be more expensive.
+Other attributes of insecticides such as their repellency are also relevant.
+Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mixtures over sequences.
+However the reduced exposure resulting from repellency also reduces the population protection provided by the insecticide [@Birget2015].
+
+Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
+Of course, using no insecticide is likely to have serious consequences for mosquito and malaria abundance.
+The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
+
+#### Conclusions
+We have developed a flexible, modern model of the evolution of insecticide resistance that can that help us understand the mechanisms by which different insecticide application strategies may be favoured.
+It replicates previous modelling work and has the potential to investigate further operational strategies for established and new insecticides.
+For assessing the relative benefits of mixtures and sequences we have demonstrated that mixtures are likely to be favoured by high insecticide effectiveness, low exposure and high costs of resistance.
+Similar processes will occur in other strategies such as rotations and mosaics.
+This accessible modelling approach can help to promote discussion and exploration of the likely effect on resistance evolution of different insecticide use strategies in tandem with the collection of empirical data to refine and test predictions.
 

### 2017-09-11 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 pre-submission edits

@@
 output: word_document
-author: "Andy South and Ian Hastings"
-date: "`r Sys.Date()`"
+#author: "Andy South and Ian Hastings"
+#date: "`r Sys.Date()`"
 bibliography: library.bib
@@
 
-```{r, eval=FALSE, include=FALSE}
-#!!!! paper protocol
-#!!!! submission MS, knit to word_document, YAML, this chunk eval=FALSE, 
-#!!!!  in WORD: delete figs, add fig numbers to fig legends, set double space, add line numbers 
+Andy South * and Ian M. Hastings
+
+Department of Parasitology, Liverpool School of Tropical Medicine, Liverpool L3 5QA, UK.
+
+southandy@gmail.com
+
+ian.hastings@lstmed.ac.uk
+
+\* Corresponding author
+
+THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome
+\pagebreak
+
+```{r, eval=TRUE, include=FALSE}
+#!!!! PAPER PROTOCOL
+#!!!! submission MS, knit to word_document 
+#!!!!  in WORD: delete figs, set double space, add line numbers 
+#!!!!  (spacing & line numbers work for pdf output but not for word)
 #!!!! submission figs, set this chunk to eval=TRUE
@@
 
-### version 14
-
-Target : Malaria journal
-
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome
-
 ## Abstract
@@
 ## Background
-
 Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
@@
 ## Methods
-
 The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
@@
 For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
-Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Figure 1).
+Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Fig 1).
 
@@
 ## Results
-
 We start by describing the results of model runs with a single insecticide which are more straightforward and move on to describing results of using two insecticides in sequences and mixtures.
@@
 These effects of inputs on single insecticide use are summarised in Table 2.
+
+**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
+
+Parameter to increase  | effect on resistance evolution | Mechanism
+----------------------- | -------- | ---------------------------------------------
+1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
+2. Exposure | faster | reduced fitness of SS overall
+3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
+4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
+5. Frequency  | faster | less change needed to reach resistance threshold
+6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
 
@@
 #### Insecticide exposure and effectiveness
-
 Increasing exposure (the proportion of insects that come into contact with the insecticides) (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy.
@@
 The mechanisms identified in the tables are covered further in the discussion.
-
-#### Linkage disequilibrium
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
-
-## Discussion
-Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
-This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
-We start by explaining the evolution of resistance to a single insecticide, which is relatively straightforward, and move on to explaining the effect of two insecticides in sequences and mixtures.
-
-When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 4B).
-
-These mechanisms, explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
-Referring to figure 1 can help explain these responses.
-Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
-Cost acts by decreasing the fitness of RR and SR among those mosquitoes  not exposed.
-
-#### Two insecticides
-Responses to changing input parameters are different when two insecticides are used either in sequence or in mixtures as compared to when used alone.
-These differences can help us to understand the dynamics of resistance evolution.
-In each panel of Figures 5-9 the solid lines, indicating that an insecticide is being used at the same time as another in a mixture, are always shallower than the same colour dashed line indicating the insecticide is being used in isolation as part of a sequence.
-The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Because resistant individuals are being killed, the selection pressure for that resistance is lower.
-Thus each insecticide reduces the selection pressure for resistance to the other and could be said to 'protect' the other.
-This protection means that evolution of resistance to one insecticide is always slower when it is used in a mixture than when used alone.
-This protection has been termed multiple intra-generational killing [@Consortium2013] because individuals have the potential to be killed by more than one insecticide.
-
-Whilst this protection ensures that resistance to each insecticide evolves more slowly in a mixture it does not guarantee that a mixture strategy is preferable  to a sequence once the resistance to both insecticides is taken into account.
-In a mixture, selection pressure is applied by both insecticides at the same time, whereas in a sequence there is no selection for the insecticide not being used.
-Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
-
-This advantage of sequence over mixtures can be removed by increasing the effectiveness of one of the insecticides.
-Increasing the effectiveness of one insecticide (Fig 5B) decreases the total time for the sequence from 120 to 90 and increases that for the mixture from 90 to 115 thus changing the order and favouring the mixture.
-The mechanism for this change can be seen by comparing the shape of the curves in Figs 5A & B.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide reaches the resistance threshold of 50% it kills fewer individuals that are resistant to the second insecticide, it loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-In both cases increasing the effectiveness of insecticides increases their protective effect when used in mixtures and favours mixtures over sequences.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide (the red curve in this case).
-The more effective insecticide prompts both a) higher selection pressure for its own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-The more effective insecticide has a faster evolution of resistance than the less effective one but this is not caused by the presence of the less effective one.
-(Indeed the presence of the less effective one is still slowing the evolution of resistance to the more effective one, as indicated by the red solid curve for use in the mixture being shallower than the red dashed curve for use alone as part of the sequence).
-
-#### Difference between exposure and effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us and we are keen to explore it in more detail.
-The most likely explanation is that increasing exposure increases overall selection pressure and lowers time to resistance, whereas increasing effectiveness increases mutual protection of the insecticides in a mixture and lowers the speed at which resistance evolves.
-
-#### Costs of resistance
-Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@ffrench-Constant2017].
-Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 9).
-
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-It makes sense that the protective effect of the other insecticide in a mixture combines with the cost of resistance to reduce selection pressure more than in a sequence.
-
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
-Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
-Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
-Also note that if the costs are set even higher than we set them here (e.g.
-above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-However, we do not expect such high costs to occur in operational settings.
-
-Declines in resistance frequencies are likely to occur when there are costs of resistance and the insecticide is not being used [@ffrench-Constant2017].
-Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where the first insecticide becomes operationally useful again.
-Consequently, a repeated sequential strategy could offer benefits over mixtures not considered here.
-To quantify these benefits one would need to use a more detailed metric than the time-to-resistance we have used.
-We quantified the number of generations until resistance thresholds are first reached.
-Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a threshold.
-
-#### Consistency with previous work 
-The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
-A recent review [@Consortium2013] found that 8 out of 10 empirical studies and 11 out of 14 modelling studies favoured mixtures over sequences.
-In the 3 other modelling studies the relative performance of mixtures and sequences were dependent on the values of other inputs as we also have shown.
-
-The review concluded that the advantage of mixtures is greatest when seven other conditions are satisfied [@Consortium2013] : i) rare starting resistance , ii) independent loci (i.e.
-no cross resistance), iii) high recombination, iv) high susceptible mortality, v) resistance is recessive, vi) similar persistence of insecticides, and vii) some of population remains untreated.
-Their points iv) and vii) agree with our results that mixtures are favoured by high effectiveness and low exposure.
-Some of the other conditions, notably i) the starting frequency of resistance and v) the dominance of resistance did not alter whether mixtures or sequences were favoured in our model [@Levick2017].
-
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
-
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
-Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
-
-One aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
-
-Poor implementation would likely cause lower effectiveness of, and exposure to, insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [Bourguet2000][@Barbosa2011] [@FAO2012] see Fig 1 of [@Levick2017].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that higher effectiveness favours mixtures, higher exposure favours sequences and higher dominance favours neither (Table 4).
-More research into how the mortality of different genotypes (susceptible, resistant and heterozygous) is affected by poor application rates or insecticide degradation over time would help make this clearer.
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the development of resistance.
-
-#### Caveats
-There is much evidence of the involvement of multiple genes in resistance mechanisms.
-Our model, in common with most previous ones, assumes a single ‘resistance’ gene per insecticide.
-In most cases we believe this is reasonable.
-Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Ffrench-Constant2004] [@Ffrench-Constant2013].
-Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
-To represent polygenic resistance a different 'quantitative genetic' approach would be required.
-A quantitative genetic model would likely lead to a more rapid initial evolution of resistance, a slower middle phase and a continuing increase.
-If polygenic resistance is shown in future to be more of a problem then these results could be compared to those of quantitative genetic models.
-
-Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
-The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-There is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-
-In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers to investigate model behaviour in areas of parameter space we have not shown.
-The user-interface allows inputs to be set for two scenarios and results to be viewed side by side.
-For example you can set a higher insecticide effectivenesses in one scenario to see how this favours mixtures over sequences (Fig 2) or increase the exposure levels to see how this favours sequences.
-
-There are other model behaviours that can be changed by modifying input files.
-For example, here we assume that insecticide interaction in a mixture is multiplicative, e.g.
-if the probability of surviving exposure to insecticide A alone is 0.3 and of surviving insecticide B alone is 0.2, then the probability of surviving exposure to a mixture of A and B is 0.3*0.2 = 0.06.
-This is consistent with recent work on aphids [@Taillebois2016], although that did also indicate some synergistic effects.
-Modifying input files would allow us to incorporate such synergistic or antagonistic effects of mixtures, i.e.
-that insecticides used together perform better or worse than when used separately.
-
-In addition to these uncertainties regarding resistance it is important to remember that our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as affecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
-There are also potential knock-on effects of resistance itself; reduced vector transmission of disease (e.g.
-by reduced lifespan of the vector) or increased transmission (e.g.
-by increasing the susceptibility  of the vector to the disease) [@Rivero2010].
-This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other operational factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture may be more expensive.
-Other attributes of insecticides such as their repellency are also relevant.
-Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mixtures over sequences.
-However the reduced exposure resulting from repellency also reduces the population protection provided by the insecticide [@Birget2015].
-
-Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Of course, using no insecticide is likely to have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
-
-#### Conclusions
-We have developed a flexible, modern model of the evolution of insecticide resistance that can that help us understand the mechanisms by which different insecticide application strategies may be favoured.
-It replicates previous modelling work and has the potential to investigate further operational strategies for established and new insecticides.
-For assessing the relative benefits of mixtures and sequences we have demonstrated that mixtures are likely to be favoured by high insecticide effectiveness, low exposure and high costs of resistance.
-Similar processes will occur in other strategies such as rotations and mosaics.
-This accessible modelling approach can help to promote discussion and exploration of the likely effect on resistance evolution of different insecticide use strategies in tandem with the collection of empirical data to refine and test predictions.
-
-### Declarations
-TODO see https://malariajournal.biomedcentral.com/submission-guidelines/preparing-your-manuscript/research-article
-
-##### Ethics approval and consent to participate
-Not applicable
-
-##### Consent for publication
-Not applicable
-
-##### Availability of data and material
-
-##### Competing interests
-
-##### Funding
-
-##### Authors' contributions
-AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
-IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
-
-##### Acknowledgements
-Thank you to Bethany Levick who developed the first version of the model code.
-
-\pagebreak
-
-**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
-2. Exposure | faster | reduced fitness of SS overall
-3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
-4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
 
@@
 6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
-
-\pagebreak
 
@@
 
+#### Linkage disequilibrium
+It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
+Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
+In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
+
+## Discussion
+Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
+This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
+We start by explaining the evolution of resistance to a single insecticide, which is relatively straightforward, and move on to explaining the effect of two insecticides in sequences and mixtures.
+
+When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 4B).
+
+These mechanisms, explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
+Referring to figure 1 can help explain these responses.
+Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
+Cost acts by decreasing the fitness of RR and SR among those mosquitoes  not exposed.
+
+#### Two insecticides
+Responses to changing input parameters are different when two insecticides are used either in sequence or in mixtures as compared to when used alone.
+These differences can help us to understand the dynamics of resistance evolution.
+In each panel of Figures 5-9 the solid lines, indicating that an insecticide is being used at the same time as another in a mixture, are always shallower than the same colour dashed line indicating the insecticide is being used in isolation as part of a sequence.
+The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
+Because resistant individuals are being killed, the selection pressure for that resistance is lower.
+Thus each insecticide reduces the selection pressure for resistance to the other and could be said to 'protect' the other.
+This protection means that evolution of resistance to one insecticide is always slower when it is used in a mixture than when used alone.
+This protection has been termed multiple intra-generational killing [@Consortium2013] because individuals have the potential to be killed by more than one insecticide.
+
+Whilst this protection ensures that resistance to each insecticide evolves more slowly in a mixture it does not guarantee that a mixture strategy is preferable  to a sequence once the resistance to both insecticides is taken into account.
+In a mixture, selection pressure is applied by both insecticides at the same time, whereas in a sequence there is no selection for the insecticide not being used.
+Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
+In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
+
+This advantage of sequence over mixtures can be removed by increasing the effectiveness of one of the insecticides.
+Increasing the effectiveness of one insecticide (Fig 5B) decreases the total time for the sequence from 120 to 90 and increases that for the mixture from 90 to 115 thus changing the order and favouring the mixture.
+The mechanism for this change can be seen by comparing the shape of the curves in Figs 5A & B.
+In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
+However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
+Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
+Once the first insecticide reaches the resistance threshold of 50% it kills fewer individuals that are resistant to the second insecticide, it loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
+A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
+The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
+In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
+In both cases increasing the effectiveness of insecticides increases their protective effect when used in mixtures and favours mixtures over sequences.
+
+Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide (the red curve in this case).
+The more effective insecticide prompts both a) higher selection pressure for its own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
+The more effective insecticide has a faster evolution of resistance than the less effective one but this is not caused by the presence of the less effective one.
+(Indeed the presence of the less effective one is still slowing the evolution of resistance to the more effective one, as indicated by the red solid curve for use in the mixture being shallower than the red dashed curve for use alone as part of the sequence).
+
+#### Difference between exposure and effectiveness
+It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
+Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
+This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
+This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us and we are keen to explore it in more detail.
+The most likely explanation is that increasing exposure increases overall selection pressure and lowers time to resistance, whereas increasing effectiveness increases mutual protection of the insecticides in a mixture and lowers the speed at which resistance evolves.
+
+#### Costs of resistance
+Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@ffrench-Constant2017].
+Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 9).
+
+Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
+It makes sense that the protective effect of the other insecticide in a mixture combines with the cost of resistance to reduce selection pressure more than in a sequence.
+
+The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
+This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
+Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
+Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
+Also note that if the costs are set even higher than we set them here (e.g.
+above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
+However, we do not expect such high costs to occur in operational settings.
+
+Declines in resistance frequencies are likely to occur when there are costs of resistance and the insecticide is not being used [@ffrench-Constant2017].
+Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where the first insecticide becomes operationally useful again.
+Consequently, a repeated sequential strategy could offer benefits over mixtures not considered here.
+To quantify these benefits one would need to use a more detailed metric than the time-to-resistance we have used.
+We quantified the number of generations until resistance thresholds are first reached.
+Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a threshold.
+
+#### Consistency with previous work 
+The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
+A recent review [@Consortium2013] found that 8 out of 10 empirical studies and 11 out of 14 modelling studies favoured mixtures over sequences.
+In the 3 other modelling studies the relative performance of mixtures and sequences were dependent on the values of other inputs as we also have shown.
+
+The review concluded that the advantage of mixtures is greatest when seven other conditions are satisfied [@Consortium2013] : i) rare starting resistance , ii) independent loci (i.e.
+no cross resistance), iii) high recombination, iv) high susceptible mortality, v) resistance is recessive, vi) similar persistence of insecticides, and vii) some of population remains untreated.
+Their points iv) and vii) agree with our results that mixtures are favoured by high effectiveness and low exposure.
+Some of the other conditions, notably i) the starting frequency of resistance and v) the dominance of resistance did not alter whether mixtures or sequences were favoured in our model [@Levick2017].
+
+Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
+One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
+That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
+This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
+This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
+see Fig 6D and Table 4).
+
+A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
+This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
+Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
+Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
+
+One aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
+Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
+This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
+@Roush1998].
+This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
+e.g.
+from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
+These include: ...
+- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
+
+Poor implementation would likely cause lower effectiveness of, and exposure to, insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
+The explanation for this apparent difference is in the effect of dominance.
+Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [Bourguet2000][@Barbosa2011] [@FAO2012] see Fig 1 of [@Levick2017].
+Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
+So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
+Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that higher effectiveness favours mixtures, higher exposure favours sequences and higher dominance favours neither (Table 4).
+More research into how the mortality of different genotypes (susceptible, resistant and heterozygous) is affected by poor application rates or insecticide degradation over time would help make this clearer.
+
+Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
+It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
+A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
+Mosquitoes with different behaviours e.g.
+feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
+This lower exposure could favour mixtures over sequences as we have indicated.
+In both cases our modelling approach could be used to investigate implications for the development of resistance.
+
+#### Caveats
+There is much evidence of the involvement of multiple genes in resistance mechanisms.
+Our model, in common with most previous ones, assumes a single ‘resistance’ gene per insecticide.
+In most cases we believe this is reasonable.
+Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Ffrench-Constant2004] [@Ffrench-Constant2013].
+Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
+To represent polygenic resistance a different 'quantitative genetic' approach would be required.
+A quantitative genetic model would likely lead to a more rapid initial evolution of resistance, a slower middle phase and a continuing increase.
+If polygenic resistance is shown in future to be more of a problem then these results could be compared to those of quantitative genetic models.
+
+Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
+The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
+holding most inputs constant at intermediate values while modifying single inputs in isolation).
+There is the potential that the system may behave differently in the field than we describe here.
+However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
+This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
+
+In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers to investigate model behaviour in areas of parameter space we have not shown.
+The user-interface allows inputs to be set for two scenarios and results to be viewed side by side.
+For example you can set a higher insecticide effectivenesses in one scenario to see how this favours mixtures over sequences (Fig 2) or increase the exposure levels to see how this favours sequences.
+
+There are other model behaviours that can be changed by modifying input files.
+For example, here we assume that insecticide interaction in a mixture is multiplicative, e.g.
+if the probability of surviving exposure to insecticide A alone is 0.3 and of surviving insecticide B alone is 0.2, then the probability of surviving exposure to a mixture of A and B is 0.3*0.2 = 0.06.
+This is consistent with recent work on aphids [@Taillebois2016], although that did also indicate some synergistic effects.
+Modifying input files would allow us to incorporate such synergistic or antagonistic effects of mixtures, i.e.
+that insecticides used together perform better or worse than when used separately.
+
+In addition to these uncertainties regarding resistance it is important to remember that our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
+Different insecticide use strategies, as well as affecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
+There are also potential knock-on effects of resistance itself; reduced vector transmission of disease (e.g.
+by reduced lifespan of the vector) or increased transmission (e.g.
+by increasing the susceptibility  of the vector to the disease) [@Rivero2010].
+This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
+Other operational factors also need to be considered.
+On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
+On the negative side a mixture may be more expensive.
+Other attributes of insecticides such as their repellency are also relevant.
+Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mixtures over sequences.
+However the reduced exposure resulting from repellency also reduces the population protection provided by the insecticide [@Birget2015].
+
+Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
+Of course, using no insecticide is likely to have serious consequences for mosquito and malaria abundance.
+The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
+
+#### Conclusions
+We have developed a flexible, modern model of the evolution of insecticide resistance that can that help us understand the mechanisms by which different insecticide application strategies may be favoured.
+It replicates previous modelling work and has the potential to investigate further operational strategies for established and new insecticides.
+For assessing the relative benefits of mixtures and sequences we have demonstrated that mixtures are likely to be favoured by high insecticide effectiveness, low exposure and high costs of resistance.
+Similar processes will occur in other strategies such as rotations and mosaics.
+This accessible modelling approach can help to promote discussion and exploration of the likely effect on resistance evolution of different insecticide use strategies in tandem with the collection of empirical data to refine and test predictions.
+
+### Declarations
+
+##### Ethics approval and consent to participate
+Not applicable
+
+##### Consent for publication
+Not applicable
+
+##### Availability of data and material
+Project name: resistance
+Project home page: https://github.com/AndySouth/resistance
+Archived version: 10.5281/zenodo.889012
+Operating system: Platform independent
+Programming language: R
+License: GNU General Public License v3.0
+
+##### Competing interests
+AS received a small grant from Bayer S.A.S. to support a few months work and attendance of the American Society of Tropical Medicine and Hygiene Annual Meeting.
+
+##### Funding
+This work was supported by: Bill and Melinda Gates Foundation, grant #37999.01 to IMH; and Wellcome Trust Institutional Strategic Strengthening fund grant #105620/Z/14/Z to Liverpool School of Tropical Medicine; sub-funding to IMH; and a small grant from Bayer S.A.S.
+to AS.
+
+##### Authors' contributions
+AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
+IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
+
+##### Acknowledgements
+Thank you to Bethany Levick who developed the first version of the model code.
+
 \pagebreak
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="The effect of model inputs on the fitness of genotypes for a single insecticide.
+```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1.
+The effect of model inputs on the fitness of genotypes for a single insecticide.
 Fitness is shown on the y-axis and the different genotypes (SS, SR, RR) on the x axis.
@@
 
-![Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
+![Figure 2.
+Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
 The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
+```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 3.
+Single insecticide effects of effectiveness, exposure, dominance and resistance restoration on resistance frequency.
 A.
@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Single insecticide use and the effect of inputs on changing resistance frequency over time.
+```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 4.
+Single insecticide effects of starting frequency and cost of resistance on resistance frequency.
 A.
@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
+```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 5.
+Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
 Exposure to the insecticides is increased from row 1 (A and B) to row 2 (C and D).
@@
 
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
+```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 6.
+Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
 Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
@@
 
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
+```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 7.
+Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
 Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
@@
 
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
+```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 8.
+Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
 Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
@@
 
-```{r Fig 9 cost of 1 resistance, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Influence of cost of resistance on time-to-resistance for mixtures and sequences.
+```{r Fig9, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 9.
+Influence of cost of resistance on time-to-resistance for mixtures and sequences.
 Row 1 (A and B) same as Fig.

### 2017-09-21 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 tweaking

@@
 These assumptions are consistent with previous modelling work.
-In support of the first assumption insecticide resistance has been termed 'pre-adaptive' meaning that resistance alleles are present in very low numbers even prior to exposure to novel insecticides [@Ffrench-Constant2013][@Liu2015] as has been documented in blowflies [Hartley2006].
+In support of the first assumption insecticide resistance has been termed 'pre-adaptive' meaning that resistance alleles are present in very low numbers even prior to exposure to novel insecticides [@Ffrench-Constant2013][@Liu2015] as has been documented in blowflies [@Hartley2006].
 In support of using a single gene representation of resistance evidence suggests that whilst polygenic resistance is common, control failures, particularly those due to target site resistance, are mostly due to single major genes [@Ffrench-Constant2013].
@@
 
-**Table 1. Parameters inflencing the development of insecticide resistance**
+**Table 1. Parameters influencing the development of insecticide resistance**
 
@@
 Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | faster | reduced fitness of SS in presence of insecticide
-2. Exposure | faster | reduced fitness of SS overall
-3. Dominance of restoration | faster | increased fitness of SR in presence of insecticide
-4. Resistance restoration | faster | increased fitness of RR in presence of insecticide
+----------------------- | ------ | ---------------------------------------------
+1. Effectiveness | faster | reduced fitness of SS & SR exposed to insecticide
+2. Exposure | faster | reduced fitness of SS & SR overall
+3. Dominance of restoration | faster | increased fitness of SR exposed to insecticide
+4. Resistance restoration | faster | increased fitness of RR & SR exposed to insecticide
 5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of RR in absence of insecticide
+6. Cost of resistance | slower | reduced fitness of RR & SR not exposed to insecticide
 
@@
 ----------------------- | ---------- | ---------------------------------------------
-1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure from the other
+1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure for the other
 2.
-Exposure | faster (but less than for single) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
+Exposure | faster (but less than for insecticide used alone) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
 However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
@@
 When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 4B).
-
-These mechanisms, explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
+Mechanisms explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
 Referring to figure 1 can help explain these responses.
@@
 #### Two insecticides
-Responses to changing input parameters are different when two insecticides are used either in sequence or in mixtures as compared to when used alone.
+Responses to input parameters are different when two insecticides are used either in sequence or in mixtures as compared to when used alone.
 These differences can help us to understand the dynamics of resistance evolution.
@@
 The more effective insecticide has a faster evolution of resistance than the less effective one but this is not caused by the presence of the less effective one.
-(Indeed the presence of the less effective one is still slowing the evolution of resistance to the more effective one, as indicated by the red solid curve for use in the mixture being shallower than the red dashed curve for use alone as part of the sequence).
+Indeed the presence of the less effective one is still slowing the evolution of resistance to the more effective one, as indicated by the red solid curve for use in the mixture being shallower than the red dashed curve for use alone as part of the sequence.
 
@@
 #### Costs of resistance
-Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@ffrench-Constant2017].
+Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@Ffrench-Constant2017].
 Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 9).
@@
 The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes becomes greater.
+This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes also increases.
 Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
@@
 
-Declines in resistance frequencies are likely to occur when there are costs of resistance and the insecticide is not being used [@ffrench-Constant2017].
+Declines in resistance frequencies are likely to occur when there are costs of resistance and the insecticide is not being used [@Ffrench-Constant2017].
 Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where the first insecticide becomes operationally useful again.
@@
 We quantified the number of generations until resistance thresholds are first reached.
-Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a threshold.
+Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a resistance threshold.
 
@@
 
-A subsequent model [@Roush1998] lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
+A subsequent model lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
 This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
 Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Our result showing that decreasing exposure favours mixtures over sequences is also consistent with figure 3 from that paper [@Roush1998].
-
+Consistent with our results [@Roush1998] also showed (Figure 3) that decreasing exposure favours mixtures over sequences.
+
+#### Poor implementation
 One aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
@@
 The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (RS) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [Bourguet2000][@Barbosa2011] [@FAO2012] see Fig 1 of [@Levick2017].
+Poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].
 Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
@@
 In most cases we believe this is reasonable.
-Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Ffrench-Constant2004] [@Ffrench-Constant2013].
+Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Roush1989][@Ffrench-Constant2004][@Ffrench-Constant2013].
 Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
-To represent polygenic resistance a different 'quantitative genetic' approach would be required.
-A quantitative genetic model would likely lead to a more rapid initial evolution of resistance, a slower middle phase and a continuing increase.
-If polygenic resistance is shown in future to be more of a problem then these results could be compared to those of quantitative genetic models.
+To represent polygenic resistance a different 'quantitative genetic' approach would be required [@Via1986][@Firko1990].
+We are only aware of preliminary published work applying a quantitative genetic approach to assessing insecticide use strategies in the context of polygenic resistance [@Via1986][@Gardner1998].
+One suggestion is that evolution of resistance to one insecticide in a mixture could be associated with a decrease in tolerance to the other [@Via1986].
+If it becomes clearer that polygenic resistance is an issue then we would welcome a comparison of the results presented here to those of quantitative genetic models.
 

### 2017-09-25 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper2 pre-sub tweaks

@@
 
-_must be 350 words, currently 348_
-
 ##### Background
@@
 We provide a concise, accessible description of a flexible model of the evolution of insecticide resistance.
-We use the model to develop a mechanistic picture the evolution of insecticide resistance and how it is likely to respond to potential insecticide use strategies.
+We use the model to develop a mechanistic picture of the evolution of insecticide resistance and how it is likely to respond to potential insecticide use strategies.
 The aim is to reach an audience unlikely to read a more detailed modelling paper.
@@
 ##### Results
-We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about selection pressure.
+We use the model to demonstrate the evolution of resistance under different scenarios and how this fits with intuitive reasoning about selection pressure.
 We show that using an insecticide in a mixture, relative to alone, always prompts slower evolution of resistance to that insecticide.
-However, when the resistance to both insecticides is considered, resistance thresholds may be reached later for a sequence relative to a mixture.
+However, when resistance to both insecticides is considered, resistance thresholds may be reached later for a sequence relative to a mixture.
 Increasing the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on favouring a mixture over a sequence because one highly effective insecticide provides more protection to another in a mixture.
@@
 ## Background
-Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
+Insecticide resistance is a problem for malaria control[@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
 Malaria alone still results in hundreds of thousands of deaths per year.
@@
 Efforts are under way to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
-The Innovative Vector Control Consortium (IVCC) was set up in 2005 to work with industry in the development of new vector control tools and particularly new insecticides with new modes of action to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
+The Innovative Vector Control Consortium (IVCC) was set up in 2005 to work with industry to develop new vector control tools and particularly new insecticides with new modes of action to address insecticide resistance in disease transmitting mosquitoes [@Hemingway2006][@IVCC2016].
 Three new insecticides are now in development [@IVCC2016] and likely to be available after 2020 [@Ranson2016].
@@
 Modelling studies have investigated the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
-[@Curtis1985][@Mani1985][@Roush1989] but much of the work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
+[@Curtis1985][@Mani1985][@Roush1989] but much of the work was done more than 25 years ago and there remained some confusion about the results [@Levick2017].
 In an earlier paper [@Levick2017] we described the technical details of a flexible model used to investigate the relative benefits of mixtures and sequences.
-Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance can be expected to change in different scenarios.
-This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012][@Consortium2013].
+For that paper we investigated thousands of scenarios to explore potential outcomes.
+In contrast here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance can be expected to respond.
+This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies and extend existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012][@Consortium2013].
 
 Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
-We assume both that genes conferring resistance are already present in the population and that resistance to each insecticide is coded by a single gene.
-These assumptions are consistent with previous modelling work.
-In support of the first assumption insecticide resistance has been termed 'pre-adaptive' meaning that resistance alleles are present in very low numbers even prior to exposure to novel insecticides [@Ffrench-Constant2013][@Liu2015] as has been documented in blowflies [@Hartley2006].
+We assume that genes conferring resistance are already present in the population and that resistance to each insecticide is coded by a single gene.
+These assumptions are consistent with previous modelling work [@Consortium2013].
+In support of the first assumption insecticide resistance has been termed 'pre-adaptive' meaning that resistance alleles are present in very low numbers, most likely as a mutation-selection balance, even prior to exposure to novel insecticides [@Ffrench-Constant2013][@Liu2015] as has been documented in blowflies [@Hartley2006].
 In support of using a single gene representation of resistance evidence suggests that whilst polygenic resistance is common, control failures, particularly those due to target site resistance, are mostly due to single major genes [@Ffrench-Constant2013].
@@
 Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
-The fitness of the SS genotype in absence of the insecticide is set at a maximum value of 1.
-All other fitnesses are less, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
+The fitness of the SS genotype in absence of the insecticide is used as the reference genotype and assigned a value of 1.
+All other fitnesses are assumed to be less than 1, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
+These inputs include the effectiveness of the insecticide against SS genotypes, the dominance of the resistance allele and the ability of resistance to restore fitness of the RR genotype in the presence of the insecticide (see Table 1).
 
@@
 For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS and resistance restoration 'restores' a portion of the fitness for RR.
-A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
+A value of 1 for resistance restoration would bring the fitness of RR back to 1, equivalent to the unexposed SS and meaning the RR genotype is completely unaffected by the insecticide.
+A value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
 We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
 Cost of resistance determines the fitness of RR in the portion of the population not exposed to the insecticide.
@@
 
-The simulation proceeds through generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
+The simulation proceeds through non-overlapping generations.
+In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness, the latter depending on their absolute fitness (Fig 1) and level of exposure.
 This acts to make the fitter alleles more common over time.
 Separate sexes and standard sexual reproduction with recombination is included.
-There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that here.
+There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that parameter here.
 
@@
 
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an on-line user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
+A major aim of this paper is to us modern techniques to make the modelling accessible to a wider community.
+The model is implemented in R [@RCoreTeam2016], the code is open-source and hosted on Github [@South2017] including the code to generate both the figures and text of this paper.
+We provide an on-line user interface [@South2017a] (Fig 2) accessing the same code and enabling readers with no coding experience to change inputs and run the model themselves.
 
@@
 This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
-Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that less change is required to reach the threshold.
+Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that a smaller cahnge in frequency is required to reach the threshold.
 
@@
 #### Caveats
-There is much evidence of the involvement of multiple genes in resistance mechanisms.
-Our model, in common with most previous ones, assumes a single ‘resistance’ gene per insecticide.
+There is evidence of the involvement of multiple genes in resistance mechanisms[@Ffrench-Constant2004][@Ffrench-Constant2013].
+Our model, in common with almost all previous ones [@Bourguet2010], assumes a single ‘resistance’ gene per insecticide.
 In most cases we believe this is reasonable.
@@
 To represent polygenic resistance a different 'quantitative genetic' approach would be required [@Via1986][@Firko1990].
+A recent review of 187 models of the evolution of resistance [@Bourguet2010] did not find any that represented quantitative multiple gene resistance.
 We are only aware of preliminary published work applying a quantitative genetic approach to assessing insecticide use strategies in the context of polygenic resistance [@Via1986][@Gardner1998].
-One suggestion is that evolution of resistance to one insecticide in a mixture could be associated with a decrease in tolerance to the other [@Via1986].
-If it becomes clearer that polygenic resistance is an issue then we would welcome a comparison of the results presented here to those of quantitative genetic models.
+If it becomes clearer that in some cases polygenic resistance is more of an issue than previously thought then there will be a need for appropriate quantitative genetic models.
+We would be keen to contribute to that development.
 
 Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
-The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-There is the potential that the system may behave differently in the field than we describe here.
+Our model is able to represent much of that uncertainty, yet in this paper we restrict our explorations to a limited region of parameter space.
+Here we hold most inputs constant at intermediate values while modifying single inputs in isolation.
+This is a deliberate tactic to develop understanding and communicate the key mechanisms.
+There is the potential that the system may behave differently in the field.
 However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
@@
 
-In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers to investigate model behaviour in areas of parameter space we have not shown.
+In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers with no coding experience to investigate model behaviour in areas of parameter space we have not shown.
 The user-interface allows inputs to be set for two scenarios and results to be viewed side by side.
@@
 
-There are other model behaviours that can be changed by modifying input files.
+There are other model behaviours that can be changed by modifying model inputs within the computer code.
 For example, here we assume that insecticide interaction in a mixture is multiplicative, e.g.
@@
 This is consistent with recent work on aphids [@Taillebois2016], although that did also indicate some synergistic effects.
-Modifying input files would allow us to incorporate such synergistic or antagonistic effects of mixtures, i.e.
+Modifying model inputs would allow us to incorporate such synergistic or antagonistic effects of mixtures, i.e.
 that insecticides used together perform better or worse than when used separately.
@@
 ##### Funding
-This work was supported by: Bill and Melinda Gates Foundation, grant #37999.01 to IMH; and Wellcome Trust Institutional Strategic Strengthening fund grant #105620/Z/14/Z to Liverpool School of Tropical Medicine; sub-funding to IMH; and a small grant from Bayer S.A.S.
-to AS.
+This work was supported by: Bill and Melinda Gates Foundation, grant #37999.01 to IMH; and Wellcome Trust Institutional Strategic Strengthening fund grant #105620/Z/14/Z to Liverpool School of Tropical Medicine; sub-funding to IMH; a small grant from Bayer S.A.S.
+to AS; and IVCC funding to IMH.
 

### 2017-09-27 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
notes and paper2 tweaks

@@
 Our model provides an accessible description of the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
-We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/MixSeqResist1).
+We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/resistmob2).
 These tools can contribute to an improved discussion about operational decisions in insecticide resistance management.
@@
 3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4.
-Resistance restoration | neither | As for dominance of restoration.
-Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
-5.
-Frequency  | neither | As for dominance and resistance restoration.
-Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
+4. Resistance restoration | neither | As for dominance of restoration. 
+5. Frequency  | neither | As for dominance and resistance restoration. 
 6.
@@
 ![Figure 2.
-Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
+Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2>.
 The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.

### 2017-12-13 inst/documents/sensiAnPaper1All.Rmd
tidy

@@
 
-### All code for the sensitivity analysis for the first paper
+### code to start the sensitivity analysis for the first paper
+### main analyses of results is in : paper1_results_figs_slimmed_50_rr
 
@@
   resistPointsMix_1 <- findResistancePoints(listOutMix, locus='either')
-  #todo - to be comparable I think this should be for when resistance to BOTH insecticides is reached
 

### 2017-12-13 inst/documents/linkage_dis_2017.Rmd
quick look at linkage

---
  title: "linkage disequilibrium"
  output: pdf_document
  #output: word_document
  author: "Andy South"
  date: "`r Sys.Date()`"
---

### documenting whats happening with linkage disequilibrium  after review of paper2

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(resistance)
  library(ggplot2)
```

### fig 8D in paper2 with a new & old insecticide, mix slower

```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left'}

listOut <- runcurtis_f2( recomb_rate=0.5, max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, maxX = 150, labelMixSeqRatio = 1 )
```

### now plot the LD for that scenario
It's not obvious to me what this shows. 
Possible that only the mixture one is meaningfull. In the sequential cases resistance to the other insecticide probably remains at 0 and therefore LD is not meaningfull.

```{r, eval=TRUE, echo=FALSE, fig.height=8, fig.width=7, fig.show='hold', fig.align='left'}

#we are mostly interested in what happens in the first 80 generations, by that time the frequency of I1 has reached 0.5 both alone and in the mixture.

#in listOut from runcurtis_f2, 1=I1, 2=I2, 3=mix
par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
par(mfrow=c(3,1))

plotlinkage(listOut$results[[3]], plot_d = TRUE, max_gen_plot = 80, main = "mixture")
plotlinkage(listOut$results[[1]], plot_d = TRUE, max_gen_plot = 80, main = "insecticide1 alone")
plotlinkage(listOut$results[[2]], plot_d = TRUE, max_gen_plot = 80, main = "insecticide2 alone")

```

### 2017-12-15 inst/documents/paper2/field_data_in.Rmd
looking at getting field data in

---
title: "getting field data into insecticide resistance model"
# output: 
#   pdf_document:
#     keep_tex: true
#     fig_caption: true
#     latex_engine: pdflatex

output: word_document
#author: "Andy South and Ian Hastings"
#date: "`r Sys.Date()`"
bibliography: library.bib
csl: biomed-central.csl
fontsize: 11pt
spacing: double
# below worked in test_linenumbers.Rmd
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
---

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(resistance)
  #library(ggplot2)
```

### Section demonstrating getting field data into the model, to be added to Malaria Journal paper.

The main model inputs relating to properties of the mosquitoes and insecticides can be derived from data on the relative survival of the different genotypes (RR,RS,SS) Table x.

**Table x1. Calculating model inputs from field data.**

Parameter  | Calculation (where RRfit, RSfit, SSfit are fitnesses of genotypes)
---------------------- | ----------------------------------------------------
1. Effectiveness | 1 - SSfit in presence of insecticide
2. Exposure | estimated
3. Resistance restoration  | (RRfit-SSfit) / Effectiveness in presence of insecticide
4. Dominance of restoration | (RSfit-SSfit)/(RRfit-SSfit) in presence of insecticide
5. Frequency  | estimated
6. Cost of resistance  |  1 - RRfit in absence of insecticide
7. Dominance of cost  | (RSfit-RRfit)/(RRfit-SSfit) in absence of insecticide

Exposure, the proportion of mosquitoes that are exposed to the insecticide, is a property of the location and will depend on the use of nets or IRS and mosquito behaviour. The frequency of resistance alleles can be measured, although when at low levels a large number of mnosquitos would need to be sampled to detect low frequencies.

As an example of making these calculations we use data for Anopheles gambiae on pyrethroid (Etofenprox) resistance associated with the Kdr mutation [@Kolaczinski2000] and carbamate (Bendiocarb) resistance associated with the Ace1 mutation [@Essandoh2013].

In both cases data were provided as numbers alive and dead by genotype. We used these to calculate genotype-specific survival values that we assume are a good indication of fitness and from that the values of our model inputs (Table x2).

Insecticide/Mutation|survival RR|survival RS|survival SS|Effectiveness|Resistance restoration|Dominance of restoration
---|---|---|---|---|---|---
Pyrethroid/Kdr|0.84|0.65|0.57|0.43|0.63|0.30
Carbamate/Ace1|0.84|0.56|0.02|0.98|0.84|0.66

These input values can then be used in the model. In Figure x insecticide 1 and 2 are set to the pyrethroid (red) and carbamate (blue) respectively. Exposure is set to 0.5 and starting frequencies to 0.01.

```{r Fig10, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 10. Running the model using field data."}

# add side panels for labels
#layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
layout( matrix(c(1:2), byrow = TRUE), heights=c(1,0.15) )

#plot margins
par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)

runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.3 , h.RS2_0B = 0.66 , exposure = 0.5 , phi.SS1_A0 = 0.43 , phi.SS2_0B = 0.98 , rr_restoration_ins1 = 0.63 , rr_restoration_ins2 = 0.84 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, labelMixSeqRatio = 1 )

#adding legend in lower panel
#plot margins
par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)

plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')

legend( 'bottom', legend=c("pyrethroid alone   ", "pyrethroid in mix   ", "carbamate alone   ", "carbamate in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )

```

Resistance to the carbamate (blue) rises quickly because it has high effectiveness and dominance. The mixture performs better than the sequence because the carbamate with it's high effectiveness provides high protection to the pyrethroid.

#References

### 2017-12-18 inst/documents/paper2/fig1_revised.Rmd
revising fig1 for malaria journal

--- 2017-09-21 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
@@
 ---
-title: "Insecticide resistance evolution with mixtures and sequences : a model-based explanation."
+title: "getting field data into insecticide resistance model"
 # output: 
@@
 
-Andy South * and Ian M. Hastings
+```{r, eval=TRUE, include=FALSE}
 
-Department of Parasitology, Liverpool School of Tropical Medicine, Liverpool L3 5QA, UK.
-
-southandy@gmail.com
-
-ian.hastings@lstmed.ac.uk
-
-\* Corresponding author
-
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome
-\pagebreak
-
-```{r, eval=TRUE, include=FALSE}
-#!!!! PAPER PROTOCOL
-#!!!! submission MS, knit to word_document 
-#!!!!  in WORD: delete figs, set double space, add line numbers 
-#!!!!  (spacing & line numbers work for pdf output but not for word)
-#!!!! submission figs, set this chunk to eval=TRUE
-#!!!! previous drafts knitted to pdf with pdf_document in YAML and this chunk eval=FALSE
-
-# options to create final publication quality figures
-# causes the document to fail, but the figs do get stored in pap2figs/
+# options to create final publication quality figures stored in pap2figs/
 library(knitr)
@@
   library(resistance)
-  library(ggplot2)
+  #library(ggplot2)
 ```
-
-## Abstract
-
-_must be 350 words, currently 348_
-
-##### Background
-Insecticide resistance threatens effective vector control, especially for mosquitoes and malaria.
-To manage resistance, recommended insecticide use strategies include mixtures, sequences and rotations.
-New insecticides are being developed and there is an opportunity to develop use strategies that limit the evolution of further resistance in the short term.
-A 2013 review of modelling and empirical studies of resistance points to the advantages of mixtures.
-However, there is limited recent, accessible modelling work addressing the evolution of resistance under different operational strategies.
-We see an opportunity to improve the level of mechanistic understanding within the operational community of how insecticide resistance can be expected to evolve in response to different strategies.
-
-We provide a concise, accessible description of a flexible model of the evolution of insecticide resistance.
-We use the model to develop a mechanistic picture the evolution of insecticide resistance and how it is likely to respond to potential insecticide use strategies.
-The aim is to reach an audience unlikely to read a more detailed modelling paper.
-The model itself, as described here, represents two independent genes coding for resistance to two insecticides.
-This allows us to represent the use of insecticides in isolation, sequence and mixtures.
-
-##### Results
-We use the model to demonstrate the evolution of resistance under different inputs and how this fits with intuitive reasoning about selection pressure.
-We show that using an insecticide in a mixture, relative to alone, always prompts slower evolution of resistance to that insecticide.
-However, when the resistance to both insecticides is considered, resistance thresholds may be reached later for a sequence relative to a mixture.
-Increasing the ability of insecticides to kill susceptible mosquitoes (effectiveness), has the most influence on favouring a mixture over a sequence because one highly effective insecticide provides more protection to another in a mixture.
-
-##### Conclusions
-Our model provides an accessible description of the process of insecticide resistance evolution and how it is likely to respond to insecticide use.
-We provide a simple user-interface allowing further exploration (https://andysouth.shinyapps.io/MixSeqResist1).
-These tools can contribute to an improved discussion about operational decisions in insecticide resistance management.
-
-## Keywords
-insecticide resistance; public health; mosquitoes; vector-borne diseases; infectious diseases; malaria; population genetics
-
-## Background
-Insecticide resistance is a problem for malaria [@WHO2012][@Ranson2016][@Hemingway2016] other vector borne diseases [@IRAC2011] and agriculture [@FAO2012].
-Malaria alone still results in hundreds of thousands of deaths per year.
-Recent malaria control efforts have centred on treated bed nets and indoor residual spraying, both reliant on insecticides.
-Treated nets were recently estimated to contribute 68% and indoor residual spraying 13% to averting more than 500 million falciparum malaria cases between 2000 and 2015 [@Bhatt2015].
-A recent malaria transmission model [@Churcher2016] predicts that even low pyrethroid resistance is likely to increase malaria incidence in Africa by reducing the performance of bed nets.
-
-The WHO produced a Global Plan for Insecticide Resistance Management in malaria vectors (GPIRM)[@WHO2012] which includes recommendations on operational strategies for managing resistance including the use of insecticide mixtures when they become available.
-Efforts are under way to develop new insecticides that will be effective in the light of existing resistance and allow additional options within insecticide resistance management.
-The Innovative Vector Control Consortium (IVCC) was set up in 2005 to work with industry in the development of new vector control tools and particularly new insecticides with new modes of action to address insecticide resistance in malaria transmitting mosquitoes [@Hemingway2006][@IVCC2016].
-Three new insecticides are now in development [@IVCC2016] and likely to be available after 2020 [@Ranson2016].
-It is important that decisions about how best to use the new insecticides, with a clear mindset of delaying the evolution of resistance, are made before insecticides are released [@Hemingway2016].
-
-A recent comprehensive review of strategies to avoid resistance evolution across pesticides and drugs [@Consortium2013] concluded that mixtures (combination of molecules) are usually the best resistance management strategy.
-This was based on both empirical and modelling work.
-Modelling studies have investigated the evolution of insecticide resistance in insecticide mixtures including in a public health context e.g.
-[@Curtis1985][@Mani1985][@Roush1989] but much of the work was done more than 20 years ago and there remained some confusion about the results [@Levick2017].
-In an earlier paper [@Levick2017] we described the technical details of a flexible model used to investigate the relative benefits of mixtures and sequences.
-Here we provide an accessible summary of the model and use selected parameter values to describe mechanistically how the evolution of resistance can be expected to change in different scenarios.
-This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies extending existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012][@Consortium2013].
-
-Our modelling approach focuses on the change in frequency of a single resistance gene for each insecticide.
-We assume both that genes conferring resistance are already present in the population and that resistance to each insecticide is coded by a single gene.
-These assumptions are consistent with previous modelling work.
-In support of the first assumption insecticide resistance has been termed 'pre-adaptive' meaning that resistance alleles are present in very low numbers even prior to exposure to novel insecticides [@Ffrench-Constant2013][@Liu2015] as has been documented in blowflies [@Hartley2006].
-In support of using a single gene representation of resistance evidence suggests that whilst polygenic resistance is common, control failures, particularly those due to target site resistance, are mostly due to single major genes [@Ffrench-Constant2013].
-
-## Methods
-The simulation represents a population of randomly mixing individuals using standard population genetic approaches to avoid the need to follow every individual.
-One gene or locus is represented for each insecticide.
-Each individual has two copies of the gene and there are two potential alleles per locus.
-Each allele confers either resistance or susceptibility to the insecticide.
-Thus individuals can either have both resistance genes (RR homozygous), both susceptible genes (SS) or one of each (SR, heterozygous).
-The combination of the two genes is termed the genotype.
-Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
-The fitness of the SS genotype in absence of the insecticide is set at a maximum value of 1.
-All other fitnesses are less, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
-These inputs include the effectiveness of the insecticide, the dominance of the resistance allele and the ability of resistance to restore fitness in the presence of the insecticide (see Table 1).
-
-**Table 1. Parameters influencing the development of insecticide resistance**
-
-Parameter  | Description
----------------------- | ----------------------------------------------------
-1. Effectiveness | proportion of susceptible (SS) insects killed by exposure to insecticide
-2. Exposure | proportion of insects exposed to insecticide
-3. Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide
-4. Dominance of restoration | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
-5. Frequency  | frequency of resistance alleles within the population
-6. Cost of resistance  |  fitness of resistant (RR) insects in absence of insecticide
-7. Dominance of cost  | sets fitness of heterozygous (SR) insects between that of SS & RR in absence of insecticide
-
-The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 1.
-In the model itself this is repeated for both insecticides and the fitnesses multiplied.
-Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS and resistance restoration 'restores' a portion of the fitness for RR.
-A value of 1 for resistance restoration would bring the fitness of RR back to 1 (equivalent to the unexposed SS), a value of 0 for resistance restoration would lead to the RR having the same fitness as SS in the presence of the insecticide and effectively there would be no resistance.
-We use 'Resistance restoration' to represent how much the RR genotype restores fitness in the presence of the insecticide.
-The 'selection coefficient' of resistance, that may be seen referred to elsewhere [@Curtis1985], can be calculated by multiplying resistance restoration by insecticide effectiveness.
-Cost of resistance determines the fitness of RR in the portion of the population not exposed to the insecticide.
-
-Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
-For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
-Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Fig 1).
-
-The simulation proceeds through generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness.
-This acts to make the fitter alleles more common over time.
-Separate sexes and standard sexual reproduction with recombination is included.
-There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that here.
-
-So far this description just considers a single insecticide and associated resistance allele.
-In the model itself a second insecticide and resistance allele are included and overall fitness is calculated by multiplying the two results.
-This allows the model to represent populations exposed to two insecticides together in a mixture.
-
-The model outputs the change in resistance allele frequencies over time measured in generations.
-We extract the number of generations at which a resistance allele frequency of 50% is reached and term this 'time-to-resistance'.
-Using a 50% threshold is consistent with previous modelling analyses e.g.
-[@Curtis1985][@Birget2015].
-Preliminary investigations showed that using other potential thresholds of 25 or 75% gave very similar results.
-
-We investigate three insecticide use strategies.
-
-1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
-1. two insecticides used in a mixture (concurrently)
-
-The model is implemented in R [@RCoreTeam2016], the code is hosted on Github [@South2017] and we provide an on-line user interface [@South2017a] (Fig 2) enabling the reader to change inputs and run the model themselves.
-
-## Results
-We start by describing the results of model runs with a single insecticide which are more straightforward and move on to describing results of using two insecticides in sequences and mixtures.
-
-### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
-This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
-Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that less change is required to reach the threshold.
-
-Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect.
-Higher values led to a slower development of resistance.
-Again this is intuitive because costs reduce the advantage of resistance and decrease selection.
-These effects of inputs on single insecticide use are summarised in Table 2.
-
-**Table 2. Effect of inputs on resistance when insecticides used singly or in sequence**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | ------ | ---------------------------------------------
-1. Effectiveness | faster | reduced fitness of SS & SR exposed to insecticide
-2. Exposure | faster | reduced fitness of SS & SR overall
-3. Dominance of restoration | faster | increased fitness of SR exposed to insecticide
-4. Resistance restoration | faster | increased fitness of RR & SR exposed to insecticide
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of RR & SR not exposed to insecticide
-
-### Two insecticides
-When two insecticides are used we can compare the relative performance of concurrent use in a mixture with sequential use.
-Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
-To help understand the mechanisms influencing this relative performance we start with a single base scenario and demonstrate changing inputs individually.
-In this base scenario we set effectiveness, exposure, resistance restoration, and dominance at 0.5, the starting frequencies of resistance to 0.01 and cost of resistance to 0.
-Resistance arises slower for sequential than mixture for this base scenario (Fig 5A).
-Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 5A).
-Thus the time-to-resistance for the mixture divided by the sequence is less than 1 at 0.8.
-
-#### Insecticide exposure and effectiveness
-Increasing exposure (the proportion of insects that come into contact with the insecticides) (Fig 5C) decreases the time-to-resistance for both the sequential and mixture strategy.
-The effect on the mixture is greater so time-to-resistance remains longer for the sequence (the ratio decreasing from 0.8 to 0.6).
-In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 5B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
-The longer time-to-resistance for the mixture results from a slower increase in resistance for the less effective insecticide.
-Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
-This points a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
-
-The result is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
-This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-
-In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 5, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 5.
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
-
-Increasing the effectiveness further of either or both insecticides increases the performance of mixtures over sequences in terms of time-to-resistance.
-Figure 6A uses the same inputs as Fig 5B.
-From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
-This again points to how more effective insecticides can slow resistance to another when used in a mixture.
-
-#### Resistance restoration and dominance
-Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
-The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
-So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
-
-#### Starting frequencies of resistance
-Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
-Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
-
-#### Costs of resistance
-Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
-The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 9C).
-The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C.
-With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 9A and 1 in Fig 9C.
-At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
-
-A summary of the effects of inputs on the evolution of resistance for mixtures is given in Table 3 and for the difference between mixtures and sequences in Table 4.
-The mechanisms identified in the tables are covered further in the discussion.
-
-**Table 3. Effect of inputs on resistance when insecticides used in a mixture**
-
-Parameter to increase  | effect on resistance evolution | Mechanism
------------------------ | ---------- | ---------------------------------------------
-1. Effectiveness | **slower** | one insecticide reduces the fitness of individuals resistant to the other thereby reducing selection pressure for the other
-2.
-Exposure | faster (but less than for insecticide used alone) | reduced fitness of individuals susceptible to one insecticide increases selection pressure for that resistance.
-However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
-3. Dominance of restoration | faster | increased fitness of heterozygotes
-4. Resistance restoration | faster | increased fitness of resistants
-5. Frequency  | faster | less change needed to reach resistance threshold
-6. Cost of resistance | slower | reduced fitness of resistants in absence of insecticide
-
-**Table 4. Effect of inputs on the difference between mixture and sequential use**
-
-Parameter to increase  | increase favours mix or sequence | Mechanism
------------------------ | -------- | ---------------------------------------------
-1. Effectiveness | mixture | Higher effectiveness gives faster resistance for sequence and slower resistance in mixture
-2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4.
-Resistance restoration | neither | As for dominance of restoration.
-Higher resistance restoration gives faster resistance in both sequences and mixtures such that the difference between them is not changed.
-5.
-Frequency  | neither | As for dominance and resistance restoration.
-Higher starting frequencies give faster resistance in both sequences and mixtures such that the difference between them is not changed.
-6.
-Cost of resistance | mixture | Higher costs slow the evolution of resistance more in a mixture than when used in sequence (but with higher costs in a sequence there is a greater chance for resistance levels to decline for the insecticide not being used).
-
-#### Linkage disequilibrium
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
-
-## Discussion
-Resistance responds to insecticide use in the model in ways that we can explain mechanistically through the process of selection.
-This mechanistic explanation can help us develop a more robust understanding of how we would expect resistance to evolve in the field.
-We start by explaining the evolution of resistance to a single insecticide, which is relatively straightforward, and move on to explaining the effect of two insecticides in sequences and mixtures.
-
-When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 4B).
-Mechanisms explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
-Referring to figure 1 can help explain these responses.
-Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
-Cost acts by decreasing the fitness of RR and SR among those mosquitoes  not exposed.
-
-#### Two insecticides
-Responses to input parameters are different when two insecticides are used either in sequence or in mixtures as compared to when used alone.
-These differences can help us to understand the dynamics of resistance evolution.
-In each panel of Figures 5-9 the solid lines, indicating that an insecticide is being used at the same time as another in a mixture, are always shallower than the same colour dashed line indicating the insecticide is being used in isolation as part of a sequence.
-The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
-Because resistant individuals are being killed, the selection pressure for that resistance is lower.
-Thus each insecticide reduces the selection pressure for resistance to the other and could be said to 'protect' the other.
-This protection means that evolution of resistance to one insecticide is always slower when it is used in a mixture than when used alone.
-This protection has been termed multiple intra-generational killing [@Consortium2013] because individuals have the potential to be killed by more than one insecticide.
-
-Whilst this protection ensures that resistance to each insecticide evolves more slowly in a mixture it does not guarantee that a mixture strategy is preferable  to a sequence once the resistance to both insecticides is taken into account.
-In a mixture, selection pressure is applied by both insecticides at the same time, whereas in a sequence there is no selection for the insecticide not being used.
-Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
-
-This advantage of sequence over mixtures can be removed by increasing the effectiveness of one of the insecticides.
-Increasing the effectiveness of one insecticide (Fig 5B) decreases the total time for the sequence from 120 to 90 and increases that for the mixture from 90 to 115 thus changing the order and favouring the mixture.
-The mechanism for this change can be seen by comparing the shape of the curves in Figs 5A & B.
-In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
-However, the increased effectiveness of insecticide 1 causes resistance to insecticide 2 to increase more slowly initially in the mixture.
-Once insecticide 1 reaches resistance of around 50% at around 50 generations, the curve for insecticide 2 becomes steeper.
-Once the first insecticide reaches the resistance threshold of 50% it kills fewer individuals that are resistant to the second insecticide, it loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
-In both cases increasing the effectiveness of insecticides increases their protective effect when used in mixtures and favours mixtures over sequences.
-
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide (the red curve in this case).
-The more effective insecticide prompts both a) higher selection pressure for its own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
-The more effective insecticide has a faster evolution of resistance than the less effective one but this is not caused by the presence of the less effective one.
-Indeed the presence of the less effective one is still slowing the evolution of resistance to the more effective one, as indicated by the red solid curve for use in the mixture being shallower than the red dashed curve for use alone as part of the sequence.
-
-#### Difference between exposure and effectiveness
-It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
-This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us and we are keen to explore it in more detail.
-The most likely explanation is that increasing exposure increases overall selection pressure and lowers time to resistance, whereas increasing effectiveness increases mutual protection of the insecticides in a mixture and lowers the speed at which resistance evolves.
-
-#### Costs of resistance
-Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@Ffrench-Constant2017].
-Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 9).
-
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
-It makes sense that the protective effect of the other insecticide in a mixture combines with the cost of resistance to reduce selection pressure more than in a sequence.
-
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
-This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes also increases.
-Thus differences due to cost are less noticeable at low resistance frequencies and become more noticeable at higher resistance frequencies.
-Thus if a lower resistance threshold was used cost would have less of an effect and if a higher resistance threshold was used cost would have more of an effect.
-Also note that if the costs are set even higher than we set them here (e.g.
-above 0.2) they can maintain resistance frequencies below 1 or lead to their decline.
-However, we do not expect such high costs to occur in operational settings.
-
-Declines in resistance frequencies are likely to occur when there are costs of resistance and the insecticide is not being used [@Ffrench-Constant2017].
-Thus the frequency of resistance for the first insecticide in a sequence, after it has been replaced, may decline to a level where the first insecticide becomes operationally useful again.
-Consequently, a repeated sequential strategy could offer benefits over mixtures not considered here.
-To quantify these benefits one would need to use a more detailed metric than the time-to-resistance we have used.
-We quantified the number of generations until resistance thresholds are first reached.
-Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a resistance threshold.
-
-#### Consistency with previous work 
-The predicted responses of resistance presented here can be put into the context of previous work and recommendations.
-A recent review [@Consortium2013] found that 8 out of 10 empirical studies and 11 out of 14 modelling studies favoured mixtures over sequences.
-In the 3 other modelling studies the relative performance of mixtures and sequences were dependent on the values of other inputs as we also have shown.
-
-The review concluded that the advantage of mixtures is greatest when seven other conditions are satisfied [@Consortium2013] : i) rare starting resistance , ii) independent loci (i.e.
-no cross resistance), iii) high recombination, iv) high susceptible mortality, v) resistance is recessive, vi) similar persistence of insecticides, and vii) some of population remains untreated.
-Their points iv) and vii) agree with our results that mixtures are favoured by high effectiveness and low exposure.
-Some of the other conditions, notably i) the starting frequency of resistance and v) the dominance of resistance did not alter whether mixtures or sequences were favoured in our model [@Levick2017].
-
-Earlier modelling work has made sometimes contrasting predictions and these can be explained by the details of their structure and input choices.
-One model predicted that "the use of mixtures is always more effective in delaying the onset of resistance often by many orders of magnitude"[@Mani1985].
-That model assumed that SS genotypes were always killed, RR always survived, and that a constant 10% of mosquitoes escaped the insecticide.
-This would be represented in our model by a very high 'insecticide effectiveness' of 1, a 'resistance restoration' of 1 and an 'exposure' of 0.9.
-This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
-
-A subsequent model lead to the following more pessimistic conclusion about the potential of mixtures : “As a result of incomplete coverage and residue decay, the mortality of susceptible homozygotes is rarely consistently high enough for pesticide mixtures to be effective” [@Roush1998].
-This "mortality of susceptible homozygotes" is equivalent to our insecticide effectiveness.
-Again this points to the importance of effectiveness and the question of whether the effectiveness levels needed to make mixtures better are attainable operationally.
-Consistent with our results [@Roush1998] also showed (Figure 3) that decreasing exposure favours mixtures over sequences.
-
-#### Poor implementation
-One aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
-This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
-@Roush1998].
-This can seem to contradict recommendations that poor implementation of insecticide interventions can promote the development of resistance.
-e.g.
-from [@FAO2012] "Certain pest control practices have consistently been shown to exacerbate the loss of susceptible pest populations and the development of resistance.
-These include: ...
-- the use of application rates that are below or above those recommended on the label, - poor coverage of the area being treated ...".
-
-Poor implementation would likely cause lower effectiveness of, and exposure to, insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
-So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
-Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that higher effectiveness favours mixtures, higher exposure favours sequences and higher dominance favours neither (Table 4).
-More research into how the mortality of different genotypes (susceptible, resistant and heterozygous) is affected by poor application rates or insecticide degradation over time would help make this clearer.
-
-Experience from agriculture where single interventions have been associated with a rapid development of resistance have led for calls for a more Integrated Vector Management composed of a series of partially effective tools [@Thomas2012].
-It is suggested that such an integrated approach is more sustainable and 'evolution-proof' [@Thomas2012].
-A similar combining of interventions has been advocated to cover mosquitoes exhibiting different behaviours [@Killeen2017].
-Mosquitoes with different behaviours e.g.
-feeding outdoors and/or on animals are likely to result in lower exposures to insecticides in nets or sprayed on walls.
-This lower exposure could favour mixtures over sequences as we have indicated.
-In both cases our modelling approach could be used to investigate implications for the development of resistance.
-
-#### Caveats
-There is much evidence of the involvement of multiple genes in resistance mechanisms.
-Our model, in common with most previous ones, assumes a single ‘resistance’ gene per insecticide.
-In most cases we believe this is reasonable.
-Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Roush1989][@Ffrench-Constant2004][@Ffrench-Constant2013].
-Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
-To represent polygenic resistance a different 'quantitative genetic' approach would be required [@Via1986][@Firko1990].
-We are only aware of preliminary published work applying a quantitative genetic approach to assessing insecticide use strategies in the context of polygenic resistance [@Via1986][@Gardner1998].
-One suggestion is that evolution of resistance to one insecticide in a mixture could be associated with a decrease in tolerance to the other [@Via1986].
-If it becomes clearer that polygenic resistance is an issue then we would welcome a comparison of the results presented here to those of quantitative genetic models.
-
-Even if the assumption that single genes are responsible for control failures is correct there is still considerable other uncertainty around the development of insecticide resistance in operational field settings [@Ranson2016].
-The model we present here is able to cover much of the range of that uncertainty, yet in this paper, to develop an understanding of the key mechanisms we restrict our explorations to a limited region of parameter space (e.g.
-holding most inputs constant at intermediate values while modifying single inputs in isolation).
-There is the potential that the system may behave differently in the field than we describe here.
-However, the results presented here are supported by an earlier analysis where we ran 10,000 unique scenarios varying more inputs [@Levick2017].
-This previous analysis also highlighted the relative importance of insecticide effectiveness and exposure in determining the relative performance of mixtures and sequences.
-
-In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers to investigate model behaviour in areas of parameter space we have not shown.
-The user-interface allows inputs to be set for two scenarios and results to be viewed side by side.
-For example you can set a higher insecticide effectivenesses in one scenario to see how this favours mixtures over sequences (Fig 2) or increase the exposure levels to see how this favours sequences.
-
-There are other model behaviours that can be changed by modifying input files.
-For example, here we assume that insecticide interaction in a mixture is multiplicative, e.g.
-if the probability of surviving exposure to insecticide A alone is 0.3 and of surviving insecticide B alone is 0.2, then the probability of surviving exposure to a mixture of A and B is 0.3*0.2 = 0.06.
-This is consistent with recent work on aphids [@Taillebois2016], although that did also indicate some synergistic effects.
-Modifying input files would allow us to incorporate such synergistic or antagonistic effects of mixtures, i.e.
-that insecticides used together perform better or worse than when used separately.
-
-In addition to these uncertainties regarding resistance it is important to remember that our model examines solely the evolution of resistance and not the control of mosquitoes or malaria.
-Different insecticide use strategies, as well as affecting the evolution of resistance, will also influence both the numbers of vectors and their propensity to transmit the disease [@Rivero2010].
-There are also potential knock-on effects of resistance itself; reduced vector transmission of disease (e.g.
-by reduced lifespan of the vector) or increased transmission (e.g.
-by increasing the susceptibility  of the vector to the disease) [@Rivero2010].
-This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance.
-Other operational factors also need to be considered.
-On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease.
-On the negative side a mixture may be more expensive.
-Other attributes of insecticides such as their repellency are also relevant.
-Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mixtures over sequences.
-However the reduced exposure resulting from repellency also reduces the population protection provided by the insecticide [@Birget2015].
-
-Our model would predict that the best strategy for limiting the development of resistance in the short term would be to use no insecticide.
-Of course, using no insecticide is likely to have serious consequences for mosquito and malaria abundance.
-The challenge is to develop good strategies for delivering mosquito and malaria control in the short term while sustaining their operational capacity in the longer term by restricting the development of resistance.
-
-#### Conclusions
-We have developed a flexible, modern model of the evolution of insecticide resistance that can that help us understand the mechanisms by which different insecticide application strategies may be favoured.
-It replicates previous modelling work and has the potential to investigate further operational strategies for established and new insecticides.
-For assessing the relative benefits of mixtures and sequences we have demonstrated that mixtures are likely to be favoured by high insecticide effectiveness, low exposure and high costs of resistance.
-Similar processes will occur in other strategies such as rotations and mosaics.
-This accessible modelling approach can help to promote discussion and exploration of the likely effect on resistance evolution of different insecticide use strategies in tandem with the collection of empirical data to refine and test predictions.
-
-### Declarations
-
-##### Ethics approval and consent to participate
-Not applicable
-
-##### Consent for publication
-Not applicable
-
-##### Availability of data and material
-Project name: resistance
-Project home page: https://github.com/AndySouth/resistance
-Archived version: 10.5281/zenodo.889012
-Operating system: Platform independent
-Programming language: R
-License: GNU General Public License v3.0
-
-##### Competing interests
-AS received a small grant from Bayer S.A.S. to support a few months work and attendance of the American Society of Tropical Medicine and Hygiene Annual Meeting.
-
-##### Funding
-This work was supported by: Bill and Melinda Gates Foundation, grant #37999.01 to IMH; and Wellcome Trust Institutional Strategic Strengthening fund grant #105620/Z/14/Z to Liverpool School of Tropical Medicine; sub-funding to IMH; and a small grant from Bayer S.A.S.
-to AS.
-
-##### Authors' contributions
-AS conceived the paper, re-factored the model code, performed the analyses and wrote the first and subsequent versions of the paper.
-IH conceived the modelling approach, developed the first version of the model in collaboration with Bethany Levick, explained concepts to AS and reviewed and commented on all manuscript versions.
-
-##### Acknowledgements
-Thank you to Bethany Levick who developed the first version of the model code.
-
-\pagebreak
-
-## Figures
 
@@
 For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3."}
+In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of restoration=0.7 which sets the fitness of the SR closer to RR than SS.
+Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
 
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.8, cost=0.3 )
 
 ```
-
-![Figure 2.
-Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2/>.
-The user can modify values of the input parameters considered in this paper using simple sliders and run the model to get graphs of resulting resistance frequency over time.
-Two scenarios (A and B) can be run and the results viewed side by side.
-This makes it easy to explore the effect of changing individual inputs.](paper2_UI1.png)
-
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 3.
-Single insecticide effects of effectiveness, exposure, dominance and resistance restoration on resistance frequency.
-A.
-Effectiveness, B.
-Exposure, C.
-Dominance of restoration, D.
-Resistance restoration.
-Increasing any of the inputs (going from red to blue) results in shorter times-to-resistance.
-For each panel the chosen input was varied in isolation with the remaining inputs set to 0.5, except for starting frequency of resistance which was set to 0.01).
-"}
-
-  library(cowplot)
-
-  #list to put plots in for cowplot
-  plotlist <- list(4)
-
-  #A effectiveness  
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.2,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.8,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  #input <- cbind(input, inputOneScenario)
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(phi.SS1_A0))) + 
-         theme_bw() +  
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +  
-         labs(colour = "effectiveness") +
-         coord_trans(y = "log10") +    
-         geom_line()
-
-  #B exposure
-  exposure <- 0.2
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.5
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  exposure <- 0.8
-  a <- setExposure(exposure = exposure, insecticideUsed='insecticide1')
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      exposure = exposure, #used for post-run analysis
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(exposure))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "exposure", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  #plot_grid( plotlist[[1]],plotlist[[2]],ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-  #C single insecticide dominance
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.2, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.8, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #C 
-  plotlist[[3]] <-ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(h.RS1_A0))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance of restoration") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #D single insecticide resistance restoration
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.2,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.8,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #D  
-  plotlist[[4]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(rr_restoration_ins1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) +
-         labs(colour = "resistance restoration", y=NULL) +
-         coord_trans(y = "log10") +   
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 4.
-Single insecticide effects of starting frequency and cost of resistance on resistance frequency.
-A.
-Starting frequency, B.
-Cost of resistance.
-Increasing the starting frequency decreases the number of generations taken to reach resistance thresholds.
-Increasing costs of resistance increases time-to-resistance."}
-
-  #single insecticide frequency fig
-
-  #A starting frequency
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.001,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.1,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #A  
-  plotlist[[1]] <-  ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(P_1))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "start\nfrequency") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  #B changing costs at dominance 0.5
-
-  a <- setExposure(exposure=0.5, insecticideUsed='insecticide1')
-
-  i1 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0,
-                      h.RS1_00 = 0.5)
-
-  i2 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.1,
-                      h.RS1_00 = 0.5 )
-
-  i3 <- setInputOneScenario( max_gen = 500,
-                      P_1 = 0.01,       
-                      h.RS1_A0 = 0.5, 
-                      h.RS2_0B = 0.5,
-                      a = a,
-                      phi.SS1_A0 = 0.5,
-                      phi.SS2_0B = 0.5,
-                      rr_restoration_ins1 = 0.5,
-                      rr_restoration_ins2 = 0.5,
-                      z.RR1_00 = 0.2,
-                      h.RS1_00 = 0.5 )
-
-  input <- cbind(i1,i2,i3)
-
-  listOut <- runModel2( input )
-
-  df_resist <- get_resistance(locus=1, listOut)
-
-  #B  
-  plotlist[[2]] <- ggplot(df_resist, aes(x=generation, y=resistance, colour=factor(z.RR1_00))) + 
-         theme_bw() +
-         theme(legend.position = "bottom", legend.key = element_blank()) + 
-         guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "cost of\nresistance") +
-         coord_trans(y = "log10") + 
-         geom_line()
-
-  cowplot::plot_grid( plotlist[[1]],plotlist[[2]], ncol=2, rel_widths=c(1,1), labels='AUTO')  
-  #cowplot::plot_grid( plotlist[[1]],plotlist[[2]],plotlist[[3]],plotlist[[4]], ncol=2, rel_widths=c(1,1), labels='AUTO')
-
-```
-
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 5.
-Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
-Exposure to the insecticides is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of one insecticide is increased from column 1 (A and C) to column2 (B and D).
-On the upper X axis 's' and 'm' indicate where the 50 percent resistance threshold is reached for the sequence and the mixture.
-In the lower right of each panel the the ratio of time-to-resistance for mixture/sequence is shown rounded to 1 decimal place to give an indication of the relative performance of mixtures and sequences.
-A.
-All control inputs equal at 0.5 : time-to-resistance is longer for sequential use, B.
-Effectiveness of insecticide1 increased from 0.5 to 0.8 : time-to-resistance is longer for the mixture, C.
-Exposure increased to 0.8 : time-to-resistance is longer for sequential use, D.
-Effectiveness of insecticide1 and exposure increased to 0.8 : time-to-resistance equal for mixture and sequence.
-Increasing effectiveness increases times-to-resistance for mixtures and improves their performance relative to sequences.
-Increasing exposure decreases times-to-resistance for mixtures and reduces their performance relative to sequences.
-"}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-#layout.show(n=9)
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.8 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 6.
-Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
-Effectiveness of insecticide 2 is increased from row 1 (A and B) to row 2 (C and D).
-The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-Effectiveness of insecticide1 increased from 0.5 to 0.8, B.
-Effectiveness of insecticide1 increased from 0.5 to 1, C.
-Effectiveness of insecticides 1 and 2 increased to 0.8, D.
-Effectiveness of insecticide2 0.8 and of insecticide1 1.
-With the effectiveness of at least one insecticide greater than or equal to 0.8, times-to-resistance are longer for the mixture in all scenarios."}
-
-#4 panel figure looking at effectiveness changes alone
-
-#test of rearranging figures on a grid
-#eff1  0.8  1
-#eff2
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.5
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness2 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 1 eff2 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 1 , phi.SS2_0B = 0.8 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 1', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 7.
-Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
-Dominance of the allele coding for resistance to insecticide 1 is increased from row 1 (A and B) to row 2 (C and D).
-Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, B.
-Resistance restoration for insecticide1 increased from 0.5 to 0.8, C.
-Dominance for insecticide1 increased to 0.8, D.
-Resistance restoration and dominance for insecticide1 increased to 0.8.
-Changing dominance and resistance restoration does not change the relative ordering of mixtures and sequences, time-to-resistance remains longest for sequences in all 4 scenarios."}
-
-#4 panel figure looking at dominance and resistance restoration
-
-#rr1  0.5  0.8
-#dom1
-#0.5    1   2
-#0.8    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#testing
-# plot(1)
-# plot(1)
-# plot(1)
-# plot(1)
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.5', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('dominance1 = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dominance1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#dom1 & rr1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.8 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.8 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 120, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('resistance restoration1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 8.
-Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
-Starting frequency of the gene conferring resistance to insecticide 1 is decreased from row 1 (A and B) to row 2 (C and D).
-Effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-A.
-All control inputs equal at 0.5, starting frequencies of resistance at 0.01, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Starting frequency of resistance for insecticide1 decreased from 0.01 to 0.001, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and starting frequency for insecticide1 decreased from 0.01 to 0.001.
-In these scenarios the starting frequencies do not change better performance of sequences at low effectiveness and mixtures at high effectiveness."}
-
-#4 panel figure looking at effectiveness and starting freq.
-
-#eff1  0.5  0.8
-#strt freq1
-#0.01    1   2
-#0.001    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('frequency1 = 0.1%', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.001 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Fig9, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 9.
-Influence of cost of resistance on time-to-resistance for mixtures and sequences.
-Row 1 (A and B) same as Fig.
-5 with no cost of resistance, row 2 (C and D) cost of resistance set to 0.15 with dominance of cost set to 0.5.
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
-The increase in cost of resistance from A.
-to C.
-removes the advantage of sequence relative to mixture.
-The ratio of time-to-resistance for the mixture divided by that for the sequence goes from 0.
-to 1.
-Costs seem to favour mixtures in these plots.
-Also be aware that costs would lead to a decline in the frequency of resistance for the first insecticide in a sequence when it stops being used."}
-
-#4 panel figure looking at effectiveness and exposure
-
-#test of rearranging figures on a grid
-#eff1  0.5  0.8
-#exp
-#0.5    1   2
-#0.8    3   4
-
-#layout(matrix(1:4,2,2, byrow = TRUE))
-
-#add lower panel for legend
-#layout(matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.5), respect=FALSE)
-
-#layout( matrix(c(1,2,3,4,5,5),3,2, byrow = TRUE), heights=c(1,1,0.2), 
-# respect=matrix(c(1,1,1,1,0,0),3,2, byrow = TRUE) )
-#layout.show(n=5)
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#effectiveness on mixture
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#effectiveness1 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('cost 1 = 0.15', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#exposure 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1  and  expos 0.8
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 150, labelMixSeqRatio = 1, z.RR1_00 = 0.15  )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-```{r Figxx not eval, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.
-All control inputs equal at 0.5, male exposure 0.8, B.
-Effectiveness for insecticide1 increased from 0.5 to 0.8, C.
-Male exposure decreased from 0.8 to 0.2, D.
-Effectiveness for insecticide1 increased from 0.5 to 0.8 and Male exposure decreased from 0.8 to 0.2"}
-
-# not evaluated at the moment
-#4 panel figure looking at male exposure prop and starting freq.
-
-#eff1  0.5  0.8
-#maleprop
-#0.8    1   2
-#0.2    3   4
-
-# add side panels for labels
-layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure = 0.8', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#base scenario
-
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#eff1 0.8
-runcurtis_f2( male_exposure_prop=0.8, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, xlabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('male exposure =0.2', line= -2, side=2, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 4, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.5 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('C.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#plot margins
-par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
-
-#freq1 0.001, eff1 0.8
-runcurtis_f2( male_exposure_prop=0.2, max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), ylab="", ylabs = FALSE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 200, labelMixSeqRatio = 1 )
-
-mtext('D.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r
-
-#adding legend in lower panel
-#plot margins
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-
-# legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide1 in mix   ", "insecticide2 alone   ", "insecticide2 in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#add for when insecticides same
-legend( 'bottom', legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   ", "insecticides1&2 in mix   "), col=c("red","blue","red","blue","purple"), lty=c(2,2,1,1,1), pch=c(NA,NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.5', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-#label
-par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
-plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
-mtext('effectiveness1 = 0.8', line= -2, side=1, cex=0.9) #side=1b,2l,3t,4r
-
-```
-
-\pagebreak
-
-## References
-<!-- now auto created here-->

### 2017-12-18 R/plot_fit_calc.r

@@ -7,12 +7,13 @@
 #' @param dominance_restoration dominance of resistance restoration 0-1
 #' @param dominance_cost dominance of resistance cost 0-1
 #' @param cost cost of resistance
-#' @param title optional title for the plot
 #' @param simple default FALSE whether to create a simpler plot e.g. for public health journal
+#' @param title optional title for the plot
 # @param yblank whether to remove y axis title & labels
 # @param size text size for RS labels, default 4
 #' 
-#' @import ggplot2 gridExtra
+#' #grid needed for title
+#' @import ggplot2 gridExtra grid
 #' 
 #' @examples 
 #' plot_fit_calc(effectiveness=0.5, resistance_restoration=0.5, dominance=0.5)
@@ -27,7 +28,8 @@ plot_fit_calc <- function ( effectiveness=0.75,
                             dominance_restoration=0.4, 
                             dominance_cost=0.6, 
                             cost=0.2,
-                            simple = FALSE){
+                            simple = FALSE,
+                            title = "Fitness calculation for each genotype in each generation for one insecticide"){
     
   
   #library(gridExtra)
@@ -56,18 +58,17 @@ plot_fit_calc <- function ( effectiveness=0.75,
   gg1 <- ggplot(dfg, aes(x = x2, y = y2)) +    
         ylim(0,1) +
         #points
-        geom_point(size=3, colour='blue') +
+        geom_point(size=3, colour='red4') +
         #arrows
-        geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2),size=0.9, arrow = arrow(length = unit(3,"mm"))) +
+        geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2),size=0.9, arrow = arrow(length = unit(3,"mm")), colour="grey40") +
         #dotted horizontal lines showing calculations
-        annotate("segment", x='SS', xend='RR', y=ss, yend=ss, linetype="dotted", colour = "blue") +
-        annotate("segment", x='SR', xend='RR', y=rr, yend=rr, linetype="dotted", colour = "blue") +
+        annotate("segment", x='SS', xend='RR', y=ss, yend=ss, linetype="dotted", colour = "red4") +
[…]

### 2017-12-18 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
responses to reviewer

@@
 For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, dominance of restoration=0.6, resistance restoration=0.5, dominance of cost=0.6 and resistance cost=0.3."}
-
-plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.6, dominance_cost=0.6, cost=0.3 )
+In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of restoration=0.7 which sets the fitness of the SR closer to RR than SS.
+Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
+
+plot_fit_calc( effectiveness=0.8, resistance_restoration=0.5, dominance_restoration=0.7, dominance_cost=0.7, cost=0.3 )
 

### 2017-12-18 R/inputs_from_gen_fit.r

@@ -18,8 +18,13 @@
 #' #see survival_from_genotype_livedead.xls
 #' #1 Kolaczinski2000 etofenprox (pyr), gives low eff 0.43, rr_ 0.63, dom 0.30
 #' inputs_from_gen_fit(RR=0.84, RS=0.65, SS=0.57)
+#' #Kolaczinski2000 etofenprox (pyr) including control mortality. eff 0.32, rr_ 1, dom 0.30
+#' inputs_from_gen_fit(RR=0.84, RS=0.65, SS=0.57, SS_noi=0.84)
 #' #Kolaczinski2000 Alpha-cypermethrin eff 0.41, rr_ 0.60, dom 0.52
 #' inputs_from_gen_fit(RR=0.84, RS=0.72, SS=0.59)
+#' #Kolaczinski2000 Alpha-cypermethrin with control mort eff 0.29, rr_ 1, dom 0.54, cost 0.19, domcost 1 (truncated from 2.06)
+#' inputs_from_gen_fit(RR=0.83, RS=0.72, SS=0.59, SS_noi=0.83) 
+#' inputs_from_gen_fit(RR=0.83, RS=0.72, SS=0.59, SS_noi=0.83, RS_noi = 0.5, RR_noi = 0.67) 
 #' #2 Essandoh2013 An. gambiae	bendiocarb	Ace1, eff 0.98, rr_ 0.84, dom 0.66
 #' inputs_from_gen_fit(RR=0.84, RS=0.56, SS=0.02)
 #' #Essandoh2013 An. coluzii	bendiocarb	Ace1, eff 0.94, rr_ 1, dom 0.36
@@ -33,25 +38,47 @@
 #' @return fitness values
 #' @export
 
-inputs_from_gen_fit <- function ( SSfit = 0.2,
-                                  RSfit = 0.4,
-                                  RRfit = 1)
+inputs_from_gen_fit <- function ( SS = 0.2,
+                                  RS = 0.4,
+                                  RR = 1,
+                                  SS_noi = NULL,
+                                  RS_noi = NULL,
+                                  RR_noi = NULL)
 {
   
-  dfin <- data.frame(eff=NA, rr_=NA, dom=NA)
+  dfin <- data.frame(eff=NA, rr_=NA, dom=NA, cost=NA, domcost=NA)
   
   #in insecticide
   # Effectiveness (1-SSfit)
[…]

### 2018-01-02 inst/documents/paper2/field_data_in.Rmd
tweak

@@
 
-### Section demonstrating getting field data into the model, to be added to Malaria Journal paper.
+### to be added to Malaria Journal paper.
+### PUT THIS SECTION AT END OF RESULTS
 
-The main model inputs relating to properties of the mosquitoes and insecticides can be derived from data on the relative survival of the different genotypes (RR,RS,SS) Table x.
+### Setting model inputs from field data
 
-**Table x1. Calculating model inputs from field data.**
+The main model inputs quantifying properties of the mosquitoes and insecticides can be derived from data on the relative survival of the different genotypes (RR,SR,SS) as described in Table 5.
 
-Parameter  | Calculation (where RRfit, RSfit, SSfit are fitnesses of genotypes)
+**Table 5. Equations to calculate model input parameters from field data. Where RRfit, SRfit, SSfit are fitnesses of genotypes. Derived from [@Levick2017].**
+
+Parameter  | Calculation 
 ---------------------- | ----------------------------------------------------
@@
 3. Resistance restoration  | (RRfit-SSfit) / Effectiveness in presence of insecticide
-4. Dominance of restoration | (RSfit-SSfit)/(RRfit-SSfit) in presence of insecticide
+4. Dominance of restoration | (SRfit-SSfit)/(RRfit-SSfit) in presence of insecticide
 5. Frequency  | estimated
 6. Cost of resistance  |  1 - RRfit in absence of insecticide
-7. Dominance of cost  | (RSfit-RRfit)/(RRfit-SSfit) in absence of insecticide
+7. Dominance of cost  | (SRfit-RRfit)/(RRfit-SSfit) in absence of insecticide
 
 Exposure, the proportion of mosquitoes that are exposed to the insecticide, is a property of the location and will depend on the use of nets or IRS and mosquito behaviour.
-The frequency of resistance alleles can be measured, although when at low levels a large number of mnosquitos would need to be sampled to detect low frequencies.
+The frequency of resistance alleles can be measured, although when at low levels a large number of mosquitoes would need to be sampled to detect low frequencies.
 
-As an example of making these calculations we use data for Anopheles gambiae on pyrethroid (Etofenprox) resistance associated with the Kdr mutation [@Kolaczinski2000] and carbamate (Bendiocarb) resistance associated with the Ace1 mutation [@Essandoh2013].
+As an example of making these calculations we use data for _Anopheles gambiae_ on pyrethroid (Alpha-cypermethrin) resistance associated with the Kdr mutation [@Kolaczinski2000] and carbamate (Bendiocarb) resistance associated with the Ace1 mutation [@Essandoh2013].
+This is not intended to be a definitive prediction of resistance evolution, sample sizes for some genotypes are low so fitnesses are approximate.
+It is intended to demonstrate how models of resistance evolution can be set to field data and hence bring the two approaches closer together.
 
-In both cases data were provided as numbers alive and dead by genotype.
-We used these to calculate genotype-specific survival values that we assume are a good indication of fitness and from that the values of our model inputs (Table x2).
+Data were provided as numbers alive and dead by genotype in both publications.
+We used these to calculate genotype-specific survival values as an indication of fitness.
+We used the non-exposed control data for the pyrethroid [@Kolaczinski2000] to estimate the cost of resistance and to rescale the fitness estimates on a scale of 0-1 where 1 is the fitness of the unexposed SS (i.e.
+we divided all pyrethroid survivals by that of the unexposed SS genotype, 0.83).
+In the absence of unexposed data for the carbamate [@Essandoh2013] we assumed no costs of resistance and that survival of the non-exposed SS would be 1.
+From these estimates we calculated the values of our model inputs (Table 6).
 
-Insecticide/Mutation|survival RR|survival RS|survival SS|Effectiveness|Resistance restoration|Dominance of restoration
----|---|---|---|---|---|---
-Pyrethroid/Kdr|0.84|0.65|0.57|0.43|0.63|0.30
-Carbamate/Ace1|0.84|0.56|0.02|0.98|0.84|0.66
+**Table 6. Model input parameters calculated from field data.**
 
-These input values can then be used in the model.
-In Figure x insecticide 1 and 2 are set to the pyrethroid (red) and carbamate (blue) respectively.
-Exposure is set to 0.5 and starting frequencies to 0.01.
+**Field data**|Pyrethroid / Kdr[@Kolaczinski2000]|rescaled Pyrethroid / Kdr[@Kolaczinski2000]|Carbamate /  Ace1[@Essandoh2013]
+------------------------------|-------|-------|-------
+exposed survival RR|0.83|1|0.84
+exposed survival SR|0.72|0.87|0.56
+exposed survival SS|0.59|0.71|0.02
+unexposed survival RR|0.67|0.81||-
+unexposed survival SR|0.50|0.60|-
+unexposed survival SS|0.83|1|-
+**Calculated model inputs**|    |    |
+Effectiveness|0.41|0.29|0.98
+Resistance restoration|0.60|1|0.84
+Dominance of restoration|0.52|0.54|0.66
+Cost of resistance|-|0.19|-
+Dominance of cost|-|1*|-
 
-```{r Fig10, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 10. Running the model using field data."}
+**For the pyrethroid the unexposed survival of the SR (0.5) was less than that for the RR (0.67) This indicates underdominance which is biologically unlikely for unexposed genotypes so we set dominance at 1 so that SR and RR genotypes have the same fitness costs.
+The authors indicated that suprisingly high mortality in the control huts may have been due to rough handling [@Kolaczinski2000] and this could contribute to the unexpected values here.*
+
+The results of using these calculated inputs in one model example are shown in Figure 11.
+Remember that this is a demonstration exercise for setting model parameters from field data and not a definitive prediction for whether using a pyrethroid and a carbamate in a mixture or sequence is likely to be preferable.
+
+```{r Fig11, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.align='left', fig.cap="Figure 11.
+Time-to-resistance for a mixture and sequence using inputs derived from field studies as outlined in Table x2.
+Alpha-cypermethrin (pyrethroid) is shown in red and Bendiocarb (carbamate) is shown in blue.
+Exposure is set to 0.8 and starting frequencies to 0.01.
+Resistance to the carbamate (blue) increases very quickly in both the mixture and sequence due to it's high effectiveness.
+Resistance to the pyrethroid (red) increases more slowly due to low effectiveness and high cost of resistance.
+In a mixture the resistance frequency of the pyrethroid even declines when resistance to the carbamate is low.
+In this illustrative example resistance to both insecticides takes longer to evolve for a mixture than a sequence."}
 
@@
 
-runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.3 , h.RS2_0B = 0.66 , exposure = 0.5 , phi.SS1_A0 = 0.43 , phi.SS2_0B = 0.98 , rr_restoration_ins1 = 0.63 , rr_restoration_ins2 = 0.84 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, labelMixSeqRatio = 1 )
+# runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.3 , h.RS2_0B = 0.66 , exposure = 0.5 , phi.SS1_A0 = 0.43 , phi.SS2_0B = 0.98 , rr_restoration_ins1 = 0.63 , rr_restoration_ins2 = 0.84 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, labelMixSeqRatio = 1 )
+
+# Alpha-cypermethrin (pyr) incl cost, exposure 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01, z.RR1_00 = 0.19, h.RS1_00 = 1,  h.RS1_A0 = 0.54 , h.RS2_0B = 0.66 , exposure = 0.8 , phi.SS1_A0 = 0.29 , phi.SS2_0B = 0.98 , rr_restoration_ins1 = 1 , rr_restoration_ins2 = 0.84 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, labelMixSeqRatio = 1 )
 
@@
 
-Resistance to the carbamate (blue) rises quickly because it has high effectiveness and dominance.
-The mixture performs better than the sequence because the carbamate with it's high effectiveness provides high protection to the pyrethroid.
+### PUT THIS SECTION IN DISCUSSION
+
+### Using field derived inputs
+
+Model behaviour when using the field derived inputs (Fig.
+11) can be explained using the understanding developed from the preceeding runs.
+Resistance to the carbamate (blue) rises quickly partly because it has high effectiveness (0.98).
+Resistance to the pyrethroid (red) increases more slowly due to it's much lower effectiveness (0.29) and high cost of resistance.
+In the mixture the resistance to the pyrethroid declines at the start when resistance to the carbamate is low.
+This illustrates the protective effect of the insecticide with the higher effectiveness.
+That protective effect causes the mixture to out-perform the sequence in this example.
+Indeed, if exposure is set lower to 0.5 then resistance to the pyrethroid declines further and the resistance threshold is not reached.
+This is a combination of the low effectiveness creating a low selection pressure for the resistance and the high cost creating a selection pressure against it.
+The low effectiveness and high cost are partly due to the surprisingly high mortality in unexposed mosquitoes in the field data [@Kolaczinski2000].
 

### 2018-01-02 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
add field inputs text and fig(11) start modifying fig nums

@@
 
-THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome
 \pagebreak
@@
 
+### Setting model inputs from field data
+
+The main model inputs quantifying properties of the mosquitoes and insecticides can be derived from data on the relative survival of the different genotypes (RR,SR,SS) as described in Table 5.
+
+**Table 5. Equations to calculate model input parameters from field data. Where RRfit, SRfit, SSfit are fitnesses of genotypes. Derived from [@Levick2017].**
+
+Parameter  | Calculation 
+---------------------- | ----------------------------------------------------
+1. Effectiveness | 1 - SSfit in presence of insecticide
+2. Exposure | estimated
+3. Resistance restoration  | (RRfit-SSfit) / Effectiveness in presence of insecticide
+4. Dominance of restoration | (SRfit-SSfit)/(RRfit-SSfit) in presence of insecticide
+5. Frequency  | estimated
+6. Cost of resistance  |  1 - RRfit in absence of insecticide
+7. Dominance of cost  | (SRfit-RRfit)/(RRfit-SSfit) in absence of insecticide
+
+Exposure, the proportion of mosquitoes that are exposed to the insecticide, is a property of the location and will depend on the use of nets or IRS and mosquito behaviour.
+The frequency of resistance alleles can be measured, although when at low levels a large number of mosquitoes would need to be sampled to detect low frequencies.
+
+As an example of making these calculations we use data for _Anopheles gambiae_ on pyrethroid (Alpha-cypermethrin) resistance associated with the Kdr mutation [@Kolaczinski2000] and carbamate (Bendiocarb) resistance associated with the Ace1 mutation [@Essandoh2013].
+This is not intended to be a definitive prediction of resistance evolution, sample sizes for some genotypes are low so fitnesses are approximate.
+It is intended to demonstrate how models of resistance evolution can be set to field data and hence bring the two approaches closer together.
+
+Data were provided as numbers alive and dead by genotype in both publications.
+We used these to calculate genotype-specific survival values as an indication of fitness.
+We used the non-exposed control data for the pyrethroid [@Kolaczinski2000] to estimate the cost of resistance and to rescale the fitness estimates on a scale of 0-1 where 1 is the fitness of the unexposed SS (i.e.
+we divided all pyrethroid survivals by that of the unexposed SS genotype, 0.83).
+In the absence of unexposed data for the carbamate [@Essandoh2013] we assumed no costs of resistance and that survival of the non-exposed SS would be 1.
+From these estimates we calculated the values of our model inputs (Table 6).
+
+**Table 6. Model input parameters calculated from field data.**
+
+**Field data**|Pyrethroid / Kdr[@Kolaczinski2000]|rescaled Pyrethroid / Kdr[@Kolaczinski2000]|Carbamate /  Ace1[@Essandoh2013]
+------------------------------|-------|-------|-------
+exposed survival RR|0.83|1|0.84
+exposed survival SR|0.72|0.87|0.56
+exposed survival SS|0.59|0.71|0.02
+unexposed survival RR|0.67|0.81||-
+unexposed survival SR|0.50|0.60|-
+unexposed survival SS|0.83|1|-
+**Calculated model inputs**|    |    |
+Effectiveness|0.41|0.29|0.98
+Resistance restoration|0.60|1|0.84
+Dominance of restoration|0.52|0.54|0.66
+Cost of resistance|-|0.19|-
+Dominance of cost|-|1*|-
+
+**For the pyrethroid the unexposed survival of the SR (0.5) was less than that for the RR (0.67) This indicates underdominance which is biologically unlikely for unexposed genotypes so we set dominance at 1 so that SR and RR genotypes have the same fitness costs.
+The authors indicated that suprisingly high mortality in the control huts may have been due to rough handling [@Kolaczinski2000] and this could contribute to the unexpected values here.*
+
+The results of using these calculated inputs in one model example are shown in Figure 11.
+Remember that this is a demonstration exercise for setting model parameters from field data and not a definitive prediction for whether using a pyrethroid and a carbamate in a mixture or sequence is likely to be preferable.
+
 ## Discussion
@@
 Other measures for quantifying resistance over time can be imagined, for example the mean resistance or proportion of generations below a resistance threshold.
+
+### Using field derived inputs
+Model behaviour when using the field derived inputs (Fig.
+11) can be explained using the understanding developed from the preceeding runs.
+Resistance to the carbamate (blue) rises quickly partly because it has high effectiveness (0.98).
+Resistance to the pyrethroid (red) increases more slowly due to it's much lower effectiveness (0.29) and high cost of resistance.
+In the mixture the resistance to the pyrethroid declines at the start when resistance to the carbamate is low.
+This illustrates the protective effect of the insecticide with the higher effectiveness.
+That protective effect causes the mixture to out-perform the sequence in this example.
+Indeed, if exposure is set lower to 0.5 then resistance to the pyrethroid declines further and the resistance threshold is not reached.
+This is a combination of the low effectiveness creating a low selection pressure for the resistance and the high cost creating a selection pressure against it.
+The low effectiveness and high cost are partly due to the surprisingly high mortality in unexposed mosquitoes in the field data [@Kolaczinski2000].
 
@@
 
-```{r Fig1, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 1.
+![Figure 1.
+Potential resistance genotypes with A) one and B) two insecticides.
+An individual has two copies of the gene and each can be either resistant (R) or susceptible (S).
+Thus for one insecticide (A) individuals can either have both resistance genes (RR homozygous), both susceptible genes (SS) or one of each (SR, heterozygous).
+For two insecticides (B) there are 9 potential genotypes, here with the gene for one insecticide represented by the first two characters (purple) and the second insecticide by the final two characters (green).
+Thus RR SR represents an individual homozygous resistant to the first insecticide and heterozygous resistant to the second.](paper2_fig1new_genotypes.tif)
+
+```{r Fig2, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 2.
 The effect of model inputs on the fitness of genotypes for a single insecticide.
@@
 
-![Figure 2.
+![Figure 3.
 Screenshot of one online model user interface, accessible at : <https://andysouth.shinyapps.io/resistmob2>.
@@
 
-```{r Fig3, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 3.
+```{r Fig4, eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.height=7.5, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 4.
 Single insecticide effects of effectiveness, exposure, dominance and resistance restoration on resistance frequency.
@@
 
+```{r Fig11, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.align='left', fig.cap="Figure 11.
+Time-to-resistance for a mixture and sequence using inputs derived from field studies as outlined in Table x2.
+Alpha-cypermethrin (pyrethroid) is shown in red and Bendiocarb (carbamate) is shown in blue.
+Exposure is set to 0.8 and starting frequencies to 0.01.
+Resistance to the carbamate (blue) increases very quickly in both the mixture and sequence due to it's high effectiveness.
+Resistance to the pyrethroid (red) increases more slowly due to low effectiveness and high cost of resistance.
+In a mixture the resistance frequency of the pyrethroid even declines when resistance to the carbamate is low.
+In this illustrative example resistance to both insecticides takes longer to evolve for a mixture than a sequence."}
+
+# add side panels for labels
+#layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
+layout( matrix(c(1:2), byrow = TRUE), heights=c(1,0.15) )
+
+#plot margins
+par(mar=c(1.5, 0, 2, 0)) #b,l,t,r default c(5, 4, 4, 2)
+
+# runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.3 , h.RS2_0B = 0.66 , exposure = 0.5 , phi.SS1_A0 = 0.43 , phi.SS2_0B = 0.98 , rr_restoration_ins1 = 0.63 , rr_restoration_ins2 = 0.84 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, labelMixSeqRatio = 1 )
+
+# Alpha-cypermethrin (pyr) incl cost, exposure 0.8
+runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01, z.RR1_00 = 0.19, h.RS1_00 = 1,  h.RS1_A0 = 0.54 , h.RS2_0B = 0.66 , exposure = 0.8 , phi.SS1_A0 = 0.29 , phi.SS2_0B = 0.98 , rr_restoration_ins1 = 1 , rr_restoration_ins2 = 0.84 , addCombinedStrategy = FALSE, strategyLabels = c('s','','a','m'), cex.axis = 0.8, addLegend=FALSE, labelMixSeqRatio = 1 )
+
+#adding legend in lower panel
+#plot margins
+par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
+
+plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m')
+
+legend( 'bottom', legend=c("pyrethroid alone   ", "pyrethroid in mix   ", "carbamate alone   ", "carbamate in mix   "), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.7, horiz =TRUE )
+
+```
+
 ```{r Figxx not eval, eval=FALSE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="A.

### 2018-01-02 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
renumbered figures in legends

@@
 
-```{r Fig4, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 4.
+```{r Fig5, eval=TRUE, echo=FALSE, fig.height=3.7, fig.width=7.5, fig.show='hold', fig.align='left', fig.cap="Figure 5.
 Single insecticide effects of starting frequency and cost of resistance on resistance frequency.
@@
 
-```{r Fig5, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 5.
+```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 6.
 Influence of insecticide effectiveness and exposure on time-to-resistance for mixtures and sequences.
@@
 
-```{r Fig6, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 6.
+```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 7.
 Influence of the effectiveness of both insecticides on time-to-resistance for mixtures and sequences.
@@
 The effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
+The 's' and 'm' on the upper X axis and 'mix/seq' in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 6.
 A.
@@
 
-```{r Fig7, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 7.
+```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 8.
 Influence of dominance and resistance restoration on time-to-resistance for mixtures and sequences.
@@
 Resistance restoration for the allele coding for resistance to insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
+The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 6.
 A.
@@
 
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 8.
+```{r Fig9, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 9.
 Influence of starting frequencies of resistance on time-to-resistance for mixtures and sequences.
@@
 Effectiveness of insecticide 1 is increased from column 1 (A and C) to column2 (B and D).
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
+The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 6.
 A.
@@
 
-```{r Fig9, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 9.
+```{r Fig10, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Figure 10.
 Influence of cost of resistance on time-to-resistance for mixtures and sequences.
 Row 1 (A and B) same as Fig.
-5 with no cost of resistance, row 2 (C and D) cost of resistance set to 0.15 with dominance of cost set to 0.5.
-The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig 5.
+6 with no cost of resistance, row 2 (C and D) cost of resistance set to 0.15 with dominance of cost set to 0.5.
+The 's' and 'm' on the upper X axis and mix/seq in the lower right of each panel stand for sequence and mixture and are explained in the legend to Fig.
+6.
 The increase in cost of resistance from A.
@@
 ```{r Fig11, eval=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.align='left', fig.cap="Figure 11.
-Time-to-resistance for a mixture and sequence using inputs derived from field studies as outlined in Table x2.
+Time-to-resistance for a mixture and sequence using inputs derived from field studies as outlined in Table 6.
 Alpha-cypermethrin (pyrethroid) is shown in red and Bendiocarb (carbamate) is shown in blue.

### 2018-01-02 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
renumber figure references in text

@@
 #!!!!  (spacing & line numbers work for pdf output but not for word)
-#!!!! submission figs, set this chunk to eval=TRUE
-#!!!! previous drafts knitted to pdf with pdf_document in YAML and this chunk eval=FALSE
-
-# options to create final publication quality figures
-# causes the document to fail, but the figs do get stored in pap2figs/
 library(knitr)
@@
 
-The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 1.
+The calculation of fitness for each genotype when considering just a single insecticide is illustrated in Figure 2.
 In the model itself this is repeated for both insecticides and the fitnesses multiplied.
@@
 For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
-Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Fig 1).
+Dominance of cost determines how the fitness for SR sits between that of SS and RR in the absence of the insecticide (Fig 2).
 
 The simulation proceeds through non-overlapping generations.
-In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness, the latter depending on their absolute fitness (Fig 1) and level of exposure.
+In each generation selection is represented by multiplying genotype frequencies in the population by their relative fitness, the latter depending on their absolute fitness (Fig 2) and level of exposure.
 This acts to make the fitter alleles more common over time.
@@
 The model is implemented in R [@RCoreTeam2016], the code is open-source and hosted on Github [@South2017] including the code to generate both the figures and text of this paper.
-We provide an on-line user interface [@South2017a] (Fig 2) accessing the same code and enabling readers with no coding experience to change inputs and run the model themselves.
+We provide an on-line user interface [@South2017a] (Fig 3) accessing the same code and enabling readers with no coding experience to change inputs and run the model themselves.
 
@@
 ### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 3 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
+For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 4 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
 This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
-Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 4A) through the even simpler mechanism that a smaller cahnge in frequency is required to reach the threshold.
-
-Cost of resistance (Fig 4B) and its dominance (not shown) had the opposite effect.
+Similarly, higher values of starting resistance frequency also resulted in shorter times to resistance thresholds (Fig 5A) through the even simpler mechanism that a smaller change in frequency is required to reach the threshold.
+
+Cost of resistance (Fig 5B) and its dominance (not shown) had the opposite effect.
 Higher values led to a slower development of resistance.
@@
 When two insecticides are used we can compare the relative performance of concurrent use in a mixture with sequential use.
-Each panel in Figures 5-9 compares mixture and sequence for a single combination of inputs.
+Each panel in Figures 6-10 compares mixture and sequence for a single combination of inputs.
 To help understand the mechanisms influencing this relative performance we start with a single base scenario and demonstrate changing inputs individually.
 In this base scenario we set effectiveness, exposure, resistance restoration, and dominance at 0.5, the starting frequencies of resistance to 0.01 and cost of resistance to 0.
-Resistance arises slower for sequential than mixture for this base scenario (Fig 5A).
+Resistance arises slower for sequential than mixture for this base scenario (Fig 6A).
 Resistance to both insecticides in the mixture follow the same path and reach the threshold at the same time (as would be expected given that they have identical input parameters).
-When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 5A).
+When used in sequence the curve (dashed) for each insecticide individually is steeper than in the mixture but, because they happen one after the other, it takes longer for both to reach the resistance threshold (Fig 6A).
 Thus the time-to-resistance for the mixture divided by the sequence is less than 1 at 0.8.
@@
 The effect on the mixture is greater so time-to-resistance remains longer for the sequence (the ratio decreasing from 0.8 to 0.6).
-In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 5B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
+In contrast keeping the exposure constant and increasing effectiveness (the proportion of SS insects that are killed by contact) of one of the insecticides (Fig 6B) results in a longer time-to-resistance for the mixture relative to the sequence (and the ratio increasing from 0.8 to 1.2).
 The longer time-to-resistance for the mixture results from a slower increase in resistance for the less effective insecticide.
@@
 
-The result is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 5A & B).
-This is opposite to what happens when used in sequence (compare the red dashed line in Fig 5A & B) and in isolation (Fig 2A), when more effective insecticides shortened times-to-resistance.
-
-In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 5, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 5.
+The result is that a more effective insecticide increases time-to-resistance when used in a mixture (compare the solid lines in Fig 6A & B).
+This is opposite to what happens when used in sequence (compare the red dashed line in Fig 6A & B) and in isolation (Fig 4A), when more effective insecticides shortened times-to-resistance.
+
+In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 6, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 6.
 Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
@@
 Increasing the effectiveness further of either or both insecticides increases the performance of mixtures over sequences in terms of time-to-resistance.
-Figure 6A uses the same inputs as Fig 5B.
-From this scenario a higher effectiveness for insecticide2 (Fig 6C) or insecticide 1 (Fig 6B) or both (Fig 6D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
+Figure 7A uses the same inputs as Fig 6B.
+From this scenario a higher effectiveness for insecticide2 (Fig 7C) or insecticide 1 (Fig 7B) or both (Fig 7D) all result in a greater positive difference in time-to-resistance for mixture over sequence.
 This again points to how more effective insecticides can slow resistance to another when used in a mixture.
@@
 #### Resistance restoration and dominance
-Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 7) in the same way that it did for sole use (Fig 3).
-The result is that increasing either or both of dominance and resistance restoration (Figs 7B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 7A) where it takes resistance longer to develop for the sequence.
+Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 8) in the same way that it did for sole use (Fig 4).
+The result is that increasing either or both of dominance and resistance restoration (Figs 8B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 8A) where it takes resistance longer to develop for the sequence.
 So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
@@
 Changing the starting frequency of resistance had a similar effect on time to resistance for both sequences and mixtures and thus had little effect on their relative performance.
-For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 8C to 8A).
-Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 8B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 8D).
+For example taking the base scenario and reducing the starting frequency of one resistance allele did not change from sequence being favoured (compare Fig 9C to 9A).
+Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 9B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 9D).
 
 #### Costs of resistance
-Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 9C, the red line is for the resistance with the higher cost).
-The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 9C).
-The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 9A (and Fig 5A) is removed in Fig 9C.
-With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 9A and 1 in Fig 9C.
-At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 9D to 9B.
+Increasing the cost of resistance to one insecticide in a mixture leads resistance to that insecticide to evolve more slowly (compare the solid red line to the solid blue line in Figure 10C, the red line is for the resistance with the higher cost).
+The resistance with the higher cost also increases more slowly in the mixture relative to when it is used in sequence (compare the solid red line to the dotted red line in Fig 10C).
+The greater effect of cost on mixtures means the advantage of sequences over mixtures in Fig 10A (and Fig 6A) is removed in Fig 10C.
+With costs set to 0 time-to-resistance for mixture divided by sequence is 0.8 in Fig 10A and 1 in Fig 10C.
+At higher insecticide effectiveness the benefit of mixture over sequence is improved as seen by comparing Fig 10D to 10B.
 
@@
 It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 8 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 8 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
+Figure 9 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
+In Figures 9 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
 
@@
 
-When single insecticide use was investigated (Figs 3-4) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 4B).
+When single insecticide use was investigated (Figs 4-5) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 5B).
 Mechanisms explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
-Referring to figure 1 can help explain these responses.
+Referring to Figure 2 can help explain these responses.
 Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
@@
 These differences can help us to understand the dynamics of resistance evolution.
-In each panel of Figures 5-9 the solid lines, indicating that an insecticide is being used at the same time as another in a mixture, are always shallower than the same colour dashed line indicating the insecticide is being used in isolation as part of a sequence.
+In each panel of Figures 6-10 the solid lines, indicating that an insecticide is being used at the same time as another in a mixture, are always shallower than the same colour dashed line indicating the insecticide is being used in isolation as part of a sequence.
 The shallower curves for the mixture can be explained by each insecticide killing individuals that are resistant to the other insecticide.
@@
 Resistance can evolve more slowly for both insecticides in sequential use, despite being faster for each, because they occur one after the other.
-In Figure 5A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
+In Figure 6A it takes 60 generations for resistance to reach the threshold for each insecticide in the sequence, and 90 when they are applied together in the mixture, so the sequence is still favoured because 2x60 is greater than 90.
 
 This advantage of sequence over mixtures can be removed by increasing the effectiveness of one of the insecticides.
-Increasing the effectiveness of one insecticide (Fig 5B) decreases the total time for the sequence from 120 to 90 and increases that for the mixture from 90 to 115 thus changing the order and favouring the mixture.
-The mechanism for this change can be seen by comparing the shape of the curves in Figs 5A & B.
+Increasing the effectiveness of one insecticide (Fig 6B) decreases the total time for the sequence from 120 to 90 and increases that for the mixture from 90 to 115 thus changing the order and favouring the mixture.
+The mechanism for this change can be seen by comparing the shape of the curves in Figs 6A & B.
 In the mixture, resistance to insecticide 1 (with the increased effectiveness) rises at a similar speed in both figures.
@@
 Once the first insecticide reaches the resistance threshold of 50% it kills fewer individuals that are resistant to the second insecticide, it loses its protective effect and thus stops slowing the rise in resistance to the second insecticide.
-A similar effect is visible when the effectiveness of both insecticides are increased (Fig 6C).
-The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 5A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 5B.
+A similar effect is visible when the effectiveness of both insecticides are increased (Fig 7C).
+The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 6A) because more individuals resistant to each insecticide are being killed by the other insecticide.
+In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 6B.
 In both cases increasing the effectiveness of insecticides increases their protective effect when used in mixtures and favours mixtures over sequences.
 
-Fig 5B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide (the red curve in this case).
+Fig 6B also demonstrates that when two insecticides have a different effectiveness in a mixture, other parameters being equal, resistance will be expected to increase faster to the more effective insecticide (the red curve in this case).
 The more effective insecticide prompts both a) higher selection pressure for its own resistance and b) greater 'protection' reducing the rise in resistance to the other insecticide.
@@
 It seems initially counter-intuitive that whereas increasing effectiveness slows the development of resistance for mixtures, increasing exposure speeds it up.
-Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 5A to 40 generations in Fig 5C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 5A to 115 in Fig 5B).
-This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 3 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 5B & 5C).
+Increasing the exposure to both insecticides results in a decrease in time-to-resistance (from 90 generations in Fig 6A to 40 generations in Fig 6C) where increasing the effectiveness of both insecticides by the same amount results in a slight increase (from 90 in Fig 6A to 115 in Fig 6B).
+This contrasts with the identical effects that exposure and effectiveness have on a single insecticide (Figs 4 A,B) and on insecticides in sequence (the red dashed lines are the same in Figs 6B & C).
 This difference, that exposure reduces times-to-resistance for a mixture and effectiveness increases it, was unexpected for us and we are keen to explore it in more detail.
@@
 Fitness costs of resistance have been widely reported(review in [@Kliot2012]) but it has been suggested that few of these are directly relevant to control failure in the field [@Ffrench-Constant2017].
-Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 9).
-
-Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 9C and D).
+Because of this likely rarity of costs we did not include them in most of our analyses (with the exception of Fig 10).
+
+Costs of resistance favour mixtures in these analyses because the evolution of resistance is slowed more by cost in a mixture than in a sequence (compare the red solid lines for a mixture to the red dotted line for a sequence in Fig 10C and D).
 It makes sense that the protective effect of the other insecticide in a mixture combines with the cost of resistance to reduce selection pressure more than in a sequence.
 
-The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 9C).
+The resistance curves for two insecticides in a mixture that differ only in their cost of resistance, diverge more as the frequency of resistance increases (Fig 10C).
 This makes intuitive sense, as the frequency of resistance genes increases the impact of fitness costs associated with those genes also increases.
@@
 This result is consistent with our prediction that the very high effectiveness is likely to favour mixtures over sequences (e.g.
-see Fig 6D and Table 4).
+see Fig 7D and Table 4).
 
@@
 One aspect of the results presented here can seem to contradict published advice and has been questioned when we have presented model results.
-Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 3 & 5, Table 2).
+Our model predicts that lower insecticide effectiveness or exposure will lead to slower spread of resistance in sole use and sequences (Figs 4 & 6, Table 2).
 This is consistent with expectations that lower effectiveness and exposure would lead to lower mortality of susceptibles and thus lower selection pressure for resistance [e.g.
@@
 Poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 3C, 7).
+Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 4C, 8).
 So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
@@
 
-In addition we provide an on-line user interface to the model (Fig 2) [@South2017a] enabling readers with no coding experience to investigate model behaviour in areas of parameter space we have not shown.
+In addition we provide an on-line user interface to the model (Fig 3) [@South2017a] enabling readers with no coding experience to investigate model behaviour in areas of parameter space we have not shown.
 The user-interface allows inputs to be set for two scenarios and results to be viewed side by side.
-For example you can set a higher insecticide effectivenesses in one scenario to see how this favours mixtures over sequences (Fig 2) or increase the exposure levels to see how this favours sequences.
+For example you can set a higher insecticide effectivenesses in one scenario to see how this favours mixtures over sequences (Fig 3) or increase the exposure levels to see how this favours sequences.
 

### 2018-01-02 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
starting to address reviewers comments

@@
 Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
-The fitness of the SS genotype in absence of the insecticide is used as the reference genotype and assigned a value of 1.
+The fitness of the SS genotype in the absence of the insecticide is used as the reference genotype and assigned a value of 1.
 All other fitnesses are assumed to be less than 1, determined by alleles, exposure to the insecticide and other inputs that can be set in the model.
@@
 Separate sexes and standard sexual reproduction with recombination is included.
-There is also the potential to set exposure rates to be different for males and females to represent their different behaviours but we do not explore that parameter here.
+There is also the potential to set exposure rates to be different for males and females to represent their different behaviours.
+Females are more likely to come into contact with insecticides on walls and nets when seeking to feed on humans compared to males that do not.
+We only explore that input briefly here because a previous analysis indicated it was not a major factor determining differences between insecticide-use strategies [@Levick2017].
 
@@
 1. single insecticide
-1. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
-1. two insecticides used in a mixture (concurrently)
-
-A major aim of this paper is to us modern techniques to make the modelling accessible to a wider community.
-The model is implemented in R [@RCoreTeam2016], the code is open-source and hosted on Github [@South2017] including the code to generate both the figures and text of this paper.
+2. two insecticides used in sequence, replacing the first with a second once the frequency of resistance alleles reaches a threshold of 50%
+3. two insecticides used in a mixture (concurrently)
+
+A major aim of this paper is to use modern techniques to make the modelling accessible to a wider community.
+The model is implemented in R [@RCoreTeam2017], the code is open-source and hosted on Github [@South2017] including the code to generate both the figures and text of this paper relying on the packages knitr[@Xie2017], shiny[@Chang2017] and ggplot2[@Wickham2009].
 We provide an on-line user interface [@South2017a] (Fig 3) accessing the same code and enabling readers with no coding experience to change inputs and run the model themselves.
@@
 ### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and its dominance all resulted in faster resistance spread (Fig 4 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
+For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and dominance of restoration all resulted in faster resistance spread (Fig 4 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
 This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
@@
 In summary increasing effectiveness can favour mixtures moving from column 1 to 2 in Figure 6, and increasing exposure can favour sequences moving from row 1 to 2 in Figure 6.
-Reduced male exposure with respect to female, as might be expected for male mosquitoes not seeking blood meals, had a similar effect to changing overall exposure.
-Reducing male exposure (not shown) increased times-to-resistance and favoured mixtures over sequences.
+We do not show the results here, but reduced male exposure with respect to female, as would be expected for male mosquitoes not seeking blood meals, had a similar effect to reducing overall exposure.
 
@@
 #### Linkage disequilibrium
-It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through a process termed linkage disequilibrium.
-Figure 9 C&D show how our model predicts that linkage disequilibrium does not lead to more rapid evolution of resistance to the newer insecticide.
-In Figures 9 C&D the newer insecticide with lower starting resistance (in red) leads to faster resistance when used alone (red dotted) than when used in a mixture (red solid).
+Positive linkage disequilibrium is said to exist when the frequency of two alleles at different loci is higher than expected if they were just selected individually.
+It has previously been suggested [@Curtis1985][@WHO2012] that using an insecticide to which there was existing resistance in a mixture with a new insecticide with little resistance could lead to more rapid evolution of resistance to the latter through linkage disequilibrium.
+We explicitly look at the example of a mixture of a new insecticide and one with pre-existing resistance (Figures 9 C&D).
+In both cases resistance to the newer insecticide with lower starting resistance (in red) increases faster when it is used alone (red dotted) than when it is used in a mixture (red solid).
+This suggests that linkage disequilibrium does not restrict the potential benefits of mixtures in all situations.
+These results support the statement in GPIRM [1] that : "Even though the risk for linkage disequilibrium exists, mixtures may still be the most attractive IRM tool …".
+The dynamics of LD is too complex to discuss in detail here but interested readers can find more discussion in [@Curtis1985][@WHO2012][@Levick2017].
 
@@
 that insecticides used together perform better or worse than when used separately.
+This would allow the model to represent cross-resistance where a resistance gene provides resistance to more than one insecticide.
+Similarly behavioural resistance is likely to offer resistance against more than one insecticide [@GPIRM2012] but until there are more data on whether behavioural resistance is genetic and heritable we would not recommend representing it in this way.
 

### 2018-01-02 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
new section on starting frequencies

@@
 Similarly taking a scenario in which time-to-resistance is longer for a mixture (Fig 9B) and decreasing the starting frequency of that resistance allele did not change the fact that the mixture was favoured (Fig 9D).
+In these analyses we set starting frequencies of resistance relatively high at 0.01 or 0.001.
+In our more detailed sensitivity analysis [@Levick2017] the lower limit for starting frequencies was 0.0001 and still it was not a key parameter in determining the relative performance of mixtures and sequences.
 
@@
 This suggests that linkage disequilibrium does not restrict the potential benefits of mixtures in all situations.
-These results support the statement in GPIRM [1] that : "Even though the risk for linkage disequilibrium exists, mixtures may still be the most attractive IRM tool …".
+These results support the statement in GPIRM [@WHO2012] that : "Even though the risk for linkage disequilibrium exists, mixtures may still be the most attractive IRM tool …".
 The dynamics of LD is too complex to discuss in detail here but interested readers can find more discussion in [@Curtis1985][@WHO2012][@Levick2017].
@@
 This would allow the model to represent cross-resistance where a resistance gene provides resistance to more than one insecticide.
-Similarly behavioural resistance is likely to offer resistance against more than one insecticide [@GPIRM2012] but until there are more data on whether behavioural resistance is genetic and heritable we would not recommend representing it in this way.
+Similarly behavioural resistance is likely to offer resistance against more than one insecticide [@WHO2012] but until there are more data on whether behavioural resistance is genetic and heritable we would not recommend representing it in this way.
 

### 2018-01-03 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
add few reponses to reviewer

@@
 This mechanistic understanding can contribute to the debate on the relative merits of different insecticide strategies and extend existing frameworks [@Roush1989][@IRAC2011][@FAO2012][@WHO2012][@Consortium2013].
+By making the model more accessible we also hope to help bridge gaps between academia and policy in insecticide resistance management [@Huijben2017].
 
@@
 The combination of the two genes is termed the genotype.
+There are 3 genotypes when considering one insecticide and 9 genotypes (e.g.
+SSRR) when considering two insecticides (Fig 1).
 Fitness is the main currency of the model, representing how much each genotype survives and reproduces.
@@
 Resistance to the less effective insecticide in the mixture increases slowly initially and then speeds up after the more effective insecticide has reached the resistance threshold.
-This points a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
+This points to a mechanism whereby the more effective insecticide initially slows the rate of evolution of resistance to the less effective one.
 
@@
 Increasing resistance restoration or its dominance decreases time-to-resistance for both sequences and mixtures (Fig 8) in the same way that it did for sole use (Fig 4).
-The result is that increasing either or both of dominance and resistance restoration (Figs 8B,C,D) doesn't change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 8A) where it takes resistance longer to develop for the sequence.
+The result is that increasing either or both of dominance and resistance restoration (Figs 8B,C,D) does not change the relative time-to-resistance for mixtures and sequences from that in the base scenario (Fig 8A) where it takes resistance longer to develop for the sequence.
 So, unlike effectiveness and exposure, the levels of resistance restoration and dominance do not effect whether a mixture or a sequence are likely to be favoured.
@@
 The identical resistance curves for each insecticide in the mixture are shallower than at the lower effectiveness (Fig 6A) because more individuals resistant to each insecticide are being killed by the other insecticide.
-In this case the 'protection' given to both insecticides by the other declines at the same rate so we don't see the change in slope seen in Fig 6B.
+In this case the 'protection' given to both insecticides by the other declines at the same rate so we do not see the change in slope seen in Fig 6B.
 In both cases increasing the effectiveness of insecticides increases their protective effect when used in mixtures and favours mixtures over sequences.
@@
 This is a combination of the low effectiveness creating a low selection pressure for the resistance and the high cost creating a selection pressure against it.
-The low effectiveness and high cost are partly due to the surprisingly high mortality in unexposed mosquitoes in the field data [@Kolaczinski2000].
+The low effectiveness and high cost are partly due to the surprisingly high mortality in unexposed mosquitoes in the field data we used [@Kolaczinski2000].
 
@@
 Consistent with our results [@Roush1998] also showed (Figure 3) that decreasing exposure favours mixtures over sequences.
+
+#### Caveats
+There is evidence of the involvement of multiple genes in resistance mechanisms[@Ffrench-Constant2004][@Ffrench-Constant2013][@Gressel2017].
+Our model, in common with almost all previous ones [@Bourguet2010], assumes a single ‘resistance’ gene per insecticide.
+In many cases we believe this is reasonable.
+Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Roush1989][@Ffrench-Constant2004][@Ffrench-Constant2013].
+One view is that control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
+There is also evidence that low insecticide doses in agriculture have led to polygenic resistance [@Gressel2017].
+To represent polygenic resistance a different 'quantitative genetic' modelling approach would be required [@Via1986][@Firko1990].
+A recent review of 187 models of the evolution of resistance [@Bourguet2010] did not find any that represented quantitative multiple gene resistance.
+We are only aware of limited published work applying a quantitative genetic approach to assessing insecticide use strategies in the context of polygenic resistance [@Via1986][@Gardner1998][@Gardner1999].
+If it becomes clearer that polygenic resistance is an issue for vector control then there will be a need for appropriate quantitative genetic models.
+We would be keen to contribute to that development.
 
@@
 Poor implementation would likely cause lower effectiveness of, and exposure to, insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-The explanation for this apparent difference is in the effect of dominance.
-Poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].
+There are two main explanations for this apparent difference, a) dominance and b) polygenic resistance.
+The effect of dominance is that poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].
 Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 4C, 8).
@@
 Similarly, the effects of poor application on the relative benefits of mixtures and sequences will depend on these trade-offs because of our result that higher effectiveness favours mixtures, higher exposure favours sequences and higher dominance favours neither (Table 4).
-More research into how the mortality of different genotypes (susceptible, resistant and heterozygous) is affected by poor application rates or insecticide degradation over time would help make this clearer.
+More research on how the mortality of different genotypes (susceptible, resistant and heterozygous) is affected by poor application rates or insecticide degradation over time would help make this clearer.
+Secondly, if resistance is predominantly polygenic rather than monogenic poor implementation is likely to promote resistance by allowing the build up of many genes of small effect [@Gressel2017].
+Indeed modelling work from 15 years ago [@Gardner1998][@Gardner1999] proposes an alternation of low and high doses with the low doses limiting the evolution of monogenic resistance and the high doses limiting that of the polygenic resistance.
+More research on the relative importance of monogenic and polygenic resistance for vector control would help here.
 
@@
 In both cases our modelling approach could be used to investigate implications for the development of resistance.
-
-#### Caveats
-There is evidence of the involvement of multiple genes in resistance mechanisms[@Ffrench-Constant2004][@Ffrench-Constant2013].
-Our model, in common with almost all previous ones [@Bourguet2010], assumes a single ‘resistance’ gene per insecticide.
-In most cases we believe this is reasonable.
-Target-site resistance has been shown to be mostly determined by single genes, and control failures in general have mostly been attributed to the effect of single major genes [@Roush1989][@Ffrench-Constant2004][@Ffrench-Constant2013].
-Control doses of insecticides in the field are likely to lead to the selection of single major genes whereas the lower doses used to select laboratory strains are more likely to lead to polygenic resistance [@Ffrench-Constant2013].
-To represent polygenic resistance a different 'quantitative genetic' approach would be required [@Via1986][@Firko1990].
-A recent review of 187 models of the evolution of resistance [@Bourguet2010] did not find any that represented quantitative multiple gene resistance.
-We are only aware of preliminary published work applying a quantitative genetic approach to assessing insecticide use strategies in the context of polygenic resistance [@Via1986][@Gardner1998].
-If it becomes clearer that in some cases polygenic resistance is more of an issue than previously thought then there will be a need for appropriate quantitative genetic models.
-We would be keen to contribute to that development.
 
@@
 #### Conclusions
-We have developed a flexible, modern model of the evolution of insecticide resistance that can that help us understand the mechanisms by which different insecticide application strategies may be favoured.
+We have developed a flexible, modern model of the evolution of insecticide resistance that can help us understand the mechanisms by which different insecticide application strategies may be favoured.
 It replicates previous modelling work and has the potential to investigate further operational strategies for established and new insecticides.
@@
 ##### Acknowledgements
-Thank you to Bethany Levick who developed the first version of the model code.
+Thank you to Bethany Levick who developed the first version of the model code in collaboration with IH.
+Thank you to David Malone, Justin McBeath Hilary Ranson, Sarah Rees, Graham Small and an anonymous reviewer for commenting constructively on an earlier draft of this paper.
+Thank you to Brian Bourke and phylopic.org for the mosquito image used in Figure 1.
 

### 2018-01-04 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
final resubmission version (minor post knit edits in word)

@@
 
-When single insecticide use was investigated (Figs 4-5) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution resistance, whereas higher values for cost led to slower evolution of resistance (Fig 5B).
+When single insecticide use was investigated (Figs 4-5) higher values of parameters that increased the selective advantage of resistance all led to more rapid evolution of resistance, whereas higher values for cost led to slower evolution of resistance (Fig 5B).
 Mechanisms explaining the evolution of resistance to single insecticide use, are summarised in Table 2.
@@
 Poor implementation would likely cause lower effectiveness of, and exposure to, insecticides, so how can this increase the potential for resistance when our model suggests it would be decreased ?
-There are two main explanations for this apparent difference, a) dominance and b) polygenic resistance.
+There are two main potential explanations for this apparent difference, a) dominance and b) polygenic resistance.
 The effect of dominance is that poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].

### 2018-01-15 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
change dominance of restoration to dominance of resistance

@@
 3. Resistance restoration  | ability of resistance (RR) to restore fitness of insects exposed to insecticide
-4. Dominance of restoration | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
+4. Dominance of resistance | sets fitness of heterozygous (SR) insects between that of SS & RR in presence of insecticide
 5. Frequency  | frequency of resistance alleles within the population
@@
 
-Dominance of restoration determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
+Dominance of resistance determines how the fitness for SR sits between that of SS and RR for those that are exposed to the insecticide.
 For those that are not exposed, fitness of SS is set to 1 by definition and resistance cost determines the fitness of RR.
@@
 ### Single insecticide
-For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and dominance of restoration all resulted in faster resistance spread (Fig 4 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
+For single insecticide use, higher values of insecticide effectiveness, exposure, resistance restoration and dominance of resistance all resulted in faster resistance spread (Fig 4 A-D) and thus fewer generations to reach a resistance threshold of 0.5.
 This makes intuitive sense as all increase the impact of the insecticide and thus increase selection for resistance.
@@
 2. Exposure | faster | reduced fitness of SS & SR overall
-3. Dominance of restoration | faster | increased fitness of SR exposed to insecticide
+3. Dominance of resistance | faster | increased fitness of SR exposed to insecticide
 4. Resistance restoration | faster | increased fitness of RR & SR exposed to insecticide
@@
 However at the same time selection pressure is reduced by reduced fitness of resistant individuals caused by the other insecticide
-3. Dominance of restoration | faster | increased fitness of heterozygotes
+3. Dominance of resistance | faster | increased fitness of heterozygotes
 4. Resistance restoration | faster | increased fitness of resistants
@@
 2. Exposure | sequence | Higher exposure gives faster resistance for sequence and mixture but the greater effect on mixture favours sequence.
-3. Dominance of restoration | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
-4. Resistance restoration | neither | As for dominance of restoration. 
+3. Dominance of resistance | neither | Higher dominance gives faster resistance in both sequences and mixtures such that the difference between them is not changed. 
+4. Resistance restoration | neither | As for dominance of resistance. 
 5. Frequency  | neither | As for dominance and resistance restoration. 
@@
 3. Resistance restoration  | (RRfit-SSfit) / Effectiveness in presence of insecticide
-4. Dominance of restoration | (SRfit-SSfit)/(RRfit-SSfit) in presence of insecticide
+4. Dominance of resistance | (SRfit-SSfit)/(RRfit-SSfit) in presence of insecticide
 5. Frequency  | estimated
@@
 Resistance restoration|0.60|1|0.84
-Dominance of restoration|0.52|0.54|0.66
+Dominance of resistance|0.52|0.54|0.66
 Cost of resistance|-|0.19|-
@@
 Referring to Figure 2 can help explain these responses.
-Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of restoration acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
+Exposure and effectiveness act by decreasing the fitness of SS and SR more than RR, dominance of resistance acts by increasing fitness of SR and resistance-restoration acts by increasing fitness of RR & SR.
 Cost acts by decreasing the fitness of RR and SR among those mosquitoes  not exposed.
@@
 The effect of dominance is that poor application can reduce the mortality of heterozygotes (SR) thus effectively increasing the dominance of the resistant allele and increasing its rate of evolution [@Bourguet2000][@Barbosa2011][@FAO2012] see Fig 1 of [@Levick2017].
-Our model results are consistent with this, showing that increasing dominance of restoration leads to faster spread of resistance (Fig 4C, 8).
+Our model results are consistent with this, showing that increasing dominance of resistance leads to faster spread of resistance (Fig 4C, 8).
 So, the effect of poor implementation of an insecticide intervention on resistance could be either positive or negative depending on the relative effects on dominance of the resistance gene, insecticide effectiveness and exposure.
@@
 Firstly the exposure input determines the proportion of the population in the left and right panels (exposed and not exposed).
-For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of restoration determines how the fitness for SR lies between that of SS and RR.
+For those that are exposed (left panel) insecticide effectiveness sets the fitness for SS, resistance restoration 'restores' a portion of the fitness for RR and dominance of resistance determines how the fitness for SR lies between that of SS and RR.
 For those that are not exposed, fitness of SS is set to 1 by definition, resistance cost determines the fitness of RR and again dominance of cost determines how the fitness for SR sits between that of SS and RR.
-In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of restoration=0.7 which sets the fitness of the SR closer to RR than SS.
+In this example effectiveness=0.8, resistance restoration=0.5 which 'restores' half of the fitness lost due to the insecticide, dominance of resistance=0.7 which sets the fitness of the SR closer to RR than SS.
 Resistance cost=0.3 which reduces fitness in the absence of the insecticide from 1 to 0.7, and dominance of cost=0.8 which sets fitness of SR close to RR."}
@@
 Exposure, C.
-Dominance of restoration, D.
+dominance of resistance, D.
 Resistance restoration.
@@
          guides(colour = guide_legend(reverse=TRUE)) + 
-         labs(colour = "dominance of restoration") +
+         labs(colour = "dominance of resistance") +
          coord_trans(y = "log10") + 

### 2018-01-15 R/plot_fit_calc.r

@@ -51,7 +51,7 @@ plot_fit_calc <- function ( effectiveness=0.75,
 
   #set up different labels for dominance in simple & other plot
   if (simple) {dom_rest_lab <- dom_cost_lab <- " dominance"} else {
-               dom_rest_lab <- " dominance of\n  restoration"
+               dom_rest_lab <- " dominance of\n  resistance"
                dom_cost_lab <- " dominance of\n  cost" }
     
   

### 2018-01-16 R/plot_fit_calc.r
fix R CMD check errors and warnings

@@ -14,10 +14,11 @@
 #' 
 #' #grid needed for title
 #' @import ggplot2 grid
+# in check: namespace dependency not required
 #' @importFrom gridExtra grid.arrange 
 #' 
 #' @examples 
-#' plot_fit_calc(effectiveness=0.5, resistance_restoration=0.5, dominance=0.5)
+#' plot_fit_calc(effectiveness=0.5, resistance_restoration=0.5, dominance_restoration=0.5)
 #' 
 #' 
 #' @return gtable plot object

### 2018-01-16 R/plot_fit_variance.r

@@ -3,8 +3,8 @@
 #' I could try to get this to accept multiple scenarios & put them on the same graph.
 #' 
 #' @import ggplot2
-# taking out because causing shinyapps problem  
-# @importFrom cowplot plot_grid
+# if in here must be in imports too taking out because causing shinyapps problem  
+#' @importFrom cowplot plot_grid
 #' @importFrom tidyr gather
 #' @examples 
 #' input <- setInputOneScenario()
@@ -74,8 +74,8 @@ plot_fit_variance <- function ( listOut,
   {
     # to remove x axis label from top plot
     plotlist[[1]] <- plotlist[[1]] + xlab("")
-    warning("cowplot::plot_grid disabled to fix shinyapps problem")
-    #cowplot::plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(1,1), labels='AUTO')    
+    #warning("cowplot::plot_grid disabled to fix shinyapps problem")
+    cowplot::plot_grid( plotlist[[1]],plotlist[[2]],ncol=1, rel_heights=c(1,1), labels='AUTO')    
   } else
   {
     print(plotlist[[1]])

### 2018-01-16 R/plot_ld_gpirm.r

@@ -1,10 +1,10 @@
 #' try to plot genotype frequencies to show linkage disequilibrium
 #' 
 #' as done in GPIRM p122
-#' BEWARE this initial version assumes dominance=1
+#' BEWARE this initial version assumes dominance of resistance=1
 
 #' @param genotype genotype frequencies by generation
-#' @param gen_num which generation to plot for
+#' @param gen which generation to plot for
 #' @param labgen whether to add label for generation number
 #' 
 #' @examples 
@@ -27,10 +27,10 @@
 #' listOut <- runModel2(input)
 #' plot_ld_gpirm(genotype=listOut$genotype[[1]], gen=seq(from = 5, to = 50, by =5))
 #' 
-#' TODO this seems to expose an error in my plotting routine
-#' OR can this plot idea not cope with our data ?
-#' is this caused by having different effectivenesses for each insecticide ?
-#' SEE whitespace between generations 20 & 40
+#' #TODO this seems to expose an error in my plotting routine
+#' #OR can this plot idea not cope with our data ?
+#' #is this caused by having different effectivenesses for each insecticide ?
+#' #SEE whitespace between generations 20 & 40
 #' #exposure 0.5, eff1 0.8, eff2 0.5, dominances 1
 #' a <- setExposure( exposure=0.5,  insecticideUsed = 'mixture' )
 #' input <- setInputOneScenario( a=a, phi.SS1_A0=0.8, phi.SS2_0B=0.5, h.RS1_A0=1, h.RS2_0B=1)

### 2018-01-16 R/resistance-package.r

@@ -15,3 +15,7 @@
 #' @importFrom utils read.csv write.csv
 
 NULL
+
+## quiets concerns of R CMD check re: the .'s that appear in pipelines
+# copied from : https://github.com/jennybc/googlesheets/blob/master/R/googlesheets.R
+if(getRversion() >= "2.15.1")  utils::globalVariables(c("."))

### 2018-01-23 inst/documents/paper2/summary_fig_pap2.Rmd
create a labelled summary fig for paper2

---
title: "summary figure for malaria journal paper"
output: 
  pdf_document:

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
---

```{r, eval=TRUE, include=FALSE}
#```{r, eval=FALSE, include=FALSE}
# options to create final publication quality figures
library(knitr)
opts_chunk$set(dev="tiff",
               dev.args=list(compression="lzw"),
               dpi=300,
               cache=FALSE,
               fig.path='')
```

```{r,echo=FALSE,results='hide',warning=FALSE}
  library(resistance)
  library(ggplot2)
```

```{r summary_fig, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', fig.cap="Comparing time-to-resistance for mixtures and sequences  A. Resistance arises slower for the mixture, B. Resistance arises slower for the sequence. All inputs set to 0.5 except for starting frequency of resistance set to 0.01, and effectiveness of insecticide 1 set to 0.4 in A and 0.8 in B."}

#2 panel fig A. sequence better, B. mix better

#eff1  0.4  0.8

# add side panels for labels
#layout( matrix(c(1:6,0,7,7,0,8,9),4,3, byrow = TRUE), heights=c(1,1,0.15,0.2), widths=c(0.1,1,1) )
# 2 side by side plus lower panl for legent
#layout( matrix(c(1:2,3,3),2,2, byrow = TRUE), heights=c(1,0.04), widths=c(1,1) )
#layout.show(n=3)
# just 2 side by side plus lower panel for legent
layout( matrix(c(1:2),2,2, byrow = TRUE), heights=c(1), widths=c(1,1) )

#eff1 0.4 seq better

#plot margins, high t just a bodge to get graphs down closer to legend
par(mar=c(1.5, 4, 15, 0)) #b,l,t,r default c(5, 4, 4, 2)

runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.4 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, addStrategyLabels = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = NULL )

mtext('A.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r

#plot margins, hight t just a bodge to get graphs down closer to legend
par(mar=c(1.5, 3, 15, 1)) #b,l,t,r default c(5, 4, 4, 2)

#eff1 0.8 mix better
runcurtis_f2( max_gen=500,  P_1 = 0.01 , P_2 = 0.01 , h.RS1_A0 = 0.5 , h.RS2_0B = 0.5 , exposure = 0.5 , phi.SS1_A0 = 0.8 , phi.SS2_0B = 0.5 , rr_restoration_ins1 = 0.5 , rr_restoration_ins2 = 0.5 , addCombinedStrategy = FALSE, addStrategyLabels = FALSE, ylab="", ylabs = FALSE, xlabs = TRUE, cex.axis = 0.8, addLegend=FALSE, main='', maxX = 140, labelMixSeqRatio = NULL )

legend( 'bottomright', legend=c("insecticide1 alone", "insecticide1 in mix", "insecticide2 alone", "insecticide2 in mix"), col=c("red","red","blue","blue"), lty=c(2,1,2,1), pch=c(NA,NA,NA,NA), bty="n", cex=0.9 )

mtext('B.',side=3, adj=0, line=1, cex=0.9) #side=1b,2l,3t,4r

#adding legend in lower panel

# par(mar=c(0, 0, 0, 0), pty='m') #b,l,t,r default c(5, 4, 4, 2)
# 
# plot( 0, type="n", axes=FALSE, ann=FALSE, pty='m') 
# 
# legend( 'bottom', inset=0.1, legend=c("insecticide1 alone   ", "insecticide2 alone   ", "insecticide1 in mix   ", "insecticide2 in mix   "), col=c("red","blue","red","blue"), lty=c(2,2,1,1), pch=c(NA,NA,NA,NA), bty="n", cex=1.05, horiz =TRUE )

```

### 2018-02-22 inst/shiny/resistmixseq/ui.r
tweak readme & UI post MJ publication

@@ -45,13 +45,13 @@ shinyUI(fluidPage( theme = "bootstrap_simplex.css",
   title = "resistance to insecticide mixtures",
 
   #navbarPage sets up navbar, title appears on left
-  navbarPage("insecticide resistance evolution with mixtures and sequences", id="selectedTab",
+  navbarPage("insecticide resistance evolution with mixtures and sequences, exploring mechanisms", id="selectedTab",
              
     # tab UI
     tabPanel("UI", 
            
   
-  h6("Modify inputs, press Run button to generate plot for either scenario. Select About for outline."),  
+  h6("Warning! This is a research tool and not for making operational decisions.  Modify inputs, press Run button to generate plot for either scenario. Select About for outline."),  
   
   #fixedRow(
   fluidRow(

### 2018-06-06 R/plot_wos.r
adding wos stuff

@@ -22,7 +22,7 @@
 #' @param outtoresistance whether to calculate and output resistance model inputs
 #' @param xreverse whether to reverse x axis for declining concentration
 #' 
-#' @import ggplot2 tidyverse
+#' @import ggplot2 tidyverse patchwork
 #' 
 #' @examples 
 #' plot_wos(mort_slope=3, dom_resist=0.9)
@@ -262,7 +262,7 @@ plot_wos <- function( conc_n = 50,
       labs(y='dominance',x='')
     if (xreverse) gg_dom <- gg_dom + scale_x_continuous(trans='reverse')    
     
-    library(patchwork)
+    #library(patchwork)
     #plot(gg / gg_timetor)
 
     plot(gg / gg_dom/ gg_timetor + plot_layout(ncol = 1, heights = c(5,1,5)))

### 2018-06-06 R/setInputSensiScenarios.r

@@ -5,6 +5,7 @@
 #' You need to pass a range for which parameters you want, as c(min,max).
 #' Scenarios will be created by selecting from a uniform distribution specified by this range.
 #' Any parameters not specified will be set to their default values.
+#' TODO it fails if no param ranges are passed i.e. you can't just create a matrix of repeated default scenarios
 #' 
 #' @param nScenarios number of scenarios to generate 
 #' @param ... parameter ranges to run, passed as name=c(min,max)

### 2018-06-11 R/inputs_from_gen_fit.r
starting refactoring wos functions

@@ -47,13 +47,15 @@ inputs_from_gen_fit <- function ( SS = 0.2,
                                   RR_noi = NULL)
 {
   
-  dfin <- data.frame(eff=rep(NA,length(SS)), rr_=NA, dom=NA, cost=NA, domcost=NA)
+  #put the input data into outfile, aids checking and allows it to accept vectors
+  dfin <- data.frame(SS=SS, SR=SR, RR=RR,
+    eff=NA, rr_=NA, dom=NA, cost=NA, domcost=NA)
   
   #in insecticide
   # Effectiveness (1-SSfit)
   # Selection coefficient = RRfit-SSfit
   # Resistance restoration = selection coefficient / effectiveness
-  # Dominance of restoration = (SRfit-SSfit)/(RRfit-SSfit)
+  # Dominance of resistance = (SRfit-SSfit)/(RRfit-SSfit)
 
   
   if (!is.null(SS_noi)) 

### 2018-06-11 R/wos_diagram.r

@@ -25,14 +25,14 @@
 #' @import ggplot2 tidyverse patchwork
 #' 
 #' @examples 
-#' plot_wos(mort_slope=3, dom_resist=0.9)
-#' dfout <- plot_wos(out=TRUE, rr_cost=0)
+#' wos_diagram(mort_slope=3, dom_resist=0.9)
+#' dfout <- wos_diagram(out=TRUE, rr_cost=0)
 #' 
 #' @return ggplot object
 #' @export
 
 
-plot_wos <- function( conc_n = 50,
+wos_diagram <- function( conc_n = 50,
                       #conc_rr_mort1 <- 0.7 # concentration killing all of rr
                       #conc_ss_mort1 <- 0.3 # concentration killing all of ss
                       #or try mort 0 because fits better in y=mx+c
@@ -130,8 +130,10 @@ plot_wos <- function( conc_n = 50,
     theme( legend.position = legendpos,
            #panel.grid = element_blank(),
            axis.text.x = element_blank(),
+           axis.line.x = element_line(arrow = arrow(length = unit(0.1, "inches"))),
            #plot.title = element_text(hjust = 0.5),
-           panel.background = element_rect(colour = 'grey60', fill=NA, size=0.5))
+           axis.line.y = element_line())
+           #panel.background = element_rect(colour = 'grey60', fill=NA, size=0.5))
   
   #shading window of selection, i'm not sure whether helpful to shade between
   #the curves or vertically
@@ -171,102 +173,67 @@ plot_wos <- function( conc_n = 50,
   # 4. Dominance of resistance | (SR-SS)/(RR-SS) 
   if (out)
   {
-    dfout <- data_frame( conc = c(concs),
-                         mort_rr = mort_rr,
-                         mort_sr = mort_sr,
-                         mort_ss = mort_ss,
-                         fit_rr = 1-mort_rr,
-                         fit_sr = 1-mort_sr,
-                         fit_ss = 1-mort_ss,  
-                         effectiveness = mort_ss,
-                         resist_restor = (fit_rr-fit_ss)/effectiveness,
-                         dom_resist = (fit_sr-fit_ss)/(fit_rr-fit_ss) )
     
-    #resist_restor & dom_resist are sometimes NaN
-    #try setting to 0
-    dfout$resist_restor=ifelse(dfout$resist_restor=='NaN',0,dfout$resist_restor)
-    dfout$dom_resist=ifelse(dfout$dom_resist=='NaN',0,dfout$dom_resist)
[…]

### 2018-06-11 R/wos_plot_input.R

@@ -1,0 +1,35 @@
+
+#gg_dom <- wos_plot_input(dfout, input='dom_resist', label='dominance')
+
+wos_plot_input <- function( dfout,
+                         name = "dom_resist",
+                         label = "dominance",
+                         title = "selected simulation inputs",
+                         addwindow = TRUE,
+                         addshading = TRUE,
+                         addlabels = TRUE,
+                         legendpos = 'bottom', #'right'
+                         plot = TRUE,
+                         xreverse = TRUE
+) {
+
+
+# plot dominance
+gg <- dfout %>%
+  ggplot(aes_string(x='conc', y=name))+
+  #geom_point(shape=1, size=0.6) + #open circle
+  geom_line(linetype=4) +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(),
+        axis.title.y = element_text(size = rel(0.7))) +      
+  ggtitle(title) +
+  scale_y_continuous(limits=c(0,1), breaks=c(0,1)) +
+  #labs(y='',x='')
+  labs(y=label,x='')
+if (xreverse) gg <- gg + scale_x_continuous(trans='reverse')  
+
+if(plot) plot(gg)
+
+invisible(gg)
+
+}

### 2018-06-11 R/wos_plot_timetor.r

@@ -1,0 +1,19 @@
+
+wos_plot_timetor <- function()
+{
+
+#plot simulation results
+gg_timetor <- dfout %>%
+  
+  ggplot(aes(x=conc, y=time_to_resistance0.5))+
+  geom_point() +
+  #geom_line(lwd=2,alpha=0.6) +
+  theme_minimal() +
+  theme(axis.text.x = element_blank()) +
+  labs(title = paste0("simulation, exposure=",exposure," starting resistance frequency=",startfreq)) +
+  scale_y_continuous(limits=c(0,NA)) +
+  labs(x='') 
+#labs(x='Time or declining insecticide concentration')
+if (xreverse) gg_timetor <- gg_timetor + scale_x_continuous(trans='reverse')
+
+}

### 2018-06-11 R/wos_sim.r

@@ -1,0 +1,113 @@
+#' wos_sim run a series of resistance evolution simulations to illustrate 'windows of selection'
+#' 
+#' 
+#' @param conc_n number of x intervals default 50, will influence smoothness of lines, converted to 0-1
+#' @param conc_rr_mort0 concentration killing none of rr
+#' @param conc_ss_mort0 concentration killing none of ss
+#' @param mort_slope = 3 slope of mortality curves
+#' @param sr whether to include heterozygotes
+#' @param dom_resistance dominance of resistance 0-1 determines position of sr between rr & ss but it's not really dominance
+#' @param rr_cost cost of resistance to rr (simply added to rr mort)
+#' @param dom_cost dominance of cost
+#' @param exposure proportion of popn exposed to insecticide only used in the simulations
+#' @param max_gen maximum generations to use in the simulations
+#' @param startfreq starting frequency to use in the simulations
+#' @param title title for ggplot
+#' @param addwindow whether to add dotted lines for opening and shutting of window
+#' @param addshading whether to shade window
+#' @param addlabels whether to add labels
+#' @param legendpos where to add legend
+#' @param plot whether to plot
+#' @param outtoresistance whether to calculate and output resistance model inputs
+#' @param xreverse whether to reverse x axis for declining concentration
+#' 
+#' @import ggplot2 tidyverse patchwork
+#' 
+#' @examples 
+#' plot_wos(mort_slope=3, dom_resist=0.9)
+#' dfout <- plot_wos(out=TRUE, rr_cost=0)
+#' 
+#' @return ggplot object
+#' @export
+
+
+wos_sim <- function( concs = c(1:5),
+                     mort_rr = c(0,0,0,0.5,1),
+                     mort_sr =c(0,0,0.5,1,1),
+                     mort_ss = c(0,0.5,1,1,1),
+                     exposure = 0.5, #only used in the simulation
+                     max_gen = 1000, #used in simulations
+                     startfreq = 0.001
+) {
+  
+dfout <- data_frame( conc = c(concs),
+                     mort_rr = mort_rr,
+                     mort_sr = mort_sr,
+                     mort_ss = mort_ss,
+                     fit_rr = 1-mort_rr,
+                     fit_sr = 1-mort_sr,
+                     fit_ss = 1-mort_ss,  
+                     effectiveness = mort_ss,
[…]

### 2018-06-13 R/wos_advantage.r
wos functions refactored

@@ -1,0 +1,76 @@
+#' wos_advantage calculate selective advantage through 'window of selection'
+#' 
+#' from mortalities of rr sr ss
+#' 
+#' @param concs concentration or time
+#' @param mort_rr rr mortalities
+#' @param mort_sr sr mortalities
+#' @param mort_ss ss mortalities
+#' @param exposure proportion of popn exposed to insecticide 
+# @param max_gen maximum generations to use in the simulations
+#' @param startfreq starting frequency to use in the simulations
+#' 
+#' @import dplyr
+#' 
+#' @examples 
+#' wos_advantage()
+#' 
+#' @return dataframe of simulation outputs
+#' @export
+
+
+wos_advantage <- function( concs = c(1:5),
+                     mort_rr = c(0,0,0,0.5,1),
+                     mort_sr =c(0,0,0.5,1,1),
+                     mort_ss = c(0,0.5,1,1,1),
+                     exposure = 0.5, #only used in the simulation
+                     #max_gen = 1000, #used in simulations
+                     startfreq = 0.001
+) {
+  
+  freq <- startfreq
+  
+  dfsim <- data_frame( conc = c(concs),
+                       freq = freq, #just put in here to make dplyr easier below
+                       mort_rr = mort_rr,
+                       mort_sr = mort_sr,
+                       mort_ss = mort_ss,
+                       fit_rr = 1-mort_rr,
+                       fit_sr = 1-mort_sr,
+                       fit_ss = 1-mort_ss )
+ 
+  dfsim <- dfsim %>% 
+    mutate(r_next_gen = exposure*(freq^2*fit_rr + 2*freq*(1-freq)*0.5*fit_sr) +
+                        (1-exposure)*(freq^2+2*freq*(1-freq)*0.5)) %>%
+  
+    mutate(s_next_gen = exposure*(2*freq*(1-freq)*0.5*fit_sr +
+                        (1-freq)^2 * fit_ss) +
+                        (1-exposure)*(2*freq*(1-freq)*0.5+(1-freq)*(1-freq))) %>%
+    
+    # S_next_gen=exposed*(2*R_freq*(1-R_freq)*0.5*fitness_RS+(1-R_freq)*(1-R_freq)*fitness_SS)+
[…]

### 2018-06-13 R/wos_diagram.r

@@ -19,14 +19,14 @@
 #' @param addlabels whether to add labels
 #' @param legendpos where to add legend
 #' @param plot whether to plot
-#' @param outtoresistance whether to calculate and output resistance model inputs
+#' @param out whether to calculate and output resistance model inputs
 #' @param xreverse whether to reverse x axis for declining concentration
 #' 
 #' @import ggplot2 tidyverse patchwork
 #' 
 #' @examples 
 #' wos_diagram(mort_slope=3, dom_resist=0.9)
-#' dfout <- wos_diagram(out=TRUE, rr_cost=0)
+#' dfsim <- wos_diagram(out=TRUE, rr_cost=0)
 #' 
 #' @return ggplot object
 #' @export
@@ -53,6 +53,7 @@ wos_diagram <- function( conc_n = 50,
                       legendpos = 'bottom', #'right'
                       plot = TRUE,
                       out = FALSE,
+                      adv = FALSE,
                       xreverse = TRUE
 ) {
   
@@ -173,68 +174,55 @@ wos_diagram <- function( conc_n = 50,
   # 4. Dominance of resistance | (SR-SS)/(RR-SS) 
   if (out)
   {
-    
-    dfout <- wos_sim(concs = c(1:5),
-                 mort_rr = c(0,0,0,0.5,1),
-                 mort_sr =c(0,0,0.5,1,1),
-                 mort_ss = c(0,0.5,1,1,1))
-    
-    gg_timetor <- wos_plot_timetor(dfout)
-    
-    gg_dom <- wos_plot_input(dfout, input='dom_resist', label='dominance')
-    
-    # plot dominance
-    gg_dom <- dfout %>%
-      ggplot(aes(x=conc, y=dom_resist))+
-      #geom_point(shape=1, size=0.6) + #open circle
-      geom_line(linetype=4) +
-      theme_minimal() +
-      theme(axis.text.x = element_blank(),
-            axis.title.y = element_text(size = rel(0.7))) +      
-      ggtitle("selected simulation inputs") +
-      scale_y_continuous(limits=c(0,1), breaks=c(0,1)) +
-      #labs(y='',x='')
-      labs(y='dominance',x='')
-    if (xreverse) gg_dom <- gg_dom + scale_x_continuous(trans='reverse')    
-    
-    # plot resistance restoration
-    gg_rr <- dfout %>%
-      ggplot(aes(x=conc, y=resist_restor))+
-      #geom_point(shape=3, size=0.6) + #shape=3 cross
-      geom_line(linetype=3) +
[…]

### 2018-06-13 R/wos_plot_input.R

@@ -1,32 +1,58 @@
+#' plot one simulation input for 'windows of selection' following simulation run
+#' 
+#' @param dfsim data  on concentration and time-to-resistance output from simulation run from wos_sim()
+#' @param x name of column of x data
+#' @param xlabel label of column of x data for plot
+#' @param y name of column of y data
+#' @param ylabel label of column of y data for plot 
+#' @param title title for ggplot
+# @param legendpos where to add legend
+#' @param plot whether to plot
+#' @param xreverse whether to reverse x axis for declining concentration
+#' 
+#' @import ggplot2 tidyverse patchwork
+#' 
+#' @examples 
+#' dfsim <- wos_diagram(out=TRUE, rr_cost=0, plot=FALSE)
+#' wos_plot_input(dfsim, y='dom_resist', ylab='dominance')
+#' 
+#' @return ggplot object
+#' @export
 
-#gg_dom <- wos_plot_input(dfout, input='dom_resist', label='dominance')
-
-wos_plot_input <- function( dfout,
-                         name = "dom_resist",
-                         label = "dominance",
+wos_plot_input <- function( dfsim,
+                         x='conc',  
+                         y = "dom_resist",
+                         xlab = NULL,
+                         ylab = NULL,                         
                          title = "selected simulation inputs",
-                         addwindow = TRUE,
-                         addshading = TRUE,
-                         addlabels = TRUE,
-                         legendpos = 'bottom', #'right'
+                         #legendpos = 'bottom', #'right'
                          plot = TRUE,
                          xreverse = TRUE
 ) {
 
 
 # plot dominance
-gg <- dfout %>%
-  ggplot(aes_string(x='conc', y=name))+
+gg <- dfsim %>%
+  ggplot(aes_string(x=x, y=y))+
   #geom_point(shape=1, size=0.6) + #open circle
   geom_line(linetype=4) +
   theme_minimal() +
+  #scale_x_continuous(sec.axis = dup_axis()) +
+
   theme(axis.text.x = element_blank(),
-        axis.title.y = element_text(size = rel(0.7))) +      
+        #axis.title.y = element_text(size = rel(0.7)),
+        #2ry x axis for ylab
+        axis.title.x.top = element_text(hjust=0)) +
+        #axis.title.y = element_text(size = rel(0.7),angle=0, hjust=0)) +
[…]

### 2018-06-13 R/wos_plot_timetor.r

@@ -1,19 +1,52 @@
-
-wos_plot_timetor <- function()
-{
+#' plot time-to-resistance for 'windows of selection' following simulation run
+#' 
+#' @param dfsim data  on concentration and time-to-resistance output from simulation run from wos_sim()
+#' @param x name of column of x data
+#' @param xlabel label of column of x data for plot
+#' @param y name of column of y data
+#' @param ylabel label of column of y data for plot 
+#' @param title title for ggplot
+# @param legendpos where to add legend
+#' @param plot whether to plot
+#' @param xreverse whether to reverse x axis for declining concentration
+#' 
+#' @import ggplot2 tidyverse patchwork
+#' 
+#' @examples 
+#' dfsim <- wos_diagram(out=TRUE, rr_cost=0, plot=FALSE)
+#' wos_plot_timetor(dfsim)
+#' 
+#' @return ggplot object
+#' @export
+#' 
+wos_plot_timetor <- function(dfsim,
+                             x='conc',
+                             y='time_to_resistance0.5',
+                             ylab='time to\nresistance',
+                             xlab = NULL,
+                             title = "simulation results",
+                             #legendpos = 'bottom', #'right'
+                             plot = TRUE,
+                             xreverse = TRUE
+) {
 
 #plot simulation results
-gg_timetor <- dfout %>%
+gg <- dfsim %>%
   
-  ggplot(aes(x=conc, y=time_to_resistance0.5))+
+  ggplot(aes_string(x=x, y=y))+
   geom_point() +
   #geom_line(lwd=2,alpha=0.6) +
   theme_minimal() +
   theme(axis.text.x = element_blank()) +
-  labs(title = paste0("simulation, exposure=",exposure," starting resistance frequency=",startfreq)) +
+  labs(title = title) +
+  #labs(title = paste0("simulation, exposure=",exposure," starting resistance frequency=",startfreq)) +
   scale_y_continuous(limits=c(0,NA)) +
-  labs(x='') 
+  labs(x='', y=ylab) 
 #labs(x='Time or declining insecticide concentration')
-if (xreverse) gg_timetor <- gg_timetor + scale_x_continuous(trans='reverse')
+if (xreverse) gg <- gg + scale_x_continuous(trans='reverse')
+
+if(plot) plot(gg)
+
+invisible(gg)
 
 }

### 2018-06-13 R/wos_sim.r

@@ -1,33 +1,18 @@
 #' wos_sim run a series of resistance evolution simulations to illustrate 'windows of selection'
 #' 
-#' 
-#' @param conc_n number of x intervals default 50, will influence smoothness of lines, converted to 0-1
-#' @param conc_rr_mort0 concentration killing none of rr
-#' @param conc_ss_mort0 concentration killing none of ss
-#' @param mort_slope = 3 slope of mortality curves
-#' @param sr whether to include heterozygotes
-#' @param dom_resistance dominance of resistance 0-1 determines position of sr between rr & ss but it's not really dominance
-#' @param rr_cost cost of resistance to rr (simply added to rr mort)
-#' @param dom_cost dominance of cost
+#' @param concs concentration or time
+#' @param mort_rr rr mortalities
+#' @param mort_sr sr mortalities
+#' @param mort_ss ss mortalities
 #' @param exposure proportion of popn exposed to insecticide only used in the simulations
 #' @param max_gen maximum generations to use in the simulations
 #' @param startfreq starting frequency to use in the simulations
-#' @param title title for ggplot
-#' @param addwindow whether to add dotted lines for opening and shutting of window
-#' @param addshading whether to shade window
-#' @param addlabels whether to add labels
-#' @param legendpos where to add legend
-#' @param plot whether to plot
-#' @param outtoresistance whether to calculate and output resistance model inputs
-#' @param xreverse whether to reverse x axis for declining concentration
 #' 
-#' @import ggplot2 tidyverse patchwork
 #' 
 #' @examples 
-#' plot_wos(mort_slope=3, dom_resist=0.9)
-#' dfout <- plot_wos(out=TRUE, rr_cost=0)
+#' wos_sim()
 #' 
-#' @return ggplot object
+#' @return dataframe of simulation outputs
 #' @export
 
 
@@ -40,7 +25,7 @@ wos_sim <- function( concs = c(1:5),
                      startfreq = 0.001
 ) {
   
-dfout <- data_frame( conc = c(concs),
+dfsim <- data_frame( conc = c(concs),
                      mort_rr = mort_rr,
                      mort_sr = mort_sr,
                      mort_ss = mort_ss,
@@ -53,20 +38,17 @@ dfout <- data_frame( conc = c(concs),
 
[…]

### 2018-06-13 inst/documents/wos_advantage_ian.r

@@ -1,0 +1,15 @@
+exposed=0.5# The proportion exposed
+R_freq=0.01; #frequency of the resistance allele. this will affect results if resistance is recessive beacuse it determines the proportion of R allelels in RR genotypes
+fitness_RR=0.5
+fitness_RS=0
+fitness_SS=0 
+
+R_next_gen=exposed*(R_freq*R_freq*fitness_RR+2*R_freq*(1-R_freq)*0.5*fitness_RS)+
+  (1-exposed)*(R_freq*R_freq+2*R_freq*(1-R_freq)*0.5);
+
+S_next_gen=exposed*(2*R_freq*(1-R_freq)*0.5*fitness_RS+(1-R_freq)*(1-R_freq)*fitness_SS)+
+  (1-exposed)*(2*R_freq*(1-R_freq)*0.5+(1-R_freq)*(1-R_freq));
+
+R_freq_next_gen=R_next_gen/(R_next_gen+S_next_gen)
+
+selective_advantage=(R_freq_next_gen/R_freq)-1 #so if no increase, advantage=0

### 2018-06-14 R/wos_advantage.r
wos

@@ -52,9 +52,12 @@ wos_advantage <- function( concs = c(1:5),
     
              
     mutate(freq_next_gen = r_next_gen/(r_next_gen+s_next_gen)) %>%
+    
+    mutate(relative_fitness=(freq_next_gen/freq)) %>%
              
     mutate(selective_advantage=(freq_next_gen/freq)-1) #so if no increase, advantage=0         
 
+  
   # exposed=0.5# The proportion exposed
   # R_freq=0.001; #frequency of the resistance allele. this will affect results if resistance is recessive beacuse it determines the proportion of R allelels in RR genotypes
   # fitness_RR=1.0

### 2018-06-14 R/wos_diagram.r

@@ -20,6 +20,8 @@
 #' @param legendpos where to add legend
 #' @param plot whether to plot
 #' @param out whether to calculate and output resistance model inputs
+#' @param adv whether to calculate and output selective advantage
+#' @param logadv whether to use log scale in selective advantage plots
 #' @param xreverse whether to reverse x axis for declining concentration
 #' 
 #' @import ggplot2 tidyverse patchwork
@@ -41,7 +43,7 @@ wos_diagram <- function( conc_n = 50,
                       mort_slope = 3, #0.5 # slope of mortality curves
                       sr = TRUE, # whether to include heterozygotes
                       dom_resistance = 0.5, #1 # dominance of resistance 0-1, it's not really dominance
-                      rr_cost = 0.1, #cost of resistance to rr (simply added to rr mort)
+                      rr_cost = 0, #.1, #cost of resistance to rr (simply added to rr mort)
                       dom_cost = 0, #dominance of cost
                       exposure = 0.5, #only used in the simulation
                       max_gen = 1000, #used in simulations
@@ -54,8 +56,11 @@ wos_diagram <- function( conc_n = 50,
                       plot = TRUE,
                       out = FALSE,
                       adv = FALSE,
+                      logadv = TRUE,
                       xreverse = TRUE
 ) {
+  
+  if (rr_cost > 0 & out) warning('running simulation with rr_cost>0 can cause it to crash') 
   
   concs <- seq(0,1,length=conc_n)
   #trying (and failing) to base directly on helps2017
@@ -95,7 +100,7 @@ wos_diagram <- function( conc_n = 50,
   sr_win_opens <- concs[max(which(mort_ss > mort_sr))]
   sr_win_closes <- concs[min(which(mort_ss > mort_sr))]
   
-  dfdiag <- data_frame( conc = c(concs,concs,concs),
+  dfdiag <- dplyr::data_frame( conc = c(concs,concs,concs),
                         genotype = c(rep('rr',conc_n),
                                      rep('ss',conc_n),
                                      rep('sr',conc_n)),
@@ -181,7 +186,8 @@ wos_diagram <- function( conc_n = 50,
                      mort_ss = mort_ss,
                      exposure = exposure,
                      max_gen = max_gen,
-                     startfreq = startfreq
+                     startfreq = startfreq,
+                     plot = FALSE
                      )
 
     dfadv <- wos_advantage(concs = concs,
[…]

### 2018-06-14 R/wos_plot_timetor.r

@@ -2,15 +2,17 @@
 #' 
 #' @param dfsim data  on concentration and time-to-resistance output from simulation run from wos_sim()
 #' @param x name of column of x data
-#' @param xlabel label of column of x data for plot
+#' @param xlab label of column of x data for plot
+#' @param xtxt whether to number x axis 
 #' @param y name of column of y data
-#' @param ylabel label of column of y data for plot 
+#' @param ylab label of column of y data for plot 
 #' @param title title for ggplot
 # @param legendpos where to add legend
 #' @param plot whether to plot
 #' @param xreverse whether to reverse x axis for declining concentration
+#' @param xlog whether to log x axis
 #' 
-#' @import ggplot2 tidyverse patchwork
+#' @import ggplot2 tidyverse patchwork scales
 #' 
 #' @examples 
 #' dfsim <- wos_diagram(out=TRUE, rr_cost=0, plot=FALSE)
@@ -24,12 +26,24 @@ wos_plot_timetor <- function(dfsim,
                              y='time_to_resistance0.5',
                              ylab='time to\nresistance',
                              xlab = NULL,
+                             xtxt = FALSE,
                              title = "simulation results",
                              #legendpos = 'bottom', #'right'
                              plot = TRUE,
-                             xreverse = TRUE
+                             xreverse = TRUE,
+                             xlog = FALSE
 ) {
 
+#to allow reverse logscale
+#library("scales")
+reverselog_trans <- function(base = exp(1)) {
+  trans <- function(x) -log(x, base)
+  inv <- function(x) base^(-x)
+  scales::trans_new(paste0("reverselog-", format(base)), trans, inv, 
+                    scales::log_breaks(base = base), 
+                    domain = c(1e-100, Inf))
+}  
+  
 #plot simulation results
 gg <- dfsim %>%
   
@@ -37,13 +51,17 @@ gg <- dfsim %>%
   geom_point() +
   #geom_line(lwd=2,alpha=0.6) +
   theme_minimal() +
-  theme(axis.text.x = element_blank()) +
+  #theme(axis.text.x = element_blank()) +
   labs(title = title) +
   #labs(title = paste0("simulation, exposure=",exposure," starting resistance frequency=",startfreq)) +
   scale_y_continuous(limits=c(0,NA)) +
   labs(x='', y=ylab) 
[…]

### 2018-06-14 R/wos_sim.r

@@ -1,29 +1,63 @@
 #' wos_sim run a series of resistance evolution simulations to illustrate 'windows of selection'
 #' 
+#' @param dfmortbygen dataframe of mortality by genotype, an alternative way of initialising function, needs genotype column
+#' @param x column name of x in dfmortbygen
+#' @param y column name of y in dfmortbygen 
 #' @param concs concentration or time
-#' @param mort_rr rr mortalities
-#' @param mort_sr sr mortalities
-#' @param mort_ss ss mortalities
+#' @param mort_rr rr mortalities 0-1
+#' @param mort_sr sr mortalities 0-1
+#' @param mort_ss ss mortalities 0-1
 #' @param exposure proportion of popn exposed to insecticide only used in the simulations
 #' @param max_gen maximum generations to use in the simulations
 #' @param startfreq starting frequency to use in the simulations
-#' 
+#' @param plot whether to plot
 #' 
 #' @examples 
 #' wos_sim()
+#' wos_sim(dfmortbygen=data_frame(conc=rep(c(1,2),3),
+#'                                genotype=rep(c('RR','SR','SS'),each=2),
+#'                                mort=c(0,0.2,0.4,0.6,0.8,1)),
+#'         x='conc',y='mort')
+#'         
 #' 
 #' @return dataframe of simulation outputs
 #' @export
 
 
-wos_sim <- function( concs = c(1:5),
+wos_sim <- function( dfmortbygen = NULL,
+                     x = NULL,
+                     y = NULL,
+                     concs = c(1:5),
                      mort_rr = c(0,0,0,0.5,1),
                      mort_sr =c(0,0,0.5,1,1),
                      mort_ss = c(0,0.5,1,1,1),
                      exposure = 0.5, #only used in the simulation
                      max_gen = 1000, #used in simulations
-                     startfreq = 0.001
+                     startfreq = 0.001,
+                     plot = TRUE
 ) {
+ 
+   
+if (!is.null(dfmortbygen))
+{
+  #removes blank rows and comments
+  #dfmortbygen <- dfmortbygen[ !is.na(dfmortbygen$genotype), ]
+  #!!(enquo(x))
+  #!! unquotes a variable name
+  dfmortbygen <- filter(dfmortbygen, !is.na(!!x))
+  
[…]

### 2018-06-22 R/wos_advantage.r
wos functions

@@ -1,7 +1,12 @@
 #' wos_advantage calculate selective advantage through 'window of selection'
 #' 
-#' from mortalities of rr sr ss
+#' Similar to wos_sim() which simulates time-to-resistance. 
+#' Requires either a dataframe with columns for genotype, x (e.g. time) and y (e.g.mortality)
+#' or vectors of x, and yrr,ysr,yss. 
 #' 
+#' @param dfmortbygen dataframe of mortality by genotype, an alternative way of initialising function, needs genotype column
+#' @param x column name of x in dfmortbygen
+#' @param y column name of y in dfmortbygen 
 #' @param concs concentration or time
 #' @param mort_rr rr mortalities
 #' @param mort_sr sr mortalities
@@ -19,7 +24,10 @@
 #' @export
 
 
-wos_advantage <- function( concs = c(1:5),
+wos_advantage <- function( dfmortbygen = NULL,
+                           x = 'concentration',
+                           y = 'mortality',
+                           concs = c(1:5),
                      mort_rr = c(0,0,0,0.5,1),
                      mort_sr =c(0,0,0.5,1,1),
                      mort_ss = c(0,0.5,1,1,1),
@@ -28,6 +36,51 @@ wos_advantage <- function( concs = c(1:5),
                      startfreq = 0.001
 ) {
   
+  #%% to this mark are copied from wos_sim() as temp. measure I must be able to rationalise ...
+  # if the input data are provided in a dataframe   
+  if (!is.null(dfmortbygen))
+  {
+    #removes blank rows and comments
+    #dfmortbygen <- dfmortbygen[ !is.na(dfmortbygen$genotype), ]
+    #!!(enquo(x))
+    #!! unquotes a variable name
+    dfmortbygen <- filter(dfmortbygen, !is.na(!!x))
+    
+    #assumes that concs are the same and repeatd for rr,sr,ss
+    concs <- filter(dfmortbygen,genotype=='RR') %>% select(!!x) %>% unlist(use.names=FALSE) 
+    mort_rr <- filter(dfmortbygen,genotype=='RR') %>% select(!!y) %>% unlist(use.names=FALSE)
+    # if no sr data
+    if (nrow(filter(dfmortbygen,genotype=='SR'))==0)
+    { 
[…]

### 2018-06-22 R/wos_diagram.r

@@ -204,13 +204,13 @@ wos_diagram <- function( conc_n = 50,
     #plot(gg / gg_dom/ gg_timetor + plot_layout(ncol = 1, heights = c(5,1,5)))
     if (plot)
     {
-      gg_timetor <- wos_plot_timetor(dfsim,
+      gg_timetor <- wos_plot_sim(dfsim,
                                      title=paste0("simulation, exposure=",exposure," starting resistance frequency=",startfreq),
                                      plot=FALSE)
       
       #selective advantage
-      #gg_adv <- wos_plot_timetor(dfadv, x='conc', y='selective_advantage', ylab='selective\nadvantage', 
-      gg_adv <- wos_plot_timetor(dfadv, x='conc', y='relative_fitness', ylab='relative\nfitness',                                  
+      #gg_adv <- wos_plot_sim(dfadv, x='conc', y='selective_advantage', ylab='selective\nadvantage', 
+      gg_adv <- wos_plot_sim(dfadv, x='conc', y='relative_fitness', ylab='relative\nfitness',                                  
                                      title=paste0("relative fitness, exposure=",exposure," starting resistance frequency=",startfreq),
                                      plot=FALSE)
       #log selective advantage

### 2018-06-22 R/wos_plot_sim.r

@@ -11,19 +11,22 @@
 #' @param plot whether to plot
 #' @param xreverse whether to reverse x axis for declining concentration
 #' @param xlog whether to log x axis
+#' @param ylog whether to log y axis
+#' @param no_sr option for dealing with data with no SR heteroxzygotes plots dominance 0&1 with diff symbols
 #' 
 #' @import ggplot2 tidyverse patchwork scales
 #' 
 #' @examples 
-#' dfsim <- wos_diagram(out=TRUE, rr_cost=0, plot=FALSE)
-#' wos_plot_timetor(dfsim)
+#' dfsim <- wos_diagram(out=TRUE, conc_n = 10, rr_cost=0, plot=FALSE)
+#' wos_plot_sim(dfsim)
 #' 
 #' @return ggplot object
 #' @export
 #' 
-wos_plot_timetor <- function(dfsim,
+wos_plot_sim <- function(dfsim,
                              x='conc',
                              y='time_to_resistance0.5',
+                             shape=NULL,
                              ylab='time to\nresistance',
                              xlab = NULL,
                              xtxt = FALSE,
@@ -31,7 +34,9 @@ wos_plot_timetor <- function(dfsim,
                              #legendpos = 'bottom', #'right'
                              plot = TRUE,
                              xreverse = TRUE,
-                             xlog = FALSE
+                             xlog = FALSE,
+                             ylog = FALSE,                         
+                             no_sr = FALSE
 ) {
 
 #to allow reverse logscale
@@ -43,11 +48,49 @@ reverselog_trans <- function(base = exp(1)) {
                     scales::log_breaks(base = base), 
                     domain = c(1e-100, Inf))
 }  
+
+# if (no_sr) symbol <- 'dom_resist'
+# else symbol <- NULL
   
+
 #plot simulation results
-gg <- dfsim %>%
+
+#diff symbols for points dominance 0&1 if no SR
+#if (!no_sr) 
+if (is.null(shape))   
+{
+  gg <- dfsim %>% ggplot(aes_string(x=x, y=y))
+} else 
+{
+  #dfsim$dominance <- factor(dfsim$dom_resist)
   
-  ggplot(aes_string(x=x, y=y))+
+  #converting shape column to a factor
+  #rlang::sym solved tidyeval faffing 
[…]

### 2018-06-22 R/wos_sim.r

@@ -1,5 +1,8 @@
 #' wos_sim run a series of resistance evolution simulations to illustrate 'windows of selection'
 #' 
+#' requires either a dataframe with columns for genotype, x (e.g. time) and y (e.g.mortality)
+#' or vectors of x, and yrr,ysr,yss
+#' 
 #' @param dfmortbygen dataframe of mortality by genotype, an alternative way of initialising function, needs genotype column
 #' @param x column name of x in dfmortbygen
 #' @param y column name of y in dfmortbygen 
@@ -9,35 +12,44 @@
 #' @param mort_ss ss mortalities 0-1
 #' @param exposure proportion of popn exposed to insecticide only used in the simulations
 #' @param max_gen maximum generations to use in the simulations
+#' @param no_thresh what to return as time-to-resistance if threshold not reached
 #' @param startfreq starting frequency to use in the simulations
+#' @param dominances only if there are no SR values which dominances to run
 #' @param plot whether to plot
 #' 
 #' @examples 
-#' wos_sim()
-#' wos_sim(dfmortbygen=data_frame(conc=rep(c(1,2),3),
+#' dfsim <- wos_sim()
+#' dfsim <- wos_sim(dfmortbygen=data_frame(conc=rep(c(1,2),3),
 #'                                genotype=rep(c('RR','SR','SS'),each=2),
 #'                                mort=c(0,0.2,0.4,0.6,0.8,1)),
 #'         x='conc',y='mort')
-#'         
+#' #with no sr data runs dom0,0.1&1         
+#' dfsim <- wos_sim(dfmortbygen=data_frame(conc=rep(c(1,2),2),
+#'                                genotype=rep(c('RR','SS'),each=2),
+#'                                mort=c(0,0.2,0.8,1)),
+#'         x='conc',y='mort')         
 #' 
 #' @return dataframe of simulation outputs
 #' @export
 
 
 wos_sim <- function( dfmortbygen = NULL,
-                     x = NULL,
-                     y = NULL,
+                     x = 'concentration',
+                     y = 'mortality',
                      concs = c(1:5),
                      mort_rr = c(0,0,0,0.5,1),
                      mort_sr =c(0,0,0.5,1,1),
[…]

### 2018-07-30 R/wos_diagram.r
wos work

@@ -7,7 +7,7 @@
 #' @param conc_ss_mort0 concentration killing none of ss
 #' @param mort_slope = 3 slope of mortality curves
 #' @param sr whether to include heterozygotes
-#' @param dom_resistance dominance of resistance 0-1 determines position of sr between rr & ss but it's not really dominance
+#' @param win_dom_strt dominance of resistance 0-1 determines position of sr between rr & ss but it's not really dominance
 #' @param rr_cost cost of resistance to rr (simply added to rr mort)
 #' @param dom_cost dominance of cost
 #' @param exposure proportion of popn exposed to insecticide only used in the simulations
@@ -28,8 +28,9 @@
 #' @import ggplot2 tidyverse patchwork
 #' 
 #' @examples 
-#' wos_diagram(mort_slope=3, dom_resist=0.9)
+#' wos_diagram(mort_slope=3, win_dom_strt=0.9)
 #' dfsim <- wos_diagram(sim=TRUE, rr_cost=0)
+#' wos_diagram(adv=TRUE,win_dom_strt=0.1)
 #' 
 #' @return ggplot object
 #' @export
@@ -43,7 +44,7 @@ wos_diagram <- function( conc_n = 50,
                       conc_ss_mort0 = 0.1, # concentration killing none of ss
                       mort_slope = 3, #0.5 # slope of mortality curves
                       sr = TRUE, # whether to include heterozygotes
-                      dom_resistance = 0.5, #1 # dominance of resistance 0-1, it's not really dominance
+                      win_dom_strt = 0.5, #1 # dominance of resistance 0-1, it's not really dominance
                       rr_cost = 0, #.1, #cost of resistance to rr (simply added to rr mort)
                       dom_cost = 0, #dominance of cost
                       exposure = 0.5, #only used in the simulation
@@ -81,14 +82,13 @@ wos_diagram <- function( conc_n = 50,
   mort_rr <- mort_rr + rr_cost*(1-mort_rr)
   
   #sr
-  #set where it crosses mort0 to be between ss & rr according to dominance
-  #(although the dominance param isn't really dominance)
-  conc_sr_mort0 <- conc_ss_mort0 + dom_resistance*(conc_rr_mort0-conc_ss_mort0)
+  #set where it crosses mort0 to be between ss & rr according to win_dom_strt
+  conc_sr_mort0 <- conc_ss_mort0 + win_dom_strt*(conc_rr_mort0-conc_ss_mort0)
[…]

### 2020-11-11 inst/documents/linkage_dis_2017.Rmd
rerun linkage dis plots

@@
   author: "Andy South"
-  date: "`r Sys.Date()`"
+  date: "originally from 2017, rerun:`r Sys.Date()`"
 ---
@@
 
-```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left'}
+```{r Fig8, eval=TRUE, echo=FALSE, fig.height=7, fig.width=7, fig.show='hold', fig.align='left', warning=FALSE}
 
